### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/baller_dep_heterogen_server_path.Rmd
First commit

---
title: "Baller_dep_heterogeneity"
author: "Erica Baller"
date: "12/12/2017"
output: html_document
runtime: shiny
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Welcome to Baller Depression Heterogeneity!  This document contains the code and output generated from 
12/12/2017: Subsetting and running mass univariate analysis on CNB cognitive scores

```{r baller_depression_heterogeneity}
#install.packages("visreg")
#install.packages("mgcv")
#install.packages("tableone")
#################################################
#### Get subset for Baller_DepHeterogeneity ####
#################################################

#This does the subsetting, runs statistics with lapply instead of loop, visualizes progress with visreg, and does fdr correction
# it then makes table 1 (demographics)
#load libraries
library(visreg)
library(mgcv)
library(tableone)

#read in csvs
demographics <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
cnb_scores <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
health <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
psych_summary <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv

#remove people with NA for race, age, or sex.  START WITH N = 9498
demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406

#remove people with NA for depression or total psych score, START WITH N = 9498
psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
psych_summary_no_NA_dep_and_smry_psych_overall <- psych_summary_no_NA_dep[!is.na(psych_summary_no_NA_dep$smry_psych_overall_rtg),] #take out those with NA for overall psych rtg, no additional people lost, N = 9411

#merge the csvs
#dem_cnb <- merge(demographics, cnb_scores, by = "bblid") #merge demographics and cnb #this is if we want to include people without full demographic data
dem_cnb <- merge(demographics_noNA_race_age_andCNBage_sex, cnb_scores, by = "bblid") #merge demographics and cnb, N = 9406
psych_health <- merge(psych_summary_no_NA_dep_and_smry_psych_overall, health, by = "bblid") #merge psych and health, N = 9411
dem_cnb_psych_health_merged <- merge (dem_cnb, psych_health, by = "bblid") #merge all 4 csvs, lost 1 person [134716] (had demographics, but no psych ratings): N = 9405

#make subsets
subset_just_dep_and_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & (smry_dep == 4), select = "bblid") #subset people who were not medically excluded and who are depressed, N = 776
subset_no_psych_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & (smry_psych_overall_rtg < 4), select = "bblid") #subset people who are psychiatrically healthy, N = 2508
subset_dep_or_no_psych_and_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & ((smry_dep == 4) | (smry_psych_overall_rtg <4))) #subset including both depressed and healthies, good for regressions, N = 3284

#would binarize depression smry score to 0 and 1-- 4 and <4
dep_binarized <- ifelse(subset_dep_or_no_psych_and_no_medicalratingExclude$smry_dep == 4, 1, 0)
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED <- cbind(subset_dep_or_no_psych_and_no_medicalratingExclude, dep_binarized) #N = 3284

#divide ageAtCNB by 12 for age
age_in_years <- subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$ageAtCnb1/12
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED <- cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED, age_in_years)

#would
#1) use gam and visreg to see effects on single columns CHECK
#2) try with a for loop CHECK
#3) try with an apply function-- can get code from toni CHECK
#4) try with `voxel` wrapper -> NOT imaging data... stuck on this part

#get CNB measure names
cnb_measure_names <- names(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED)[grep("_z", names(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED))] #get the names of all the columns with _z in the name
#cnb_measure_names_lm <- paste0(cnb_measure_names, "_lm", sep='')
#cnb_measure_names_gam <- paste0(cnb_measure_names, "_gam", sep='')

####GLM/GAM comparing depressed with non-depressed, using Lapply.  Only looking at depression, binarized beyond this point

#using lm, results stored in list
CNB_cog_score_stats_lm_dep_binarized <- lapply(cnb_measure_names, function(x) 
{
  lm(substitute(i ~ dep_binarized + sex + age_in_years, list(i = as.name(x))), data = subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED)
})

#using gam, results stored in list
CNB_cog_score_stats_gam_dep_binarized <- lapply(cnb_measure_names, function(x) 
{
  gam(substitute(i ~ dep_binarized + sex + age_in_years, list(i = as.name(x))), data = subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED)
})

#add names for each list element
names(CNB_cog_score_stats_lm_dep_binarized) <- cnb_measure_names
names(CNB_cog_score_stats_gam_dep_binarized) <- cnb_measure_names

#####Only looking at gam models beyond this point#####
#Look at model summaries
models_lm <- lapply(CNB_cog_score_stats_lm_dep_binarized, summary)
models_gam <- lapply(CNB_cog_score_stats_gam_dep_binarized, summary)

#Pull p-values
p_lm <- sapply(CNB_cog_score_stats_lm_dep_binarized, function(v) summary(v)$coef[,"Pr(>|t|)"][2]) #get the p value for dep binarized
p_gam <- sapply(CNB_cog_score_stats_gam_dep_binarized, function(v) summary(v)$p.table[2,4])

#Convert to data frame
p_lm <- as.data.frame(p_lm)
p_gam <- as.data.frame(p_gam)

#Print original p-values to three decimal places
p_round_lm <- round(p_lm,3)
p_round_gam <- round(p_gam,3)

#FDR correct p-values
pfdr_lm <- p.adjust(p_lm[,1],method="fdr")
pfdr_gam <- p.adjust(p_gam[,1],method="fdr")

#Convert to data frame
pfdr_lm <- as.data.frame(pfdr_lm)
row.names(pfdr_lm) <- cnb_measure_names
pfdr_gam <- as.data.frame(pfdr_gam)
row.names(pfdr_gam) <- cnb_measure_names

#To print fdr-corrected p-values to three decimal places
pfdr_round_lm <- round(pfdr_lm,3)
pfdr_round_gam <- round(pfdr_gam,3)

#List the NMF components that survive FDR correction
CNB_fdr_lm <- row.names(pfdr_lm)[pfdr_lm<0.05]
CNB_fdr_gam <- row.names(pfdr_gam)[pfdr_gam<0.05]

#make a data frame with names and fdr values (rounded to 3 decimals)
CNB_names_and_fdr_values_lm <- data.frame(cbind(CNB_fdr_lm, round(pfdr_lm[pfdr_lm<0.05],3)))
CNB_names_and_fdr_values_gam <- data.frame(cbind(CNB_fdr_gam, round(pfdr_gam[pfdr_gam<0.05],3)))

#add titles to names_and_fdr tables
names(CNB_names_and_fdr_values_lm) <- c("CNB_measure", "p_FDR_corr")
names(CNB_names_and_fdr_values_gam) <- c("CNB_measure", "p_FDR_corr")

#write the results of the mass univariate stats to files
write.csv(CNB_names_and_fdr_values_lm, file = "/data/joy/BBL/projects/ballerDepHeterogen/results/n3284_dep776_nondep2508_mass_univ_FDR_corrected_lm_20171211.csv")
write.csv(CNB_names_and_fdr_values_gam, file = "/data/joy/BBL/projects/ballerDepHeterogen/results/n3284_dep776_nondep2508_mass_univ_FDR_corrected_gam_20171211.csv")

#checkmodel with visreg
#lapply(CNB_cog_score_stats_lm_dep_binarized, function(x) {visreg(x)}) 
#lapply(CNB_cog_score_stats_gam_dep_binarized, function(x) {visreg(x)})

#Make table 1 (demographics)
#subset demographics
subset_demographics_for_table1 <- data.frame(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex, subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$race, subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$medu1, subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years, subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$dep_binarized)
listVars <- c("Gender", "Race", "Maternal Ed", "Age") #Gender 1 = male, 2 = female, Race 1 = caucasian, Maternal Ed = years, age = years, dep 1 = dep, 0 = non_dep
names(subset_demographics_for_table1) <- c(listVars, "Depression")

#Change categorical values to have names
subset_demographics_for_table1$Gender <- ifelse(subset_demographics_for_table1$Gender == 1, "Male", "Female")
subset_demographics_for_table1$Depression <- ifelse(subset_demographics_for_table1$Depression == 1, "Depressed", "Non-depressed")
subset_demographics_for_table1$Race <- ifelse(subset_demographics_for_table1$Race == "1", "Caucasian","Non-caucasian")

#make variable list
table_titles <- c("Non-depressed", "depressed", "p-value")

#Define Categorical Variables
cat_variables <- c("Gender", "Race", "Depression")

#create demographics table
demographics_table <- CreateTableOne(vars = listVars, data = subset_demographics_for_table1, factorVars = cat_variables, strata = c("Depression"))
#print(demographics_table, showAllLevels = TRUE)

```
## Including CNB measures that are FDR corrected

```{r CNB_names_and_fdr_values}
print(CNB_names_and_fdr_values_lm)
print(CNB_names_and_fdr_values_gam)
```

## Including Tables

Table 1: Demographics no levels shown
```{r R_demo, echo=TRUE}

print(demographics_table)
```

Table 1: Demographics, levels shown
```{r R_demo_levels, echo=TRUE}
print(demographics_table, showAllLevels = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/20180104_group_selection_and_demo_by_gender_matching_age_and_race.R

@@ -1,0 +1,252 @@
+#install.packages("visreg")
+#install.packages("mgcv")
+#install.packages("tableone")
+#################################################
+#### Get subset for Baller_DepHeterogeneity ####
+#################################################
+
+#This does the subsetting, runs statistics with lapply instead of loop, visualizes progress with visreg, and does fdr correction
+# it then makes table 1 (demographics)
+#load libraries
+library(visreg)
+library(mgcv)
+library(tableone)
+
+#read in csvs
+demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
+cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
+health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
+psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
+
+#remove people with NA for race, age, or sex.  START WITH N = 9498
+demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
+demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
+demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
[…]

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/DataPrep.R

@@ -1,0 +1,130 @@
+#################
+### LOAD DATA ###
+#################
+
+#Demographic data (n=1629)
+data.demo <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/demographics/n1601_demographics_go1_20161212.csv", header=TRUE) 
+
+##Clinical data
+#Screening diagnoses (n=1601)
+data.diag <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/clinical/n1601_goassess_psych_summary_vars_20131014.csv", header=TRUE, na.strings=".")
+
+#Bifactors and correlated traits (n=1601)
+data.factors <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/clinical/n1601_goassess_clinical_factor_scores_20161212.csv", header=TRUE, na.strings=".")
+
+#State trait anxiety data (n=1391)
+data.stai <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/clinical/n1601_stai_pre_post_itemwise_smry_factors_20170131.csv", header=TRUE, na.strings=".")
+
+#Exclusion data (n=1601)
+#Health exclusion (use the new healthExcludev2 variable)
+data.healthExclude <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/health/n1601_health_20170421.csv", header=TRUE, na.strings=".")
+
+#T1 QA exclusion (n=1601)
+data.t1QA <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/neuroimaging/t1struct/n1601_t1QaData_20170306.csv", header=TRUE, na.strings=".")
+
+##################
+#### DATA PREP ###
+##################
+
+#Transform the age variable from months to years
+data.demo$age <- (data.demo$ageAtScan1)/12
+
+#Recode male as 0 and female as 1
+data.demo$sex[which(data.demo$sex==1)] <- 0
+data.demo$sex[which(data.demo$sex==2)] <- 1
+
+##################
+### MERGE DATA ###
+##################
+dataMerge1 <-merge(data.demo,data.diag, by=c("bblid","scanid"), all=TRUE) 
+dataMerge2 <-merge(dataMerge1,data.factors, by=c("bblid","scanid"), all=TRUE) 
+dataMerge3 <-merge(dataMerge2,data.stai, by=c("bblid","scanid"), all=TRUE) 
[…]

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/GamAnalyses.R

@@ -1,0 +1,45 @@
+e##########################################
+#### GAM MODELS FOR PREMATURITY STUDY ####
+##########################################
+
+#Load data
+data.NMF <- read.csv("/data/joy/BBL/projects/pncPreterm/subjectData/n282_Prematurity_allData.csv", header=TRUE, na.strings = "NA")
+
+#Remove those who are missing maternal level of education data
+data.NMF <- data.NMF[!is.na(data.NMF$medu1),]
+
+#Load library
+library(mgcv)
+
+#Get NMF variable names
+nmfComponents <- names(data.NMF)[grep("Nmf26",names(data.NMF))]
+
+#Run gam models
+NmfModels <- lapply(nmfComponents, function(x) {
+  gam(substitute(i ~ s(age) + sex + medu1 + ga, list(i = as.name(x))), method="REML", data = data.NMF)
+})
+
+#Look at model summaries
+models <- lapply(NmfModels, summary)
+
+#Pull p-values
+p <- sapply(NmfModels, function(v) summary(v)$p.table[4,4])
+
+#Convert to data frame
+p <- as.data.frame(p)
+
+#Print original p-values to three decimal places
+p_round <- round(p,3)
+
+#FDR correct p-values
+pfdr <- p.adjust(p[,1],method="fdr")
+
+#Convert to data frame
+pfdr <- as.data.frame(pfdr)
+
+#To print fdr-corrected p-values to three decimal places
+pfdr_round <- round(pfdr,3)
+
+#List the NMF components that survive FDR correction
+Nmf_fdr <- row.names(pfdr)[pfdr<0.05]
+

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra20180115.R

@@ -1,0 +1,110 @@
+library(visreg)
+library(mgcv)
+library(tableone)
+library(dplyr)
+library(plm)
+library(MatchIt)
+
+#read in csvs
+demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
+cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
+health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
+psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
+
+#remove people with NA for race, age, or sex.  START WITH N = 9498
+demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
+demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
+demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
+demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406
+
+#remove people with NA for depression or total psych score, START WITH N = 9498
+psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
[…]

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Sunny_matching_script.R

@@ -1,0 +1,412 @@
+library(MASS)
+library(Matching)
+library(tidyr)
+library(readr)
+library(effsize)
+library(dplyr)
+
+###Loading Data
+alldata<-read.csv("OlfactionFromStata_2017_06-19.csv") %>% 
+  group_by(group) %>%
+  arrange(group,bblid)
+View(alldata)
+str(alldata)
+colnames(alldata)
+
+#Cleaning up variables
+alldata$anyspectrum<-as.factor(alldata$anyspectrum)
+
+###Calculating and Z-transforming CNB Categories 
+# Creating normalized summary variables
+zacexe <- as.vector(
+  scale(
+    rowMeans(
+      data.frame(alldata$zaclnb, alldata$zacpcet, alldata$zacpcpt), na.rm = TRUE)
+  ))
+
+zacmem <- as.vector(
+  scale(
+    rowMeans(
+      data.frame(alldata$zaccpf, alldata$zaccpw, alldata$zacvolt), na.rm = TRUE)
+  ))
+
+zaccog <- as.vector(
+  scale(
+    rowMeans(
+      data.frame(alldata$zacpvrt, alldata$zacpmat, alldata$zacplot), na.rm = TRUE)
+  ))
+
+zacsoc <- as.vector(
+  scale(
+    rowMeans(
+      data.frame(alldata$zacadt, alldata$zacer, alldata$zacedf), na.rm = TRUE)
+  ))
+
+zacall <- as.vector(
+  scale(
+    rowMeans(
+      data.frame(zacexe, zacmem, zaccog, zacsoc), na.rm = TRUE)
+  ))
+
+#Merging with existing data
+alldata$zacexe = zacexe
+alldata$zacmem = zacmem
+alldata$zaccog = zaccog
+alldata$zacsoc = zacsoc
+alldata$zacall = zacall
+
+alldata<-data.frame(alldata,zacall)
+alldata$zacall.1 <- NULL
+
+#Checking
+colnames(alldata)
+summary(alldata$zacall)
+sd(alldata$zacall, na.rm = TRUE)
+
+###Selecting Important Variables for Analysis
+dat <- select(alldata, bblid, group, sex, race, age, edu, avgedu, hand, smoker, anysmoke,
+              id, disc, lyral, cit, 
+              zacexe, zacmem, zaccog, zacsoc, zacall, wratstd, gafc, comtgenotype,
+              positive, negative, sipstot, anyspectrum, psyscale)
+
+View(dat)
+table(dat$group)
+
[…]

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/subsetting_for_initial_group.R

@@ -1,0 +1,123 @@
+#install.packages("visreg")
+#################################################
+#### Get subset for Baller_Dep_Heterogeneity ####
+#################################################
+
+#load libraries
+library("visreg")
+library(mgcv)
+
+#read in csvs
+demographics <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",")
+cnb_scores <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",")
+health <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",")
+psych_summary <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",")
+
+#remove people with NA for race, age, or sex.  START WITH N = 9498
+demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
+demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
+demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
+demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406
+
+#remove people with NA for depression or total psych score, START WITH N = 9498
+psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
+psych_summary_no_NA_dep_and_smry_psych_overall <- psych_summary_no_NA_dep[!is.na(psych_summary_no_NA_dep$smry_psych_overall_rtg),] #take out those with NA for overall psych rtg, no additional people lost, N = 9411
+
+#merge the csvs
+#dem_cnb <- merge(demographics, cnb_scores, by = "bblid") #merge demographics and cnb #this is if we want to include people without full demographic data
[…]

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/subsetting_for_initial_group_with_lapply.R

@@ -1,0 +1,155 @@
+#install.packages("visreg")
+#install.packages("mgcv")
+#install.packages("tableone")
+#################################################
+#### Get subset for Baller_DepHeterogeneity ####
+#################################################
+
+#This does the subsetting, runs statistics with lapply instead of loop, visualizes progress with visreg, and does fdr correction
+# it then makes table 1 (demographics)
+#load libraries
+library(visreg)
+library(mgcv)
+library(tableone)
+
+#read in csvs
+demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
+cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
+health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
+psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
+
+#remove people with NA for race, age, or sex.  START WITH N = 9498
+demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
+demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
+demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
[…]

### 2018-01-22 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/subsetting_for_initial_group_with_lapply_server_path.R

@@ -1,0 +1,152 @@
+#install.packages("visreg")
+#install.packages("mgcv")
+#install.packages("tableone")
+#################################################
+#### Get subset for Baller_DepHeterogeneity ####
+#################################################
+
+#This does the subsetting, runs statistics with lapply instead of loop, visualizes progress with visreg, and does fdr correction
+# it then makes table 1 (demographics)
+#load libraries
+library(visreg)
+library(mgcv)
+library(tableone)
+
+#read in csvs
+demographics <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
+cnb_scores <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
+health <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
+psych_summary <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
+
+#remove people with NA for race, age, or sex.  START WITH N = 9498
+demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
+demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
+demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
[…]

### 2018-02-12 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra_rmarkdown_20180122.Rmd
updates including hydra prep scripts

---
title: "Prep_for_hyrda_with_matching"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
The purpose of this script is to read in CNB data, subset it, merge it, and write files to use in Hydra.  Before doing hydra, however, we must address the vastly different ages, maternal education and race between depressed and control groups.  Unfortunately, the matching removes a high number of our patients.

```{r subset, echo = TRUE}
library(visreg)
library(mgcv)
library(tableone)
library(dplyr)
library(plm)
library(MatchIt)

#read in csvs
demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv

#remove people with NA for race, age, or sex.  START WITH N = 9498
demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406

#remove people with NA for depression or total psych score, START WITH N = 9498
psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
psych_summary_no_NA_dep_and_smry_psych_overall <- psych_summary_no_NA_dep[!is.na(psych_summary_no_NA_dep$smry_psych_overall_rtg),] #take out those with NA for overall psych rtg, no additional people lost, N = 9411

#merge the csvs
#merge demographics and cnb #this is if we want to include people without full demographic data
dem_cnb <- merge(demographics_noNA_race_age_andCNBage_sex, cnb_scores, by = "bblid") #merge demographics and cnb, N = 9406
psych_health <- merge(psych_summary_no_NA_dep_and_smry_psych_overall, health, by = "bblid") #merge psych and health, N = 9411
dem_cnb_psych_health_merged <- merge (dem_cnb, psych_health, by = "bblid") #merge all 4 csvs, lost 1 person [134716] (had demographics, but no psych ratings): N = 9405

#make subsets
subset_just_dep_and_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & (smry_dep == 4), select = "bblid") #subset people who were not medically excluded and who are depressed, N = 776
subset_no_psych_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & (smry_psych_overall_rtg < 4), select = "bblid") #subset people who are psychiatrically healthy, N = 2508
subset_dep_or_no_psych_and_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & ((smry_dep == 4) | (smry_psych_overall_rtg <4))) #subset including both depressed and healthies, good for regressions, N = 3284

#would binarize depression smry score to -1 (less than 4, not depressed) and 1 (score 4 , depressed)
dep_binarized <- ifelse(subset_dep_or_no_psych_and_no_medicalratingExclude$smry_dep == 4, 1, -1)
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED <- cbind(subset_dep_or_no_psych_and_no_medicalratingExclude, dep_binarized) #N = 3284

#make depression and gender into factor scores
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$dep_binarized <- as.factor(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$dep_binarized)
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex <- as.factor(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex)

#divide ageAtCNB by 12 for age
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years <- subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$ageAtCnb1/12

#age demeaned and squared, from Toni
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$ageSq <- as.numeric(I(scale(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years, scale = FALSE, center = TRUE)^2))

#Subset only variables needed for hydra analysis 
#(BBLID, cognitive variables, depression), also do by males(1555)/females(1729) separately
subset_bblidAndCog_features <- data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[14:39], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[77]))
subset_bblidAndCog_features_males <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[14:39], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[77])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 1)
subset_bblidAndCog_features_females <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[14:39], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[77])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 2)

#subset of covariates (BBLID, sex, age in years), also do by males/females
subset_bblidAndCovariates <- data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1:2], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[78]))
subset_bblidAndCovariates_males <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1:2], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[78])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 1)
subset_bblidAndCovariates_females <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1:2], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[78])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 2)

#save files for hyd
write.csv(subset_bblidAndCog_features, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_all_gender/Features.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCovariates, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_all_gender/Covariates.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCog_features_males, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_males/Features.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCovariates_males, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_males/Covariates.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCog_features_females, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_females/Features.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCovariates_females, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_females/Covariates.csv", row.names = FALSE, quote = FALSE)

```

## Including Plots

###

```{r pressure, echo=TRUE}
#match with matchit (starting n = 3284, males = 1555, females 1729)
data.unmatched = subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED
data.unmatched$unmatchedRows =rownames(data.unmatched)
dataset = data.unmatched
# Some preprocessing
dataset = dplyr::select(dataset, sex, age_in_years, ageSq, medu1, race, dep_binarized, unmatchedRows)
#dataset = dplyr::filter(dataset, !is.na(group))
# Dep: 1, Health = 0 
dataset$dep_binarized = 1*(dataset$dep_binarized==1) 
#"male": 1, "female": 0
dataset$sex = 1*(dataset$sex==1)
# Remove subjects with NA for maternal edu, new N = 3256, males = 1539, females 1717
dataset <- dataset[!is.na(dataset$medu1),]

# Plot prematch
plot(dataset$age_in_years,jitter(dataset$medu1, factor=3), col=2*dataset$race+dataset$dep_binarized+1,pch=16, ylab="Maternal Edu", xlab="Age")
legend("bottomright",c("Non-white, non-depressed", "Non-white, depressed", "White, non-depressed", "White, depressed"),pch=rep(16, 4), col=c(1,2,3,4))

#GAM for propensity score
ps.model =gam(dep_binarized ~s(age_in_years) +s(medu1) + race + sex, data=dataset, family=binomial)
ps =exp(predict(ps.model))/(1 +exp(predict(ps.model)))
m.out <-matchit(dep_binarized ~ age_in_years, data=dataset, method="nearest", exact=c("race", "medu1", "sex"), distance="mahalanobis")
plot(m.out)

#return the matched dataset. N = 1518, males = 518, females = 1000
m.data <- match.data(m.out)

# Test for significant difference in age between groups
t.test(age_in_years~dep_binarized, data=m.data)
t.test(race~dep_binarized, data=m.data)
t.test(sex~dep_binarized, data=m.data)
t.test(medu1~dep_binarized, data=m.data)

# Re-plot
plot(m.data$age_in_years,jitter(m.data$medu1, factor=3), col=2*m.data$race+m.data$dep_binarized+1,pch=16, ylab="Maternal Edu", xlab="Age")
legend("bottomright",c("Non-white, non-depressed", "Non-white, depressed", "White, non-depressed", "White, depressed"),pch=rep(16, 4), col=c(1,2,3,4))

# Make the final matched data set
data.matched = data.unmatched[data.unmatched$unmatchedRows%in%m.data$unmatchedRows,]
data.matched$unmatchedRows = NULL

saveRDS(data.matched, file='/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydraMatched_age_race_medu1_sex_20180115_matched.rds')
```

### 2018-02-12 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/20180104_group_selection_and_demo_by_gender_matching_age_and_race.R

@@ -157,7 +157,7 @@ subset_demographics_for_table1_females$Race <- ifelse(subset_demographics_for_ta
 
 #make variable list
 table_titles_males <- c("Males Non-depressed", "Males depressed", "p-value")
-table_titles_males <- c("Females Non-depressed", "Females depressed", "p-value")
+table_titles_females <- c("Females Non-depressed", "Females depressed", "p-value")
 
 #Define Categorical Variables
 cat_variables <- c("Race", "Depression")

### 2018-02-12 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra20180124.R

@@ -46,10 +46,49 @@ subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years <- 
 #age demeaned and squared, from Toni
 subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$ageSq <- as.numeric(I(scale(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years, scale = FALSE, center = TRUE)^2))
 
-#race binarized for plotting purposes
+#race binarized for plotting purposes, caucasian 1, non-caucasion 0
 subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$race_binarized <- ifelse(as.factor(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$race) == 1, 1, 0)
 
-#Subset only variables needed for hydra analysis 
+#remove people with NA in their cognitive measures
+subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED <- subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[complete.cases(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[14:39]),]
+
+#GAM to get residuals for hydra since too hard to match on age and race
+cnb_measure_names <- names(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED)[grep("_z", names(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED))] #get the names of all the columns with _z in the name
+CNB_cog_scores_residuals <- subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED
+
+CNB_cog_score_stats_gam <- lapply(cnb_measure_names, function(x) 
+{
+  #gam(substitute(i ~ s(age_in_years) + race_binarized, list(i = as.name(x))), data = subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED)
+  gam(substitute(i ~ s(age_in_years) + sex, list(i = as.name(x))), data = subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED)
+  
+}) 
+
+names(CNB_cog_score_stats_gam) <- cnb_measure_names
+
+#set CNB_cog_score_residuals to original data frame and remove NAs
[…]

### 2018-02-12 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra_20180212.R

@@ -5,6 +5,13 @@ library(dplyr)
 library(plm)
 library(MatchIt)
 
+
+
+###### This script reads in demographics, cnb_scores, health and psych summaries, merges them, removes NAs, codes and separates by depression#####
+####Also preps for hydra, both using a typical GAM model and matching (we lose a lot of people) and also with residuals plotted so we don't have to match#########
+########Also provides matched data sets and tests them, if we decide to use them.  It significantly reduces N to match (dataset from 3022 to 1424)
+
+
 #read in csvs
 demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
 cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
@@ -46,10 +53,15 @@ subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years <- 
 #age demeaned and squared, from Toni
 subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$ageSq <- as.numeric(I(scale(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years, scale = FALSE, center = TRUE)^2))
 
-#race binarized for plotting purposes
+#race binarized for plotting purposes, caucasian 1, non-caucasion 0
 subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$race_binarized <- ifelse(as.factor(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$race) == 1, 1, 0)
 
-#Subset only variables needed for hydra analysis 
+#remove people with NA in their cognitive measures
[…]

### 2020-10-29 dac_summary.Rmd
Add files via upload

---
title: "Dac Summary"
output:
  html_document
---

## Dac summary
First run on: 20201008
Spreadsheet:

This script goes through data that has been returned from the DAC and summarizes it
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r summary}
#preprocessing - because these are sent

#Load original MS database from melissa
#load("/Users/eballer/BBL/msdepression/data/database_2019-03-04.RData")

#read in melissa's xls: n=30915
melissa_xls <- read.csv("/Users/eballer/BBL/msdepression/data/mri_file_upd_w_empi.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"))

#load our df
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninpatientswithmsmaskdataaddedmoredxcodes.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"))

data2 <- read.csv("/Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/investigatingdepressioninpatientswithmultiplesclerosis10-23medsinonerow (1).csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"))
#subset EMPI, accession codes, F-codes, Phq2 and phq9 #n = 42101
data_empi_acc_f_phq <- data#subset(data, select = c("EMPI", "ACCESSION_NUM", "PHQ.2", "PHQ.9"))

##########################
### some preprocessing ###
##########################

#change accession name for ease of merging with Rdata if using
#names(data_empi_acc_f_phq) <- gsub(pattern = "ACCESSION_NUM", replacement = "AccessionNumber", names(data_empi_acc_f_phq))

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
data_empi_acc_f_phq[28:42] <- apply(X = data_empi_acc_f_phq[28:42], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)

#get empis of people who have wacky deidentified empis
db_empi_acc <- subset(melissa_xls, select = c("ACCESSION_NUM", "EMPI"))

#change EMPI to "patient ID"
names(db_empi_acc) <- gsub(pattern = "EMPI", replacement = "PatientID", names(db_empi_acc))
                      #subset(database, select = c("AccessionNumber", "PatientID"))

#merge with our database, PatientID is De-identified EMPI, n = 30799
data_empi_acc_f_phq_ptID <- merge(data_empi_acc_f_phq, db_empi_acc, by = "ACCESSION_NUM")

# number of unique scanners and scans
print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(which(!duplicated(data_empi_acc_f_phq$PROC_CODE)))))
print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(which(!duplicated(data_empi_acc_f_phq$DEPARTMENT_NAME)))))
print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(which(!duplicated(data_empi_acc_f_phq$MODALITY)))))

#phq-2;
print(paste0("N PHQ-2 = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))))

#print histogram of phq-2s for people

hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

hist(data_empi_acc_f_phq_ptID$PHQ.2[which(!is.na(data_empi_acc_f_phq_ptID$PHQ.2))], main = paste0("PHQ-2 Histo in Melissa's subgroup #: n = ", length(which(!is.na(data_empi_acc_f_phq_ptID$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

#phq-9
print(paste0("N PHQ-9 = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))))

#print histogram of phq-9s for people
hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

hist(data_empi_acc_f_phq_ptID$PHQ.9[which(!is.na(data_empi_acc_f_phq_ptID$PHQ.9))], main = paste0("PHQ-9 Histo In Melissa's group: n = ", length(which(!is.na(data_empi_acc_f_phq_ptID$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

#both phq2 and 9

num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
print(num_both_phq2_and_9)

#looking at just people with phq-2 and 9
phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=5855
phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n=606

#checking logic for order of subsetting - unique EMPI then ph2/9 == phq2/9 then EMPI subsetting - this works
unique_phq2_empis <- length(which(!duplicated(phq2_subset$EMPI))) #n=797
unique_phq9_empis <- length(which(!duplicated(phq9_subset$EMPI))) #n=107

#age histogram
hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")

#tsh
hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")

#hemoglobin
hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")

#folate
hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)

#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#race
hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
     xlab = "white = 1, other = 2", 
     breaks = 2)

#subet people with unique EMPIs

#melissa's group
data_empi_acc_f_phq_unique_EMPI_m <- data_empi_acc_f_phq_ptID[!duplicated(data_empi_acc_f_phq_ptID$PatientID),]
print(paste0("N unique EMPIs in Melissa's group = ", dim(data_empi_acc_f_phq_unique_EMPI_m)[1]))

#my group
data_empi_acc_f_phq_unique_EMPI <- data_empi_acc_f_phq[!duplicated(data_empi_acc_f_phq$EMPI),]
print(paste0("N unique EMPIS in my group = ", dim(data_empi_acc_f_phq_unique_EMPI)[1]))

#histos
hist(data_empi_acc_f_phq_unique_EMPI$PHQ.2[which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.2))], main = paste0("PHQ-2 Histo my group w/unique EMPI : n = ", length(which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.2)))), xlab = "PHQ-2", breaks = 6)

hist(data_empi_acc_f_phq_unique_EMPI$PHQ.9[which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.9))], main = paste0("PHQ-9 Histo my group w/unique EMPI : n = ", length(which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

#age histogram
hist(data_empi_acc_f_phq_unique_EMPI$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq_unique_EMPI$MRI_ENC_AGE))], main = paste0("Age Histo my group w/unique EMPI: n = ", length(which(!is.na(data_empi_acc_f_phq_unique_EMPI$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")

```

### 2020-10-29 lab_notebook.Rmd
Add files via upload

---
title: "R Notebook for MSDepression Project"
output:
  html_notebook
---

#### 20201008

First, need to summarize data

1) Got data into csv format
  - Opened /Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/*xls in excel and saved as csv
2)
```{move_data bash}
cd /Users/eballer/BBL/msdepression/data/
mkdir dac
cp /Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/*csv .
```

3) In R, started /Users/eballer/BBL/msdepression/dac_summary.Rmd to summarize data

** during the last iteration of de-identifying the DAC, the PHQ-9 scores all vanished. Will need to ask for them to be re-added.
** also, number of people dropped by 50% (from 36277 to 16383)

#### 20201012

Continued with descriptives

  - Read in Melissa's data sheet.
  - subsetted EMPI and Accession # (since our EMPI's are deidentified)
  - Merged with my datasheet on Accession #
  - "PatientID == EMPI"

Merged Melissa's data with mine to get empis.

#### 20201013

[x] Figure out how many people with repeat mrns - to remove doubles. Will need to remove duplicates.

#### 20201027

Going back and forth a lot with DAC people, should have a new update soon. Big thing will be to parse PHQ2/9. We will get all of them but have to tie them to their appropriate scans

### 2020-11-05 dac_summary.Rmd
Add files via upload
- added details combining my and melissa's data. Made lots of changes to unique/duplicated code.

@@
 #load our df
-data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninpatientswithmsmaskdataaddedmoredxcodes.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"))
+data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninpatientswithmsmaskdataaddedmoredxcodes.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL")) #n=42101, with 42044 unique accession codes
 
 data2 <- read.csv("/Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/investigatingdepressioninpatientswithmultiplesclerosis10-23medsinonerow (1).csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"))
-#subset EMPI, accession codes, F-codes, Phq2 and phq9 #n = 42101
-data_empi_acc_f_phq <- data#subset(data, select = c("EMPI", "ACCESSION_NUM", "PHQ.2", "PHQ.9"))
+#subset EMPI, accession codes, F-codes, Phq2 and phq9 #n = 29226, with 21378 unique accession codes
+
+#drawn from Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninpatientswithmsmaskdataaddedmoredxcodes.csv, which was split into 2 b/c the spreadsheet was too big 
+data3_p1 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/20201029_msdep_pt1.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep = ",") # n = 28556 with 20871 unique accession codes
+
+data3_p2 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/20201029_msdep_pt2.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep = ",")
+
+data3 <- rbind(data3_p1, data3_p2)
+write.csv(x = data3, file = "/Users/eballer/BBL/msdepression/data/dac/msdep_20201029_merged.csv")
+
+#set data_empi_acc_f_phq to data3
+data_empi_acc_f_phq <- data3#subset(data, select = c("EMPI", "ACCESSION_NUM", "PHQ.2", "PHQ.9"))
+
+#number of unique in melissa's
+m_access_n <- length(unique(melissa_xls$ACCESSION_NUM)) #n = 29183
+
+#number of unique in my group
+eb_access_n <- length(unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 21027
+
+#number of overlap
+
+melissa_peeps_in_eb_group <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 15137
+
+eb_peeps_in_melissa_group <- sum(unique(data_empi_acc_f_phq$ACCESSION_NUM) %in% unique(melissa_xls$ACCESSION_NUM)) #n=15137
+
+print(paste0("Number of Melissa's people in EB group : ", melissa_peeps_in_eb_group, " out of ", m_access_n, " unique accession numbers")) #suggesting that Melissa has a lot of people that don't have ms defined the same way we do
+
+print(paste0("Number of EB's people in Melissa's group : ", eb_peeps_in_melissa_group, " out of ", eb_access_n, " unique accession numbers"))
+
+overlap_group <- data_empi_acc_f_phq[which(unique(data_empi_acc_f_phq$ACCESSION_NUM) %in% unique(melissa_xls$ACCESSION_NUM)),] #n=15137
 
@@
 # number of unique scanners and scans
-print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(which(!duplicated(data_empi_acc_f_phq$PROC_CODE)))))
-print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(which(!duplicated(data_empi_acc_f_phq$DEPARTMENT_NAME)))))
-print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(which(!duplicated(data_empi_acc_f_phq$MODALITY)))))
+print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE))))
+print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME))))
+print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY))))
 
 #phq-2;
-print(paste0("N PHQ-2 = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))))
+print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2))))
 
@@
 #phq-9
-print(paste0("N PHQ-9 = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))))
+print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9))))
 
@@
 num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
-print(num_both_phq2_and_9)
+print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9))
 
@@
 #checking logic for order of subsetting - unique EMPI then ph2/9 == phq2/9 then EMPI subsetting - this works
-unique_phq2_empis <- length(which(!duplicated(phq2_subset$EMPI))) #n=797
-unique_phq9_empis <- length(which(!duplicated(phq9_subset$EMPI))) #n=107
+unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=1005
+unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=70
 
@@
 
-#subet people with unique EMPIs
+#f-codes - people usually have BOTH icd9/10 or neither
+num_depression <- length(which(!is.na(data_empi_acc_f_phq$ICD9))) #n=11131
+print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
 
-#melissa's group
-data_empi_acc_f_phq_unique_EMPI_m <- data_empi_acc_f_phq_ptID[!duplicated(data_empi_acc_f_phq_ptID$PatientID),]
-print(paste0("N unique EMPIs in Melissa's group = ", dim(data_empi_acc_f_phq_unique_EMPI_m)[1]))
+num_unique_EMPI_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & unique(data_empi_acc_f_phq$EMPI))) #n = 593
+print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep))
 
-#my group
-data_empi_acc_f_phq_unique_EMPI <- data_empi_acc_f_phq[!duplicated(data_empi_acc_f_phq$EMPI),]
-print(paste0("N unique EMPIS in my group = ", dim(data_empi_acc_f_phq_unique_EMPI)[1]))
+num_phq2_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2))) #n = 4855
+print(paste0("N with ICD 9/10 codes for depression and phq2 = ", num_phq2_and_dep))
+
+num_phq9_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9))) #n=669
+print(paste0("N with ICD 9/10 codes for depression and phq9 = ", num_phq9_and_dep))
+
+#fcode histos
+hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2))], 
+     main = paste0("PHQ2 and Dep Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = paste0("PHQ2"))
+
+hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9))], 
+     main = paste0("PHQ9 and Dep Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = paste0("PHQ9"))
+
+############ 11/5/2020 #############
+#########################
+##### overlap group #####
+#########################
+
+#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
+overlap_group$ACCESSION_NUM <- as.integer(overlap_group$ACCESSION_NUM)
+overlap_group[28:42] <- apply(X = overlap_group[28:42], FUN = as.integer, MARGIN = 2)
+overlap_group$sex_binarized <- ifelse(overlap_group$SEX == "MALE", 1, 2)
+overlap_group$race_binarized <- ifelse(overlap_group$RACE == "WHITE", 1, 2)
+
+#remove people who do not have phq2 or 9
+overlap_group_with_phq2 <- overlap_group[which(!is.na(overlap_group$PHQ.2)),] #4864
+overlap_group_with_phq9 <- overlap_group[which(!is.na(overlap_group$PHQ.9)),] #376
+
+#subset people with unique scans (these are one time scanners)
+overlap_group_no_duplicate_EMPI <- overlap_group[which(!duplicated(overlap_group$EMPI)),] #2675
+
+overlap_group_with_phq2_no_duplicate_EMPI <- overlap_group_with_phq2[which(!duplicated(overlap_group_with_phq2$EMPI)),] #670
+
+overlap_group_with_phq9_no_duplicate_EMPI <- overlap_group_with_phq9[which(!duplicated(overlap_group_with_phq9$EMPI)),] #44
+
+#####take duplicated people and take first instance
+
+#get duplicated people
+overlap_group_with_duplicate_EMPI <- overlap_group[which(duplicated(overlap_group$EMPI)),]
+
+overlap_group_with_phq2_w_duplicate_EMPI <- overlap_group_with_phq2[which(duplicated(overlap_group_with_phq2$EMPI)),]
+
+overlap_group_with_phq9_w_duplicate_EMPI <- overlap_group_with_phq9[which(duplicated(overlap_group_with_phq9$EMPI)),]
+
+#get match - i.e. first instance of each
+overlap_group_with_duplicated_EMPI_first_instance <- overlap_group_with_duplicate_EMPI[match(unique(overlap_group_with_duplicate_EMPI$EMPI), overlap_group_with_duplicate_EMPI$EMPI),]
+
+overlap_group_with_phq2_duplicated_EMPI_first_instance <- overlap_group_with_phq2_w_duplicate_EMPI[match(unique(overlap_group_with_phq2_w_duplicate_EMPI$EMPI), overlap_group_with_phq2_w_duplicate_EMPI$EMPI),]
+
+overlap_group_with_phq9_duplicated_EMPI_first_instance <- overlap_group_with_phq9_w_duplicate_EMPI[match(unique(overlap_group_with_phq9_w_duplicate_EMPI$EMPI), overlap_group_with_phq9_w_duplicate_EMPI$EMPI),]
+
+#combine groups
+overlap_group_unique_EMPI_combined <- rbind(overlap_group_no_duplicate_EMPI, overlap_group_with_duplicated_EMPI_first_instance)
+
+overlap_group_phq2_unique_empi_combined <- rbind(overlap_group_with_phq2_no_duplicate_EMPI, overlap_group_with_phq2_duplicated_EMPI_first_instance)
+
+overlap_group_phq9_unique_empi_combined <- rbind(overlap_group_with_phq9_no_duplicate_EMPI, overlap_group_with_phq9_duplicated_EMPI_first_instance)
+
+#summarize
+print(paste0("N unique EMPIs in EB and Melissa overlap group = ", length(unique(overlap_group$EMPI)), " out of ", dim(overlap_group)[1])) #
+
+print(paste0("N unique EMPIs in EB and Melissa overlap group w/phq2 = ", dim(overlap_group_phq2_unique_empi_combined)[1], " out of ", dim(overlap_group)[1])) #
+
+print(paste0("N unique EMPIs in EB and Melissa overlap group w/phq9 = ", dim(overlap_group_phq9_unique_empi_combined)[1], " out of ", dim(overlap_group)[1])) #
 
 #histos
-hist(data_empi_acc_f_phq_unique_EMPI$PHQ.2[which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.2))], main = paste0("PHQ-2 Histo my group w/unique EMPI : n = ", length(which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.2)))), xlab = "PHQ-2", breaks = 6)
+hist(overlap_group_phq2_unique_empi_combined$PHQ.2, main = paste0("PHQ-2 Histo combined group w/unique EMPI vs all EMPI: n = ", dim(overlap_group_phq2_unique_empi_combined)[1], "/", length(which(!is.na(overlap_group$PHQ.2)))), xlab = "PHQ-2", breaks = 6)
 
-hist(data_empi_acc_f_phq_unique_EMPI$PHQ.9[which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.9))], main = paste0("PHQ-9 Histo my group w/unique EMPI : n = ", length(which(!is.na(data_empi_acc_f_phq_unique_EMPI$PHQ.9)))), xlab = "PHQ-9", breaks = 30)
+#histos
+hist(overlap_group_phq9_unique_empi_combined$PHQ.9, main = paste0("PHQ-9 Histo combined group w/unique EMPI vs all EMPI: n = ", dim(overlap_group_phq9_unique_empi_combined)[1], "/", length(which(!is.na(overlap_group$PHQ.9)))), xlab = "PHQ-9", breaks = 6)
 
-#age histogram
-hist(data_empi_acc_f_phq_unique_EMPI$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq_unique_EMPI$MRI_ENC_AGE))], main = paste0("Age Histo my group w/unique EMPI: n = ", length(which(!is.na(data_empi_acc_f_phq_unique_EMPI$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")
+#########################
+### Erica's group  ####
+#########################
+data_e <- data3
+
+#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
+data_e$ACCESSION_NUM <- as.integer(data_e$ACCESSION_NUM)
+data_e[28:42] <- apply(X = data_e[28:42], FUN = as.integer, MARGIN = 2)
+data_e$sex_binarized <- ifelse(data_e$SEX == "MALE", 1, 2)
+data_e$race_binarized <- ifelse(data_e$RACE == "WHITE", 1, 2)
+
+#remove people who do not have phq2 or 9
+data_e_with_phq2 <- data_e[which(!is.na(data_e$PHQ.2)),] #4864
+data_e_with_phq9 <- data_e[which(!is.na(data_e$PHQ.9)),] #376
+
+#subset people with unique scans (these are one time scanners)
+data_e_no_duplicate_EMPI <- data_e[which(!duplicated(data_e$EMPI)),]
+
+data_e_with_phq2_no_duplicate_EMPI <- data_e_with_phq2[which(!duplicated(data_e_with_phq2$EMPI)),]
+
+data_e_with_phq9_no_duplicate_EMPI <- data_e_with_phq9[which(!duplicated(data_e_with_phq9$EMPI)),]
+
+#####take duplicated people and take first instance
+
+#get duplicated people
+data_e_with_duplicate_EMPI <- data_e[which(duplicated(data_e$EMPI)),]
+
+data_e_with_phq2_w_duplicate_EMPI <- data_e_with_phq2[which(duplicated(data_e_with_phq2$EMPI)),]
+
+data_e_with_phq9_w_duplicate_EMPI <- data_e_with_phq9[which(duplicated(data_e_with_phq9$EMPI)),]
+
+#get match - i.e. first instance of each
+data_e_with_duplicated_EMPI_first_instance <- data_e_with_duplicate_EMPI[match(unique(data_e_with_duplicate_EMPI$EMPI), data_e_with_duplicate_EMPI$EMPI),]
+
+data_e_with_phq2_duplicated_EMPI_first_instance <- data_e_with_phq2_w_duplicate_EMPI[match(unique(data_e_with_phq2_w_duplicate_EMPI$EMPI), data_e_with_phq2_w_duplicate_EMPI$EMPI),]
+
+data_e_with_phq9_duplicated_EMPI_first_instance <- data_e_with_phq9_w_duplicate_EMPI[match(unique(data_e_with_phq9_w_duplicate_EMPI$EMPI), data_e_with_phq9_w_duplicate_EMPI$EMPI),]
+
+#combine groups
+data_e_unique_EMPI_combined <- rbind(data_e_no_duplicate_EMPI, data_e_with_duplicated_EMPI_first_instance)
+
+data_e_phq2_unique_empi_combined <- rbind(data_e_with_phq2_no_duplicate_EMPI, data_e_with_phq2_duplicated_EMPI_first_instance)
+
+data_e_phq9_unique_empi_combined <- rbind(data_e_with_phq9_no_duplicate_EMPI, data_e_with_phq9_duplicated_EMPI_first_instance)
+
+#summarize
+#print(paste0("N unique EMPIs in EB group = ", dim(data_e_unique_EMPI_combined)[1], " out of ", dim(data_e)[1])) #
+
+print(paste0("N unique EMPIs in EB group = ", length(unique(data_e$EMPI)), " out of ", dim(data_e)[1]))
+
+print(paste0("N unique EMPIs in EB group w/phq2 = ", dim(data_e_phq2_unique_empi_combined)[1], " out of ", dim(data_e)[1])) #
+
+print(paste0("N unique EMPIs in EB group w/phq9 = ", dim(data_e_phq9_unique_empi_combined)[1], " out of ", dim(data_e)[1])) #
+
+print(paste0("N in EB group w/phq2 = ", length(which(!is.na(data_e_no_duplicate_EMPI$PHQ.2))), " out of ", dim(data_e)[1])) #
+
+print(paste0("N in EB group w/phq9 = ", length(which(!is.na(data_e_no_duplicate_EMPI$PHQ.9))), " out of ", dim(data_e)[1])) #
+
+print(paste0("N unique EMPIs in Melissa's group = ", length(unique(melissa_xls$EMPI)), " out of ", dim(melissa_xls)[1]))
+
+print(paste0("N in EB + Melissa overlap group w/depression ICD 9/10 = ", length(which(!is.na(overlap_group$ICD9))), " out of ", dim(overlap_group)[1])) #
+
+print(paste0("N in EB & Melissa's group with dep ICD 9/10 and phq2 = ", length(which(!is.na(overlap_group_with_phq2$ICD9))), " out of ", dim(overlap_group)[1]))
+
+print(paste0("N in EB & Melissa's group with dep ICD 9/10 and phq9 = ", length(which(!is.na(overlap_group_with_phq9$ICD9))), " out of ", dim(overlap_group)[1]))
+
+#histos
+hist(data_e_phq2_unique_empi_combined$PHQ.2, main = paste0("PHQ-2 Histo EB group w/unique EMPI vs all EMPI: n = ", dim(data_e_phq2_unique_empi_combined)[1], "/", length(which(!is.na(data_e$PHQ.2)))), xlab = "PHQ-2", breaks = 6)
+
+#histos
+hist(data_e_phq9_unique_empi_combined$PHQ.9, main = paste0("PHQ-9 Histo EB group w/unique EMPI vs all EMPI: n = ", dim(data_e_phq9_unique_empi_combined)[1], "/", length(which(!is.na(data_e$PHQ.9)))), xlab = "PHQ-9", breaks = 6)
 

### 2020-11-24 lab_notebook.Rmd
Add files via upload
update

@@
 Going back and forth a lot with DAC people, should have a new update soon. Big thing will be to parse PHQ2/9. We will get all of them but have to tie them to their appropriate scans
+
+#### 20201117
+
+Lots of back and forth with DAC, significantly less people in my pull that limited its pull to people who had G35 codes only. Went back to emails with Sunil (Taki involved). See email below.
+
+---
+
+I think my focus was more on the images and using the challenging radiology data.
+So I took a quick look at the codes and files that I might have sent or used. Unfortunately, I can’t remember what the last file I sent was and I don’t have access to the lothal drive anymore.
+
+Criteria:
+Don’t see any dates used. Guessing data probably was up to 7/2018(around the time I ran the report)
+I looked for patients with order associated diagnosis (%340% or %G35%)  which we mentioned that it is something that is not always populated.
+And specific proc codes or cpt codes:
+-          '70551','70552','70553','MHDI', 'IMGMR0128','IMGMR0047','IMGMR0061','IMGMR0054','70551','70552','70553','MHDI', 'IMGMR0128','IMGMR0047','IMGMR0061','IMGMR0054'
+Performing Depts: 361, 547, 6004, 6013
+
+--If I remember correctly, we added additional proc code and cpt codes in because the data output was very low and associated diagnosis was not always populated.
+Also, Taki, Melissa and I talked about some of the possible discrepancies.
+Which was really helpful.
+
+Recommendation: I recommend taking a couple patients and looking at their records.
+
+For ex: I reviewed a patient: Zylstra, Whitney
+I don’t see any ms diagnosis on the main snapshot problem list on the chart and started assuming she doesn’t have MS.
+I looked further and found that there was a procedure encounter diag on 9/11/2017 with diag of “G35”.
+
+But if you look at the orders, I found a couple orders with the procode listed above but the associated diagnosis was not 340 or G35.
+I’m not clinical and I could be wrong but I would think this patient would be on your report.
+
+Thanks
+Sunil
+
+cid:image001.gif@01D291AA.9A09B140
+Sunil Thomas
+Senior Clinical Information Analyst
+(: 215-662-4784
+📧: Sunil.Thomas@pennmedicine.upenn.edu
+
+“Be water, my friend.”
+
+This is the best I have for now guys. I hope this helps!
+
+---

### 2020-11-24 dac_summary.Rmd
Add files via upload
updates

@@
 melissa_peeps_in_eb_group <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 15137
+
+#make a column that indicates whether or not accession number in melissa's is in my group
+melissa_xls$overlap <- ((melissa_xls$ACCESSION_NUM) %in% (data_empi_acc_f_phq$ACCESSION_NUM))
+
+#grab unique accession numbers that are NOT in the overlap group 
+unique_access_nums_only_melissas_group <- melissa_xls$ACCESSION_NUM[which(!melissa_xls$overlap)]
+names(unique_access_nums_only_melissas_group) <- c("count", "accession_num")
+
+#write table, n = 14785
+write.table(file = "/Users/eballer/BBL/msdepression/data/dac/unique_access_nums_in_melissas_group_from_mri_file_upd_w_empi.csv", x = unique_access_nums_only_melissas_group, sep = ",", col.names = TRUE, row.names = FALSE)
 

### 2020-11-24 dacSummary.R

@@ -1,0 +1,1 @@
+# This script takes in data from DAC output and summarizes statistics

### 2021-01-26 dac_summary.Rmd
Add files via upload

@@
 output:
-  html_document
+  html_document: default
+  pdf_document: default
 ---
@@
 
-#set data_empi_acc_f_phq to data3
-data_empi_acc_f_phq <- data3#subset(data, select = c("EMPI", "ACCESSION_NUM", "PHQ.2", "PHQ.9"))
+#from Jan 2021 pull
+data4_20210119 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
+
+#set data_empi_acc_f_phq to data4
+data_empi_acc_f_phq <- data4_20210119#subset(data, select = c("EMPI", "ACCESSION_NUM", "PHQ.2", "PHQ.9"))
 
@@
 #number of unique in my group
-eb_access_n <- length(unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 21027
+eb_access_n <- length(unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 21027 -42840
 
@@
 
-melissa_peeps_in_eb_group <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 15137
+melissa_peeps_in_eb_group <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 15137 -> 25857
 
@@
 
-eb_peeps_in_melissa_group <- sum(unique(data_empi_acc_f_phq$ACCESSION_NUM) %in% unique(melissa_xls$ACCESSION_NUM)) #n=15137
+eb_peeps_in_melissa_group <- sum(unique(data_empi_acc_f_phq$ACCESSION_NUM) %in% unique(melissa_xls$ACCESSION_NUM)) #n=15137 -> 25857
 
@@
 data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
-data_empi_acc_f_phq[28:42] <- apply(X = data_empi_acc_f_phq[28:42], FUN = as.integer, MARGIN = 2)
+data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
+data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
 data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
@@
 #f-codes - people usually have BOTH icd9/10 or neither
-num_depression <- length(which(!is.na(data_empi_acc_f_phq$ICD9))) #n=11131
+#num_depression <- length(which(!is.na(data_empi_acc_f_phq$ICD9))) #n=11131
+num_depression <- length(grep("F3", data_empi_acc_f_phq$ICD10))
 print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
 
-num_unique_EMPI_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & unique(data_empi_acc_f_phq$EMPI))) #n = 593
+#num_unique_EMPI_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & unique(data_empi_acc_f_phq$EMPI))) #n = 593
+# compares overlap of # people with F3* ICD 10 and the unique EMPIs
+#depression df
+dep_df <- grep("F3", data_empi_acc_f_phq$ICD10)
+unique_empi <- match(unique(data_empi_acc_f_phq$EMPI), data_empi_acc_f_phq$EMPI)
+num_unique_EMPI_and_dep <- sum(unique_empi %in% dep_df) #n = 6386
+
 print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep))
 
-num_phq2_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2))) #n = 4855
+num_phq2_and_dep <- sum(dep_df %in% which(!is.na(data_empi_acc_f_phq$PHQ.2))) #n = 6295
 print(paste0("N with ICD 9/10 codes for depression and phq2 = ", num_phq2_and_dep))
 
-num_phq9_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9))) #n=669
+num_phq9_and_dep <- sum(dep_df %in% which(!is.na(data_empi_acc_f_phq$PHQ.9))) #n=932
 print(paste0("N with ICD 9/10 codes for depression and phq9 = ", num_phq9_and_dep))
@@
 #fcode histos
-hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2))], 
-     main = paste0("PHQ2 and Dep Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = paste0("PHQ2"))
+hist(data_empi_acc_f_phq$PHQ.2[dep_df %in% which(!is.na(data_empi_acc_f_phq$PHQ.2))], 
+     main = paste0("PHQ2 and Dep Histo : n = ", num_phq2_and_dep), xlab = paste0("PHQ2"))
 
-hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9))], 
-     main = paste0("PHQ9 and Dep Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = paste0("PHQ9"))
+hist(data_empi_acc_f_phq$PHQ.9[dep_df %in% which(!is.na(data_empi_acc_f_phq$PHQ.9))], 
+     main = paste0("PHQ2 and Dep Histo : n = ", num_phq9_and_dep), xlab = paste0("PHQ9"))
 
@@
 #########################
-data_e <- data3
+data_e <- data4_20210119
 
@@
 #remove people who do not have phq2 or 9
-data_e_with_phq2 <- data_e[which(!is.na(data_e$PHQ.2)),] #4864
-data_e_with_phq9 <- data_e[which(!is.na(data_e$PHQ.9)),] #376
+data_e_with_phq2 <- data_e[which(!is.na(data_e$PHQ.2)),] #13820
+data_e_with_phq9 <- data_e[which(!is.na(data_e$PHQ.9)),] #1202
 
@@
 
+################################
+#### Subset by MS Attending ####
+################################
+
+#obtained from: unique(data_empi_acc_f_phq$Provider[grep("PRUIT", data_empi_acc_f_phq$Provider)])
+ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, LINDA", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
+
+#repeat demographics with only those w/MS providers
+
+data_e <- data4_20210119[which(data_empi_acc_f_phq$Provider %in% ms_providers),]
+dep_df <- grep("F3", data_e$ICD10)
+unique_empi <- match(unique(data_e$EMPI), data_e$EMPI)
+num_unique_EMPI_and_dep <- sum(unique_empi %in% dep_df) #n = 469
+
+#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
+data_e$ACCESSION_NUM <- as.integer(data_e$ACCESSION_NUM)
+data_e[28:42] <- apply(X = data_e[28:42], FUN = as.integer, MARGIN = 2)
+data_e$sex_binarized <- ifelse(data_e$SEX == "MALE", 1, 2)
+data_e$race_binarized <- ifelse(data_e$RACE == "WHITE", 1, 2)
+
+#remove people who do not have phq2 or 9
+data_e_with_phq2 <- data_e[which(!is.na(data_e$PHQ.2)),] #13820
+data_e_with_phq9 <- data_e[which(!is.na(data_e$PHQ.9)),] #1202
+
+#subset people with unique scans (these are one time scanners)
+data_e_no_duplicate_EMPI <- data_e[which(!duplicated(data_e$EMPI)),]
+
+data_e_with_phq2_no_duplicate_EMPI <- data_e_with_phq2[which(!duplicated(data_e_with_phq2$EMPI)),]
+
+data_e_with_phq9_no_duplicate_EMPI <- data_e_with_phq9[which(!duplicated(data_e_with_phq9$EMPI)),]
+
+#####take duplicated people and take first instance
+
+#get duplicated people
+data_e_with_duplicate_EMPI <- data_e[which(duplicated(data_e$EMPI)),]
+
+data_e_with_phq2_w_duplicate_EMPI <- data_e_with_phq2[which(duplicated(data_e_with_phq2$EMPI)),]
+
+data_e_with_phq9_w_duplicate_EMPI <- data_e_with_phq9[which(duplicated(data_e_with_phq9$EMPI)),]
+
+#get match - i.e. first instance of each
+data_e_with_duplicated_EMPI_first_instance <- data_e_with_duplicate_EMPI[match(unique(data_e_with_duplicate_EMPI$EMPI), data_e_with_duplicate_EMPI$EMPI),]
+
+data_e_with_phq2_duplicated_EMPI_first_instance <- data_e_with_phq2_w_duplicate_EMPI[match(unique(data_e_with_phq2_w_duplicate_EMPI$EMPI), data_e_with_phq2_w_duplicate_EMPI$EMPI),]
+
+data_e_with_phq9_duplicated_EMPI_first_instance <- data_e_with_phq9_w_duplicate_EMPI[match(unique(data_e_with_phq9_w_duplicate_EMPI$EMPI), data_e_with_phq9_w_duplicate_EMPI$EMPI),]
+
+#combine groups
+data_e_unique_EMPI_combined <- rbind(data_e_no_duplicate_EMPI, data_e_with_duplicated_EMPI_first_instance)
+
+data_e_phq2_unique_empi_combined <- rbind(data_e_with_phq2_no_duplicate_EMPI, data_e_with_phq2_duplicated_EMPI_first_instance)
+
+data_e_phq9_unique_empi_combined <- rbind(data_e_with_phq9_no_duplicate_EMPI, data_e_with_phq9_duplicated_EMPI_first_instance)
+
+#summarize
+#print(paste0("N unique EMPIs in EB group = ", dim(data_e_unique_EMPI_combined)[1], " out of ", dim(data_e)[1])) #
+
+print(paste0("This looks at data set that only includes people with MS PROVIDERS:"))
+print(paste0(ms_providers))
+
+#unique accession #s
+print(paste0("Number of unique accession #s n= ", length(unique(data_e$ACCESSION_NUM))))
+
+#dep_df includes anyone with ICD 9/10 codes for depression.
+print(paste0("N with ICD 9/10 codes for depression and phq2- ", sum(!is.na(data_e[dep_df,]$PHQ.2))))
+print(paste0("N with ICD 9/10 codes for depression and phq9- ", sum(!is.na(data_e[dep_df,]$PHQ.9))))
+
+print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep))
+
+print(paste0("N PHQ2 = ", sum(!is.na(data_e$PHQ.2)), " out of ", dim(data_e)[1]))
+
+print(paste0("N PHQ9 = ", sum(!is.na(data_e$PHQ.9)), " out of ", dim(data_e)[1]))
+
+print(paste0("N unique EMPIs in EB group = ", length(unique(data_e$EMPI)), " out of ", dim(data_e)[1]))
+
+print(paste0("N unique EMPIs in EB group w/phq2 = ", dim(data_e_phq2_unique_empi_combined)[1], " out of ", dim(data_e)[1])) #
+
+print(paste0("N unique EMPIs in EB group w/phq9 = ", dim(data_e_phq9_unique_empi_combined)[1], " out of ", dim(data_e)[1])) #
+
+print(paste0("N in EB group w/phq2 = ", sum(!is.na(data_e_no_duplicate_EMPI$PHQ.2)), " out of ", dim(data_e)[1])) #
+
+print(paste0("N in EB group w/phq9 = ", sum(!is.na(data_e_no_duplicate_EMPI$PHQ.9)), " out of ", dim(data_e)[1])) #
+
+#histos
+hist(data_e_phq2_unique_empi_combined$PHQ.2, main = paste0("PHQ-2 Histo EB group w/unique EMPI vs all EMPI: n = ", dim(data_e_phq2_unique_empi_combined)[1], "/", length(which(!is.na(data_e$PHQ.2)))), xlab = "PHQ-2", breaks = 6)
+
+#histos
+hist(data_e_phq9_unique_empi_combined$PHQ.9, main = paste0("PHQ-9 Histo EB group w/unique EMPI vs all EMPI: n = ", dim(data_e_phq9_unique_empi_combined)[1], "/", length(which(!is.na(data_e$PHQ.9)))), xlab = "PHQ-9", breaks = 6)
+
 ```

### 2021-01-26 lab_notebook.Rmd
Add files via upload

@@
 ---
+
+#### 20210126
+
+A lot of success.
+Went back and forth with DAC, decided to ask for basically everyone, and we would post sort for providers that were associated with MS clinic, assuming that patients who received a diagnosis of MS from one of these people actually had it.
+
+Updates dac_summary to include not only a summary of the returned DAC pull, but also of the subset of people with MS providers.
+dac_summary.Rmd can be found on github, and summary table on: https://docs.google.com/presentation/d/1pDWXYUmq6sq7Xe08HaWdxLELa-qtQtABCw8RIIZ_7to/edit#slide=id.gb446eacf6a_0_0

### 2021-02-05 dac_summary.Rmd
Add files via upload

@@
 data_empi_acc_f_phq <- data4_20210119#subset(data, select = c("EMPI", "ACCESSION_NUM", "PHQ.2", "PHQ.9"))
+
+############ Start to synthesie data ###########
 
@@
 
-print(paste0("N in EB group w/phq2 = ", sum(!is.na(data_e_no_duplicate_EMPI$PHQ.2)), " out of ", dim(data_e)[1])) #
+print(paste0("N in EB group w/phq2, one time scanners = ", sum(!is.na(data_e_no_duplicate_EMPI$PHQ.2)), " out of ", dim(data_e)[1])) #
 
-print(paste0("N in EB group w/phq9 = ", sum(!is.na(data_e_no_duplicate_EMPI$PHQ.9)), " out of ", dim(data_e)[1])) #
+print(paste0("N in EB group w/phq9, one time scanners = ", sum(!is.na(data_e_no_duplicate_EMPI$PHQ.9)), " out of ", dim(data_e)[1])) #
 
+########################
+####### overlap ########
+########################
+
+eb_icd_and_phq2 <- data_e$ACCESSION_NUM[which(!is.na(data_e[dep_df,]$PHQ.2))] #n = 741
+eb_icd_and_phq9 <- data_e$ACCESSION_NUM[which(!is.na(data_e[dep_df,]$PHQ.9))] #n = 106
+
+#overlap of groups
+n_melissa <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM))
+
+#unique empis
+n_melissa_unique_empi <- length(unique(melissa_xls$EMPI[(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM))]))
+
+#overlap of people with just icds
+n_melissa_icd <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM[dep_df]))
+n_melissa_icd_unique_empi <- length(unique(melissa_xls$EMPI[unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM[dep_df])]))
+
+#overlap of people with phq2, phq9
+n_melissa_phq2 <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.2))]))
+n_melissa_phq9 <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.9))]))
+
+#unique people with phq2/9
+n_melissa_phq2_unique_empi <- length(unique(melissa_xls$EMPI[(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.2))]))]))
+
+n_melissa_phq9_unique_empi <- length(unique(melissa_xls$EMPI[(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.9))]))]))
+
+#overlap of icd dx + phq2 or 9
+n_melissa_peeps_in_eb_icd_phq2_group <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(eb_icd_and_phq2)) #n = 615
+n_melissa_peeps_in_eb_icd_phq9_group <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(eb_icd_and_phq9)) #n = 87
+
+#overlap of unique subjects + phq2, phq9 + icd9/10
+n_mellisa_peeps_in_eb_icd_phq2_group_unique_empi <- length(unique(melissa_xls$EMPI[(unique(melissa_xls$ACCESSION_NUM) %in% unique(eb_icd_and_phq2))]))
+
+n_mellisa_peeps_in_eb_icd_phq9_group_unique_empi <- length(unique(melissa_xls$EMPI[(unique(melissa_xls$ACCESSION_NUM) %in% unique(eb_icd_and_phq9))]))
+######
+
+print(paste0("n people in Melissa and EB overlap group: ", n_melissa, " out of ", length(unique(data_e$ACCESSION_NUM))))
+
+print(paste0("n people in Melissa and EB overlap group, unique empi: ", n_melissa_unique_empi, " out of ", length(unique(data_e$ACCESSION_NUM))))
+
+print(paste0("n with icd9/10 in Melissa and EB overlap groups: ", n_melissa_icd, " out of ", length(unique(data_e$ACCESSION_NUM[dep_df]))))
+
+print(paste0("n people w/icd9/10 in Melissa and EB overlap group with unique empi: ", n_melissa_icd_unique_empi, " out of ", length(unique(data_e$ACCESSION_NUM))))
+
+print(paste0("n with phq2 in Melissa and EB overlap groups: ", n_melissa_phq2, " out of ", length(unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.2))]))))
+
+print(paste0("n with phq9 in Melissa and EB overlap groups: ", n_melissa_phq9, " out of ", length(unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.9))]))))
+
+print(paste0("n with phq2 in Melissa and EB overlap groups, unique empi: ", n_melissa_phq2_unique_empi, " out of ", length(unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.2))]))))
+
+print(paste0("n with phq9 in Melissa and EB overlap groups, unique empi: ", n_melissa_phq9_unique_empi, " out of ", length(unique(data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.9))]))))
+
+print(paste0("n with icd and phq2 in Melissa and EB overlap groups: ", n_melissa_peeps_in_eb_icd_phq2_group, " out of ", length(unique(eb_icd_and_phq2))))
+
+print(paste0("n with icd and phq9 in Melissa and EB groups: ", n_melissa_peeps_in_eb_icd_phq9_group, " out of ", length(unique(eb_icd_and_phq9))))
+
+print(paste0("n with icd and phq2 in Melissa and EB overlap groups, unique empi: ", n_mellisa_peeps_in_eb_icd_phq2_group_unique_empi, " out of ", length(unique(eb_icd_and_phq2))))
+
+print(paste0("n with icd and phq9 in Melissa and EB groups, unique empi: ", n_mellisa_peeps_in_eb_icd_phq9_group_unique_empi, " out of ", length(unique(eb_icd_and_phq9))))
+
+###########
 #histos

### 2021-02-05 lab_notebook.Rmd
Add files via upload

@@
 dac_summary.Rmd can be found on github, and summary table on: https://docs.google.com/presentation/d/1pDWXYUmq6sq7Xe08HaWdxLELa-qtQtABCw8RIIZ_7to/edit#slide=id.gb446eacf6a_0_0
+
+#### 20210129
+
+[x] figure out who has unique icd9/10 and phq2/9
+[x] figure out what the overlap in # of scans between Melissa and my new group are
+
+#### 20210208
+[] check logic of unique EMPIs, looks like something wonky is going on - less unique empis in my group than in the overlap group. Obviously they should be the same if not better

### 2021-02-08 dac_summary.Rmd
Add files via upload

@@
 dep_df <- grep("F3", data_e$ICD10)
-unique_empi <- match(unique(data_e$EMPI), data_e$EMPI)
+unique_empi <- which(unique(data_e$EMPI) %in% data_e$EMPI) #index of the locations of unique empis
 num_unique_EMPI_and_dep <- sum(unique_empi %in% dep_df) #n = 469
@@
 #summarize
-#print(paste0("N unique EMPIs in EB group = ", dim(data_e_unique_EMPI_combined)[1], " out of ", dim(data_e)[1])) #
+print(paste0("N unique EMPIs in EB group = ", length(unique_empi), " out of ", dim(data_e)[1])) #
 
@@
 
-eb_icd_and_phq2 <- data_e$ACCESSION_NUM[which(!is.na(data_e[dep_df,]$PHQ.2))] #n = 741
-eb_icd_and_phq9 <- data_e$ACCESSION_NUM[which(!is.na(data_e[dep_df,]$PHQ.9))] #n = 106
+eb_icd_and_phq2 <- data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.2[dep_df]))] #n = 741
+eb_icd_and_phq9 <- data_e$ACCESSION_NUM[which(!is.na(data_e$PHQ.9[dep_df]))] #n = 106
 
@@
 #unique empis
-n_melissa_unique_empi <- length(unique(melissa_xls$EMPI[(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM))]))
+n_melissa_unique_empi <- sum(unique(melissa_xls$EMPI) %in% unique(data_e$EMPI))
+n_melissa_unique_accession <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_e$ACCESSION_NUM))
 
@@
 
-print(paste0("n people in Melissa and EB overlap group, unique empi: ", n_melissa_unique_empi, " out of ", length(unique(data_e$ACCESSION_NUM))))
+print(paste0("n people in Melissa and EB overlap group, unique empi: ", n_melissa_unique_empi, " out of ", length(unique(data_e$EMPI))))
+
+print(paste0("n people in Melissa and EB overlap group, unique accession #: ", n_melissa_unique_accession, " out of ", length(unique(data_e$ACCESSION_NUM))))
 

### 2021-02-08 lab_notebook.Rmd
Add files via upload

--- 2021-01-26 lab_notebook.Rmd
@@
 dac_summary.Rmd can be found on github, and summary table on: https://docs.google.com/presentation/d/1pDWXYUmq6sq7Xe08HaWdxLELa-qtQtABCw8RIIZ_7to/edit#slide=id.gb446eacf6a_0_0
+
+#### 20210129
+
+[x] figure out who has unique icd9/10 and phq2/9
+[x] figure out what the overlap in # of scans between Melissa and my new group are
+
+#### 20210208
+[x] check logic of unique EMPIs, looks like something wonky is going on - less unique empis in my group than in the overlap group. Obviously they should be the same if not better
+   - fixed, using accession #s and not empis to look for unique empis. Now that I look for unique EMPIs, we are all good!
+
+#### 20210209
+[] post to channel tomorrow

### 2021-04-15 final_dac_summary_pull_20210119.Rmd
Add files via upload

---
title: "Dac summary 01/19/2021"
author: "Erica Baller"
date: "4/15/2021"
output: html_document
---

title: "Dac Summary"
output:
  html_document: default
  pdf_document: default
---

## Dac summary

Welcome to the MS project! This is a really exciting project dedicated to understanding psychopathology, particular depression in patients with MS. It is unqiue in a variety of ways, primarily that it uses clinical images that are research grade.

The team so far is big.
PIs- Ted Satterthwaite and Taki Shinohara
Senior Scientists - Azeez Adebimpe and Matt Cieslak
Research Specialist- Timothy Robert-Fitzgerald
Graduate students - Melissa Martin
Data Analysist - Sidney Covitz
Collaborators - Aaron Alexander-Bloch and Jenna Young
DAC Pull done by - Victoria Rautman(2020/2021) and Sunil Thomas (2018)

Date of FINAL Pull: 1/19/2021
Spreadsheet: /Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv 
Summary of request to Victoria @ DAC:
---
9/2/2020

DAC Request
Non-funded IRB Approved Research  (enter IRB number): 843669
Criteria	Display?	Description / Exclusions / Limitations / Filters
MRN	Y	Unique identifier requested, but can be masked MRN
Visit ID	Y	
Patient Class(es)
Please select only which class(es) you will need.		x  Inpatient
 X   Outpatient
x  Emergency
Age or DOB ranges	Y	
Gender	Y	
Race	Y	
Department(s)
Provide department numbers not just names.	N	
Provider(s)
Provide ID’s not just names.	N	
Date(s)
Include in the specific range and date types (eg, admit, order, result)	1/1/2010-
12/31/20	All scans based on date of first treatment with glatiramer acetate
Procedure
Please include the specific procedure codes.  (ICD9/ICD10 is preferred for inpatient)	Y	88.91 (Magnetic Resonance Imaging of Brain)

Diagnosis
Please include the specific ICD-9/ICD-10 codes including all decimal points. Do not simply include ranges or wildcards.	Y	Multiple sclerosis (ICD-9: 340; ICD-10: G35), Depressive disorders (ICD-9: 296.99, 296.21, 296.22, 296.23, 296.24, 296.25, 296.26, 296.20, 296.31, 296.32, 296.33, 296.34, 296.35, 296.26, 296.30, 300.4, 293.83, 311, ICD-10: F32.0, F32.1, F32.2, F32.3, F32.4, F32.5, F32.6, F33.0, F33.1, F33.2, F33.41, F33.42, F33.9, F34.1, F06.31, F06.32, F06.34, F32.8, F32.9, major depressive disorder, persistent depressive disorder, depressive disorder due to another medical condition, other specified depressive disorder, unspecified depressive disorder)
Orders	N	
Medication
Please list as it is ordered within the UPHS EMR’s – medication id’s preferred	Y	Glatiramer Acetate (Copaxone)
Lab Result
Please list the lab as it is ordered within the UPHS EMR’s.	Y	
Other	Y	EDSS scores (if present)
Other	Y	All medications patients are receiving
Other	Y	All PACS accession numbers available
Other	Y	PHQ-2 and PHQ-9 if present
---
9/16/2020:
Per her note to me: We also specified that we wanted to pull images with ORIG_CPT values of 70551, 70552, 70553, MHDI, IMGMR0128, and we wanted scans with department ID 361, 547, 6004, or 6013. And I may have mentioned this before but the ICD9 and ICD10 codes for MS patients are 340 and G35, respectively

*For us, we are additionally interested in ALL ICD9/10 codes, ALL medications they are on at the time of scan, specific labs (as indicated), EDSS , PHQ-9 and 2 scores. The particular reason for this is that we are going to be looking at the relationship of brain imaging to depression in the MS group.

Mock columns: DE_PAT_ID	PAT_ID	EMPI	HUP_MRN	PAT_NM_WMRN	SEX	RACE	ETHNIC_GROUP	ORDER_ID	ORDERING_DTTM	BEGIN_EXAM_DTTM	END_EXAM_DTTM	TECH_USER_ID	TECHNOLOGIST	ACCESSION_NUM	PROC_ID	PROC_NAME	PROC_CODE	ORIG_CPT	PERFORMING_CSN_ID	COVERAGE_ID	FIN_CLASS_NAME	VISIT_EPM_ID	PERFORMING_DEP_ID	DEPARTMENT_NAME	MODALITY	ASSOCIATED_ICD9	ASSOCIATED_ICD10	PAT_AGE_AT_EXAM	MRI_ENC_AGE	CURRENT_AGE	All ICD-9 or 10 codes	All medication codes	wbc	rbc	hemoglobin	csf studies…	phq-9 score (and date)	phq-2 score (and date)	b12	folate 	TSH	RPR Vit D

---
 1/8/2021 email update (me to Victoria Rautman):
 What we realized is that the best way to define inclusion for the MS group is to focus on people who were seen at some point in neurology clinic.

If you’d be so kind, I’d be grateful for *hopefully the last pull. We would use the previous one, but instead of using the ICD 9/10 codes associated with the actual scan, we would use the location of clinic visits, and pull ALL scans, before or after diagnosis.

New updates:
 -Pull all the scans for patients who have ever been seen by “NEUROLOGY SOUTH PAVILLION” and “NEUROLOGY HUP” departments for clinic.

-Please include provider name associated with these department visits.

-Don’t worry about doing exclusions based on ICD9/10 codes for the scans themselves, please include all and we will sort through it ourselves

***Please also include vitamin D levels in addition to other labs***

The rest of the query would be the same as the most recent query you sent back.
---

Radiology sent images to Taki and group 4/2021

This script goes through data that has been returned from the DAC on 1/19 and summarizes it
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r summary}
#preprocessing - because these are sent

#load our df

#from Jan 2021 pull
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")

ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")

data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
table(data_empi_acc_f_phq$Provider)

##########################
### some preprocessing ###
##########################

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)

########################################
#### Start counting / making subsets ###
########################################

# number of unique scanners and scans
print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 11
print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME)))) #n = 3
print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY)))) #n = 11
print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 17067

#### PHQ-2 and PHQ-9 summaries
#phq-2;
print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)))) #n = 4701

#phq-9
print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)))) #n = 274

#both phq2 and 9

num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0

#looking at just people with phq-2 and 9
phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4701
phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n=274

#Number of unique EMPIs in phq2/phq9 subsets
unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76

#### ICD 9/10 for Depression
dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2339
num_depression <- dim(dep_df)[1]
unique_empi <- which(unique(data_empi_acc_f_phq$EMPI) %in% data_empi_acc_f_phq$EMPI) #index of the locations of unique empis; n = 3885
num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 633

print(paste0("N with ICD 9/10 codes for depression = ", num_depression))

print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of ", dim(data_empi_acc_f_phq)[1]))

##### PHQ2 or 9 + ICD 9/10 Depression
num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 1049
print(paste0("N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = ", num_phq2_and_dep, " out of ", dim(dep_df)[1]))

num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=153
print(paste0("N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = ", num_phq9_and_dep, " out of ", dim(dep_df)[1]))

#### # unique people with dep dx and phq2/9 -> working on this
dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 266
print(paste0("N with ICD 9/10 codes for depression and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(dep_df)[1]))

dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 51
print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))

####

#demographics and lab historgrams
#print histogram of phq-2/9s for people

hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

#fcode histos
hist(dep_df$PHQ.2, main = paste0("PHQ2 and Dep Histo : n = ", num_phq2_and_dep), xlab = paste0("PHQ2"))

hist(dep_df$PHQ.9,main = paste0("PHQ9 and Dep Histo : n = ", num_phq9_and_dep), xlab = paste0("PHQ9"))

###

#age histogram
hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")

#tsh
hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")

#hemoglobin
hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")

#folate
hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)

#vitamin D
hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "Vitamin D", breaks = 10)

#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#race
hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
     xlab = "white = 1, other = 2", 
     breaks = 2)

#f-codes - people usually have BOTH icd9/10 or neither

```

### 2021-04-16 final_dac_summary_pull_20210119.Rmd
Add files via upload

@@
 
+#keep people with MS providers 
 data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
+
+#keep people with ICD10 code G35 (this is the same # as people with IC9 - 340)
+data_empi_acc_f_phq <- data_empi_acc_f_phq[grep("G35", data_empi_acc_f_phq$ICD10),] #n = 16830
 table(data_empi_acc_f_phq$Provider)
@@
 # number of unique scanners and scans
-print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 11
+print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 10
 print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME)))) #n = 3
 print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY)))) #n = 11
-print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 17067
+print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 16830
 
@@
 #phq-2;
-print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)))) #n = 4701
+print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)))) #n = 4606
 
 #phq-9
-print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)))) #n = 274
+print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)))) #n = 261
 
@@
 #### ICD 9/10 for Depression
-dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2339
+dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
 num_depression <- dim(dep_df)[1]
 unique_empi <- which(unique(data_empi_acc_f_phq$EMPI) %in% data_empi_acc_f_phq$EMPI) #index of the locations of unique empis; n = 3885
-num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 633
+num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
 
@@
 ##### PHQ2 or 9 + ICD 9/10 Depression
-num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 1049
+num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 960
 print(paste0("N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = ", num_phq2_and_dep, " out of ", dim(dep_df)[1]))
 
-num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=153
+num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=140
 print(paste0("N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = ", num_phq9_and_dep, " out of ", dim(dep_df)[1]))
@@
 #### # unique people with dep dx and phq2/9 -> working on this
-dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 266
+dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 201
 print(paste0("N with ICD 9/10 codes for depression and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(dep_df)[1]))
 
-dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 51
+dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 43
 print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))

### 2021-05-04 final_dac_summary_pull_20210119.Rmd
added stuff for Tim's piece

@@
 print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 16830
+print(paste0("Number of unique EMPIS n = ", length(unique(data_empi_acc_f_phq$EMPI)))) #n = 3737
 
 #### PHQ-2 and PHQ-9 summaries
+
+#looking at just people with phq-2 and 9
+phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4606
+
+phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n= 261
+
 #phq-2;
-print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)))) #n = 4606
+print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)), "; n = ", length(which(phq2_subset$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
 
 #phq-9
-print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)))) #n = 261
+print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
+             length(which(phq9_subset$PHQ.9 < 5)), "/",
+             length(which(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)), "/",
+             length(which(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)), "/",
+             length(which(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)), "/",
+             length(which(phq9_subset$PHQ.9 >= 20)))) #n = 261
 
@@
 
-#looking at just people with phq-2 and 9
-phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4701
-phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n=274
-
 #Number of unique EMPIs in phq2/phq9 subsets
 unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
+print(paste0("N with phq2 (Unique EMPI) n = ", unique_phq2_empis, " out of ", dim(phq2_subset)[1], "; n - ", length(unique(phq2_subset$EMPI[phq2_subset$PHQ.2 >= 3])), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
+
 unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76
+print(paste0("N with phq9 (Unique EMPI) n = ", unique_phq9_empis, " out of ", dim(phq9_subset)[1], "; noDep/mild/mod/mod-severe/severe = ", 
+             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 < 5])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)])), "/",
+             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 >= 20]))))
 
@@
 
+#phq-2;
+print(paste0("N dep + PHQ-2 = ", sum(!is.na(dep_df$PHQ.2)), "; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
+
+#phq-9
+print(paste0("N dep + PHQ-9 = ", sum(!is.na(dep_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
+             length(which(dep_df$PHQ.9 < 5)), "/",
+             length(which(dep_df$PHQ.9 >= 5 & dep_df$PHQ.9 < 10)), "/",
+             length(which(dep_df$PHQ.9 >= 10 & dep_df$PHQ.9 < 15)), "/",
+             length(which(dep_df$PHQ.9 >= 15 & dep_df$PHQ.9 < 20)), "/",
+             length(which(dep_df$PHQ.9 >= 20)))) #n = 261
+
 #### # unique people with dep dx and phq2/9 -> working on this
@@
 print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))
+
+####
+
+#### ICD 9/10 for Healthy Group
+healthy_df <- data_empi_acc_f_phq[(-grep("F3", data_empi_acc_f_phq$ICD10)),] #n = 14707
+num_healthy <- dim(healthy_df)[1]
+unique_empi <- which(unique(healthy_df$EMPI) %in% healthy_df$EMPI) #index of the locations of unique empis; n = 3244
+num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
+
+print(paste0("N without Depression/ICD 9/10 codes for healthy = ", num_healthy))
+
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", dim(healthy_df)[1]))
+
+##### PHQ2 or 9 + No Depression ICD9/10
+num_phq2_and_healthy <- length(which(!is.na(healthy_df$PHQ.2))) #n = 3646
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = ", num_phq2_and_healthy, " out of ", dim(healthy_df)[1]))
+
+#phq-2;
+print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 47
+
+num_phq9_and_healthy <- length(which(!is.na(healthy_df$PHQ.9))) #n=121
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = ", num_phq9_and_healthy, " out of ", dim(healthy_df)[1]))
+
+#phq-9
+print(paste0("N healthy w/PHQ-9 = ", sum(!is.na(healthy_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
+             length(which(healthy_df$PHQ.9 < 5)), "/",
+             length(which(healthy_df$PHQ.9 >= 5 & healthy_df$PHQ.9 < 10)), "/",
+             length(which(healthy_df$PHQ.9 >= 10 & healthy_df$PHQ.9 < 15)), "/",
+             length(which(healthy_df$PHQ.9 >= 15 & healthy_df$PHQ.9 < 20)), "/",
+             length(which(healthy_df$PHQ.9 >= 20)))) #n = 55/34/21/1/10
+
+#### # unique people without dep dx and phq2/9 
+dep_dx_phq2_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.2)])) #n = 201
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(healthy_df)[1]))
+
+dep_dx_phq9_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.9)])) #n = 43
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(healthy_df)[1]))
 
@@
 
+#wbc
+hist(data_empi_acc_f_phq$WBC[which(!is.na(data_empi_acc_f_phq$WBC))], main = paste0("WBC Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$WBC)))), xlab = "WBC")
+
 #folate
@@
 #vitamin D
-hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "Vitamin D", breaks = 10)
+hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$VitD)))), xlab = "Vitamin D", breaks = 10)
 
@@
 ```
+```{r frequencies_of_dep_in_phq2_vs_9}
+num_dep_phq2 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & (data_empi_acc_f_phq$PHQ.2 >= 3))) #n = 130
+num_dep_phq9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.9) & (data_empi_acc_f_phq$PHQ.9 >= 5))) #n = 184; https://www.sciencedirect.com/science/article/pii/B9780128179338000293 recommends using cut point of 11
+num_total_phq2 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))
+num_total_phq9 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.9))) 
+dep_total_phq_table <- data.frame(matrix(data = c(num_dep_phq2, num_dep_phq9, num_total_phq2, num_total_phq9), nrow = 2, ncol = 2))
+names(dep_total_phq_table) <- c("dep", "healthy")
+row.names(dep_total_phq_table) <- c("phq2", "phq9")
+chisq_dep_phq2_v_phq9 <- chisq.test(dep_total_phq_table)
+chisq_dep_phq2_v_phq9
+print("observed")
+chisq_dep_phq2_v_phq9$observed
+print("expected")
+chisq_dep_phq2_v_phq9$expected
+print("residuals")
+chisq_dep_phq2_v_phq9$residuals
+```
+
+```{r sex_differences_regressions}
+
+### looking at anyone with phq2s or 9s
+males.phq2 <- phq2_subset[which(phq2_subset$SEX == "MALE"),] #n = 951, mean = 0.20
+females.phq2 <- phq2_subset[which(phq2_subset$SEX == "FEMALE"),] #n = 3655, mean = 0.26
+m_v_f_phq2 <- t.test(males.phq2$PHQ.2, females.phq2$PHQ.2) #p = 0.03 ***
+
+males.phq9 <- phq9_subset[which(phq9_subset$SEX == "MALE"),] #n = 80, mean = 8.63
+females.phq9 <- phq9_subset[which(phq9_subset$SEX == "FEMALE"),] #n = 181, mean = 9.13
+m_v_f_phq9 <- t.test(males.phq9$PHQ.9, females.phq9$PHQ.9)
+
+#### Looking at depressed group
+## interstingly, not the 2:1 ratio we usually see in general world
+## phq2 means suggest people are not depressed 
+## phq2 means suggest people meet criteria for mild depression
+
+males_dep <- dep_df[(which(dep_df$SEX == "MALE")),] #n = 419/total(4087; 10%) F; mean PHQ2 = 0.61, mean PHQ9 = 13.05
+females_dep <- dep_df[(which(dep_df$SEX == "FEMALE")),] #n = 1704/total(12743; 13%)total F; mean = 0.56; mean PHQ9 = 10.61
+m_v_f_dep_phq2 <- t.test(males_dep$PHQ.2, females_dep$PHQ.2) #p = 0.66
+m_v_f_dep_phq9 <- t.test(males_dep$PHQ.9, females_dep$PHQ.9) #p = 0.24
+
+#chi sq comparing # of pts dx with depression in males vs females; col 1- male dep n, female dep n; col 2- male total n, female total n
+### fascinating - only 12% of people even gave a dx of depression in their charts. More likely to have it if female than male; also phq2 mean only slightly higher in females. 
+num_dep_m <- dim(males_dep)[1] #n = 419
+num_dep_f <-  dim(females_dep)[1]#n = 1704
+num_total_m <- length(which(data_empi_acc_f_phq$SEX == "MALE")) #n = 4087
+num_total_f <- length(which(data_empi_acc_f_phq$SEX == "FEMALE")) #n = 
+m_v_f_table <- data.frame(matrix(data = c(num_dep_m, num_dep_f, num_total_m, num_total_f), nrow = 2, ncol = 2))
+names(m_v_f_table) <- c("dep", "total")
+row.names(m_v_f_table) <- c("male", "female")
+chisq_m_v_f_dist <- chisq.test(m_v_f_table) #p = 4 x 10^-6
+chisq_m_v_f_dist
+print("observed")
+chisq_m_v_f_dist$observed
+print("expected")
+chisq_m_v_f_dist$expected
+print("residuals")
+chisq_m_v_f_dist$residuals
+#chisq comparing # of
+
+```
+
+```{r summary_psychosis}
+#preprocessing - because these are sent
+
+#load our df
+
+#from Jan 2021 pull
+data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
+
+ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
+
+#keep people with MS providers 
+data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
+
+#keep people with ICD10 code G35 (this is the same # as people with IC9 - 340)
+data_empi_acc_f_phq <- data_empi_acc_f_phq[grep("G35", data_empi_acc_f_phq$ICD10),] #n = 16830
+table(data_empi_acc_f_phq$Provider)
+
+##########################
+### some preprocessing ###
+##########################
+
+#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
+data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
+data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
+data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
+data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
+data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
+
+########################################
+#### Start counting / making subsets ###
+########################################
+
+# number of unique scanners and scans
+print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 10
+print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME)))) #n = 3
+print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY)))) #n = 11
+print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 16830
+print(paste0("Number of unique EMPIS n = ", length(unique(data_empi_acc_f_phq$EMPI)))) #n = 3737
+
+#### PHQ-2 and PHQ-9 summaries
+
+#looking at just people with phq-2 and 9
+phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4606
+
+phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n= 261
+
+#phq-2;
+print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)), "; n = ", length(which(phq2_subset$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
+
+#phq-9
+print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
+             length(which(phq9_subset$PHQ.9 < 5)), "/",
+             length(which(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)), "/",
+             length(which(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)), "/",
+             length(which(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)), "/",
+             length(which(phq9_subset$PHQ.9 >= 20)))) #n = 261
+
+#both phq2 and 9
+
+num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
+print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0
+
+#Number of unique EMPIs in phq2/phq9 subsets
+unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
+print(paste0("N with phq2 (Unique EMPI) n = ", unique_phq2_empis, " out of ", dim(phq2_subset)[1], "; n - ", length(unique(phq2_subset$EMPI[phq2_subset$PHQ.2 >= 3])), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
+
+unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76
+print(paste0("N with phq9 (Unique EMPI) n = ", unique_phq9_empis, " out of ", dim(phq9_subset)[1], "; noDep/mild/mod/mod-severe/severe = ", 
+             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 < 5])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)])), "/",
+             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 >= 20]))))
+
+#### ICD 9/10 for Psychosis
+psychosis_df <- data_empi_acc_f_phq[grep("F2", data_empi_acc_f_phq$ICD10),] #n = 2123
+num_psychosis <- dim(psychosis_df)[1]
+unique_empi <- which(unique(data_empi_acc_f_phq$EMPI) %in% data_empi_acc_f_phq$EMPI) #index of the locations of unique empis; n = 3885
+num_unique_EMPI_and_psychosis <- length(unique(psychosis_df$EMPI)) #n = 493
+
+print(paste0("N with ICD 9/10 codes for psychosis = ", num_psychosis))
+
+print(paste0("N with ICD 9/10 codes for psychosis and unique EMPI = ", num_unique_EMPI_and_psychosis, " out of ", dim(data_empi_acc_f_phq)[1]))
+
+##### PHQ2 or 9 + ICD 9/10 psychosis
+num_phq2_and_psychosis <- length(which(!is.na(psychosis_df$PHQ.2))) #n = 960
+print(paste0("N with ICD 9/10 codes for psychosis and phq2 (ALL SCANS) = ", num_phq2_and_psychosis, " out of ", dim(psychosis_df)[1]))
+
+num_phq9_and_psychosis <- length(which(!is.na(psychosis_df$PHQ.9))) #n=140
+print(paste0("N with ICD 9/10 codes for psychosis and phq9 (ALL SCANS) = ", num_phq9_and_psychosis, " out of ", dim(psychosis_df)[1]))
+
+#phq-2;
+print(paste0("N psychosis + PHQ-2 = ", sum(!is.na(psychosis_df$PHQ.2)), "; n = ", length(which(psychosis_df$PHQ.2 >= 3)), " meeting criteria for psychosis (PHQ2) >= 3")) #n = 4606
+
+#phq-9
+print(paste0("N psychosis + PHQ-9 = ", sum(!is.na(psychosis_df$PHQ.9)), "; nopsychosis/mild/mod/mod-severe/severe = ", 
+             length(which(psychosis_df$PHQ.9 < 5)), "/",
+             length(which(psychosis_df$PHQ.9 >= 5 & psychosis_df$PHQ.9 < 10)), "/",
+             length(which(psychosis_df$PHQ.9 >= 10 & psychosis_df$PHQ.9 < 15)), "/",
+             length(which(psychosis_df$PHQ.9 >= 15 & psychosis_df$PHQ.9 < 20)), "/",
+             length(which(psychosis_df$PHQ.9 >= 20)))) #n = 261
+
+#### # unique people with psychosis dx and phq2/9 -> working on this
+psychosis_dx_phq2_unique <- length(unique(psychosis_df$EMPI[!is.na(psychosis_df$PHQ.2)])) #n = 201
+print(paste0("N with ICD 9/10 codes for psychosis and phq2 (unique) = ", psychosis_dx_phq2_unique, " out of ", dim(psychosis_df)[1]))
+
+psychosis_dx_phq9_unique <- length(unique(psychosis_df$EMPI[!is.na(psychosis_df$PHQ.9)])) #n = 43
+print(paste0("N with ICD 9/10 codes for psychosis and phq9 (unique) = ", psychosis_dx_phq9_unique, " out of ", dim(psychosis_df)[1]))
+
+####
+
+#### ICD 9/10 for Healthy Group
+healthy_df <- data_empi_acc_f_phq[(-grep("F3", data_empi_acc_f_phq$ICD10)),] #n = 14707
+num_healthy <- dim(healthy_df)[1]
+unique_empi <- which(unique(healthy_df$EMPI) %in% healthy_df$EMPI) #index of the locations of unique empis; n = 3244
+num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
+
+print(paste0("N without psychosis/ICD 9/10 codes for healthy = ", num_healthy))
+
+print(paste0("N without psychosis/ ICD 9/10 codes for healthy and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", dim(healthy_df)[1]))
+
+##### PHQ2 or 9 + No psychosis ICD9/10
+num_phq2_and_healthy <- length(which(!is.na(healthy_df$PHQ.2))) #n = 3646
+print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = ", num_phq2_and_healthy, " out of ", dim(healthy_df)[1]))
+
+#phq-2;
+print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for psychosis (PHQ2) >= 3")) #n = 47
+
+num_phq9_and_healthy <- length(which(!is.na(healthy_df$PHQ.9))) #n=121
+print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = ", num_phq9_and_healthy, " out of ", dim(healthy_df)[1]))
+
+#phq-9
+print(paste0("N healthy w/PHQ-9 = ", sum(!is.na(healthy_df$PHQ.9)), "; nopsychosis/mild/mod/mod-severe/severe = ", 
+             length(which(healthy_df$PHQ.9 < 5)), "/",
+             length(which(healthy_df$PHQ.9 >= 5 & healthy_df$PHQ.9 < 10)), "/",
+             length(which(healthy_df$PHQ.9 >= 10 & healthy_df$PHQ.9 < 15)), "/",
+             length(which(healthy_df$PHQ.9 >= 15 & healthy_df$PHQ.9 < 20)), "/",
+             length(which(healthy_df$PHQ.9 >= 20)))) #n = 55/34/21/1/10
+
+#### # unique people without psychosis dx and phq2/9 
+psychosis_dx_phq2_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.2)])) #n = 201
+print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq2 (unique) = ", psychosis_dx_phq2_unique, " out of ", dim(healthy_df)[1]))
+
+psychosis_dx_phq9_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.9)])) #n = 43
+print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq9 (unique) = ", psychosis_dx_phq9_unique, " out of ", dim(healthy_df)[1]))
+
+####
+
+#demographics and lab historgrams
+#print histogram of phq-2/9s for people
+
+hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)
+
+hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)
+
+#fcode histos - no one with it at this time. 
+#hist(psychosis_df$PHQ.2, main = paste0("PHQ2 and psychosis Histo : n = ", num_phq2_and_psychosis), xlab = paste0("PHQ2"))
+
+#hist(psychosis_df$PHQ.9,main = paste0("PHQ9 and psychosis Histo : n = ", num_phq9_and_psychosis), xlab = paste0("PHQ9"))
+
+###
+
+#age histogram
+hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")
+
+#tsh
+hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")
+
+#hemoglobin
+hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")
+
+#wbc
+hist(data_empi_acc_f_phq$WBC[which(!is.na(data_empi_acc_f_phq$WBC))], main = paste0("WBC Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$WBC)))), xlab = "WBC")
+
+#folate
+hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)
+
+#vitamin D
+hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$VitD)))), xlab = "Vitamin D", breaks = 10)
+
+#sex
+hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
+     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ),
+     xlab = "Male = 1, Female = 2", 
+     breaks = 2)
+
+#race
+hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
+     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
+     xlab = "white = 1, other = 2", 
+     breaks = 2)
+
+#f-codes - people usually have BOTH icd9/10 or neither
+
+```
+```{r comparing_out_group_with_Tim_group}
+##### checking to see whether we have clinical data for patients who have already been segmented as of 5/4/2021
+### had to do some parsing of tim's spreadsheet: /Users/eballer/BBL/msdepression/data/insurance_images_for_depression.csv
+### used the following commands to extract the empi and the date -> in tim's they are in the format sub-***** and ses-YYYYMMDD
+
+## more insurance_images_for_depression.csv | cut -f1 -d ','| perl -pe 's/sub-(.*)/$1/' > tim_empi
+## more insurance_images_for_depression.csv | cut -f2 -d ','| perl -pe 's/ses-(....)(..)(..)/$2\/$3\/$1/' > tim_dates
+## more tim_dates | perl -pe 's/(.*)20(..)/$1$2/' | perl -pe 's/(0)(.)\/(.*)\/(.*)/$2\/$3\/$4/' | perl -pe 's/(.*)\/(0)(.*)\/(.*)/$1\/$3\/$4/' > tim_dates_no_zero_pad
+
+segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression_with_accession.csv", header = F, sep = ",") #n = 4468
+names(segmented) <- c("EMPI", "EXAM_DATE", "SES", "ACCESSION_NUM")
+
+#look at the overlap of EMPIs. Won't be able to actually check the matching of scan to person without either accession #s or changing the format of the date. 
+number_of_people_in_tims_group_that_are_also_in_mine <- length(which(segmented$ACCESSION_NUM %in% data_empi_acc_f_phq$ACCESSION_NUM)) #2919
+overlap_group <- data_empi_acc_f_phq[which(data_empi_acc_f_phq$ACCESSION_NUM %in% segmented$ACCESSION_NUM),]
+
+#### PHQ-2 and PHQ-9 summaries
+
+#looking at just people with phq-2 and 9
+phq2_subset <- overlap_group[which(!is.na(overlap_group$PHQ.2)),] #n=4606
+
+phq9_subset <- overlap_group[which(!is.na(overlap_group$PHQ.9)),] #n= 261
+
+#phq-2;
+print(paste0("N PHQ-2 = ", sum(!is.na(overlap_group$PHQ.2)), "; n = ", length(which(phq2_subset$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
+
+#phq-9
+print(paste0("N PHQ-9 = ", sum(!is.na(overlap_group$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
+             length(which(phq9_subset$PHQ.9 < 5)), "/",
+             length(which(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)), "/",
+             length(which(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)), "/",
+             length(which(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)), "/",
+             length(which(phq9_subset$PHQ.9 >= 20)))) #n = 261
+
+#both phq2 and 9
+
+num_both_phq2_and_9 <- length(which(!is.na(overlap_group$PHQ.2) & !is.na(overlap_group$PHQ.9)))
+print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0
+
+#Number of unique EMPIs in phq2/phq9 subsets
+unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
+print(paste0("N with phq2 (Unique EMPI) n = ", unique_phq2_empis, " out of ", dim(phq2_subset)[1], "; n - ", length(unique(phq2_subset$EMPI[phq2_subset$PHQ.2 >= 3])), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
+
+unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76
+print(paste0("N with phq9 (Unique EMPI) n = ", unique_phq9_empis, " out of ", dim(phq9_subset)[1], "; noDep/mild/mod/mod-severe/severe = ", 
+             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 < 5])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)])), "/",
+             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)])), "/",
+             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 >= 20]))))
+
+#### ICD 9/10 for Depression
+dep_df <- overlap_group[grep("F3", overlap_group$ICD10),] #n = 2123
+num_depression <- dim(dep_df)[1]
+unique_empi <- which(unique(overlap_group$EMPI) %in% overlap_group$EMPI) #index of the locations of unique empis; n = 3885
+num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
+
+print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
+
+print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of ", dim(overlap_group)[1]))
+
+##### PHQ2 or 9 + ICD 9/10 Depression
+num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 960
+print(paste0("N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = ", num_phq2_and_dep, " out of ", dim(dep_df)[1]))
+
+num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=140
+print(paste0("N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = ", num_phq9_and_dep, " out of ", dim(dep_df)[1]))
+
+#phq-2;
+print(paste0("N dep + PHQ-2 = ", sum(!is.na(dep_df$PHQ.2)), "; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
+
+#phq-9
+print(paste0("N dep + PHQ-9 = ", sum(!is.na(dep_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
+             length(which(dep_df$PHQ.9 < 5)), "/",
+             length(which(dep_df$PHQ.9 >= 5 & dep_df$PHQ.9 < 10)), "/",
+             length(which(dep_df$PHQ.9 >= 10 & dep_df$PHQ.9 < 15)), "/",
+             length(which(dep_df$PHQ.9 >= 15 & dep_df$PHQ.9 < 20)), "/",
+             length(which(dep_df$PHQ.9 >= 20)))) #n = 261
+
+#### # unique people with dep dx and phq2/9 -> working on this
+dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 201
+print(paste0("N with ICD 9/10 codes for depression and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(dep_df)[1]))
+
+dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 43
+print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))
+
+####
+
+#### ICD 9/10 for Healthy Group
+healthy_df <- overlap_group[(-grep("F3", overlap_group$ICD10)),] #n = 14707
+num_healthy <- dim(healthy_df)[1]
+unique_empi <- which(unique(healthy_df$EMPI) %in% healthy_df$EMPI) #index of the locations of unique empis; n = 3244
+num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
+
+print(paste0("N without Depression/ICD 9/10 codes for healthy = ", num_healthy))
+
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", dim(healthy_df)[1]))
+
+##### PHQ2 or 9 + No Depression ICD9/10
+num_phq2_and_healthy <- length(which(!is.na(healthy_df$PHQ.2))) #n = 3646
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = ", num_phq2_and_healthy, " out of ", dim(healthy_df)[1]))
+
+#phq-2;
+print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 47
+
+num_phq9_and_healthy <- length(which(!is.na(healthy_df$PHQ.9))) #n=121
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = ", num_phq9_and_healthy, " out of ", dim(healthy_df)[1]))
+
+#phq-9
+print(paste0("N healthy w/PHQ-9 = ", sum(!is.na(healthy_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
+             length(which(healthy_df$PHQ.9 < 5)), "/",
+             length(which(healthy_df$PHQ.9 >= 5 & healthy_df$PHQ.9 < 10)), "/",
+             length(which(healthy_df$PHQ.9 >= 10 & healthy_df$PHQ.9 < 15)), "/",
+             length(which(healthy_df$PHQ.9 >= 15 & healthy_df$PHQ.9 < 20)), "/",
+             length(which(healthy_df$PHQ.9 >= 20)))) #n = 55/34/21/1/10
+
+#### # unique people without dep dx and phq2/9 
+dep_dx_phq2_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.2)])) #n = 201
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(healthy_df)[1]))
+
+dep_dx_phq9_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.9)])) #n = 43
+print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(healthy_df)[1]))
+
+####
+
+#demographics and lab historgrams
+#print histogram of phq-2/9s for people
+
+hist(overlap_group$PHQ.2[which(!is.na(overlap_group$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(overlap_group$PHQ.2)))), xlab = "PHQ-2", breaks = 8)
+
+hist(overlap_group$PHQ.9[which(!is.na(overlap_group$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(overlap_group$PHQ.9)))), xlab = "PHQ-9", breaks = 30)
+
+#fcode histos
+hist(dep_df$PHQ.2, main = paste0("PHQ2 and Dep Histo : n = ", num_phq2_and_dep), xlab = paste0("PHQ2"))
+
+hist(dep_df$PHQ.9,main = paste0("PHQ9 and Dep Histo : n = ", num_phq9_and_dep), xlab = paste0("PHQ9"))
+
+###
+
+#age histogram
+hist(overlap_group$MRI_ENC_AGE[which(!is.na(overlap_group$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(overlap_group$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")
+
+#tsh
+hist(overlap_group$TSH[which(!is.na(overlap_group$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(overlap_group$TSH)))), xlab = "TSH")
+
+#hemoglobin
+hist(overlap_group$hemoglobin[which(!is.na(overlap_group$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(overlap_group$hemoglobin)))), xlab = "hemoglobin")
+
+#wbc
+hist(overlap_group$WBC[which(!is.na(overlap_group$WBC))], main = paste0("WBC Histo : n = ", length(which(!is.na(overlap_group$WBC)))), xlab = "WBC")
+
+#folate
+hist(overlap_group$FOLATE[which(!is.na(overlap_group$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(overlap_group$FOLATE)))), xlab = "folate", breaks = 10)
+
+#vitamin D
+hist(overlap_group$VitD[which(!is.na(overlap_group$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(overlap_group$VitD)))), xlab = "Vitamin D", breaks = 10)
+
+#sex
+hist(overlap_group$sex_binarized[which(!is.na(overlap_group$sex_binarized))], 
+     main = paste0("Sex Histo : n = ", length(which(!is.na(overlap_group$sex_binarized))), "; M = ", length(which(overlap_group$sex_binarized == 1)), "; F = ", length(which(overlap_group$sex_binarized == 2)) ),
+     xlab = "Male = 1, Female = 2", 
+     breaks = 2)
+
+#race
+hist(overlap_group$race_binarized[which(!is.na(overlap_group$race_binarized))], 
+     main = paste0("Race Histo : n = ", length(which(!is.na(overlap_group$race))), "; White = ", length(which(overlap_group$race_binarized == 1)), "; NW = ", length(which(overlap_group$race_binarized == 2))),
+     xlab = "white = 1, other = 2", 
+     breaks = 2)
+
+#f-codes - people usually have BOTH icd9/10 or neither
+
+```

### 2021-10-13 dsi_studio_bash.sh
Add files via upload

@@ -1,0 +1,58 @@
+#!/bin/bash
+
+#### Welcome to the Region_making party #########
+### Pre: Must have a file with full paths to the lesioned data 
+### Post: Within each directory specified by the lesioned data, will have a directory that has all tractography (ROA, ROI, Full)
+### Uses: For use in MS depression - take a subject's mimosa lesions and generate the fiber tracts (individual fascicles) that run through it
+#dependencies: Using dsi studio downloaded 9/2021
+
+export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+template='/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+default='/Users/eballer/BBL/msdepression/data/mimosa_file_paths'
+fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa path file" $default
+    lesion_file=$default
+else  
+    lesion_file=$1
+fi
+echo "File being read is "$lesion_file
+echo "... Starting to make lesions ..."
+
+lesion_paths=$(cat $lesion_file)
+
+for lesion in ${lesion_paths}; do
+    echo "Working on file" $lesion
+    # make directory within parent directory for tracts
+    parent_dir=$(echo $lesion | perl -pe s'/(.*)\/.*/$1/g')
+    mask_prefix=$(echo $lesion | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
+    if [ ! -d $parent_dir/fiber_tracking_maps ] 
+    then
+        mkdir $parent_dir/fiber_tracking_maps
+    fi
+
+    fiber_bundle_type_list=(
+        "association"
+        "cerebellum"
+        "cranial_nerve"
+        "projection"
+        "commissural")
+    for fiber_bundle_type in "${fiber_bundle_type_list[@]}"; do
+        echo "Fiber bundle type is " $fiber_bundle_type
+        echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
+        for fascicle in ${fascicles}; do
[…]

### 2021-10-13 lab_notebook/lab_notebook.Rmd
Add files via upload

---
title: "R Notebook for MSDepression Project"
output:
  html_notebook
---

#### 20201008

First, need to summarize data

1) Got data into csv format
  - Opened /Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/*xls in excel and saved as csv
2)
```{move_data bash}
cd /Users/eballer/BBL/msdepression/data/
mkdir dac
cp /Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/*csv .
```

3) In R, started /Users/eballer/BBL/msdepression/dac_summary.Rmd to summarize data

** during the last iteration of de-identifying the DAC, the PHQ-9 scores all vanished. Will need to ask for them to be re-added.
** also, number of people dropped by 50% (from 36277 to 16383)

#### 20201012

Continued with descriptives

  - Read in Melissa's data sheet.
  - subsetted EMPI and Accession # (since our EMPI's are deidentified)
  - Merged with my datasheet on Accession #
  - "PatientID == EMPI"

Merged Melissa's data with mine to get empis.

#### 20201013

[x] Figure out how many people with repeat mrns - to remove doubles. Will need to remove duplicates.

#### 20201027

Going back and forth a lot with DAC people, should have a new update soon. Big thing will be to parse PHQ2/9. We will get all of them but have to tie them to their appropriate scans

#### 20201117

Lots of back and forth with DAC, significantly less people in my pull that limited its pull to people who had G35 codes only. Went back to emails with Sunil (Taki involved). See email below.

---

I think my focus was more on the images and using the challenging radiology data.
So I took a quick look at the codes and files that I might have sent or used. Unfortunately, I can’t remember what the last file I sent was and I don’t have access to the lothal drive anymore.

Criteria:
Don’t see any dates used. Guessing data probably was up to 7/2018(around the time I ran the report)
I looked for patients with order associated diagnosis (%340% or %G35%)  which we mentioned that it is something that is not always populated.
And specific proc codes or cpt codes:
-          '70551','70552','70553','MHDI', 'IMGMR0128','IMGMR0047','IMGMR0061','IMGMR0054','70551','70552','70553','MHDI', 'IMGMR0128','IMGMR0047','IMGMR0061','IMGMR0054'
Performing Depts: 361, 547, 6004, 6013

--If I remember correctly, we added additional proc code and cpt codes in because the data output was very low and associated diagnosis was not always populated. Also, Taki, Melissa and I talked about some of the possible discrepancies. Which was really helpful.

Recommendation: I recommend taking a couple patients and looking at their records.

For ex: I reviewed a patient: Zylstra, Whitney
I don’t see any ms diagnosis on the main snapshot problem list on the chart and started assuming she doesn’t have MS.
I looked further and found that there was a procedure encounter diag on 9/11/2017 with diag of “G35”.

But if you look at the orders, I found a couple orders with the procode listed above but the associated diagnosis was not 340 or G35. I’m not clinical and I could be wrong but I would think this patient would be on your report.

Thanks
Sunil

cid:image001.gif@01D291AA.9A09B140
Sunil Thomas
Senior Clinical Information Analyst
(: 215-662-4784
📧: Sunil.Thomas@pennmedicine.upenn.edu

“Be water, my friend.”

This is the best I have for now guys. I hope this helps!

---

#### 20210126

A lot of success. Went back and forth with DAC, decided to ask for basically everyone, and we would post sort for providers that were associated with MS clinic, assuming that patients who received a diagnosis of MS from one of these people actually had it.

Updates dac_summary to include not only a summary of the returned DAC pull, but also of the subset of people with MS providers. dac_summary.Rmd can be found on github, and summary table on: https://docs.google.com/presentation/d/1pDWXYUmq6sq7Xe08HaWdxLELa-qtQtABCw8RIIZ_7to/edit#slide=id.gb446eacf6a_0_0

#### 20210129

[x] figure out who has unique icd9/10 and phq2/9
[x] figure out what the overlap in # of scans between Melissa and my new group are

#### 20210208
[x] check logic of unique EMPIs, looks like something wonky is going on - less unique empis in my group than in the overlap group. Obviously they should be the same if not better
   - fixed, using accession #s and not empis to look for unique empis. Now that I look for unique EMPIs, we are all good!

#### 20210209
[x] post to channel tomorrow

#### 20210223
[x] Melissa caught that one of our MS providers was in incorrectly. This added a bunch of scans. JACOBS, DINA A. is the right person, not JACOBS, LINDA. Reran my stuff and updated my ms google doc.

Mellisa also realized that our group should include those with a visit with an MS provider AND ICD9/10 MS code.

Today:
3737 patients
28348 total accession numbers
17038 accession numbers we don’t already have and will be requesting
3737 unique empis from the new pull who have seen an MS attending with a corresponding MS diagnosis

#### The future
-at some point, will have to make sure to clarify we used ms diagnoses and the ms provider for our final #s

#### 20210421

- Have been summarizing MS data and some interesting findings are coming about: 
1] "Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): 10"
[1] "Number of unique department names (ie. RAD HUP PCAM Ground): 3"
[1] "Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): 11"
[1] "Number of unique accession #s n= 16830"
[1] "Number of unique EMPIS n = 3737"
[1] "N PHQ-2 = 4606; n = 130 meeting criteria for depression (PHQ2) >= 3"
[1] "N PHQ-9 = 261; noDep/mild/mod/mod-severe/severe = 77/73/65/21/25"
[1] "num people with both phq2 and 9 = 0"
[1] "N with phq2 (Unique EMPI) n = 899 out of 4606; n - 32 meeting criteria for depression (PHQ2) >= 3"
[1] "N with phq9 (Unique EMPI) n = 68 out of 261; noDep/mild/mod/mod-severe/severe = 18/16/18/8/8"
[1] "N with ICD 9/10 codes for depression = 2123"
[1] "N with ICD 9/10 codes for depression and unique EMPI = 493 out of 16830"
[1] "N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = 960 out of 2123"
[1] "N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = 140 out of 2123"
[1] "N dep + PHQ-2 = 960; n = 83 meeting criteria for depression (PHQ2) >= 3"
[1] "N dep + PHQ-9 = 140; noDep/mild/mod/mod-severe/severe = 22/39/44/20/15"
[1] "N with ICD 9/10 codes for depression and phq2 (unique) = 201 out of 2123"
[1] "N with ICD 9/10 codes for depression and phq9 (unique) = 43 out of 2123"
[1] "N without Depression/ICD 9/10 codes for healthy = 14707"
[1] "N without Depression/ ICD 9/10 codes for healthy and unique EMPI = 3244 out of 14707"
[1] "N without Depression/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = 3646 out of 14707"
[1] "N healthy w/PHQ-2 = 3646; n = 47 meeting criteria for depression (PHQ2) >= 3"
[1] "N without Depression/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = 121 out of 14707"
[1] "N healthy w/PHQ-9 = 121; noDep/mild/mod/mod-severe/severe = 55/34/21/1/10"
[1] "N without Depression/ ICD 9/10 codes for healthy and phq2 (unique) = 698 out of 14707"
[1] "N without Depression/ ICD 9/10 codes for healthy and phq9 (unique) = 25 out of 14707"

- Starting to look at statistical relationships as well

- takeaways. ~12% of people have a chart dx of MS, even though the literature says much more
   - PHQ2 and phq9s characterize MS differently. On average, people are not depressed (PHQ2 ~ 0.5), but are mildly depressed with (PHQ9 ~9). What does this mean?
   - PHQ9s include 7 more qs than 2s, including a lot of somatic symptoms (fatigue, appetite, psychomotor agitation/retardation)

- To dos:
  [] Chi sq comparing frequency of depression dx by phq2s vs phq9s
  [] Finding out if there is some sort of a time where some people got phq2s at some point, and some got phq9s -> is there a reason? Same population, different results
  [] literature review of ms depression, understand how depression was previously dx/described; n's, any explanation for somatic sxs?
      - does depression get better with MS tx? Suggesting that MS depression dx could have a lot to do with somatic stuff
      - Another thought - if PHQ9 is pos, but 2 is negative, even though we don't have indiv measures, it is actually splitting out anhedonia from other depressive sxs, which is interesting
  [x] lit search on evidence behind phq2 and 9
  [x] matt schindler meeting
    - to prepare 10 minute summary of the research I've reviewed
       - what has been done
       - what is the state of the field?
    - summary of our sample so far
    - prelim analyses and qs
    - Ask: are there markers of progression in the literature or that you use? Could we extract from EMR? Or would we need NLP?
    - What do you all notice? ANy differences in depression in pts you start on certain txs versus other than we can test in real world?

- also, we do have longitudinal data in a subset, could compare phq2/9 for brains within same subject with more/less MS.

[x] meeting with everyone

#### 20210527

- We are getting really into our collab with harvard and neurology!
- on 5/28, I'll present my work so far. For visualization purposes, I've used a combo of afni (for masking) and fslview (for visualizing), and dsi studio (for making the WM tracts)

- First, we got a "depression network" from harvard, based on a paper from Shan Siddiqui. It is from a preprint:
https://www.dropbox.com/sh/lj5mbfdsig5e4xa/AABfln6Ma_2ialx_KXNHR0oQa?dl=0

- Next, Matt took this into dsi studio, made his mask, and then ran the fibertracking and got some beautiful pictures

- Then, I took the depression network, masked it with our MNI mask, and did visualization in fslview

- Making the masks (had to do this for my figures because the mask from harvard didn't quite line up with our MNI, so I had to make sure I masked the harvard one with the mni template so it looked pretty)

    3dcalc -a MNI152_T1_2mm_brain.nii.gz -expr 'ispositive(a)' -prefix mni_positive_mask.nii
    3dcalc -a mni_positive_mask.nii -b AllFX_wmean.nii -expr 'a*b' -prefix mni_x_AllFx_mean.nii

- visualizing the overlaps 
  - fslview_deprecated
  - open each of the masks. Make sure the eye is clicked to keep it "on"
  - the info button allows you to change the scale from grey to hot (red)
  - the threshold is at the top. When changing it, make sure you are clicking on the image you want to use

- dsistudio to make the fiber tracks
  - open dsistudio
  - click 3
  - open the image /Users/eballer/BBL/msdepression/templates/dti/1904002815.src.gz.odf8.f5.gqi.1.25.fib.gz
  - Step t3d, click "enable autotrack"
  - click "fibertracking"
  - so pretty!
  - To do special fibers, unhighlight the full brain
    -   Step t3d, select the fiber you want to use
    -   click "fibertracking"
  - To do a region of avoidance, go to step T3a
    -   click the white piece of paper
    - click ROA
    - in t3b, make a box, circle, or whatever on the brain
    - in T3d, click the same fiber tracking button, and now you have only the fibers that were not hit by your cutting!
    - click on and off the two fiber tracts to see what you gain/lose!
    - if you don't like what you have, delete it

#### 20210806

- Going to work on algorithm with Matt as Tim reprocesses all of the scans
- Doing this by taking a good, qc'd sample of mimosa, finding a way to register it to T1, and how to register T1 to MNI space (to piggyback lesion maps)
- can do the same with probability maps, though starting with lesion maps (as if they are just rois) could be a good way to start
- Tasks

[x] I need datalad on pmacs
   - so we can put our data here
   - steps:
```{bash}
ssh -Y eballer@sciget.pmacs.upenn.edu
cd ~
source pmacs-setup-project-user.sh # this was copied from the cubic-setup-project-user.sh script per Azeez
```

[x] I need to download hcp data into datalad

[x] connectomedb account
[x] dsistudio is local
   - use it to test reconstruction
[x] download 3 or 4 hcp scans
- things were going okay and then my dsi studio had a bug. Now I basically need to wait for my new computer to come in.

20210826
  [x] Chi sq comparing frequency of depression dx by phq2s vs phq9s
  [x] Finding out if there is some sort of a time where some people got phq2s at some point, and some got phq9s -> is there a reason? Same population, different results - not really
  [] literature review of ms depression, understand how depression was previously dx/described; n's, any explanation for somatic sxs?
      - does depression get better with MS tx? Suggesting that MS depression dx could have a lot to do with somatic stuff
      - Another thought - if PHQ9 is pos, but 2 is negative, even though we don't have indiv measures, it is actually splitting out anhedonia from other depressive sxs, which is interesting
[x] figure out how to register lesions to hcp scans
[x] ask Phil cook for how he registers older brains (with some WM lesions that can interfere with registration) onto MNI

20211007 
[x] made a script to take lesions (these are hand drawn) and use dsi studio to make subtraction maps. Now all I need are real files!

script: "/Users/eballer/BBL/msdepression/scripts/dsi_studio_bash.sh

```{bash dsi_studio_bash}
#!/bin/bash

#### Welcome to the Region_making party #########
### Pre: Must have a file with full paths to the lesioned data 
### Post: Within each directory specified by the lesioned data, will have a directory that has all tractography (ROA, ROI, Full)
### Uses: For use in MS depression - take a subject's mimosa lesions and generate the fiber tracts (individual fascicles) that run through it
#dependencies: Using dsi studio downloaded 9/2021

export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
template='/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz'
default='/Users/eballer/BBL/msdepression/data/mimosa_file_paths'
fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'

if [ $# == 0 ]
then
    echo "We will use the default mimosa path file" $default
    lesion_file=$default
else  
    lesion_file=$1
fi
echo "File being read is "$lesion_file
echo "... Starting to make lesions ..."

lesion_paths=$(cat $lesion_file)

for lesion in ${lesion_paths}; do
    echo "Working on file" $lesion
    # make directory within parent directory for tracts
    parent_dir=$(echo $lesion | perl -pe s'/(.*)\/.*/$1/g')
    mask_prefix=$(echo $lesion | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
    if [ ! -d $parent_dir/fiber_tracking_maps ] 
    then
        mkdir $parent_dir/fiber_tracking_maps
    fi

    fiber_bundle_type_list=(
        "association"
        "cerebellum"
        "cranial_nerve"
        "projection"
        "commissural")
    for fiber_bundle_type in "${fiber_bundle_type_list[@]}"; do
        echo "Fiber bundle type is " $fiber_bundle_type
        echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
        mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
        fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
        for fascicle in ${fascicles}; do
            echo "Fascicle is  " $fascicle
            fascicle_root_name=$(echo $fascicle | perl -pe s'/(.*)\/(.*).tt.gz/$2/g') 
            dsi_studio --action=ana --source=$template --tract=$fascicle --roa=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.tt.gz
            dsi_studio --action=ana --source=$template --tract=$fascicle --roi=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.tt.gz
            dsi_studio --action=ana --source=$template --tract=$fascicle --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_full.tt.gz
        done
        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roa=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROA.nii.gz
        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roi=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROI.nii.gz
        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_full_ROI.nii.gz
    done
done

```

20211012
 [] Working on getting a sample MS mimosa output into HCP space

    - Perfect subject in MS group
        - T1w: 
        ** with skull
        /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/t1_n4.nii.gz
        ** skull stripped
        /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/t1_n4_brain.nii.gz
        ** brainmask
        /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/t1_n4_brainmask.nii.gz
        - mimosa file: /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz
    - HCP T1 : /Users/eballer/BBL/msdepression/templates/dti/mni_icbm152_nlin_asym_09a/mni_icbm152_t1_tal_nlin_asym_09a.nii

  - recommended script from Phil Cook: 
    * antsRegistrationSyN.sh

### ---------- Narrative from Phil and Matt ------    
    If you have raw data, I would do antsBrainExtraction.sh, N4BiasFieldCorrection, then antsRegistrationSyN.sh (brain <-> brain)

Matt Cieslak  10:06 AM
would you do anything differently if you had a mask of small MS lesions?

Phil Cook  10:08 AM
You can mask them out by inverting and then passing mask with -x
10:08
So basically -x (brain_mask - lesions)

Matt Cieslak  10:09 AM
for the masks, we should use the skull-on, n4 image and -x’ed mask from antsBrainExtraction minus the lesions? as input to antsRegistrationSyN.sh (edited)

Phil Cook  10:10 AM
If the lesions are small, masking them out should be sufficient
10:12
For larger lesions, you can try to impute "normal" tissue in various ways, then do the registration
10:17
There's different strategies for skull on vs off.
Some people use the skull-on image with a tight brain mask in fixed and moving space, to compensate for areas where the brain mask was not expansive enough.
I usually take the skull off, so that bits of skull don't get pulled inside the mask during nonlinear registration, and use a mask that is dilated a bit so that the edge of the brain is not the edge of the mask. Reasoning being that I want small misalignments at the edge of the brain to feature in the metric and its gradients

Matt Cieslak  10:19 AM
ok that’s what I thought
10:19
Sounds good @Erica Baller?
10:19
thanks as always @Phil Cook!

Erica Baller  10:20 AM
Mostly yes!
10:25
In theory it makes sense - practically a little more challenging. Is there any documentation or script that actually does this kind of thing step by step that I could use to get the more technical pieces?

Phil Cook  10:36 AM
https://github.com/ntustison/antsBrainExtractionExample
https://github.com/stnava/ANTsTutorial
The scripts are not too hard to use. An example N4 call:
N4BiasFieldCorrection -d 3 -i <input> -x <brain mask> -w <brain_mask minus lesions> -s 2 -c  [ 50x50x50x50,0.0000000001 ] -b [ 180 ] -o <output> --verbose 1
ntustison/antsBrainExtractionExample
Stars
4
Language
R
Added by GitHub
stnava/ANTsTutorial
Stars
27
Language
HTML
Added by GitHub
10:37
The trick with -x and -w for N4 is that you can include them in the mask (-x) but exclude them from the bias computation (-w is zero in lesions). This way they won't mess up bias correction but will get corrected. If you exclude voxels from -x, they do not get corrected
10:41
The shrink factor "-s" controls how to downsample the image. This makes the computation faster. For 1mm data I use 2, for sub-1mm data 4. The other parameter of interest is the initial spline spacing 180mm. Every level doubles resolution, so 180x90x45x22.5 . For a more detailed bias field, reduce this number or add an extra level. The default in ANTs scripts is 200 for adult human data, but I have found that bias fields have gotten worse over time, so I typically use 160-180
### --------- Done with excerpted text --------

Making my mask (for brain_mask minus lesions)

```{bash make_substraction_mask}
cd /Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg
3dcalc -a t1_n4_brainmask.nii.gz -b mimosa_binary_mask_0.25.nii.gz -expr 'a-b' -prefix t1_n4_brain_minus_mimosa_binary_mask.nii.gz
```

Now Matt is asking if we need to dilate the mask more.

So I did
```{bash}
3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3

3dcalc -a t1_n4_brainmask.nii.gz -b mimosa_binary_mask_0.25.nii.gz -expr 'a-b' -prefix t1_n4_brain_minus_mimosa_binary_mask.nii.gz

3dcalc -a mni_icbm152_t1_tal_nlin_asym_09a.nii -b mni_icbm152_t1_tal_nlin_asym_09a_mask.nii -expr 'a*b' -prefix mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
```
 The input going in will be:
 T1w: /Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg/t1_n4_brain.nii.gz  
 Mask:/Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg/t1_n4_brainmask_d3.nii.gz
 MNI HCP template: /Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg/mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii

 Transferred my directory onto pmacs, where ants should theoretically be
```{bash}
scp -r erica_created_for_hcp_reg/ eballer@transfer.pmacs.upenn.edu:/project/msdepression

ssh -Y eballer@scisub.pmacs.upenn.edu
cd /project/ms
xbash

sh ants_registration_code.sh

```

Current code in ants_registration_code.sh
```{bash}
#!/bin/bash
### ANTS registration for MS Depression
##pre- requires ms patient T1, (eventually their mimosa), mni brain: mni_icbm152_t1_tal_nlin_asym_09a.
nii, mask (currently using the one that was dilated in afni, d3
## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
##Post - ms participant T1w registered to icbm_template space (and later mimosa as well)
##Uses - We have all of these great ms depression scans but they are in native space. This will get th
em into mni space so they can be used with HCP templates, both structural and diffusion
##Dependencies: Using ANTs/2.3.5

# Set up paths
#export ANTSPATH=/home/eballer/
#export PATH=${ANTSPATH}:$PATH

#### Run registration

#default T1: t1_n4_brain.nii.gz 
t1=t1_n4_brain.nii.gz 
#default brain mask: t1_n4_brainmask_d3.nii.gz
brain_mask=t1_n4_brainmask_d3.nii.gz
#default target/template hcp brain" mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
mni_target_t1=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii

#output
outfile_prefix=ms_t1_to_mni_icbm152
antsRegistrationSyN.sh -d 3 -f ${mni_target_t1} -m ${t1} -o ${outfile_prefix} -x ${brain_mask}

#now actually do the transform
mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp.nii.gz #output file prefix
mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
seWarp, InverseWarped. I picked the one that matched my pnc output the closes

antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2mni_warp} -t ${affine_mat} -n GenericLabel
```

20211013
  [x] made a copy of the pmacs stuff and put it locally for ease of making pictures
 [x] take the new lesion file, put it in dsi studio, and run the tractography to get the track loss. 
 [] please please please git this
 [x] can also try to run with script from command line WORKS!
 [x] checked R/L orientation - cooking with gas

[x] there is a weird line in my pictures, so am going to try with different interpolation (Multilabel and Nearest Neighbor)
  - made new script: /project/msdepression/erica_created_for_hcp_reg/ants_transform_code.sh
  - to run:
  - ssh -Y eballer@scisub.pmacs.upenn.edu
  - module load ANTs/2.3.5
  - xbash
  - sh ants_transform_code.sh
  - if this doesn't work, type "export PATH=${ANTSPATH}:$PATH" at the command line

```{bash ants_transform_code.sh}
export PATH=${ANTSPATH}:$PATH
mimosaa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp_gl.nii.gz #output file prefix
mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
seWarp, InverseWarped. I picked the one that matched my pnc output the closes

antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2m
ni_warp} -t ${affine_mat} -n GenericLabel

mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp_ml.nii.gz #output file prefix
mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
seWarp, InverseWarped. I picked the one that matched my pnc output the closes

antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2m
ni_warp} -t ${affine_mat} -n MultiLabel

mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp_nn.nii.gz #output file prefix
mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
seWarp, InverseWarped. I picked the one that matched my pnc output the closes

antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2m
ni_warp} -t ${affine_mat} -n NearestNeighbor
```

### 2021-10-13 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
 20211012
- [] Working on getting a sample MS mimosa output into HCP space
+ [x] Working on getting a sample MS mimosa output into HCP space
 
@@
  [x] take the new lesion file, put it in dsi studio, and run the tractography to get the track loss. 
- [] please please please git this
+ [x] please please please git this
  [x] can also try to run with script from command line WORKS!
@@
 export PATH=${ANTSPATH}:$PATH
-mimosaa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
 mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp_gl.nii.gz #output file prefix

### 2021-10-13 ants_registration_code.sh
Add files via upload

@@ -1,0 +1,33 @@
+#!/bin/bash
+### ANTS registration for MS Depression
+##pre- requires ms patient T1, (eventually their mimosa), mni brain: mni_icbm152_t1_tal_nlin_asym_09a.nii, mask (currently using the one that was dilated in afni, d3
+## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
+##Post - ms participant T1w registered to icbm_template space (and later mimosa as well)
+##Uses - We have all of these great ms depression scans but they are in native space. This will get them into mni space so they can be used with HCP templates, both structural and diffusion
+##Dependencies: Using ANTs/2.3.5
+
+# Set up paths
+#export ANTSPATH=/home/eballer/
+#export PATH=${ANTSPATH}:$PATH
+
+#### Run registration
+
+#default T1: t1_n4_brain.nii.gz 
+t1=t1_n4_brain.nii.gz 
+#default brain mask: t1_n4_brainmask_d3.nii.gz
+brain_mask=t1_n4_brainmask_d3.nii.gz
+#default target/template hcp brain" mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
+mni_target_t1=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
+
+#output
+outfile_prefix=ms_t1_to_mni_icbm152
+antsRegistrationSyN.sh -d 3 -f ${mni_target_t1} -m ${t1} -o ${outfile_prefix} -x ${brain_mask}
+
+#now actually do the transform
+mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp.nii.gz #output file prefix
+mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
+affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
+ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped InverseWarp, InverseWarped. I picked the one that matched my pnc output the closes
+
+antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2mni_warp} -t ${affine_mat} -n GenericLabel

### 2021-10-13 ants_registration_code.sh
Add files via upload

@@ -30,4 +30,4 @@ mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
 affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
 ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped InverseWarp, InverseWarped. I picked the one that matched my pnc output the closes
 
-antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2mni_warp} -t ${affine_mat} -n GenericLabel
+antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2mni_warp} -t ${affine_mat} -n MultiLabel

### 2021-10-14 parse_mimosa_file.sh
convert melissa files to paths for depression mimosa

@@ -1,0 +1,3 @@
+more 1yr_report_address.csv | grep "mimosa" | grep ",,75.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_75_paths
+more 1yr_report_address.csv | grep "mimosa" | grep ",,100.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_100_paths
+cat mimosa_100_paths mimosa_75_paths > mimosa_100_and_75_paths

### 2021-10-15 ants_registration_code.sh
ants code updated with new export paths. Have not run this version yet with some of the exports

@@ -1,33 +1,87 @@
 #!/bin/bash
 ### ANTS registration for MS Depression
-##pre- requires ms patient T1, (eventually their mimosa), mni brain: mni_icbm152_t1_tal_nlin_asym_09a.nii, mask (currently using the one that was dilated in afni, d3
-## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
-##Post - ms participant T1w registered to icbm_template space (and later mimosa as well)
+##pre- requires ms mimosa file that contains the directory, mni brain: mni_icbm152_t1_tal_nlin_asym_09a.nii
+##Post - ms participant T1w and mimosa registered to icbm_template space
 ##Uses - We have all of these great ms depression scans but they are in native space. This will get them into mni space so they can be used with HCP templates, both structural and diffusion
+#Steps
+## 1 - take all file paths to data with good qc, extract file paths, and make shadow directories with linksin my local data directory
+## 2 - make the dilated mask
+     #####mask (currently using the one that was dilated in afni, d3
+     ## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
+## 3 - for each subject, run the registration, and then apply the transpforms
 ##Dependencies: Using ANTs/2.3.5
-
 # Set up paths
-#export ANTSPATH=/home/eballer/
+module load ANTs/2.3.5
+#export ANTSPATH="/appl/ANTs/2.3.5"
+#export PATH=${ANTSPATH}:$PATH
+export ANTSPATH=/appl/ANTs-2.3.5/bin
+
+#######if it doesn't work, fix this ###
 #export PATH=${ANTSPATH}:$PATH
 
-#### Run registration
 
-#default T1: t1_n4_brain.nii.gz 
-t1=t1_n4_brain.nii.gz 
-#default brain mask: t1_n4_brainmask_d3.nii.gz
-brain_mask=t1_n4_brainmask_d3.nii.gz
-#default target/template hcp brain" mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
-mni_target_t1=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
 
-#output
+
+#set default templatesfiles/paths
+
[…]

### 2021-10-15 lab_notebook.Rmd

@@
 #### 20210209
-[] post to channel tomorrow
+[x] post to channel tomorrow
+
+#### 20210223
+[x] Melissa caught that one of our MS providers was in incorrectly. This added a bunch of scans. JACOBS, DINA A. is the right person, not JACOBS, LINDA. Reran my stuff and updated my ms google doc.
+
+Mellisa also realized that our group should include those with a visit with an MS provider AND ICD9/10 MS code.
+
+Today:
+3737 patients
+28348 total accession numbers
+17038 accession numbers we don’t already have and will be requesting
+3737 unique empis from the new pull who have seen an MS attending with a corresponding MS diagnosis
+
+#### The future
+-at some point, will have to make sure to clarify we used ms diagnoses and the ms provider for our final #s
+
+#### 20210421
+
+- Have been summarizing MS data and some interesting findings are coming about: 
+1] "Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): 10"
+[1] "Number of unique department names (ie. RAD HUP PCAM Ground): 3"
+[1] "Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): 11"
+[1] "Number of unique accession #s n= 16830"
+[1] "Number of unique EMPIS n = 3737"
+[1] "N PHQ-2 = 4606; n = 130 meeting criteria for depression (PHQ2) >= 3"
+[1] "N PHQ-9 = 261; noDep/mild/mod/mod-severe/severe = 77/73/65/21/25"
+[1] "num people with both phq2 and 9 = 0"
+[1] "N with phq2 (Unique EMPI) n = 899 out of 4606; n - 32 meeting criteria for depression (PHQ2) >= 3"
+[1] "N with phq9 (Unique EMPI) n = 68 out of 261; noDep/mild/mod/mod-severe/severe = 18/16/18/8/8"
+[1] "N with ICD 9/10 codes for depression = 2123"
+[1] "N with ICD 9/10 codes for depression and unique EMPI = 493 out of 16830"
+[1] "N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = 960 out of 2123"
+[1] "N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = 140 out of 2123"
+[1] "N dep + PHQ-2 = 960; n = 83 meeting criteria for depression (PHQ2) >= 3"
+[1] "N dep + PHQ-9 = 140; noDep/mild/mod/mod-severe/severe = 22/39/44/20/15"
+[1] "N with ICD 9/10 codes for depression and phq2 (unique) = 201 out of 2123"
+[1] "N with ICD 9/10 codes for depression and phq9 (unique) = 43 out of 2123"
+[1] "N without Depression/ICD 9/10 codes for healthy = 14707"
+[1] "N without Depression/ ICD 9/10 codes for healthy and unique EMPI = 3244 out of 14707"
+[1] "N without Depression/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = 3646 out of 14707"
+[1] "N healthy w/PHQ-2 = 3646; n = 47 meeting criteria for depression (PHQ2) >= 3"
+[1] "N without Depression/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = 121 out of 14707"
+[1] "N healthy w/PHQ-9 = 121; noDep/mild/mod/mod-severe/severe = 55/34/21/1/10"
+[1] "N without Depression/ ICD 9/10 codes for healthy and phq2 (unique) = 698 out of 14707"
+[1] "N without Depression/ ICD 9/10 codes for healthy and phq9 (unique) = 25 out of 14707"
+
+- Starting to look at statistical relationships as well
+
+- takeaways. ~12% of people have a chart dx of MS, even though the literature says much more
+   - PHQ2 and phq9s characterize MS differently. On average, people are not depressed (PHQ2 ~ 0.5), but are mildly depressed with (PHQ9 ~9). What does this mean?
+   - PHQ9s include 7 more qs than 2s, including a lot of somatic symptoms (fatigue, appetite, psychomotor agitation/retardation)
+
+- To dos:
+  [] Chi sq comparing frequency of depression dx by phq2s vs phq9s
+  [] Finding out if there is some sort of a time where some people got phq2s at some point, and some got phq9s -> is there a reason? Same population, different results
+  [] literature review of ms depression, understand how depression was previously dx/described; n's, any explanation for somatic sxs?
+      - does depression get better with MS tx? Suggesting that MS depression dx could have a lot to do with somatic stuff
+      - Another thought - if PHQ9 is pos, but 2 is negative, even though we don't have indiv measures, it is actually splitting out anhedonia from other depressive sxs, which is interesting
+  [x] lit search on evidence behind phq2 and 9
+  [x] matt schindler meeting
+    - to prepare 10 minute summary of the research I've reviewed
+       - what has been done
+       - what is the state of the field?
+    - summary of our sample so far
+    - prelim analyses and qs
+    - Ask: are there markers of progression in the literature or that you use? Could we extract from EMR? Or would we need NLP?
+    - What do you all notice? ANy differences in depression in pts you start on certain txs versus other than we can test in real world?
+
+- also, we do have longitudinal data in a subset, could compare phq2/9 for brains within same subject with more/less MS.
+
+[x] meeting with everyone
+
+#### 20210527
+
+- We are getting really into our collab with harvard and neurology!
+- on 5/28, I'll present my work so far. For visualization purposes, I've used a combo of afni (for masking) and fslview (for visualizing), and dsi studio (for making the WM tracts)
+
+- First, we got a "depression network" from harvard, based on a paper from Shan Siddiqui. It is from a preprint:
+https://www.dropbox.com/sh/lj5mbfdsig5e4xa/AABfln6Ma_2ialx_KXNHR0oQa?dl=0
+
+- Next, Matt took this into dsi studio, made his mask, and then ran the fibertracking and got some beautiful pictures
+
+- Then, I took the depression network, masked it with our MNI mask, and did visualization in fslview
+
+- Making the masks (had to do this for my figures because the mask from harvard didn't quite line up with our MNI, so I had to make sure I masked the harvard one with the mni template so it looked pretty)
+
+    3dcalc -a MNI152_T1_2mm_brain.nii.gz -expr 'ispositive(a)' -prefix mni_positive_mask.nii
+    3dcalc -a mni_positive_mask.nii -b AllFX_wmean.nii -expr 'a*b' -prefix mni_x_AllFx_mean.nii
+
+- visualizing the overlaps 
+  - fslview_deprecated
+  - open each of the masks. Make sure the eye is clicked to keep it "on"
+  - the info button allows you to change the scale from grey to hot (red)
+  - the threshold is at the top. When changing it, make sure you are clicking on the image you want to use
+
+- dsistudio to make the fiber tracks
+  - open dsistudio
+  - click 3
+  - open the image /Users/eballer/BBL/msdepression/templates/dti/1904002815.src.gz.odf8.f5.gqi.1.25.fib.gz
+  - Step t3d, click "enable autotrack"
+  - click "fibertracking"
+  - so pretty!
+  - To do special fibers, unhighlight the full brain
+    -   Step t3d, select the fiber you want to use
+    -   click "fibertracking"
+  - To do a region of avoidance, go to step T3a
+    -   click the white piece of paper
+    - click ROA
+    - in t3b, make a box, circle, or whatever on the brain
+    - in T3d, click the same fiber tracking button, and now you have only the fibers that were not hit by your cutting!
+    - click on and off the two fiber tracts to see what you gain/lose!
+    - if you don't like what you have, delete it
+
+#### 20210806
+
+- Going to work on algorithm with Matt as Tim reprocesses all of the scans
+- Doing this by taking a good, qc'd sample of mimosa, finding a way to register it to T1, and how to register T1 to MNI space (to piggyback lesion maps)
+- can do the same with probability maps, though starting with lesion maps (as if they are just rois) could be a good way to start
+- Tasks
+
+[x] I need datalad on pmacs
+   - so we can put our data here
+   - steps:
+```{bash}
+ssh -Y eballer@sciget.pmacs.upenn.edu
+cd ~
+source pmacs-setup-project-user.sh # this was copied from the cubic-setup-project-user.sh script per Azeez
+```
+
+[x] I need to download hcp data into datalad
+
+[x] connectomedb account
+[x] dsistudio is local
+   - use it to test reconstruction
+[x] download 3 or 4 hcp scans
+- things were going okay and then my dsi studio had a bug. Now I basically need to wait for my new computer to come in.
+
+20210826
+  [x] Chi sq comparing frequency of depression dx by phq2s vs phq9s
+  [x] Finding out if there is some sort of a time where some people got phq2s at some point, and some got phq9s -> is there a reason? Same population, different results - not really
+  [] literature review of ms depression, understand how depression was previously dx/described; n's, any explanation for somatic sxs?
+      - does depression get better with MS tx? Suggesting that MS depression dx could have a lot to do with somatic stuff
+      - Another thought - if PHQ9 is pos, but 2 is negative, even though we don't have indiv measures, it is actually splitting out anhedonia from other depressive sxs, which is interesting
+[x] figure out how to register lesions to hcp scans
+[x] ask Phil cook for how he registers older brains (with some WM lesions that can interfere with registration) onto MNI
+
+20211007 
+[x] made a script to take lesions (these are hand drawn) and use dsi studio to make subtraction maps. Now all I need are real files!
+
+script: "/Users/eballer/BBL/msdepression/scripts/dsi_studio_bash.sh
+
+```{bash dsi_studio_bash}
+#!/bin/bash
+
+#### Welcome to the Region_making party #########
+### Pre: Must have a file with full paths to the lesioned data 
+### Post: Within each directory specified by the lesioned data, will have a directory that has all tractography (ROA, ROI, Full)
+### Uses: For use in MS depression - take a subject's mimosa lesions and generate the fiber tracts (individual fascicles) that run through it
+#dependencies: Using dsi studio downloaded 9/2021
+
+export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+template='/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+default='/Users/eballer/BBL/msdepression/data/mimosa_file_paths'
+fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa path file" $default
+    lesion_file=$default
+else  
+    lesion_file=$1
+fi
+echo "File being read is "$lesion_file
+echo "... Starting to make lesions ..."
+
+lesion_paths=$(cat $lesion_file)
+
+for lesion in ${lesion_paths}; do
+    echo "Working on file" $lesion
+    # make directory within parent directory for tracts
+    parent_dir=$(echo $lesion | perl -pe s'/(.*)\/.*/$1/g')
+    mask_prefix=$(echo $lesion | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
+    if [ ! -d $parent_dir/fiber_tracking_maps ] 
+    then
+        mkdir $parent_dir/fiber_tracking_maps
+    fi
+
+    fiber_bundle_type_list=(
+        "association"
+        "cerebellum"
+        "cranial_nerve"
+        "projection"
+        "commissural")
+    for fiber_bundle_type in "${fiber_bundle_type_list[@]}"; do
+        echo "Fiber bundle type is " $fiber_bundle_type
+        echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
+        for fascicle in ${fascicles}; do
+            echo "Fascicle is  " $fascicle
+            fascicle_root_name=$(echo $fascicle | perl -pe s'/(.*)\/(.*).tt.gz/$2/g') 
+            dsi_studio --action=ana --source=$template --tract=$fascicle --roa=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.tt.gz
+            dsi_studio --action=ana --source=$template --tract=$fascicle --roi=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.tt.gz
+            dsi_studio --action=ana --source=$template --tract=$fascicle --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_full.tt.gz
+        done
+        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roa=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROA.nii.gz
+        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roi=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROI.nii.gz
+        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_full_ROI.nii.gz
+    done
+done
+
+```
+
+20211012
+ [x] Working on getting a sample MS mimosa output into HCP space
+
+    - Perfect subject in MS group
+        - T1w: 
+        ** with skull
+        /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/t1_n4.nii.gz
+        ** skull stripped
+        /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/t1_n4_brain.nii.gz
+        ** brainmask
+        /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/t1_n4_brainmask.nii.gz
+        - mimosa file: /Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz
+    - HCP T1 : /Users/eballer/BBL/msdepression/templates/dti/mni_icbm152_nlin_asym_09a/mni_icbm152_t1_tal_nlin_asym_09a.nii
+
+  - recommended script from Phil Cook: 
+    * antsRegistrationSyN.sh
+
+### ---------- Narrative from Phil and Matt ------    
+    If you have raw data, I would do antsBrainExtraction.sh, N4BiasFieldCorrection, then antsRegistrationSyN.sh (brain <-> brain)
+
+Matt Cieslak  10:06 AM
+would you do anything differently if you had a mask of small MS lesions?
+
+Phil Cook  10:08 AM
+You can mask them out by inverting and then passing mask with -x
+10:08
+So basically -x (brain_mask - lesions)
+
+Matt Cieslak  10:09 AM
+for the masks, we should use the skull-on, n4 image and -x’ed mask from antsBrainExtraction minus the lesions? as input to antsRegistrationSyN.sh (edited)
+
+Phil Cook  10:10 AM
+If the lesions are small, masking them out should be sufficient
+10:12
+For larger lesions, you can try to impute "normal" tissue in various ways, then do the registration
+10:17
+There's different strategies for skull on vs off.
+Some people use the skull-on image with a tight brain mask in fixed and moving space, to compensate for areas where the brain mask was not expansive enough.
+I usually take the skull off, so that bits of skull don't get pulled inside the mask during nonlinear registration, and use a mask that is dilated a bit so that the edge of the brain is not the edge of the mask.
+Reasoning being that I want small misalignments at the edge of the brain to feature in the metric and its gradients
+
+Matt Cieslak  10:19 AM
+ok that’s what I thought
+10:19
+Sounds good @Erica Baller?
+10:19
+thanks as always @Phil Cook!
+
+Erica Baller  10:20 AM
+Mostly yes!
+10:25
+In theory it makes sense - practically a little more challenging.
+Is there any documentation or script that actually does this kind of thing step by step that I could use to get the more technical pieces?
+
+Phil Cook  10:36 AM
+https://github.com/ntustison/antsBrainExtractionExample
+https://github.com/stnava/ANTsTutorial
+The scripts are not too hard to use. An example N4 call:
+N4BiasFieldCorrection -d 3 -i <input> -x <brain mask> -w <brain_mask minus lesions> -s 2 -c  [ 50x50x50x50,0.0000000001 ] -b [ 180 ] -o <output> --verbose 1
+ntustison/antsBrainExtractionExample
+Stars
+4
+Language
+R
+Added by GitHub
+stnava/ANTsTutorial
+Stars
+27
+Language
+HTML
+Added by GitHub
+10:37
+The trick with -x and -w for N4 is that you can include them in the mask (-x) but exclude them from the bias computation (-w is zero in lesions).
+This way they won't mess up bias correction but will get corrected.
+If you exclude voxels from -x, they do not get corrected
+10:41
+The shrink factor "-s" controls how to downsample the image.
+This makes the computation faster.
+For 1mm data I use 2, for sub-1mm data 4.
+The other parameter of interest is the initial spline spacing 180mm.
+Every level doubles resolution, so 180x90x45x22.5 .
+For a more detailed bias field, reduce this number or add an extra level.
+The default in ANTs scripts is 200 for adult human data, but I have found that bias fields have gotten worse over time, so I typically use 160-180
+### --------- Done with excerpted text --------
+
+Making my mask (for brain_mask minus lesions)
+
+```{bash make_substraction_mask}
+cd /Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg
+3dcalc -a t1_n4_brainmask.nii.gz -b mimosa_binary_mask_0.25.nii.gz -expr 'a-b' -prefix t1_n4_brain_minus_mimosa_binary_mask.nii.gz
+```
+
+Now Matt is asking if we need to dilate the mask more.
+
+So I did
+```{bash}
+3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
+
+3dcalc -a t1_n4_brainmask.nii.gz -b mimosa_binary_mask_0.25.nii.gz -expr 'a-b' -prefix t1_n4_brain_minus_mimosa_binary_mask.nii.gz
+
+3dcalc -a mni_icbm152_t1_tal_nlin_asym_09a.nii -b mni_icbm152_t1_tal_nlin_asym_09a_mask.nii -expr 'a*b' -prefix mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
+```
+ The input going in will be:
+ T1w: /Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg/t1_n4_brain.nii.gz  
+ Mask:/Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg/t1_n4_brainmask_d3.nii.gz
+ MNI HCP template: /Users/eballer/BBL/msdepression/templates/erica_created_for_hcp_reg/mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
+
+ Transferred my directory onto pmacs, where ants should theoretically be
+```{bash}
+scp -r erica_created_for_hcp_reg/ eballer@transfer.pmacs.upenn.edu:/project/msdepression
+
+ssh -Y eballer@scisub.pmacs.upenn.edu
+cd /project/ms
+xbash
+
+sh ants_registration_code.sh
+
+```
+
+Current code in ants_registration_code.sh
+```{bash}
+#!/bin/bash
+### ANTS registration for MS Depression
+##pre- requires ms patient T1, (eventually their mimosa), mni brain: mni_icbm152_t1_tal_nlin_asym_09a.
+nii, mask (currently using the one that was dilated in afni, d3
+## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
+##Post - ms participant T1w registered to icbm_template space (and later mimosa as well)
+##Uses - We have all of these great ms depression scans but they are in native space. This will get th
+em into mni space so they can be used with HCP templates, both structural and diffusion
+##Dependencies: Using ANTs/2.3.5
+
+# Set up paths
+#export ANTSPATH=/home/eballer/
+#export PATH=${ANTSPATH}:$PATH
+
+#### Run registration
+
+#default T1: t1_n4_brain.nii.gz 
+t1=t1_n4_brain.nii.gz 
+#default brain mask: t1_n4_brainmask_d3.nii.gz
+brain_mask=t1_n4_brainmask_d3.nii.gz
+#default target/template hcp brain" mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
+mni_target_t1=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii
+
+#output
+outfile_prefix=ms_t1_to_mni_icbm152
+antsRegistrationSyN.sh -d 3 -f ${mni_target_t1} -m ${t1} -o ${outfile_prefix} -x ${brain_mask}
+
+#now actually do the transform
+mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp.nii.gz #output file prefix
+mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
+affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
+ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
+seWarp, InverseWarped. I picked the one that matched my pnc output the closes
+
+antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2mni_warp} -t ${affine_mat} -n GenericLabel
+```
+
+20211013
+  [x] made a copy of the pmacs stuff and put it locally for ease of making pictures
+ [x] take the new lesion file, put it in dsi studio, and run the tractography to get the track loss. 
+ [x] please please please git this
+ [x] can also try to run with script from command line WORKS!
+ [x] checked R/L orientation - cooking with gas
+
+[x] there is a weird line in my pictures, so am going to try with different interpolation (Multilabel and Nearest Neighbor)
+  - made new script: /project/msdepression/erica_created_for_hcp_reg/ants_transform_code.sh
+  - to run:
+  - ssh -Y eballer@scisub.pmacs.upenn.edu
+  - module load ANTs/2.3.5
+  - xbash
+  - sh ants_transform_code.sh
+  - if this doesn't work, type "export PATH=${ANTSPATH}:$PATH" at the command line
+
+```{bash ants_transform_code.sh}
+export PATH=${ANTSPATH}:$PATH
+mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp_gl.nii.gz #output file prefix
+mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
+affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
+ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
+seWarp, InverseWarped. I picked the one that matched my pnc output the closes
+
+antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2m
+ni_warp} -t ${affine_mat} -n GenericLabel
+
+mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp_ml.nii.gz #output file prefix
+mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
+affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
+ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
+seWarp, InverseWarped. I picked the one that matched my pnc output the closes
+
+antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2m
+ni_warp} -t ${affine_mat} -n MultiLabel
+
+mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp_nn.nii.gz #output file prefix
+mni_hcp_path=mni_icbm152_t1_tal_nlin_asym_09axbrainmask.nii #template
+affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
+ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped Inver
+seWarp, InverseWarped. I picked the one that matched my pnc output the closes
+
+antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_hcp_path} -t ${ms2m
+ni_warp} -t ${affine_mat} -n NearestNeighbor
+```
+
+  20211014
+
+  [] Docker dsi studio
+    - Tim downloaded 10/13
+
+  [] Start making mimosa registrations for melissa's group
+    - Melissa's QC'ed people 
+  /project/pennsive_pacs2/repos/insurance/1yr_report_address.csv
+
+  - first, filter the group to keep the people who have good mimosas (assumes that their T1s are also good, and pull out directory paths
+  - make a file that contains people whose mimosas have quality 75, 100, and then combine
+
+```{bash parse_mimosa_file.sh}
+
+#n = 2336, corresponding to  958 subjects
+more 1yr_report_address.csv | grep "mimosa" | grep ",,75.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_75_paths
+more 1yr_report_address.csv | grep "mimosa" | grep ",,100.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_100_paths
+cat mimosa_100_paths mimosa_75_paths > mimosa_100_and_75_paths
+```
+   New script for /project/msdepression/scripts/ants_registration_code.sh on pmacs. too nervous to git it
+
+```{bash ants_registration_code.sh}
+
+#read in directories
+directories=$(cat $qc_file_path)
+
+#extract name of directory and create shadow directory in my own data directory, with appropriate file names
+for directory in $directories; do
+        sub_sess_run=$(echo ${directory} | perl -pe 's/.*(sub.*)/$1/') #grab sub/sess/run
+        echo $sub_sess_run
+        echo "making dir" ${new_data_path}/${sub_sess_run}
+        mkdir -p ${new_data_path}/${sub_sess_run}
+        ln -s ${directory}/* ${new_data_path}/${sub_sess_run}
+        cp ${mni_t1_path} ${new_data_path}/${sub_sess_run}
+
+        #go into directory for afni stuff
+        cd ${new_data_path}/${sub_sess_run}
+
+        #make the dilated mask
+        3dmask_tool -input ${brain_mask} -prefix ${brain_mask_dilated} -dilate_input ${mask_dilation}
+
+        #multiply with mni t1 to get appropriate coverage
+        3dcalc -a ${brain_mask_dilated} -b ${mni_t1_root}.nii -expr 'a*b' -prefix ${mni_t1_target}
+
+        # a typical file 
+#### Run registration
+
+        antsRegistrationSyN.sh -d 3 -f ${mni_t1_target} -m ${t1} -o ${outfile_prefix} -x ${brain_mask_dilated}
+
+        #now actually do the transform${brain
+        mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+        mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp.nii.gz #output file prefix
+        affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
+        ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped InverseWarp, InverseWarped. I picked the one that matched my pnc output the closes
+
+        antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_t1_target} -t ${ms2mni_warp} -t ${affine_mat} -n MultiLabel
+        cd ${base_dir}
+done
+
+```

### 2021-10-15 lab_notebook/lab_notebook.Rmd

@@
 ```
+
+  20211014
+
+  [] Docker dsi studio
+    - Tim downloaded 10/13
+
+  [] Start making mimosa registrations for melissa's group
+    - Melissa's QC'ed people 
+  /project/pennsive_pacs2/repos/insurance/1yr_report_address.csv
+
+  - first, filter the group to keep the people who have good mimosas (assumes that their T1s are also good, and pull out directory paths
+  - make a file that contains people whose mimosas have quality 75, 100, and then combine
+
+```{bash parse_mimosa_file.sh}
+
+#n = 2336, corresponding to  958 subjects
+more 1yr_report_address.csv | grep "mimosa" | grep ",,75.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_75_paths
+more 1yr_report_address.csv | grep "mimosa" | grep ",,100.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_100_paths
+cat mimosa_100_paths mimosa_75_paths > mimosa_100_and_75_paths
+```
+   New script for /project/msdepression/scripts/ants_registration_code.sh on pmacs. too nervous to git it
+
+```{bash ants_registration_code.sh}
+
+#read in directories
+directories=$(cat $qc_file_path)
+
+#extract name of directory and create shadow directory in my own data directory, with appropriate file names
+for directory in $directories; do
+        sub_sess_run=$(echo ${directory} | perl -pe 's/.*(sub.*)/$1/') #grab sub/sess/run
+        echo $sub_sess_run
+        echo "making dir" ${new_data_path}/${sub_sess_run}
+        mkdir -p ${new_data_path}/${sub_sess_run}
+        ln -s ${directory}/* ${new_data_path}/${sub_sess_run}
+        cp ${mni_t1_path} ${new_data_path}/${sub_sess_run}
+
+        #go into directory for afni stuff
+        cd ${new_data_path}/${sub_sess_run}
+
+        #make the dilated mask
+        3dmask_tool -input ${brain_mask} -prefix ${brain_mask_dilated} -dilate_input ${mask_dilation}
+
+        #multiply with mni t1 to get appropriate coverage
+        3dcalc -a ${brain_mask_dilated} -b ${mni_t1_root}.nii -expr 'a*b' -prefix ${mni_t1_target}
+
+        # a typical file 
+#### Run registration
+
+        antsRegistrationSyN.sh -d 3 -f ${mni_t1_target} -m ${t1} -o ${outfile_prefix} -x ${brain_mask_dilated}
+
+        #now actually do the transform${brain
+        mimosa_path=mimosa_binary_mask_0.25.nii.gz #this is from mimosa output
+        mimosa_mni_hcp_path=mimosa_binary_mask_0.25_mni_hcp.nii.gz #output file prefix
+        affine_mat=ms_t1_to_mni_icbm1520GenericAffine.mat #affine output from last step
+        ms2mni_warp=ms_t1_to_mni_icbm1521Warp.nii.gz #warp from last step. There are a few, Warp, Warped InverseWarp, InverseWarped. I picked the one that matched my pnc output the closes
+
+        antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_t1_target} -t ${ms2mni_warp} -t ${affine_mat} -n MultiLabel
+        cd ${base_dir}
+done
+
+```
+
+   20211015
+   [x] Need to test ants registration code. If all goes well, hope to run this over the weekend
+   [x]just need to add ants to my home directory somehow. But everything up to that seems to work
+
+   Last update 20211015
+
+   /project/msdepression/scripts/ants_registration_code.sh
+```{bash ants_registration_code}
+
+#!/bin/bash
+### ANTS registration for MS Depression
+##pre- requires ms mimosa file that contains the directory, mni brain: mni_icbm152_t1_tal_nlin_asym_09a.nii
+##Post - ms participant T1w and mimosa registered to icbm_template space
+##Uses - We have all of these great ms depression scans but they are in native space. This will get them into mni space so t
+hey can be used with HCP templates, both structural and diffusion
+#Steps
+## 1 - take all file paths to data with good qc, extract file paths, and make shadow directories with linksin my local data 
+directory
+## 2 - make the dilated mask
+     #####mask (currently using the one that was dilated in afni, d3
+     ## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
+## 3 - for each subject, run the registration, and then apply the transpforms
+##Dependencies: Using ANTs/2.3.5
+# Set up paths
+module load ANTs/2.3.5
+#export ANTSPATH="/appl/ANTs/2.3.5"
+#export PATH=${ANTSPATH}:$PATH
+export ANTSPATH=/appl/ANTs-2.3.5/bin
+
+#######if it doesn't work, fix this ###
+#export PATH=${ANTSPATH}:$PATH
+
+#set default templatesfiles/paths
+
+base_dir="/project/msdepression/scripts"
+
+#Primary directorie - CHANGE THESE IF ANYTHING EVER CHANGES, IT WILL UPDATE FILE NAMES BELOW
+
+qc_file_path='/project/msdepression/data/melissa_martin_files/csv/minimelissa_list_for_testing_n3'
+mni_t1_root='mni_icbm152_t1_tal_nlin_asym_09a'
+mask_dilation=3
+brain_mask_root="t1_n4_brainmask"
+t1="t1_n4_reg_brain_ws.nii.gz"
+mimosa_root="mimosa_binary_mask_0.25"
+new_data_path='/project/msdepression/data/subj_directories/'
+mni_t1_direc="/project/msdepression/templates/mni_icbm152_nlin_asym_09a/"
+
+#set outfile prefix
+outfile_prefix=ms_t1_to_mni_icbm152
+
+#for transform
+mimosa_path=${mimosa_root}.nii.gz #this is from mimosa output
+mimosa_mni_hcp_path=${mimosa_root}_mni_hcp.nii.gz #output file prefix
+affine_mat=${outfile_prefix}0GenericAffine.mat #affine output from last step
+ms2mni_warp=${outfile_prefix}1Warp.nii.gz #warp from last step. There are a few, Warp, Warped InverseWarp, InverseWarped. I 
+picked the one that matched my pnc output the closes
+
+#making some secondary files
+mni_t1_path="${mni_t1_direc}/${mni_t1_root}.nii"
+mni_t1_target="${mni_t1_root}xbrainmask.nii"
+brain_mask="${brain_mask_root}.nii.gz"
+brain_mask_dilated="${brain_mask_root}_d${mask_dilation}.nii.gz"
+
+#read in directories
+directories=$(cat $qc_file_path)
+
+#extract name of directory and create shadow directory in my own data directory, with appropriate file names
+for directory in $directories; do
+        sub_sess_run=$(echo ${directory} | perl -pe 's/.*(sub.*)/$1/') #grab sub/sess/run
+        echo $sub_sess_run
+        echo "making dir" ${new_data_path}/${sub_sess_run}
+        mkdir -p ${new_data_path}/${sub_sess_run}
+        ln -s ${directory}/* ${new_data_path}/${sub_sess_run}
+        cp ${mni_t1_direc}/${mni_t1_root}* ${new_data_path}/${sub_sess_run}
+
+        #go into directory for afni stuff
+        cd ${new_data_path}/${sub_sess_run}
+
+        	#make the dilated mask
+        	3dmask_tool -input ${brain_mask} -prefix ${brain_mask_dilated} -dilate_input ${mask_dilation}
+
+        	#multiply with mni t1 to get appropriate coverage - this is the problem!!! 
+        	3dcalc -a ${mni_t1_root}.nii -b ${mni_t1_root}_mask.nii -expr 'a*b' -prefix ${mni_t1_target}
+
+		#### Run registration
+
+        	antsRegistrationSyN.sh -d 3 -f ${mni_t1_target} -m ${t1} -o ${outfile_prefix} -x ${brain_mask_dilated}
+
+        	#now actually do the transform${brain
+
+       	 	antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_t1_target} -t ${ms2mni_wa
+rp} -t ${affine_mat} -n MultiLabel
+
+	cd ${base_dir}
+done
+
+```

### 2021-11-03 indiv_mimosa_lesion_dsi_studio_script.sh
indiv mimosa lesion now deletes whole fiber_tracking directory rather than avoiding it

@@ -30,17 +30,32 @@ else
     	echo "$parent_dir"
 	mask_prefix=$(echo ${lesion} | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
     	echo $mask_prefix
-	if [ ! -d $parent_dir/fiber_tracking_maps ] 
-    	then
-        	mkdir $parent_dir/fiber_tracking_maps
-    	fi
+
+	#if directory exists, delete it
+
+	#if [ -d $parent_dir/fiber_tracking_maps ] 
+    	#then
+	
+	# delete fiber tracking maps directory
+
+	echo "Deleting ${parent_dir}/fiber_tracking_maps/"
+	rm -rf ${parent_dir}/fiber_tracking_maps
+#	fi
+
+	#make new fiber tracking maps directory
+	echo "Making new ${parent_dir}/fiber_tracking_maps"
+        mkdir ${parent_dir}/fiber_tracking_maps
+ 		
 
     #make sub directories for each bundle type, so within fiber_tracking_maps, there will be subdirectories for each type of map
 	fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
     	for fiber_bundle_type in ${fiber_bundle_type_list}; do
         	echo "Fiber bundle type is " $fiber_bundle_type
         	echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
-        	mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+
+		
+		#make new directory
+		mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
         	fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
     
 

### 2021-11-03 make_streamline_volumes_for_template.sh

@@ -1,0 +1,34 @@
+
+### Make streamline volumes ###
+
+### Pre: All template streamlines must have been made in dsi studio (these are the nii files)
+### Post: fiber_volume_values.csv that contains the name of the fiber bundle and its volume
+### Uses: Will be getting the volumes of the streamlines for use in later density analyses
+### Dependencies: This script requires afni to be installed.
+
+module load afni_openmp/20.1
+export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
+num_cores=1
+
+directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
+outfile=${directory}/fiber_volume_values.csv
+
+#remove outfile if it exists
+rm -f ${outfile}
+
+#make the file again
+touch ${outfile}
+echo "fascicle,volume" >> ${outfile}
+   
+#go through each fiber, get volume, and store it in csv
+fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
+for fiber_bundle_type in ${fiber_bundle_type_list}; do
+	echo "Fiber bundle type is " $fiber_bundle_type
+fascicle_volumes=$(ls ${directory}/${fiber_bundle_type}/*.nii*)
+        for fascicle in ${fascicle_volumes}; do
+            fascicle_prefix=$(echo $fascicle | perl -pe 's/.*\/(.*).nii.*/$1/g')
+            volume=$(3dmaskave -quiet -mask SELF -sum ${fascicle})
+            echo "${fascicle_prefix},$volume" >> ${outfile}
+        done
+done
+

### 2021-11-05 roa_ratio_regressions.R
made roi data frame for future merger

@@ -1,0 +1,54 @@
+### ROA Ratio Regressions ###
+
+#####
+## Pre: /project/msdepression/data/melissa_martin_files/csv/roi_fiber_volumes_n2336, which contains full paths to the ROI files, can be ammended to do ro
+## Post: Analysis
+## Uses: 1) reads in the ROIs, puts them in a data frame (columns are accession num/date/each fascicle), one row per subjected per date
+#        2) Performers univariate analyses, depression diagnosis versus not
+## Dependencies: R 3.6.3
+
+#include packages
+library(dbplyr)
+
+#set directories
+homedir<-'/project/msdepression/'
+
+#analysis type
+region_type="roi"
+
+#read in template fascicle volumes
+template_full_volumes <- read.csv(paste0(homedir, '/templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv'), header = T)
+fascicle_names <- template_full_volumes$fascicle
+
+### Read in file with all full paths to region volumes
+region_volume_csv <- read.table(paste0(homedir, '/data/melissa_martin_files/csv/', region_type, '_fiber_volumes_n2336'), header = F, stringsAsFactors = F)
+region_volume_csv$V1 <- as.character(region_volume_csv$V1)
+
+#outfile, named for number of subjects
+outfile <- paste0(homedir, "/results/fascicle_volumes_all_subjects_", region_type, "_n", dim(region_volume_csv)[1], ".csv")
+
+#make data frame, 1 row per subject, num rows = num rows in region_volume_csv
+volumes_df <- data.frame(matrix(nrow = dim(region_volume_csv)[1], ncol = 89))
+names(volumes_df) <- c("ACCESSION_NUM", "EXAM_DATE", as.character(fascicle_names))
+#loop through region_volume_csv, read in df corresponding to each row of region_volume_csv, extract accession num and date,
+## and transpose fiber volumes from column to rows. append accession num/date/region_volumes row into big matrix
+
+for (subj in 1:dim(region_volume_csv)[1]) {
+  file_path <- region_volume_csv$V1[subj]
+  region_volume_data <- read.csv(file_path, header = T, sep = ",")
+  accession_num<- region_volume_data[1,1]
+  date <- region_volume_data[1,2]
+  volumes <- t(region_volume_data[,4])
+  volumes[is.na(volumes)] <- 0
+  volumes <- volumes/template_full_volumes$volume
[…]

### 2021-11-05 lab_notebook/lab_notebook.Rmd
lab notebook additions

@@
        	 	antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_t1_target} -t ${ms2mni_wa
-rp} -t ${affine_mat} -n MultiLabel
+rp} -t ${affine_mat} -n GenericLabel
 
@@
 ```
+
+20211019
+Still having issues with transform, running slow, realizing that I had multilable instead of generic label. Will try that. Still worried I will have trouble getting it to work.
+
+Got it to run by putting a bsub inside the script!
+
+20211026
+- Now trying to get dsi studio to run on pmacs. 
+- scp -r /Users/eballer/msdepression/templates/dti to /project/msdepression/templates
+
+Made a list of the paths to all the mimosa masks
+ls /project/msdepression/data/subj_directories/*/*//*/mimosa*mn* > mimosa_binary_masks_hcp_space_20211026_n2336
+
+Then moved to melissa's data directory
+cp mimosa_binary_masks_hcp_space_20211026_n2336 ../melissa_martin_files/csv/.
+
+20211027
+ - fixing dsi studio. Goal today is to figure out how to bsub it because it takes a long time. steps
+
+```{bash}
+ssh -Y eballer@scisub.pmacs.upenn.edu
+cd /project/msdepression/scripts
+xbash
+
+```
+
+ Also needed to generate a list of niftis (i.e. tracts is voxel space). Did this locally and scp'ed to templates/dti
+
+```{bash}
+#!/bin/bash
+
+#### Welcome to the Region_making party #########
+### Pre: Must have a file with full paths to the lesioned data 
+### Post: Within each directory specified by the lesioned data, will have a directory that has all tractography (ROA, ROI, Full)
+### Uses: For use in MS depression - take a subject's mimosa lesions and generate the fiber tracts (individual fascicles) that run through it
+#dependencies: Using dsi studio downloaded 9/2021
+
+export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+template='/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+default='/Users/eballer/BBL/msdepression/data/mimosa_file_paths'
+fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa path file" $default
+    lesion_file=$default
+else  
+    lesion_file=$1
+fi
+echo "File being read is "$lesion_file
+echo "... Starting to make lesions ..."
+
+lesion_paths=$(cat $lesion_file)
+
+for lesion in ${lesion_paths}; do
+    echo "Working on file" $lesion
+    # make directory within parent directory for tracts
+    #parent_dir=$(echo $lesion | perl -pe s'/(.*)\/.*/$1/g')
+    mask_prefix=$(echo $lesion | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
+    if [ ! -d $parent_dir/fiber_tracking_maps ] 
+    then
+        mkdir $parent_dir/fiber_tracking_maps
+    fi
+
+    fiber_bundle_type_list=(
+        "association"
+        "cerebellum"
+        "cranial_nerve"
+        "projection"
+        "commissural")
+    for fiber_bundle_type in "${fiber_bundle_type_list[@]}"; do
+        echo "Fiber bundle type is " $fiber_bundle_type
+        echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        #mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
+        for fascicle in ${fascicles}; do
+            echo "Fascicle is  " $fascicle
+            fascicle_root_name=$(echo $fascicle | perl -pe s'/(.*)\/(.*).tt.gz/$2/g') 
+          #  dsi_studio --action=ana --source=$template --tract=$fascicle --roa=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.tt.gz
+          #  dsi_studio --action=ana --source=$template --tract=$fascicle --roi=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.tt.gz
+           # dsi_studio --action=ana --source=$template --tract=$fascicle --output=${fascicle_directory}/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_full.tt.gz
+            dsi_studio --action=ana --source=$template --tract=$fascicle --output=${fascicle_directory}/${fiber_bundle_type}/${fascicle_root_name}.nii.gz
+        done
+        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roa=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROA.nii.gz
+        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roi=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROI.nii.gz
+        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_full_ROI.nii.gz
+    done
+done
+```
+ 20211028
+
+ Big moves! Currently running these two scripts and they have been git'ed
+```{bash dsi_studio_bash.sh}
+#!/bin/bash
+
+#### Welcome to the Region_making party #########
+### Pre: Must have a file with full paths to the lesioned data 
+### Post: Within each directory specified by the lesioned data, will have a directory that has all tractography (ROA, ROI, Full)
+### Uses: For use in MS depression - take a subject's mimosa lesions and generate the fiber tracts (individual fascicles) that run through it
+#dependencies: Using dsi studio from docker, sif created by Tim 10/26/2021
+#export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+set -euf -o pipefail
+
+export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
+num_cores=1
+
+#set default paths 
+template='/project/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+default='/project/msdepression/data/melissa_martin_files/csv/mimosa_binary_masks_hcp_space_20211026_n2336'
+#default='/project/msdepression/data/melissa_martin_files/csv/erica_mini_mimosa_paths_n3'
+fascicle_directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa path file" $default
+    lesion_file=$default
+else  
+    lesion_file=$1
+fi
+echo "File being read is "$lesion_file
+echo "... Starting to make lesions ..."
+
+lesion_paths=$(cat $lesion_file)
+
+# loop through each mimosa lesion map; lesion paths contain full paths to mimosa files in hcp space
+job_count=1
+for lesion in ${lesion_paths}; do
+	echo "working on ${lesion} ..."  
+	bsub -J "job_${job_count}" -n ${num_cores} -o /project/msdepression/scripts/logfiles/out_roa_${job_count}.out /project/msdepression/scripts/indiv_mimosa_lesion_dsi_s
+tudio_script.sh $lesion
+   	((job_count+=1))
+done
+
+```
+
+and this calls 
+```{bash indiv_mimosa_lesion_dsi_studio_script.sh}
+
+### The Inner Script... Doing all the dsi studio action for each individual mimosa brain ###
+
+### Pre: path to an individual's mimosa lesion in HCP/MNI space
+### Post: Filtered tracts; ROA and ROI voxel maps/niftis that can be used for data analysis
+### Uses: In trying to find a way to speed everything along, it made the most sense to separate this piece of the script out. 
+#          - it takes a subject's mimosa path, and makes a fiber_tracking_map directory as well as sub directories for each bundle type
+#          - it then runs ROA/ROI filtration and makes output maps.
+
+#set -euf -o pipefail
+module load singularity
+export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
+num_cores=1
+
+#set default directory/templates. This will change when we use local, personal dwi
+template='/project/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+fascicle_directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+if [ $# == 0 ]
+then
+    echo "You did not enter a file. Make sure your call makes sense"
+else  
+    lesion=$1
+
+	echo "File being read is "$lesion
+	echo "... Starting to make tracts"
+# make directory within parent directory for tracts
+    	parent_dir=$(echo ${lesion} | perl -pe s'/(.*)\/.*/$1/g')
+    	echo "$parent_dir"
+	mask_prefix=$(echo ${lesion} | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
+    	echo $mask_prefix
+	if [ ! -d $parent_dir/fiber_tracking_maps ] 
+    	then
+        	mkdir $parent_dir/fiber_tracking_maps
+    	fi
+
+    #make sub directories for each bundle type, so within fiber_tracking_maps, there will be subdirectories for each type of map
+	fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
+    	for fiber_bundle_type in ${fiber_bundle_type_list}; do
+        	echo "Fiber bundle type is " $fiber_bundle_type
+        	echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        	mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        	fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
+
+	# set a job counter to be used to name the jobs
+		job_count=1
+		for fascicle in ${fascicles}; do
+            		echo "Fascicle is  " $fascicle
+            		fascicle_root_name=$(echo $fascicle | perl -pe s'/(.*)\/(.*).tt.gz/$2/g') 
+   			echo  "Fascicle root name is ${fascicle_root_name}" 
+    	    		singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roa=
+$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.tt.gz
+            		singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roi=
+$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.tt.gz
+
+     			singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roa=
+$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.nii.gz
+           		singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roi=
+$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.nii.gz
+
+		# I have removed this line to save space. All hcp fibers are in the directory: /project/msdepression/templates/dti/HCP_YA1065_tractography/ . Can uncomment t
+his if at some point I need to make them for each individual, like when we have their personal DWI
+#singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --output=$parent_dir/fiber_tr
+acking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_full.tt.gz
+			((job_count+=1))
+		done
+    	done
+fi
+```
+
+20211029
+[x] 3dmaskave, get the # of voxels into csvs
+
+scripts are made, just have no more space on pmacs, so waiting for them to make space for me and then I can do this.
+
+20211103
+[x] start working on getting these files into r, MAKE SURE TO ADD A ZERO in the read in line = NA -> 0!!!
+
+20211104
+[x] ton of weird errors that are inconsistent with afni. Need to figure out if they are real or fake.
+
+Here is how I figured out the failed subj/session #
+```{bash}
+more 3dmaskav_*.out | grep "ERROR" | perl -pe 's/.*\/project.*\/(sub-.*\/ses-.*)\/run.*/$1/' | sort -u > failed_3dmaskave_n243_20211103
+```
+
+Have identified the people who failed, it looks like their lesions are empty for these
+```{bash}
+more failed_3dmaskave_n243_20211103 | perl -pe 's/(.*)/\/project\/msdepression\/data\/subj_directories\/$1\/run-001\/mimosa_binary_mask_0.25_mni_hcp.nii.gz/' > failed_full_mimosa_paths_n243
+```
+
+[x]Going to try rerunning these specifically from the dsi_studio command, using these files instead of what is already there
+
+Have to make a file that contains all the full volumes of each of the fascicles
+
+```{bash make_streamline_volumes_for_template.sh}
+### Make streamline volumes ###
+
+### Pre: All template streamlines must have been made in dsi studio (these are the nii files)
+### Post: fiber_volume_values.csv that contains the name of the fiber bundle and its volume
+### Uses: Will be getting the volumes of the streamlines for use in later density analyses
+### Dependencies: This script requires afni to be installed.
+
+module load afni_openmp/20.1
+export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
+num_cores=1
+
+directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
+outfile=${directory}/fiber_volume_values.csv
+
+#remove outfile if it exists
+rm -f ${outfile}
+
+#make the file again
+touch ${outfile}
+echo "fascicle,volume" >> ${outfile}
+
+#go through each fiber, get volume, and store it in csv
+fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
+for fiber_bundle_type in ${fiber_bundle_type_list}; do
+	echo "Fiber bundle type is " $fiber_bundle_type
+fascicle_volumes=$(ls ${directory}/${fiber_bundle_type}/*.nii*)
+        for fascicle in ${fascicle_volumes}; do
+            fascicle_prefix=$(echo $fascicle | perl -pe 's/.*\/(.*).nii.*/$1/g')
+            volume=$(3dmaskave -quiet -mask SELF -sum ${fascicle})
+            echo "${fascicle_prefix},$volume" >> ${outfile}
+        done
+done
+```
+
+output of the above script
+```{bash}
+fascicle,volume
+AF_L,35716
+AF_R,15249
+C_FPH_L,2569
+C_FPH_R,2760
+C_FP_L,14457
+C_FP_R,21386
+C_PH_L,3171
+C_PHP_L,1900
+C_PHP_R,4612
+C_PH_R,2593
+C_R_L,2666
+C_R_R,5181
+EMC_L,7344
+EMC_R,5641
+FAT_L,27142
+FAT_R,15510
+IFOF_L,30311
+IFOF_R,34880
+ILF_L,27401
+ILF_R,36327
+MdLF_L,7637
+MdLF_R,5744
+PAT_L,10921
+PAT_R,13976
+SLF1_L,9836
+SLF1_R,13482
+SLF2_L,26947
+SLF2_R,37091
+SLF3_L,10053
+SLF3_R,15526
+UF_L,7607
+UF_R,7744
+VOF_L,5681
+VOF_R,4079
+CB_L,45470
+CB_R,43251
+ICP_L,4963
+ICP_R,4449
+MCP,16514
+SCP,4315
+V,11299
+CNIII_L,563
+CNIII_R,486
+CNII_L,1097
+CNII_R,1415
+CNVIII_L,1034
+CNVIII_R,968
+CNVII_L,298
+CNVII_R,276
+CNV_L,817
+CNV_R,995
+AR_L,3082
+AR_R,1935
+CBT_L,4609
+CBT_R,6146
+CPT_F_L,18693
+CPT_F_R,34955
+CPT_O_L,9022
+CPT_O_R,7015
+CPT_P_L,24377
+CPT_P_R,27829
+CS_A_L,26114
+CS_A_R,21542
+CS_P_L,5526
+CS_P_R,6231
+CS_S_L,31034
+CS_S_R,34141
+CST_L,20476
+CST_R,16388
+DRTT_L,14986
+DRTT_R,17163
+F_L,5175
+F_R,8018
+ML_L,10871
+ML_R,9139
+OR_L,6876
+OR_R,5510
+RST_L,14766
+RST_R,11890
+TR_A_L,20628
+TR_A_R,14840
+TR_P_L,4402
+TR_P_R,4748
+TR_S_L,39767
+TR_S_R,37036
+AC,26517
+CC,390601
+```
+
+2021104
+[x] need to see if total number of fibers in a bundle = ROA + ROI - NO
+[] if all these work, our first analyses will be depressed vs non total , total fiber volume lost in lesions, then specificity. WOuld look at ratios ROI/total
+
+To initialize Rstudio on pmacs
+```{bash}
+source  /project/bbl_projects/apps/default_modules.sh
+conda activate rstudio
+xbash
+rstudio
+### OR ###
+cd /project/msdepression/scripts
+source load_rstudio
+```
+
+To make list of all roi files
+```{bash}
+cd /project/msdepression/data/melissa_martin_files/csv/
+more mimosa_binary_masks_hcp_space_20211026_n2336 | perl -pe 's/(.*)mimosa_bin.*/$1fiber_tracking_maps\/fiber_volume_values_roi\.csv/' > roi_fiber_volumes_n2336
+```
+
+Rscript to make df with lesions
+
+```{r Roi_ratio_regressions}
+### ROA Ratio Regressions ###
+
+#####
+## Pre: /project/msdepression/data/melissa_martin_files/csv/roi_fiber_volumes_n2336, which contains full paths to the ROI files, can be ammended to do ro
+## Post: Analysis
+## Uses: 1) reads in the ROIs, puts them in a data frame (columns are accession num/date/each fascicle), one row per subjected per date
+#        2) Performers univariate analyses, depression diagnosis versus not
+## Dependencies: R 3.6.3
+
+#include packages
+library(dbplyr)
+
+#set directories
+homedir<-'/project/msdepression/'
+
+#analysis type
+region_type="roi"
+
+#read in template fascicle volumes
+template_full_volumes <- read.csv(paste0(homedir, '/templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv'), header = T)
+fascicle_names <- template_full_volumes$fascicle
+
+### Read in file with all full paths to region volumes
+region_volume_csv <- read.table(paste0(homedir, '/data/melissa_martin_files/csv/', region_type, '_fiber_volumes_n2336'), header = F, stringsAsFactors = F)
+region_volume_csv$V1 <- as.character(region_volume_csv$V1)
+
+#outfile, named for number of subjects
+outfile <- paste0(homedir, "/results/fascicle_volumes_all_subjects_", region_type, "_n", dim(region_volume_csv)[1], ".csv")
+
+#make data frame, 1 row per subject, num rows = num rows in region_volume_csv
+volumes_df <- data.frame(matrix(nrow = dim(region_volume_csv)[1], ncol = 89))
+names(volumes_df) <- c("ACCESSION_NUM", "EXAM_DATE", as.character(fascicle_names))
+#loop through region_volume_csv, read in df corresponding to each row of region_volume_csv, extract accession num and date,
+## and transpose fiber volumes from column to rows. append accession num/date/region_volumes row into big matrix
+
+for (subj in 1:dim(region_volume_csv)[1]) {
+  file_path <- region_volume_csv$V1[subj]
+  region_volume_data <- read.csv(file_path, header = T, sep = ",")
+  accession_num<- region_volume_data[1,1]
+  date <- region_volume_data[1,2]
+  volumes <- t(region_volume_data[,4])
+  volumes[is.na(volumes)] <- 0
+  volumes <- volumes/template_full_volumes$volume
+  row_to_add <- cbind(accession_num, date, volumes)
+  volumes_df[subj,]<- row_to_add
+}
+
+write.csv(file=outfile, volumes_df, quote = FALSE, row.names = FALSE)
+
+```
+
+To see the frequencies of the lesions
+```{bash MC plot}
+library(reshape2)
+library(ggplot2)
+df <- read.csv("fascicle_volumes_all_subjects_roi_n2336.csv", header=TRUE)
+
+m.df <- melt(df, id.vars = c("ACCESSION_NUM", "EXAM_DATE"))
+lesioned <- subset(m.df, value > 0)
+
+ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
+```
+
+20211107
+[] need to change the date format in my regular dac pull
+[] need to divide the ROIs by the total volume 
+[] need to get total # of lesions - this will be tough, need taki's code. Not to do right now - sydney to look into

### 2021-11-13 final_dac_summary_pull_20210119.Rmd
prelim analyses

@@
 author: "Erica Baller"
-date: "4/15/2021"
-output: html_document
+date: "9/29/2021"
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 Date of FINAL Pull: 1/19/2021
-Spreadsheet: /Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv 
+Spreadsheet: /Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv 
 Summary of request to Victoria @ DAC:
@@
 knitr::opts_chunk$set(echo = TRUE)
-
+library(dplyr)
+library(visreg)
+library(lme4)
+library(lmerTest)
+#functions
+source("~/BBL/msdepression/scripts/msdepression_functions.R")
 ```
@@
 #from Jan 2021 pull
-data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
+data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
 
@@
 data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
-
+#get date into better format
+data_empi_acc_f_phq$EXAM_DATE <- as.Date(data_empi_acc_f_phq$BEGIN_EXAM_DTTM, format = "%m/%d/%y")
+data_empi_acc_f_phq$EXAM_DATE <- as.numeric(gsub(data_empi_acc_f_phq$EXAM_DATE, pattern = "-", replacement = ""))
 ########################################
@@
 
+print(paste0("Mean dep + PHQ-2 = ", round(mean(!is.na(dep_df$PHQ.2)),3), "[",round(sd(!is.na(dep_df$PHQ.2)),3), "]; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 79
+
 #phq-9
@@
 
+print(paste0("Mean dep + PHQ-9 = ", round(mean(!is.na(dep_df$PHQ.9)),3), "[",round(sd(!is.na(dep_df$PHQ.9)),3), "]; n = ", length(which(dep_df$PHQ.9 >= 10)), " meeting criteria for depression (PHQ9) >= 10")) #n = 79
+
 #### # unique people with dep dx and phq2/9 -> working on this
@@
 print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 47
+
+print(paste0("Mean/SD healthy w/PHQ-2 = ", round(mean(!is.na(healthy_df$PHQ.2)),3), "[",round(sd(!is.na(healthy_df$PHQ.2)),3), "]; n = ", length(which(healthy_df$PHQ.2>= 3)), " meeting criteria for depression (PHQ2) >= 3"))
 
@@
 
+print(paste0("Mean/SD healthy w/PHQ-9 = ", round(mean(!is.na(healthy_df$PHQ.9)),3), "[",round(sd(!is.na(healthy_df$PHQ.9)),3), "]; n = ", length(which(healthy_df$PHQ.9>= 10)), " meeting criteria for depression (PHQ9) >= 10"))
+
+#statistical testing
+t.test(!is.na(healthy_df$PHQ.2), !is.na(dep_df$PHQ.2))
+t.test(!is.na(healthy_df$PHQ.9), !is.na(dep_df$PHQ.9))
+
+#test first instance of phq2/9
+df_unique_empi_healthy <- healthy_df %>% 
+ group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+df_unique_empi_dep <- dep_df %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+t.test(!is.na(df_unique_empi_healthy$PHQ.2), !is.na(df_unique_empi_dep$PHQ.2))
+t.test(!is.na(df_unique_empi_healthy$PHQ.9), !is.na(df_unique_empi_dep$PHQ.9))
+
+print(paste0("Mean/SD depressed w/PHQ-2/9; healthy w/ph2/9 = ",
+             round(mean(!is.na(df_unique_empi_dep$PHQ.2)),3),"[",round(sd(!is.na(df_unique_empi_dep$PHQ.2)),3), "];", 
+             round(mean(!is.na(df_unique_empi_dep$PHQ.9)),3),
+              "[",round(sd(!is.na(df_unique_empi_dep$PHQ.9)),3), "];",
+             round(mean(!is.na(df_unique_empi_healthy$PHQ.2)),3),
+             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.2)),3), "];",
+             round(mean(!is.na(df_unique_empi_healthy$PHQ.9)),3),
+             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.9)),3), "]"))
+
 #### # unique people without dep dx and phq2/9 
@@
 
+#race
+hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
+     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
+     xlab = "white = 1, other = 2", 
+     breaks = 2)
+
+#f-codes - people usually have BOTH icd9/10 or neither
+
+#Gender breakdown
 #sex
@@
 
-#race
-hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
-     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
-     xlab = "white = 1, other = 2", 
+#sex with depression diagnosis
+hist(dep_df$sex_binarized[which(!is.na(dep_df$sex_binarized))], 
+     main = paste0("Sex Histo in ICD9/10 depression dx group : n = ", length(which(!is.na(dep_df$sex_binarized))), "; M = ", length(which(dep_df$sex_binarized == 1)), "; F = ", length(which(dep_df$sex_binarized == 2)) ),
+     xlab = "Male = 1, Female = 2", 
      breaks = 2)
 
-#f-codes - people usually have BOTH icd9/10 or neither
-
-```
+#sex w/ phq2
+hist(phq2_subset$sex_binarized[which(!is.na(phq2_subset$sex_binarized))], 
+     main = paste0("Sex Histo w/phq2: n = ", length(which(!is.na(phq2_subset$sex_binarized))), "; M = ", length(which(phq2_subset$sex_binarized == 1)), "; F = ", length(which(phq2_subset$sex_binarized == 2)) ),
+     xlab = "Male = 1, Female = 2", 
+     breaks = 2)
+
+#sex w/phq9
+hist(phq9_subset$sex_binarized[which(!is.na(phq9_subset$sex_binarized))], 
+     main = paste0("Sex Histo w/phq9: n = ", length(which(!is.na(phq9_subset$sex_binarized))), "; M = ", length(which(phq9_subset$sex_binarized == 1)), "; F = ", length(which(phq9_subset$sex_binarized == 2)) ),
+     xlab = "Male = 1, Female = 2", 
+     breaks = 2)
+
+```
+
 ```{r frequencies_of_dep_in_phq2_vs_9}
 num_dep_phq2 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & (data_empi_acc_f_phq$PHQ.2 >= 3))) #n = 130
-num_dep_phq9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.9) & (data_empi_acc_f_phq$PHQ.9 >= 5))) #n = 184; https://www.sciencedirect.com/science/article/pii/B9780128179338000293 recommends using cut point of 11
+num_dep_phq9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.9) & (data_empi_acc_f_phq$PHQ.9 >= 5))) #n = 184;
+#https://www.sciencedirect.com/science/article/pii/B9780128179338000293 recommends using cut point of 11
 num_total_phq2 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))
@@
 chisq_dep_phq2_v_phq9$residuals
+
+```
+```{r chi_sq_dep_versus_healthy_PHQ2_and_9}
+num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
+num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
+num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
+num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
+total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
+total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
+total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
+total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;
+
+#### phq2 symptomatic versus asymptomatic
+phq2_table <- data.frame(matrix(data = c(num_dep_phq2,(total_dep_phq2-num_dep_phq2), num_healthy_phq2,   (total_healthy_phq2-num_healthy_phq2)), nrow = 2, ncol = 2))
+names(phq2_table) <- c("dep", "healthy")
+row.names(phq2_table) <- c("symptomatic", "asymptomatic")
+chisq_dep_phq2 <- chisq.test(phq2_table)
+chisq_dep_phq2
+print("observed")
+chisq_dep_phq2$observed
+print("expected")
+chisq_dep_phq2$expected
+print("residuals")
+chisq_dep_phq2$residuals
+
+#### phq9 symptomatic versus asymptomatic
+phq9_table <- data.frame(matrix(data = c(num_dep_phq9,(total_dep_phq9-num_dep_phq9), num_healthy_phq9,   (total_healthy_phq9 - num_healthy_phq9)), nrow = 2, ncol = 2))
+names(phq9_table) <- c("dep", "healthy")
+row.names(phq9_table) <- c("symptomatic", "asymptomatic")
+chisq_dep_phq9 <- chisq.test(phq9_table)
+chisq_dep_phq9
+print("observed")
+chisq_dep_phq9$observed
+print("expected")
+chisq_dep_phq9$expected
+print("residuals")
+chisq_dep_phq9$residuals
+
+################### Unique
+
+num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
+num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
+num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
+num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
+total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
+total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
+total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
+total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;
+
+#patients who are depressed who score on phq2/9
+num_dep_phq2_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.2 >= 3)]))
+num_dep_phq9_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.9 >= 10)]))
+
+#persons who are characterized as healthy but score on PHQ2/9
+num_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.2 >= 3)]))
+num_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.9 >= 10)]))
+
+#total # of people who have phq2/9
+asym_dep_phq2_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.2))])) - num_dep_phq2_unique
+asym_dep_phq9_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.9))])) - num_dep_phq9_unique
+asym_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.2))])) - num_healthy_phq2_unique
+asym_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.9))])) - num_healthy_phq9_unique
+
+#### phq2 symptomatic versus asymptomatic
+phq2_table_unique <- data.frame(matrix(data = c(num_dep_phq2_unique,asym_dep_phq2_unique, num_healthy_phq2_unique, asym_healthy_phq2_unique), nrow = 2, ncol = 2))
+names(phq2_table_unique) <- c("dep", "healthy")
+row.names(phq2_table_unique) <- c("symptomatic", "asymptomatic")
+chisq_dep_phq2_unique <- chisq.test(phq2_table_unique)
+chisq_dep_phq2_unique
+print("observed")
+chisq_dep_phq2_unique$observed
+print("expected")
+chisq_dep_phq2_unique$expected
+print("residuals")
+chisq_dep_phq2_unique$residuals
+
+#### phq9 symptomatic versus asymptomatic
+phq9_table_unique <- data.frame(matrix(data = c(num_dep_phq9_unique,asym_dep_phq9_unique, num_healthy_phq9_unique,   asym_healthy_phq9_unique), nrow = 2, ncol = 2))
+names(phq9_table_unique) <- c("dep", "healthy")
+row.names(phq9_table_unique) <- c("symptomatic", "asymptomatic")
+chisq_dep_phq9_unique <- chisq.test(phq9_table_unique)
+chisq_dep_phq9_unique
+print("observed")
+chisq_dep_phq9_unique$observed
+print("expected")
+chisq_dep_phq9_unique$expected
+print("residuals")
+chisq_dep_phq9_unique$residuals
+
+```
+
+This next step, put together 9/29/2021, aims to define our control and case groups based on the following definitions-
+
+Healthy group - No F33* code, and PHQ2 and 9 always 0
+
+Lifetime depression - EITHER F33* diagnosis OR PHQ2 >= 3 OR PHQ9 >= 10
+   - of note:
+    *PHQ9 0-4 - no depression
+    * 5-9 - mild depression
+    * 10-14 - moderate depression
+    * 15-19 - mod/severe
+    * 20+ - severe
+
+Logic
+  1) First, separate out people with ICD10 code (F3*) into depressed and non depressed
+       - dep_df is people with an F3* code
+       - healthy df is people without
+       - look for unique EMPIs in each of these groups
+
+  2) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ2 
+        - Depression is PHQ2 >=3 at any point
+        - Healthy if PHQ2 = 0
+        - Look for unique EMPIs in these groups
+
+  3) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ9
+        - Depression is PHQ9 >= 10 at any point
+        - Healthy if PHQ9 = 0
+        - Look for unique EMPIs in these groups
+
+  4) Combine for depressed group and healthy
+        - Look for people with PHQ2 or 9 meeting criteria for depression that aren't in depression group, and add them -> dep_num_icd10_plus_phq2_and_phq9
+        - remove people from healthy_df if people had EMPIs in phq2 or 9 depressed group -> healthy_num_no_icd10_plus_phq2_and_phq9
+        - make separate group that only includes healthies for whom we have phq2 or 9 -> healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9
+
+```{r defining_control_and_dep_groups}
+
+##########################################################
+#### Step 1- make F3* depressed and healthy groups* ######
+##########################################################
+
+########
+### separate out depressed group
+########
+print("separate out depressed group")
+dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
+num_depression <- dim(dep_df)[1]
+num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
+print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
+print(paste0("N with ICD 9/10 codes for depression and unique EMPI in whole depression group = ", num_unique_EMPI_and_dep, " out of ", num_depression))
+
+#######
+### separate out healthy group
+#######
+print("separate out healthy group")
+healthy_df <- data_empi_acc_f_phq[-grep("F3", data_empi_acc_f_phq$ICD10),] #n = 14707
+num_healthy <- dim(healthy_df)[1] 
+num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
+print(paste0("N with WITHOUT depression ICD 9/10 codes (i.e. \"healthy\"): ", num_healthy))
+print(paste0("N with WITHOUT ICD 9/10 depression codes for depression and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", num_healthy))
+
+##########################################################
+#### Step 2- Identify PHQ2 >=3 in healthy group     ######
+##########################################################
+
+print("Get people in healthy group with phq2s for dep/healthy")
+# num in healthy group with PHQ2
+phq2_subset <- healthy_df[which(!is.na(healthy_df$PHQ.2)),] #n=3646, unique 698
+depressed_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 >= 3),] #n = 47 
+depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
+healthy_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 == 0),] #n = 3326
+healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631
+
+print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]"))
+
+##########################################################
+#### Step 3- Identify PHQ9 >=10 in healthy group    ######
+##########################################################
+
+print("Get people in healthy group with phq9s for dep/healthy")
+# num in healthy group with PHQ9
+phq9_subset <- healthy_df[which(!is.na(healthy_df$PHQ.9)),] #n=121
+depressed_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 >= 10),] #n = 32 
+depressed_unique_phq9_empis <- unique(depressed_phq9_subset$EMPI) #n=9
+healthy_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 == 0),] #n = 21
+healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n =4
+
+print(paste0("PHQ9 breakdown in healthy group (total n with score = ", dim(phq9_subset)[1], "[unique = ", length(unique(phq9_subset$EMPI)),"]): PHQ9 >= 10 :", dim(depressed_phq9_subset)[1], "[unique = ", length(depressed_unique_phq9_empis), "]; PHQ9 == 0 : ", dim(healthy_phq9_subset)[1], "[unique = ", length(unique(healthy_phq9_subset$EMPI)),"]"))
+
+##########################################################
+#Step 4 - Combine for depressed group and healthy groups #
+##########################################################
+
+print("Combine groups for new cohort")
+gain_in_dep_group <- length(depressed_unique_phq2_empis) + length(depressed_unique_phq9_empis)
+dep_num_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_dep + gain_in_dep_group # we can do this because we know they are each exclusive - num_unique_empi_and_dep from icd10 group, and no overlap in phq2 and 9 group
+print(paste0("number of UNIQUE people with an icd10 code for Depression OR scored positive on PHQ2(>=3) or PHQ9(>=10) : ", dep_num_icd10_plus_phq2_and_phq9, ", for a gain of n = ", gain_in_dep_group))
+
+healthy_num_no_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_healthy - gain_in_dep_group #take number of people who don't have depression diagnosis and don't have a depression phq2/9
+
+print(paste0("number of UNIQUE people with NO icd10 code for Depression and NEVER had a PHQ2(>=3) or PHQ9(>=10) : ", healthy_num_no_icd10_plus_phq2_and_phq9, ", for a loss of n = ", gain_in_dep_group))
+
+healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9 <- length(healthy_unique_phq2_empis) + length(healthy_unique_phq9_empis)
+print(paste0("number of UNIQUE people with NO icd10 code for Depression and HAS HAD AT LEAST 1 PHQ2 or 9 that was 0) : ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
+
+print(paste0("Cleanest cohort (depressed in icd9/10 or Phq2/9; healthy by at least one phq2/9 and never depression dx ---->>>>  Depressed: ", dep_num_icd10_plus_phq2_and_phq9, "; Healthy: ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
+
+##### THESE ARE THE DATAFRAMES I WANT TO USE GOING FORWARD ###
+#all empis for depressed group include people with depresssion dx, and people who got in for phq2 or phq9
+empis_for_depressed_group <- append(
+  append(unique(dep_df$EMPI), depressed_unique_phq2_empis), depressed_unique_phq9_empis)
+
+final_depressed_group_withICD_AND_depressed_phq2_or_9 <- 
+  data_empi_acc_f_phq %>%
+filter(EMPI %in% empis_for_depressed_group)
+
+#------
+empis_for_healthy_group <- append(healthy_unique_phq2_empis, healthy_unique_phq9_empis)
+
+final_healthy_group_withNOICD_AND_phq2_or_9_0<- 
+  data_empi_acc_f_phq %>%
+filter(EMPI %in% empis_for_healthy_group)
+
+#------
+### add a column to the original dataframe with a 0 if not in any group, -1 if healthy, and 1 if depressed
+data_empi_acc_f_phq$depGroupVar = 0
+data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_depressed_group)] = 1
+data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_healthy_group)] = -1
+data_empi_acc_f_phq$depGroupVar = as.factor(data_empi_acc_f_phq$depGroupVar)
+
 ```
@@
 
-#race
-hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
-     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
-     xlab = "white = 1, other = 2", 
-     breaks = 2)
-
-#f-codes - people usually have BOTH icd9/10 or neither
-
-```
+```
+
 ```{r comparing_out_group_with_Tim_group}
@@
 
-segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression_with_accession.csv", header = F, sep = ",") #n = 4468
+#segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression_with_accession.csv", header = F, sep = ",") #n = 4468
+segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression3.csv", header = F, sep = ",") #n = 2833
+
 names(segmented) <- c("EMPI", "EXAM_DATE", "SES", "ACCESSION_NUM")
@@
 overlap_group <- data_empi_acc_f_phq[which(data_empi_acc_f_phq$ACCESSION_NUM %in% segmented$ACCESSION_NUM),]
+
+print(paste0("Num people in previously curated group and Erica's group -> ", dim(overlap_group)[1]))
 
@@
 
-print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of ", dim(overlap_group)[1]))
+print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of total depressed - ", dim(dep_df)[1]))
 
@@
 ```
+
+```{r depression_vs_healthy_melissa_qc_group}
+
+### this chunk can be run separately from the rest of the analysis
+### A lot of it is repeat, but I wanted it to be independent
+### Started 20211119
+
+homedir <- "/Users/eballer/BBL/msdepression/"
+source("~/BBL/msdepression/scripts/msdepression_functions.R")
+
+##########################################################
+####                   Prepare the date               ####
+##########################################################
+#load our df
+
+#from Jan 2021 pull
+data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
+
+ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
+
+#keep people with MS providers 
+data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
+
+#keep people with ICD10 code G35 (this is the same # as people with IC9 - 340)
+data_empi_acc_f_phq <- data_empi_acc_f_phq[grep("G35", data_empi_acc_f_phq$ICD10),] #n = 16830
+table(data_empi_acc_f_phq$Provider)
+
+##########################
+### some preprocessing ###
+##########################
+
+#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
+data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
+data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
+data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
+data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
+data_empi_acc_f_phq$osex <- ordered(data_empi_acc_f_phq$sex_binarized)
+data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
+#get date into better format
+data_empi_acc_f_phq$EXAM_DATE <- as.Date(data_empi_acc_f_phq$BEGIN_EXAM_DTTM, format = "%m/%d/%y")
+data_empi_acc_f_phq$EXAM_DATE <- as.numeric(gsub(data_empi_acc_f_phq$EXAM_DATE, pattern = "-", replacement = ""))
+
+##########################################################
+#### Step 1- make F3* depressed and healthy groups* ######
+##########################################################
+  ### This is identical to up above, but repeated so I can run this chunk separately
+
+########
+### separate out depressed group
+########
+print("separate out depressed group")
+dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
+num_depression <- dim(dep_df)[1]
+num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
+print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
+print(paste0("N with ICD 9/10 codes for depression and unique EMPI in whole depression group = ", num_unique_EMPI_and_dep, " out of ", num_depression))
+
+#find number of participants who meet criteria for depression in the PHQ2/9 symptom rating
+print(paste0("N with ICD 9/10 codes for depression, PHQ2 >= 3, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.2 >= 3])))))
+print(paste0("N with ICD 9/10 codes for depression, PHQ9 >= 10, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.9 >= 10])))))
+
+#######
+### separate out healthy group
+#######
+print("separate out healthy group")
+healthy_df <- data_empi_acc_f_phq[-grep("F3", data_empi_acc_f_phq$ICD10),] #n = 14707
+num_healthy <- dim(healthy_df)[1] 
+num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
+print(paste0("N with WITHOUT depression ICD 9/10 codes (i.e. \"healthy\"): ", num_healthy))
+print(paste0("N with WITHOUT ICD 9/10 depression codes for depression and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", num_healthy))
+
+##########################################################
+#### Step 2- Identify PHQ2 >=3 in healthy group     ######
+##########################################################
+
+print("Get people in healthy group with phq2s for dep/healthy")
+# num in healthy group with PHQ2
+phq2_subset <- healthy_df[which(!is.na(healthy_df$PHQ.2)),] #n=3646, unique 698
+depressed_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 >= 3),] #n = 47 
+depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
+healthy_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 == 0),] #n = 3326
+healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631
+
+print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]"))
+
+##########################################################
+#### Step 3- Identify PHQ9 >=10 in healthy group    ######
+##########################################################
+
+print("Get people in healthy group with phq9s for dep/healthy")
+# num in healthy group with PHQ9
+phq9_subset <- healthy_df[which(!is.na(healthy_df$PHQ.9)),] #n=121
+depressed_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 >= 10),] #n = 32 
+depressed_unique_phq9_empis <- unique(depressed_phq9_subset$EMPI) #n=9
+healthy_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 == 0),] #n = 21
+healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n =4
+
+print(paste0("PHQ9 breakdown in healthy group (total n with score = ", dim(phq9_subset)[1], "[unique = ", length(unique(phq9_subset$EMPI)),"]): PHQ9 >= 10 :", dim(depressed_phq9_subset)[1], "[unique = ", length(depressed_unique_phq9_empis), "]; PHQ9 == 0 : ", dim(healthy_phq9_subset)[1], "[unique = ", length(unique(healthy_phq9_subset$EMPI)),"]"))
+
+##########################################################
+#Step 4 - Combine for depressed group and healthy groups #
+##########################################################
+
+print("Combine groups for new cohort")
+gain_in_dep_group <- length(depressed_unique_phq2_empis) + length(depressed_unique_phq9_empis)
+dep_num_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_dep + gain_in_dep_group # we can do this because we know they are each exclusive - num_unique_empi_and_dep from icd10 group, and no overlap in phq2 and 9 group
+print(paste0("number of UNIQUE people with an icd10 code for Depression OR scored positive on PHQ2(>=3) or PHQ9(>=10) : ", dep_num_icd10_plus_phq2_and_phq9, ", for a gain of n = ", gain_in_dep_group))
+
+healthy_num_no_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_healthy - gain_in_dep_group #take number of people who don't have depression diagnosis and don't have a depression phq2/9
+
+print(paste0("number of people with NO icd10 code for Depression and NEVER had a PHQ2(>=3) or PHQ9(>=10) : ", healthy_num_no_icd10_plus_phq2_and_phq9, ", for a loss of n = ", gain_in_dep_group))
+
+healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9 <- length(healthy_unique_phq2_empis) + length(healthy_unique_phq9_empis)
+print(paste0("number of UNIQUE people with NO icd10 code for Depression and HAS HAD AT LEAST 1 PHQ2 or 9 that was 0) : ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
+
+print(paste0("Cleanest cohort (depressed in icd9/10 or Phq2/9; healthy by at least one phq2/9 and never depression dx UNIQUE ---->>>>  Depressed: ", dep_num_icd10_plus_phq2_and_phq9, "; Healthy: ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
+
+##### THESE ARE THE DATAFRAMES I WANT TO USE GOING FORWARD ###
+#all empis for depressed group include people with depresssion dx, and people who got in for phq2 or phq9
+empis_for_depressed_group <- append(
+  append(unique(dep_df$EMPI), depressed_unique_phq2_empis), depressed_unique_phq9_empis)
+
+final_depressed_group_withICD_AND_depressed_phq2_or_9 <- 
+  data_empi_acc_f_phq %>%
+filter(EMPI %in% empis_for_depressed_group)
+
+#------
+empis_for_healthy_group <- append(healthy_unique_phq2_empis, healthy_unique_phq9_empis)
+
+final_healthy_group_withNOICD_AND_phq2_or_9_0<- 
+  data_empi_acc_f_phq %>%
+filter(EMPI %in% empis_for_healthy_group)
+
+#------
+### add a column to the original dataframe with a 0 if not in any group, -1 if healthy, and 1 if depressed
+data_empi_acc_f_phq$depGroupVar = 0
+data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_depressed_group)] = 1
+data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_healthy_group)] = -1
+data_empi_acc_f_phq$depGroupVar = ordered(data_empi_acc_f_phq$depGroupVar)
+
+##########################################################
+#             Step 5 - Read in Fascicle info             #
+##########################################################
+
+fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
+
+#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
+fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
+
+#fascicle_mapping, returns numerical mapping as well as names of tracts
+fascicle_bundle_mapping <- get_fascicle_bundle_mapping()
+##########################################################
+#             Step 6 - Merge with data_empi              #
+##########################################################
+df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962
+
+##########################################################
+#                 Step 7 - Histos                        #
+##########################################################
+##### Histograms for PHQ2/9 healthy and depressed
+
+hist(df_demo_and_fascicles$PHQ.2[which(!is.na(df_demo_and_fascicles$PHQ.2))], main = paste0("PHQ-2 In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.2)))), xlab = "PHQ-2", breaks = 8)
+
+hist(df_demo_and_fascicles$PHQ.9[which(!is.na(df_demo_and_fascicles$PHQ.9))], main = paste0("PHQ-9 In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.9)))), xlab = "PHQ-9", breaks = 8)
+
+hist(df_demo_and_fascicles$PHQ.2[which(!is.na(df_demo_and_fascicles$PHQ.2) & (df_demo_and_fascicles$depGroupVar == 1))], main = paste0("PHQ-2/Depressed In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.2) & df_demo_and_fascicles$depGroupVar == 1))), xlab = "PHQ-2", breaks = 8)
+
+hist(df_demo_and_fascicles$PHQ.9[which(!is.na(df_demo_and_fascicles$PHQ.9) & (df_demo_and_fascicles$depGroupVar == 1))], main = paste0("PHQ-9/Depressed In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.9) & df_demo_and_fascicles$depGroupVar == 1))), xlab = "PHQ-9", breaks = 8)
+
+hist(df_demo_and_fascicles$PHQ.2[which(!is.na(df_demo_and_fascicles$PHQ.2) & (df_demo_and_fascicles$depGroupVar == -1))], main = paste0("PHQ-2/Healthy In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.2) & df_demo_and_fascicles$depGroupVar == -1))), xlab = "PHQ-2", breaks = 8)
+
+hist(df_demo_and_fascicles$PHQ.9[which(!is.na(df_demo_and_fascicles$PHQ.9) & (df_demo_and_fascicles$depGroupVar == -1))], main = paste0("PHQ-9/Healthy In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.9) & df_demo_and_fascicles$depGroupVar == -1))), xlab = "PHQ-9", breaks = 8)
+
+##########################################################
+#   Step 8 - Regressions in dep/non dep                  #
+##########################################################
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthy = 713, total n = 1106
+
+#lm
+fascicle_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_lm) <- fascicle_names
+
+#anova
+fascicle_anova <- lapply(fascicle_lm, anova)
+
+#fdr corrected
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_lm, visreg)
+```
+
+```{r phq2_correlation_fascicles}
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthy = 713, total n = 1106
+
+#filter those with phq2
+with_phq2 <- dep_and_healthy_groups_for_ICD_analysis %>% filter(!is.na(PHQ.2))
+
+#lm
+fascicle_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PHQ.2, list(i = as.name(x))), data = with_phq2)
+})
+names(fascicle_lm) <- fascicle_names
+
+#anova
+fascicle_anova <- lapply(fascicle_lm, anova)
+
+#fdr corrected
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
+fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
+print(fascicle_anova_p05_unc)
+
+#visreg
+#sapply(fascicle_lm, visreg)
+```
+
+```{r phq9_correlation_fascicles}
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthy = 713, total n = 1106
+
+#filter those with phq2
+with_phq9 <- dep_and_healthy_groups_for_ICD_analysis %>% filter(!is.na(PHQ.9))
+
+#lm
+fascicle_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
+})
+names(fascicle_lm) <- fascicle_names
+
+#anova
+fascicle_anova <- lapply(fascicle_lm, anova)
+
+#fdr corrected
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
+fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
+print(fascicle_anova_p05_unc)
+
+#visreg
+#sapply(fascicle_lm, visreg)
+```
+
+```{r depression_main_effect_unique_subjects}
+##### repeat above analysis but use only the first instance of each EMPI, sorted by date
+df_unique_empi <- dep_and_healthy_groups_for_ICD_analysis %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n = 300, 115 depressed, 185 healthy
+
+#lm
+fascicle_lm_unique <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ depGroupVar, list(i = as.name(x))), data = df_unique_empi)
+})
+names(fascicle_lm_unique) <- fascicle_names
+
+#anova
+fascicle_anova_unique <- lapply(fascicle_lm_unique, anova)
+
+#fdr corrected
+fascicle_anova_unique_fdr <- fdr_anova_generic(fascicle_anova_unique, 1)
+print(fascicle_anova_unique_fdr)#anova values < 0.05, uncorrected
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_unique_p <- sapply(fascicle_anova_unique, function(v) v$"Pr(>F)"[1])
+fascicle_anova_unique_p05_unc <- as.data.frame(fascicle_anova_unique_p[fascicle_anova_unique_p < 0.05])
+print(fascicle_anova_unique_p05_unc)
+
+```
+```{r depression_main_effect_mixed_model}
+##### repeat above analysis but use only the first instance of each EMPI, sorted by date
+df_mm <-df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]
+
+#lm
+fascicle_lm_mm <- lapply(fascicle_names, function(x) 
+{
+  lmerTest::lmer(substitute(i ~ depGroupVar + (1 | EMPI), list(i = as.name(x))), data = df_mm)
+})
+names(fascicle_lm_mm) <- fascicle_names
+
+#anova
+fascicle_anova_mm <- lapply(fascicle_lm_mm, anova)
+
+#fdr corrected
+fascicle_anova_mm_fdr <- fdr_anova_generic(fascicle_anova_mm, 1)
+print(fascicle_anova_mm_fdr)#anova values < 0.05, uncorrected
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_mm_p <- sapply(fascicle_anova_mm, function(v) v$"Pr(>F)"[1])
+fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
+print(fascicle_anova_mm_p05_unc)
+
+```
+
+```{r age_effects_fascicles}
+########### age effect whole group
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthhy = 713, total n = 1106
+
+#lm
+fascicle_age_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_age_lm) <- fascicle_names
+
+#anova
+fascicle_age_anova <- lapply(fascicle_age_lm, anova)
+
+#fdr corrected
+fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
+print(fascicle_age_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+```
+
+```{r sex_effects_fascicles}
+########### sex effect whole group
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthhy = 713, total n = 1106
+
+#lm
+fascicle_sex_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sex_lm) <- fascicle_names
+
+#anova
+fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)
+
+#fdr corrected
+fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
+print(fascicle_sex_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+```
+
+```{r sex_by_depression_fascicles}
+#sex by depression
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthhy = 713, total n = 1106
+
+#lm
+fascicle_sexdep_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ osex*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sexdep_lm) <- fascicle_names
+
+#anova
+fascicle_sexdep_anova <- lapply(fascicle_sexdep_lm, anova)
+
+#fdr corrected
+fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 3)
+print(fascicle_sexdep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+```
+
+```{r visreg_models}
+
+#visreg
+#sapply(fascicle_lm, visreg)
+#sapply(fascicle_age_lm, visreg)
+#sapply(fascicle_sex_lm, visreg)
+#sapply(fascicle_sexdep_lm, visreg)
+#sapply(fascicle_lm_unique, visreg)
+
+```

### 2021-11-13 final_dac_summary_pull_20210119.Rmd
prelim analyses

@@
 library(lmerTest)
+library(ggplot2)
 #functions
@@
 data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_healthy_group)] = -1
-data_empi_acc_f_phq$depGroupVar = ordered(data_empi_acc_f_phq$depGroupVar)
+data_empi_acc_f_phq$depGroupVar = as.factor(data_empi_acc_f_phq$depGroupVar)
 
@@
 ##########################################################
-#   Step 8 - Regressions in dep/non dep                  #
-##########################################################
+#                Step 8 - Regressions                    #
+##########################################################
+
+```
+
+```{r dep_vs_healthy_regressions}
+
 #isolate out only the depressed group and healthy group
@@
 #visreg
-#sapply(fascicle_lm, visreg)
+sapply(fascicle_lm, visreg)
 ```
@@
 ```
+```{r phq2_correlation_fascicles_dep}
+#isolate out only the depressed group and healthy group
+dep_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar == 1,] #n depressed = 393, healthy = 713, total n = 1106
+
+#filter those with phq2
+with_phq2_dep <- dep_groups_for_ICD_analysis %>% filter(!is.na(PHQ.2) & PHQ.2 != 0) %>% filter(depGroupVar == 1)
+
+#lm
+fascicle_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PHQ.2, list(i = as.name(x))), data = with_phq2_dep)
+})
+names(fascicle_lm) <- fascicle_names
+
+#anova
+fascicle_anova <- lapply(fascicle_lm, anova)
+
+#fdr corrected
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
+fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
+print(fascicle_anova_p05_unc)
+
+#visreg
+#sapply(fascicle_lm, visreg)
+```
 
@@
   lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
+})
+names(fascicle_lm) <- fascicle_names
+
+#anova
+fascicle_anova <- lapply(fascicle_lm, anova)
+
+#fdr corrected
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
+fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
+print(fascicle_anova_p05_unc)
+
+#visreg
+#sapply(fascicle_lm, visreg)
+```
+```{r phq9_correlation_fascicles_dep}
+#isolate out only the depressed group and healthy group
+dep_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar == 1,] #n depressed = 393, healthy = 713, total n = 1106
+
+#filter those with phq2
+with_phq9_dep <- dep_groups_for_ICD_analysis %>% filter(!is.na(PHQ.9) & PHQ.9 != 0)
+
+#lm
+fascicle_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9_dep)
 })
@@
 #isolate out only the depressed group and healthy group
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthhy = 713, total n = 1106
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthy = 713, total n = 1106
+
+#### age histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$PAT_AGE_AT_EXAM,
+main = paste0("Age In Good Mimosa Group : n = ", dim(dep_and_healthy_groups_for_ICD_analysis)[1]), xlab = "Age", breaks = 10)
 
@@
 print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_age_lm,visreg)
 ```
@@
 dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthhy = 713, total n = 1106
+
+#### sex histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
+main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)
 
@@
 print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_sex_lm,visreg)
 ```
@@
 ```
+```{r age_by_depression_fascicles}
+#sex by depression
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthhy = 713, total n = 1106
+
+#lm
+fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_agedep_lm) <- fascicle_names
+
+#anova
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+
+#fdr corrected
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+```
 

### 2021-11-17 final_dac_summary_pull_20210119.Rmd
making subsets for Tim

@@
 library(ggplot2)
+library(reshape2)
 #functions
@@
 
+#------ Group for Tim to check quality issues ----------
+write.csv(data_empi_acc_f_phq, "/Users/eballer/BBL/msdepression/data/dac/data_empi_acc_f_q.csv", col.names = T, sep = ",")
+
+subset <- subset(data_empi_acc_f_phq, select = c("ACCESSION_NUM","EMPI","EXAM_DATE","depGroupVar"))
+
+melissa_qc_data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/just_empi_date_rating", sep = ",", header = F) #n = 2841
+names(melissa_qc_data) <- c("EMPI", "EXAM_DATE", "rating")
+subset_with_qc <- merge(subset, melissa_qi_data, by = c("EMPI", "EXAM_DATE"))
+
+write.csv(subset_with_qc, "/Users/eballer/BBL/msdepression/data/dac/subset_with_qc_for_tim_20211117.csv", col.names = T)
+
 ##########################################################
@@
 ##########################################################
-#                Step 8 - Regressions                    #
+#          Step 8 - Visualize all Tracts                 #
+##########################################################
+just_dep_and_empi <- subset(data_empi_acc_f_phq, select = c("EMPI", "depGroupVar"))
+added_dep_to_fascicles <- merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI"))
+melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
+lesioned <- subset(melted_df, value > 0)
+lesioned_dx <- subset(lesioned, depGroupVar != 0)
+healthy <- subset(lesioned, depGroupVar == -1)
+depressed <- subset(lesioned, depGroupVar == 1)
+
+q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
+r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
+x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
+y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
+z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
+print(q)
+print(r)
+print(x)
+print(y)
+print(z)
+
+##########################################################
+#                Step 9 - Regressions                    #
 ##########################################################
@@
 
+```{r age_effects_fascicles_unique}
+########### age effect whole group
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#### age histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$PAT_AGE_AT_EXAM,
+main = paste0("Age In Good Mimosa Group : n = ", dim(dep_and_healthy_groups_for_ICD_analysis)[1]), xlab = "Age", breaks = 10)
+
+#lm
+fascicle_age_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_age_lm) <- fascicle_names
+
+#anova
+fascicle_age_anova <- lapply(fascicle_age_lm, anova)
+
+#fdr corrected
+fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
+print(fascicle_age_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_age_lm,visreg)
+```
+
+```{r age_effects_fascicles_mm}
+########### age effect whole group
+
+#isolate out only the depressed group and healthy group
+
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthy = 713, total n = 1106
+
+#### age histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$PAT_AGE_AT_EXAM,
+main = paste0("Age In Good Mimosa Group : n = ", dim(dep_and_healthy_groups_for_ICD_analysis)[1]), xlab = "Age", breaks = 10)
+
+#lm
+fascicle_age_lm <- lapply(fascicle_names, function(x) 
+{
+  lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_age_lm) <- fascicle_names
+
+#anova
+fascicle_age_anova <- lapply(fascicle_age_lm, anova)
+
+#fdr corrected
+fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
+print(fascicle_age_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_age_lm,visreg)
+```
+
 ```{r sex_effects_fascicles}
@@
 
+```{r sex_effects_fascicles_unique}
+########### sex effect whole group
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#### sex histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
+main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)
+
+#lm
+fascicle_sex_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sex_lm) <- fascicle_names
+
+#anova
+fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)
+
+#fdr corrected
+fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
+print(fascicle_sex_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_sex_lm,visreg)
+```
+
+```{r sex_effects_fascicles_mm}
+########### sex effect whole group
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,] #n depressed = 393, healthhy = 713, total n = 1106
+
+#### sex histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
+main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)
+
+#lm
+fascicle_sex_lm <- lapply(fascicle_names, function(x) 
+{
+  lmerTest::lmer(substitute(i ~ osex + (1|EMPI), list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sex_lm) <- fascicle_names
+
+#anova
+fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)
+
+#fdr corrected
+fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
+print(fascicle_sex_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_sex_lm,visreg)
+```
+
 ```{r sex_by_depression_fascicles}

### 2021-11-18 msdepression_regressions_started_20211118.Rmd
branched out of my big dac summary script for ease of doing things

--- 2021-11-17 final_dac_summary_pull_20210119.Rmd
@@
 ---
-title: "Dac summary 01/19/2021"
+title: "MS Regressions"
 author: "Erica Baller"
-date: "9/29/2021"
-output:
-  pdf_document: default
-  html_document: default
----
-
-title: "Dac Summary"
+date: "11/18/2021"
 output:
@@
 
-```{r summary}
-#preprocessing - because these are sent
-
+```{r depression_vs_healthy_melissa_qc_group}
+
+homedir <- "/Users/eballer/BBL/msdepression/"
+
+##########################################################
+####                   Prepare the date               ####
+##########################################################
 #load our df
@@
 #from Jan 2021 pull
-data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
+data <- read.csv(paste0(homedir,"/data/dac/investigatingdepressioninmspatients_dates_right_format.csv"), header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
 
@@
 
-#keep people with MS providers 
-data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
-
-#keep people with ICD10 code G35 (this is the same # as people with IC9 - 340)
-data_empi_acc_f_phq <- data_empi_acc_f_phq[grep("G35", data_empi_acc_f_phq$ICD10),] #n = 16830
-table(data_empi_acc_f_phq$Provider)
+columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")
 
@@
 ##########################
-
-#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
-data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
-data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
-data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
-data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
-data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
-#get date into better format
-data_empi_acc_f_phq$EXAM_DATE <- as.Date(data_empi_acc_f_phq$BEGIN_EXAM_DTTM, format = "%m/%d/%y")
-data_empi_acc_f_phq$EXAM_DATE <- as.numeric(gsub(data_empi_acc_f_phq$EXAM_DATE, pattern = "-", replacement = ""))
-########################################
-#### Start counting / making subsets ###
-########################################
-
-# number of unique scanners and scans
-print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 10
-print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME)))) #n = 3
-print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY)))) #n = 11
-print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 16830
-print(paste0("Number of unique EMPIS n = ", length(unique(data_empi_acc_f_phq$EMPI)))) #n = 3737
-
-#### PHQ-2 and PHQ-9 summaries
-
-#looking at just people with phq-2 and 9
-phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4606
-
-phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n= 261
-
-#phq-2;
-print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)), "; n = ", length(which(phq2_subset$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-#phq-9
-print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
-             length(which(phq9_subset$PHQ.9 < 5)), "/",
-             length(which(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)), "/",
-             length(which(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)), "/",
-             length(which(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)), "/",
-             length(which(phq9_subset$PHQ.9 >= 20)))) #n = 261
-
-#both phq2 and 9
-
-num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
-print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0
-
-#Number of unique EMPIs in phq2/phq9 subsets
-unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
-print(paste0("N with phq2 (Unique EMPI) n = ", unique_phq2_empis, " out of ", dim(phq2_subset)[1], "; n - ", length(unique(phq2_subset$EMPI[phq2_subset$PHQ.2 >= 3])), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76
-print(paste0("N with phq9 (Unique EMPI) n = ", unique_phq9_empis, " out of ", dim(phq9_subset)[1], "; noDep/mild/mod/mod-severe/severe = ", 
-             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 < 5])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)])), "/",
-             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 >= 20]))))
-
-#### ICD 9/10 for Depression
-dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
-num_depression <- dim(dep_df)[1]
-unique_empi <- which(unique(data_empi_acc_f_phq$EMPI) %in% data_empi_acc_f_phq$EMPI) #index of the locations of unique empis; n = 3885
-num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
-
-print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
-
-print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of ", dim(data_empi_acc_f_phq)[1]))
-
-##### PHQ2 or 9 + ICD 9/10 Depression
-num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 960
-print(paste0("N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = ", num_phq2_and_dep, " out of ", dim(dep_df)[1]))
-
-num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=140
-print(paste0("N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = ", num_phq9_and_dep, " out of ", dim(dep_df)[1]))
-
-#phq-2;
-print(paste0("N dep + PHQ-2 = ", sum(!is.na(dep_df$PHQ.2)), "; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-print(paste0("Mean dep + PHQ-2 = ", round(mean(!is.na(dep_df$PHQ.2)),3), "[",round(sd(!is.na(dep_df$PHQ.2)),3), "]; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 79
-
-#phq-9
-print(paste0("N dep + PHQ-9 = ", sum(!is.na(dep_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
-             length(which(dep_df$PHQ.9 < 5)), "/",
-             length(which(dep_df$PHQ.9 >= 5 & dep_df$PHQ.9 < 10)), "/",
-             length(which(dep_df$PHQ.9 >= 10 & dep_df$PHQ.9 < 15)), "/",
-             length(which(dep_df$PHQ.9 >= 15 & dep_df$PHQ.9 < 20)), "/",
-             length(which(dep_df$PHQ.9 >= 20)))) #n = 261
-
-print(paste0("Mean dep + PHQ-9 = ", round(mean(!is.na(dep_df$PHQ.9)),3), "[",round(sd(!is.na(dep_df$PHQ.9)),3), "]; n = ", length(which(dep_df$PHQ.9 >= 10)), " meeting criteria for depression (PHQ9) >= 10")) #n = 79
-
-#### # unique people with dep dx and phq2/9 -> working on this
-dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 201
-print(paste0("N with ICD 9/10 codes for depression and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(dep_df)[1]))
-
-dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 43
-print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))
-
-####
-
-#### ICD 9/10 for Healthy Group
-healthy_df <- data_empi_acc_f_phq[(-grep("F3", data_empi_acc_f_phq$ICD10)),] #n = 14707
-num_healthy <- dim(healthy_df)[1]
-unique_empi <- which(unique(healthy_df$EMPI) %in% healthy_df$EMPI) #index of the locations of unique empis; n = 3244
-num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
-
-print(paste0("N without Depression/ICD 9/10 codes for healthy = ", num_healthy))
-
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", dim(healthy_df)[1]))
-
-##### PHQ2 or 9 + No Depression ICD9/10
-num_phq2_and_healthy <- length(which(!is.na(healthy_df$PHQ.2))) #n = 3646
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = ", num_phq2_and_healthy, " out of ", dim(healthy_df)[1]))
-
-#phq-2;
-print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 47
-
-print(paste0("Mean/SD healthy w/PHQ-2 = ", round(mean(!is.na(healthy_df$PHQ.2)),3), "[",round(sd(!is.na(healthy_df$PHQ.2)),3), "]; n = ", length(which(healthy_df$PHQ.2>= 3)), " meeting criteria for depression (PHQ2) >= 3"))
-
-num_phq9_and_healthy <- length(which(!is.na(healthy_df$PHQ.9))) #n=121
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = ", num_phq9_and_healthy, " out of ", dim(healthy_df)[1]))
-
-#phq-9
-print(paste0("N healthy w/PHQ-9 = ", sum(!is.na(healthy_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
-             length(which(healthy_df$PHQ.9 < 5)), "/",
-             length(which(healthy_df$PHQ.9 >= 5 & healthy_df$PHQ.9 < 10)), "/",
-             length(which(healthy_df$PHQ.9 >= 10 & healthy_df$PHQ.9 < 15)), "/",
-             length(which(healthy_df$PHQ.9 >= 15 & healthy_df$PHQ.9 < 20)), "/",
-             length(which(healthy_df$PHQ.9 >= 20)))) #n = 55/34/21/1/10
-
-print(paste0("Mean/SD healthy w/PHQ-9 = ", round(mean(!is.na(healthy_df$PHQ.9)),3), "[",round(sd(!is.na(healthy_df$PHQ.9)),3), "]; n = ", length(which(healthy_df$PHQ.9>= 10)), " meeting criteria for depression (PHQ9) >= 10"))
-
-#statistical testing
-t.test(!is.na(healthy_df$PHQ.2), !is.na(dep_df$PHQ.2))
-t.test(!is.na(healthy_df$PHQ.9), !is.na(dep_df$PHQ.9))
-
-#test first instance of phq2/9
-df_unique_empi_healthy <- healthy_df %>% 
- group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-df_unique_empi_dep <- dep_df %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-t.test(!is.na(df_unique_empi_healthy$PHQ.2), !is.na(df_unique_empi_dep$PHQ.2))
-t.test(!is.na(df_unique_empi_healthy$PHQ.9), !is.na(df_unique_empi_dep$PHQ.9))
-
-print(paste0("Mean/SD depressed w/PHQ-2/9; healthy w/ph2/9 = ",
-             round(mean(!is.na(df_unique_empi_dep$PHQ.2)),3),"[",round(sd(!is.na(df_unique_empi_dep$PHQ.2)),3), "];", 
-             round(mean(!is.na(df_unique_empi_dep$PHQ.9)),3),
-              "[",round(sd(!is.na(df_unique_empi_dep$PHQ.9)),3), "];",
-             round(mean(!is.na(df_unique_empi_healthy$PHQ.2)),3),
-             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.2)),3), "];",
-             round(mean(!is.na(df_unique_empi_healthy$PHQ.9)),3),
-             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.9)),3), "]"))
-
-#### # unique people without dep dx and phq2/9 
-dep_dx_phq2_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.2)])) #n = 201
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(healthy_df)[1]))
-
-dep_dx_phq9_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.9)])) #n = 43
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(healthy_df)[1]))
-
-####
-
-#demographics and lab historgrams
-#print histogram of phq-2/9s for people
-
-hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)
-
-hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)
-
-#fcode histos
-hist(dep_df$PHQ.2, main = paste0("PHQ2 and Dep Histo : n = ", num_phq2_and_dep), xlab = paste0("PHQ2"))
-
-hist(dep_df$PHQ.9,main = paste0("PHQ9 and Dep Histo : n = ", num_phq9_and_dep), xlab = paste0("PHQ9"))
-
-###
-
-#age histogram
-hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")
-
-#tsh
-hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")
-
-#hemoglobin
-hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")
-
-#wbc
-hist(data_empi_acc_f_phq$WBC[which(!is.na(data_empi_acc_f_phq$WBC))], main = paste0("WBC Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$WBC)))), xlab = "WBC")
-
-#folate
-hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)
-
-#vitamin D
-hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$VitD)))), xlab = "Vitamin D", breaks = 10)
-
-#race
-hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
-     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
-     xlab = "white = 1, other = 2", 
-     breaks = 2)
-
-#f-codes - people usually have BOTH icd9/10 or neither
-
-#Gender breakdown
-#sex
-hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
-     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ),
-     xlab = "Male = 1, Female = 2", 
-     breaks = 2)
-
-#sex with depression diagnosis
-hist(dep_df$sex_binarized[which(!is.na(dep_df$sex_binarized))], 
-     main = paste0("Sex Histo in ICD9/10 depression dx group : n = ", length(which(!is.na(dep_df$sex_binarized))), "; M = ", length(which(dep_df$sex_binarized == 1)), "; F = ", length(which(dep_df$sex_binarized == 2)) ),
-     xlab = "Male = 1, Female = 2", 
-     breaks = 2)
-
-#sex w/ phq2
-hist(phq2_subset$sex_binarized[which(!is.na(phq2_subset$sex_binarized))], 
-     main = paste0("Sex Histo w/phq2: n = ", length(which(!is.na(phq2_subset$sex_binarized))), "; M = ", length(which(phq2_subset$sex_binarized == 1)), "; F = ", length(which(phq2_subset$sex_binarized == 2)) ),
-     xlab = "Male = 1, Female = 2", 
-     breaks = 2)
-
-#sex w/phq9
-hist(phq9_subset$sex_binarized[which(!is.na(phq9_subset$sex_binarized))], 
-     main = paste0("Sex Histo w/phq9: n = ", length(which(!is.na(phq9_subset$sex_binarized))), "; M = ", length(which(phq9_subset$sex_binarized == 1)), "; F = ", length(which(phq9_subset$sex_binarized == 2)) ),
-     xlab = "Male = 1, Female = 2", 
-     breaks = 2)
-
-```
-
-```{r frequencies_of_dep_in_phq2_vs_9}
-num_dep_phq2 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & (data_empi_acc_f_phq$PHQ.2 >= 3))) #n = 130
-num_dep_phq9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.9) & (data_empi_acc_f_phq$PHQ.9 >= 5))) #n = 184;
-#https://www.sciencedirect.com/science/article/pii/B9780128179338000293 recommends using cut point of 11
-num_total_phq2 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))
-num_total_phq9 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.9))) 
-dep_total_phq_table <- data.frame(matrix(data = c(num_dep_phq2, num_dep_phq9, num_total_phq2, num_total_phq9), nrow = 2, ncol = 2))
-names(dep_total_phq_table) <- c("dep", "healthy")
-row.names(dep_total_phq_table) <- c("phq2", "phq9")
-chisq_dep_phq2_v_phq9 <- chisq.test(dep_total_phq_table)
-chisq_dep_phq2_v_phq9
-print("observed")
-chisq_dep_phq2_v_phq9$observed
-print("expected")
-chisq_dep_phq2_v_phq9$expected
-print("residuals")
-chisq_dep_phq2_v_phq9$residuals
-
-```
-```{r chi_sq_dep_versus_healthy_PHQ2_and_9}
-num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
-num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
-num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
-num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
-total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
-total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
-total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
-total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;
-
-#### phq2 symptomatic versus asymptomatic
-phq2_table <- data.frame(matrix(data = c(num_dep_phq2,(total_dep_phq2-num_dep_phq2), num_healthy_phq2,   (total_healthy_phq2-num_healthy_phq2)), nrow = 2, ncol = 2))
-names(phq2_table) <- c("dep", "healthy")
-row.names(phq2_table) <- c("symptomatic", "asymptomatic")
-chisq_dep_phq2 <- chisq.test(phq2_table)
-chisq_dep_phq2
-print("observed")
-chisq_dep_phq2$observed
-print("expected")
-chisq_dep_phq2$expected
-print("residuals")
-chisq_dep_phq2$residuals
-
-#### phq9 symptomatic versus asymptomatic
-phq9_table <- data.frame(matrix(data = c(num_dep_phq9,(total_dep_phq9-num_dep_phq9), num_healthy_phq9,   (total_healthy_phq9 - num_healthy_phq9)), nrow = 2, ncol = 2))
-names(phq9_table) <- c("dep", "healthy")
-row.names(phq9_table) <- c("symptomatic", "asymptomatic")
-chisq_dep_phq9 <- chisq.test(phq9_table)
-chisq_dep_phq9
-print("observed")
-chisq_dep_phq9$observed
-print("expected")
-chisq_dep_phq9$expected
-print("residuals")
-chisq_dep_phq9$residuals
-
-################### Unique
-
-num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
-num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
-num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
-num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
-total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
-total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
-total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
-total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;
-
-#patients who are depressed who score on phq2/9
-num_dep_phq2_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.2 >= 3)]))
-num_dep_phq9_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.9 >= 10)]))
-
-#persons who are characterized as healthy but score on PHQ2/9
-num_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.2 >= 3)]))
-num_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.9 >= 10)]))
-
-#total # of people who have phq2/9
-asym_dep_phq2_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.2))])) - num_dep_phq2_unique
-asym_dep_phq9_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.9))])) - num_dep_phq9_unique
-asym_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.2))])) - num_healthy_phq2_unique
-asym_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.9))])) - num_healthy_phq9_unique
-
-#### phq2 symptomatic versus asymptomatic
-phq2_table_unique <- data.frame(matrix(data = c(num_dep_phq2_unique,asym_dep_phq2_unique, num_healthy_phq2_unique, asym_healthy_phq2_unique), nrow = 2, ncol = 2))
-names(phq2_table_unique) <- c("dep", "healthy")
-row.names(phq2_table_unique) <- c("symptomatic", "asymptomatic")
-chisq_dep_phq2_unique <- chisq.test(phq2_table_unique)
-chisq_dep_phq2_unique
-print("observed")
-chisq_dep_phq2_unique$observed
-print("expected")
-chisq_dep_phq2_unique$expected
-print("residuals")
-chisq_dep_phq2_unique$residuals
-
-#### phq9 symptomatic versus asymptomatic
-phq9_table_unique <- data.frame(matrix(data = c(num_dep_phq9_unique,asym_dep_phq9_unique, num_healthy_phq9_unique,   asym_healthy_phq9_unique), nrow = 2, ncol = 2))
-names(phq9_table_unique) <- c("dep", "healthy")
-row.names(phq9_table_unique) <- c("symptomatic", "asymptomatic")
-chisq_dep_phq9_unique <- chisq.test(phq9_table_unique)
-chisq_dep_phq9_unique
-print("observed")
-chisq_dep_phq9_unique$observed
-print("expected")
-chisq_dep_phq9_unique$expected
-print("residuals")
-chisq_dep_phq9_unique$residuals
-
-```
-
-This next step, put together 9/29/2021, aims to define our control and case groups based on the following definitions-
-
-Healthy group - No F33* code, and PHQ2 and 9 always 0
-
-Lifetime depression - EITHER F33* diagnosis OR PHQ2 >= 3 OR PHQ9 >= 10
-   - of note:
-    *PHQ9 0-4 - no depression
-    * 5-9 - mild depression
-    * 10-14 - moderate depression
-    * 15-19 - mod/severe
-    * 20+ - severe
-
-Logic
-  1) First, separate out people with ICD10 code (F3*) into depressed and non depressed
-       - dep_df is people with an F3* code
-       - healthy df is people without
-       - look for unique EMPIs in each of these groups
-
-  2) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ2 
-        - Depression is PHQ2 >=3 at any point
-        - Healthy if PHQ2 = 0
-        - Look for unique EMPIs in these groups
-
-  3) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ9
-        - Depression is PHQ9 >= 10 at any point
-        - Healthy if PHQ9 = 0
-        - Look for unique EMPIs in these groups
-
-  4) Combine for depressed group and healthy
-        - Look for people with PHQ2 or 9 meeting criteria for depression that aren't in depression group, and add them -> dep_num_icd10_plus_phq2_and_phq9
-        - remove people from healthy_df if people had EMPIs in phq2 or 9 depressed group -> healthy_num_no_icd10_plus_phq2_and_phq9
-        - make separate group that only includes healthies for whom we have phq2 or 9 -> healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9
-
-```{r defining_control_and_dep_groups}
+#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
+#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as
+
+#goal is to keep people who were seen by 
+ data_empi_acc_f_phq <- data %>% 
+   filter(Provider %in% ms_providers) %>%       #n=17,067
+   filter(grepl("G35", ICD10)) %>%              #n=16830
+   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
+   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
+   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
+   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
+   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
+   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
+   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = ""))
 
@@
 ##########################################################
+  ### This is identical to up above, but repeated so I can run this chunk separately
 
@@
 print("separate out depressed group")
-dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
+dep_df <- data_empi_acc_f_phq %>% filter(grepl("F3", ICD10)) #n = 2123
 num_depression <- dim(dep_df)[1]
@@
 
+#find number of participants who meet criteria for depression in the PHQ2/9 symptom rating
+print(paste0("N with ICD 9/10 codes for depression, PHQ2 >= 3, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.2 >= 3]))))) #n=23
+print(paste0("N with ICD 9/10 codes for depression, PHQ9 >= 10, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.9 >= 10])))))#n=26
+
 #######
@@
 print("separate out healthy group")
-healthy_df <- data_empi_acc_f_phq[-grep("F3", data_empi_acc_f_phq$ICD10),] #n = 14707
+healthy_df <- data_empi_acc_f_phq %>% filter(!grepl("F3", ICD10)) #n = 14707
 num_healthy <- dim(healthy_df)[1] 
@@
 # num in healthy group with PHQ2
-phq2_subset <- healthy_df[which(!is.na(healthy_df$PHQ.2)),] #n=3646, unique 698
-depressed_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 >= 3),] #n = 47 
+phq2_subset <- healthy_df %>% filter(!is.na(PHQ.2)) #n=3646, unique 698
+depressed_phq2_subset <- phq2_subset %>% filter(PHQ.2 >= 3) #n = 47 
 depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
-healthy_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 == 0),] #n = 3326
+healthy_phq2_subset <- phq2_subset %>% filter(PHQ.2 == 0) #n = 3326
 healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631
 
-print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]"))
+print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]")) #PHQ2 breakdown in healthy group (total n with score = 3646[unique = 698]): PHQ2 >= 3 :47[unique = 10]; PHQ2 == 0 : 3326[unique = 631]
 
@@
 # num in healthy group with PHQ9
-phq9_subset <- healthy_df[which(!is.na(healthy_df$PHQ.9)),] #n=121
-depressed_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 >= 10),] #n = 32 
+phq9_subset <- healthy_df %>% filter(!is.na(PHQ.9)) #n=121 (unique 5)
+depressed_phq9_subset <- phq9_subset %>% filter(PHQ.9 >= 10) #n = 32 
 depressed_unique_phq9_empis <- unique(depressed_phq9_subset$EMPI) #n=9
-healthy_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 == 0),] #n = 21
-healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n =4
+healthy_phq9_subset <- phq9_subset %>% filter(PHQ.9 == 0) #n = 21
+healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n = 4
 
@@
 
-print(paste0("number of UNIQUE people with NO icd10 code for Depression and NEVER had a PHQ2(>=3) or PHQ9(>=10) : ", healthy_num_no_icd10_plus_phq2_and_phq9, ", for a loss of n = ", gain_in_dep_group))
+print(paste0("number of people with NO icd10 code for Depression and NEVER had a PHQ2(>=3) or PHQ9(>=10) : ", healthy_num_no_icd10_plus_phq2_and_phq9, ", for a loss of n = ", gain_in_dep_group))
 
@@
 
-print(paste0("Cleanest cohort (depressed in icd9/10 or Phq2/9; healthy by at least one phq2/9 and never depression dx ---->>>>  Depressed: ", dep_num_icd10_plus_phq2_and_phq9, "; Healthy: ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
+print(paste0("Cleanest cohort (depressed in icd9/10 or Phq2/9; healthy by at least one phq2/9 and never depression dx UNIQUE ---->>>>  Depressed: ", dep_num_icd10_plus_phq2_and_phq9, "; Healthy: ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
 
@@
 
-```
-
-```{r sex_differences_regressions}
-
-### looking at anyone with phq2s or 9s
-males.phq2 <- phq2_subset[which(phq2_subset$SEX == "MALE"),] #n = 951, mean = 0.20
-females.phq2 <- phq2_subset[which(phq2_subset$SEX == "FEMALE"),] #n = 3655, mean = 0.26
-m_v_f_phq2 <- t.test(males.phq2$PHQ.2, females.phq2$PHQ.2) #p = 0.03 ***
-
-males.phq9 <- phq9_subset[which(phq9_subset$SEX == "MALE"),] #n = 80, mean = 8.63
-females.phq9 <- phq9_subset[which(phq9_subset$SEX == "FEMALE"),] #n = 181, mean = 9.13
-m_v_f_phq9 <- t.test(males.phq9$PHQ.9, females.phq9$PHQ.9)
-
-#### Looking at depressed group
-## interstingly, not the 2:1 ratio we usually see in general world
-## phq2 means suggest people are not depressed 
-## phq2 means suggest people meet criteria for mild depression
-
-males_dep <- dep_df[(which(dep_df$SEX == "MALE")),] #n = 419/total(4087; 10%) F; mean PHQ2 = 0.61, mean PHQ9 = 13.05
-females_dep <- dep_df[(which(dep_df$SEX == "FEMALE")),] #n = 1704/total(12743; 13%)total F; mean = 0.56; mean PHQ9 = 10.61
-m_v_f_dep_phq2 <- t.test(males_dep$PHQ.2, females_dep$PHQ.2) #p = 0.66
-m_v_f_dep_phq9 <- t.test(males_dep$PHQ.9, females_dep$PHQ.9) #p = 0.24
-
-#chi sq comparing # of pts dx with depression in males vs females; col 1- male dep n, female dep n; col 2- male total n, female total n
-### fascinating - only 12% of people even gave a dx of depression in their charts. More likely to have it if female than male; also phq2 mean only slightly higher in females. 
-num_dep_m <- dim(males_dep)[1] #n = 419
-num_dep_f <-  dim(females_dep)[1]#n = 1704
-num_total_m <- length(which(data_empi_acc_f_phq$SEX == "MALE")) #n = 4087
-num_total_f <- length(which(data_empi_acc_f_phq$SEX == "FEMALE")) #n = 
-m_v_f_table <- data.frame(matrix(data = c(num_dep_m, num_dep_f, num_total_m, num_total_f), nrow = 2, ncol = 2))
-names(m_v_f_table) <- c("dep", "total")
-row.names(m_v_f_table) <- c("male", "female")
-chisq_m_v_f_dist <- chisq.test(m_v_f_table) #p = 4 x 10^-6
-chisq_m_v_f_dist
-print("observed")
-chisq_m_v_f_dist$observed
-print("expected")
-chisq_m_v_f_dist$expected
-print("residuals")
-chisq_m_v_f_dist$residuals
-#chisq comparing # of
-
-```
-
-```{r summary_psychosis}
-#preprocessing - because these are sent
-
-#load our df
-
-#from Jan 2021 pull
-data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
-
-ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
-
-#keep people with MS providers 
-data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
-
-#keep people with ICD10 code G35 (this is the same # as people with IC9 - 340)
-data_empi_acc_f_phq <- data_empi_acc_f_phq[grep("G35", data_empi_acc_f_phq$ICD10),] #n = 16830
-table(data_empi_acc_f_phq$Provider)
-
-##########################
-### some preprocessing ###
-##########################
-
-#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
-data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
-data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
-data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
-data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
-data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
-
-########################################
-#### Start counting / making subsets ###
-########################################
-
-# number of unique scanners and scans
-print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 10
-print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME)))) #n = 3
-print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY)))) #n = 11
-print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 16830
-print(paste0("Number of unique EMPIS n = ", length(unique(data_empi_acc_f_phq$EMPI)))) #n = 3737
-
-#### PHQ-2 and PHQ-9 summaries
-
-#looking at just people with phq-2 and 9
-phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4606
-
-phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n= 261
-
-#phq-2;
-print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)), "; n = ", length(which(phq2_subset$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-#phq-9
-print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
-             length(which(phq9_subset$PHQ.9 < 5)), "/",
-             length(which(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)), "/",
-             length(which(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)), "/",
-             length(which(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)), "/",
-             length(which(phq9_subset$PHQ.9 >= 20)))) #n = 261
-
-#both phq2 and 9
-
-num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
-print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0
-
-#Number of unique EMPIs in phq2/phq9 subsets
-unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
-print(paste0("N with phq2 (Unique EMPI) n = ", unique_phq2_empis, " out of ", dim(phq2_subset)[1], "; n - ", length(unique(phq2_subset$EMPI[phq2_subset$PHQ.2 >= 3])), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76
-print(paste0("N with phq9 (Unique EMPI) n = ", unique_phq9_empis, " out of ", dim(phq9_subset)[1], "; noDep/mild/mod/mod-severe/severe = ", 
-             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 < 5])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)])), "/",
-             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 >= 20]))))
-
-#### ICD 9/10 for Psychosis
-psychosis_df <- data_empi_acc_f_phq[grep("F2", data_empi_acc_f_phq$ICD10),] #n = 2123
-num_psychosis <- dim(psychosis_df)[1]
-unique_empi <- which(unique(data_empi_acc_f_phq$EMPI) %in% data_empi_acc_f_phq$EMPI) #index of the locations of unique empis; n = 3885
-num_unique_EMPI_and_psychosis <- length(unique(psychosis_df$EMPI)) #n = 493
-
-print(paste0("N with ICD 9/10 codes for psychosis = ", num_psychosis))
-
-print(paste0("N with ICD 9/10 codes for psychosis and unique EMPI = ", num_unique_EMPI_and_psychosis, " out of ", dim(data_empi_acc_f_phq)[1]))
-
-##### PHQ2 or 9 + ICD 9/10 psychosis
-num_phq2_and_psychosis <- length(which(!is.na(psychosis_df$PHQ.2))) #n = 960
-print(paste0("N with ICD 9/10 codes for psychosis and phq2 (ALL SCANS) = ", num_phq2_and_psychosis, " out of ", dim(psychosis_df)[1]))
-
-num_phq9_and_psychosis <- length(which(!is.na(psychosis_df$PHQ.9))) #n=140
-print(paste0("N with ICD 9/10 codes for psychosis and phq9 (ALL SCANS) = ", num_phq9_and_psychosis, " out of ", dim(psychosis_df)[1]))
-
-#phq-2;
-print(paste0("N psychosis + PHQ-2 = ", sum(!is.na(psychosis_df$PHQ.2)), "; n = ", length(which(psychosis_df$PHQ.2 >= 3)), " meeting criteria for psychosis (PHQ2) >= 3")) #n = 4606
-
-#phq-9
-print(paste0("N psychosis + PHQ-9 = ", sum(!is.na(psychosis_df$PHQ.9)), "; nopsychosis/mild/mod/mod-severe/severe = ", 
-             length(which(psychosis_df$PHQ.9 < 5)), "/",
-             length(which(psychosis_df$PHQ.9 >= 5 & psychosis_df$PHQ.9 < 10)), "/",
-             length(which(psychosis_df$PHQ.9 >= 10 & psychosis_df$PHQ.9 < 15)), "/",
-             length(which(psychosis_df$PHQ.9 >= 15 & psychosis_df$PHQ.9 < 20)), "/",
-             length(which(psychosis_df$PHQ.9 >= 20)))) #n = 261
-
-#### # unique people with psychosis dx and phq2/9 -> working on this
-psychosis_dx_phq2_unique <- length(unique(psychosis_df$EMPI[!is.na(psychosis_df$PHQ.2)])) #n = 201
-print(paste0("N with ICD 9/10 codes for psychosis and phq2 (unique) = ", psychosis_dx_phq2_unique, " out of ", dim(psychosis_df)[1]))
-
-psychosis_dx_phq9_unique <- length(unique(psychosis_df$EMPI[!is.na(psychosis_df$PHQ.9)])) #n = 43
-print(paste0("N with ICD 9/10 codes for psychosis and phq9 (unique) = ", psychosis_dx_phq9_unique, " out of ", dim(psychosis_df)[1]))
-
-####
-
-#### ICD 9/10 for Healthy Group
-healthy_df <- data_empi_acc_f_phq[(-grep("F3", data_empi_acc_f_phq$ICD10)),] #n = 14707
-num_healthy <- dim(healthy_df)[1]
-unique_empi <- which(unique(healthy_df$EMPI) %in% healthy_df$EMPI) #index of the locations of unique empis; n = 3244
-num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
-
-print(paste0("N without psychosis/ICD 9/10 codes for healthy = ", num_healthy))
-
-print(paste0("N without psychosis/ ICD 9/10 codes for healthy and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", dim(healthy_df)[1]))
-
-##### PHQ2 or 9 + No psychosis ICD9/10
-num_phq2_and_healthy <- length(which(!is.na(healthy_df$PHQ.2))) #n = 3646
-print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = ", num_phq2_and_healthy, " out of ", dim(healthy_df)[1]))
-
-#phq-2;
-print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for psychosis (PHQ2) >= 3")) #n = 47
-
-num_phq9_and_healthy <- length(which(!is.na(healthy_df$PHQ.9))) #n=121
-print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = ", num_phq9_and_healthy, " out of ", dim(healthy_df)[1]))
-
-#phq-9
-print(paste0("N healthy w/PHQ-9 = ", sum(!is.na(healthy_df$PHQ.9)), "; nopsychosis/mild/mod/mod-severe/severe = ", 
-             length(which(healthy_df$PHQ.9 < 5)), "/",
-             length(which(healthy_df$PHQ.9 >= 5 & healthy_df$PHQ.9 < 10)), "/",
-             length(which(healthy_df$PHQ.9 >= 10 & healthy_df$PHQ.9 < 15)), "/",
-             length(which(healthy_df$PHQ.9 >= 15 & healthy_df$PHQ.9 < 20)), "/",
-             length(which(healthy_df$PHQ.9 >= 20)))) #n = 55/34/21/1/10
-
-#### # unique people without psychosis dx and phq2/9 
-psychosis_dx_phq2_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.2)])) #n = 201
-print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq2 (unique) = ", psychosis_dx_phq2_unique, " out of ", dim(healthy_df)[1]))
-
-psychosis_dx_phq9_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.9)])) #n = 43
-print(paste0("N without psychosis/ ICD 9/10 codes for healthy and phq9 (unique) = ", psychosis_dx_phq9_unique, " out of ", dim(healthy_df)[1]))
-
-####
-
-#demographics and lab historgrams
-#print histogram of phq-2/9s for people
-
-hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)
-
-hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)
-
-#fcode histos - no one with it at this time. 
-#hist(psychosis_df$PHQ.2, main = paste0("PHQ2 and psychosis Histo : n = ", num_phq2_and_psychosis), xlab = paste0("PHQ2"))
-
-#hist(psychosis_df$PHQ.9,main = paste0("PHQ9 and psychosis Histo : n = ", num_phq9_and_psychosis), xlab = paste0("PHQ9"))
-
-###
-
-#age histogram
-hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")
-
-#tsh
-hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")
-
-#hemoglobin
-hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")
-
-#wbc
-hist(data_empi_acc_f_phq$WBC[which(!is.na(data_empi_acc_f_phq$WBC))], main = paste0("WBC Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$WBC)))), xlab = "WBC")
-
-#folate
-hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)
-
-#vitamin D
-hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$VitD)))), xlab = "Vitamin D", breaks = 10)
-
-#race
-hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
-     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
-     xlab = "white = 1, other = 2", 
-     breaks = 2)
-
-#f-codes - people usually have BOTH icd9/10 or neither
-
-#Gender breakdown
-#sex
-hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
-     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ),
-     xlab = "Male = 1, Female = 2", 
-     breaks = 2)
-
-```
-
-```{r comparing_out_group_with_Tim_group}
-##### checking to see whether we have clinical data for patients who have already been segmented as of 5/4/2021
-### had to do some parsing of tim's spreadsheet: /Users/eballer/BBL/msdepression/data/insurance_images_for_depression.csv
-### used the following commands to extract the empi and the date -> in tim's they are in the format sub-***** and ses-YYYYMMDD
-
-## more insurance_images_for_depression.csv | cut -f1 -d ','| perl -pe 's/sub-(.*)/$1/' > tim_empi
-## more insurance_images_for_depression.csv | cut -f2 -d ','| perl -pe 's/ses-(....)(..)(..)/$2\/$3\/$1/' > tim_dates
-## more tim_dates | perl -pe 's/(.*)20(..)/$1$2/' | perl -pe 's/(0)(.)\/(.*)\/(.*)/$2\/$3\/$4/' | perl -pe 's/(.*)\/(0)(.*)\/(.*)/$1\/$3\/$4/' > tim_dates_no_zero_pad
-
-#segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression_with_accession.csv", header = F, sep = ",") #n = 4468
-segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression3.csv", header = F, sep = ",") #n = 2833
-
-names(segmented) <- c("EMPI", "EXAM_DATE", "SES", "ACCESSION_NUM")
-
-#look at the overlap of EMPIs. Won't be able to actually check the matching of scan to person without either accession #s or changing the format of the date. 
-number_of_people_in_tims_group_that_are_also_in_mine <- length(which(segmented$ACCESSION_NUM %in% data_empi_acc_f_phq$ACCESSION_NUM)) #2919
-overlap_group <- data_empi_acc_f_phq[which(data_empi_acc_f_phq$ACCESSION_NUM %in% segmented$ACCESSION_NUM),]
-
-print(paste0("Num people in previously curated group and Erica's group -> ", dim(overlap_group)[1]))
-
-#### PHQ-2 and PHQ-9 summaries
-
-#looking at just people with phq-2 and 9
-phq2_subset <- overlap_group[which(!is.na(overlap_group$PHQ.2)),] #n=4606
-
-phq9_subset <- overlap_group[which(!is.na(overlap_group$PHQ.9)),] #n= 261
-
-#phq-2;
-print(paste0("N PHQ-2 = ", sum(!is.na(overlap_group$PHQ.2)), "; n = ", length(which(phq2_subset$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-#phq-9
-print(paste0("N PHQ-9 = ", sum(!is.na(overlap_group$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
-             length(which(phq9_subset$PHQ.9 < 5)), "/",
-             length(which(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)), "/",
-             length(which(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)), "/",
-             length(which(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)), "/",
-             length(which(phq9_subset$PHQ.9 >= 20)))) #n = 261
-
-#both phq2 and 9
-
-num_both_phq2_and_9 <- length(which(!is.na(overlap_group$PHQ.2) & !is.na(overlap_group$PHQ.9)))
-print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0
-
-#Number of unique EMPIs in phq2/phq9 subsets
-unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
-print(paste0("N with phq2 (Unique EMPI) n = ", unique_phq2_empis, " out of ", dim(phq2_subset)[1], "; n - ", length(unique(phq2_subset$EMPI[phq2_subset$PHQ.2 >= 3])), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76
-print(paste0("N with phq9 (Unique EMPI) n = ", unique_phq9_empis, " out of ", dim(phq9_subset)[1], "; noDep/mild/mod/mod-severe/severe = ", 
-             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 < 5])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 5 & phq9_subset$PHQ.9 < 10)])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 10 & phq9_subset$PHQ.9 < 15)])), "/",
-             length(unique(phq9_subset$EMPI[(phq9_subset$PHQ.9 >= 15 & phq9_subset$PHQ.9 < 20)])), "/",
-             length(unique(phq9_subset$EMPI[phq9_subset$PHQ.9 >= 20]))))
-
-#### ICD 9/10 for Depression
-dep_df <- overlap_group[grep("F3", overlap_group$ICD10),] #n = 2123
-num_depression <- dim(dep_df)[1]
-unique_empi <- which(unique(overlap_group$EMPI) %in% overlap_group$EMPI) #index of the locations of unique empis; n = 3885
-num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
-
-print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
-
-print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of total depressed - ", dim(dep_df)[1]))
-
-##### PHQ2 or 9 + ICD 9/10 Depression
-num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 960
-print(paste0("N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = ", num_phq2_and_dep, " out of ", dim(dep_df)[1]))
-
-num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=140
-print(paste0("N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = ", num_phq9_and_dep, " out of ", dim(dep_df)[1]))
-
-#phq-2;
-print(paste0("N dep + PHQ-2 = ", sum(!is.na(dep_df$PHQ.2)), "; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 4606
-
-#phq-9
-print(paste0("N dep + PHQ-9 = ", sum(!is.na(dep_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
-             length(which(dep_df$PHQ.9 < 5)), "/",
-             length(which(dep_df$PHQ.9 >= 5 & dep_df$PHQ.9 < 10)), "/",
-             length(which(dep_df$PHQ.9 >= 10 & dep_df$PHQ.9 < 15)), "/",
-             length(which(dep_df$PHQ.9 >= 15 & dep_df$PHQ.9 < 20)), "/",
-             length(which(dep_df$PHQ.9 >= 20)))) #n = 261
-
-#### # unique people with dep dx and phq2/9 -> working on this
-dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 201
-print(paste0("N with ICD 9/10 codes for depression and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(dep_df)[1]))
-
-dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 43
-print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))
-
-####
-
-#### ICD 9/10 for Healthy Group
-healthy_df <- overlap_group[(-grep("F3", overlap_group$ICD10)),] #n = 14707
-num_healthy <- dim(healthy_df)[1]
-unique_empi <- which(unique(healthy_df$EMPI) %in% healthy_df$EMPI) #index of the locations of unique empis; n = 3244
-num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
-
-print(paste0("N without Depression/ICD 9/10 codes for healthy = ", num_healthy))
-
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", dim(healthy_df)[1]))
-
-##### PHQ2 or 9 + No Depression ICD9/10
-num_phq2_and_healthy <- length(which(!is.na(healthy_df$PHQ.2))) #n = 3646
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (ALL SCANS) = ", num_phq2_and_healthy, " out of ", dim(healthy_df)[1]))
-
-#phq-2;
-print(paste0("N healthy w/PHQ-2 = ", sum(!is.na(healthy_df$PHQ.2)), "; n = ", length(which(healthy_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 47
-
-num_phq9_and_healthy <- length(which(!is.na(healthy_df$PHQ.9))) #n=121
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (ALL SCANS) = ", num_phq9_and_healthy, " out of ", dim(healthy_df)[1]))
-
-#phq-9
-print(paste0("N healthy w/PHQ-9 = ", sum(!is.na(healthy_df$PHQ.9)), "; noDep/mild/mod/mod-severe/severe = ", 
-             length(which(healthy_df$PHQ.9 < 5)), "/",
-             length(which(healthy_df$PHQ.9 >= 5 & healthy_df$PHQ.9 < 10)), "/",
-             length(which(healthy_df$PHQ.9 >= 10 & healthy_df$PHQ.9 < 15)), "/",
-             length(which(healthy_df$PHQ.9 >= 15 & healthy_df$PHQ.9 < 20)), "/",
-             length(which(healthy_df$PHQ.9 >= 20)))) #n = 55/34/21/1/10
-
-#### # unique people without dep dx and phq2/9 
-dep_dx_phq2_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.2)])) #n = 201
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(healthy_df)[1]))
-
-dep_dx_phq9_unique <- length(unique(healthy_df$EMPI[!is.na(healthy_df$PHQ.9)])) #n = 43
-print(paste0("N without Depression/ ICD 9/10 codes for healthy and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(healthy_df)[1]))
-
-####
-
-#demographics and lab historgrams
-#print histogram of phq-2/9s for people
-
-hist(overlap_group$PHQ.2[which(!is.na(overlap_group$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(overlap_group$PHQ.2)))), xlab = "PHQ-2", breaks = 8)
-
-hist(overlap_group$PHQ.9[which(!is.na(overlap_group$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(overlap_group$PHQ.9)))), xlab = "PHQ-9", breaks = 30)
-
-#fcode histos
-hist(dep_df$PHQ.2, main = paste0("PHQ2 and Dep Histo : n = ", num_phq2_and_dep), xlab = paste0("PHQ2"))
-
-hist(dep_df$PHQ.9,main = paste0("PHQ9 and Dep Histo : n = ", num_phq9_and_dep), xlab = paste0("PHQ9"))
-
-###
-
-#age histogram
-hist(overlap_group$MRI_ENC_AGE[which(!is.na(overlap_group$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(overlap_group$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")
-
-#tsh
-hist(overlap_group$TSH[which(!is.na(overlap_group$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(overlap_group$TSH)))), xlab = "TSH")
-
-#hemoglobin
-hist(overlap_group$hemoglobin[which(!is.na(overlap_group$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(overlap_group$hemoglobin)))), xlab = "hemoglobin")
-
-#wbc
-hist(overlap_group$WBC[which(!is.na(overlap_group$WBC))], main = paste0("WBC Histo : n = ", length(which(!is.na(overlap_group$WBC)))), xlab = "WBC")
-
-#folate
-hist(overlap_group$FOLATE[which(!is.na(overlap_group$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(overlap_group$FOLATE)))), xlab = "folate", breaks = 10)
-
-#vitamin D
-hist(overlap_group$VitD[which(!is.na(overlap_group$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(overlap_group$VitD)))), xlab = "Vitamin D", breaks = 10)
-
-#sex
-hist(overlap_group$sex_binarized[which(!is.na(overlap_group$sex_binarized))], 
-     main = paste0("Sex Histo : n = ", length(which(!is.na(overlap_group$sex_binarized))), "; M = ", length(which(overlap_group$sex_binarized == 1)), "; F = ", length(which(overlap_group$sex_binarized == 2)) ),
-     xlab = "Male = 1, Female = 2", 
-     breaks = 2)
-
-#race
-hist(overlap_group$race_binarized[which(!is.na(overlap_group$race_binarized))], 
-     main = paste0("Race Histo : n = ", length(which(!is.na(overlap_group$race))), "; White = ", length(which(overlap_group$race_binarized == 1)), "; NW = ", length(which(overlap_group$race_binarized == 2))),
-     xlab = "white = 1, other = 2", 
-     breaks = 2)
-
-#f-codes - people usually have BOTH icd9/10 or neither
-
-```
-
-```{r depression_vs_healthy_melissa_qc_group}
-
-### this chunk can be run separately from the rest of the analysis
-### A lot of it is repeat, but I wanted it to be independent
-### Started 20211119
-
-homedir <- "/Users/eballer/BBL/msdepression/"
-source("~/BBL/msdepression/scripts/msdepression_functions.R")
-
-##########################################################
-####                   Prepare the date               ####
-##########################################################
-#load our df
-
-#from Jan 2021 pull
-data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")
-
-ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
-
-#keep people with MS providers 
-data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
-
-#keep people with ICD10 code G35 (this is the same # as people with IC9 - 340)
-data_empi_acc_f_phq <- data_empi_acc_f_phq[grep("G35", data_empi_acc_f_phq$ICD10),] #n = 16830
-table(data_empi_acc_f_phq$Provider)
-
-##########################
-### some preprocessing ###
-##########################
-
-#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
-data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
-data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
-data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
-data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
-data_empi_acc_f_phq$osex <- ordered(data_empi_acc_f_phq$sex_binarized)
-data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
-#get date into better format
-data_empi_acc_f_phq$EXAM_DATE <- as.Date(data_empi_acc_f_phq$BEGIN_EXAM_DTTM, format = "%m/%d/%y")
-data_empi_acc_f_phq$EXAM_DATE <- as.numeric(gsub(data_empi_acc_f_phq$EXAM_DATE, pattern = "-", replacement = ""))
-
-##########################################################
-#### Step 1- make F3* depressed and healthy groups* ######
-##########################################################
-  ### This is identical to up above, but repeated so I can run this chunk separately
-
-########
-### separate out depressed group
-########
-print("separate out depressed group")
-dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
-num_depression <- dim(dep_df)[1]
-num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
-print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
-print(paste0("N with ICD 9/10 codes for depression and unique EMPI in whole depression group = ", num_unique_EMPI_and_dep, " out of ", num_depression))
-
-#find number of participants who meet criteria for depression in the PHQ2/9 symptom rating
-print(paste0("N with ICD 9/10 codes for depression, PHQ2 >= 3, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.2 >= 3])))))
-print(paste0("N with ICD 9/10 codes for depression, PHQ9 >= 10, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.9 >= 10])))))
-
-#######
-### separate out healthy group
-#######
-print("separate out healthy group")
-healthy_df <- data_empi_acc_f_phq[-grep("F3", data_empi_acc_f_phq$ICD10),] #n = 14707
-num_healthy <- dim(healthy_df)[1] 
-num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
-print(paste0("N with WITHOUT depression ICD 9/10 codes (i.e. \"healthy\"): ", num_healthy))
-print(paste0("N with WITHOUT ICD 9/10 depression codes for depression and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", num_healthy))
-
-##########################################################
-#### Step 2- Identify PHQ2 >=3 in healthy group     ######
-##########################################################
-
-print("Get people in healthy group with phq2s for dep/healthy")
-# num in healthy group with PHQ2
-phq2_subset <- healthy_df[which(!is.na(healthy_df$PHQ.2)),] #n=3646, unique 698
-depressed_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 >= 3),] #n = 47 
-depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
-healthy_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 == 0),] #n = 3326
-healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631
-
-print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]"))
-
-##########################################################
-#### Step 3- Identify PHQ9 >=10 in healthy group    ######
-##########################################################
-
-print("Get people in healthy group with phq9s for dep/healthy")
-# num in healthy group with PHQ9
-phq9_subset <- healthy_df[which(!is.na(healthy_df$PHQ.9)),] #n=121
-depressed_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 >= 10),] #n = 32 
-depressed_unique_phq9_empis <- unique(depressed_phq9_subset$EMPI) #n=9
-healthy_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 == 0),] #n = 21
-healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n =4
-
-print(paste0("PHQ9 breakdown in healthy group (total n with score = ", dim(phq9_subset)[1], "[unique = ", length(unique(phq9_subset$EMPI)),"]): PHQ9 >= 10 :", dim(depressed_phq9_subset)[1], "[unique = ", length(depressed_unique_phq9_empis), "]; PHQ9 == 0 : ", dim(healthy_phq9_subset)[1], "[unique = ", length(unique(healthy_phq9_subset$EMPI)),"]"))
-
-##########################################################
-#Step 4 - Combine for depressed group and healthy groups #
-##########################################################
-
-print("Combine groups for new cohort")
-gain_in_dep_group <- length(depressed_unique_phq2_empis) + length(depressed_unique_phq9_empis)
-dep_num_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_dep + gain_in_dep_group # we can do this because we know they are each exclusive - num_unique_empi_and_dep from icd10 group, and no overlap in phq2 and 9 group
-print(paste0("number of UNIQUE people with an icd10 code for Depression OR scored positive on PHQ2(>=3) or PHQ9(>=10) : ", dep_num_icd10_plus_phq2_and_phq9, ", for a gain of n = ", gain_in_dep_group))
-
-healthy_num_no_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_healthy - gain_in_dep_group #take number of people who don't have depression diagnosis and don't have a depression phq2/9
-
-print(paste0("number of people with NO icd10 code for Depression and NEVER had a PHQ2(>=3) or PHQ9(>=10) : ", healthy_num_no_icd10_plus_phq2_and_phq9, ", for a loss of n = ", gain_in_dep_group))
-
-healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9 <- length(healthy_unique_phq2_empis) + length(healthy_unique_phq9_empis)
-print(paste0("number of UNIQUE people with NO icd10 code for Depression and HAS HAD AT LEAST 1 PHQ2 or 9 that was 0) : ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
-
-print(paste0("Cleanest cohort (depressed in icd9/10 or Phq2/9; healthy by at least one phq2/9 and never depression dx UNIQUE ---->>>>  Depressed: ", dep_num_icd10_plus_phq2_and_phq9, "; Healthy: ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))
-
-##### THESE ARE THE DATAFRAMES I WANT TO USE GOING FORWARD ###
-#all empis for depressed group include people with depresssion dx, and people who got in for phq2 or phq9
-empis_for_depressed_group <- append(
-  append(unique(dep_df$EMPI), depressed_unique_phq2_empis), depressed_unique_phq9_empis)
-
-final_depressed_group_withICD_AND_depressed_phq2_or_9 <- 
-  data_empi_acc_f_phq %>%
-filter(EMPI %in% empis_for_depressed_group)
-
-#------
-empis_for_healthy_group <- append(healthy_unique_phq2_empis, healthy_unique_phq9_empis)
-
-final_healthy_group_withNOICD_AND_phq2_or_9_0<- 
-  data_empi_acc_f_phq %>%
-filter(EMPI %in% empis_for_healthy_group)
-
-#------
-### add a column to the original dataframe with a 0 if not in any group, -1 if healthy, and 1 if depressed
-data_empi_acc_f_phq$depGroupVar = 0
-data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_depressed_group)] = 1
-data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_healthy_group)] = -1
-data_empi_acc_f_phq$depGroupVar = as.factor(data_empi_acc_f_phq$depGroupVar)
-
 #------ Group for Tim to check quality issues ----------
@@
 names(melissa_qc_data) <- c("EMPI", "EXAM_DATE", "rating")
-subset_with_qc <- merge(subset, melissa_qi_data, by = c("EMPI", "EXAM_DATE"))
+
+bad_flairs <- read.csv("/Users/eballer/BBL/msdepression/data/dac/missing_flair.csv", sep = ",", header = T) #n = 1561
+bad_flairs$EXAM_DATE <- gsub(pattern = "-", replacement = "", x = bad_flairs$date)
+
+subset_with_qc <- merge(subset, melissa_qc_data, by = c("EMPI", "EXAM_DATE"))
+subset_excluded <- merge(subset, bad_flairs, by = c("EMPI", "EXAM_DATE")) #n = 1973;  healthy = 457, uncategorized = 1284, depressed = 232
+
+unique <- subset_excluded %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()  #n (total) = 458, healthy = 114, uncat = 282. dep = 62
 
 write.csv(subset_with_qc, "/Users/eballer/BBL/msdepression/data/dac/subset_with_qc_for_tim_20211117.csv", col.names = T)
+write.csv(subset_with_qc, "/Users/eballer/BBL/msdepression/data/dac/subset_bad_flairs_for_tim_20211117.csv", col.names = T)
 
@@
 
+write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
+
 #fascicle_mapping, returns numerical mapping as well as names of tracts
@@
 
+```
+
+##########################################################
+#          Step 8 - Visualize all Tracts                 #
+##########################################################
+
+```{r visualize_tracts}
 ##########################################################
@@
 
+```
+
 ##########################################################
@@
 
-```
-
 ```{r dep_vs_healthy_regressions}
+##########################################################
+#                Step 9 - Regressions                    #
+##########################################################
 
@@
 {
-  lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
 })
@@
 ```
+
 ```{r age_by_depression_fascicles}

### 2021-12-01 msdepression_regressions_started_20211118.Rmd
figuring out why some people are broken

@@
    mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
-   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = ""))
+   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
+   mutate(EMPI = as.factor(EMPI))
 
@@
 
-#fdr corrected
+#fdr corrected - depression alone
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected - interaction
 fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
@@
 
+```
+
+```{r age_by_depression_unique}
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#lm
+fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
+{
+   lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_agedep_lm) <- fascicle_names
+
+#anova
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+
+#fdr corrected - depression alone
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected - interaction
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+```
+```{r depression_unique_all_fascicles_together}
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  ungroup()
+##
+
+dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+ggplot(data = dep_and_healthy_groups_for_ICD_analysis, aes(x=PAT_AGE_AT_EXAM,y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
+  geom_point(position = "jitter", show.legend = F) + 
+  geom_line(show.legend = F) + 
+  facet_wrap(~EMPI)
+
+#lm
+fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
+{
+   lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_agedep_lm) <- fascicle_names
+
+#anova
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+
+#fdr corrected - depression alone
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected - interaction
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+```
+
+```{r depression_mm_all_fascicles_together}
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  ungroup()
+##
+
+### thank you Bart! 11/30/2021
+ggplot(data = dep_and_healthy_groups_for_ICD_analysis, aes(x=PAT_AGE_AT_EXAM,y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
+  geom_point(position = "jitter", show.legend = F) + 
+  geom_line(show.legend = F) + 
+  facet_wrap(~depGroupVar)
+
+ggplot(data = dep_and_healthy_groups_for_ICD_analysis, aes(x=PAT_AGE_AT_EXAM,y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
+  geom_point(position = "jitter", show.legend = F) + 
+  geom_line(show.legend = F) + 
+  facet_wrap(~EMPI)
+
+dep_vs_healthy <- lmerTest::lmer(total_fascicle_prop_lost ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+dep_easy <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+ggplot(data = dep_and_healthy_groups_for_ICD_analysis2, aes(x=`(Intercept)`, y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
+  geom_point(position = "jitter", show.legend = F) + 
+  geom_line(show.legend = F) + 
+  facet_wrap(~EMPI)
+
+#lm
+fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
+{
+   lmerTest::lmer(substitute(i ~ depGroupVar + (1|EMPI) + PAT_AGE_AT_EXAM , list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_agedep_lm) <- fascicle_names
+
+#anova
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+
+#fdr corrected - depression alone
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected - interaction
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
 ```

### 2021-12-01 lab_notebook/lab_notebook.Rmd
adding

@@
 20211107
-[] need to change the date format in my regular dac pull
-[] need to divide the ROIs by the total volume 
+[x] need to divide the ROIs by the total volume
+
+20211109
+[x] need to change the date format in my regular dac pull
+ - pulled the results down to my local machine
+```{bash}
+scp eballer@transfer.pmacs.upenn.edu:/project/msdepression/results/fascicle_volumes_all_subjects_roi_n2336.csv .
+```
+ [x] age
+ [x] sex
+ [x] depression
+ [x] repeated measures
+ [x] exclusions flow chart
+   - compare # good phq2s/9s in the direction of depression
+ [x] meeting with matt taki setup
+
+ 20211117
+  making a list of my people with depressed/not depressed with melissa rating
+```{bash}
+#done locally on /Users/eballer/BBL/msdepression/data/dac
+more 1yr_report_address.csv | grep "mimosa" | perl -pe 's/.*sub-(.*)\/ses-(.*)\/run.*gz,.*,(.*),20.*/$1,$2,$3/' > just_empi_date_rating
+
+# to count how many people are unique in our gorup of people with qc, n = 1059
+ more 1yr_report_address.csv | grep "mimosa" | perl -pe 's/.*sub-(.*)\/ses-(.*)\/run.*gz,.*,(.*),20.*\/pro.*/$1,$2,$3/' | cut -f 1 -d ',' | sort -u | wc
+
+```
+
+ 20211118
+ - my dac summary thing is getting too meaty and I can't find anything anymore. Making a new rmd with analyses.
+ [x] msdepression_regressions_started_20211118.Rmd
+
+20211123
+[] make regression function
+
 [] need to get total # of lesions - this will be tough, need taki's code. Not to do right now - sydney to look into
+
+20211201
+[x] trying to figure out why some people completely bottom out with their lesions. Looks like some fail registration, though I'm not really sure why.
+[] make depression mask
+
+plan to calculate depression network overlap:
+
+1) take the depression network map and 3dresample it into the voxel space of any of the fascicle masks
+2) for each fascicle mask, multiply the mask by the resampled depression network.
+then count the nonzero voxels.
+The number of nonzero voxels is the volume overlap of that fascicle with the depression network
+3) plot the histogram of depression network overlap volumes. Hopefully there will be a notch in it where we can say anything beneath X mm^3 is not connected to the depression network
+
+20211202
+[] figure out why registration broke

### 2021-12-02 lab_notebook/lab_notebook.Rmd
group analyses wiwth depression network

@@
 
+on local machine, make binary mask
+```{bash}
+cd /Users/eballer/BBL/msdepression/templates/harvard_depression
+3dcalc -a Depression_Clust_mask_rois.nii.gz -expr 'ispositive(a)' -prefix Depression_Clust_mask_roi_binarized.nii.gz
+scp -r harvard_depression eballer@transfer.pmacs.upenn.edu:/project/msdepression/templates/.
+```
+
+On cluster
+```{bash}
+cd /project/msdepression/templates/harvard_depression
+3dresample -master AF_L.nii.gz -prefix "resampled_hcp_space_harvard_dep_mask.nii.gz" -input Depression_Clust_mask_roi_binarized.nii.gz
+```
+
+code for calc_vol_fascicles_within_depression_mask.sh 
+```{bash calc_vol_fascicles_within_depression_mask}
+#### calculate volume of fascicles within the depression network ####
+### Pre: All streamlines must have been made in dsi studio (these are the nii files)
+### Post: A depression mask resampled to template space, file <streamline_volume_within_dep_network>, containing each fascicle and its corresponding % within depression netw
+ork
+### Uses: Trying to figure out which fascicles are in the depression network for dimensionality reduction for use in later regressions (i.e. decrease # comparisons by only l
+ooking at fascicles within depression network)
+### Dependencies: This script requires afni to be installed (currently using 20.1)
+
+#set outfile
+outfile="/project/msdepression/results/streamline_volume_within_dep_network.csv"
+
+#set template directories, will be making our resampled mask from within this
+template="Depression_Clust_mask_roi_binarized.nii.gz"
+template_dir="/project/msdepression/templates/harvard_depression"
+resampled_filename="resampled_${template}"
+template_fullpath="/project/msdepression/templates/harvard_depression/${template}"
+
+#set fascicle directory
+fascicle_direc="/project/msdepression/templates/dti/HCP_YA1065_tractography/"
+fascicle_type="association"
+fascicle_for_resampling="AF_L.nii.gz"
+
+#set home directory
+homedir="/project/msdepression/scripts/"
+
+#initiate csvs
+rm -f $outfile
+touch $outfile
+echo "fascicle,num_voxels_total,non_zero_voxels_in_dep_map,prop_in_mask" >> $outfile
+
+#resample Depression_Clust_amsk_roi_binarized.nii.gz to match sample fascicle - AF_L.nii.gz
+if [ -f "${template_dir}/${resampled_filename}" ]; then
+	rm -f ${template_dir}/${resampled_filename}
+fi
+3dresample -master ${fascicle_direc}/${fascicle_type}/${fascicle_for_resampling} -prefix ${template_dir}/${resampled_filename} -input ${template_fullpath} -rmode NN
+
+#go through each fiber, get volume, and store it in csv
+fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"
+for fiber_bundle_type in ${fiber_bundle_type_list}; do
+    echo "Fiber bundle type is " $fiber_bundle_type
+    fascicle_volumes=$(ls ${fascicle_direc}/${fiber_bundle_type}/*.nii* | grep -v "dep_mask") #list all files, exclude depression ones
+    for fascicle in ${fascicle_volumes}; do
+	    fascicle_prefix=$(echo $fascicle | perl -pe 's/.*\/(.*).nii.gz/$1/g')
+	    num_voxels_whole_mask=$(3dmaskave -quiet -mask SELF -sum ${fascicle})
+
+	    #remove previous mask if already made
+	    if [ -f "${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz" ]; then
+		    rm -f ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz
+	    fi
+
+	    3dcalc -a ${template_dir}/${resampled_filename} -b ${fascicle} -expr 'a*b' -prefix ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz
+	    num_nonzero_volumes=$(3dmaskave -quiet -mask SELF -sum ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}xdep_mask.nii.gz)
+	    prop_in_mask=$(echo ${num_nonzero_volumes}/${num_voxels_whole_mask} | bc -l)
+	    echo "num_voxels" $num_voxels_whole_mask "num_nonzero" $num_nonzero "prop_in_mask" $prop_in_mask
+
+	    echo "${fascicle_prefix},${num_voxels_whole_mask},${num_nonzero_volumes},${prop_in_mask}" >> $outfile
+    done
+done
+```
+
+Moved back to home directory
+```{bash}
+#from local machine
+scp eballer@transfer.pmacs.upenn.edu:/project/msdepression/results/streamline_volume_within_dep_network.csv .
+
+#moved 
+```
+
+get volume in dep network per fascicle
 20211202

### 2021-12-02 msdepression_regressions_started_20211118.Rmd
group analyses wiwth depression network

@@
 
-#fascicle_mapping, returns numerical mapping as well as names of tracts
+########
+#fascicle volumes in depression network
+########
+fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), header = T, sep = ",")
+
+#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
+fascicle_volumes_dep_network <- fascicle_volumes_dep_network %>%
+  replace(is.na(.), 0) %>%
+  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))
+
+write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
+
+#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
+#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all.csv", row.names = F, quote = F, col.names = F)
+
+#fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)]
+#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)
+
+#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
 fascicle_bundle_mapping <- get_fascicle_bundle_mapping()
+
 ##########################################################
@@
 
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep 10% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
 #visreg
-sapply(fascicle_lm, visreg)
+#sapply(fascicle_lm, visreg)
 ```
@@
 fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep 10% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
 fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
@@
 print(fascicle_anova_fdr_association)
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep 10% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
 
@@
 
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep 10% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
 #uncorrected
@@
 
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep 10% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector ==2)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
 #uncorrected
@@
 ```
+
 ```{r depression_main_effect_mixed_model}
 ##### repeat above analysis but use only the first instance of each EMPI, sorted by date
-df_mm <-df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]
+df_mm <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]
 
@@
 fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network 10%
+fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
 fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
@@
 #isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+#also make columns for the mean across the full depression network, and the 10% cutoff depression network
+#remove rows where the total_fascicle_prop_lost is 0 - this was usually a failure of normalization
 dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
   mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_10_percent =
+           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+  filter(total_fascicle_prop_lost > 0) %>%
   ungroup()
-##
 
@@
 
+dep_vs_healthy_net <- lmerTest::lmer(total_fascicle_prop_lost_indepnet ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+dep_easy_dep_net <- lm(total_fascicle_prop_lost_indepnet ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+dep_vs_healthy_net_10_percent <- lmerTest::lmer(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+dep_easy_dep_net_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
 ggplot(data = dep_and_healthy_groups_for_ICD_analysis2, aes(x=`(Intercept)`, y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
@@
 
+#fdr corrected nondep, number of fascicles is total, 40, 5 correct - C_PHP_L, SLF1_L, UF_L, R, DRTT_L
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected dep, number of fascicles is total = 47, num correct =30
+#take home - areas of worsening disease over age much more likely to be in depression network
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
 #fdr corrected - interaction
+fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
+{
+   lmerTest::lmer(substitute(i ~ depGroupVar*PAT_AGE_AT_EXAM + (1|EMPI) , list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_agedep_lm) <- fascicle_names
+
+#anova
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+
 fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
@@
 fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep 10%
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected nondep, number of fascicles is total, 40, 4 correct - C_PHP_L, UF_L, R, DRTT_L
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected dep, number of fascicles is total = 47, num correct = 29
+#take home - areas of worsening disease over age much more likely to be in depression network
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
 print(fascicle_anova_fdr_association)

### 2021-12-04 msdepression_regressions_started_20211118.Rmd
looking at dep network alone

@@
 
-#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
+fascicle_names_nondep_network<- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 0)]
+
+fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
 #write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all.csv", row.names = F, quote = F, col.names = F)
 
-#fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)]
+fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)]
 #write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)
@@
 
+```{r testing_whether_ms_disease_more_likely_in_depression_mask}
+### tests whether there is higher proportion of vertices lost in depression network versus not
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_10_percent =
+           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowMeans(across(fascicle_names_nondep_network))) %>%
+  filter(total_fascicle_prop_lost > 0) %>%
+  #group_by(EMPI) %>% 
+  #arrange(EXAM_DATE) %>% 
+  #slice(1) %>% 
+  ungroup()
+
+volume_lost_dep_vs_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)
+
+volume_lost_dep10_vs_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)
+
+#unique
+### tests whether there is higher proportion of vertices lost in depression network versus not, ends up being n=298; healthy = 184, depressed = 114 
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_10_percent =
+           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowMeans(across(fascicle_names_nondep_network))) %>%
+  filter(total_fascicle_prop_lost > 0) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)
+
+volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)
+
+###### Depression network x depression group interaction - not quite there. 
+#I want to ask whether depressed MS patients are more likely to have more damage in depression networks than none.
+So I know MS preferentially affects depression networks.
+Maybe I answered this already
+depression_network_and_dx_subgroup <- data.frame(subset(dep_and_healthy_groups_for_ICD_analysis, select = c("EMPI", "EXAM_DATE", "depGroupVar","total_fascicle_prop_lost_nondep_network", "total_fascicle_prop_lost_indepnet","total_fascicle_prop_lost_indepnet_10_percent")))
+#melted_df <- melt(depression_network_and_dx_subgroup, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
+
+#loss doesn't seem to be different between those with depression and those without
+loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = depression_network_and_dx_subgroup) #non sig
+loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = depression_network_and_dx_subgroup) #non sig
+loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = depression_network_and_dx_subgroup) #non sig
+```
+
 ```{r visreg_models}

### 2021-12-04 lab_notebook/lab_notebook.Rmd
adding musings about overlap of dep network and ms network

@@
 get volume in dep network per fascicle
+
+20211203
+So something cool is happening here, but unexpected.
+The areas in the brain that have the most significant disease are much more likely to be in the depression network versus outside of it.
+p<1*10^-15.
+Doesn't matter whether or not you are depressed, they seem to share the same network.
+Also, areas that tend to worsen with age also are more likely to be in the same network.
+Will be interesting to figure out why some people don't get depressed, if people all have disease there.
+Also, why is there overlap?
+Interesting that inflammatory attack of WM happens to be in the depression network - is this why inflammation makes depression?
+Is there something structurally that happens along these white matter tracts?
+Weaker BBB?
+Different energetic characteristics?
+Is the same inflammation/depression network overlap seen in other inflammatory white matter diseases with a high burden of depression, or just MS?
+Will higher numbers allow the individual fascicle difference prediction to come up?
+Or is this really just a case where HYDRA will identify subtypes (which I imagine might be the case)
 20211202
-[] figure out why registration broke
+[] figure out why registration broke - not sure

### 2022-02-17 msdepression_regressions_started_20211118.Rmd
updated msdepression script

@@
 
+data_processed_as_of_20220209 <- read.csv(paste0(homedir, "/data/msdepression/data/erica_dac_pull/accession_numbers_tim_updtes_20220209.csv"), header = TRUE, stringsAsFactors = FALSE, sep = ",") #n=2738 new subjects added, 1124 were previously in Melissa's group, 2732 were already in our sample.
+Total n = 3862.
+Total accession # = 16,853
+
 ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
@@
 
+```{r depression_unique_all_fascicles_together_last_scan}
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice_tail(n = 1) %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  ungroup()
+##
+
+dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+ggplot(data = dep_and_healthy_groups_for_ICD_analysis, aes(x=PAT_AGE_AT_EXAM,y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
+  geom_point(position = "jitter", show.legend = F) + 
+  geom_line(show.legend = F) + 
+  facet_wrap(~EMPI)
+
+#lm
+fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
+{
+   lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_agedep_lm) <- fascicle_names
+
+#anova
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+
+#fdr corrected - depression alone
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected - interaction
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+```
+
 ```{r depression_mm_all_fascicles_together}
@@
 print(fascicle_anova_fdr_association)
+
+### Chi sq - compare # fdr corrected values in dep network to outside
+table_for_chisq <-data.frame(rbind(c(5,35), c(30,17)), row.names = c("nondep_net", "dep_net"))
+names(table_for_chisq) <- c("fdr05", "NS")
+dep_network_age_effects_chi <- chisq.test(table_for_chisq) #p = 3.374 x 10^-6
 
@@
 
-volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == -1]) #mean dep = 28%, mean non_dep = 25%, p = 0.12
+
+volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network) #mean dep_net = 35%, mean non_dep_net =  17%, p 2.2 x 10^-16
 
 volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
-dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network) #mean dep net = 31%, mean non_dep_net = 17%, p 6.6 * 10^-15
+
+#df for visreg
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "prop_lost", "network")
+df_for_visreg$network <- as.factor(df_for_visreg$network)
+
+plot <- ggplot(df_for_visreg, aes(x=network, y = prop_lost, col=network)) +
+  geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="red") + 
+  scale_x_discrete(labels=c("1" = "Dep Network", "2" = "Non Dep Network"))
 
@@
 loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = depression_network_and_dx_subgroup) #non sig
-loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = depression_network_and_dx_subgroup) #non sig
+loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = depression_network_and_dx_subgroup) #non sig, trend 0.09, mean dep = 19% loss, healthy 17% loss
 ```

### 2022-03-01 K23_prelim_data_started_20220225.Rmd
K23 prelim analyses

---
title: "MS Prelim Data K23"
author: "Erica Baller"
date: "2/25/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
```

## MS preliminary data

Ideas:

Overall:
[] demographics - tables made below
[] imaging parameters

Aim 1: lesion network mapping
[] some picture showing the # of people who are scanned and the # of people with depression. Maybe an adaptation of Syndey's picture somehow. 
[] mimosa picture
[] picture of filtering pipeline
[] comparison of some of the histograms for high volume loss similar to prev literature(CN II, slf), not so much in other areas
[] results of age analysis [maybe/maybe not, to prove the method works]?
[] results of disease in vs outside of depression network (corrected), lead with this?
[] results total proportion lost (single p)
[] results of dep vs controls across all regions,(uncorrected)
[] results of dep vs controls in depression network (uncorrected)

[] prelim network enrichment via spin-testing or just citation?
- do I want to validate here or somewhere else with neurosynth maps?

Aim 2: prospective
[] power analysis
[] CNB (Bart has great Figure 3)
[] Goassess

Aim 3
[] Hydra picture
[] hydra results from my paper
[] acquisition protocol (T1, t2, flair, dwi)

## R Markdown

The first chunk here is reading in appropriate data frames and preprocessing. Two big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not. Inclusion criteria below:

Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
  1) Scanned under MS protocol (42K)
  2) MS Diagnosis (17K)
  3) Seen by MS doctor (16K)

Inclusion criteria for Depressed Group
  1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder)
        - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition
  2) Screened positive for depression on PHQ2 or PHQ9 at any point
  3) Prescribed antidepressant medications per NAMI website
        - https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

Inclusion criteria for Healthy Group
  1) No F* diagnoses
  2) At least 1 PHQ2 or PHQ9 with a documented 0
  3) Free of psychiatric medications per NAMI website

Exclusions
  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

Imaging exclusion
  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.

Final N demographics 
  Depressed: 859 (232 unique)
  Super clean healthy: 604 (148 unique)

(unique, with good MIMoSA):

                         Healthy          Depressed             p      test
  n                                148            232                     
  Race (%)         Caucasian       108 ( 73.0)    173 ( 74.6)   0.821     
                   Non-caucasian    40 ( 27.0)     59 ( 25.4)             
  Sex (%)          Female          117 ( 79.1)    199 ( 85.8)   0.117     
                   Male             31 ( 20.9)     33 ( 14.2)             
  Age (mean (SD))                46.57 (12.66)  48.75 (11.75)   0.089     
  Depression (%)   1               148 (100.0)      0 (  0.0)  <0.001     
                   2                 0 (  0.0)    232 (100.0)             
  PHQ2 (mean (SD))                0.00 (0.00)    0.58 (1.53)   <0.001     
  PHQ9 (mean (SD))                0.00 (NA)     10.43 (9.02)       NA

Networks

  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1). It is a thresholded mask. To use it, I binarized it, and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle. Most fascicles were either entirely overlapping or entirely outside the depression mask. We consider the fascicles that had any intersection with the binary depression mask to be within the depression mask, or "indepnet." 47 out of 87 fascicles are in this network. We have an even tighter mask "indepnet_10percent", which indicates that at least 10% of a fascicle's volume intersected or overlapped with the depression mask. 15 are in this network. "in_nondep_net" refers to brain regios that completely avoided the depression network (i.e. had no shared volume of overlap). 40 are in this network

Association fascicles:
 [1] "AF_L"     "AF_R"     "C_FPH_L"  "C_FPH_R"  "C_FP_L"   "C_FP_R"   "C_PH_L"   "C_PHP_L"  "C_PHP_R"  "C_PH_R"  
[11] "C_R_L"    "C_R_R"    "EMC_L"    "EMC_R"    "FAT_L"    "FAT_R"    "IFOF_L"   "IFOF_R"   "ILF_L"    "ILF_R"   
[21] "MdLF_L"   "MdLF_R"   "PAT_L"    "PAT_R"    "SLF1_L"   "SLF1_R"   "SLF2_L"   "SLF2_R"   "SLF3_L"   "SLF3_R"  
[31] "UF_L"     "UF_R"     "VOF_L"    "VOF_R"

Depression network:
 [1] "AF_L"    "AF_R"    "C_PHP_R" "EMC_L"   "EMC_R"   "FAT_L"   "FAT_R"   "IFOF_L"  "IFOF_R"  "ILF_R"   "MdLF_L" 
[12] "MdLF_R"  "PAT_L"   "PAT_R"   "SLF2_L"  "SLF2_R"  "SLF3_L"  "SLF3_R"  "V"       "AR_R"    "CBT_L"   "CBT_R"  
[23] "CPT_F_L" "CPT_F_R" "CPT_O_L" "CPT_O_R" "CPT_P_L" "CPT_P_R" "CS_A_L"  "CS_A_R"  "CS_P_L"  "CS_P_R"  "CS_S_L" 
[34] "CS_S_R"  "CST_L"   "CST_R"   "DRTT_R"  "ML_L"    "RST_L"   "RST_R"   "TR_A_L"  "TR_A_R"  "TR_P_L"  "TR_P_R" 
[45] "TR_S_L"  "TR_S_R"  "CC"

Depression 10% network
 [1] "AF_L"    "AF_R"    "C_PHP_R" "FAT_L"   "FAT_R"   "MdLF_L"  "MdLF_R"  "SLF2_L"  "SLF2_R"  "SLF3_R"  "CBT_L"  
[12] "CBT_R"   "CPT_O_R" "CS_P_R"  "TR_P_R"

Notebook

  I start out with a proof of concept analyses, to demonstrate that our lesion filtering works. I create histograms looking at the proportion of volume lost in each fascicle across subjects. 
We recapitulate earlier work (will need to cite) and show that higher volume loss with our method reflects regions previously known to have higher disease burden.These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis.

Analyses
  There are two main types of analyses I performed, t.tests (for example, to compare inside versus outside depression network) and linear models (for example, looking at the relationship of depression to each individual fascicle). With respect to T-tests, I assume unequal variance for all. For linear models, I have looked at the first scan (considered Unique for unique empi), the last scan previously (not in this notebook), and mixed models with EMPI as a repeated measure (all at the end of the notebook). I like the unique one the best. The general pattern is to look at the total proportion of volume lost across fascicles, then within depression network, within depression 10% network, and within non depression network. Then, I look at each individual fascicle, as well as separately consider the association fascicles (per DSI studio, these are the ones that connect association cortices).

  I attempted to do FDR correction for all analyses where I looked at each of 87 fascicles separately, or bucketed by association network or within depression network. If i was not successful, I included uncorrected (P<0.05).

  A note about age. I have tried with and without age in these models. Age is so phenomenally strong as an effect (thank goodness) but it actually tends to suck up some of the depression variance so currently it is not in the depression model. Given the age of our participants (range from 20s to 80s), there are I imagine at least two big age effects going on 1) progressive disease and 2) natural (or unnatural aging). Gams would probably be better at looking at this relationship and could probably be its own thing.

  And this is only in WM. MS is also associated with GM atrophy. I can only imagine the IMCO paper that could emerge out of looking at the relationship between white matter loss and gray matter atrophy as a function of age. But certainly not for this K! Also sounds like lots of people are interested in this.

  I have also looked at sex differences in individual fascicles. Underwhelming right now.

Results
  Depression group: 232 with depression, 148 healthy. PHQ2
  Looking at each person's first scan:

    *Age effects:
      -Age mean 47 (range 20-83). 
      -Almost everything corrects (54 out of 87 fascicles, 25 in association network)
      -Take home - as expected, the older you are, the worse white matter fascicle invovlement you have

    *Sex effects:
      - In the final good MIMoSA group, there were 64 unique males and 316 unique females, 83% women
      - Individual fascicle results underwhelming, only 3 regions meet P<0.05 (unc), SCP, CPT_O_R, OR_R
      - no association or within dep network findings
      - Analyses to consider at a later date
        [] looking at overall fascicle prop lost (rather than inidiv region)
        [] dep * sex
        [] some sort of matching

    *Depression over all fascicles:
      - The exciting news of the day! With our new depression group (moving the meds people), we have findings
      - total proportion lost - p = 0.049, depressed (27%) more loss than healthy(24%)
      - total proportion lost within depression network - trend (0.059) between depressed (35%) and healthy(30%)
      - total proportion lost within depression 10% network - p =  0.049 (depressed (33%) have more loss than healthy (27%))
      - total proportion lost outside depression network- p = 0.037 (depressed (18%) more than healthy(16%))
      - take home - people with a lifetime history of depression have more disease than nondepressed individuals. This finding is not specific to the depression network, but is overall.

    *MS disease and depression network overlap:
      - Ms disease within depression network (35%) versus outside (17%), p < 2.2* 10^-16
      - MS disease within depression 10% network (31%) versus outside (17%), p < 2.2 * 10^-16
      - Linear Regressions (essentially matching t tests above): 
         +depressed v healthy within dep network in (p = 0.061); 
         +dep vs healthy within dep 10% (p = 0.049);
         +dep vs healthy within non dep network (p = 0.037)
      - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis. Suggests a final possible common pathway? Shared biology? This may also in part explain why rates of depression in MS are much higher than rates of depression in other autoimmune conditions. 
    *Depression by individual fascicle:
      - Nothing corrects at this time, unfortunately. 
      - P<0.05, uncorrected findings (* indicates association network):
          -AF_R	0.01168289	*		
          -FAT_R	0.01828607 *			
          -SLF2_R	0.01901378 *		
          -VOF_R	0.04741569 *			
          -CPT_F_R	0.02887163 			
          -CPT_O_L	0.02238344			
          -CPT_P_R	0.02053104			
          -CS_S_R	0.03783654			
          -DRTT_R	0.00695066			
          -ML_R	0.02511932	
    Fascicle loss x PHQ9:
     - future analysis?

Limitations/Future directions

  All diagnostic info is obtained from EMR fields, not notes, nor direct chart review. We do not know if patients are relapsing/remiting, nor do we know their active depression status. Through careful review, though patients sometimes have many phq2s listed, they are always identical.

  It would be great if we also had better dimensional scores. The PHQ2s are largely 0s, and the lack of spread makes interpretation really challenging. Though the PHQ9s definitely have better spread, there just aren't many of them. Doing really good cognitive and clinical phenotyping, or even getting a different or better set of measures into the MS clinic would be ideal (Aim 2).

  With respect to networks, we have used the "depression" one from Harvard because I was inspired by the lesion network mapping work. However, the mask is pretty broad, covering and strongly overlapping with the executive network. There is bias toward having regions that functionally connect with the L DLPFC b/c it is a TMS target. We are considering neurosynth as a possible place to look at different depression masks.

  Interestingly, a key topic at the ACTRIMS 2022 conference was psychiatric comorbidity. Though we are focusing on depression, it sounds like anxiety is just as bad if not worse or more highly comorbid with MS. Would be worth considering not only a DSM breakdown but dimensionally looking at internalizing disorders for the K. Also more in line with rdoc but would require changing aims a bit.

  Future analyses: age x dep, sex x dep

```{r read_in_df_and_subset}
homedir <- "/Users/eballer/BBL/msdepression/"

data <- read.csv(paste0(homedir, "/data/drugs_data/parsable_msdepression.csv"), sep = ",", header = TRUE) #n=16830, only has people with MS ICD10 code G35 and seen by MS provider

columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <- data %>% 
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(EMPI = as.factor(EMPI)) %>%
   mutate(On.Psych.Meds = ifelse(On.Psych.Meds == "True", 1, 0)) %>%
   mutate(On.Antidepressants = ifelse(On.Antidepressants == "True", 1, 0)) %>%
   mutate(Has.PHQ2 = ifelse(Has.PHQ2 == "True", 1, 0)) %>%
   mutate(Has.PHQ9 = ifelse(Has.PHQ9 == "True", 1, 0)) %>%
   mutate(Has.depdx = ifelse(grepl("F3", ICD10), 1, 0)) %>%
   mutate(PHQ.2_modsev_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 >= 3), 1, 0)) %>%
   mutate(PHQ.9_modsev_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 >= 10), 1, 0)) %>%
   mutate(PHQ.2_mild_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 > 0 & PHQ.2 < 3), 1, 0)) %>%
   mutate(PHQ.9_mild_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 > 0 & PHQ.9 < 10), 1, 0)) %>%
   mutate(PHQ.2_zero = ifelse((!is.na(PHQ.2) & PHQ.2 == 0), 1, 0)) %>%
   mutate(PHQ.9_zero = ifelse((!is.na(PHQ.9) & PHQ.9 == 0), 1, 0)) %>%
   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
   rowwise(ACCESSION_NUM) %>%
   mutate(depGroupVar =
            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
   ungroup()

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])

write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

########
#fascicle volumes in depression network
########
fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), header = T, sep = ",")

#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
  replace(is.na(.), 0) %>%
  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))

write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

fascicle_names_nondep_network<- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 0)]

fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask > 0)]

#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all.csv", row.names = F, quote = F, col.names = F)

fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)]
#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
fascicle_bundle_mapping <- get_fascicle_bundle_mapping()

##########################################################
#             Merge with data_empi              #
##########################################################
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962

##########################################################
#                 Histos                        #
##########################################################
##### Histograms for PHQ2/9 healthy and depressed

hist(df_demo_and_fascicles$PHQ.2[which(!is.na(df_demo_and_fascicles$PHQ.2))], main = paste0("PHQ-2 In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

hist(df_demo_and_fascicles$PHQ.9[which(!is.na(df_demo_and_fascicles$PHQ.9))], main = paste0("PHQ-9 In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.9)))), xlab = "PHQ-9", breaks = 8)

phq2_and_dep <- df_demo_and_fascicles %>% filter(depGroupVar == 2) %>% drop_na(PHQ.2) 
phq9_and_dep <- df_demo_and_fascicles %>% filter(depGroupVar == 2) %>% drop_na(PHQ.9) 
phq2_and_healthy <- df_demo_and_fascicles %>% filter(Has.depdx == 0) %>% drop_na(PHQ.2) 
phq9_and_healthy <- df_demo_and_fascicles %>% filter(Has.depdx == 0) %>% drop_na(PHQ.9)

hist(phq2_and_dep$PHQ.2, main = paste0("PHQ-2/Depressed In Good Mimosa Group : n = ", dim(phq2_and_dep)[1]), xlab = "PHQ-2", breaks = 8)

hist(phq9_and_dep$PHQ.9, main = paste0("PHQ-9/Depressed In Good Mimosa Group : n = ", dim(phq9_and_dep)[1]), xlab = "PHQ-9", breaks = 8)

hist(phq2_and_healthy$PHQ.2, main = paste0("PHQ-2/Healthy In Good Mimosa Group : n = ", dim(phq2_and_healthy)[1]), xlab = "PHQ-2", breaks = 8)
hist(phq9_and_healthy$PHQ.9, main = paste0("PHQ-9/Healthy In Good Mimosa Group : n = ", dim(phq9_and_healthy)[1]), xlab = "PHQ-9", breaks = 8)

```

## Depression versus healthy analysis

```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment

make_demographics_table_ms(data_frame = data_empi_acc_f_phq)

df_unique_empi <- data_empi_acc_f_phq %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

make_demographics_table_ms(data_frame = df_unique_empi)

df_just_healthy_and_depressed <- df_unique_empi %>% filter(depGroupVar != 0)

make_demographics_table_ms(data_frame = df_just_healthy_and_depressed)

#just people with good mimosa
df_good_mimosa <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
make_demographics_table_ms(data_frame = df_good_mimosa)

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)
```

```{r proof_of_concept}

#this section creates histograms looking at the proportion of volume lost in each fascicle across subjects
#we recapitulate earlier work and show that higher volume loss with our method reflects regions previously known to have higher volume loss
#These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis

#have to pull out just the pieces I'm interested in
just_dep_and_empi <- subset(data_empi_acc_f_phq, select = c("EMPI", "depGroupVar"))
added_dep_to_fascicles <- merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI"))
melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))

#exclude regions where there was no lesion at all
lesioned <- subset(melted_df, value > 0)

#only include people in healthy or depressed group
lesioned_dx <- subset(lesioned, depGroupVar != 0)

#separate out healthy and depressed
healthy <- subset(lesioned, depGroupVar == 1)
depressed <- subset(lesioned, depGroupVar == 2)

q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

print(q)
print(r)
print(r_color)
print(x)
print(y)
print(z)
```

```{r age_effects_fascicles_unique}
########### age effect whole group

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#### age histo ###
hist(dep_and_healthy_groups_for_ICD_analysis$PAT_AGE_AT_EXAM,
main = paste0("Age In Good Mimosa Group : n = ", dim(dep_and_healthy_groups_for_ICD_analysis)[1]), xlab = "Age", breaks = 10)

#lm
fascicle_age_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_age_lm) <- fascicle_names

#anova
fascicle_age_anova <- lapply(fascicle_age_lm, anova)

#fdr corrected
fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
print(fascicle_age_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#visreg
#sapply(fascicle_age_lm,visreg)
```

```{r sex_effects_fascicles_unique}

### really not interesting at all. 3 areas that are uncorrected: SCP, CPT_O_R, OR_R

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#### sex histo ###
hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_sex_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_sex_lm) <- fascicle_names

#anova
fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)

#fdr corrected
fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
print(fascicle_sex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_sex_anova_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
fascicle_sex_anova_p05_unc <- as.data.frame(fascicle_sex_anova_p[fascicle_sex_anova_p < 0.05])
print(fascicle_sex_anova_p05_unc)

```

```{r depression_unique_all_fascicles_together_in_dep_networks}

#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
summary(dep_vs_healthy) #p=0.494

#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049

#look at within depression network specifically - trend, p = 0.059
total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 36%, mean non_dep = 32%, p = 0.071

#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
#yes, p = 0.048
total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048

#look at outside depression network, yes p = 0.037
total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.037

### Age actually makes things worse
#dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
#summary(dep_vs_healthy_w_age)

```

```{r depressed_versus_healthy_unique_subjs_indiv_fascicle}
##### repeat above analysis but use only the first instance of each EMPI, sorted by date
df_unique_empi <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#lm
fascicle_lm_unique <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM, list(i = as.name(x))), data = df_unique_empi)
})
names(fascicle_lm_unique) <- fascicle_names

#anova
fascicle_anova_unique <- lapply(fascicle_lm_unique, anova)

#fdr corrected
fascicle_anova_unique_fdr <- fdr_anova_generic(fascicle_anova_unique, 1)
print(fascicle_anova_unique_fdr)#anova values < 0.05, uncorrected

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector ==2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_unique_p <- sapply(fascicle_anova_unique, function(v) v$"Pr(>F)"[1])
fascicle_anova_unique_p05_unc <- as.data.frame(fascicle_anova_unique_p[fascicle_anova_unique_p < 0.05])
print(fascicle_anova_unique_p05_unc)

```
```{r testing_whether_ms_disease_more_likely_in_depression_mask_unique}
### tests whether there is higher proportion of vertices lost in depression network versus not

dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ### when I filter and remove fascicles where there was no loss at all, the depression vs healthy findings go away
#  filter(total_fascicle_prop_lost > 0) %>%
  ungroup()

#test whether there is differentially more volume loss in depression versus nondepression network
volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 35%, mean non_dep_net =  17%, p 2.2 x 10^-16

volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 31%, mean non_dep_net = 17%, p 2.2 * 10^-16

#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network), c(rep(1,298), rep(2,298))))
names(df_for_visreg) <- c("EMPI", "prop_lost", "network")
df_for_visreg$network <- as.factor(df_for_visreg$network)

plot <- ggplot(df_for_visreg, aes(x=network, y = prop_lost, col=network)) +
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.05)) + 
  geom_boxplot(width=0.1) + theme_minimal() + 
  stat_summary(fun.y=median, geom="point", size=2, color="red") + 
  scale_x_discrete(labels=c("1" = "Dep Network", "2" = "Non Dep Network"))

print(plot)
###### Depression network x depression group interaction - It is there, depression always worse. 
#these lms largely match up with the t.tests above. Sig differences between depressed (more disease) within and outside of mask

loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #non sig, p = 0.061
loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.0492
loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.037
```

All the mixed models are below
```{r age_effects_indiv_fascicles_mm}

### This subscript compares proportion of fascicles lost across age for each fascicle of the 80+. Looks lonly at the people in the depressed versus healthy group. Importantly, I ran this looking at everyone, at the results are the same.
### Separately considers findings just within the association cortex

#isolate out only the depressed group and healthy group
df_mm <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
#n depressed = 859, healthy = 604, total n = 1463

#lm
fascicle_lm_mm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM + (1 | EMPI), list(i = as.name(x))), data = df_mm)
})
names(fascicle_lm_mm) <- fascicle_names

#anova
fascicle_anova_mm <- lapply(fascicle_lm_mm, anova)

#fdr corrected
fascicle_anova_mm_fdr <- fdr_anova_generic(fascicle_anova_mm, 1)
print(fascicle_anova_mm_fdr)#anova values < 0.05, uncorrected

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network 10%
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_mm_p <- sapply(fascicle_anova_mm, function(v) v$"Pr(>F)"[1])
fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
print(fascicle_anova_mm_p05_unc)

```

```{r sex_effects_fascicles_mm}
########### sex effect whole group

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% filter(depGroupVar != 0) 
#healthhy = 713, total n = 1106

#### sex histo ###
hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_sex_lm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ osex + (1|EMPI), list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_sex_lm) <- fascicle_names

#anova
fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)

#fdr corrected
fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
print(fascicle_sex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#visreg
#sapply(fascicle_sex_lm,visreg)
```

```{r depressed_versus_healthy_indiv_fascicles_mm}

### This subscript compares proportion of fascicles lost for each fascicle of the 80+ between those in the depressed and the clean healthy group. Done as both a mixed model as well as using unique individuals alone.

#Of note, adding age into this model makes things WORSE

### Separately considers findings just within the association cortex
#isolate out only the depressed group and healthy group
df_mm <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
#n depressed = 859, healthy = 604, total n = 1463

#lm
fascicle_lm_mm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ depGroupVar + (1 | EMPI), list(i = as.name(x))), data = df_mm)
})
names(fascicle_lm_mm) <- fascicle_names

#anova
fascicle_anova_mm <- lapply(fascicle_lm_mm, anova)

#fdr corrected
fascicle_anova_mm_fdr <- fdr_anova_generic(fascicle_anova_mm, 1)
print(fascicle_anova_mm_fdr)#anova values < 0.05, uncorrected

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network 10%
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_mm_p <- sapply(fascicle_anova_mm, function(v) v$"Pr(>F)"[1])
fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
print(fascicle_anova_mm_p05_unc)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-03-03 K23_prelim_data_started_20220225.Rmd
addt'l analyses

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 Overall:
+
 [] demographics - tables made below
+
 [] imaging parameters
@@
 The thrust of these mutate commands is to define who has depression and who does not.
+
+Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion.
+Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses.
+For these preliminary analysis, our inclusion/exclusion are more stringent.
+Of note, the graph below is irrespective of whether they had good/bad MIMoSA.
+As such, the MIMoSA subset is smaller.
+It does however highlight the
+
+<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>
+
 Inclusion criteria below:
@@
   A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1).
-It is a thresholded mask.
+It is an unthresholded mask (which I think is in Ts per their supplement).
 To use it, I binarized it, and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle.
@@
 had no shared volume of overlap).
-40 are in this network
+40 are in this network.
 
@@
 
-Notebook
+Notebook Overview
 
@@
 Results
-  Depression group: 232 with depression, 148 healthy. PHQ2
+  Depression group: 232 with depression, 148 healthy
   Looking at each person's first scan:
@@
       - no association or within dep network findings
+      - no age*sex effects
       - Analyses to consider at a later date
-        [] looking at overall fascicle prop lost (rather than inidiv region)
-        [] dep * sex
         [] some sort of matching
@@
       - total proportion lost - p = 0.049, depressed (27%) more loss than healthy(24%)
-      - total proportion lost within depression network - trend (0.059) between depressed (35%) and healthy(30%)
-      - total proportion lost within depression 10% network - p =  0.049 (depressed (33%) have more loss than healthy (27%))
-      - total proportion lost outside depression network- p = 0.037 (depressed (18%) more than healthy(16%))
+      - total proportion lost within depression network (47 total) - trend (0.059) between depressed (35%) and healthy(30%)
+      - total proportion lost within depression 10% network (15 fascicles total) - p = 0.049 (depressed (33%) have more loss than healthy (27%))
+      - total proportion lost within depression 5% network (24 fascicles total) -  p = 0.056 (depressed (32%) have more loss than healthy (27%))
+      - total proportion lost outside depression network (40 fascicles total)- p = 0.037 (depressed (18%) more than healthy(16%))
       - take home - people with a lifetime history of depression have more disease than nondepressed individuals. This finding is not specific to the depression network, but is overall.
@@
       - MS disease within depression 10% network (31%) versus outside (17%), p < 2.2 * 10^-16
+      - MS disease within depression 5% network (30%) versus outside (17%), p < 2.2 * 10^-16
       - Linear Regressions (essentially matching t tests above): 
@@
 
-fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)]
+fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
 #write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)
+
+fascicle_names_dep_network_5_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)] #n=24
 
@@
 hist(phq9_and_healthy$PHQ.9, main = paste0("PHQ-9/Healthy In Good Mimosa Group : n = ", dim(phq9_and_healthy)[1]), xlab = "PHQ-9", breaks = 8)
+
+######## Histograms for depression network
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20)
 
@@
 
-print(q)
-print(r)
+#print(q)
+#print(r)
 print(r_color)
-print(x)
-print(y)
-print(z)
+#print(x)
+#print(y)
+#print(z)
 ```
@@
            rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+   mutate(total_fascicle_prop_lost_indepnet_5_percent =
+           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
   mutate(total_fascicle_prop_lost_nondep_network =
@@
 total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048
+
+#look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
+#yes, 
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.056
 
@@
 fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
-fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
-print(fascicle_anova_fdr_association)
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
 
@@
 fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector ==2)]
-fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
-print(fascicle_anova_fdr_association)
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only dep 5% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)]
+fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
 
@@
            rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_5_percent =
+           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
   mutate(total_fascicle_prop_lost_nondep_network =
@@
 dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 31%, mean non_dep_net = 17%, p 2.2 * 10^-16
+
+volume_lost_dep5_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 17%, p 2.2 * 10^-16
 
@@
 loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.037
+```
+```{r depressionx}
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_10_percent =
+           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowMeans(across(fascicle_names_nondep_network))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy) #p=0.494
+
+#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049
+
+#look at within depression network specifically - trend, p = 0.059
+total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 36%, mean non_dep = 32%, p = 0.071
+
+#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
+#yes, p = 0.048
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048
+
+#look at outside depression network, yes p = 0.037
+total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.037
+
+### Age actually makes things worse
+#dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+#summary(dep_vs_healthy_w_age)
+
 ```
@@
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Lastly, this section will explore interactions (age * dep), (sex * dep)
+
+```{r age_by_depression_unique}
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>%
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#lm
+fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
+{
+   lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_agedep_lm) <- fascicle_names
+
+#anova
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+
+#fdr corrected - depression alone
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected - interaction
+fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
+print(fascicle_agedep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network 10%
+fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_agedep_anova, function(v) v$"Pr(>F)"[3])
+fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
+print(fascicle_anova_p05_unc)
+```
+
+```{r sex_by_depression_unique}
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>%
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#lm
+fascicle_sexdep_lm <- lapply(fascicle_names, function(x) 
+{
+   lm(substitute(i ~ osex*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+
+})
+names(fascicle_sexdep_lm) <- fascicle_names
+
+#anova
+fascicle_sexdep_anova <- lapply(fascicle_sexdep_lm, anova)
+
+#fdr corrected - depression alone
+fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 2)
+print(fascicle_sexdep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[2])
+fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
+print(fascicle_anova_p05_unc)
+
+#fdr corrected - interaction
+fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 3)
+print(fascicle_sexdep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network 10%
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[3])
+fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
+print(fascicle_anova_p05_unc)
+```

### 2022-03-03 K23_prelim_data_started_20220225.Rmd
updated with phq9 stuff, didn't work

@@
     Fascicle loss x PHQ9:
-     - future analysis?
+     - Nothing significant by individual fascicle, within association network, by fascicle within the depression network
+     - similarly, no significant findings looks at means of all fascicle loss, within dep network, 10% and 5% and non-depression network
 
@@
 
+```{r phq9_correlation_fascicles}
+
+####### NOTHING SIGNIFICANT!!!! ########3
+#isolate out only the depressed group and healthy group
+with_phq9 <- df_demo_and_fascicles %>%
+  filter(depGroupVar != 0) %>%
+  filter(!is.na(PHQ.9)) %>%
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_10_percent =
+           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_5_percent =
+           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowMeans(across(fascicle_names_nondep_network))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#lm
+fascicle_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
+})
+names(fascicle_lm) <- fascicle_names
+
+#anova
+fascicle_anova <- lapply(fascicle_lm, anova)
+
+#fdr corrected
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only dep 10% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
+fascicle_anova_p05_unc <- fascicle_anova_p[!is.na(fascicle_anova_p) & fascicle_anova_p < 0.05]
+print(fascicle_anova_p05_unc)
+
+######### overall fascicle proportion affected - nothing is significant
+overall_prop_affected <- lm(total_fascicle_prop_lost~PHQ.9, data=with_phq9)
+overall_prop_affected_depnet <- lm(total_fascicle_prop_lost_indepnet~PHQ.9, data=with_phq9)
+overall_prop_affected_depnet_10percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~PHQ.9, data=with_phq9)
+overall_prop_affected_depnet_5percent <- lm(total_fascicle_prop_lost_indepnet_5_percent~PHQ.9, data=with_phq9)
+overall_prop_affected_nondep <- lm(total_fascicle_prop_lost_nondep_network~PHQ.9, data=with_phq9)
+```
+
+```{r phq9_regressions_just_dep_group}
+##### NOTHING SIGNIFICANT ########
+##################################
+## Just within depression group ##
+##################################
+#isolate out only the depressed group and healthy group
+with_phq9 <- df_demo_and_fascicles %>%
+  filter(depGroupVar == 2) %>%
+  filter(!is.na(PHQ.9)) %>%
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_10_percent =
+           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_5_percent =
+           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowMeans(across(fascicle_names_nondep_network))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+#lm
+fascicle_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
+})
+names(fascicle_lm) <- fascicle_names
+
+#anova
+fascicle_anova <- lapply(fascicle_lm, anova)
+
+#fdr corrected
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only dep 10% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#uncorrected
+fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
+fascicle_anova_p05_unc <- fascicle_anova_p[!is.na(fascicle_anova_p) & fascicle_anova_p < 0.05]
+print(fascicle_anova_p05_unc)
+
+######### overall fascicle proportion affected - nothing is significant
+overall_prop_affected <- lm(total_fascicle_prop_lost~PHQ.9, data=with_phq9)
+overall_prop_affected_depnet <- lm(total_fascicle_prop_lost_indepnet~PHQ.9, data=with_phq9)
+overall_prop_affected_depnet_10percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~PHQ.9, data=with_phq9)
+overall_prop_affected_depnet_5percent <- lm(total_fascicle_prop_lost_indepnet_5_percent~PHQ.9, data=with_phq9)
+overall_prop_affected_nondep <- lm(total_fascicle_prop_lost_nondep_network~PHQ.9, data=with_phq9)
+
+#visreg
+#sapply(fascicle_lm, visreg)
+```
 All the mixed models are below

### 2022-03-15 K23_prelim_data_started_20220225.Rmd
graphs

@@
 library(reshape2)
+library(pwr)
+library(mgcv)
 #functions
@@
       - no association or within dep network findings
-      - no age*sex effects
+      - Sex effects interaction model (fascicle ~ depGroupVar*osex + PAT_AGE_AT_EXAM)
+        - Dep effects - same as above, 12 fascicles (p<0.05 unc)
+        - Sex effects - same as above, with the edition of CNVIII_R (p<0.05 unc)
+        - age*sex interactions - SLF1_L and CNVIII_R (p<0.05 unc)
       - Analyses to consider at a later date
@@
 Through careful review, though patients sometimes have many phq2s listed, they are always identical.
+It is also possible that people who are on antidepressants are prescribed them not for depression.
+I tried a sensitivity analysis and removed people on meds from heatlhy group but didn't include them in the depression group.
+Unique n in depression group went from 232->115 and all findings go away.
+
+  In order to do this, you need to change the mutate command:
+
+##------
+#goal is to keep people who were seen by
+
+#### add this line #####
+   mutate(dep_by_dx_phq_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq | true_healthy, 1, 0)) %>%
+   rowwise(ACCESSION_NUM) %>%
+
+#### remove this line or comment out
+    # mutate(depGroupVar = sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq))) %>%
+
+#### add this line
+   mutate(depGroupVar = sum(c(dep_by_dx_phq_healthy_phq0_no_psych_meds,dep_by_dx_phq))) %>%
+   ungroup()
+### -----
 
@@
 dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
-summary(dep_vs_healthy) #p=0.494
+summary(dep_vs_healthy) #p=0.494, cohen d = 0.2
+
+#power analysis to predict sample required for K, u = num df, v = denom df, f2=0.01 (cohen d = 0.02/2) 
+#prop in each fascicle predicting dep (87), number of predictors (age, sex, ef, ess?) = 4, u = 87-4 = 83
+pwr.f2.test(u=83, f2 = 0.02, sig.level = 0.5, power = 0.8)
 
@@
 total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049
+
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
+df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
+
+plot <- ggplot(df_for_visreg, aes(x=diagnosis, y = prop_lost, col=diagnosis)) +
+  geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))
+
+print(plot)
+
+#bar plot
+Diagnosis <- c("Depressed", "Non-Depressed")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
 
@@
 #summary(dep_vs_healthy_w_age)
+
+#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
 
@@
 
-```
+### plot AF_R
+
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
+df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
+
+plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
+  ggtitle("Arcuate Fasciculus R") + 
+ # geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))
+
+print(plot)
+
+### Plot OR_R	0.69024050
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$OR_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$OR_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
+df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
+
+plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
+  ggtitle("Optic Radiation R") + 
+ # geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) 
+print(plot)
+
+### grouped bar plot
+
+Diagnosis <- c(rep("Depressed" , 2) , rep("Non-Depressed" , 2))
+Fascicle <- rep(c("Arcuate Fasciculus" , "Superior Cerebellar Peduncle"), 2)
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))),
+          (sd(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))))
+
+data <- data.frame(Diagnosis,Fascicle,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
+    geom_bar(position="dodge", stat="identity") + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())
+
+print(barplot)
+
+```
+
 ```{r testing_whether_ms_disease_more_likely_in_depression_mask_unique}
@@
   stat_summary(fun.y=median, geom="point", size=2, color="red") + 
-  scale_x_discrete(labels=c("1" = "Dep Network", "2" = "Non Dep Network"))
+  scale_x_discrete(labels=c("1" = "Depression Network", "2" = "Non-Depression Network")) + 
+  ylab("Proportion of Fascicles Affected") + 
+  theme(axis.text =element_text(size=14), axis.title.x = element_blank())
 
@@
 loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.037
-```
-```{r depressionx}
-#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
-  filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
-  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_10_percent =
-           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
-  mutate(total_fascicle_prop_lost_nondep_network =
-           rowMeans(across(fascicle_names_nondep_network))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
-summary(dep_vs_healthy) #p=0.494
-
-#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
-total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049
-
-#look at within depression network specifically - trend, p = 0.059
-total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 36%, mean non_dep = 32%, p = 0.071
-
-#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
-#yes, p = 0.048
-total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048
-
-#look at outside depression network, yes p = 0.037
-total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.037
-
-### Age actually makes things worse
-#dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-#summary(dep_vs_healthy_w_age)
-
 ```
@@
 ```
+Interaction models below
+
+```{r sexdep_effects_fascicles_unique}
+
+### Trying to figure out if there are any sex * depression effects, covarying for age given that men tend to get worse over time. dep*sex uncorrected in SLF1_L and CNVIII_R p< 0.05)
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#### sex histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
+main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)
+
+#lm
+fascicle_sexdep_lm <- lapply(fascicle_names, function(x) 
+{
+  lm(substitute(i ~ depGroupVar*osex + PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sexdep_lm) <- fascicle_names
+
+#anova
+fascicle_sexdep_anova <- lapply(fascicle_sexdep_lm, anova)
+
+##### depGroupVar ####
+### 12 areas, nearly identical to reg model
+
+#fdr corrected
+fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 1)
+print(fascicle_sexdep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_sexdep_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[1])
+fascicle_sexdep_anova_p05_unc <- as.data.frame(fascicle_sexdep_anova_p[fascicle_sexdep_anova_p < 0.05])
+print(fascicle_sexdep_anova_p05_unc) ### CNVIII p < 0.05
+
+##### sex differences ####
+## same as above, just add in CNVIII_R
+
+#fdr corrected
+fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 2)
+print(fascicle_sexdep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_sexdep_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[2])
+fascicle_sexdep_anova_p05_unc <- as.data.frame(fascicle_sexdep_anova_p[fascicle_sexdep_anova_p < 0.05])
+print(fascicle_sexdep_anova_p05_unc) ### CNVIII p < 0.05
+
+#######interaction ####
+
+### SLF1_L and CNVIII_R
+
+#fdr corrected interaction
+fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 4)
+print(fascicle_sexdep_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,4)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,4)
+print(fascicle_anova_fdr_association)
+
+#uncorrected
+fascicle_sexdep_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[4])
+fascicle_sexdep_anova_p05_unc <- as.data.frame(fascicle_sexdep_anova_p[fascicle_sexdep_anova_p < 0.05])
+print(fascicle_sexdep_anova_p05_unc) #p uncorrected slf1_L p = 0.0488, CNVIII_R 0.01)
+
+```

### 2022-03-22 K23_prelim_data_started_20220225.Rmd
graphs

@@
 #sapply(fascicle_age_lm,visreg)
+visreg(fascicle_age_lm$AF_R,ylab = "Proportion of Affected Streamlines", xlab = "Age in Years",
+       line=list(col="red"),
+       points=list(col="black"))
+
 ```
@@
 
-plot <- ggplot(df_for_visreg, aes(x=diagnosis, y = prop_lost, col=diagnosis)) +
+plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
   geom_violin() + 
@@
 #bar plot
-Diagnosis <- c("Depressed", "Non-Depressed")
+Diagnosis <- c("MS w/Depression", "MS No Depression")
 value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
@@
   ylab("Proportion of Fascicles Affected") +
+  ylim(0,0.5) +
   theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Proportion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
 
@@
 Diagnosis <- c(rep("Depressed" , 2) , rep("Non-Depressed" , 2))
-Fascicle <- rep(c("Arcuate Fasciculus" , "Superior Cerebellar Peduncle"), 2)
+Fascicle <- rep(c("Depression Network (AF)" , "Non-depression network (SCP)"), 2)
 value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
@@
 print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
+    geom_bar(position="dodge", stat="identity") + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())
+
+print(barplot)
+
+barplot2 <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
+    geom_bar(position="dodge", stat="identity", colour = "black", width=0.9,size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) + 
+    scale_fill_manual(values = c("#cdcdcd", "#FFFFFF", "#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+
+    ylab("Proportion of Fascicles Affected") +
+    theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+          axis.title.y = element_text(size=20),
+          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+          panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty graph
+    scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot2)
 
@@
 print(plot)
+
+########## Bar plot
+Network <- c("Depression Network","Non-Depression Network")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])))
+
+data <- data.frame(Network,value,sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Network,fill=Network, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of Fascicles Affected") +
+  ylim(0,0.5) +
+  theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
 ###### Depression network x depression group interaction - It is there, depression always worse. 
@@
 
-```
+### grouped bar and line plot
+
+Diagnosis <- c(rep("Depressed" , 2) , rep("Non-Depressed" , 2))
+Sex <- rep(c("Male" , "Female"), 2)
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)))),
+         (sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)))),
+         (sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)))),
+         (sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)))))
+
+data <- data.frame(Diagnosis,Sex,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(fill=Sex, y=value, x=Diagnosis)) + 
+    geom_bar(position="dodge", stat="identity") + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of SLF Affected") +
+  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())
+
+print(barplot)
+
+lineplot <- ggplot(data=data, aes(fill=Sex, y=value, x=Diagnosis, group = Sex)) +
+
+    geom_line(aes(color=Sex, size=0.9)) +
+    geom_point(aes(color=Sex, size=0.9)) +
+    ylim(0.05,0.30)+
+    ylab("Proportion of SLF Affected") +
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1) + 
+    theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank())
+
+print(lineplot)
+
+```

### 2022-06-23 msdepression/scripts/Descriptives_potential_paper_1_20220622.Rmd
afni command to make yeo networks

---
title: "Descriptives for potential paper 1 "
author: "Erica Baller"
date: "Started 2022 06 22"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
```

## Descriptives
[x] Demographics of full group unique
[x] Fascicles most likely to be hit
[] Networks most likely to be affected (map Yeo 7 networks to corresponding WM tracts)
[x] Age regressions (unique)
  [x] overall 
  [] networks
[x] Age regressions (mixed model)
[] Sex differences
  [x] overall
  [] networks
[] agexsex unique
  [x] overall
  [] networks
[] agexsex mm
  [x] overall - YAY! Now we have CORRECTED FINDINGS: age corrected (55 fascicles), 1 sex diff (Fornix_R) and 1 agexsex diff (Fornix_R)
  [] networks

## Descriptives

Considering doing a paper that just describes the cohort, and the method for lesion fascicle mapping. Would describe the sample, the method, fascicles most likely to be hit, what networks they are a part of. Could also look at age regressions (considering each person once, and then a mixed model that looks at trajectory). Lastly, sex differences (first episode + mixed model). Will also check age x sex interactions.

The first chunk here is reading in appropriate data frames and preprocessing. Two big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

Inclusion criteria below (not separating out depressed from non, but could do sensitivity analysis I guess):

Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
  1) Scanned under MS protocol (42K)
  2) MS Diagnosis (17K)
  3) Seen by MS doctor (16K)

Exclusions
  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.

Imaging exclusion
  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.

Final N demographics

Networks

  Will use yeo 7 networks.For each network, I will binarized it, and then used it as an ROI and calculate, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the network mask to the whole volume of the fascicle. If overlap, considered in yeo network x, if not, not in it.

Association fascicles:
 [1] "AF_L"     "AF_R"     "C_FPH_L"  "C_FPH_R"  "C_FP_L"   "C_FP_R"   "C_PH_L"   "C_PHP_L"  "C_PHP_R"  "C_PH_R"  
[11] "C_R_L"    "C_R_R"    "EMC_L"    "EMC_R"    "FAT_L"    "FAT_R"    "IFOF_L"   "IFOF_R"   "ILF_L"    "ILF_R"   
[21] "MdLF_L"   "MdLF_R"   "PAT_L"    "PAT_R"    "SLF1_L"   "SLF1_R"   "SLF2_L"   "SLF2_R"   "SLF3_L"   "SLF3_R"  
[31] "UF_L"     "UF_R"     "VOF_L"    "VOF_R"

Notebook Overview

  I start out with a proof of concept analyses, to demonstrate that our lesion filtering works. I create histograms looking at the proportion of volume lost in each fascicle across subjects. We recapitulate earlier work (will need to cite) and show that higher volume loss with our method reflects regions previously known to have higher disease burden.These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis.

Analyses

  There are two main types of analyses I performed, t.tests (for example, to compare inside versus outside depression network) and linear models (for example, looking at the relationship of depression to each individual fascicle). With respect to T-tests, I assume unequal variance for all. For linear models, I have looked at the first scan (considered Unique for unique empi), the last scan previously (not in this notebook), and mixed models with EMPI as a repeated measure (all at the end of the notebook). I like the unique one the best. The general pattern is to look at the total proportion of volume lost across fascicles, then within depression network, within depression 10% network, and within non depression network. Then, I look at each individual fascicle, as well as separately consider the association fascicles (per DSI studio, these are the ones that connect association cortices).

  I attempted to do FDR correction for all analyses where I looked at each of 87 fascicles separately, or bucketed by association network. If i was not successful, I included uncorrected (P<0.05).

  A note about age. I have tried with and without age in these models. Age is so phenomenally strong as an effect. Given the age of our participants (range from 20s to 80s), there are I imagine at least two big age effects going on 1) progressive disease and 2) natural (or unnatural aging). Gams would probably be better at looking at this relationship and could probably be its own thing.

  I have also looked at sex differences in individual fascicles as well as agexsex interactions. All are done using 1st scan alone as well as with mixed model using lmer.

Results
  *Demographics
    total n unique subjects = 783 (F = 609, M = 174) 
    total scans = 2962 (F = 2270, M = 692)

  *Age effects (unique):
      -Age mean 47 (range 20-83). 
      -Almost everything corrects (59 out of 87 fascicles, 26 in association network)
      -Take home - as expected, the older you are, the worse white matter fascicle invovlement you have

  *Age effects (gam unique)
      - 62 out of 87 fascicles correct, 26 in association network

  *Age effects (mixed model)
      - 57 out of 87 correct, 27 in association network

  *Sex effects (unique):
      - In the final good MIMoSA group, there were 609 unique females and 174 unique males, 77% women
      - Individual fascicle results underwhelming, only 3 regions meet P<0.05 (unc), 
          SCP	0.020859515			
          CPT_F_R	0.040590173			
          CPT_O_R	0.049616343			
          CPT_P_R	0.032079145			
          DRTT_L	0.047328252			
          DRTT_R	0.008453751			
          ML_R	0.033199761			
          OR_R	0.039151847

      - Analyses to consider at a later date
        [] some sort of matching

  *Sex effects (mixed model):
      - no action
      - Individual fascicle results w/ 12 that meet P<0.05 (uncorrected)
          ICP_L	0.016204905			
          MCP	0.002505451			
          SCP	0.005273532			
          V	0.018452360			
          CNVIII_L	0.005692622			
          CPT_F_R	0.047149209			
          CPT_O_R	0.031170467			
          CPT_P_R	0.025315969			
          CST_R	0.039553847			
          DRTT_L	0.043957171

  * Age*sex (unique)
      - no action

  * Age * sex (mixed model) 
  lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM*osex + (1|EMPI), list(i = as.name(x))), data = df_mm)
      - Action!
      - Age term: 55/87 correct
      - Sex Term: 1 (Fornix_R) corrects
      - Age*Sex intx term: 1 (fornix_R) corrects
      - Age *sex uncorrected (5):
          UF_L	0.0114639855			
          CNVIII_R	0.0163558293			
          CNVII_R	0.0175275993			
          CBT_L	0.0260870383			
          F_R	0.0002210451

Limitations/Future directions

```{r read_in_df_and_subset}
homedir <- "/Users/eballer/BBL/msdepression/"

data <- read.csv(paste0(homedir, "/data/drugs_data/parsable_msdepression.csv"), sep = ",", header = TRUE) #n=16830, only has people with MS ICD10 code G35 and seen by MS provider

columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

 data_empi_acc_f_phq <- data %>% 
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(EMPI = as.factor(EMPI)) %>%
   mutate(On.Psych.Meds = ifelse(On.Psych.Meds == "True", 1, 0)) %>%
   mutate(On.Antidepressants = ifelse(On.Antidepressants == "True", 1, 0)) %>%
   mutate(Has.PHQ2 = ifelse(Has.PHQ2 == "True", 1, 0)) %>%
   mutate(Has.PHQ9 = ifelse(Has.PHQ9 == "True", 1, 0)) %>%
   mutate(Has.depdx = ifelse(grepl("F3", ICD10), 1, 0)) %>%
   mutate(PHQ.2_modsev_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 >= 3), 1, 0)) %>%
   mutate(PHQ.9_modsev_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 >= 10), 1, 0)) %>%
   mutate(PHQ.2_mild_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 > 0 & PHQ.2 < 3), 1, 0)) %>%
   mutate(PHQ.9_mild_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 > 0 & PHQ.9 < 10), 1, 0)) %>%
   mutate(PHQ.2_zero = ifelse((!is.na(PHQ.2) & PHQ.2 == 0), 1, 0)) %>%
   mutate(PHQ.9_zero = ifelse((!is.na(PHQ.9) & PHQ.9 == 0), 1, 0)) %>%
   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
   rowwise(ACCESSION_NUM) %>%
   mutate(depGroupVar =
            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
   ungroup()

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])

write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

##########################################################
#             Merge with data_empi              #
##########################################################
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962

##########################################################
#             Histograms                        #
##########################################################

#all subjects/all scans, n  = 2962
hist(df_demo_and_fascicles$PAT_AGE_AT_EXAM, main = "Age at Exam, all subjects, multiple scans", xlab = "Age")
hist(df_demo_and_fascicles$sex_binarized, main = "Sex binarized, all subjects, multiple scans", xlab = "Sex", breaks = 2)

#unique, first scan, n = 783
df_unique_empi <- df_demo_and_fascicles %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

hist(df_unique_empi$PAT_AGE_AT_EXAM, main = "Age at First Exam", xlab = "Age")
hist(df_unique_empi$sex_binarized, main = "Sex binarized at First Exam", xlab = "Sex", breaks = 2)

```

```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment

make_demographics_table_all_subjs(data_frame = df_demo_and_fascicles, binarized_race = 1) #simple race 1 = just binarized, otherwise it will give you the full distribution of race
make_demographics_table_all_subjs(data_frame = df_unique_empi, binarized_race = 1)

```

```{r proof_of_concept}

#this section creates histograms looking at the proportion of volume lost in each fascicle across subjects
#we recapitulate earlier work and show that higher volume loss with our method reflects regions previously known to have higher volume loss
#These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis

#have to pull out just the pieces I'm interested in; this is for ALL subjects
just_empi <- subset(df_demo_and_fascicles, select = c("EMPI"))
added_to_fascicles <- merge(just_empi, fascicle_proportions, by = c("EMPI"))
melted_df <- melt(added_to_fascicles, id.vars = c("EMPI", "EXAM_DATE"))

#exclude regions where there was no lesion at all
lesioned <- subset(melted_df, value > 0)

histogram_plot_most_affected_fascicles<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

print(histogram_plot_most_affected_fascicles)

#have to pull out just the pieces I'm interested in; this is for unique subjects
just_empi <- subset(df_unique_empi, select = c("EMPI"))
added_to_fascicles <- merge(just_empi, fascicle_proportions, by = c("EMPI"))
melted_df <- melt(added_to_fascicles, id.vars = c("EMPI", "EXAM_DATE"))

#exclude regions where there was no lesion at all
lesioned <- subset(melted_df, value > 0)

histogram_plot_most_affected_fascicles_unique<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

print(histogram_plot_most_affected_fascicles_unique)

```

```{r age_effects_fascicles_unique}
########### age effect whole group

#### age histo ###
hist(df_unique_empi$PAT_AGE_AT_EXAM,
main = paste0("Age In Good Mimosa Group : n = ", dim(df_unique_empi)[1]), xlab = "Age", breaks = 10)

#lm
fascicle_age_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = df_unique_empi)
})
names(fascicle_age_lm) <- fascicle_names

#anova
fascicle_age_anova <- lapply(fascicle_age_lm, anova)

#fdr corrected
fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
print(fascicle_age_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#visreg
#sapply(fascicle_age_lm,visreg)
#visreg(fascicle_age_lm$AF_R,ylab = "Proportion of Affected Streamlines", xlab = "Age in Years",
#       line=list(col="red"),
 #      points=list(col="black"))

```
```{r age_effects_fascicles_unique_gam}
########### age effect whole group

#### age histo ###
hist(df_unique_empi$PAT_AGE_AT_EXAM,
main = paste0("Age In Good Mimosa Group : n = ", dim(df_unique_empi)[1]), xlab = "Age", breaks = 10)

#lm
fascicle_age_gam <- lapply(fascicle_names, function(x) 
{
  gam(substitute(i ~ s(PAT_AGE_AT_EXAM), list(i = as.name(x))), data = df_unique_empi)
})
names(fascicle_age_gam) <- fascicle_names

#anova
fascicle_age_anova <- lapply(fascicle_age_gam, anova)

fascicle_age_anova_smooth_p <- sapply(fascicle_age_anova, function(x)
{
  x$s.pv
})

#fdr corrected
fascicle_age_anova_fdr <- fdr_anova_generic_p_already_extracted(fascicle_age_anova_smooth_p)
print(fascicle_age_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_age_anova_smooth_p[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic_p_already_extracted(fascicle_anova_w_fiber_mapping)
print(fascicle_anova_fdr_association)

```

```{r sex_effects_fascicles_unique}

#### no sig differences, but uncorrected Men usually worse disease than women
#### sex histo ###
hist(df_unique_empi$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(df_unique_empi$sex_binarized == 1), "/", sum(df_unique_empi$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_sex_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ osex, list(i = as.name(x))), data = df_unique_empi)
})
names(fascicle_sex_lm) <- fascicle_names

#anova
fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)

#fdr corrected
fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
print(fascicle_sex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_sex_anova_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
fascicle_sex_anova_p05_unc <- as.data.frame(fascicle_sex_anova_p[fascicle_sex_anova_p < 0.05])
print(fascicle_sex_anova_p05_unc)

#sapply(fascicle_sex_lm,visreg)
#visreg(fascicle_age_lm$AF_R,ylab = "Proportion of Affected Streamlines", xlab = "Age in Years",
       #line=list(col="red"),
       #points=list(col="black"))
```

All the mixed models are below
```{r age_effects_indiv_fascicles_mm}

### This subscript compares proportion of fascicles lost across age for each fascicle of the 80+. Looks lonly at the people in the depressed versus healthy group. Importantly, I ran this looking at everyone, at the results are the same.
### Separately considers findings just within the association cortex

#isolate out only the depressed group and healthy group
df_mm <- df_demo_and_fascicles
#n depressed = 859, healthy = 604, total n = 1463

#lm
fascicle_lm_mm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM + (1 | EMPI), list(i = as.name(x))), data = df_mm)
})
names(fascicle_lm_mm) <- fascicle_names

#anova
fascicle_anova_mm <- lapply(fascicle_lm_mm, anova)

#fdr corrected
fascicle_anova_mm_fdr <- fdr_anova_generic(fascicle_anova_mm, 1)
print(fascicle_anova_mm_fdr)#anova values < 0.05, uncorrected

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_mm_p <- sapply(fascicle_anova_mm, function(v) v$"Pr(>F)"[1])
fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
print(fascicle_anova_mm_p05_unc)

```

```{r sex_effects_fascicles_mm}
########### sex effect whole group

#isolate out only the depressed group and healthy group
df_mm <- df_demo_and_fascicles 
#healthhy = 713, total n = 1106

#### sex histo ###
hist(df_mm$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(df_mm$sex_binarized == 1), "/", sum(df_mm$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_sex_lm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ osex + (1|EMPI), list(i = as.name(x))), data = df_mm)
})
names(fascicle_sex_lm) <- fascicle_names

#anova
fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)

#fdr corrected
fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
print(fascicle_sex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_sex_anova_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
fascicle_sex_anova_p05_unc <- as.data.frame(fascicle_sex_anova_p[fascicle_sex_anova_p < 0.05])
print(fascicle_sex_anova_p05_unc)
#visreg
#sapply(fascicle_sex_lm,visreg)
```

Interaction models below

```{r agesex_effects_fascicles_unique}

### Trying to figure out if there are any sex * depression effects, covarying for age given that men tend to get worse over time. dep*sex uncorrected in SLF1_L and CNVIII_R p< 0.05)

hist(df_unique_empi$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(df_unique_empi$sex_binarized == 1), "/", sum(df_unique_empi$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_agesex_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PAT_AGE_AT_EXAM*osex, list(i = as.name(x))), data = df_unique_empi)
})
names(fascicle_agesex_lm) <- fascicle_names

#anova
fascicle_agesex_anova <- lapply(fascicle_agesex_lm, anova)

### age term

#fdr corrected
fascicle_agesex_anova_fdr <- fdr_anova_generic(fascicle_agesex_anova, 1)
print(fascicle_agesex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_agesex_anova_p <- sapply(fascicle_agesex_anova, function(v) v$"Pr(>F)"[1])
fascicle_agesex_anova_p05_unc <- as.data.frame(fascicle_agesex_anova_p[fascicle_agesex_anova_p < 0.05])
print(fascicle_agesex_anova_p05_unc) ### CNVIII p < 0.05

##### sex term ####

#fdr corrected
fascicle_agesex_anova_fdr <- fdr_anova_generic(fascicle_agesex_anova, 2)
print(fascicle_agesex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_agesex_anova_p <- sapply(fascicle_agesex_anova, function(v) v$"Pr(>F)"[2])
fascicle_agesex_anova_p05_unc <- as.data.frame(fascicle_agesex_anova_p[fascicle_agesex_anova_p < 0.05])
print(fascicle_agesex_anova_p05_unc) ### CNVIII p < 0.05

#######interaction age*sex term ####

#fdr corrected interaction
fascicle_agesex_anova_fdr <- fdr_anova_generic(fascicle_agesex_anova, 3)
print(fascicle_agesex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_agesex_anova_p <- sapply(fascicle_agesex_anova, function(v) v$"Pr(>F)"[3])
fascicle_agesex_anova_p05_unc <- as.data.frame(fascicle_agesex_anova_p[fascicle_agesex_anova_p < 0.05])
print(fascicle_agesex_anova_p05_unc)

```

```{r agesex_effects_fascicles_mm}

### Trying to figure out if there are any age x sex interactions w/ a mixed model. Yes!!

### age restults still robust, but now 1 sex diff (Fornix_R) & one agexsex diff (Fornix_R). Still 55 areas that age correct

df_mm <- df_demo_and_fascicles

hist(df_unique_empi$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(df_unique_empi$sex_binarized == 1), "/", sum(df_unique_empi$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_agesex_lm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM*osex + (1|EMPI), list(i = as.name(x))), data = df_mm)
})
names(fascicle_agesex_lm) <- fascicle_names

#anova
fascicle_agesex_anova <- lapply(fascicle_agesex_lm, anova)

### age term

#fdr corrected
fascicle_agesex_anova_fdr <- fdr_anova_generic(fascicle_agesex_anova, 1)
print(fascicle_agesex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_agesex_anova_p <- sapply(fascicle_agesex_anova, function(v) v$"Pr(>F)"[1])
fascicle_agesex_anova_p05_unc <- as.data.frame(fascicle_agesex_anova_p[fascicle_agesex_anova_p < 0.05])
print(fascicle_agesex_anova_p05_unc) ### CNVIII p < 0.05

##### sex term ####

#fdr corrected
fascicle_agesex_anova_fdr <- fdr_anova_generic(fascicle_agesex_anova, 2)
print(fascicle_agesex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_agesex_anova_p <- sapply(fascicle_agesex_anova, function(v) v$"Pr(>F)"[2])
fascicle_agesex_anova_p05_unc <- as.data.frame(fascicle_agesex_anova_p[fascicle_agesex_anova_p < 0.05])
print(fascicle_agesex_anova_p05_unc) ###

#######interaction age*sex term ####

#fdr corrected interaction
fascicle_agesex_anova_fdr <- fdr_anova_generic(fascicle_agesex_anova, 3)
print(fascicle_agesex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agesex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_agesex_anova_p <- sapply(fascicle_agesex_anova, function(v) v$"Pr(>F)"[3])
fascicle_agesex_anova_p05_unc <- as.data.frame(fascicle_agesex_anova_p[fascicle_agesex_anova_p < 0.05])
print(fascicle_agesex_anova_p05_unc) #p uncorrected slf1_L p = 0.0488, CNVIII_R 0.01)

```

### 2022-06-23 msdepression/scripts/Group_level_analyses/msdepression_regressions_started_20211118.Rmd
afni command to make yeo networks

--- 2022-02-17 msdepression_regressions_started_20211118.Rmd
@@
    lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
-# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
 })
@@
 
-dep_vs_healthy <- lmerTest::lmer(total_fascicle_prop_lost ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-
-dep_easy <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-
-dep_vs_healthy_net <- lmerTest::lmer(total_fascicle_prop_lost_indepnet ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-
-dep_easy_dep_net <- lm(total_fascicle_prop_lost_indepnet ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-
-dep_vs_healthy_net_10_percent <- lmerTest::lmer(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-
-dep_easy_dep_net_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+dep_vs_healthy <- lmerTest::lmer(total_fascicle_prop_lost ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis) #trend, p = 0.15
+
+#dep_easy <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+dep_vs_healthy_net <- lmerTest::lmer(total_fascicle_prop_lost_indepnet ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis) #trend, p = 0.14
+
+#dep_easy_dep_net <- lm(total_fascicle_prop_lost_indepnet ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
+
+dep_vs_healthy_net_10_percent <- lmerTest::lmer(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.23
+
+#dep_easy_dep_net_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
 
@@
 #anova
-fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)
+fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova) #p unc= C_PH_L, p = 0.03, PAT_R p=0.04, slf1_R p = 0.09, VOF_R p = 0.03
 

### 2022-06-23 msdepression/scripts/K23_prelim_data_started_20220225.Rmd
afni command to make yeo networks

--- 2022-03-22 K23_prelim_data_started_20220225.Rmd
@@
 Diagnosis <- c(rep("Depressed" , 2) , rep("Non-Depressed" , 2))
-Fascicle <- rep(c("Depression Network (AF)" , "Non-depression network (SCP)"), 2)
+Fascicle <- rep(c("Depression Network (AF)" , "Region Outside Depression Network (C_PH)"), 2)
 value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
            mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+           mean(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
 
 sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
            (sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))),
-          (sd(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))))
+          (sd(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))))
 
@@
     geom_bar(position="dodge", stat="identity", colour = "black", width=0.9,size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1) + 
     scale_fill_manual(values = c("#cdcdcd", "#FFFFFF", "#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
@@
   theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+# Black and white
+barplot <- ggplot(data, aes(x=Network,fill=Network, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Proportion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
 

### 2022-06-23 msdepression/scripts/dac_summary.Rmd
afni command to make yeo networks

--- 2021-02-08 dac_summary.Rmd
@@
 #obtained from: unique(data_empi_acc_f_phq$Provider[grep("PRUIT", data_empi_acc_f_phq$Provider)])
-ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, LINDA", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
+ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")
 
@@
 
-data_e <- data4_20210119[which(data_empi_acc_f_phq$Provider %in% ms_providers),]
+data_e <- data4_20210119[which(data4_20210119$Provider %in% ms_providers),]
+
 dep_df <- grep("F3", data_e$ICD10)

### 2022-06-23 msdepression/scripts/final_dac_summary_pull_20210119.Rmd
afni command to make yeo networks

--- 2021-11-17 final_dac_summary_pull_20210119.Rmd
@@
 healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n =4
+healthyish_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 <= 4),] #n = 55
+healthyish_unique_phq9_empis <- unique(healthyish_phq9_subset$EMPI) #n =9
 
@@
 names(melissa_qc_data) <- c("EMPI", "EXAM_DATE", "rating")
+
+bad_flairs <- read.csv("/Users/eballer/BBL/msdepression/data/dac/missing_flair.csv", sep = ",", header = T) #n = 1561
+bad_flairs$EXAM_DATE <- gsub(pattern = "-", replacement = "", x = bad_flairs$date)
+
 subset_with_qc <- merge(subset, melissa_qi_data, by = c("EMPI", "EXAM_DATE"))
+subset_excluded <- merge(subset, bad_flairs, by = c("EMPI", "EXAM_DATE")) #n = 1973;  healthy = 457, uncategorized = 1284, depressed = 232
+
+unique <- subset_excluded %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()  #n (total) = 458, healthy = 114, uncat = 282. dep = 62
 
 write.csv(subset_with_qc, "/Users/eballer/BBL/msdepression/data/dac/subset_with_qc_for_tim_20211117.csv", col.names = T)
+write.csv(subset_with_qc, "/Users/eballer/BBL/msdepression/data/dac/subset_bad_flairs_for_tim_20211117.csv", col.names = T)
 

### 2022-06-23 msdepression/scripts/make_list_of_good_subj_paths_from_melissa_20211014.R

@@ -1,0 +1,14 @@
+library(dplyr)
+melissa_file <- read.csv("/Users/eballer/BBL/msdepression/data/melissa_martin_files/csv/1yr_report_address.csv", sep = ",", header = T)
+
+#remove rows who have a score of below 75 in column $rating
+pass_qc <- melissa_file[melissa_file$rating>=75.0,]
+
+#make a new column to indicate whether the scan has mimosa or is t1_n4_brain.nii.gz
+pass_qc$mimosa_or_t1 = 0
+
+#assign 1s if patient has pattern mimosa_binary_mask_0.25.nii.gz or t1_n4_brain.nii.gz
+pass_qc$mimosa_or_t1[grep(pattern = "mimosa_binary_mask_0.25.nii.gz", x = pass_qc$scan)] <- 1 #n = 2365
+pass_qc$mimosa_or_t1[grep(pattern = "t1_n4_brain.nii.gz", x = pass_qc$scan)] <- 1 #added 672, new n = 3037
+
+new_df <- pass_qc[pass_qc$mimosa_or_t1 == 1, ]

### 2022-06-23 msdepression/scripts/make_nii_files_dsi_studio.sh

@@ -1,0 +1,59 @@
+#!/bin/bash
+
+#### Welcome to the Region_making party #########
+### Pre: Must have a file with full paths to the lesioned data 
+### Post: Within each directory specified by the lesioned data, will have a directory that has all tractography (ROA, ROI, Full)
+### Uses: For use in MS depression - take a subject's mimosa lesions and generate the fiber tracts (individual fascicles) that run through it
+#dependencies: Using dsi studio downloaded 9/2021
+
+export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+template='/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+default='/Users/eballer/BBL/msdepression/data/mimosa_file_paths'
+fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa path file" $default
+    lesion_file=$default
+else  
+    lesion_file=$1
+fi
+echo "File being read is "$lesion_file
+echo "... Starting to make lesions ..."
+
+lesion_paths=$(cat $lesion_file)
+
+for lesion in ${lesion_paths}; do
+    echo "Working on file" $lesion
+    # make directory within parent directory for tracts
+    #parent_dir=$(echo $lesion | perl -pe s'/(.*)\/.*/$1/g')
+    mask_prefix=$(echo $lesion | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
+    if [ ! -d $parent_dir/fiber_tracking_maps ] 
+    then
+        mkdir $parent_dir/fiber_tracking_maps
+    fi
+
+    fiber_bundle_type_list=(
+        "association"
+        "cerebellum"
+        "cranial_nerve"
+        "projection"
+        "commissural")
+    for fiber_bundle_type in "${fiber_bundle_type_list[@]}"; do
+        echo "Fiber bundle type is " $fiber_bundle_type
+        echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        #mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
+        fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
+        for fascicle in ${fascicles}; do
[…]

### 2022-06-23 msdepression/scripts/make_streamline_volumes_single_subj.sh

@@ -1,0 +1,60 @@
+### Make streamline volumes ###
+
+### Pre: All streamlines must have been made in dsi studio (these are the nii files)
+### Post: One csv that contains the name of the streamline as one column, and the volume of the streamline in a second column, separated by a ','
+### Uses: Will be getting the volumes of the streamlines for use in later density analyses
+### Dependencies: This script requires afni to be installed.
+
+#module load afni
+
+#module load singularity
+#export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+#export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
+num_cores=1
+
+#set default directory/templates. This will change when we use local, personal dwi
+#template='/project/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+#fascicle_directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
+fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+output_csv="${fascicle_directory}/fiber_volume_values.csv"
+
+#initiate csv
+rm -f $output_csv
+touch $output_csv
+echo "fascicle,volume" >> $output_csv
+#if [ $# == 0 ]
+#then
+ #   echo "You did not enter a file. Make sure your call makes sense"
+  #  echo "for now, using default $fascicle_directory"
+ #else  
+ #   lesion=$1
+    #parent_dir=$(echo ${lesion} | perl -pe s'/(.*)\/.*/$1/g')
+    parent_dir=${fascicle_directory}
+#	echo "File being read is "$l
+#	echo "... Starting to make tracts"
+# make directory within parent directory for tracts
+   
+    echo "$parent_dir"
+	#mask_prefix=$(echo ${lesion} | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
+    #echo $mask_prefix
+	#if [ ! -d $parent_dir/fiber_tracking_maps ] 
+    #	then
+     #   	mkdir $parent_dir/fiber_tracking_maps
+    #	fi
+
+    #make sub directories for each bundle type, so within fiber_tracking_maps, there will be subdirectories for each type of map
+	fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
+    	for fiber_bundle_type in ${fiber_bundle_type_list}; do
[…]

### 2022-06-23 msdepression/scripts/make_streamline_volumes_single_subj_local_comp.sh

@@ -1,0 +1,60 @@
+### Make streamline volumes ###
+
+### Pre: All streamlines must have been made in dsi studio (these are the nii files)
+### Post: One csv that contains the name of the streamline as one column, and the volume of the streamline in a second column, separated by a ','
+### Uses: Will be getting the volumes of the streamlines for use in later density analyses
+### Dependencies: This script requires afni to be installed.
+
+#module load afni
+
+#module load singularity
+#export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
+#export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
+num_cores=1
+
+#set default directory/templates. This will change when we use local, personal dwi
+#template='/project/msdepression/templates/dti/HCP1065.1mm.fib.gz'
+#fascicle_directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
+fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'
+
+output_csv="${fascicle_directory}/fiber_volume_values.csv"
+
+#initiate csv
+rm -f $output_csv
+touch $output_csv
+echo "fascicle,volume_full" >> $output_csv
+#if [ $# == 0 ]
+#then
+ #   echo "You did not enter a file. Make sure your call makes sense"
+  #  echo "for now, using default $fascicle_directory"
+ #else  
+ #   lesion=$1
+    #parent_dir=$(echo ${lesion} | perl -pe s'/(.*)\/.*/$1/g')
+    parent_dir=${fascicle_directory}
+#	echo "File being read is "$l
+#	echo "... Starting to make tracts"
+# make directory within parent directory for tracts
+   
+    echo "$parent_dir"
+	#mask_prefix=$(echo ${lesion} | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
+    #echo $mask_prefix
+	#if [ ! -d $parent_dir/fiber_tracking_maps ] 
+    #	then
+     #   	mkdir $parent_dir/fiber_tracking_maps
+    #	fi
+
+    #make sub directories for each bundle type, so within fiber_tracking_maps, there will be subdirectories for each type of map
+	fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
+    	for fiber_bundle_type in ${fiber_bundle_type_list}; do
[…]

### 2022-06-23 msdepression/scripts/parse_melissa_file_to_prepare_for_ANTS_202111014.py

@@ -1,0 +1,38 @@
+import csv
+import re
+import os
+import copy
+
+def add_to_outputs(image, outputs):
+    sub = re.search('sub-[0-9]+', path).group(0)
+    ses = re.search('ses-[0-9]+', path).group(0)
+    run = re.search('run-[0-9]+', path).group(0)
+    image_type = os.path.basename(image).replace(".nii.gz", "")
+    if sub not in outputs:
+        outputs[sub] = {}
+    if ses not in outputs[sub]:
+        outputs[sub][ses] = {}
+    if run not in outputs[sub][ses]:
+        outputs[sub][ses][run] = {}
+    outputs[sub][ses][run][image_type] = image
+    return outputs
+
+
+good_outputs = {}
+with open('1yr_report_address.csv') as csvfile:
+    reader = csv.reader(csvfile, quotechar='"')
+    next(reader) # skip header
+    for row in reader:
+        path = row[0]
+        rating = float(row[2])
+        if rating >= 75 and ('t1_n4_reg_brain_ws.nii.gz' in path or 'flair_n4_brain_ws.nii.gz' in path or 'mimosa_binary_mask_0.25.nii.gz' in path):
+            good_outputs = add_to_outputs(path, good_outputs)
+
+# make sure t1, flair, mask all reviewed
+for sub, v in copy.deepcopy(good_outputs).items(): # copy because can't change dictionary size during iteration
+    for ses, vv in v.items():
+        for run, vvv in vv.items():
+            if len(good_outputs[sub][ses][run]) != 3:
+                del good_outputs[sub][ses][run]
+            else: # then if you want to print it out
+                print(sub, ses, run, good_outputs[sub][ses][run])

### 2022-06-23 msdepression/scripts/streamline_volumes_all_subjs.sh

@@ -1,0 +1,30 @@
+#!/bin/bash
+
+#### Welcome to the Region_making party #########
+### Pre: Must have a file with full paths to the lesioned data 
+### Post: Within each directory specified by the lesioned data, we will have a file that contains the volumes of the streamlines in each subject
+### Uses: For use in MS depression - iterates through each subject and session
+#dependencies: Using dsi studio downloaded 9/2021
+
+default='/project/msdepression/data/melissa_martin_files/csv/mimosa_binary_masks_hcp_space_20211026_n2336'
+num_cores=1
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa path file" $default
+    lesion_file=$default
+else  
+    lesion_file=$1
+fi
+echo "File being read is "$lesion_file
+echo "... Starting to make lesions ..."
+
+lesion_paths=$(cat $lesion_file)
+
+job_count=1
+for lesion in ${lesion_paths}; do
+    echo "Working on file" $lesion
+    bsub -J "job+${job_count}" -n ${num_cores} -o /project/msdepression/scripts/logfiles/3dmaskav_${job_count}.out make_streamline_volumes_single_subj_pmacs.sh $lesion
+    ((job_count+=1))
+done
+  

### 2022-06-24 calc_vol_fascicles_within_yeo_7_net.sh
figuring out overlap between networks and fascicles

@@ -1,0 +1,68 @@
+#### calculate volume of fascicles within the yeo 7 networks ####
+### Pre: All streamlines must have been made in dsi studio (these are the nii files)
+### Post: Yeo7 network mask resampled to template space, file <streamline_volume_within_${network}_network>, containing each fascicle and its corresponding % within yeo network
+### Uses: Trying to figure out which fascicles are in the yeo 7 network for dimensionality reduction for use in later regressions (i.e. decrease # comparisons by only looking at fascicles within yeo 7 networks)
+### Dependencies: This script requires afni to be installed (currently using 20.1)
+
+#assemble yeo 7 networks
+
+yeo_7_net_list=("VIS"
+                "MOT"
+                "DA"
+                "VA"
+                "LIM"
+                "FP"
+                "DM")
+
+for network in "${yeo_7_net_list[@]}"; do
+	#set outfile
+	outfile="/project/msdepression/results/streamline_volume_within_${network}_network.csv"
+
+	#set template directories, will be making our resampled mask from within this
+	template="yeo_7_net_${network}.nii"
+	template_dir="/project/msdepression/templates/Yeo_JNeurophysiol11_MNI152/"
+	resampled_filename="resampled_${template}"
+	template_fullpath="${template_dir}/${template}"
+
+	#set fascicle directory
+	fascicle_direc="/project/msdepression/templates/dti/HCP_YA1065_tractography/"
+	fascicle_type="association"
+	fascicle_for_resampling="AF_L.nii.gz"
+
+	#set home directory
+	homedir="/project/msdepression/scripts/"
+
+	#initiate csvs
+	rm -f $outfile
+	touch $outfile
+	echo "fascicle,num_voxels_total,non_zero_voxels_in_${network}_map,prop_in_mask" >> $outfile
+
+	#resample yeo_7_net_${network}.nii to match sample fascicle - AF_L.nii.gz
+	if [ -f "${template_dir}/${resampled_filename}" ]; then
+		rm -f ${template_dir}/${resampled_filename}
+	fi
+	3dresample -master ${fascicle_direc}/${fascicle_type}/${fascicle_for_resampling} -prefix ${template_dir}/${resampled_filename} -input ${template_fullpath} -rmode NN
+
[…]

### 2022-06-24 calc_vol_fascicles_within_yeo_7_net.sh
now have files with list of overlaps/binarized

@@ -1,8 +1,11 @@
 #### calculate volume of fascicles within the yeo 7 networks ####
 ### Pre: All streamlines must have been made in dsi studio (these are the nii files)
-### Post: Yeo7 network mask resampled to template space, file <streamline_volume_within_${network}_network>, containing each fascicle and its corresponding % within yeo network
+### Post: Yeo7 network mask resampled to template space, file <streamline_volume_within_${network}_network>, containing each fascicle and its corresponding % within yeo networ. Additionally, will have a file that contains all of the overlaps per each network, and another with a binarized (yes/no)k
 ### Uses: Trying to figure out which fascicles are in the yeo 7 network for dimensionality reduction for use in later regressions (i.e. decrease # comparisons by only looking at fascicles within yeo 7 networks)
 ### Dependencies: This script requires afni to be installed (currently using 20.1)
+
+#output directory
+outdir="/project/msdepression/results/"
 
 #assemble yeo 7 networks
 
@@ -14,6 +17,12 @@ yeo_7_net_list=("VIS"
                 "FP"
                 "DM")
 
+#set up output file names
+all_net_overlap_proportions="${outdir}/yeo_7_network_overlap_proportions.txt"
+all_net_overlap_proportions_binarized="${outdir}/yeo_7_network_overlap_proportions_binarized.txt"
+
+#set counter to keep track of whether it is the first iteration (and you need to do a special column add), or just take the last line and pr it
+counter=1
 for network in "${yeo_7_net_list[@]}"; do
 	#set outfile
 	outfile="/project/msdepression/results/streamline_volume_within_${network}_network.csv"
@@ -35,7 +44,7 @@ for network in "${yeo_7_net_list[@]}"; do
 	#initiate csvs
 	rm -f $outfile
 	touch $outfile
-	echo "fascicle,num_voxels_total,non_zero_voxels_in_${network}_map,prop_in_mask" >> $outfile
+	echo "fascicle,num_voxels_total,non_zero_voxels_in_${network}_map,prop_in_mask_${network}" >> $outfile
 
 	#resample yeo_7_net_${network}.nii to match sample fascicle - AF_L.nii.gz
 	if [ -f "${template_dir}/${resampled_filename}" ]; then
[…]

### 2022-06-29 msdepression/scripts/afni_commands_to_construct_yeo_indiv_vol_masks.sh
changed the ispositive command
It was redundant, so went from ispositive(a*amongst(a,{network_name}) to amongst(a,{network_name} which returns the binary mask.

@@ -2,8 +2,8 @@
 
 # pre: Yeo cortical parcellations (7 network) in volumetric space: https://surfer.nmr.mgh.harvard.edu/fswiki/CorticalParcellation_Yeo2011
 # post: 7 individual network maps: VIS, MOT, DA, VA, LIM, FP, DM, labeled yeo_7_vol_[network abbreviation]
-# uses: This script takes the parcellations in volume (Yeo2011_7Networks_MNI152_FreeSurferConformed1mm_LiberalMask.nii) and makes 7 individual maps. In the Yeo2011 volume mask, each network is given a number (1-7), with 1 corresponding to VIS, and 7 corresponding to DM. In order to make individual maps, I need to separate out each network in the map based on intensity. Will try to code in the following way: 1) make a vector where each item is the name of the network 2) make a loop that iterates 1-7, have a second counter that goes in the reverse direction 3) For each iteration, the count corresponds to both the intensity in the mask as well as its position in the name vector 4) To separate out the values with a specific intensity, need to do some weird math (b/c I can't just pull out intensity directly: If my count is 1, I need to pull out only values that are 1 from the map. This is equivalent to ispositive(a-[1-1])+isnegative(a-[1+1]). For 2, ispositive(a-[2-1]+isnegative(a-[2+1]) 
-# dependencies: Afni 
+# uses: This script takes the parcellations in volume (Yeo2011_7Networks_MNI152_FreeSurferConformed1mm_LiberalMask.nii) and makes 7 individual maps. In the Yeo2011 volume mask, each network is given a number (1-7), with 1 corresponding to VIS, and 7 corresponding to DM. In order to make individual maps, I need to separate out each network in the map based on intensity. Will try to code in the following way: 1) make a vector where each item is the name of the network 2) make a loop that counts the networks 3) go through each network and do math to pull out only the network with a particular intensity. It uses the command "amongst", which gives back the voxels where the intensity is equivalent to the counter in a binarized mask
+# dependencies: Afni
 
 #make list of yeo network abbreviations
 yeo_7_net_list=("VIS"
@@ -44,7 +44,9 @@ echo "Making individual network maps"
 #have to write out yeo_7_net_list because bash can't handle too many extractions. Prob because I'm not referencing an object but a name when pointing network list to yeo_7_net_list
 for network in "${yeo_7_net_list[@]}"; do
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/20180104_group_selection_and_demo_by_gender_matching_age_and_race.R
no reason to have this in directory

@@ -1,252 +1,0 @@
-#install.packages("visreg")
-#install.packages("mgcv")
-#install.packages("tableone")
-#################################################
-#### Get subset for Baller_DepHeterogeneity ####
-#################################################
-
-#This does the subsetting, runs statistics with lapply instead of loop, visualizes progress with visreg, and does fdr correction
-# it then makes table 1 (demographics)
-#load libraries
-library(visreg)
-library(mgcv)
-library(tableone)
-
-#read in csvs
-demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
-cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
-health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
-psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
-
-#remove people with NA for race, age, or sex.  START WITH N = 9498
-demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
-demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
-demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/DataPrep.R

@@ -1,130 +1,0 @@
-#################
-### LOAD DATA ###
-#################
-
-#Demographic data (n=1629)
-data.demo <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/demographics/n1601_demographics_go1_20161212.csv", header=TRUE) 
-
-##Clinical data
-#Screening diagnoses (n=1601)
-data.diag <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/clinical/n1601_goassess_psych_summary_vars_20131014.csv", header=TRUE, na.strings=".")
-
-#Bifactors and correlated traits (n=1601)
-data.factors <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/clinical/n1601_goassess_clinical_factor_scores_20161212.csv", header=TRUE, na.strings=".")
-
-#State trait anxiety data (n=1391)
-data.stai <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/clinical/n1601_stai_pre_post_itemwise_smry_factors_20170131.csv", header=TRUE, na.strings=".")
-
-#Exclusion data (n=1601)
-#Health exclusion (use the new healthExcludev2 variable)
-data.healthExclude <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/health/n1601_health_20170421.csv", header=TRUE, na.strings=".")
-
-#T1 QA exclusion (n=1601)
-data.t1QA <- read.csv("/data/joy/BBL/studies/pnc/n1601_dataFreeze/neuroimaging/t1struct/n1601_t1QaData_20170306.csv", header=TRUE, na.strings=".")
-
-##################
-#### DATA PREP ###
-##################
-
-#Transform the age variable from months to years
-data.demo$age <- (data.demo$ageAtScan1)/12
-
-#Recode male as 0 and female as 1
-data.demo$sex[which(data.demo$sex==1)] <- 0
-data.demo$sex[which(data.demo$sex==2)] <- 1
-
-##################
-### MERGE DATA ###
-##################
-dataMerge1 <-merge(data.demo,data.diag, by=c("bblid","scanid"), all=TRUE) 
-dataMerge2 <-merge(dataMerge1,data.factors, by=c("bblid","scanid"), all=TRUE) 
-dataMerge3 <-merge(dataMerge2,data.stai, by=c("bblid","scanid"), all=TRUE) 
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/GamAnalyses.R

@@ -1,45 +1,0 @@
-e##########################################
-#### GAM MODELS FOR PREMATURITY STUDY ####
-##########################################
-
-#Load data
-data.NMF <- read.csv("/data/joy/BBL/projects/pncPreterm/subjectData/n282_Prematurity_allData.csv", header=TRUE, na.strings = "NA")
-
-#Remove those who are missing maternal level of education data
-data.NMF <- data.NMF[!is.na(data.NMF$medu1),]
-
-#Load library
-library(mgcv)
-
-#Get NMF variable names
-nmfComponents <- names(data.NMF)[grep("Nmf26",names(data.NMF))]
-
-#Run gam models
-NmfModels <- lapply(nmfComponents, function(x) {
-  gam(substitute(i ~ s(age) + sex + medu1 + ga, list(i = as.name(x))), method="REML", data = data.NMF)
-})
-
-#Look at model summaries
-models <- lapply(NmfModels, summary)
-
-#Pull p-values
-p <- sapply(NmfModels, function(v) summary(v)$p.table[4,4])
-
-#Convert to data frame
-p <- as.data.frame(p)
-
-#Print original p-values to three decimal places
-p_round <- round(p,3)
-
-#FDR correct p-values
-pfdr <- p.adjust(p[,1],method="fdr")
-
-#Convert to data frame
-pfdr <- as.data.frame(pfdr)
-
-#To print fdr-corrected p-values to three decimal places
-pfdr_round <- round(pfdr,3)
-
-#List the NMF components that survive FDR correction
-Nmf_fdr <- row.names(pfdr)[pfdr<0.05]
-

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra20180115.R

@@ -1,155 +1,0 @@
-library(visreg)
-library(mgcv)
-library(tableone)
-library(dplyr)
-library(plm)
-library(MatchIt)
-
-#read in csvs
-demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
-cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
-health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
-psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
-
-#remove people with NA for race, age, or sex.  START WITH N = 9498
-demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
-demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
-demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
-demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406
-
-#remove people with NA for depression or total psych score, START WITH N = 9498
-psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra20180124.R

@@ -1,267 +1,0 @@
-library(visreg)
-library(mgcv)
-library(tableone)
-library(dplyr)
-library(plm)
-library(MatchIt)
-
-#read in csvs
-demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
-cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
-health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
-psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
-
-#remove people with NA for race, age, or sex.  START WITH N = 9498
-demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
-demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
-demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
-demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406
-
-#remove people with NA for depression or total psych score, START WITH N = 9498
-psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra_20180212.R

@@ -1,295 +1,0 @@
-library(visreg)
-library(mgcv)
-library(tableone)
-library(dplyr)
-library(plm)
-library(MatchIt)
-
-###### This script reads in demographics, cnb_scores, health and psych summaries, merges them, removes NAs, codes and separates by depression#####
-####Also preps for hydra, both using a typical GAM model and matching (we lose a lot of people) and also with residuals plotted so we don't have to match#########
-########Also provides matched data sets and tests them, if we decide to use them.  It significantly reduces N to match (dataset from 3022 to 1424)
-
-
-#######################################################
-############ READ IN, MERGE AND SUBSET DATA############
-#######################################################
-
-#read in csvs
-demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
-cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
-health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
-psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
-
-#remove people with NA for race, age, or sex.  START WITH N = 9498
-demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Sunny_matching_script.R

@@ -1,412 +1,0 @@
-library(MASS)
-library(Matching)
-library(tidyr)
-library(readr)
-library(effsize)
-library(dplyr)
-
-###Loading Data
-alldata<-read.csv("OlfactionFromStata_2017_06-19.csv") %>% 
-  group_by(group) %>%
-  arrange(group,bblid)
-View(alldata)
-str(alldata)
-colnames(alldata)
-
-#Cleaning up variables
-alldata$anyspectrum<-as.factor(alldata$anyspectrum)
-
-###Calculating and Z-transforming CNB Categories 
-# Creating normalized summary variables
-zacexe <- as.vector(
-  scale(
-    rowMeans(
-      data.frame(alldata$zaclnb, alldata$zacpcet, alldata$zacpcpt), na.rm = TRUE)
-  ))
-
-zacmem <- as.vector(
-  scale(
-    rowMeans(
-      data.frame(alldata$zaccpf, alldata$zaccpw, alldata$zacvolt), na.rm = TRUE)
-  ))
-
-zaccog <- as.vector(
-  scale(
-    rowMeans(
-      data.frame(alldata$zacpvrt, alldata$zacpmat, alldata$zacplot), na.rm = TRUE)
-  ))
-
-zacsoc <- as.vector(
-  scale(
-    rowMeans(
-      data.frame(alldata$zacadt, alldata$zacer, alldata$zacedf), na.rm = TRUE)
-  ))
-
-zacall <- as.vector(
-  scale(
-    rowMeans(
-      data.frame(zacexe, zacmem, zaccog, zacsoc), na.rm = TRUE)
-  ))
-
-#Merging with existing data
-alldata$zacexe = zacexe
-alldata$zacmem = zacmem
-alldata$zaccog = zaccog
-alldata$zacsoc = zacsoc
-alldata$zacall = zacall
-
-alldata<-data.frame(alldata,zacall)
-alldata$zacall.1 <- NULL
-
-#Checking
-colnames(alldata)
-summary(alldata$zacall)
-sd(alldata$zacall, na.rm = TRUE)
-
-###Selecting Important Variables for Analysis
-dat <- select(alldata, bblid, group, sex, race, age, edu, avgedu, hand, smoker, anysmoke,
-              id, disc, lyral, cit, 
-              zacexe, zacmem, zaccog, zacsoc, zacall, wratstd, gafc, comtgenotype,
-              positive, negative, sipstot, anyspectrum, psyscale)
-
-View(dat)
-table(dat$group)
-
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/subsetting_for_initial_group.R

@@ -1,123 +1,0 @@
-#install.packages("visreg")
-#################################################
-#### Get subset for Baller_Dep_Heterogeneity ####
-#################################################
-
-#load libraries
-library("visreg")
-library(mgcv)
-
-#read in csvs
-demographics <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",")
-cnb_scores <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",")
-health <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",")
-psych_summary <- read.csv("/Users/test/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",")
-
-#remove people with NA for race, age, or sex.  START WITH N = 9498
-demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
-demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
-demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
-demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406
-
-#remove people with NA for depression or total psych score, START WITH N = 9498
-psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
-psych_summary_no_NA_dep_and_smry_psych_overall <- psych_summary_no_NA_dep[!is.na(psych_summary_no_NA_dep$smry_psych_overall_rtg),] #take out those with NA for overall psych rtg, no additional people lost, N = 9411
-
-#merge the csvs
-#dem_cnb <- merge(demographics, cnb_scores, by = "bblid") #merge demographics and cnb #this is if we want to include people without full demographic data
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/subsetting_for_initial_group_with_lapply.R

@@ -1,155 +1,0 @@
-#install.packages("visreg")
-#install.packages("mgcv")
-#install.packages("tableone")
-#################################################
-#### Get subset for Baller_DepHeterogeneity ####
-#################################################
-
-#This does the subsetting, runs statistics with lapply instead of loop, visualizes progress with visreg, and does fdr correction
-# it then makes table 1 (demographics)
-#load libraries
-library(visreg)
-library(mgcv)
-library(tableone)
-
-#read in csvs
-demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
-cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
-health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
-psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
-
-#remove people with NA for race, age, or sex.  START WITH N = 9498
-demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
-demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
-demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
[…]

### 2022-06-29 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/subsetting_for_initial_group_with_lapply_server_path.R

@@ -1,152 +1,0 @@
-#install.packages("visreg")
-#install.packages("mgcv")
-#install.packages("tableone")
-#################################################
-#### Get subset for Baller_DepHeterogeneity ####
-#################################################
-
-#This does the subsetting, runs statistics with lapply instead of loop, visualizes progress with visreg, and does fdr correction
-# it then makes table 1 (demographics)
-#load libraries
-library(visreg)
-library(mgcv)
-library(tableone)
-
-#read in csvs
-demographics <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
-cnb_scores <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
-health <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
-psych_summary <- read.csv("/data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv
-
-#remove people with NA for race, age, or sex.  START WITH N = 9498
-demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
-demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
-demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
[…]

### 2022-06-29 calc_vol_fascicles_within_depression_mask_3_09.sh
new script for 3_09 mask and filters out any maps with mask in them

@@ -1,14 +1,14 @@
-#### calculate volume of fascicles within the depression network ####
+#### calculate volume of fascicles within the depression network, thresholded at 3.09 ####
 ### Pre: All streamlines must have been made in dsi studio (these are the nii files)
 ### Post: A depression mask resampled to template space, file <streamline_volume_within_dep_network>, containing each fascicle and its corresponding % within depression network
 ### Uses: Trying to figure out which fascicles are in the depression network for dimensionality reduction for use in later regressions (i.e. decrease # comparisons by only looking at fascicles within depression network)
 ### Dependencies: This script requires afni to be installed (currently using 20.1)
 
 #set outfile
-outfile="/project/msdepression/results/streamline_volume_within_dep_network.csv"
+outfile="/project/msdepression/results/streamline_volume_within_dep_network_3_09.csv"
 
 #set template directories, will be making our resampled mask from within this
-template="Depression_Clust_mask_roi_binarized.nii.gz"
+template="Dep_clust_T_3_09_binarized.nii"
 template_dir="/project/msdepression/templates/harvard_depression"
 resampled_filename="resampled_${template}"
 template_fullpath="/project/msdepression/templates/harvard_depression/${template}"
@@ -26,7 +26,7 @@ rm -f $outfile
 touch $outfile
 echo "fascicle,num_voxels_total,non_zero_voxels_in_dep_map,prop_in_mask" >> $outfile
 
-#resample Depression_Clust_mask_roi_binarized.nii.gz to match sample fascicle - AF_L.nii.gz
+#resample Depression mask to match sample fascicle - AF_L.nii.gz
 if [ -f "${template_dir}/${resampled_filename}" ]; then
 	rm -f ${template_dir}/${resampled_filename}
 fi

### 2022-07-01 msdepression/scripts/Descriptives_potential_paper_1_20220622.Rmd
Add files via upload

@@
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---
@@
 Results
-  *Demographics
+
+  *Demographics*
+
     total n unique subjects = 783 (F = 609, M = 174) 
@@
 
-  *Age effects (unique):
+  *Age effects (unique): *
+
       -Age mean 47 (range 20-83). 
@@
 
-  *Age effects (gam unique)
+  *Age effects (gam unique)*
+
       - 62 out of 87 fascicles correct, 26 in association network
 
-  *Age effects (mixed model)
+  *Age effects (mixed model)*
+
       - 57 out of 87 correct, 27 in association network
 
-  *Sex effects (unique):
+  *Sex effects (unique):*
+
       - In the final good MIMoSA group, there were 609 unique females and 174 unique males, 77% women
-      - Individual fascicle results underwhelming, only 3 regions meet P<0.05 (unc), 
+      - Individual fascicle results underwhelming, P<0.05 (unc), 
           SCP	0.020859515			
@@
 
-  *Sex effects (mixed model):
+  *Sex effects (mixed model):*
+
       - no action
@@
 
-  * Age*sex (unique)
+  *Agexsex (unique)*
+
       - no action
 
-  * Age * sex (mixed model) 
+  *Agexsex (mixed model) *
+
   lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM*osex + (1|EMPI), list(i = as.name(x))), data = df_mm)
@@
 #fdr corrected only association cortex
+fascicle_bundle_mapping=get_fascicle_bundle_mapping()
 fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]

### 2022-07-01 msdepression/scripts/K23_prelim_data_using_dep_T_3_09_mask_20220630.Rmd
Add files via upload

--- 2022-06-23 msdepression/scripts/K23_prelim_data_started_20220225.Rmd
@@
 ---
-title: "MS Prelim Data K23"
+title: "MS Prelim Data K23, using Depression T=3.09 mask"
 author: "Erica Baller"
-date: "2/25/2022"
+date: "6/30/22"
 output:
@@
 
-Ideas:
-
-Overall:
-
-[] demographics - tables made below
-
-[] imaging parameters
-
-Aim 1: lesion network mapping
-[] some picture showing the # of people who are scanned and the # of people with depression. Maybe an adaptation of Syndey's picture somehow. 
-[] mimosa picture
-[] picture of filtering pipeline
-[] comparison of some of the histograms for high volume loss similar to prev literature(CN II, slf), not so much in other areas
-[] results of age analysis [maybe/maybe not, to prove the method works]?
-[] results of disease in vs outside of depression network (corrected), lead with this?
-[] results total proportion lost (single p)
-[] results of dep vs controls across all regions,(uncorrected)
-[] results of dep vs controls in depression network (uncorrected)
-
-[] prelim network enrichment via spin-testing or just citation?
-- do I want to validate here or somewhere else with neurosynth maps?
-
-Aim 2: prospective
-[] power analysis
-[] CNB (Bart has great Figure 3)
-[] Goassess
-
-Aim 3
-[] Hydra picture
-[] hydra results from my paper
-[] acquisition protocol (T1, t2, flair, dwi)
-
 ## R Markdown
@@
 As such, the MIMoSA subset is smaller.
-It does however highlight the
 
 <img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>
+
+Lastly, unlike the prelim K data analyses, this uses a depression mask that I generated which is slightly more leniant than the one I used for my first iteration (Matt thresholded that at 3.49.
+I threshold this one a 3.09).The only change I made it changing which file is fed into the depression analysis.
 
@@
 
-Depression network:
- [1] "AF_L"    "AF_R"    "C_PHP_R" "EMC_L"   "EMC_R"   "FAT_L"   "FAT_R"   "IFOF_L"  "IFOF_R"  "ILF_R"   "MdLF_L" 
-[12] "MdLF_R"  "PAT_L"   "PAT_R"   "SLF2_L"  "SLF2_R"  "SLF3_L"  "SLF3_R"  "V"       "AR_R"    "CBT_L"   "CBT_R"  
-[23] "CPT_F_L" "CPT_F_R" "CPT_O_L" "CPT_O_R" "CPT_P_L" "CPT_P_R" "CS_A_L"  "CS_A_R"  "CS_P_L"  "CS_P_R"  "CS_S_L" 
-[34] "CS_S_R"  "CST_L"   "CST_R"   "DRTT_R"  "ML_L"    "RST_L"   "RST_R"   "TR_A_L"  "TR_A_R"  "TR_P_L"  "TR_P_R" 
-[45] "TR_S_L"  "TR_S_R"  "CC"
-
-Depression 10% network
- [1] "AF_L"    "AF_R"    "C_PHP_R" "FAT_L"   "FAT_R"   "MdLF_L"  "MdLF_R"  "SLF2_L"  "SLF2_R"  "SLF3_R"  "CBT_L"  
-[12] "CBT_R"   "CPT_O_R" "CS_P_R"  "TR_P_R"
+Depression network (* where there are new regions as compared to the 3.49 mask, now we are up to 62 regions):
+[1] "AF_L"   "AF_R"    "C_FP_L"*  "C_FP_R"*  "C_PHP_L"* "C_PHP_R" "C_R_R"*   "EMC_L"   "EMC_R"   "FAT_L"   "FAT_R" [12] "IFOF_L"  "IFOF_R"  "ILF_L"*   "ILF_R"   "MdLF_L"  "MdLF_R"  "PAT_L" "PAT_R"   "SLF1_L"*  "SLF1_R"*  "SLF2_L" [23] "SLF2_R"  "SLF3_L"  "SLF3_R"  "VOF_R"*   "CB_L"*    "CB_R"*    "MCP"*     "V"   "AR_L"*    "AR_R"    "CBT_L" [34] "CBT_R"   "CPT_F_L" "CPT_F_R" "CPT_O_L" "CPT_O_R" "CPT_P_L" "CPT_P_R" "CS_A_L"  "CS_A_R"  "CS_P_L"  "CS_P_R"
+[45] "CS_S_L"  "CS_S_R"  "CST_L"   "CST_R"   "DRTT_L"*  "DRTT_R"  "F_R"*   "ML_L"    "RST_L"   "RST_R"   "TR_A_L" 
+[56] "TR_A_R"  "TR_P_L"  "TR_P_R"  "TR_S_L"  "TR_S_R"  "AC"*     "CC"
+
+Depression 10% network (stars where there are new regions as compared to the 3.49 mask, 21 regions)
+ [1] "AF_L" "AF_R"    "C_PHP_R" "EMC_R"*   "FAT_L"*   "FAT_R"   "MdLF_L"  "MdLF_R"  "PAT_L"*   "PAT_R"*   "SLF2_L"
+[12] "SLF2_R"  "SLF3_R"  "CBT_L"   "CBT_R"   "CPT_O_R" "CS_P_R"  "TR_P_L"*  "TR_P_R"  "TR_S_L"*  "TR_S_R" *
 
@@
       - Sex effects interaction model (fascicle ~ depGroupVar*osex + PAT_AGE_AT_EXAM)
-        - Dep effects - same as above, 12 fascicles (p<0.05 unc)
-        - Sex effects - same as above, with the edition of CNVIII_R (p<0.05 unc)
         - age*sex interactions - SLF1_L and CNVIII_R (p<0.05 unc)
@@
 
-    *Depression over all fascicles:
+    *Depression over all fascicles (also, for within/outside dep network, used rowMeans and rowSums, doesn't change anything w/r/t statistics, you just get sums instead of means for means between groups):
       - The exciting news of the day! With our new depression group (moving the meds people), we have findings
       - total proportion lost - p = 0.049, depressed (27%) more loss than healthy(24%)
-      - total proportion lost within depression network (47 total) - trend (0.059) between depressed (35%) and healthy(30%)
-      - total proportion lost within depression 10% network (15 fascicles total) - p = 0.049 (depressed (33%) have more loss than healthy (27%))
-      - total proportion lost within depression 5% network (24 fascicles total) -  p = 0.056 (depressed (32%) have more loss than healthy (27%))
-      - total proportion lost outside depression network (40 fascicles total)- p = 0.037 (depressed (18%) more than healthy(16%))
+      - total proportion lost within depression network (xx total) - trend (0.054) between depressed (33%) and healthy(28%)
+      - total proportion lost within depression 10% network (xx fascicles total) - p = 0.055 (depressed (32%) have more loss than healthy (27%))
+      - total proportion lost within depression 5% network (xx fascicles total) -  p = 0.053 (depressed (32%) have more loss than healthy (27%))
+      - total proportion lost outside depression network (xx fascicles total)- p = 0.042 (depressed (14%) more than healthy(13%))
       - take home - people with a lifetime history of depression have more disease than nondepressed individuals. This finding is not specific to the depression network, but is overall.
@@
     *MS disease and depression network overlap:
-      - Ms disease within depression network (35%) versus outside (17%), p < 2.2* 10^-16
-      - MS disease within depression 10% network (31%) versus outside (17%), p < 2.2 * 10^-16
-      - MS disease within depression 5% network (30%) versus outside (17%), p < 2.2 * 10^-16
+      - Ms disease within depression network (31%) versus outside (14%), p < 2.2* 10^-16
+      - MS disease within depression 10% network (30%) versus outside (14%), p < 2.2 * 10^-16
+      - MS disease within depression 5% network (30%) versus outside (14%), p < 2.2 * 10^-16
       - Linear Regressions (essentially matching t tests above): 
-         +depressed v healthy within dep network in (p = 0.061); 
-         +dep vs healthy within dep 10% (p = 0.049);
-         +dep vs healthy within non dep network (p = 0.037)
+         +depressed v healthy within dep network in (p = 0.055); 
+         +dep vs healthy within dep 10% (p = 0.053);
+         +dep vs healthy within non dep network (p = 0.042)
       - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis.
@@
       - Nothing corrects at this time, unfortunately. 
-      - P<0.05, uncorrected findings (* indicates association network):
-          -AF_R	0.01168289	*		
-          -FAT_R	0.01828607 *			
-          -SLF2_R	0.01901378 *		
-          -VOF_R	0.04741569 *			
-          -CPT_F_R	0.02887163 			
-          -CPT_O_L	0.02238344			
-          -CPT_P_R	0.02053104			
-          -CS_S_R	0.03783654			
-          -DRTT_R	0.00695066			
-          -ML_R	0.02511932	
+      - P<0.05, uncorrected findings:
+          -AF_R	0.03235420			
+          -FAT_R	0.03337075			
+          -SLF2_R	0.04066933			
+          -CPT_F_R	0.04932598			
+          -CPT_O_L	0.04346334			
+          -CPT_P_R	0.03471933			
+          -DRTT_R	0.01914199			
+          -ML_R	0.01259474			
+          -TR_S_R	0.01724321	
     Fascicle loss x PHQ9:
@@
      - similarly, no significant findings looks at means of all fascicle loss, within dep network, 10% and 5% and non-depression network
+
+     Depression by sex: there is an ineraction, interestingly:
+     - Males with depression have higher loss in depression network, and males without depression have less loss in depression network, as compared to females.
+I.e.
+females, depression yes/no less related to lesions in/out dep network
+
+     * Comparison of proportion overlap between depression network 3.49, 3.09
+
+     * Proportion lost per volume x proportion overlap in depression network (prop lost * dep proportion)
+     - 3.09;  only uncorrected values when looking at the relationship between depression diagnosis and lesion hit (can only look at within depression map in this analysis)
+     -  AF_R_3_09	0.012274944
+        FAT_R_3_09	0.022765660			
+        SLF2_R_3_09	0.021441998			
+        VOF_R_3_09	0.049555998			
+        CPT_F_R_3_09	0.031783931			
+        CPT_O_L_3_09	0.023932661			
+        CPT_P_R_3_09	0.020599230			
+        CS_S_R_3_09	0.042603557			
+        DRTT_R_3_09	0.008120982			
+        TR_A_R_3_09	0.049682290
 
@@
 
-write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
+#write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
 
@@
 ########
-fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), header = T, sep = ",")
+
+#### THIS WAS CHANGED 6/30 from streamline_volume_within_dep_network.csv to streamline_volume_within_dep_network_3_09.csv
+
+fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
 
@@
 
-write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
+#changed this 6/30 by adding 3_09, need this for bundle mapping
+fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
+write.table(fascicle_volumes_dep_network, file=fascicle_names_dep_network_grouping, sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
 
@@
 #fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
-#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all.csv", row.names = F, quote = F, col.names = F)
+#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all_3_09.csv", row.names = F, quote = F, col.names = F)
 
 fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
-#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)
+#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent_3_09.csv", row.names = F, quote = F, col.names = F)
 
@@
 #fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
-fascicle_bundle_mapping <- get_fascicle_bundle_mapping()
+fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)
 
@@
 #this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
-hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20)
+hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
 
@@
 dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
-summary(dep_vs_healthy) #p=0.494, cohen d = 0.2
+summary(dep_vs_healthy) #p=0.0494
 
@@
 #test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
-total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049, cohen d = 0.17
 
@@
 
-#look at within depression network specifically - trend, p = 0.059
-total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 36%, mean non_dep = 32%, p = 0.071
+#look at within depression network specifically - trend, p = 0.054
+total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
 
 #look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
-#yes, p = 0.048
-total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048
+#yes, p = 0.055
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.055
 
 #look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
-#yes, 
-total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.056
-
-#look at outside depression network, yes p = 0.037
-total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.037
-
-### Age actually makes things worse
-#dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-#summary(dep_vs_healthy_w_age)
+#yes, p=0.053
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.053
+
+#look at outside depression network, yes p = 0.042
+total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 14%, mean non_dep = 13%, p = 0.042
+
+### Age actually makes things worse, switched age and dep group var around, values are the same
+#however, when I send an anova over the linear model, depression group var becomes significant
+dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ PAT_AGE_AT_EXAM + depGroupVar , data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_w_age) #dep group var p = 0.101
+dep_vs_healthy_w_age_anova=anova(dep_vs_healthy_w_age) # p = 0.044
+
+#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
+
+```
+
+```{r depression_unique_all_fascicles_together_in_dep_networks_sum_rather_than_mean}
+
+#### SUM DOESN"T CHANGE STATS AT ALL
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowSums(across(fascicle_names))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowSums(across(fascicle_names_dep_network_all))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_10_percent =
+           rowSums(across(fascicle_names_dep_network_10_percent))) %>%
+   mutate(total_fascicle_prop_lost_indepnet_5_percent =
+           rowSums(across(fascicle_names_dep_network_5_percent))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowSums(across(fascicle_names_nondep_network))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy) #p=0.0494
+
+#power analysis to predict sample required for K, u = num df, v = denom df, f2=0.01 (cohen d = 0.02/2) 
+#prop in each fascicle predicting dep (87), number of predictors (age, sex, ef, ess?) = 4, u = 87-4 = 83
+pwr.f2.test(u=83, f2 = 0.02, sig.level = 0.5, power = 0.8)
+
+#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049, cohen d = 0.17
+
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
+df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
+
+plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
+  geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))
+
+print(plot)
+
+#bar plot
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of Fascicles Affected") +
+  ylim(0,0.5) +
+  theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Proportion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
+
+#look at within depression network specifically - trend, p = 0.054
+total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 20.2, mean non_dep = 17.4, p = 0.054
+
+#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
+#yes, p = 0.055
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 6.7, mean non_dep = 5.6, p = 0.055
+
+#look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
+#yes, p=0.053
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 10.2, mean non_dep = 8.6, p = 0.053
+
+#look at outside depression network, yes p = 0.042
+total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 3.6, mean non_dep = 3.2, p = 0.042
+
+### Age actually makes things worse, switched age and dep group var around, values are the same
+#however, when I send an anova over the linear model, depression group var becomes significant
+dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ PAT_AGE_AT_EXAM + depGroupVar , data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_w_age) #dep group var p = 0.101
+dep_vs_healthy_w_age_anova=anova(dep_vs_healthy_w_age) # p = 0.044
 
@@
 #test whether there is differentially more volume loss in depression versus nondepression network
-volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 35%, mean non_dep_net =  17%, p 2.2 x 10^-16
+volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 31%, mean non_dep_net =  14%, p 2.2 x 10^-16
 
 volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
-dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 31%, mean non_dep_net = 17%, p 2.2 * 10^-16
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 14%, p 2.2 * 10^-16
 
 volume_lost_dep5_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent,
-dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 17%, p 2.2 * 10^-16
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 14%, p 2.2 * 10^-16
 
@@
 
-loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #non sig, p = 0.061
-loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.0492
-loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.037
+loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #non sig, p = 0.055
+loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.056
+loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.042
+
 ```
@@
 
+#uncorrected
+fascicle_anova_mm_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
+fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
+print(fascicle_anova_mm_p05_unc)
 #visreg
@@
 ```
+```{r compare_proportion_overlap_in_dep_3_49_and_3_09_maps}
+fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
+
+#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
+fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
+
+dep_map_3_49 <-
+  read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), sep = ",", header = T)
+dep_map_3_09 <-
+  read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), sep = ",", header = T)
+
+dep_map_together <- data.frame(dep_map_3_49$fascicle, round(dep_map_3_49$prop_in_mask,3), round(dep_map_3_09$prop_in_mask,3))
+names(dep_map_together) <- c("fascicle", "dep_map_3_49_prop", "dep_map_3_09_prop")
+
+#get fascicle names vectors
+#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
+fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
+names_dep_map_3_49 <- gsub("^(.*)", "\\1_3_49", fascicle_names)
+names_dep_map_3_09 <- gsub("^(.*)", "\\1_3_09", fascicle_names)
+
+#multiply each row (Margin=2 equals by row), each %overlap w/depression map of that fascicle, this is like a dysconnectome score, but doesn't take into account the STRENGTH(T) of the voxels in the depression mask.
+Using the third column to the end of the fascicle_proportions data_frame because the first 2 are empi and date
+lesion_prop_x_prop_in_dep_mask_3_49<-sweep(fascicle_proportions[,3:(dim(fascicle_proportions)[2])], MARGIN=2, dep_map_together$dep_map_3_49_prop, `*`)
+names(lesion_prop_x_prop_in_dep_mask_3_49) <- names_dep_map_3_49
+
+lesion_prop_x_prop_in_dep_mask_3_09<-sweep(fascicle_proportions[,3:(dim(fascicle_proportions)[2])], MARGIN=2, dep_map_together$dep_map_3_09_prop, `*`)
+names(lesion_prop_x_prop_in_dep_mask_3_09) <- names_dep_map_3_09
+
+#merge them all together, big data frame, the proportion of each fascicle that is affected per person, and the proportion of that fascicle with the proportion of depression overlap for that fascicle
+
+fascicle_volumes_dep_w_lesion_prop_x_prop_in_dep_mask <- data.frame(fascicle_proportions, lesion_prop_x_prop_in_dep_mask_3_49, lesion_prop_x_prop_in_dep_mask_3_09)
+
+df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_volumes_dep_w_lesion_prop_x_prop_in_dep_mask, by = c("EMPI", "EXAM_DATE")) #n = 2962
+
+#get unique first scan, n = 380
+
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>%
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+### for 3_09
+
+#### map the fiber bundles
+fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
+
+fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)
+
+#0 is not in, 1 is between 0 and 10% mask, 2 is 10% mask, >0 is all
+names_dep_map_3_09_excluding_na_cols <- names_dep_map_3_09[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
+
+#lm
+
+fascicle_lm_unique <- lapply(names_dep_map_3_09_excluding_na_cols, function(x) 
+{
+  lm(substitute(i ~ depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis, na.action=na.omit)
+})
+names(fascicle_lm_unique) <- names_dep_map_3_09_excluding_na_cols
+
+#anova
+fascicle_anova_unique <- lapply(fascicle_lm_unique, anova)
+
+#fdr corrected
+fascicle_anova_unique_fdr <- fdr_anova_generic(fascicle_anova_unique, 1)
+print(fascicle_anova_unique_fdr)#anova values < 0.05, uncorrected
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+#fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+#fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+#print(fascicle_anova_fdr)
+
+#fdr corrected only dep 10% mask
+#fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector ==2)]
+#fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+#print(fascicle_anova_fdr)
+
+#fdr corrected only dep 5% mask
+#fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)]
+#fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+#print(fascicle_anova_fdr)
+
+#uncorrected
+fascicle_anova_unique_p <- sapply(fascicle_anova_unique, function(v) v$"Pr(>F)"[1])
+fascicle_anova_unique_p05_unc <- as.data.frame(fascicle_anova_unique_p[fascicle_anova_unique_p < 0.05 & !is.na(fascicle_anova_unique_p)])
+#rownames(fascicle_anova_unique_p05_unc) <- names_dep_map_3_09_excluding_na_cols[which(fascicle_anova_unique_p < 0.05)]
+print(fascicle_anova_unique_p05_unc)
+
+### for 3_49
+
+#### map the fiber bundles
+fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv")
+
+fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)
+
+#0 is not in, 1 is between 0 and 10% mask, 2 is 10% mask, >0 is all
+names_dep_map_3_49_excluding_na_cols <- names_dep_map_3_49[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]
+
+#lm
+
+fascicle_lm_unique <- lapply(names_dep_map_3_49_excluding_na_cols, function(x) 
+{
+  lm(substitute(i ~ depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis, na.action=na.omit)
+})
+names(fascicle_lm_unique) <- names_dep_map_3_49_excluding_na_cols
+
+#anova
+fascicle_anova_unique <- lapply(fascicle_lm_unique, anova)
+
+#fdr corrected
+fascicle_anova_unique_fdr <- fdr_anova_generic(fascicle_anova_unique, 1)
+print(fascicle_anova_unique_fdr)#anova values < 0.05, uncorrected
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#fdr corrected only dep network
+#fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
+#fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+#print(fascicle_anova_fdr)
+
+#fdr corrected only dep 10% mask
+#fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector ==2)]
+#fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+#print(fascicle_anova_fdr)
+
+#fdr corrected only dep 5% mask
+#fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)]
+#fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+#print(fascicle_anova_fdr)
+
+#uncorrected
+fascicle_anova_unique_p <- sapply(fascicle_anova_unique, function(v) v$"Pr(>F)"[1])
+fascicle_anova_unique_p05_unc <- as.data.frame(fascicle_anova_unique_p[fascicle_anova_unique_p < 0.05 & !is.na(fascicle_anova_unique_p)])
+#rownames(fascicle_anova_unique_p05_unc) <- names_dep_map_3_49_excluding_na_cols[which(fascicle_anova_unique_p < 0.05)]
+print(fascicle_anova_unique_p05_unc)
+
+```

### 2022-07-01 msdepression/scripts/msdepression_functions.R

@@ -862,6 +862,8 @@ get_fascicle_bundle_mapping <- function(){
   # returns vector with association (34 tracts) = 1; cerebellum (7 t) = 2
   # cranial nerves(10) = 3; projection (34) = 4; commissural (2) = 5
   dep_network_grouping <- read.csv(paste0(homedir,"/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), header = T, sep = ",")
+  yeo_7_network <- fascicle_volumes_yeo_7_binarized <- read.csv(paste0(homedir, "results/yeo_7_network_overlap_proportions.txt"), header = T, sep = ",")
+  yeo_7_network_binarized <- fascicle_volumes_yeo_7_binarized <- read.csv(paste0(homedir, "results/yeo_7_network_overlap_proportions_binarized.txt"), header = T, sep = ",")
   numerical_vector <- c(rep(1,34), rep(2,7), rep(3,10), rep(4,34), rep(5,2))
   name_vector <- c(rep("association",34), 
                    rep("cerebellum",7), 
@@ -869,7 +871,90 @@ get_fascicle_bundle_mapping <- function(){
                    rep("projection",34), 
                    rep("commissural",2))
   inDepNetwork_vector <- dep_network_grouping$inDepMask 
-  fascicle_bundle_list <- list(numerical_vector, name_vector, inDepNetwork_vector)
-  names(fascicle_bundle_list) <- c("numerical_vector", "name_vector", "inDepNetwork_vector")
+  fascicle_bundle_list <- list(numerical_vector, 
+                               name_vector, 
+                               inDepNetwork_vector,
+                               yeo_7_network$prop_in_mask_VIS, 
+                               yeo_7_network$prop_in_mask_MOT, 
+                               yeo_7_network$prop_in_mask_DA, 
+                               yeo_7_network$prop_in_mask_VA,
+                               yeo_7_network$prop_in_mask_LIM,
+                               yeo_7_network$prop_in_mask_FP,
+                               yeo_7_network$prop_in_mask_DM,
+                               yeo_7_network_binarized$VIS, 
+                               yeo_7_network_binarized$MOT, 
+                               yeo_7_network_binarized$DA, 
+                               yeo_7_network_binarized$VA,
+                               yeo_7_network_binarized$LIM,
+                               yeo_7_network_binarized$FP,
+                               yeo_7_network_binarized$DM)
+  names(fascicle_bundle_list) <- c("numerical_vector", 
+                                   "name_vector", 
[…]

### 2022-07-20 msdepression/scripts/Descriptives_potential_paper_1_20220622.Rmd
Add files via upload

@@
 just_empi <- subset(df_demo_and_fascicles, select = c("EMPI"))
-added_to_fascicles <- merge(just_empi, fascicle_proportions, by = c("EMPI"))
+added_to_fascicles <- unique(merge(just_empi, fascicle_proportions, by = c("EMPI")))
 melted_df <- melt(added_to_fascicles, id.vars = c("EMPI", "EXAM_DATE"))
@@
 print(histogram_plot_most_affected_fascicles_unique)
+
+#####
+#### look at mean fascicle volume across participants and transpose
+
+# drop exam date and empi, and transpose
+fascicles_alone_df <- t(fascicle_proportions[,!(colnames(fascicle_proportions) %in% c("EMPI", "EXAM_DATE"))])
+fascicle_names <- rownames(fascicles_alone_df)
+
+#take row means
+fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
+names(fascicles_alone_rowMeans) <- c("fascicle", "value")
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bar graphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected") + 
+  geom_bar(stat="identity") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+sorted_ggplot
 

### 2022-07-20 msdepression/scripts/K23_prelim_data_using_dep_T_3_09_mask_20220630.Rmd
Add files via upload

@@
 It is an unthresholded mask (which I think is in Ts per their supplement).
-To use it, I binarized it, and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle.
+To use it, I binarized it, thresholded clusters (3.09) and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle.
 Most fascicles were either entirely overlapping or entirely outside the depression mask.
-We consider the fascicles that had any intersection with the binary depression mask to be within the depression mask, or "indepnet." 47 out of 87 fascicles are in this network.
+We consider the fascicles that had any intersection with the binary depression mask to be within the depression mask, or "indepnet." 62 out of 87 fascicles are in this network.
 We have an even tighter mask "indepnet_10percent", which indicates that at least 10% of a fascicle's volume intersected or overlapped with the depression mask.
-15 are in this network.
+21 are in this network.
 "in_nondep_net" refers to brain regios that completely avoided the depression network (i.e.
 had no shared volume of overlap).
-40 are in this network.
+25 are in this network.
 
@@
 
+fascicle_names_dep_network_15_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.15)] #n=13
+
 #fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
@@
 hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
+
+a<-ggplot(data = fascicle_volumes_dep, aes(x = fascicle, y=prop_in_mask), main = "Proportion of Volume per Fascicle Affected Dep net 3_09") + 
+  geom_bar(stat="identity") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(a)
+
+#sort based on overlap
+fascicle_values_dep_sorted_by_overlap <- fascicle_volumes_dep
+fascicle_values_dep_sorted_by_overlap$prop_in_mask[is.na(fascicle_values_dep_sorted_by_overlap$prop_in_mask)] = 0
+fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask)])
+
+plot_all_fascicles_r_and_l_sep <-ggplot(data = fascicle_values_dep_sorted_by_overlap, aes(x = fascicle, y=prop_in_mask), main = "Proportion of Volume per Fascicle Affected Dep net 3_09") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.4) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(plot_all_fascicles_r_and_l_sep)
+
+#get top quartile
+quartile <- round((dim(fascicle_values_dep_sorted_by_overlap)[1] / 4),0)
+dep_network_names_top_quartile<- (fascicle_values_dep_sorted_by_overlap[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile]
+
+```
+
+```{r mean_R_and_L_dep_network}
+
+#this will make a data frame that contains each of the fascicles with r & l sides, and take a mean. It will then add them to a data frame with the midline structures and make a histogram of them
+
+#some preprocessing because there is one section of the data frame (C_PH_L, C_PHP_L, C_PHP_R and C_PH_R) where the L , L, R, R is out of sync with the rest. 
+fascicle_dep_data_frame_rows_switched <- fascicle_volumes_dep
+
+#make nas into 0
+fascicle_dep_data_frame_rows_switched$prop_in_mask[is.na(fascicle_dep_data_frame_rows_switched$prop_in_mask)] = 0
+
+#reorder by moving C_PH_R up before C_PHP_L, and shifting the others down
+
+c_ph_r_row <- fascicle_dep_data_frame_rows_switched[10,]
+c_php_rows <- fascicle_dep_data_frame_rows_switched[8:9,]
+
+fascicle_dep_data_frame_rows_switched[8,] <- c_ph_r_row
+fascicle_dep_data_frame_rows_switched[9:10,] <- c_php_rows
+
+#get names of fascicles w/r and l
+fascicle_names_w_r_and_l_split <- fascicle_dep_data_frame_rows_switched$fascicle[(grep("_", fascicle_dep_data_frame_rows_switched$fascicle))]
+
+#get names of fascicles that are midline
+fascicle_names_midline <- fascicle_dep_data_frame_rows_switched$fascicle[grep("_", fascicle_dep_data_frame_rows_switched$fascicle, invert=TRUE)]
+
+#separate out right and left, important because for some reason they hae a different order
+fascicle_names_r <- fascicle_names_w_r_and_l_split[(grep("_R$", fascicle_names_w_r_and_l_split))]
+fascicle_names_l <- fascicle_names_w_r_and_l_split[(grep("_L$", fascicle_names_w_r_and_l_split))]
+
+#get the fascicle names of lateral structures, we will use the right side as default
+fascicle_names_of_lateral_structs <- gsub("_R$", "", fascicle_names_r)
+
+#get values of right and left side structs
+fascicle_values_r <- fascicle_dep_data_frame_rows_switched$prop_in_mask[grep("_R$", fascicle_names_w_r_and_l_split)]
+fascicle_values_l <- fascicle_dep_data_frame_rows_switched$prop_in_mask[grep("_L$", fascicle_names_w_r_and_l_split)]
+
+#take mean of r and l side and put in new df
+df_r_and_l_mean <-  (fascicle_values_r + fascicle_values_l)/2
+df_r_and_l <- data.frame(fascicle_names_of_lateral_structs, df_r_and_l_mean)
+names(df_r_and_l) <- c("fascicle", "prop_in_mask")
+
+#pull out midline names and values
+fascicle_values_midline <- fascicle_dep_data_frame_rows_switched$prop_in_mask[grep("_", fascicle_dep_data_frame_rows_switched$fascicle, invert=TRUE)]
+df_midline<- data.frame(fascicle_names_midline, fascicle_values_midline)
+names(df_midline) <- c("fascicle", "prop_in_mask")
+
+#combine data frames
+fascicle_values_dep_not_lateralized <- data.frame(rbind(df_r_and_l, df_midline))
+
+#sort based on overlap
+fascicle_values_dep_sorted_by_overlap <- fascicle_values_dep_not_lateralized
+fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask)])
+
+#fascicle_values_dep_sorted_by_overlap <- fascicle_values_dep_not_lateralized[order(fascicle_values_dep_not_lateralized$prop_in_mask),]
+
+######## Histograms for depression network
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net, not lateralized", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot<-ggplot(data = fascicle_values_dep_sorted_by_overlap, aes(x = fascicle, y=prop_in_mask), main = "Proportion of Volume per Fascicle Affected Dep net 3_09") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.4) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(sorted_ggplot)
+
+##new depression network, top quartile of the 46 -> top 12. This step makes a new df after ordering the previous data frame in reverse order, highest numbers at top, and then takes the top quartile)
+quartile <- round((dim(fascicle_values_dep_sorted_by_overlap)[1] / 4),0)
+dep_network_names_top_quartile_r_and_l_meaned<- (fascicle_values_dep_sorted_by_overlap[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile]
 
@@
 just_dep_and_empi <- subset(data_empi_acc_f_phq, select = c("EMPI", "depGroupVar"))
-added_dep_to_fascicles <- merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI"))
+added_dep_to_fascicles <- unique(merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI")))
+
 melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
@@
 #print(z)
+
+###########################################################################################################
+##### Looking at distribution of disease across fascicles across group and within healthy and depressed ###
+###########################################################################################################
+
+#### look at mean fascicle volume across participants and transpose
+# drop exam date and empi, and transpose
+fascicles_alone_df <- t(added_dep_to_fascicles[,!(colnames(added_dep_to_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
+fascicle_names <- rownames(fascicles_alone_df)
+
+#take row means
+fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
+names(fascicles_alone_rowMeans) <- c("fascicle", "value")
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
+
+#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
+order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(sorted_ggplot)
+
+######### 
+#############
+#### look at mean fascicle volume across healthy participants and transpose
+#############
+
+# drop exam date and empi, and transpose
+healthy_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 1),]
+fascicles_alone_df <- t(healthy_group_fascicles[,!(colnames(healthy_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
+fascicle_names <- rownames(fascicles_alone_df)
+
+#take row means
+fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
+names(fascicles_alone_rowMeans) <- c("fascicle", "value")
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines in Non Depressed", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot_healthy<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected in Non Depressed") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(sorted_ggplot_healthy)
+
+# ordered by mean value overall
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
+sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(sorted_ggplot_healthy_ordered)
+
+###########
+#### look at mean fascicle volume across depressed participants and transpose
+###########
+# drop exam date and empi, and transpose
+depressed_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 2),]
+fascicles_alone_df <- t(depressed_group_fascicles[,!(colnames(depressed_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
+fascicle_names <- rownames(fascicles_alone_df)
+
+#take row means
+fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
+names(fascicles_alone_rowMeans) <- c("fascicle", "value")
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected Depressed", breaks = 20)
+
+sorted_ggplot_depressed<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected Depressed") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(sorted_ggplot_depressed)
+
+# ordered by mean value overall
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
+sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(sorted_ggplot_depressed_ordered)
 ```
@@
            rowMeans(across(fascicle_names_nondep_network))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_15_percent =
+           rowMeans(across(fascicle_names_dep_network_15_percent))) %>%
   group_by(EMPI) %>% 
@@
 total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.053
+
+#look at within depression network in fascicles where at least 15% of the volume of the fascicle was in dep network
+#yes, p=0.054, mean dep = 30%, mean nondep = 25%
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_15percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_15_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_15_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
 
@@
   ylab("Proportion of Fascicles Affected") +
-  ylim(0,0.5) +
+  ylim(0,40) +
   theme(axis.text = element_text(size = 12, colour = "black"))
@@
         panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) #comment out if you want colored graph
 
@@
 
+#fdr corrected only dep 15% mask
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_volumes_dep_network$prop_in_mask >= 0.15)]
+fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
 #uncorrected
@@
 ### tests whether there is higher proportion of vertices lost in depression network versus not
+
+nondep15_net <- fascicle_names[!(fascicle_names %in% fascicle_names_dep_network_15_percent)]
 
@@
            rowMeans(across(fascicle_names_nondep_network))) %>%
+  mutate(total_fascicle_prop_lost_dep_network_15_percent = 
+           rowMeans(across(fascicle_names_dep_network_15_percent))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network_15_percent = 
+           rowMeans(across(nondep15_net))) %>%
   group_by(EMPI) %>% 
@@
 dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 14%, p 2.2 * 10^-16
+
+volume_lost_dep15_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_dep_network_15_percent,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 28%, mean non_dep_net = 14%, p 2.2 * 10^-16
+
+volume_lost_dep15_vs_nondep_network_15_percent_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_dep_network_15_percent,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network_15_percent, var.equal = F) #mean dep 15 net = 28%, mean all fascicles not in 15% net = 26%, p=0.13
 
@@
 
-```{r sexdep_effects_fascicles_unique}
+```{r sexdep_effects_fascicles_unique_age_cov}
 

### 2022-07-29 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
Add files via upload

---
title: "MS Depression Paper"
author: "Erica Baller"
date: '2022-07-28'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
```

## MS Depression

Welcome to the MS Depression Data, installment 1 analysis markdown! This project was designed to ask a relatively simple question - are patients with MS who are depressed more likely to have lesions in fascicles that connect areas previously associated with depression. The literature are sparse in this domain. It is very clear that depression occurs in up to 50% of patients with MS, and some studies have suggested that it relates to lesion burden. But to my knowledge there are no papers that evaluate how lesions disrupt white matter tracts that serve as the backbone for functional networks. Over the course of the next few chunks, I'll walk through our analysis. It combines imaging data from research grade scans acquired at 3T as part of a clinical protocol (T1 and T2 Flair), and clinical data from the electronic medical record.

## R Markdown

Part 1: 
The first chunk here is reading in appropriate data frames and preprocessing. Two big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller.

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

Inclusion criteria below:

Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
  1) Scanned under MS protocol (42K)
  2) MS Diagnosis (17K)
  3) Seen by MS doctor (16K)

Inclusion criteria for Depressed Group
  1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder)
        - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition
  2) Screened positive for depression on PHQ2 or PHQ9 at any point
  3) Prescribed antidepressant medications per NAMI website
        - https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

Inclusion criteria for Healthy Group
  1) No F* diagnoses
  2) At least 1 PHQ2 or PHQ9 with a documented 0
  3) Free of psychiatric medications per NAMI website

Exclusions
  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

Imaging exclusion
  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.

Final N demographics 
  Depressed: 859 (232 unique)
  Super clean healthy: 604 (148 unique)

(unique, with good MIMoSA):

                         Healthy          Depressed             p      test
  n                                148            232                     
  Race (%)         Caucasian       108 ( 73.0)    173 ( 74.6)   0.821     
                   Non-caucasian    40 ( 27.0)     59 ( 25.4)             
  Sex (%)          Female          117 ( 79.1)    199 ( 85.8)   0.117     
                   Male             31 ( 20.9)     33 ( 14.2)             
  Age (mean (SD))                46.57 (12.66)  48.75 (11.75)   0.089     
  Depression (%)   1               148 (100.0)      0 (  0.0)  <0.001     
                   2                 0 (  0.0)    232 (100.0)             
  PHQ2 (mean (SD))                0.00 (0.00)    0.58 (1.53)   <0.001     
  PHQ9 (mean (SD))                0.00 (NA)     10.43 (9.02)       NA

Networks

  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1). It is an unthresholded mask (he initially sent one where the unit is r, but then sent us one where unit is T). 
  Per Siddiqi's article: The location of each lesion or brain stimulation site (Fig. 2a–c, top panels) was mapped to an underlying brain circuit using a large normative connectome database (n=1000) and previously validated methods (Fig. 2a–c, bottom panels)5. The normative connectome was used to estimate connectivity of each lesion or stimulation site to every voxel in the brain. At each voxel, a Pearson r-value was computed for the correlation between depression score and lesion or stimulation site connectivity to that voxel (Fig. 2a–c, right panels), yielding a population-derived “circuit map” for each of the 14 datasets (Supplementary Figure 1). We generated a combined depression circuit map by taking the mean of all 14 circuit maps, weighted by the sample size of each dataset (Fig. 5a). Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted)

  To use it, I thresholded the clusters (3.09), binarized it and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle. Most fascicles were either entirely overlapping or entirely outside the depression mask, though the range of overlap was anywhere from 1-30%. This required us to make some choices. We considered an option where the "depression mask" was anywhere that ever overlapped, and non-depression were areas where there was no overlap. We also looked at quartiles, and settled on the top 25% (top quartile), i.e. the top 25% of fascicles with the highest proportion of network overlap were considered in the depression network. Everything outside of that was considered "non_depression" network, or nondep_net. Cranial nerves are too small to reliably be assess with our method, and so they were removed from the analysis.In total, 77 fascicles were evaluated.

Depression top quartile, r&l separated
 [1] "SLF3_R"  "AF_R"    "AF_L"    "SLF2_R"  "CBT_R"   "FAT_R"   "CS_P_R"  "C_PHP_R" "MdLF_L"  "FAT_L"   "SLF2_L"  "MdLF_R"  "CBT_L"   "TR_P_L"  "TR_S_R" 
[16] "CPT_O_R" "TR_P_R"  "PAT_L"   "PAT_R"   "EMC_R"   "TR_S_L"  "CC"

Notebook Overview

Analyses
  There are two main types of analyses I performed, t.tests (for example, to compare inside versus outside depression network) and linear models (for example, looking at the relationship of depression to each individual fascicle). With respect to T-tests, I assume unequal variance for all. For linear models, I look only at the first scan. The general pattern is to look at the total proportion of volume lost across fascicles, then within depression network, and within non depression network. Lastly, I look at overall mean lesion burden irrespective of location by counting the number of nonzero voxels in each individual's mimosa mask.

Results
  Depression group: 232 with depression, 148 healthy

```{r read_in_df_and_subset}
homedir <- "/Users/eballer/BBL/msdepression/"

data <- read.csv(paste0(homedir, "/data/drugs_data/parsable_msdepression.csv"), sep = ",", header = TRUE) #n=16830, only has people with MS ICD10 code G35 and seen by MS provider

columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <- data %>% 
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(EMPI = as.factor(EMPI)) %>%
   mutate(On.Psych.Meds = ifelse(On.Psych.Meds == "True", 1, 0)) %>%
   mutate(On.Antidepressants = ifelse(On.Antidepressants == "True", 1, 0)) %>%
   mutate(Has.PHQ2 = ifelse(Has.PHQ2 == "True", 1, 0)) %>%
   mutate(Has.PHQ9 = ifelse(Has.PHQ9 == "True", 1, 0)) %>%
   mutate(Has.depdx = ifelse(grepl("F3", ICD10), 1, 0)) %>%
   mutate(PHQ.2_modsev_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 >= 3), 1, 0)) %>%
   mutate(PHQ.9_modsev_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 >= 10), 1, 0)) %>%
   mutate(PHQ.2_mild_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 > 0 & PHQ.2 < 3), 1, 0)) %>%
   mutate(PHQ.9_mild_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 > 0 & PHQ.9 < 10), 1, 0)) %>%
   mutate(PHQ.2_zero = ifelse((!is.na(PHQ.2) & PHQ.2 == 0), 1, 0)) %>%
   mutate(PHQ.9_zero = ifelse((!is.na(PHQ.9) & PHQ.9 == 0), 1, 0)) %>%
   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
   rowwise(ACCESSION_NUM) %>%
   mutate(depGroupVar =
            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
   ungroup()

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
#write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

########
#fascicle volumes in depression network
########

fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
  replace(is.na(.), 0) %>%
  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))

#Write out the new network groupings to a file, will be used later
fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
write.table(fascicle_volumes_dep_network, file=fascicle_names_dep_network_grouping, sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

#fascicle_names_nondep_network<- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 0)]

#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask > 0)]
#fascicle_names_dep_network_all_no_cranial_nerves <- fascicle_names_dep_network_all[grep("CN", fascicle_names_dep_network_all, invert = TRUE)]

#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all_3_09.csv", row.names = F, quote = F, col.names = F)

#fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent_3_09.csv", row.names = F, quote = F, col.names = F)

#fascicle_names_dep_network_5_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)] #n=24

#fascicle_names_dep_network_15_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.15)] #n=13

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether any overlap with depression network (anything above 1 is overlap), as well as whether the fascicle overlaps, and the proportion of overlap with yeo networks
fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)

##########################################################
#             Read in Lesion Volumes                     #
##########################################################

# contains the volume of lesioned voxels, i.e. a sum of all the areas of each individual's binary mask that are 1

lesion_volumes <- read.csv(paste0(homedir, "results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE)

##########################################################
#             Merge with data_empi              #
##########################################################
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962
df_demo_and_fascicles <- merge(df_demo_and_fascicles, lesion_volumes, by = c("EMPI", "EXAM_DATE")) #n = 2962
```

```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment

make_demographics_table_ms(data_frame = data_empi_acc_f_phq)

df_unique_empi <- data_empi_acc_f_phq %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

make_demographics_table_ms(data_frame = df_unique_empi)

df_just_healthy_and_depressed <- df_unique_empi %>% filter(depGroupVar != 0)

make_demographics_table_ms(data_frame = df_just_healthy_and_depressed)

#just people with good mimosa
df_good_mimosa <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
make_demographics_table_ms(data_frame = df_good_mimosa)

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)
```
```{r compare_overall_volume_of_lesions_dep_vs_non_dep}

#separate out only people in depression group, and take first instance
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

dep_vs_healthy_t_test_lesion_volume <-  t.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #p=0.07

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
#data$Fascicle <- pretty_names

# Grouped
barplot_lesion_volume <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
  ylab("Volume of Lesion") +
  theme(title = element_text(size=14),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
print(barplot_lesion_volume)

```

```{r visualize_depression_network}
##########################################################
#                 Histos                        #
##########################################################

######## Histograms for depression network
#sort based on overlap
fascicle_values_dep_sorted_by_overlap <- fascicle_volumes_dep
fascicle_values_dep_sorted_by_overlap$prop_in_mask[is.na(fascicle_values_dep_sorted_by_overlap$prop_in_mask)] = 0
fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask)])

#remove fascicles with cranial nerves
fascicles_r_and_l_sep_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

plot_all_fascicles_r_and_l_sep_no_cranial_nerves <-ggplot(data = fascicles_r_and_l_sep_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask))+
  ggtitle("Proportion of Volume Overlap per Fascicle 3_09") + 
  geom_bar(stat="identity") +
  xlab("Fascicle") +
  ylab("Overlap with Depression Mask") +
  ylim(0,0.4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_no_cranial_nerves)

### remove all rows with cranial nerves in them
# of note, we do not have all cranial nerves represented, we only have, CNII (optic), III(oculomotor), V (trigeminal), VII(facial), VIII(vestibulocochlear)
fascicle_values_dep_no_cranial_nerves = fascicle_values_dep_sorted_by_overlap[(grep(pattern = "CN", x = fascicle_values_dep_sorted_by_overlap$fascicle, invert = TRUE)),]

#figure out how many fascicles that is, then 
quartile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 4),0)

#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile] #n=19

#get a network that includes all of the fascicles NOT in dep 25% network
dep_network_names_bottom_3_quartiles_no_cranial_nerves <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)] #n=68

#plot values within the depression network, and values outside of it

fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
  geom_bar(stat="identity") +
  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, top quartile") + 
  ylim(0,0.4) +
  ylab("Proportion in Depression Mask") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)

#bottom 3 quartiles
fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
  geom_bar(stat="identity") +
  ylab("Proportion in Depression Mask") +
  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
  ylim(0,0.4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs

```

```{r proof_of_concept}

#this section creates histograms looking at the proportion of volume lost in each fascicle across subjects
#we recapitulate earlier work and show that higher volume loss with our method reflects regions previously known to have higher volume loss
#These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis

data_empi_acc_f_phq_unique <- data_empi_acc_f_phq %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#have to pull out just the pieces I'm interested in
just_dep_and_empi <- subset(data_empi_acc_f_phq, select = c("EMPI", "depGroupVar"))
added_dep_to_fascicles <- unique(merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI")))

melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))

#exclude regions where there was no lesion at all
lesioned <- subset(melted_df, value > 0)

#only include people in healthy or depressed group
lesioned_dx <- subset(lesioned, depGroupVar != 0)

#separate out healthy and depressed
healthy <- subset(lesioned, depGroupVar == 1)
depressed <- subset(lesioned, depGroupVar == 2)

q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

#print(q)
#print(r)
print(r_color)
#print(x)
#print(y)
#print(z)

###########################################################################################################
##### Looking at distribution of disease across fascicles across group and within healthy and depressed ###
###########################################################################################################

#### look at mean fascicle volume across participants and transpose
# drop exam date and empi, and transpose
fascicles_alone_df <- t(added_dep_to_fascicles[,!(colnames(added_dep_to_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

#take row means
fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
names(fascicles_alone_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans

#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected") + 
  ylab("Proportion") + 
  geom_bar(stat="identity") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(sorted_ggplot)

######### 
#############
#### look at mean fascicle volume across healthy participants and transpose
#############

# drop exam date and empi, and transpose
healthy_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 1),]
fascicles_alone_df <- t(healthy_group_fascicles[,!(colnames(healthy_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

#take row means
fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
names(fascicles_alone_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
  geom_bar(stat="identity") +
  ylab("Proportion") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(sorted_ggplot_healthy_ordered)

###########
#### look at mean fascicle volume across depressed participants and transpose
###########
# drop exam date and empi, and transpose
depressed_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 2),]
fascicles_alone_df <- t(depressed_group_fascicles[,!(colnames(depressed_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

#take row means
fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
names(fascicles_alone_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
  ylab("Proportion Affected") +
  geom_bar(stat="identity") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(sorted_ggplot_depressed_ordered)

##### look at mean fascicles affected in dep vs non dep network
######
#plot values within the depression network, and values outside of it

fascicles_dep_25_percent_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Dep net, top quartile") + 
  geom_bar(stat="identity") +
  ylab("Proportion") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_top_quartile)

#bottom 3 quartiles
fascicles_dep_bottom_3_quartiles_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected, bottom three quartiles") + 
  ylab("Proportion") +
  geom_bar(stat="identity") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)

```

```{r depression_vs_healthy_overall_proportion_of_fascicles_affected}

#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#test whether proportion lost overall different between depressed and healthy- No, p = 0.050, Dep = 31%, Non dep = 27%
total_prop_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054

#look at within depression network using top quartiles, yes p = 0.047, Dep = 32%, non dep = 26%
total_prop_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)

#look at bottom 75%, no p = 0.057, dep = 30%, non dep = 27%
total_prop_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057

############
#bar plot
############

#over all fascicles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean Proportion of Fascicles Affected") +
  ggtitle(paste0("Total Proportion Affected by dx, p=", round(total_prop_lost_dep_vs_healthy_unique$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph

print(barplot_overall)

#in top quartile

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean proportion of Fascicles Affected") +
    ggtitle(paste0("Total Propotion Affected by dx in Dep 25%, p=", round(total_prop_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) 
print(barplot_top_quartile)

#in bottom 3 quartiles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean Proportion of Fascicles Affected") +
    ggtitle(paste0("Total Proportion Affected by dx, Bottom 75% p=", round(total_prop_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52))

print(barplot_bottom_3_quartiles)

```

```{r depression_vs_healthy_overall_volume_of_fascicles_affected}

#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#read in fiber volume values
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#test whether volortion lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054

#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)

#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057

############
#bar plot
############

#over all fascicles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
  ggtitle(paste0("Total Volume Affected by dx, p=", round(total_volume_lost_dep_vs_healthy_unique$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) #comment out if you want colored graph

print(barplot_overall)

#in top quartile

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx in Dep 25%, p=", round(total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) 
print(barplot_top_quartile)

#in bottom 3 quartiles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx, Bottom 75% p=", round(total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000))

print(barplot_bottom_3_quartiles)

```

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop}
# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups

##############################
## FASCICLES TO BE INCLUDED ##
##############################

#define the depression network (all areas that intersect with depression network, no cranial nerves)
fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77

#######################
### depression map  ###
#######################

### define depression map
fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#filter to exclude fascicles w/cranial nerves
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
  replace(is.na(.), 0) %>%
  filter(fascicle%in% fascicle_names_no_cranial_nerves)

#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column

df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))

#######################
### real data x dx  ###
#######################

#then, make the data frame
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{
  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
  eval(parse(text=t_test_command))
})
names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves

#pull t-statistic
dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)

#convert to data frame
df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
row.names(df_t_values) <- fascicle_names_no_cranial_nerves

#pull p-values
dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)

#convert to data frame
df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves

#FDR correct p-values
dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")

#Convert to data frame
df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves

#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_overlap_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_t_values$dep_vs_healthy_t_values)
names(df_fascicle_dep_mask_overlap_t_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_t")

#make copy to sort
df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_t_value

df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") + 
  geom_bar(stat="identity") +
  ylim(0,5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_t_sorted_by_dep_overlap)

# plot fascicle overlap x t value
#statistic, p = 0.010
dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_t_value)
print(summary(dep_overlap_x_dep_vs_healthy_t))

scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_t)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Proportion of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis calculated in proportions (T)") + 
  ggtitle(paste0("Fascicle/depression mask Proportion overlap by dx T, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_t_val_by_dep_net_overlap)

```

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume}

#define the depression network (all areas that intersect with depression network, no cranial nerves)
fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77

#######################
### depression map  ###
#######################

### define depression map
fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#filter to exclude fascicles w/cranial nerves
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
  replace(is.na(.), 0) %>%
  filter(fascicle%in% fascicle_names_no_cranial_nerves)

df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))

########################
## Get Volume Values ###
########################

#read in fiber volume values
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

#isolate out only the depressed group and healthy group, fascicles should be in volumes rather than proportions
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{
  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
  eval(parse(text=t_test_command))
})
names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves

#pull t-statistic
dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)

#convert to data frame
df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
row.names(df_t_values) <- fascicle_names_no_cranial_nerves

#pull p-values
dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)

#convert to data frame
df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves

#FDR correct p-values
dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")

#Convert to data frame
df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves

#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_volume_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
names(df_fascicle_dep_mask_volume_t_value) <- c("fascicle", "dep_vs_healthy_t", "volume_in_dep_mask")

df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask <- df_fascicle_dep_mask_volume_t_value

#order by volume in depression mask
df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$volume_in_dep_mask)])

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by volume of dep network") + 
  geom_bar(stat="identity") +
  ylim(0,5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_t_sorted_by_dep_overlap)

#statistic volume, p = 0.031
dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value)
print(summary(dep_overlap_x_dep_vs_healthy_t))

scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_t_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis (T)") + 
  ggtitle(paste0("Fascicle/depression mask volume by dx T, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_t_val_by_dep_net_overlap)

```

```{r bootstrapping_proportion_and_volume_x_dep_t_analyses}

##### proportion
# 95% confidence intervals for the correlation
set.seed(1)
tval <- function(formula, data, indices) {
  d <- data[indices,] # allows boot to select sample
  fit <- lm(formula, data=d)
  return(summary(fit)$coefficients[2,3])
}

data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_t_value, select = c("depression_mask_overlap","dep_vs_healthy_t"))
results_prop <- boot(data=data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap, statistic=tval,
   R=10000, formula=dep_vs_healthy_t~depression_mask_overlap)
print(results_prop) #actual 2.62, CI 0.701, 4.553)
plot(results_prop)
boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="bca", index=1)
print(boot_results_prop.ci)

#manually calculating P from CI, from interwebs
est=results_prop$t0
l=boot_results_prop.ci$bca[4]
u=boot_results_prop.ci$bca[5]

#1) SE=(u-l)/2*1.96
se=((u-l)/(2*1.96))
#2) z=Est/SE
z=est/se
#3) p=exp(-0.717*z - 0.416*z^2)
p=exp((-0.717*z) - (0.416*(z*z)))
print(p)

##### volume
# 95% confidence intervals for the correlation
set.seed(1)
data_frame_dep_v_nondep_t_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_t_value, select = c("dep_vs_healthy_t", "volume_in_dep_mask"))

results_vol <- boot(data=data_frame_dep_v_nondep_t_x_vol_dep_mask, statistic=tval,
   R=10000, formula=dep_vs_healthy_t~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="bca", index=1)
print(boot_results_vol.ci)

#get p from CI
#manually calculating P from CI, from interwebs
est=results_vol$t0
l=boot_results_vol.ci$bca[4]
u=boot_results_vol.ci$bca[5]

#1) SE=(u-l)/2*1.96
se=((u-l)/(2*1.96))
#2) z=Est/SE
z=est/se
#3) p=exp(-0.717*z - 0.416*z^2)
p=exp((-0.717*z) - (0.416*(z*z)))
print(p)

```

### 2022-08-01 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
Add files via upload

@@
 Results
-  Depression group: 232 with depression, 148 healthy
+  Depression group: 232 with depression, 148 healthy:
+
+  A) Lesion volume between depressed and controls
+
+    - P=0.067, NS
+
+  B) Depressed vs Controls Proportions
+
+    - Overall dep vs non-dep, p = 0.050 (NS), Dep = 31%, Non dep = 27%
+    - Dep vs non, top 25%, **p = 0.047**, Dep = 32%, non dep = 26% 
+    - Dep vs non, bottom 75%, no p = 0.057, dep = 30%, non dep = 27%
+
+  C) Depressed vs Controls Volume
+
+    - Overall dep vs nondep: **p = 0.045**, Dep = 9382, Non dep = 8260
+    - Dep vs nondep in top 25% yes **p = 0.034**, Dep = 4829, non dep = 3914
+    - Dep vs nondep in bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
+
+  D) T values in fascicle associated with Overlap in Depression map
+
+    - T values correlated w/proportion of overlap of fascicle in depression map
+        - p of lm (0.010),
+        - 95%   ( 0.701,  4.553 )
+        - boostrap p calc from CI - **0.008**
+
+    - T values correlated w/volume of overlap with depression network (i.e. # voxels that fascicle shares w/depression map),
+        - p = 0.031
+        - 95%   ( 1.052,  3.724 )
+        - bootstrap p calc from CI- **0.001**
+
+I think it works best when sticking with *volumes* rather than proportions
 
@@
 
-#fascicle_names_nondep_network<- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 0)]
-
-#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask > 0)]
-#fascicle_names_dep_network_all_no_cranial_nerves <- fascicle_names_dep_network_all[grep("CN", fascicle_names_dep_network_all, invert = TRUE)]
-
-#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
-#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all_3_09.csv", row.names = F, quote = F, col.names = F)
-
-#fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
-#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent_3_09.csv", row.names = F, quote = F, col.names = F)
-
-#fascicle_names_dep_network_5_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)] #n=24
-
-#fascicle_names_dep_network_15_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.15)] #n=13
-
 #fascicle_mapping, returns numerical mapping as well as names of tracts, and whether any overlap with depression network (anything above 1 is overlap), as well as whether the fascicle overlaps, and the proportion of overlap with yeo networks
@@
 ##########################################################
+#merge whole data set with fascicle proportions
 df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962
+
+#merge with the df that has lesion volumes, 1 per perosn
 df_demo_and_fascicles <- merge(df_demo_and_fascicles, lesion_volumes, by = c("EMPI", "EXAM_DATE")) #n = 2962
@@
 data <- data.frame(Diagnosis,value, sem)
-#data$Fascicle <- pretty_names
 
@@
 #  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
-print(barplot_lesion_volume)
+print(barplot_lesion_volume) #p=0.067
 
@@
 
-#test whether volortion lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
+#test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
 total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
@@
 p=exp((-0.717*z) - (0.416*(z*z)))
-print(p)
+print(p) #p=0.008
 
@@
 p=exp((-0.717*z) - (0.416*(z*z)))
-print(p)
+print(p) # #p=0.001
 

### 2022-08-01 lab_notebook/lab_notebook.Rmd
through july 2022

@@
 
-20211203
+#### 20211203
 So something cool is happening here, but unexpected.
@@
 [] figure out why registration broke - not sure
+
+#### 20220621
+Such a long lag!
+Back to now trying to figure out how to do spin tests with ms data.
+This was raised to address the observation that people, overall, are more likely to have disease in the depression network, but the depression network is so big that we want to see how specific this finding it.
+Thus the spin test.
+
+Currently, the idea is as follows:
+
+Surface to volume -> spinning -> volume to surface:
+
+[x] 1) Take depression voxel mask Depression_Clust_mask_roi_binarized.nii.gz (done locally, /Users/eballer/BBL/msdepression/templates/harvard_depression)
+
+    - of note, Depression_Clust_mask_roi_binarized.nii.gz was generated from a mask made by Matt (Depression_Clust_mask_rois.nii.gz)
+    - command from 3dinfo: 3dclust -1Dformat -nosum -1dindex 0 -1tindex 0 -1noneg -2thresh -0.1317 0.1317 -dxyz=1 -savemask Clust_mask 1.01 20 /Users/mcieslak/Desktop/AllFX_wmean.nii
+
+```{bash}
+3dcalc -a Depression_Clust_mask_roi_binarized.nii.gz -expr 'ispositive(a)*ispositive(x)' -prefix Depression_Clust_mask_roi_binarized_l.nii
+3dcalc -a Depression_Clust_mask_roi_binarized.nii.gz -expr 'ispositive(a)*isnegative(x)' -prefix Depression_Clust_mask_roi_binarized_r.nii
+```
+
+    - output: L & R hemisphere volume images
+        - Depression_Clust_mask_roi_binarized_l.nii
+        - Depression_Clust_mask_roi_binarized_r.nii
+
+[] 2) Project to vertex space using vol2surf from imco, fs5 space
+
+```{bash}
+scp Depression_Clust_mask_roi_binarized_*.nii eballer@transfer.pmacs.upenn.edu:/project/msdepression/templates/harvard_depression/.
+cp /project/imco/from_chead/coupling_test_eb_20200918/code/vol2surf* /project/msdepression/scripts/. 
+cp /project/imco/from_chead/coupling_test_eb_20200918/code/transformMatrix.R /project/msdepression/scripts/.
+```
+    - need to register the volume to some template (hcp) and use that I think
+OLD
+```{bash}
+lta_convert --infsl ${boldDir}/${bblid}_${datexscanid}_seq2struct.mat \
+	--outreg ${targDir}/${bblid}/${datexscanid}/${bblid}_${datexscanid}_seq2struct.dat \
+	--subject ${bblid}/${datexscanid} \
+	--src ${boldDir}/${bblid}_${datexscanid}_referenceVolume.nii.gz \
+	--trg ${boldDir}/${bblid}_${datexscanid}_targetVolume.nii.gz \
+	>> ${targDir}/${bblid}/${datexscanid}/${bblid}_${datexscanid}_projection.log 2>&1
+```
+
+NEW (believe I use diff flags), think about using ants_registration_code.sh on pmacs
+```{bash}
+lta_convert --infsl ${boldDir}/${bblid}_${datexscanid}_seq2struct.mat \
+	--outreg ${targDir}/${bblid}/${datexscanid}/${bblid}_${datexscanid}_seq2struct.dat \
+	--subject ${bblid}/${datexscanid} \
+	--src ${boldDir}/${bblid}_${datexscanid}_referenceVolume.nii.gz \
+	--trg ${boldDir}/${bblid}_${datexscanid}_targetVolume.nii.gz \
+	>> ${targDir}/${bblid}/${datexscanid}/${bblid}_${datexscanid}_projection.log 2>&1
+```
+
+    - output: L & R hemisphere vertex images, all binarized
+[] 3) Spin 1000 times per hemisphere
+    - output 1000 L & R hemisphere surfaces
+[] 4) Surface back to voxel
+    - output 1000 L & R hemisphere volumes
+[] 5) Merge L & R volumes together
+    - output 1000 spun depression network volumes
+
+[] 6) Calculate volume of fascicles within each spun depression mask  
+    - have to wrap script: calc_vol_fascicles_within_dep_mask.sh, which outputs a spreadsheet with last column being proportion of each of 87 fascicles within depression mask
+    - output table (column is proportion of fascicle in dep network, row is fascicle:
+                 Spin 1 prop   Spin 2 prop   Spin 3 prop   Spin 4 prop ... Spin 1000 prop
+    Fascicle 1    0.21
+    Fascicle 2    0
+    Fascicle 3    0.33
+    Fascicle 4    0.1
+    ...
+    Fascicle 87   0
+
+[] 7) Take above spreadsheet, make binarized
+    - Will assume that if the fascicle has ANY overlap depression network, ie spin prop > 0, will be in dep net
+    - output table (column is proportion of fascicle in dep network, row is fascicle:
+
+                 Spin 1 prop   Spin 2 prop   Spin 3 prop   Spin 4 prop ... Spin 1000 prop
+    Fascicle 1    1
+    Fascicle 2    0
+    Fascicle 3    1
+    Fascicle 4    1
+    ...
+    Fascicle 87   0
+
+[] 8) Enrichment/calculations
+    - Now, for each spin, will be calculating the dep vs nondep enrichment.
+    - Two ways to do this
+        1) Means (i.e calculate mean of depression overlap[proportion] in real data and each spin, and mean of non-dep overlap)
+          - output
+
+                                 Spin 1    Spin 2    Spin 3    Spin 4 ... Spin 1000
+          Dep Mean % overlap
+          Nondep Mean % overlap
+
+        2) T tests - two sample T at each spin
+          - output
+                                Spin 1    Spin 2    Spin 3    Spin 4 ... Spin 1000
+          Dep vs nondep T       3.2       1.2       4.5       1.1
+
+[] 9) Sort, calculate top 5%
+    a) Means: Do separately for depression and non-depression/kinda like an enrichment
+    b) Ts: Just do 1 time, sort
+
+    Will try both, put Ts first as more straightforward
+
+-----------
+#### 20220622 
+Put a hold on surface projections above as I need people to come back from HBM. Instead, working on descriptives for potential paper 1
+
+Rmarkdown: /Users/eballer/BBL/msdepression/scripts/Descriptive_potential_paper_1_20220622.Rmd
+
+Redid overall tracts, age, sex, agexsex unique/mixed models.
+
+#### 20220623
+Downloaded Yeo volume masks: https://surfer.nmr.mgh.harvard.edu/fswiki/CorticalParcellation_Yeo2011
+
+Separated out each functional network
+```{bash afni_commands_to_construct_yeo_indiv_vol_masks.sh}
+#!/bin/sh
+
+# pre: Yeo cortical parcellations (7 network) in volumetric space: https://surfer.nmr.mgh.harvard.edu/fswiki/CorticalParcellation_Yeo2011
+# post: 7 individual network maps: VIS, MOT, DA, VA, LIM, FP, DM, labeled yeo_7_vol_[network abbreviation]
+# uses: This script takes the parcellations in volume (Yeo2011_7Networks_MNI152_FreeSurferConformed1mm_LiberalMask.nii) and makes 7 individual maps.
+In the Yeo2011 volume mask, each network is given a number (1-7), with 1 corresponding to VIS, and 7 corresponding to DM.
+In order to make individual maps, I need to separate out each network in the map based on intensity.
+Will try to code in the following way: 1) make a vector where each item is the name of the network 2) make a loop that counts the networks 3) go through each network and do math to pull out only the network with a particular intensity.
+It uses the command "amongst", which gives back the voxels where the intensity is equivalent to the counter, multiplies it times the whole mask (so now every voxel contains the intensity of interest, everything else is 0), and then binarizes.
+# dependencies: Afni
+
+#make list of yeo network abbreviations
+yeo_7_net_list=("VIS"
+                "MOT"
+                "DA"
+                "VA"
+                "LIM"
+                "FP"
+                "DM")
+
+#network path
+network_path=/Users/eballer/BBL/msdepression/templates/Yeo_JNeurophysiol11_MNI152/
+
+#current working directory
+cwd=`pwd`
+echo $cwd
+
+#set volume mask [change this if I ever want to use a diff mask
+volume_mask=Yeo2011_7Networks_MNI152_FreeSurferConformed1mm_LiberalMask.nii
+
+#assign generic network_prefix_list variable; this is in case I decide I want to do this for 17 networks
+network_prefix=yeo_7_net
+network_list=$(echo ${network_prefix}_list)
+
+echo "network list = ${network_list}"
+
+#set counter, this will be used in loop
+network_counter=1
+
+echo "Volume mask = ${volume_mask}"
+
+echo "Deleting old masks"
+rm -f ${network_path}/${network_prefix}_*.nii
+
+echo "Making individual network maps"
+#for loop; goes through each network, pulls out the mask that corresponds to that network number, and stores it with the appropriate prefix
+
+#have to write out yeo_7_net_list because bash can't handle too many extractions. Prob because I'm not referencing an object but a name when pointing network list to yeo_7_net_list
+for network in "${yeo_7_net_list[@]}"; do
+        echo "making ${network} mask, counter = ${network_counter}"
+
+#make binary masks of each level of the masks; amongst just takes only the values given in the next segment, which can be 1 or more
+        afni_command=("3dcalc -a ${network_path}/${volume_mask} -expr 'ispositive(a*amongst(a,${network_counter}))' -prefix ${network_path}/${network_prefix}_${network}.nii")
+        echo $afni_command
+        eval $afni_command
+        ((network_counter++))
+done
+
+#go back to current working directory
+#cd $cwd
+```
+
+#### 20220624
+[x] Move masks from local to pmacs, log onto pmacs, etc
+```{bash}
+scp yeo_7_net_* eballer@transfer.pmacs.upenn.edu:/project/msdepression/templates/Yeo_JNeurophysiol11_MNI152/.
+ssh -Y eballer@scisub.pmacs.upenn.edu
+cd /project/msdepression/scripts
+xbash
+module load afni[tab complete]
+```
+
+[x] Map each of the fascicles to the corresponding functional network (i.e. use as ROI like I did with depression)
+[x] make list of which fascicles go in which network
+```{bash calc_vol_fascicles_within_yeo_7_net.sh}
+#### calculate volume of fascicles within the yeo 7 networks ####
+### Pre: All streamlines must have been made in dsi studio (these are the nii files)
+### Post: Yeo7 network mask resampled to template space, file <streamline_volume_within_${network}_network>, containing each fascicle and its corresponding % within yeo network.
+Additionally, will have a file that contains all of the overlaps per each network, and another with a binarized (yes/no)
+### Uses: Trying to figure out which fascicles are in the yeo 7 network for dimensionality reduction for use in later regressions (i.e.
+decrease # comparisons by only looking at fascicles within yeo 7 networks)
+### Dependencies: This script requires afni to be installed (currently using 20.1)
+
+#output directory
+outdir="/project/msdepression/results/"
+
+#assemble yeo 7 networks
+
+yeo_7_net_list=("VIS"
+                "MOT"
+                "DA"
+                "VA"
+                "LIM"
+                "FP"
+                "DM")
+
+#set up output file names
+all_net_overlap_proportions="${outdir}/yeo_7_network_overlap_proportions.txt"
+all_net_overlap_proportions_binarized="${outdir}/yeo_7_network_overlap_proportions_binarized.txt"
+
+#set counter to keep track of whether it is the first iteration (and you need to do a special column add), or just take the last line and pr it
+counter=1
+for network in "${yeo_7_net_list[@]}"; do
+	#set outfile
+	outfile="/project/msdepression/results/streamline_volume_within_${network}_network.csv"
+
+	#set template directories, will be making our resampled mask from within this
+	template="yeo_7_net_${network}.nii"
+	template_dir="/project/msdepression/templates/Yeo_JNeurophysiol11_MNI152/"
+	resampled_filename="resampled_${template}"
+	template_fullpath="${template_dir}/${template}"
+
+	#set fascicle directory
+	fascicle_direc="/project/msdepression/templates/dti/HCP_YA1065_tractography/"
+	fascicle_type="association"
+	fascicle_for_resampling="AF_L.nii.gz"
+
+	#set home directory
+	homedir="/project/msdepression/scripts/"
+
+	#initiate csvs
+	rm -f $outfile
+	touch $outfile
+	echo "fascicle,num_voxels_total,non_zero_voxels_in_${network}_map,prop_in_mask_${network}" >> $outfile
+
+	#resample yeo_7_net_${network}.nii to match sample fascicle - AF_L.nii.gz
+	if [ -f "${template_dir}/${resampled_filename}" ]; then
+		rm -f ${template_dir}/${resampled_filename}
+	fi
+	3dresample -master ${fascicle_direc}/${fascicle_type}/${fascicle_for_resampling} -prefix ${template_dir}/${resampled_filename} -input ${template_fullpath} -rmode NN
+
+	#go through each fiber, get volume, and store it in csv
+	fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"
+	for fiber_bundle_type in ${fiber_bundle_type_list}; do
+	    echo "Fiber bundle type is " $fiber_bundle_type
+	    fascicle_volumes=$(ls ${fascicle_direc}/${fiber_bundle_type}/*.nii* | grep -v "_mask") #list all files, exclude ones with mask in them
+	    for fascicle in ${fascicle_volumes}; do
+		    fascicle_prefix=$(echo $fascicle | perl -pe 's/.*\/(.*).nii.gz/$1/g')
+		    num_voxels_whole_mask=$(3dmaskave -quiet -mask SELF -sum ${fascicle})
+
+		    #remove previous mask if already made
+		    if [ -f "${fascicle_direc}/${fascicle_type}/${fascicle_prefix}x${network}_mask.nii.gz" ]; then
+			    rm -f ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}x${network}_mask.nii.gz
+		    fi
+
+		    3dcalc -a ${template_dir}/${resampled_filename} -b ${fascicle} -expr 'a*b' -prefix ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}x${network}_mask.nii.gz
+		    num_nonzero_volumes=$(3dmaskave -quiet -mask SELF -sum ${fascicle_direc}/${fascicle_type}/${fascicle_prefix}x${network}_mask.nii.gz)
+		    prop_in_mask=$(echo ${num_nonzero_volumes}/${num_voxels_whole_mask} | bc -l)
+		    echo "num_voxels" $num_voxels_whole_mask "num_nonzero" $num_nonzero "prop_in_mask" $prop_in_mask
+		    echo "${fascicle_prefix},${num_voxels_whole_mask},${num_nonzero_volumes},${prop_in_mask}" >> $outfile
+	    done
+	done
+
+	#add column to output file
+	if [ ${counter} == 1 ]; then
+		#remove the file if it exists
+		rm -rf ${all_net_overlap_proportions}
+		#take the 1st and 4th columns and write new file
+		cat ${outfile} | cut -d ',' -f 1,4 > ${all_net_overlap_proportions} 
+	else
+		#paste the last column into the building table
+		cat $outfile | cut -d ',' -f 4 > to_add.txt
+		pr -mts, ${all_net_overlap_proportions} to_add.txt > temp.txt
+
+		#remove to_add file
+		rm to_add.txt
+
+		#move temp to the file, so everything is up to date
+		mv temp.txt ${all_net_overlap_proportions}
+
+	fi
+	((counter++))
+done
+
+#binarize the all_net_overlap_proportions file, takes every value and puts a 0 - this takes the ,, cases and makes them 0s. And then it takes any decimals (meaning some overlap, makes them 1s)
+cat ${all_net_overlap_proportions} | perl -pe 's/,/,0/g' | perl -pe 's/0\.\d+/1/g' > ${all_net_overlap_proportions_binarized}
+```
+
+#### 20220628
+[x] run analyses looking at depression, age, sex, within fp
+```{bash}
+sh #doesn't work with zsh
+scp eballer@transfer.pmacs.upenn.edu:/project/msdepression/results/yeo* /Users/eballer/BBL/msdepression/results/.
+```
+
+check out /Users/eballer/msdepression/scripts/yeo_7_network_analyses_20220628
+
+#### 20220629
+realized that my depression network analyses were using the clusterized, binarized mask that was thresholded (not the general map) that is so big and gross.
+Will need to ask matt how he picked the range of -0.1317 to 0.1317.
+From Shan, this range is in R's.
+With a sample size of 713, the corresponding T value is 3.50 (or p<0.0005 uncorrected)
+
+ Meeting with Ted
+
+3dclust -1Dformat -nosum -1dindex 0 -1tindex 0 -1noneg -2thresh -0.1317 0.1317 -dxyz=1 -savemask Clust_mask 1.01 20 /Users/mcieslak/Desktop/AllFX_wmean.nii.
+For my analyses, I used: Depression_Clust_mask_roi_binarized.nii.gz, which is this mask copied.
+Give an underlay, check if cerebellum .
+1) overlap % with binarized dep network (3.09) , which is redo the depression network stuff but use different brain map.
+2) average t value for each fascicle  3) average t value for overlap zone of fascicle w/depression mask .
+4) Possibly weight it by the % lesion affecting fascicle by depression T (so this is a depression dysconnectome score.
+this is how much of the fascicle is affected multiplied by the T value of the overlap depression/fascicle area).
+Brain Smash – Spin test
+
+- gonna remake the depression mask at T = 3.09 (p <0.001); r = 0.115 (with df 713)
+
+```{bash}
+
+#clusterize and binarize in 1 step
+3dclust -1Dformat -nosum -1dindex 0 -1tindex 0 -1noneg -2thresh  -3.09 3.09 -dxyz=1 -binary -savemask Dep_clust_T_3_09_binarized.nii 1.01 20 /Users/eballer/BBL/msdepression/templates/harvard_depression/DepressionCircuit_t.nii
+
+#nonbinarized
+3dclust -1Dformat -nosum -1dindex 0 -1tindex 0 -1noneg -2thresh  -3.09 3.09 -dxyz=1 -savemask Dep_clust_T_3_09.nii 1.01 20 /Users/eballer/BBL/msdepression/templates/harvard_depression/DepressionCircuit_t.nii
+
+#calculate a mask that is the T values within the clusters within the depression mask
+3dcalc -a Dep_clust_T_3_09_binarized.nii -b DepressionCircuit_t.nii -expr 'a*b' -prefix Dep_clust_T_3_09xDep_circuit_T.nii
+
+#copy to pmacs
+scp Dep_clust_T_3_09* eballer@transfer.pmacs.upenn.edu:/project/msdepression/templates/harvard_depression/.
+scp DepressionCircuit_t.nii eballer@transfer.pmacs.upenn.edu:/project/msdepression/templates/harvard_depression/.
+
+#go onto pmacs
+ssh -Y eballer@scisub.pmacs.upenn.edu
+
+#go into scripts directory
+cd /project/msdepression/scripts/
+
+#copy 
+cp calc_vol_fascicles_within_depression_mask.sh calc_vol_fascicles_within_depression_mask_3_09.sh
+
+#change line in calc_vol_fascicles_within_depression_mask_3_09.sh
+vim calc_vol_fascicles_within_depression_mask_3_09.sh
+
+# copy results back local
+scp eballer@transfer.pmacs.upenn.edu:/project/msdepression/results/streamline_volume_within_dep_network_3_09.csv /Users/eballer/BBL/msdepression/results/.
+```
+
+#### 20220630
+ [x] do analyses with new masks, see how it compares. More values in there, results largely unchanged
+
+#### 20220701-20
+Whoops, forgot to record!
+Trying to address a few questions - we will use the 3.09 mask going forward.
+There are clearly differential proportions of fascicles within the depression network.
+So I sorted them by % overlap, it looks generally distributed, plan to take top quartile and consider that to be the depression mask.
+
+Then the question becomes, what is the "nondepression" mask.
+Is it anything not in depression mask?
+Is it areas where there is NO overlap?
+Either way, there is much higher disease in areas within the depression mask versus outside of it.
+
+Then the question is, the cranial nerves basically have no overlap. Is this because they are poorly sampled by mimosa/brain scan? 
+If they are well sampled, positives for including them - you can say that disease is more likely in cortical versus cranial nerve tracts.
+Given that MS affects any part of the CNS, the fact that there is some relative sparing of the cranial nerves could be meaningful
+Cons: If they are not well-sampled, it is useless. Also, if the depression map never went down that far, then it would be hard to say to include them or not. 
+Also, there would be no way for the depression map to include them, since the depression map only includes things with bold signal, and cranial nerves aren't actually connecting brain to brain.
+
+- of note, cerebellum is in mask, and it is sampling inferiorly enough to grab brainstem (the overlays I have aren't lining up great, but it does look like mask lines up with brainstem)
+
+- had a quick panic attack when I saw that our depression circuit on my local machine didn't line up with anything I had locally.
+But I double checked.
+On pmacs, in script: calc_vol_fascicles_within_depression_mask.sh, there is a line where I resample the depression mask to be in the same space as the AF_L.nii, which lines up with MNI152_T1_1mm mask (/Users/eballer/BBL/msdepression/templates/dti/mni_icbm152_t1_nlin_asym_09a/mni_icbm152_t1_tal_nlin_asym_09a.nii).
+There is some crap outside the brain for this, should disappear when multiplied with mimosa mask.
+Similarly, the fascicles themselves (i.e.
+the volume in AF_L look great and line up perfectly).
+I don't think this harvard depression map is the best, to be honest.
+Could also be that the resampling would be better not w/NN but something else.
+
+[x] sensitivity analysis looking just at brain regions within cerebrum (i.e. drop cranial nerves)
+
+  Results:
+
+  When comparing lesion burden in depressed vs healthy
+    1) Overall lesion burden, findings no longer significant
+
+    2) Lesion burden within the 25% mask, patients with depression have MORE disease (32%) than healthy patients (27%) within this network (p=0.047)
+
+    3) Lesion burden within nondep network, depressed patients have more disease 23% vs 20%  p = 0.049
+
+    4) Lesion burden within bottom 75%, inverse relationship, depressed patients have less disease (23%) than healthies (27%) p = 0.023
+       - could be in line w/the fox work that suggests that disease in non-dep network could be protective
+
+  When looking at network enrichment
+    1) in the absence of the cranial nerves, more disease in areas than interact w/depression network vs places that never had any overlap  
+      - #mean dep_net = 31%, mean non_dep_net =  22%, p 8.58 x 10^-12
+
+    2) in the absence of the cranial nerves, more disease in the top 25% than in areas that never intersected with depression map 
+      - #mean dep net = 30%, mean non_dep_net = 22%, p 7.3* 10^-8
+
+    3) in the absense of cranial nerves, the difference between the top 25% and the bottom 75% is no longer significant
+      - #mean dep net = 30%, mean bottom 3 quartiles = 29, p = 0.52
+
+some notes: If I exclude cranial nerves up front, the "depression" network will have 19 regions (rather than 22)
+[x] message matt about what to do with depression network, inclined to leave it as is right now, but in the future, probably use something else
+
+using round- lm (dep vs healthy p 0.049, but t test between two groups doesn't work)
+
+[x] come up with EXACT verbiage on how to describe depression mask versus lesion masks
+  - 1) First, we have to define what the fascicles are considered to connect the depression mask
+  - We start with a functional mask from harvard. It was generated using lesion network mapping.
+  - 14 datasets were used, including stroke, dbs (parkinson's), and tms datasets
+  - Per article: The location of each lesion or brain stimulation site (Fig.
+2a–c, top panels) was mapped to an underlying brain circuit using a large normative connectome database (n=1000) and previously validated methods (Fig.
+2a–c, bottom panels)5.
+The normative connectome was used to estimate connectivity of each lesion or stimulation site to every voxel in the brain.
+At each voxel, a Pearson r-value was computed for the correlation between depression score and lesion or stimulation site connectivity to that voxel (Fig.
+2a–c, right panels), yielding a population-derived “circuit map” for each of the 14 datasets (Supplementary Figure 1).
+We generated a combined depression circuit map by taking the mean of all 14 circuit maps, weighted by the sample size of each dataset (Fig.
+5a).
+Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted)
+
+  - We take that mask, threshold it at 3.09 (p=0.001, uncorrected), and binarize it. 
+  - We then go through each of the 87 fascicles, and calculate what proportion of each fascicle overlaps with the depression network, scaled by the overall volume of the fascicle
+  - We then rank these fascicles by their overlap
+  - We consider the final fascicle "mask" to be the top 25% of fascicles (i.e. the top quartile according to proportion of fascicle affected)
+  - These fascicles (22 total), are then called the depression "mask"
+  - after this step, the proportion of overlap is no longer integrated into analysis, it is solely used to rank the fascicles to define the network.
+
+  -2) Having defined which fascicles are associated with the depression mask, we can now look subject data in individual fascicles, as well as comparing fascicles within the WM depression network versus outside of it.
+
+#### 20220722
+
+[] make sure graphs have value for y axis
+
+[] speak with matt 
+  - about mask in general, seems crappy  
+     - it is, but it is what we have
+  [x] email Shan, mni 2009b
+  [x] Yes to exclude cranial nerves
+[o] how about nondep mask looking at 25% low end OR 3.09 in opposite direction?
+ - looking at bottom quartile, no cranial nerves - no overall diffs in depression/healthies, but more disease w/in dep network in depressed, no differences in disease in bottom 75% or bottom quartile
+ - findings of more disease in depression network than outside in MS go away. But honestly, I'm okay with this. I want to use the no cranial nerves mask, can say it is due to poor coverage?
+ {} will instead making 2 graphs, T values per fascicle (bar graph), ranked by overlap with depression map
+ {} plot, x axis overlap w/depression map, y axis t value of group difference
+
+[x] another potential avenue, ranking each of the clusters based on their mean t value(essentially don't binarize map), and using that to multiply w/overlap proportion - crap,
+
+[x] make some pictures of the actual mask coverage.
+
+[] message bart re: mixed model w/gam
+
+[x] Make a plot of overlap of each tract w/depression network on x axis (so each fiber gets a space, 77 total), y axis is the t value of depression vs control w/in that fascicle.
+Try to fit a line, see how it looks.
+Might be an easier way than randomly thresholding.
+
+#### 20220726-27
+[x] redid these analyses instead looking at how much volume was affected in each fascicle (not correcting for size of the fascicle). 
+   - Results
+   - Overall fascicle volume affected greater in depressed vs nondep
+   - Driven by more fascicle loss w/in fibers that connect depression network
+   - No differences between diagnostic groups in bottom 25th percentil or bottom 75%
+   - For each fascicle, the higher its overlap w/the depression network, the bigger the the patient > control T test t value is
+
+#### 20220728
+[x] Going to now just compare overall volume of the lesions (irrespective of fascicles that run through) for each subject
+  - Ideal finding would be that the overall volume of lesions isn't different between groups, but that it is only when you look at streamlines affected that you get the money. 
+  - Alternatively, if the lesion volume also discriminates, that is cool, too.
+Way to test effect size?
+Whether they can be viewed together?
+Wonder how it correlates with symptoms (i.e.
+fascicle view helps correlate with dimensions of psychopathology)
+ - after making the file, I'll read it in to r, merge with the big spreadsheet, and then just do the patients vs controls analysis
+
+```{bash}
+ssh -Y eballer@scisub.pmacs.upenn.edu
+cd /project/msdepression/scripts
+xbash
+module load afni_openmp/20.1 
+sh get_volume_of_mimosa_lesions.sh 
+#results written to /project/msdepression/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes
+exit pmacs
+cd /Users/eballer/BBL/msdepression/results
+scp eballer@transfer.pmacs.upenn.edu:/project/msdepression/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes .
+```
+
+```{bash get_volume_of_mimosa_lesions.sh }
+#!/bin/bash
+
+#########     Welcome to  get_volume_of_mimosa_lesions.sh !  ##########3
+## The purpose of this script is to extract the overall volume of lesion burden for each mimosa mask. 
+## The goal is to use the output to answer the question: Does overall lesion burden (quantified by volume of le
+sions), irrespective of location, relate to depression diagnosis?
+
+### Pre: File containing full paths to the mimosa binary masks:  /project/msdepression/data/melissa_martin_file
+s/csv/mimosa_binary_masks_hcp_space_20211026_n2336
+### Post: File containing ACCESSION_NUM, EXAM_DATA, volume_of_mimosa_lesions
+### Uses: 	1) Reads in the file with full paths
+###		2) Extracts the accession number and exam date
+###		3) Performs 3dmaskave -quiet -mask SELF -sum ${binary_map} and writes to a file
+### Dependencies: 1) Requires afni, so you need to xbash, and module load afni<tab complete>
+
+### Opening Script: If nothing is entered, let person know we are using default mimosa path file, otherwise use
+ file on command line
+
+echo "Greetings and welcome to the get_volume_of_mimosa_lesions script"
+echo "This script takes a file with full paths to mimosa binary maps, and returns a file that has the accession
+ number, exam data, and overall volume of lesions in that map"
+#set default path
+default_mimosa_path="/project/msdepression/data/melissa_martin_files/csv/mimosa_binary_masks_hcp_space_20211026
+_n2336"
+
+echo "Default mimosa path $default_mimosa_path"
+#set output directory for the volumes file
+output_direc="/project/msdepression/results/"
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa files path" $default_mimosa_path
+    mimosa_files_path=$default_mimosa_path
+else  
+    mimosa_files_path=$1
+fi
+
+echo "Mimosa files path" $mimosa_files_path
+#### initiate new file
+
+#append "_volumes" to the end of the path
+suffix=$(echo ${mimosa_files_path} | perl -pe s'/.*\/(.*)/$1/g')
+output_csv=$output_direc/$suffix"_volumes"
+
+#get rid of old file if it exists, write new one and put in headers
+rm -f $output_csv
+touch $output_csv
+echo "ACCESSION_NUM,EXAM_DATE,volume_of_mimosa_lesions" >> $output_csv
+
+#loop through each file in the mimosa_file_path, extract accession/exam date, calculate volume, write to file
+mimosa_files=$(cat ${mimosa_files_path})
+for mimosa_file in $mimosa_files; do
+    	echo "Mimosa file is" $mimosa_file
+
+	#pulls out accession number and date, separated them by a comma
+    	subj_sess=$(echo ${mimosa_file} | perl -pe s'/.*sub-(.*)\/ses-(.*)\/r.*/$1,$2/g')
+
+	#gets volume of binary mask
+	volume=$(3dmaskave -quiet -mask SELF -sum ${mimosa_file})
+
+	#writes to output file
+	echo "${subj_sess},$volume" >> $output_csv
+done
+
+```
+
+```{r bootstrap}
+
+##### proportion
+# 95% confidence intervals for the correlation
+set.seed(1)
+tval <- function(formula, data, indices) {
+  d <- data[indices,] # allows boot to select sample
+  fit <- lm(formula, data=d)
+  return(summary(fit)$coefficients[2,3])
+}
+
+data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_t_value, select = c("depression_mask_overlap","dep_vs_healthy_t"))
+results_prop <- boot(data=data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap, statistic=tval,
+   R=10000, formula=dep_vs_healthy_t~depression_mask_overlap)
+print(results_prop) #actual 2.62, CI 0.701, 4.553)
+plot(results_prop)
+boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="bca", index=1)
+print(boot_results_prop.ci)
+
+#manually calculating P from CI, from interwebs
+est=results_prop$t0
+l=boot_results_prop.ci$bca[4]
+u=boot_results_prop.ci$bca[5]
+
+#1) SE=(u-l)/2*1.96
+se=((u-l)/(2*1.96))
+#2) z=Est/SE
+z=est/se
+#3) p=exp(-0.717*z - 0.416*z^2)
+p=exp((-0.717*z) - (0.416*(z*z)))
+print(p) #p=0.008
+
+##### volume
+# 95% confidence intervals for the correlation
+set.seed(1)
+data_frame_dep_v_nondep_t_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_t_value, select = c("dep_vs_healthy_t", "volume_in_dep_mask"))
+
+results_vol <- boot(data=data_frame_dep_v_nondep_t_x_vol_dep_mask, statistic=tval,
+   R=10000, formula=dep_vs_healthy_t~volume_in_dep_mask)
+print(results_vol)
+plot(results_vol)
+boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="bca", index=1)
+print(boot_results_vol.ci)
+
+#get p from CI
+#manually calculating P from CI, from interwebs
+est=results_vol$t0
+l=boot_results_vol.ci$bca[4]
+u=boot_results_vol.ci$bca[5]
+
+#1) SE=(u-l)/2*1.96
+se=((u-l)/(2*1.96))
+#2) z=Est/SE
+z=est/se
+#3) p=exp(-0.717*z - 0.416*z^2)
+p=exp((-0.717*z) - (0.416*(z*z)))
+print(p) #p=0.001
+
+```
+
+    Take home:
+
+Paper figures
+1) Demographics/how samples were calculated
+2) Method
+3) method for calculating depression map
+4) No difference in overall lesion volume between the groups (violin, the figure with SEM makes it look like there is a difference)
+5) Overall lesion breakdown across all groups, then divided by depressed & healthy (maybe, maybe not)
+6) If you look at the volume of the streamlines affected, Depressed have more volume loss than controls
+7) And specifically, that loss difference is driven by loss in depression network (areas with top percentile of overlap with depression network drive it)
+8) The more overlap a fascicle has (in either volume of proportion of streamline) with the depression network, the higher the T value of the diagnostic differences (this could also be two part figure, with 5 above)
+
+[x] taki wants me to do bootstrap to get CIs for the t value ~ overlap with depression network analysis. Boot in R can do this
+    Take data, resample subjects, w/ replacement, refit, do 1000 times or 10,000 times and confidence interval
+    https://machinelearningmastery.com/a-gentle-introduction-to-the-bootstrap-method/
+    https://www.statmethods.net/advstats/bootstrapping.html

### 2022-08-01 msdepression/scripts/yeo_7_network_analyses_20220628.Rmd
Yeo networks

--- 2022-03-22 K23_prelim_data_started_20220225.Rmd
@@
 ---
-title: "MS Prelim Data K23"
+title: "MS Yeo 7 Network Analyses"
 author: "Erica Baller"
-date: "2/25/2022"
+date: "6/28/2022"
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---
@@
 
-## MS preliminary data
-
-Ideas:
-
-Overall:
-
-[] demographics - tables made below
-
-[] imaging parameters
-
-Aim 1: lesion network mapping
-[] some picture showing the # of people who are scanned and the # of people with depression. Maybe an adaptation of Syndey's picture somehow. 
-[] mimosa picture
-[] picture of filtering pipeline
-[] comparison of some of the histograms for high volume loss similar to prev literature(CN II, slf), not so much in other areas
-[] results of age analysis [maybe/maybe not, to prove the method works]?
-[] results of disease in vs outside of depression network (corrected), lead with this?
-[] results total proportion lost (single p)
-[] results of dep vs controls across all regions,(uncorrected)
-[] results of dep vs controls in depression network (uncorrected)
-
-[] prelim network enrichment via spin-testing or just citation?
-- do I want to validate here or somewhere else with neurosynth maps?
-
-Aim 2: prospective
-[] power analysis
-[] CNB (Bart has great Figure 3)
-[] Goassess
-
-Aim 3
-[] Hydra picture
-[] hydra results from my paper
-[] acquisition protocol (T1, t2, flair, dwi)
+## MS yeo network analyses
+
+In doing my preliminary analyses for my K23, I became interested in how networks other than the depression network relate to MS.
+I specifically wondered about the classic yeo 7 networks, especially the frontoparietal network given previous depression literature as well as my own work.
+Furthermore, I have noticed that there are a lot of interesting sex differences, age differences, and depression differences all by fascicle, but they don't reach correction, suggesting that there might be benefit of dimensionality reduction.
+Thus, enter this new project.
+
+The idea here is that I took the yeo volume map (the more liberal one: (**/Users/eballer/BBL/msdepression/templates/Yeo_JNeurophysiol11_MNI152/Yeo2011_7Networks_MNI152_FreeSurferConformed1mm_LiberalMask.nii.gz**), and separated it out into individual networks on my local machine (**/Users/eballer/BBL/msdepression/scripts/afni_commands_to_construct_yeo_indiv_vol_masks.sh**), and binarized them.
+I then moved them to pmacs and treated them as "lesions" to get the proportion of fascicle overlap with each of the 87 fascicles (**/project/msdepression/scripts/calc_vol_fascicles_within_yeo_7_net.sh**).
+The results, stored here (**/project/msdepression/results/yeo_7_network_overlap_proportions.txt**), go through each of 87 fascicles and identify the proportion of overlap of that fascicle's volume with each network.
+For example, if SLF2_L has a value of .27 in the FP network, it indicates that 27% of the fibers in the SLF2_L fascicle overlap in volume with the frontoparietal network volume mask.
+In the result (**/project/msdepression/results/yeo_7_network_overlap_proportions_binarized.txt**), results are binarized - if there is ANY overlap (i.e.
+if there is at least one voxel of overlap between the fascicle and the network, the value is 1, otherwise the value is 0).
+This method is cursory, I can also imagine a situation when I scale based on the proportion of overlap.
+It is also important to know that certain fibers overlap with multiple networks; this is NOT exclusionary.
+Certain fibers are more integrated.
+Many ways to go about thinking about this.
+
+In the FIRST analysis here, I am just going to look at the frontoparietal network in the simplest condition - look at fascicles with a 1 in the FP column.
+The idea would be to look at overall effects, age effects, sex differences, depression diagnoses in fibers within and outside the FP network.
+I.e., comparing rows with a 1 for FP to a 0 for FP.
+Once I work this out, will repeat for other networks.
 
@@
 
-  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1).
-It is an unthresholded mask (which I think is in Ts per their supplement).
-To use it, I binarized it, and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle.
-Most fascicles were either entirely overlapping or entirely outside the depression mask.
-We consider the fascicles that had any intersection with the binary depression mask to be within the depression mask, or "indepnet." 47 out of 87 fascicles are in this network.
-We have an even tighter mask "indepnet_10percent", which indicates that at least 10% of a fascicle's volume intersected or overlapped with the depression mask.
-15 are in this network.
-"in_nondep_net" refers to brain regios that completely avoided the depression network (i.e.
-had no shared volume of overlap).
-40 are in this network.
+  The networks are generated from the Yeo 7 networks (as above).
 
@@
 [12] "CBT_R"   "CPT_O_R" "CS_P_R"  "TR_P_R"
+
+FP Network (n = 54)
+[1] "AF_L"    "AF_R"    "C_FPH_L" "C_FPH_R" "C_FP_L"  "C_FP_R"  "C_PHP_L" "C_PHP_R" "C_R_L"   "C_R_R"   "EMC_L"  
+[12] "EMC_R"   "FAT_L"   "FAT_R"   "IFOF_L"  "IFOF_R"  "ILF_L"   "ILF_R"   "MdLF_L"  "PAT_L"   "PAT_R"   "SLF1_L" 
+[23] "SLF1_R"  "SLF2_L"  "SLF2_R"  "SLF3_L"  "SLF3_R"  "UF_L"    "UF_R"    "CBT_L"   "CBT_R"   "CPT_F_L" "CPT_F_R"
+[34] "CPT_P_L" "CPT_P_R" "CS_A_L"  "CS_A_R"  "CS_P_L"  "CS_P_R"  "CS_S_L"  "CS_S_R"  "CST_R"   "DRTT_L"  "DRTT_R" 
+[45] "RST_L"   "RST_R"   "TR_A_L"  "TR_A_R"  "TR_P_L"  "TR_P_R"  "TR_S_L"  "TR_S_R"  "AC"      "CC"
+
+NonFP Network(n=33)
+ [1] "C_PH_L"   "C_PH_R"   "MdLF_R"   "VOF_L"    "VOF_R"    "CB_L"     "CB_R"     "ICP_L"    "ICP_R"    "MCP"     
+[11] "SCP"      "V"        "CNIII_L"  "CNIII_R"  "CNII_L"   "CNII_R"   "CNVIII_L" "CNVIII_R" "CNVII_L"  "CNVII_R" 
+[21] "CNV_L"    "CNV_R"    "AR_L"     "AR_R"     "CPT_O_L"  "CPT_O_R"  "CST_L"    "F_L"      "F_R"      "ML_L"    
+[31] "ML_R"     "OR_L"     "OR_R"
+
+FP_network_10percent (n = 16)
+ [1] "AF_L"    "AF_R"    "C_FPH_R" "C_FP_R"  "C_PHP_R" "C_R_L"   "C_R_R"   "FAT_R"   "SLF2_L"  "SLF2_R"  "SLF3_R" 
+[12] "CPT_F_R" "CS_A_L"  "RST_R"   "TR_A_L"  "TR_A_R"
+
+nonFP_network_10percent (n = 71)
+[1] "C_FPH_L"  "C_FP_L"   "C_PH_L"   "C_PHP_L"  "C_PH_R"   "EMC_L"    "EMC_R"    "FAT_L"    "IFOF_L"   "IFOF_R"  
+[11] "ILF_L"    "ILF_R"    "MdLF_L"   "MdLF_R"   "PAT_L"    "PAT_R"    "SLF1_L"   "SLF1_R"   "SLF3_L"   "UF_L"    
+[21] "UF_R"     "VOF_L"    "VOF_R"    "CB_L"     "CB_R"     "ICP_L"    "ICP_R"    "MCP"      "SCP"      "V"       
+[31] "CNIII_L"  "CNIII_R"  "CNII_L"   "CNII_R"   "CNVIII_L" "CNVIII_R" "CNVII_L"  "CNVII_R"  "CNV_L"    "CNV_R"   
+[41] "AR_L"     "AR_R"     "CBT_L"    "CBT_R"    "CPT_F_L"  "CPT_O_L"  "CPT_O_R"  "CPT_P_L"  "CPT_P_R"  "CS_A_R"  
+[51] "CS_P_L"   "CS_P_R"   "CS_S_L"   "CS_S_R"   "CST_L"    "CST_R"    "DRTT_L"   "DRTT_R"   "F_L"      "F_R"     
+[61] "ML_L"     "ML_R"     "OR_L"     "OR_R"     "RST_L"    "TR_P_L"   "TR_P_R"   "TR_S_L"   "TR_S_R"   "AC"      
+[71] "CC"
 
@@
     *Age effects:
-      -Age mean 47 (range 20-83). 
-      -Almost everything corrects (54 out of 87 fascicles, 25 in association network)
-      -Take home - as expected, the older you are, the worse white matter fascicle invovlement you have
-
+     FP network 54 fascicles, 48 correct
+     Non fp network 33 fascicle, 7 correct
     *Sex effects:
-      - In the final good MIMoSA group, there were 64 unique males and 316 unique females, 83% women
-      - Individual fascicle results underwhelming, only 3 regions meet P<0.05 (unc), SCP, CPT_O_R, OR_R
-      - no association or within dep network findings
-      - Sex effects interaction model (fascicle ~ depGroupVar*osex + PAT_AGE_AT_EXAM)
-        - Dep effects - same as above, 12 fascicles (p<0.05 unc)
-        - Sex effects - same as above, with the edition of CNVIII_R (p<0.05 unc)
-        - age*sex interactions - SLF1_L and CNVIII_R (p<0.05 unc)
-      - Analyses to consider at a later date
-        [] some sort of matching
-
-    *Depression over all fascicles:
-      - The exciting news of the day! With our new depression group (moving the meds people), we have findings
-      - total proportion lost - p = 0.049, depressed (27%) more loss than healthy(24%)
-      - total proportion lost within depression network (47 total) - trend (0.059) between depressed (35%) and healthy(30%)
-      - total proportion lost within depression 10% network (15 fascicles total) - p = 0.049 (depressed (33%) have more loss than healthy (27%))
-      - total proportion lost within depression 5% network (24 fascicles total) -  p = 0.056 (depressed (32%) have more loss than healthy (27%))
-      - total proportion lost outside depression network (40 fascicles total)- p = 0.037 (depressed (18%) more than healthy(16%))
-      - take home - people with a lifetime history of depression have more disease than nondepressed individuals. This finding is not specific to the depression network, but is overall.
-
-    *MS disease and depression network overlap:
-      - Ms disease within depression network (35%) versus outside (17%), p < 2.2* 10^-16
-      - MS disease within depression 10% network (31%) versus outside (17%), p < 2.2 * 10^-16
-      - MS disease within depression 5% network (30%) versus outside (17%), p < 2.2 * 10^-16
-      - Linear Regressions (essentially matching t tests above): 
-         +depressed v healthy within dep network in (p = 0.061); 
-         +dep vs healthy within dep 10% (p = 0.049);
-         +dep vs healthy within non dep network (p = 0.037)
-      - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis.
-Suggests a final possible common pathway?
-Shared biology?
-This may also in part explain why rates of depression in MS are much higher than rates of depression in other autoimmune conditions.
-    *Depression by individual fascicle:
-      - Nothing corrects at this time, unfortunately. 
-      - P<0.05, uncorrected findings (* indicates association network):
-          -AF_R	0.01168289	*		
-          -FAT_R	0.01828607 *			
-          -SLF2_R	0.01901378 *		
-          -VOF_R	0.04741569 *			
-          -CPT_F_R	0.02887163 			
-          -CPT_O_L	0.02238344			
-          -CPT_P_R	0.02053104			
-          -CS_S_R	0.03783654			
-          -DRTT_R	0.00695066			
-          -ML_R	0.02511932	
+     - no fdr corrected values
+
+    *MS disease and FP network overlap:
+    - w/in FP network by diagnosis, p = 0.63
+    - w/in FP_10% by diagnosis, p = 0.64
+    - nonFP, p = 0.27
+    - non FP 10%, 0.49
+
+    *Depression by individual FP fascicles:
+
     Fascicle loss x PHQ9:
-     - Nothing significant by individual fascicle, within association network, by fascicle within the depression network
-     - similarly, no significant findings looks at means of all fascicle loss, within dep network, 10% and 5% and non-depression network
-
-Limitations/Future directions
-
-  All diagnostic info is obtained from EMR fields, not notes, nor direct chart review.
-We do not know if patients are relapsing/remiting, nor do we know their active depression status.
-Through careful review, though patients sometimes have many phq2s listed, they are always identical.
-It is also possible that people who are on antidepressants are prescribed them not for depression.
-I tried a sensitivity analysis and removed people on meds from heatlhy group but didn't include them in the depression group.
-Unique n in depression group went from 232->115 and all findings go away.
-
-  In order to do this, you need to change the mutate command:
-
-##------
-#goal is to keep people who were seen by
-
-#### add this line #####
-   mutate(dep_by_dx_phq_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq | true_healthy, 1, 0)) %>%
-   rowwise(ACCESSION_NUM) %>%
-
-#### remove this line or comment out
-    # mutate(depGroupVar = sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq))) %>%
-
-#### add this line
-   mutate(depGroupVar = sum(c(dep_by_dx_phq_healthy_phq0_no_psych_meds,dep_by_dx_phq))) %>%
-   ungroup()
-### -----
-
-  It would be great if we also had better dimensional scores.
-The PHQ2s are largely 0s, and the lack of spread makes interpretation really challenging.
-Though the PHQ9s definitely have better spread, there just aren't many of them.
-Doing really good cognitive and clinical phenotyping, or even getting a different or better set of measures into the MS clinic would be ideal (Aim 2).
-
-  With respect to networks, we have used the "depression" one from Harvard because I was inspired by the lesion network mapping work.
-However, the mask is pretty broad, covering and strongly overlapping with the executive network.
-There is bias toward having regions that functionally connect with the L DLPFC b/c it is a TMS target.
-We are considering neurosynth as a possible place to look at different depression masks.
-
-  Interestingly, a key topic at the ACTRIMS 2022 conference was psychiatric comorbidity.
-Though we are focusing on depression, it sounds like anxiety is just as bad if not worse or more highly comorbid with MS.
-Would be worth considering not only a DSM breakdown but dimensionally looking at internalizing disorders for the K.
-Also more in line with rdoc but would require changing aims a bit.
-
-  Future analyses: age x dep, sex x dep
 
@@
 
-write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
+#write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
 
 ########
-#fascicle volumes in depression network
+#fascicle volumes in yeo networks
 ########
-fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), header = T, sep = ",")
+
+#read in file with proportions
+fascicle_volumes_yeo_7 <- read.csv(paste0(homedir, "results/yeo_7_network_overlap_proportions.txt"), header = T, sep = ",", na.strings="0")
+
+#make titles nice
+names(fascicle_volumes_yeo_7) <-gsub("prop_in_mask_", "prop_", names(fascicle_volumes_yeo_7))
+
+#make Nas into 0
+fascicle_volumes_yeo_7[is.na(fascicle_volumes_yeo_7)]=0
+
+#read in file with binarized yes/no 
+fascicle_volumes_yeo_7_binarized <- read.csv(paste0(homedir, "results/yeo_7_network_overlap_proportions_binarized.txt"), header = T, sep = ",")
 
 #add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
-fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
-  replace(is.na(.), 0) %>%
-  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))
-
-write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
-
-fascicle_names_nondep_network<- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 0)]
-
-fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask > 0)]
+#fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
+ # replace(is.na(.), 0) %>%
+#  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))
+
+#write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
+
+fascicle_names_FP_network<- fascicle_volumes_yeo_7_binarized$fascicle[which(fascicle_volumes_yeo_7_binarized$FP == 1)]
+
+fascicle_names_nonFP_network<- fascicle_volumes_yeo_7_binarized$fascicle[which(fascicle_volumes_yeo_7_binarized$FP == 0)]
+
+fascicle_names_FP_network_10percent<- fascicle_volumes_yeo_7$fascicle[which(fascicle_volumes_yeo_7$prop_FP > 0.1)]
+
+fascicle_names_nonFP_network_10percent<- fascicle_volumes_yeo_7$fascicle[which(fascicle_volumes_yeo_7$prop_FP <=0.1)]
 
@@
 
-fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
+#fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
 #write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)
 
-fascicle_names_dep_network_5_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)] #n=24
+#fascicle_names_dep_network_5_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)] #n=24
 
 #fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
-fascicle_bundle_mapping <- get_fascicle_bundle_mapping()
+#fascicle_bundle_mapping <- get_fascicle_bundle_mapping()
 
@@
 ######## Histograms for depression network
-#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
-hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20)
-
+
+hist(x=fascicle_volumes_yeo_7$prop_VIS, main = "Histogram Affected Streamlines in VIS Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20, xlim=c(0,0.7), ylim=c(0,70))
+
+hist(x=fascicle_volumes_yeo_7$prop_MOT, main = "Histogram Affected Streamlines in MOT Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20, xlim=c(0,0.7), ylim=c(0,70))
+
+hist(x=fascicle_volumes_yeo_7$prop_DA, main = "Histogram Affected Streamlines in DA Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20, xlim=c(0,0.7), ylim=c(0,70))
+
+hist(x=fascicle_volumes_yeo_7$prop_VA, main = "Histogram Affected Streamlines in VA Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20, xlim=c(0,0.7), ylim=c(0,70))
+
+hist(x=fascicle_volumes_yeo_7$prop_LIM, main = "Histogram Affected Streamlines in LIM Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20, xlim=c(0,0.7), ylim=c(0,70))
+
+hist(x=fascicle_volumes_yeo_7$prop_DM, main = "Histogram Affected Streamlines in DM Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20, xlim=c(0,0.7), ylim=c(0,70))
+
+hist(x=fascicle_volumes_yeo_7$prop_FP, main = "Histogram Affected Streamlines in FP Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20, xlim=c(0,0.7), ylim=c(0,70))
+
+#no breaks, trying to standardize visualization
+
+hist(x=fascicle_volumes_yeo_7$prop_VIS, main = "Histogram Affected Streamlines in VIS Net", xlab = "Proportion Streamlines per Fascicle Affected", xlim=c(0,0.6), ylim=c(0,80))
+
+hist(x=fascicle_volumes_yeo_7$prop_MOT, main = "Histogram Affected Streamlines in MOT Net", xlab = "Proportion Streamlines per Fascicle Affected", xlim=c(0,0.6), ylim=c(0,80))
+
+hist(x=fascicle_volumes_yeo_7$prop_DA, main = "Histogram Affected Streamlines in DA Net", xlab = "Proportion Streamlines per Fascicle Affected", xlim=c(0,0.6), ylim=c(0,80))
+
+hist(x=fascicle_volumes_yeo_7$prop_VA, main = "Histogram Affected Streamlines in VA Net", xlab = "Proportion Streamlines per Fascicle Affected", xlim=c(0,0.6), ylim=c(0,80))
+
+hist(x=fascicle_volumes_yeo_7$prop_LIM, main = "Histogram Affected Streamlines in LIM Net", xlab = "Proportion Streamlines per Fascicle Affected", xlim=c(0,0.6), ylim=c(0,80))
+
+hist(x=fascicle_volumes_yeo_7$prop_DM, main = "Histogram Affected Streamlines in DM Net", xlab = "Proportion Streamlines per Fascicle Affected", xlim=c(0,0.6), ylim=c(0,80))
+
+hist(x=fascicle_volumes_yeo_7$prop_FP, main = "Histogram Affected Streamlines in FP Net", xlab = "Proportion Streamlines per Fascicle Affected", xlim=c(0,0.6), ylim=c(0,80))
 ```
@@
 #print(z)
+
+### FP only; Just pull out FP data
+
+#drop columns not in FP
+just_fp_fascicles <- subset(x = fascicle_proportions, select = c("EMPI", "EXAM_DATE", fascicle_names_FP_network))
+
+added_dep_to_fascicles <- merge(just_dep_and_empi, just_fp_fascicles, by = c("EMPI"))
+melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
+
+#exclude regions where there was no lesion at all
+lesioned <- subset(melted_df, value > 0)
+
+#only include people in healthy or depressed group
+lesioned_dx <- subset(lesioned, depGroupVar != 0)
+
+#separate out healthy and depressed
+healthy <- subset(lesioned, depGroupVar == 1)
+depressed <- subset(lesioned, depGroupVar == 2)
+
+fp<-ggplot(lesioned, aes(x=value, fill=depGroupVar, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
+print(fp)
+
+#drop columns not in FP 10% 
+just_fp_fascicles <- subset(x = fascicle_proportions, select = c("EMPI", "EXAM_DATE", fascicle_names_FP_network_10percent))
+
+added_dep_to_fascicles <- merge(just_dep_and_empi, just_fp_fascicles, by = c("EMPI"))
+melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
+
+#exclude regions where there was no lesion at all
+lesioned <- subset(melted_df, value > 0)
+
+#only include people in healthy or depressed group
+lesioned_dx <- subset(lesioned, depGroupVar != 0)
+
+#separate out healthy and depressed
+healthy <- subset(lesioned, depGroupVar == 1)
+depressed <- subset(lesioned, depGroupVar == 2)
+
+fp_10<-ggplot(lesioned, aes(x=value, fill=depGroupVar, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
+print(fp_10)
+######## Non FP networks
+#drop columns in FP 
+just_nonfp_fascicles <- subset(x = fascicle_proportions, select = c("EMPI", "EXAM_DATE", fascicle_names_nonFP_network))
+
+added_dep_to_fascicles <- merge(just_dep_and_empi, just_nonfp_fascicles, by = c("EMPI"))
+melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
+
+#exclude regions where there was no lesion at all
+lesioned <- subset(melted_df, value > 0)
+
+#only include people in healthy or depressed group
+lesioned_dx <- subset(lesioned, depGroupVar != 0)
+
+#separate out healthy and depressed
+healthy <- subset(lesioned, depGroupVar == 1)
+depressed <- subset(lesioned, depGroupVar == 2)
+
+nonfp<-ggplot(lesioned, aes(x=value, fill=depGroupVar, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
+print(nonfp)
+
+#drop columns in FP 10% 
+just_nonfp_fascicles <- subset(x = fascicle_proportions, select = c("EMPI", "EXAM_DATE", fascicle_names_nonFP_network_10percent))
+
+added_dep_to_fascicles <- merge(just_dep_and_empi, just_nonfp_fascicles, by = c("EMPI"))
+melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
+
+#exclude regions where there was no lesion at all
+lesioned <- subset(melted_df, value > 0)
+
+#only include people in healthy or depressed group
+lesioned_dx <- subset(lesioned, depGroupVar != 0)
+
+#separate out healthy and depressed
+healthy <- subset(lesioned, depGroupVar == 1)
+depressed <- subset(lesioned, depGroupVar == 2)
+
+nonfp_10<-ggplot(lesioned, aes(x=value, fill=depGroupVar, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
+print(nonfp_10)
+
 ```
@@
 
-#lm
-fascicle_age_lm <- lapply(fascicle_names, function(x) 
+#lm fp
+fascicle_age_lm <- lapply(fascicle_names_FP_network, function(x) 
 {
@@
 })
-names(fascicle_age_lm) <- fascicle_names
+names(fascicle_age_lm) <- fascicle_names_FP_network
 
@@
 print(fascicle_age_anova_fdr)
-
-#fdr corrected only association cortex
-fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
-fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
-print(fascicle_anova_fdr_association)
 
@@
        points=list(col="black"))
+
+#lm fp 10 %
+fascicle_age_lm <- lapply(fascicle_names_FP_network_10percent, function(x) 
+{
+  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_age_lm) <- fascicle_names_FP_network_10percent
+
+#anova
+fascicle_age_anova <- lapply(fascicle_age_lm, anova)
+
+#fdr corrected
+fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
+print(fascicle_age_anova_fdr)
+
+#visreg
+#sapply(fascicle_age_lm,visreg)
+visreg(fascicle_age_lm$AF_R,ylab = "Proportion of Affected Streamlines", xlab = "Age in Years",
+       line=list(col="red"),
+       points=list(col="black"))
+
+####### Non FP
+#lm non fp 
+fascicle_age_lm <- lapply(fascicle_names_nonFP_network, function(x) 
+{
+  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_age_lm) <- fascicle_names_nonFP_network
+
+#anova
+fascicle_age_anova <- lapply(fascicle_age_lm, anova)
+
+#fdr corrected
+fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
+print(fascicle_age_anova_fdr)
+
+#visreg
+#sapply(fascicle_age_lm,visreg)
+#visreg(fascicle_age_lm$AF_R,ylab = "Proportion of Affected Streamlines", xlab = "Age in Years",
+ #      line=list(col="red"),
+  #     points=list(col="black"))
+
+#lm nonfp 10 %
+fascicle_age_lm <- lapply(fascicle_names_nonFP_network_10percent, function(x) 
+{
+  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_age_lm) <- fascicle_names_nonFP_network_10percent
+
+#anova
+fascicle_age_anova <- lapply(fascicle_age_lm, anova)
+
+#fdr corrected
+fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
+print(fascicle_age_anova_fdr)
+
+#visreg
+#sapply(fascicle_age_lm,visreg)
+#visreg(fascicle_age_lm$AF_R,ylab = "Proportion of Affected Streamlines", xlab = "Age in Years",
+ #      line=list(col="red"),
+  #     points=list(col="black"))
 
@@
 
-#lm
-fascicle_sex_lm <- lapply(fascicle_names, function(x) 
+#lm fp
+fascicle_sex_lm <- lapply(fascicle_names_FP_network, function(x) 
 {
@@
 })
-names(fascicle_sex_lm) <- fascicle_names
+names(fascicle_sex_lm) <- fascicle_names_FP_network
 
@@
 
-#fdr corrected only association cortex
-fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
-fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
-print(fascicle_anova_fdr_association)
-
-#fdr corrected only association cortex
-fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
-fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
-print(fascicle_anova_fdr_association)
+#uncorrected
+fascicle_sex_anova_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
+fascicle_sex_anova_p05_unc <- as.data.frame(fascicle_sex_anova_p[fascicle_sex_anova_p < 0.05])
+print(fascicle_sex_anova_p05_unc)
+
+#lm fp 10
+fascicle_sex_lm <- lapply(fascicle_names_FP_network_10percent, function(x) 
+{
+  lm(substitute(i ~ osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sex_lm) <- fascicle_names_FP_network_10percent
+
+#anova
+fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)
+
+#fdr corrected
+fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
+print(fascicle_sex_anova_fdr)
+
+#uncorrected
+fascicle_sex_anova_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
+fascicle_sex_anova_p05_unc <- as.data.frame(fascicle_sex_anova_p[fascicle_sex_anova_p < 0.05])
+print(fascicle_sex_anova_p05_unc)
+
+####
+#lm nonfp
+fascicle_sex_lm <- lapply(fascicle_names_nonFP_network, function(x) 
+{
+  lm(substitute(i ~ osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sex_lm) <- fascicle_names_nonFP_network
+
+#anova
+fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)
+
+#fdr corrected
+fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
+print(fascicle_sex_anova_fdr)
+
+#uncorrected
+fascicle_sex_anova_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
+fascicle_sex_anova_p05_unc <- as.data.frame(fascicle_sex_anova_p[fascicle_sex_anova_p < 0.05])
+print(fascicle_sex_anova_p05_unc)
+
+#lm nonfp 10
+fascicle_sex_lm <- lapply(fascicle_names_nonFP_network_10percent, function(x) 
+{
+  lm(substitute(i ~ osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_sex_lm) <- fascicle_names_nonFP_network_10percent
+
+#anova
+fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)
+
+#fdr corrected
+fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
+print(fascicle_sex_anova_fdr)
 
@@
   mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
-  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_10_percent =
-           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
-   mutate(total_fascicle_prop_lost_indepnet_5_percent =
-           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
-  mutate(total_fascicle_prop_lost_nondep_network =
-           rowMeans(across(fascicle_names_nondep_network))) %>%
+  mutate(total_fascicle_prop_lost_FP_network = 
+           rowMeans(across(fascicle_names_FP_network))) %>%
+  mutate(total_fascicle_prop_lost_FP_network_10percent =
+           rowMeans(across(fascicle_names_FP_network_10percent))) %>%
+  mutate(total_fascicle_prop_lost_nonFP_network =
+           rowMeans(across(fascicle_names_nonFP_network))) %>%
+  mutate(total_fascicle_prop_lost_nonFP_network_10percent =
+           rowMeans(across(fascicle_names_nonFP_network_10percent))) %>%
   group_by(EMPI) %>% 
@@
 summary(dep_vs_healthy) #p=0.494, cohen d = 0.2
-
-#power analysis to predict sample required for K, u = num df, v = denom df, f2=0.01 (cohen d = 0.02/2) 
-#prop in each fascicle predicting dep (87), number of predictors (age, sex, ef, ess?) = 4, u = 87-4 = 83
-pwr.f2.test(u=83, f2 = 0.02, sig.level = 0.5, power = 0.8)
 
@@
 
-#look at within depression network specifically - trend, p = 0.059
-total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 36%, mean non_dep = 32%, p = 0.071
-
-#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
-#yes, p = 0.048
-total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048
-
-#look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
-#yes, 
-total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.056
-
-#look at outside depression network, yes p = 0.037
-total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.037
-
-### Age actually makes things worse
-#dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
-#summary(dep_vs_healthy_w_age)
-
-#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
+#look at fp network specifically - trend, p = 0.063
+total_volume_lost_dep_vs_healthy_unique_in_FP_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_FP_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_FP_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #healthy = 28% dep = 32%
+
+#look at within FP network in fascicles where at least 10% of the volume of the fascicle was in FP network
+#yes, p = 0.064
+total_volume_lost_dep_vs_healthy_unique_in_FP_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_FP_network_10percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_FP_network_10percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #healthy = 22% dep = 26%
+
+#look at outside FPnetwork, yes p = 0.027
+total_volume_lost_dep_vs_healthy_unique_in_nonFP_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nonFP_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nonFP_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #healthy = 17% dep = 20%
+
+#look at outside FP network, yes p = 0.049
+total_volume_lost_dep_vs_healthy_unique_in_nonFP_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nonFP_network_10percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nonFP_network_10percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #healthy = 24 % dep = 28%
 
@@
 
+fascicle_bundle_mapping=get_fascicle_bundle_mapping()
 #fdr corrected only association cortex
@@
 
-#fdr corrected only dep 5% mask
-fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)]
-fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+#fdr corrected only vis network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$VIS == 1)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only mot network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$MOT == 1)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only da network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$DA == 1)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only VA network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$VA == 1)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only LIM network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$LIM == 1)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only FP network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$FP == 1)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr)
+
+#fdr corrected only DM network
+fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$DM == 1)]
+fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
 print(fascicle_anova_fdr)
@@
 
-### plot AF_R
-
-df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
-names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
-df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
-
-plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
-  ggtitle("Arcuate Fasciculus R") + 
- # geom_violin() + 
-  geom_jitter(shape=16, position=position_jitter(0.05)) + 
-  geom_boxplot(width=0.1) + theme_minimal() + 
-  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
-  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))
-
-print(plot)
-
-### Plot OR_R	0.69024050
-df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$OR_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$OR_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
-names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
-df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
-
-plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
-  ggtitle("Optic Radiation R") + 
- # geom_violin() + 
-  geom_jitter(shape=16, position=position_jitter(0.05)) + 
-  geom_boxplot(width=0.1) + theme_minimal() + 
-  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
-  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) 
-print(plot)
-
-### grouped bar plot
-
-Diagnosis <- c(rep("Depressed" , 2) , rep("Non-Depressed" , 2))
-Fascicle <- rep(c("Depression Network (AF)" , "Non-depression network (SCP)"), 2)
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))),
-          (sd(dep_and_healthy_groups_for_ICD_analysis$SCP[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))))
-
-data <- data.frame(Diagnosis,Fascicle,value, sem)
-#data$Fascicle <- pretty_names
-
-# Grouped
-barplot <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
-    geom_bar(position="dodge", stat="identity") + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
-  ylab("Proportion of Fascicles Affected") +
-  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())
-
-print(barplot)
-
-# Grouped
-barplot <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
-    geom_bar(position="dodge", stat="identity") + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
-  ylab("Proportion of Fascicles Affected") +
-  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())
-
-print(barplot)
-
-barplot2 <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
-    geom_bar(position="dodge", stat="identity", colour = "black", width=0.9,size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) + 
-    scale_fill_manual(values = c("#cdcdcd", "#FFFFFF", "#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-
-    ylab("Proportion of Fascicles Affected") +
-    theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-          axis.title.y = element_text(size=20),
-          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-          panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty graph
-    scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
-
-print(barplot2)
-
-```
-
-```{r testing_whether_ms_disease_more_likely_in_depression_mask_unique}
-### tests whether there is higher proportion of vertices lost in depression network versus not
-
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
-  filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
-  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_10_percent =
-           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_5_percent =
-           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
-  mutate(total_fascicle_prop_lost_nondep_network =
-           rowMeans(across(fascicle_names_nondep_network))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ### when I filter and remove fascicles where there was no loss at all, the depression vs healthy findings go away
-#  filter(total_fascicle_prop_lost > 0) %>%
-  ungroup()
-
-#test whether there is differentially more volume loss in depression versus nondepression network
-volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 35%, mean non_dep_net =  17%, p 2.2 x 10^-16
-
-volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
-dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 31%, mean non_dep_net = 17%, p 2.2 * 10^-16
-
-volume_lost_dep5_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent,
-dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 17%, p 2.2 * 10^-16
-
-#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
-df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network), c(rep(1,298), rep(2,298))))
-names(df_for_visreg) <- c("EMPI", "prop_lost", "network")
-df_for_visreg$network <- as.factor(df_for_visreg$network)
-
-plot <- ggplot(df_for_visreg, aes(x=network, y = prop_lost, col=network)) +
-  geom_violin() + 
-  geom_jitter(shape=16, position=position_jitter(0.05)) + 
-  geom_boxplot(width=0.1) + theme_minimal() + 
-  stat_summary(fun.y=median, geom="point", size=2, color="red") + 
-  scale_x_discrete(labels=c("1" = "Depression Network", "2" = "Non-Depression Network")) + 
-  ylab("Proportion of Fascicles Affected") + 
-  theme(axis.text =element_text(size=14), axis.title.x = element_blank())
-
-print(plot)
-
-########## Bar plot
-Network <- c("Depression Network","Non-Depression Network")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet),
-           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])))
-
-data <- data.frame(Network,value,sem)
-#data$Fascicle <- pretty_names
-
-# Grouped
-barplot <- ggplot(data, aes(x=Network,fill=Network, y=value)) + 
-  geom_bar(stat="identity", position=position_dodge(0.9)) + 
-  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
-  ylab("Proportion of Fascicles Affected") +
-  ylim(0,0.5) +
-  theme(axis.text = element_text(size = 12, colour = "black"))
-
-print(barplot)
-
-###### Depression network x depression group interaction - It is there, depression always worse. 
-#these lms largely match up with the t.tests above. Sig differences between depressed (more disease) within and outside of mask
-
-loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #non sig, p = 0.061
-loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.0492
-loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.037
 ```

### 2022-08-03 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
Add files via upload

@@
 library(boot)
+library(boot.pval)
 #functions
@@
 
-  B) Depressed vs Controls Proportions
-
-    - Overall dep vs non-dep, p = 0.050 (NS), Dep = 31%, Non dep = 27%
+  B) Depressed vs Controls Fascicle Affected - Proportions
+
+    - Dep vs non-dep, total fascicle proportion affected p = 0.050 (NS), Dep = 31%, Non dep = 27%
     - Dep vs non, top 25%, **p = 0.047**, Dep = 32%, non dep = 26% 
@@
 
-  C) Depressed vs Controls Volume
-
-    - Overall dep vs nondep: **p = 0.045**, Dep = 9382, Non dep = 8260
+  C) Depressed vs Controls Fascicle Affected - Volume
+
+    - Dep vs nondep, total fascicle volume affected: **p = 0.045**, Dep = 9382, Non dep = 8260
     - Dep vs nondep in top 25% yes **p = 0.034**, Dep = 4829, non dep = 3914
@@
         - 95%   ( 0.701,  4.553 )
-        - boostrap p calc from CI - **0.008**
+        - boostrap p calc from CI - **0.0087**
 
@@
         - 95%   ( 1.052,  3.724 )
-        - bootstrap p calc from CI- **0.001**
+        - bootstrap p calc from CI- **1*10^-16**
 
 I think it works best when sticking with *volumes* rather than proportions
+
+#### Promis data, late edition
+
+We decided to look at some of the promis data (which measures broad functional categories, though not specifically psychiatric diagnosis) to see if a) measures differed by diagnostic groups and b) whether there were dimensional effects.
+Unfortunately, the PROMIS picked up in 2018, which was also the same year that Melissa ended her pull, so our n's are small.
+
+Current ns for people with at least 1 promis score= 
+Total n promis scores: 1694 
+Total n unique individuals w/promis scores: 501
+
+Different scores have different # of overall scans
+                        Total                      General_Health                  Quality_of_Life                  Physical_Health 
+                            1694                             1693                             1503                             1503 
+          Mental_Health_and_Mood   Social_Activities_Satisfaction   Carrying_Out_Social_Activities Carrying_Out_Physical_Activities 
+                            1503                             1503                             1501                             1499 
+              Emotional_Problems                  Fatique_Average     PostOp.PROMIS.Physical.Score       PostOp.PROMIS.Mental.Score 
+                            1499                             1499                             1440                             1442
+
+W/good mimosa:
+Total n promis scores : 398
+Total n unique individuals w/promis scores: 116
+Total n unique indivs in nondepressed group = 36
+Total n unique individuals in depressed group = 46
+
+Method:
+To use the promis scores, I cleaned them in a separate script (promis_score_cleaning_2010-2022.Rmd)
+In this script, I link each scan with its most proximal promis score.
+The challenge here is that there is no time linking.
+Summary of the time (in days and years) between scanning and a promis (these are absolute values, do not indicate before or after).
+
+Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+    1.0   376.8   763.0   873.3  1229.8  2487.0 
+    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
+0.002747 1.035027 2.096154 2.399132 3.378434 6.832418
+
+Next, I looked at overall differences between diagnostic groups on each of the PROMIS measures, correcting for multiple comparisons w/FDR (Q<0.05).
+
+Then, I looked at the relationship between promis scores and fascicle measures (overall, 1620 analyses done):
+
+Proportion:
+Proportion within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
+    - all participants (116)
+    - Depressed + nondepressed (82)
+    - Non-depressed alone (36)
+    - Depressed alone (46)
+
+Proportion of individual fascicle affected related to promis scores (777 analyses per group (11 promis domains x 77 individual fascicles))
+  - all participants (116)
+  - Depressed + nondepressed (82)
+  - Non-depressed alone (36)
+  - Depressed alone (46)
+
+Volume:
+
+Volume within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
+   - all participants (116)
+    - Depressed + nondepressed (82)
+    - Non-depressed alone (36)
+    - Depressed alone (46)
+
+Volume of individual fascicle affected related to promis scores  (777 analyses per group (11 promis domains x 77 individual fascicles))
+  - all participants (116)
+  - Depressed + nondepressed (82)
+  - Non-depressed alone (36)
+  - Depressed alone (46)
+
+Promis results:
+Interestingly, across all 11 scores, with the exception of fatigue, Depressed patients have lower scores (lower = worse functioning), than non-depressed patients.
+controls.
+They all fdr correct.
+What I love about this is fatigue not being different.
+If fatigue was worse in depressed patients, one could argue that quality of life, health, emotional problems, social activities, etc are all worse because of underlying fatigue.
+Similarly, as fatigue is a component of the PHQ9 and depression screening, this could be confounding.
+That they were different in a lot of measures but not fatigue is informative.
+
+Quality_of_Life	0.015199144			
+Physical_Health	0.012579264			
+Mental_Health_and_Mood	0.012579264			
+Social_Activities_Satisfaction	0.001063449			
+Carrying_Out_Social_Activities	0.011375799			
+Carrying_Out_Physical_Activities	0.031616473			
+Emotional_Problems	0.011270683			
+Fatique_Average	0.231497853			
+PostOp.PROMIS.Physical.Score	0.041172446			
+PostOp.PROMIS.Mental.Score	0.001253404
+
+Proportion
+Proportion within versus outside depression map correlated w/promis scores - NO SIGNIFICANT FINDINGS
+    - all participants (116)
+    - Depressed + nondepressed (82)
+    - Non-depressed alone (36)
+    - Depressed alone (46)
+
+Proportion of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
+  - all participants (116)
+  - Depressed + nondepressed (82)
+  - Non-depressed alone (36)
+  - Depressed alone (46)
+
+Volume     
+Volume within versus outside depression map correlated w/promis scores -  NO SIGNIFICANT FINDINGS
+   - all participants (116)
+    - Depressed + nondepressed (82)
+    - Non-depressed alone (36)
+    - Depressed alone (46)
+
+Volume of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
+  - all participants (116)
+  - Depressed + nondepressed (82)
+  - Non-depressed alone (36)
+  - Depressed alone (46)
 
@@
    mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
+   mutate(BEGIN_EXAM_DTTM = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% #make sure to actually save it as date object for later processing
    mutate(EMPI = as.factor(EMPI)) %>%
@@
 df_demo_and_fascicles <- merge(df_demo_and_fascicles, lesion_volumes, by = c("EMPI", "EXAM_DATE")) #n = 2962
+
+##########################################################
+#            Read in PROMIS Data              #
+##########################################################
+promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_cleaned_n1694_unique_indiv_n501_2015-2022.rds"))
+promis_score_names <- names(promis_data[7:16])
+
+df_people_with_promis_data <- df_demo_and_fascicles %>%
+  filter(EMPI %in% promis_data$EMPI) #n=398
+
+unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=116
+
+```
+```{r make_df_with_volume_instead_of_proportion}
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#read in fiber volume values
+fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+
+```
+
+```{r promis_data_handling}
+#go through each person w/promis data and append the columns to their row that reflects the closest collected PROMIS
+
+#function, take an empi and date, returns a column of promis scores
+get_closest_promis_to_ms_exam_date <- function(subj_empi, subj_exam_date) {
+  df_closest_promis <- promis_data %>%
+    filter(EMPI == subj_empi) %>%
+    mutate(distance_from_scan = abs(RECORDED_TIME-subj_exam_date)) %>%
+    filter(abs(RECORDED_TIME-subj_exam_date)==min(abs(RECORDED_TIME-subj_exam_date))) %>%
+    slice(1)
+  return(df_closest_promis)
+}
+
+#make dataframe to store promis info
+promis_info_closest_proximity <- data.frame(matrix(NA, nrow=dim(df_people_with_promis_data)[1], ncol=(dim(promis_data)[2] + 1)))
+names(promis_info_closest_proximity) <- c(names(promis_data), "distance_from_scan")
+
+for (line in 1:dim(df_people_with_promis_data)[1]) {
+  #pull out subject empi and exam date
+  print(line)
+  subj_empi <- df_people_with_promis_data$EMPI[line]
+  subj_exam_date <- df_people_with_promis_data$BEGIN_EXAM_DTTM[line]
+  closest_promis_data <- get_closest_promis_to_ms_exam_date(subj_empi = subj_empi, subj_exam_date = subj_exam_date)
+  promis_info_closest_proximity[line,] = closest_promis_data
+}
+
+#drop unnecessary columns c("DE_PAT_ID", "EMPI", "HUP_MRN", "SEX", "RACE")
+promis_info_closest_proximity_just_scores <- promis_info_closest_proximity[,6:dim(promis_info_closest_proximity)[2]]
+
+#combine with df_people_with_promis_data
+df_people_with_promis_data_scores_added = cbind(df_people_with_promis_data, promis_info_closest_proximity_just_scores)
+
+#summarize the distance from scans
+
+summary(df_people_with_promis_data_scores_added$distance_from_scan) #mean 873.3 days, median 763 days, range 1 to 2487 days, convert to years
+hist(df_people_with_promis_data_scores_added$distance_from_scan) #convert to years
+
+summary(df_people_with_promis_data_scores_added$distance_from_scan/364) #mean 2.40 ys, median 2.10 years, range 0.003 y to 6.83 years
+hist(df_people_with_promis_data_scores_added$distance_from_scan/364) #convert to years
+
+#sub volume for proportion
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_only <- df_people_with_promis_data_scores_added[ , (names(df_people_with_promis_data_scores_added) %in% fascicle_names)]
+df_everything_else <- df_people_with_promis_data_scores_added[ , !(names(df_people_with_promis_data_scores_added) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+df_demo_and_fascicle_volumes_full_w_promis <- cbind(df_everything_else, df_fascicle_volumes)
+
 ```
@@
 make_demographics_table_ms(data_frame = df_good_mimosa_unique)
-```
+
+### with promis data
+make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added)
+
+## overall
+#just good mimosa
+df_good_mimosa_promis <- df_people_with_promis_data_scores_added %>% filter(depGroupVar != 0)
+make_demographics_table_ms_w_promis(df_good_mimosa_promis) #n=287 scans
+
+#unique
+df_people_with_promis_data_scores_added_unique <- df_people_with_promis_data_scores_added %>%
+  filter(depGroupVar != 0) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added_unique)
+
+```
+
 ```{r compare_overall_volume_of_lesions_dep_vs_non_dep}
@@
 #read in fiber volume values
-fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+#fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
 
 #extract data frame that only includes the fascicle proportion lost
-df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
-df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+#df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+#df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
 
 #multiply each row by the vector of total volumes
-df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
+#df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+ #                 STATS = fiber_volume_values$volume_full, FUN = "*")
 
 #combine back with data_fram
-df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
 
@@
 
-##### proportion
-# 95% confidence intervals for the correlation
-set.seed(1)
+######### functions
 tval <- function(formula, data, indices) {
@@
 
+##### proportion
+# 95% confidence intervals for the correlation
+
+set.seed(1)
 data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_t_value, select = c("depression_mask_overlap","dep_vs_healthy_t"))
@@
 plot(results_prop)
-boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="bca", index=1)
+boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="perc", index=1)
 print(boot_results_prop.ci)
-
-#manually calculating P from CI, from interwebs
-est=results_prop$t0
-l=boot_results_prop.ci$bca[4]
-u=boot_results_prop.ci$bca[5]
-
-#1) SE=(u-l)/2*1.96
-se=((u-l)/(2*1.96))
-#2) z=Est/SE
-z=est/se
-#3) p=exp(-0.717*z - 0.416*z^2)
-p=exp((-0.717*z) - (0.416*(z*z)))
-print(p) #p=0.008
+pval_prop<-boot.pval(boot_res=results_prop, theta_null = 0, type = "perc", pval_precision = NULL)
+print(pval_prop) #p=0.0087
 
@@
 plot(results_vol)
-boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="bca", index=1)
+boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
 print(boot_results_vol.ci)
-
-#get p from CI
-#manually calculating P from CI, from interwebs
-est=results_vol$t0
-l=boot_results_vol.ci$bca[4]
-u=boot_results_vol.ci$bca[5]
-
-#1) SE=(u-l)/2*1.96
-se=((u-l)/(2*1.96))
-#2) z=Est/SE
-z=est/se
-#3) p=exp(-0.717*z - 0.416*z^2)
-p=exp((-0.717*z) - (0.416*(z*z)))
-print(p) # #p=0.001
-
-```
+pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
+print(pval_vol) #p=1*10^-16
+
+```
+
+##### PROMIS DATA
+
+```{r compare_overall_promis_scores_by_diagnosis}
+
+#separate out only people in depression group, and take first instance
+dep_and_healthy_groups_for_ICD_analysis <-  df_people_with_promis_data_scores_added %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=82
+
+dep_vs_healthy_t <- lapply(promis_score_names, function(x) 
+{
+  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
+  eval(parse(text=t_test_command))
+})
+names(dep_vs_healthy_t) <- promis_score_names
+print(dep_vs_healthy_t)
+
+#pull t-statistic
+dep_vs_healthy_t_values <- sapply(dep_vs_healthy_t, function(x) x$statistic)
+
+#convert to data frame
+df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
+row.names(df_t_values) <- promis_score_names
+print(df_t_values)
+
+#pull p-values
+dep_vs_healthy_p_values <- sapply(dep_vs_healthy_t, function(x) x$p.value)
+
+#convert to data frame
+df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
+row.names(df_dep_vs_healthy_p_values) <- promis_score_names
+print(df_dep_vs_healthy_p_values)
+#FDR correct p-values
+dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
+
+plots <- lapply(promis_score_names, function(x) 
+{
+
+  Diagnosis <- c("MS w/Depression", "MS No Depression")
+  dep_group <- eval(parse(text=paste0("dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]")))
+  nondep_group <- eval(parse(text=paste0("dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]")))
+  value <- c(mean(dep_group, na.rm=TRUE),
+           mean(nondep_group, na.rm=TRUE))
+  sem=c((sd(dep_group, na.rm = TRUE)/sqrt(sum(dep_group, na.rm = TRUE))), (sd(nondep_group, na.rm = TRUE)/sqrt(sum(nondep_group, na.rm = TRUE))))
+
+  data <- data.frame(Diagnosis,value,sem)
+
+  # Grouped
+  barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+    geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+      geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+    scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+    ggtitle(paste0(x, " by Diagnosis")) + #, p=", round(dep_vs_healthy_t_test$p.value,3))) +
+    ylab(paste0("Promis ", x)) +
+    theme(title = element_text(size=14),
+          axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+          axis.title.y = element_text(size=20),
+          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+          panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+  #  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+  print(barplot) #
+
+})
+
+```
+
+#### PROMIS Proportion analyses
+
+```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+ # filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_dep_and_nondep}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_nondep}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+  filter(depGroupVar == 1)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_dep}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+  filter(depGroupVar == 2)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_irrespective_of_dx}
+
+#total n = 116, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=116
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_depressed_and_nondepressed}
+
+#total n = 82, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=82
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_nondepressed}
+
+#total n = 36, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+  filter(depGroupVar == 1)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=36
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_depressed}
+
+#total n = 46, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
+  filter(depGroupVar == 2)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=46
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+#### PROMIS VOLUME ANALYSES
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
+ # filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_dep_and_nondep}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_nondep}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
+  filter(depGroupVar == 1)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_dep}
+
+#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
+  filter(depGroupVar == 2)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
+
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_proportion_lost_categories
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_irrespective_of_dx_vol}
+
+#total n = 116, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=116
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_depressed_and_nondepressed_vol}
+
+#total n = 82, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=82
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_nondepressed_vol}
+
+#total n = 36, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
+  filter(depGroupVar == 1)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=36
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```
+
+```{r dimensional_promis_effects_unique_depressed_vol}
+
+#total n = 46, nothing significant
+
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
+  filter(depGroupVar == 2)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=46
+
+#ones to try
+#PostOp.PROMIS.Mental.Score
+#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
+# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
+# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
+
+#lm
+promis_lm <- lapply(promis_score_names, function(promis_score) {
+
+  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
+  {
+    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
+       data = dep_and_healthy_groups_for_ICD_analysis)
+  })
+  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
+
+  #anova
+  fascicle_anova <- lapply(fascicle_lm, anova)
+
+  #fdr corrected
+  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
+  print(fascicle_anova_fdr)
+})
+
+```

### 2022-08-03 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
     https://www.statmethods.net/advstats/bootstrapping.html
+
+#### 20220802
+
+Promis scores
+1) First, clean up the promise scores
+    [x] get rid of text in the fields (good/bad/whatever)
+    [x] change date to be same format as our main spreadsheet
+
+2) Get a sense of how many people in our sample have promise scores
+  - sort promise scores by empi and date, take unique, merge with existing data set and count
+
+  https://stackoverflow.com/questions/43472234/fastest-way-to-find-nearest-value-in-vector
+1) First, get a sense of how many people in our sample
+
+n = all promis - 1694 scores, 501 unique subjects. 
+398 in our overall pull (granted this pull went up to 2022, our ms pull was in 2021), 116 unique empis
+depressed promis - 46 unique
+nondepressed promis - 36 unique
+depressed + nondepressed prop - 82 unique
+
+Look at distribution
+
+MS lesion age, predict age in a ridge regression. Predict how old someone is how bad their lesions are. Your disease progression is better or worse than espected. 
+#### 20220803
+
+[x] still working on getting pvalue from boot - yes, it works! Needed boot.pval extra package, and switched to perc instead of bca
+  - we originally used bca, but the pval command was breaking for some reason we couldn't quite get (something about being out of range). Per Taki-
+
+  I took a quick peek and if you're using the method described by Altman it totally makes sense you'll get something slightly different; he made a normal assumption and used that for the interval inversion.
+This, rather, is totally empirical.
+As for why it's not working with bca, that's a different question.
+I'm honestly not sure.
+We could try to troubleshoot this together, but TBH I think i'd just slide back into using the  percentile bootstrap since it's legit here as well.
+The bca is considered better, but it's not the kind of thing that keeps me up at night personally.
+Does that make sense?
+
+----------
+1,554 analyses total! Nothing significant
+[x] fascicles x promis score in each individual fascicle (both proportion and volume) -NS
+[x] broken up by dx
+[x] within dep net
+[x] outside dep net
+----------
+
+-----------
+66 analysis total, nothing significant
+[x] fascicles x promis score in each network (total, depression 25, non dep 75%, both proportion and volume)
+[x] broken up by dx
+[x] within dep net
+[x] outside dep net

### 2022-08-05 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
Add files via upload

@@
 library(boot.pval)
+library(lsr)
 #functions
@@
 
-  D) T values in fascicle associated with Overlap in Depression map
+  D) Linear models of values in fascicle associated with Overlap in Depression map
 
@@
 
+    - Cohen'd correlated w/proportion of overlap of fascicle w/depression map
+        - p of lm (0.001)
+        -
+
     - T values correlated w/volume of overlap with depression network (i.e. # voxels that fascicle shares w/depression map),
@@
 
+    - Cohen'd correlated w/volume of overlap of fascicle w/depression map
+        - p of lm (p < 2.2 * 10^-16)
+        - Largest effect in corpus callosum
 I think it works best when sticking with *volumes* rather than proportions
@@
 df_good_mimosa_unique <- df_demo_and_fascicles %>% 
-  filter(depGroupVar != 0) %>%
+#  filter(depGroupVar != 0) %>%
   group_by(EMPI) %>% 
@@
 ## overall
-#just good mimosa
+#just good mimosa w/promis
 df_good_mimosa_promis <- df_people_with_promis_data_scores_added %>% filter(depGroupVar != 0)
@@
 {
+
   t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
@@
 
+```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop_effect_size}
+# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups
+
+##############################
+## FASCICLES TO BE INCLUDED ##
+##############################
+
+#define the depression network (all areas that intersect with depression network, no cranial nerves)
+fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
+
+#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
+fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
+fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
+
+#######################
+### depression map  ###
+#######################
+
+### define depression map
+fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
+
+#filter to exclude fascicles w/cranial nerves
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
+  replace(is.na(.), 0) %>%
+  filter(fascicle%in% fascicle_names_no_cranial_nerves)
+
+#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column
+
+df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))
+
+#######################
+### real data x dx  ###
+#######################
+
+#then, make the data frame
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+# lapply to get results of t-test in each fascicle
+#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
+{
+  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
+  eval(parse(text=text_for_eval))
+})
+names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves
+
+#convert to data frame
+df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)
+
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_overlap_d_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_d_values)
+names(df_fascicle_dep_mask_overlap_d_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_d")
+
+#make copy to sort
+df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_d_value
+
+df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])
+
+# bar graph of T-values of fascicles, sorted by depression overlap
+plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") +
+  geom_bar(stat="identity") +
+  ylim(-0.1,0.5) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(plot_fascicle_by_d_sorted_by_dep_overlap)
+
+# plot fascicle overlap x t value
+#statistic, p = 0.010
+dep_overlap_x_dep_vs_healthy_d <- lm(dep_vs_healthy_d ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_d_value)
+print(summary(dep_overlap_x_dep_vs_healthy_d))
+
+scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_d_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_d)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Proportion of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis calculated in proportions (D)") + 
+  ggtitle(paste0("Fascicle/depression mask Proportion overlap by dx Eff Size, p = ", round(summary(dep_overlap_x_dep_vs_healthy_d)$coefficients[2,4],3))) + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_d_val_by_dep_net_overlap)
+
+```
+
 ```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume}
@@
 
+```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume_no_CC}
+
+#define the depression network (all areas that intersect with depression network, no cranial nerves)
+fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
+
+#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
+fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
+fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
+
+#######################
+### depression map  ###
+#######################
+
+### define depression map
+fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
+
+#filter to exclude fascicles w/cranial nerves
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
+  replace(is.na(.), 0) %>%
+  filter(fascicle%in% fascicle_names_no_cranial_nerves)
+
+df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
+
+########################
+## Get Volume Values ###
+########################
+
+#read in fiber volume values
+fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+
+#isolate out only the depressed group and healthy group, fascicles should be in volumes rather than proportions
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+# lapply to get results of t-test in each fascicle
+#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
+{
+  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
+  eval(parse(text=t_test_command))
+})
+names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves
+
+#pull t-statistic
+dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)
+
+#convert to data frame
+df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
+row.names(df_t_values) <- fascicle_names_no_cranial_nerves
+
+#pull p-values
+dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)
+
+#convert to data frame
+df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
+row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves
+
+#FDR correct p-values
+dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
+
+#Convert to data frame
+df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
+row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves
+
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_volume_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
+names(df_fascicle_dep_mask_volume_t_value) <- c("fascicle", "dep_vs_healthy_t", "volume_in_dep_mask")
+
+#remove corpus callosum
+df_fascicle_dep_mask_volume_t_value_no_cc <- df_fascicle_dep_mask_volume_t_value %>%
+  filter(fascicle != "CC")
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask <- df_fascicle_dep_mask_volume_t_value_no_cc
+
+#order by volume in depression mask
+df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$volume_in_dep_mask)])
+
+# bar graph of T-values of fascicles, sorted by depression overlap
+plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by volume of dep network, no CC") +
+  geom_bar(stat="identity") +
+  ylim(0,5) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(plot_fascicle_by_t_sorted_by_dep_overlap)
+
+#statistic volume, p = 0.031
+dep_overlap_x_dep_vs_healthy_t_no_cc <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value_no_cc)
+print(summary(dep_overlap_x_dep_vs_healthy_t_no_cc))
+
+scatter_t_val_by_dep_net_overlap_no_cc <- ggplot(data=df_fascicle_dep_mask_volume_t_value_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle(paste0("Fascicle/depression mask volume by dx T, no CC, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t_no_cc)$coefficients[2,4],3))) + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_t_val_by_dep_net_overlap_no_cc)
+
+```
+
+```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_vol_effect_size_no_CC}
+# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups
+
+##############################
+## FASCICLES TO BE INCLUDED ##
+##############################
+
+#define the depression network (all areas that intersect with depression network, no cranial nerves)
+fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
+
+#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
+fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
+fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
+
+#######################
+### depression map  ###
+#######################
+
+### define depression map
+fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
+
+#filter to exclude fascicles w/cranial nerves
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
+  replace(is.na(.), 0) %>%
+  filter(fascicle%in% fascicle_names_no_cranial_nerves)
+
+df_fascicle_and_vol_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
+
+########################
+## Get Volume Values ###
+########################
+
+#read in fiber volume values
+fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+
+###############################
+#######################
+### real data x dx  ###
+#######################
+
+#then, make the data frame
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+# lapply to get results of t-test in each fascicle
+#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
+{
+  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
+  eval(parse(text=text_for_eval))
+})
+names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves
+
+#convert to data frame
+df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)
+
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(df_fascicle_and_vol_in_mask$fascicle, df_fascicle_and_vol_in_mask$non_zero_voxels_in_dep_map, df_d_values)
+names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")
+
+df_fascicle_dep_mask_volume_overlap_d_value_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>%
+  filter(fascicle != "CC")
+
+#make copy to sort
+df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask <- df_fascicle_dep_mask_volume_overlap_d_value_no_cc
+
+df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$volume_in_dep_mask)])
+
+# bar graph of eff size of fascicles, sorted by depression overlap
+plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network, no CC") +
+  geom_bar(stat="identity") +
+  #ylim(-0.1,0.5) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(plot_fascicle_by_d_sorted_by_dep_overlap)
+
+# plot fascicle overlap x Eff size value
+#statistic, p=001
+dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc )
+print(summary(dep_overlap_x_dep_vs_healthy_d))
+
+scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc , aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis calculated in Volumes (D)") + 
+  ggtitle(paste0("Fascicle/depression mask Volume overlap by dx Eff Size, No CC p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_d_val_by_dep_net_overlap)
+
+```
+
 ```{r bootstrapping_proportion_and_volume_x_dep_t_analyses}
@@
 tval <- function(formula, data, indices) {
-  d <- data[indices,] # allows boot to select sample
-  fit <- lm(formula, data=d)
+  df <- data[indices,] # allows boot to select sample
+  fit <- lm(formula, data=df)
   return(summary(fit)$coefficients[2,3])
@@
 
+##### effect size
+# 95% confidence intervals for the correlation
+
+set.seed(1)
+data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_d_value, select = c("depression_mask_overlap","dep_vs_healthy_d"))
+results_prop <- boot(data=data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap, statistic=tval,
+   R=10000, formula=dep_vs_healthy_d~depression_mask_overlap)
+print(results_prop) #actual 2.62, CI 0.589, 4.669)
+plot(results_prop)
+boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="perc", index=1)
+print(boot_results_prop.ci)
+pval_prop<-boot.pval(boot_res=results_prop, theta_null = 0, type = "perc", pval_precision = NULL)
+print(pval_prop) #p=0.0083
+
 ##### volume
@@
    R=10000, formula=dep_vs_healthy_t~volume_in_dep_mask)
+print(results_vol)
+plot(results_vol)
+boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
+print(boot_results_vol.ci)
+pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
+print(pval_vol) #p=1*10^-16
+
+##### volume
+#effect sizes
+# 95% confidence intervals for the correlation
+set.seed(1)
+data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))
+
+results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
+   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
+print(results_vol)
+plot(results_vol)
+boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
+print(boot_results_vol.ci)
+pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
+print(pval_vol) #p=1*10^-16
+
+### with no corpus callosumm p = 1* 10^-16, CI(1.727, 5.363)
+# 95% confidence intervals for the correlation
+set.seed(1)
+results_vol <- boot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc, statistic=tval,
+   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
 print(results_vol)
@@
 })
-
 ```

### 2022-08-05 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
 [x] outside dep net
+
+#### 20220805
+[x] trying out correlations with effect sizes instead of Ts, wow this makes results even better, they all work
+
+#### 20220810
+[x] make sure the mutate command is okay, somehow I think we might be undercounting the healthy group.
+Okay ish now, but somehow the depGroupVar has a lower number than chaining together this command:  df_good_mimosa_unique %>% filter(Has.depdx == FALSE) %>% filter(On.Antidepressants == FALSE) %>% filter(PHQ.2_zero == TRUE).
+Double checked the logic and it looks okay
+- I think it has something to do with when I include people if they have a positive PHQ2 OR 9 (PHQ2 | PHQ9): Checked this, the | works just fine
+- Will need to make sure to write the draw io to match my mutate commands. Must be that when I do the filter separately, I am counting some 0s that I shouldn't be counting in phq2.

### 2022-08-13 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
graphics

@@
   - Depressed alone (46)
+
+#### Using volume rather than proportion to determine inside vs outside depression network
+It occurred to me that defining the top 25% depression network has been done by looking at proportion overlap, where everything else seems to work best in volume.
+It made me wonder what would happen if I defined my top 25% network based on the absolute volume of overlap a fascicle had with the depression network, and ranked based on that (rather than the % of fascicle that overlapped).
+It tests a slightly different hypothesis, where the depression network is though of as a collection of fascicles that have the most dead ending or overlap, i.e.
+contribute the most volume to the network.
+Proportion, on the other hand, thinks more about the specificity of the fascicle - i.e.
+if the majority of a fascicle dead ends in the network, its activity is probably highly specific to the network.
+In a volume only, the value of that fascicle's contribution would be neglibile.
+Using volume = house of representatives.
+using overlap - senate.
+
+Or put another way, if you do by volume, big fascicles get the most potential votes, by virtue of their bigness.
+It would be entirely possible for the AF, for example, to have some of the highest overlap with respect to volume, but if that volume is only 10% of the AF fascicle, it means that 90% dead end outside of the depression network.
+If a tiny fascicle is entirely hit, at least we know there is some specificity to its existence.
+Conversely, if the AF accounts for 99% of the volume overlap w/the depression network, even if it only 10% of the volume of the AF, it seems like it should be valued highly as part of the depression network.
+Luckily, results from both work out well.
+
+Results after defining the depresion 25% network based on the volume rather than proportion overlap: i.e.
+a fascicle is in the "depression network" if the volume that it shares with the depression network is in the top 25%.
+  - Irrespective of network, overall disease by volume: Dep > nondep (p=0.045). 
+  - Within dep network: dep disease > nondep disease i, p=0.0.034,
+  - Outside dep network (bottom 75%), dep disease ~= nondep disease, p = 0.071 (NS)
+
+dep_network_names_top_quartile_no_cranial_nerves_by_vol
+ [1] CC      AF_L    SLF2_R  SLF3_R  FAT_L   TR_S_R  SLF2_L  AF_R    TR_S_L  FAT_R   CS_S_R  CPT_F_R CPT_P_L ILF_R   CPT_P_R CS_S_L  PAT_R   MdLF_L  CBT_R
+
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol
+  [1] C_FPH_L C_FPH_R C_FP_L  C_FP_R  C_PH_L  C_PHP_L C_PHP_R C_PH_R  C_R_L   C_R_R   EMC_L   EMC_R   IFOF_L  IFOF_R  ILF_L   MdLF_R  PAT_L   SLF1_L  SLF1_R 
+[20] SLF3_L  UF_L    UF_R    VOF_L   VOF_R   CB_L    CB_R    ICP_L   ICP_R   MCP     SCP     V       AR_L    AR_R    CBT_L   CPT_F_L CPT_O_L CPT_O_R CS_A_L 
+[39] CS_A_R  CS_P_L  CS_P_R  CST_L   CST_R   DRTT_L  DRTT_R  F_L     F_R     ML_L    ML_R    OR_L    OR_R    RST_L   RST_R   TR_A_L  TR_A_R  TR_P_L  TR_P_R 
+[58] AC
 
@@
 df_good_mimosa_unique <- df_demo_and_fascicles %>% 
-#  filter(depGroupVar != 0) %>%
   group_by(EMPI) %>% 
@@
 make_demographics_table_ms(data_frame = df_good_mimosa_unique)
+
+df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() 
+make_demographics_table_ms(data_frame = df_good_mimosa_unique_dep_and_healthy)
 
@@
 barplot_lesion_volume <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
     geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
@@
   ylab("Volume of Lesion") +
-  theme(title = element_text(size=14),
+  theme(title = element_text(size=16),
         axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
         axis.title.y = element_text(size=20),
+        aspect.ratio = 5/5,
         axis.line = element_line(size = 1.5, colour = "black", linetype=1),
@@
 
-```
-
-```{r visualize_depression_network}
+barplot_lesion_volume_range <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
+  ylab("Volume of Lesion") +
+  ylim(0,20000) +
+  theme(title = element_text(size=16, hjust = 0.2),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        aspect.ratio = 5/5,
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+ # scale_y_continuous(expand = c(0, 0), limits = c(0, 20000)) #comment out if you want colored graph
+print(barplot_lesion_volume_range) #p=0.067
+
+##### violin plot
+
+#extract overall volume and diagnosis
+df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
+  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
+  filter(depGroupVar != 0) %>% 
+  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
+
+violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, color=depGroupVar)) +#, fill=depGroupVar)) + 
+  geom_violin() + #draw_quantiles = c(0.25, 0.5, 0.75)) + 
+  stat_summary(fun= "mean", geom = "pointrange", colour = "black") +
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
+  ylab("Volume of Lesions") +
+  theme(plot.title = element_text(size=16, hjust = 0.8),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+print(violin_lesion_volume) #p=0.067
+
+jitter_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, color=depGroupVar)) +#, fill=depGroupVar)) + 
+  geom_jitter(shape=16, position=position_jitter(0.15)) + 
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
+  ylab("Volume of Lesions") +
+  theme(plot.title = element_text(size=16, hjust = 0.8),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+print(jitter_lesion_volume) #p=0.067
+
+violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
+  geom_violin() + 
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
+  ylab("Volume of Lesions") +
+  theme(plot.title = element_text(size=16, hjust = 0.8),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+print(violin_lesion_volume) #p=0.067
+
+```
+
+```{r visualize_depression_network_proportion}
 ##########################################################
@@
 
+```{r visualize_depression_network_volume}
+##########################################################
+#                 Histos                        #
+##########################################################
+
+######## Histograms for depression network
+#sort based on overlap
+fascicle_values_dep_sorted_by_overlap <- fascicle_volumes_dep
+fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map[is.na(fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map)] = 0
+fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map)])
+
+#remove fascicles with cranial nerves
+fascicles_r_and_l_sep_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+plot_all_fascicles_r_and_l_sep_no_cranial_nerves <-ggplot(data = fascicles_r_and_l_sep_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
+  ggtitle("Volume Overlap per Fascicle 3_09") + 
+  geom_bar(stat="identity", fill="006300") +
+  xlab("Fascicle") +
+  ylab("Overlap with Depression Mask") +
+  ylim(0,40000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(plot_all_fascicles_r_and_l_sep_no_cranial_nerves)
+
+### remove all rows with cranial nerves in them
+# of note, we do not have all cranial nerves represented, we only have, CNII (optic), III(oculomotor), V (trigeminal), VII(facial), VIII(vestibulocochlear)
+fascicle_values_dep_no_cranial_nerves = fascicle_values_dep_sorted_by_overlap[(grep(pattern = "CN", x = fascicle_values_dep_sorted_by_overlap$fascicle, invert = TRUE)),]
+
+#figure out how many fascicles that is, then 
+quartile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 4),0)
+
+#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
+dep_network_names_top_quartile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile] #n=19
+
+#get a network that includes all of the fascicles NOT in dep 25% network
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)] #n=68
+
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol),]
+
+plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
+  geom_bar(stat="identity", fill="violet") +
+  ggtitle("Volume per Fascicle Affected Dep net 3_09, top quartile") + 
+  ylab("Volume in Depression Mask") +
+  ylim(0,40000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)
+
+#bottom 3 quartiles
+fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves<- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol),]
+
+plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
+  geom_bar(stat="identity", fill="darkgreen") +
+  ylab("Volume in Depression Mask") +
+  ylim(0,40000) +
+  ggtitle("Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs
+
+```
+
 ```{r proof_of_concept}
@@
 r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
-r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
+r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + 
+  geom_histogram(color="darkgreen") + 
+  facet_wrap(~variable)
 x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
@@
 
-#### look at mean fascicle volume across participants and transpose
+#### look at mean fascicle volume overlap across participants and transpose
 # drop exam date and empi, and transpose
@@
 
+write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")
+
 #sort based on overlap
@@
 
-sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
   ggtitle("Proportion of Volume per Fascicle Affected") + 
@@
   geom_bar(stat="identity") + 
+ # scale_fill_continuous(type="viridis") + 
+  scale_fill_gradient(low = "red", high = "yellow") +
   ylim(0,0.8) +
@@
         axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
 print(sorted_ggplot)
@@
 #############
-#### look at mean fascicle volume across healthy participants and transpose
+#### look at mean fascicle overlap across healthy participants and transpose
 #############
@@
 
+write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77_nondep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")
+
 #sort based on overlap
@@
 fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
-sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
   ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
   geom_bar(stat="identity") + 
-  ylab("Proportion") +
+  scale_fill_continuous(type="viridis") +
+  #scale_fill_gradient(low = "red", high = "yellow") +
+  ylab("Proportion Affected") +
   ylim(0,0.8) +
@@
         axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
 print(sorted_ggplot_healthy_ordered)
@@
 
+write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77_dep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")
+
 #sort based on overlap
@@
 fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
-sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
   ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
@@
   geom_bar(stat="identity") + 
+  #scale_fill_gradient(low = "red", high = "yellow") +
+  scale_fill_continuous(type="viridis") + #, color="orange") +
   ylim(0,0.8) +
@@
         axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(sorted_ggplot_depressed_ordered)
+
+### color by quartile
+#num_buckets = round(length(fascicle_values_sorted_by_overlap$fascicle)/4,0) #get the number of quartiles
+#color_columns= rep(x = c("#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072"), times=1, each=num_buckets) #assign each quartile a color
+#color_columns= rep(x = c("violet", "blue", "green", "red"), times=1, each=num_buckets) #assign each quartile a color
+
+#if (length(fascicle_values_sorted_by_overlap$fascicle) %% 4 != 0){
+ # color_columns=c("#8DD3C7", color_columns) #add one more to the beginning of the vector since the number is odd, so mod is odd number
+#}
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
+sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
+  geom_bar(stat="identity") + 
+  scale_fill_continuous(type = "viridis") + 
+  ylab("Proportion Affected") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
 print(sorted_ggplot_depressed_ordered)
@@
   ggtitle("Proportion of Volume per Fascicle Affected Dep net, top quartile") + 
-  geom_bar(stat="identity") +
+  geom_bar(stat="identity", fill="darkblue") +
   ylab("Proportion") +
@@
         axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
 print(plot_all_fascicles_r_and_l_sep_top_quartile)
@@
   ylab("Proportion") +
-  geom_bar(stat="identity") +
+  geom_bar(stat="identity", fill="red") +
   ylim(0,0.8) +
@@
         axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
 print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)
 
+############ Look at mean overall volume across all brain regions
+#### look at mean fascicle volume overlap across participants and transpose
+# drop exam date and empi, and transpose
+fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full[,(colnames(df_demo_and_fascicle_volumes_full) %in% fascicle_names_no_cranial_nerves)])
+fascicle_names <- rownames(fascicles_alone_df_vol)
+
+#take row means
+fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
+names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")
+
+#GET RID OF CRANIAL NERVES
+fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans
+
+#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
+order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
+  ggtitle("Volume per Fascicle Affected") + 
+  ylab("Volume") + 
+  geom_bar(stat="identity") + 
+  scale_fill_continuous(type="viridis") + 
+  ylim(0,230000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(sorted_ggplot)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
+  ggtitle("Volume per Fascicle Affected, exclude CC") + 
+  ylab("Volume") + 
+  geom_bar(stat="identity") + 
+  scale_fill_continuous(type="viridis") + 
+  ylim(0,30000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(sorted_ggplot)
+
+##### healthy volume
+df_demo_and_fascicle_volumes_full_nondep <- df_demo_and_fascicle_volumes_full %>% filter(depGroupVar==1)
+
+# drop exam date and empi, and transpose
+fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full_nondep[,(colnames(df_demo_and_fascicle_volumes_full_nondep) %in% fascicle_names_no_cranial_nerves)])
+fascicle_names <- rownames(fascicles_alone_df_vol)
+
+#take row means
+fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
+names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")
+
+#GET RID OF CRANIAL NERVES
+fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77_nondep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans
+
+#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
+order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
+  ggtitle("Volume per Fascicle Affected Nondepressed") + 
+  ylab("Volume") + 
+  geom_bar(stat="identity") + 
+  scale_fill_continuous(type="viridis") + 
+  ylim(0,230000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(sorted_ggplot)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
+  ggtitle("Volume per Fascicle Affected Nondepressed, exclude CC") + 
+  ylab("Volume") + 
+  geom_bar(stat="identity") + 
+  scale_fill_continuous(type="viridis") + 
+  ylim(0,30000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(sorted_ggplot)
+
+##### depressed volume
+# drop exam date and empi, and transpose
+
+df_demo_and_fascicle_volumes_full_dep <- df_demo_and_fascicle_volumes_full %>% filter(depGroupVar==2)
+
+fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full_dep[,(colnames(df_demo_and_fascicle_volumes_full_dep) %in% fascicle_names_no_cranial_nerves)])
+fascicle_names <- rownames(fascicles_alone_df_vol)
+
+#take row means
+fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
+names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")
+
+#GET RID OF CRANIAL NERVES
+fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77_dep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans
+
+#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
+order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
+  ggtitle("Volume per Fascicle Affected Depressed") + 
+  ylab("Volume") + 
+  geom_bar(stat="identity") + 
+  scale_fill_continuous(type="viridis") + 
+  ylim(0,230000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(sorted_ggplot)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
+  ggtitle("Volume per Fascicle Affected Depressed, exclude CC") + 
+  ylab("Volume") + 
+  geom_bar(stat="identity") + 
+  scale_fill_continuous(type="viridis") + 
+  ylim(0,30000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(sorted_ggplot)
 ```
@@
 barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Mean volume of Fascicles Affected") +
+  ggtitle(paste0("Total Volume Affected by dx, p=", round(total_volume_lost_dep_vs_healthy_unique$p.value,3))) + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 16),
+        aspect.ratio=5/5,
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000)) #comment out if you want colored graph
+
+print(barplot_overall)
+
+#in top quartile
+
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+
+# Grouped
+
+barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Mean volume of Fascicles Affected") +
+    ggtitle(paste0("Total Volume Affected by dx in Dep 25%, p=", round(total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 16), 
+        aspect.ratio=5/5,
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000)) 
+print(barplot_top_quartile)
+
+#in bottom 3 quartiles
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+
+# Grouped
+
+barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Mean volume of Fascicles Affected") +
+    ggtitle(paste0("Total Volume Affected by dx, Bottom 75% p=", round(total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 16), 
+        aspect.ratio=5/5,
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000))
+
+print(barplot_bottom_3_quartiles)
+
+######## Violin plots
+
+df_mean_fascicle_volumes_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
+dplyr::select(total_fascicle_vol_lost, total_fascicle_vol_lost_indepnet_top_quartile, total_fascicle_vol_lost_nondep_bottom_3_quartiles, depGroupVar) %>%
+  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
+
+jitter_mean_fascicle_volume <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost, color=depGroupVar)) +
+  geom_jitter(shape=16, position=position_jitter(0.15)) + 
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected by Diagnosis")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,40000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(jitter_mean_fascicle_volume) #p=0.067
+
+jitter_mean_fascicle_volume_dep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_indepnet_top_quartile, color=depGroupVar)) +
+  geom_jitter(shape=16, position=position_jitter(0.15)) + 
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected Dep Net by Diagnosis")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,40000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(jitter_mean_fascicle_volume_dep) #p=0.067
+
+jitter_mean_fascicle_volume_nondep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_nondep_bottom_3_quartiles, color=depGroupVar)) +
+  geom_jitter(shape=16, position=position_jitter(0.15)) + 
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected Nondep Net by Diagnosis")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,40000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(jitter_mean_fascicle_volume_nondep) #p=0.067
+
+##### Geom violin
+
+violin_mean_fascicle_volume <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost, fill=depGroupVar)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected by Diagnosis")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,42000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_mean_fascicle_volume) #p=0.067
+
+violin_mean_fascicle_volume_dep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_indepnet_top_quartile, fill=depGroupVar)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected Dep Net by Diagnosis")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,42000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_mean_fascicle_volume_dep) #p=0.067
+
+violin_mean_fascicle_volume_nondep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_nondep_bottom_3_quartiles,  fill=depGroupVar)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected Nondep Net by Diagnosis")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,42000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_mean_fascicle_volume_nondep) 
+```
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_dep_net_by_volume}
+
+#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#read in fiber volume values
+#fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+#df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+#df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+#df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+ #                 STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves_by_vol))) %>%
+  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
+
+#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
+
+#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
+total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
+
+############
+#bar plot
+############
+
+#over all fascicles
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Mean volume of Fascicles Affected") +
+  ggtitle(paste0("Total Volume Affected by dx, p=", round(total_volume_lost_dep_vs_healthy_unique$p.value,3))) + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 14), 
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) #comment out if you want colored graph
+
+print(barplot_overall)
+
+#in top quartile
+
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+
+# Grouped
+
+barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Mean volume of Fascicles Affected") +
+    ggtitle(paste0("Total Volume Affected by dx in Dep 25%, p=", round(total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 14), 
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+  #scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) 
+print(barplot_top_quartile)
+
+#in bottom 3 quartiles
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+
+# Grouped
+
+barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Mean volume of Fascicles Affected") +
+    ggtitle(paste0("Total Volume Affected by dx, Bottom 75% p=", round(total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 14), 
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000))
+
+print(barplot_bottom_3_quartiles)
+
+```
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_dep_network_fascicles_determined_by_vol}
+
+#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#read in fiber volume values
+#fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+#df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+#df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+#df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+ #                 STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
+
+#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
+
+#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
+total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
+
+############
+#bar plot
+############
+
+#over all fascicles
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
   geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
@@
 ```
-
 ```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop}
@@
 
-```
-
-```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume_no_CC}
-
-#define the depression network (all areas that intersect with depression network, no cranial nerves)
-fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
-
-#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
-fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
-
-#######################
-### depression map  ###
-#######################
-
-### define depression map
-fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
-
-#filter to exclude fascicles w/cranial nerves
-fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
-  replace(is.na(.), 0) %>%
-  filter(fascicle%in% fascicle_names_no_cranial_nerves)
-
-df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
-
-########################
-## Get Volume Values ###
-########################
-
-#read in fiber volume values
-fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
-
-#extract data frame that only includes the fascicle proportion lost
-df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
-df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
-
-#multiply each row by the vector of total volumes
-df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
-
-#combine back with data_fram
-df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
-
-#isolate out only the depressed group and healthy group, fascicles should be in volumes rather than proportions
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-# lapply to get results of t-test in each fascicle
-#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
-fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
-{
-  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
-  eval(parse(text=t_test_command))
-})
-names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves
-
-#pull t-statistic
-dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)
-
-#convert to data frame
-df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
-row.names(df_t_values) <- fascicle_names_no_cranial_nerves
-
-#pull p-values
-dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)
-
-#convert to data frame
-df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
-row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves
-
-#FDR correct p-values
-dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
-
-#Convert to data frame
-df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
-row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves
-
-#######################
-### plot differences ##
-#######################
-
-#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
-
-df_fascicle_dep_mask_volume_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
-names(df_fascicle_dep_mask_volume_t_value) <- c("fascicle", "dep_vs_healthy_t", "volume_in_dep_mask")
-
 #remove corpus callosum
@@
 
-#statistic volume, p = 0.031
+#statistic volume, 
 dep_overlap_x_dep_vs_healthy_t_no_cc <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value_no_cc)
@@
 
-```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_vol_effect_size_no_CC}
+Effect size in volume, w /cc (eff size NOT done in proportions)
+
+```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_vol_effect_size}
 # the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups
@@
 
-df_fascicle_dep_mask_volume_overlap_d_value_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>%
-  filter(fascicle != "CC")
-
 #make copy to sort
-df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask <- df_fascicle_dep_mask_volume_overlap_d_value_no_cc
+df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask <- df_fascicle_dep_mask_volume_overlap_d_value
 
@@
 # bar graph of eff size of fascicles, sorted by depression overlap
-plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network, no CC") +
+plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") +
   geom_bar(stat="identity") +
@@
 #statistic, p=001
-dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc )
+dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )
 print(summary(dep_overlap_x_dep_vs_healthy_d))
 
-scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc , aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
+scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
   geom_point(size=1) + 
@@
   xlab("Volume of Fascicle in Depression Network") + 
-  ylab("Effect of diagnosis calculated in Volumes (D)") + 
-  ggtitle(paste0("Fascicle/depression mask Volume overlap by dx Eff Size, No CC p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
-  theme(plot.title=element_text(hjust=0.5))
+  ylab("Effect Size (D)") + 
+  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
+#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
+  theme(plot.title=element_text(hjust=0.5),
+        axis.text = element_text(size=12, color = "black"),
+        axis.title = element_text(size=14, color = "black"),
+        axis.line = element_line(color="black", size=1.3),
+        panel.background = element_blank())
 print(scatter_d_val_by_dep_net_overlap)
 
-```
-
-```{r bootstrapping_proportion_and_volume_x_dep_t_analyses}
+#no p value
+scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect Size (D)") + 
+#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
+  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
+  theme(plot.title=element_text(hjust=0.5),
+        axis.text = element_text(size=12, color = "black"),
+        axis.title = element_text(size=14, color = "black"),
+        axis.line = element_line(color="black", size=1.3),
+        panel.background = element_blank())
+print(scatter_d_val_by_dep_net_overlap)
+
+#no corpus callosum
+df_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>% filter(fascicle != "CC")
+
+dep_overlap_x_dep_vs_healthy_vol_d_no_cc <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_no_cc )
+print(summary(dep_overlap_x_dep_vs_healthy_d))
+
+scatter_d_val_by_dep_net_overlap_no_cc <- ggplot(data=dep_overlap_x_dep_vs_healthy_vol_d_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect Size (D)") + 
+ ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d_no_cc)$coefficients[2,4],3))) + 
+#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, excluding CC")) + 
+  theme(plot.title=element_text(hjust=0.5),
+        axis.text = element_text(size=12, color = "black"),
+        axis.title = element_text(size=14, color = "black"),
+        axis.line = element_line(color="black", size=1.3),
+        panel.background = element_blank())
+print(scatter_d_val_by_dep_net_overlap_no_cc)
+
+#no p value
+scatter_d_val_by_dep_net_overlap_no_cc <- ggplot(data=dep_overlap_x_dep_vs_healthy_vol_d_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect Size (D)") + 
+# ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d_no_cc)$coefficients[2,4],3))) + 
+ ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, excluding CC")) + 
+  theme(plot.title=element_text(hjust=0.5),
+        axis.text = element_text(size=12, color = "black"),
+        axis.title = element_text(size=14, color = "black"),
+        axis.line = element_line(color="black", size=1.3),
+        panel.background = element_blank())
+print(scatter_d_val_by_dep_net_overlap_no_cc)
+
+```
+
+```{r bootstrapping_proportion_and_volume_x_dep_t_analyses_and_eff_sizes}
 
@@
 # 95% confidence intervals for the correlation
+
+data_frame_dep_v_nondep_d_x_vol_dep_mask_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>%
+  filter(fascicle != "CC") %>%
+  dplyr::select("dep_vs_healthy_d", "volume_in_dep_mask")
+
 set.seed(1)
-results_vol <- boot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc, statistic=tval,
+results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask_no_cc, statistic=tval,
    R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)

### 2022-08-13 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
 - Will need to make sure to write the draw io to match my mutate commands. Must be that when I do the filter separately, I am counting some 0s that I shouldn't be counting in phq2.
+
+Checks
+- x <-data_empi_acc_f_phq %>% filter(!Has.depdx) %>% filter(!On.Psych.Meds) %>% filter(PHQ.2_zero | PHQ.9_zero)
+- length(unique(x$EMPI)) = 488
+- length(data_empi_acc_f_phq$EMPI[data_empi_acc_f_phq$true_healthy== 1]) = 2608
+- length(unique(data_empi_acc_f_phq$EMPI[data_empi_acc_f_phq$true_healthy== 1])) = 488
+
+- maybe some people with phq2_0 are actually Na?
+- x <-data_empi_acc_f_phq %>% filter(!Has.depdx) %>% filter(!On.Psych.Meds) %>% filter(PHQ.2_zero==1) = 2595, 485 unique
+- x <-data_empi_acc_f_phq %>% filter(!Has.depdx) %>% filter(!On.Psych.Meds) %>% filter(PHQ.9_zero==1) = 13, 3 unique
+- 13+2595 = 2608; 485+3 = 488
+- these all work out
+
+- so is it if you take unique first, some columns are missing? but why would that give you more phq2s that are okay
+
+- Maybe it is merge?
+First, from ?
+merge : The rows in the two data frames that match on the specified columns are extracted, and joined together.
+If there is more than one match, all possible matches contribute one row each.
+< this should be handled when I take unique later>
+
+- seems to be working now when I call unique at the end. DOuble check what happens when I call unique at the beginning.
+
+- tried getting unique values in the data_empi_acc_f_phq before merging w/ lesion volumes:
+data_empi_acc_f_phq <- data %>% 
+   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
+   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
+   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
+   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
+   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
+   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
+   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
+   mutate(BEGIN_EXAM_DTTM = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% #make sure to actually save it as date object for later processing
+   mutate(EMPI = as.factor(EMPI)) %>%
+
+  ###
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() %>%
+  ###
+   mutate(On.Psych.Meds = ifelse(On.Psych.Meds == "True", 1, 0)) %>%
+   mutate(On.Antidepressants = ifelse(On.Antidepressants == "True", 1, 0)) %>%
+   mutate(Has.PHQ2 = ifelse(Has.PHQ2 == "True", 1, 0)) %>%
+   mutate(Has.PHQ9 = ifelse(Has.PHQ9 == "True", 1, 0)) %>%
+   mutate(Has.depdx = ifelse(grepl("F3", ICD10), 1, 0)) %>%
+   mutate(PHQ.2_modsev_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 >= 3), 1, 0)) %>%
+   mutate(PHQ.9_modsev_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 >= 10), 1, 0)) %>%
+   mutate(PHQ.2_mild_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 > 0 & PHQ.2 < 3), 1, 0)) %>%
+   mutate(PHQ.9_mild_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 > 0 & PHQ.9 < 10), 1, 0)) %>%
+   mutate(PHQ.2_zero = ifelse((!is.na(PHQ.2) & PHQ.2 == 0), 1, 0)) %>%
+   mutate(PHQ.9_zero = ifelse((!is.na(PHQ.9) & PHQ.9 == 0), 1, 0)) %>%
+   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
+   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
+   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
+   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
+   rowwise(ACCESSION_NUM) %>%
+   mutate(depGroupVar =
+            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
+   ungroup()
+
+-Did not work because it took out people who didn't have good mimosas, so by the time you merge w/good mimosas, only 158 scans left.
+
+- Now going to look at the empis that make it to the 168 group and 148 group
+
+x <- df_demo_and_fascicles %>% filter(Has.depdx==FALSE) %>% filter(On.Antidepressants == FALSE) %>% filter(PHQ.2_zero == TRUE) %>% group_by(EMPI) %>% 
++     arrange(EXAM_DATE) %>% 
++     slice(1) %>% 
++     ungroup() 
+> dim(x)
+[1] 168 159
+> x_empi <- x$EMPI
+
+y_empi <- df_good_mimosa_unique$EMPI[df_good_mimosa_unique$true_healthy==TRUE]
+> length(y_empi)
+[1] 148
+
+> z_empi <- setdiff(x_empi,y_empi)
+> z_empi
+ [1] "1000443363" "1000576093" "1000616211" "1000680873" "1000731479" "1000758897" "1000782730" "1001091352" "1001387420" "1001760723" "1001776245" "1003146950"
+[13] "1003642810" "1003709298" "1003867245" "1003873441" "1004564376" "1004566947" "1004641094" "8000342932" "8440021590"
+> length(z_empi)
+[1] 21
+
+> df_demo_and_fascicles$depGroupVar[df_demo_and_fascicles$EMPI == 1000443363]
+[1] 0
+> df_demo_and_fascicles$On.Antidepressants[df_demo_and_fascicles$EMPI == 1000443363]
+[1] 0
+> df_demo_and_fascicles$Has.depdx[df_demo_and_fascicles$EMPI == 1000443363]
+[1] 0
+> df_demo_and_fascicles$PHQ.2_zero[df_demo_and_fascicles$EMPI == 1000443363]
+[1] 1
+\> df_demo_and_fascicles$PHQ.9[df_demo_and_fascicles$EMPI == 1000443363]
+[1] NA
+
+AH! I figured it out. I didn't filter out people who are ALSO NOT ON PSYCH MEDS. My mistake in that line was not including not on psych meds.
+
+x <- df_demo_and_fascicles %>% filter(Has.depdx==FALSE) %>% filter(On.Psych.Meds== FALSE) %>% filter(PHQ.2_zero == TRUE) %>% group_by(EMPI) %>%  arrange(EXAM_DATE) %>%   slice(1) %>% ungroup
+> dim(x)
+[1] 147 159
+
+GOT IT !!!!
+
+#### 20220808
+
+Another strange idea that is not needed right now, but what about defining depression even more data driven.
+
+Tract filtration done normally.
+For each tract, instead of looking at its overalp in proportion or volume in depression network, do some sort of cortical parcellation (schaefer for example), and calculate the volume or proportion of overlap of each full tract to the parcellation.
+calculate prop overlap of each individual fiber to its overlap with each of schaefer regions (so like 200 * 77 matrix).
+Map to yeo networks, compare between healthy and control.
+Or you can hydra on the yeo networks, or w/in outside depression network, or just look at indiv fascicles.
+This actually isn't that thought out.
+Or some sort of spin for network correspondence between network structure in depression as opposed to controls.
+
+[x] look at what happens when you define depression diagnosis by volume rather than proportion overlap.
+I.e.
+a fascicle is in the "depression network" if the volume that it shares with the depression network is in the top 25%.
+In this situation, we still see that overall, depressed have more disease (i.e.
+more volume affected) irrespective of dx (p=0.045).
+  [x] Within dep network (as defined by volume), 
+  [x] dep disease > nondep disease in dep network, p=0.0.034,
+  [x] dep disease and nondep disease in bottom 75%, p = 0.071 (NS)
+
+dep_network_names_top_quartile_no_cranial_nerves_by_vol
+ [1] CC      AF_L    SLF2_R  SLF3_R  FAT_L   TR_S_R  SLF2_L  AF_R    TR_S_L  FAT_R   CS_S_R  CPT_F_R CPT_P_L ILF_R   CPT_P_R CS_S_L  PAT_R   MdLF_L  CBT_R
+
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol
+  [1] C_FPH_L C_FPH_R C_FP_L  C_FP_R  C_PH_L  C_PHP_L C_PHP_R C_PH_R  C_R_L   C_R_R   EMC_L   EMC_R   IFOF_L  IFOF_R  ILF_L   MdLF_R  PAT_L   SLF1_L  SLF1_R 
+[20] SLF3_L  UF_L    UF_R    VOF_L   VOF_R   CB_L    CB_R    ICP_L   ICP_R   MCP     SCP     V       AR_L    AR_R    CBT_L   CPT_F_L CPT_O_L CPT_O_R CS_A_L 
+[39] CS_A_R  CS_P_L  CS_P_R  CST_L   CST_R   DRTT_L  DRTT_R  F_L     F_R     ML_L    ML_R    OR_L    OR_R    RST_L   RST_R   TR_A_L  TR_A_R  TR_P_L  TR_P_R 
+[58] AC
+
+I actually think we should use proportions for this, because if we define depression just based on volume, which does work, it is never possible that the volume of small fascicles will contribute meaningfully, even though they may play a big role.
+Additionally, if you just use volume, you could get a fiber which donates a ton to depression network by volume, but half of that fascicle provides information to outside the depression network.
+It seems a little more logical to define values as within the depression network based on how much of the total fascicle dead ends into the depression network as compared to its total volume.
+But luckily the findings are the same, irrespective of whether you define the 25% based on volume or proportion.
+
+One could also make a case that even if, lets say, only 30% of the AF lands in the depression network, but the 99% of the depression network is occupied by AF streamlines, then it should be ranked really high, as opposed to a tiny fascicle where the whole thing lands in the depression network, but it only accounts for 1% of the network.
+And it is just a good question of what matters.
+Thank goodness they both work.
+
+Making DSI studio tracts
+- if it doesn't work to add the depression network as roi, make sure to add the asym template from the harvard depression network first, and then the best mask to use is Dep_clust_T_3_09_binarized.nii.
+
+##### 20220810
+[] do proportion effect size plot if you'd like, not done right now
+[] Figures
+  [] Overall fascicles, highlight optic radiations, etc, use viridis and cool colors
+    - have to do volume, not just proportion.
+I am thinking - a viridis of some sort to show overall volume of fascicles across people, then a bar to show top chunk, and separate out a graph of top 25% and bottom 75%.
+    - Could also find a way to show volume in each fascicle in dep vs non
+    - Going to also try in afni to multiple each fascicle by its count, sum them up, and display over template brain. 
+```{bash}
+# go to directory
+cd /Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography
+
+#make new direc if it doesn't exist
+mkdir -p niftis_excluding_cranial_nerves
+
+#copy non-cranial nerves
+cp association/*nii* cerebellum/*nii* commissural/*nii* projection/*nii* niftis_excluding_cranial_nerves/.
+
+#cd into directory
+cd niftis_excluding_cranial_nerves
+
+#read in file with proportions
+
+```
+
+```{bash make_afni_map_w_fascicles_colored_by_prop_or_volume}
+#!/bin/bash
+
+##### Welcome to make_afni_map_w_fascicles_colored_by_prop_overlap ###
+### Pre: A file that has two columns, no headers. First column is the abbreviation of a fascicle, second is proportion overlap (or volume overlap)
+### Post: A nifti where each fascicle is colored by the proportion of overlap, or any value you give it
+### Uses: I wanted to find a way to show, in a brain, which fascicles have the most overlap with lesions (though this can be done w/depression, non-dep, volume).
+This will go through a table, and for each fascicle, multiply it times the value in the second column, add a suffix, and then at the end, sum them all to make a final image
+### Dependencies: Afni
+
+#change this if doing on pmacs
+homedir="/Users/eballer/BBL/msdepression/"
+
+default=${homedir}/results/fascicle_mean_proportion_overlap_nfasc_77.csv
+default_suffix="prop"
+
+if [ $# == 0 ]
+then
+    echo "We will use the default path file" $default
+    fascicle_file=$default
+else  
+    fascicle_file=$1
+fi
+
+echo "File being read is "$fascicle_file
+
+fascicles=$(cat $fascicle_file)
+
+cd ${homedir}/templates/dti/HCP_YA1065_tractography/niftis_excluding_cranial_nerves
+
+(pwd)
+for fascicle in ${fascicles}; do
+        # extract fascicle name and value
+        fascicle_name=$(echo ${fascicle} | perl -pe 's/(.*),.*/$1/')
+        fascicle_value=$(echo ${fascicle} | perl -pe 's/.*,(.*)/$1/')
+
+        # remove nifti w/values if already there
+        rm -rf ${fascicle_name}_${default_suffix}.nii.gz
+
+        #take the fascicle .nii, multiply it with the value
+        command=$(echo 3dcalc -a ${fascicle_name}.nii.gz -expr \'a*${fascicle_value}\' -prefix ${fascicle_name}_${default_suffix}.nii.gz)
+        echo "Here is the command: "${command}
+
+        #run command, have to do separately to get variable into afni command
+        eval "${command}";
+done
+
+#remove old file
+rm -rf fascicle_${default_suffix}_sum.nii.gz
+rm -rf fascicle_${default_suffix}_max.nii.gz
+
+all_files=$(ls *${default_suffix}.nii.gz)
+
+#take sum
+command_sum=$(3dMean -prefix fascicle_${default_suffix}_sum.nii.gz -sum ${all_files})
+eval "${command_sum}";
+
+#take max
+command_max=$(3dMean -prefix fascicle_${default_suffix}_max.nii.gz -max ${all_files})
+eval "${command_max}";
+
+cd ${homedir}/scripts
+```
+
+  [] Age effects perhaps
+  [] Rank of fascicles based on overlap in proportion (and or volume, or both) w/ depression network
+  [x] lesion overall burden
+  [x] overall fascicle comparison, broken down by 25/75 % (bar graphs)
+  [x] tract scatter plot
+
+From Phil, how to color the fascicles in dsi studio
+
+**This might well be a suboptimal solution, but you can write a text file with one RGB triplet per fiber.
+In the GUI this is done from the "Tracts Misc" menu "save tract colors".
+From the command line you'd need to count the tracts in the file and then write the text out in a script.
+You can use the command line to open the GUI with the tracts and associated color files.
+It might be worth asking on the DSI Studio user group if nobody else here has a better solution.
+Frank is very responsive and adds improvements all the time, so my convoluted way might already be out of date**
+
+#### 20220812
+- added code to look at mean volume (not just proportion) over all fascicles
+- looking at the maps w/volume, really hard to see differences b/c CC is just so much bigger than everything else
+- for funsies, did a subtraction between nondep/dep prop and volume masks - these give cool pictures and really make it clear that there is more right sided disease
+```{bash}
+3dcalc -a fascicle_prop_dep_max.nii.gz -b fascicle_prop_nondep_max.nii.gz -expr 'a-b' -prefix dep_gr_nondep_prop_max.nii.gz
+3dcalc -a fascicle_prop_dep_sum.nii.gz -b fascicle_prop_nondep_sum.nii.gz -expr 'a-b' -prefix dep_gr_nondep_prop_sum.nii.gz
+3dcalc -a fascicle_vol_dep_max.nii.gz -b fascicle_vol_nondep_max.nii.gz -expr 'a-b' -prefix dep_gr_nondep_vol_max.nii.gz
+3dcalc -a fascicle_vol_dep_sum.nii.gz -b fascicle_vol_nondep_sum.nii.gz -expr 'a-b' -prefix dep_gr_nondep_vol_sum.nii.gz
+```
+
+Steps to do the dsi studio things
+- Design color system where each proportion is mapped to an RGB scheme
+  - Okay, so red is 255, 0, 0
+  - Yellow is 255,255,0
+  - So to get the scale I want, I need to only alter the 255. 
+  - If I want to do this in proportions, it means that there need to be 100 increments, each corresponding to a proportion
+  - So 255/100, means every increment of 2.55 will equal 1 increment of proportion.
+So, to get the green color, take the proportion, multiply by 100 to get it into whole numbers, multiply * 2.55, round
+  - i.e. if the proportion of overlay is 0.02, that means that it is 2% total. 2% * 2.55 = 5.1 -> round, 5. New RGB color is 255 5 0
+
+- For each fascicle, count the # of streamlines. This will become number of rows in color file
+- For each fascicle, make a file in the style Superior_Longitudinal_Fasciculus3_R_color.txt, with each row containing the RGB color scheme, space separated, with # of rows corresponding to number of fibers in fascicle
+- alternatively, could have two types of files, file 1 with fascicle/rgb mapping, file type 2 would be each fascicle with color written out.
+Then could use perl to extra the 3 items, replace them, write to new file, and then overwrite initial file.
+
+- looking for command line to get count of each file name
+- dsi_studio --action=rec --source=*.src.gz --rotate_to=dwi_sum.nii.gz --save_nii=*_aligned.nii.gz
+
+- to get fascicle full names, I went to DSI github (https://github.com/frankyeh/DSI-Studio/issues/72), and found this list:
+
+Arcuate_Fasciculus_L Arcuate_Fasciculus_R Cingulum_Frontal_Parahippocampal_L Cingulum_Frontal_Parahippocampal_R Cingulum_Frontal_Parietal_L Cingulum_Frontal_Parietal_R Cingulum_Parahippocampal_Parietal_L Cingulum_Parahippocampal_Parietal_R Cingulum_Parahippocampal_L Cingulum_Parahippocampal_R Cingulum_Parolfactory_L Cingulum_Parolfactory_R Frontal_Aslant_Tract_L Frontal_Aslant_Tract_R Inferior_Fronto_Occipital_Fasciculus_L Inferior_Fronto_Occipital_Fasciculus_R Inferior_Longitudinal_Fasciculus_L Inferior_Longitudinal_Fasciculus_R Middle_Longitudinal_Fasciculus_L Middle_Longitudinal_Fasciculus_R Parietal_Aslant_Tract_L Parietal_Aslant_Tract_R Superior_Longitudinal_Fasciculus1_L Superior_Longitudinal_Fasciculus1_R Superior_Longitudinal_Fasciculus2_L Superior_Longitudinal_Fasciculus2_R Superior_Longitudinal_Fasciculus3_L Superior_Longitudinal_Fasciculus3_R Uncinate_Fasciculus_L Uncinate_Fasciculus_R Vertical_Occipital_Fasciculus_L Vertical_Occipital_Fasciculus_R Corticospinal_Tract_L Corticospinal_Tract_R Corticostriatal_Tract_Anterior_L Corticostriatal_Tract_Anterior_R Corticostriatal_Tract_Posterior_L Corticostriatal_Tract_Posterior_R Corticostriatal_Tract_Superior_L Corticostriatal_Tract_Superior_R Thalamic_Radiation_Anterior_L Thalamic_Radiation_Anterior_R Thalamic_Radiation_Posterior_L Thalamic_Radiation_Posterior_R Thalamic_Radiation_Superior_L Thalamic_Radiation_Superior_R Fornix_L Fornix_R Optic_Radiation_L Optic_Radiation_R Reticular_Tract_L Reticular_Tract_R Corpus_Callosum_Forceps_Minor Corpus_Callosum_Body Corpus_Callosum_Tapetum Corpus_Callosum_Forceps_Major
+
+I stuck it in a file and used a perl command to make it look pretty
+
+```{bash}
+cd /Users/eballer/BBL/msdepression/dti/HCP_YA1065_tractography
+vim fascicle_full_names
+
+#copied values in there and closed out
+more fascicle_full_names | perl -pe 's/ /\n/g' > fascicle_full_names_formatted
+
+```
+
+Now looks like this (but only 56 fascicles), so not right list:
+Arcuate_Fasciculus_L
+Arcuate_Fasciculus_R
+Cingulum_Frontal_Parahippocampal_L
+Cingulum_Frontal_Parahippocampal_R
+Cingulum_Frontal_Parietal_L
+Cingulum_Frontal_Parietal_R
+Cingulum_Parahippocampal_Parietal_L
+Cingulum_Parahippocampal_Parietal_R
+Cingulum_Parahippocampal_L
+Cingulum_Parahippocampal_R
+Cingulum_Parolfactory_L
+Cingulum_Parolfactory_R
+Frontal_Aslant_Tract_L
+Frontal_Aslant_Tract_R
+Inferior_Fronto_Occipital_Fasciculus_L
+Inferior_Fronto_Occipital_Fasciculus_R
+Inferior_Longitudinal_Fasciculus_L
+Inferior_Longitudinal_Fasciculus_R
+Middle_Longitudinal_Fasciculus_L
+Middle_Longitudinal_Fasciculus_R
+Parietal_Aslant_Tract_L
+Parietal_Aslant_Tract_R
+Superior_Longitudinal_Fasciculus1_L
+Superior_Longitudinal_Fasciculus1_R
+Superior_Longitudinal_Fasciculus2_L
+Superior_Longitudinal_Fasciculus2_R
+Superior_Longitudinal_Fasciculus3_L
+Superior_Longitudinal_Fasciculus3_R
+Uncinate_Fasciculus_L
+Uncinate_Fasciculus_R
+Vertical_Occipital_Fasciculus_L
+Vertical_Occipital_Fasciculus_R
+Corticospinal_Tract_L
+Corticospinal_Tract_R
+Corticostriatal_Tract_Anterior_L
+Corticostriatal_Tract_Anterior_R
+Corticostriatal_Tract_Posterior_L
+Corticostriatal_Tract_Posterior_R
+Corticostriatal_Tract_Superior_L
+Corticostriatal_Tract_Superior_R
+Thalamic_Radiation_Anterior_L
+Thalamic_Radiation_Anterior_R
+Thalamic_Radiation_Posterior_L
+Thalamic_Radiation_Posterior_R
+Thalamic_Radiation_Superior_L
+Thalamic_Radiation_Superior_R
+Fornix_L
+Fornix_R
+Optic_Radiation_L
+Optic_Radiation_R
+Reticular_Tract_L
+Reticular_Tract_R
+Corpus_Callosum_Forceps_Minor
+Corpus_Callosum_Body
+Corpus_Callosum_Tapetum
+Corpus_Callosum_Forceps_Major

### 2022-08-13 Descriptives_potential_paper_1_20220622.Rmd
Add files via upload

--- 2022-07-20 msdepression/scripts/Descriptives_potential_paper_1_20220622.Rmd
@@
 {
-  gam(substitute(i ~ s(PAT_AGE_AT_EXAM), list(i = as.name(x))), data = df_unique_empi)
+  gam(substitute(i ~ s(PAT_AGE_AT_EXAM, k=4), list(i = as.name(x))), data = df_unique_empi)
 })
@@
 
+visreg(fascicle_age_gam$FAT_L)
 ```

### 2022-08-13 K23_prelim_data_using_dep_T_3_09_mask_20220630.Rmd
Add files via upload

--- 2022-07-20 msdepression/scripts/K23_prelim_data_using_dep_T_3_09_mask_20220630.Rmd
@@
 
+Depression top quartile, r&l separated
+ [1] "SLF3_R"  "AF_R"    "AF_L"    "SLF2_R"  "CBT_R"   "FAT_R"   "CS_P_R"  "C_PHP_R" "MdLF_L"  "FAT_L"   "SLF2_L"  "MdLF_R"  "CBT_L"   "TR_P_L"  "TR_S_R" 
+[16] "CPT_O_R" "TR_P_R"  "PAT_L"   "PAT_R"   "EMC_R"   "TR_S_L"  "CC"
+
+Depression top quartile, r&l meaned
+[1] AF    SLF3  SLF2  FAT   MdLF  CST   CPT_O PAT   C_PHP CS_A  CC    CPT_F
+
 Notebook Overview
@@
       - MS disease within depression 5% network (30%) versus outside (14%), p < 2.2 * 10^-16
+      - MS disease within depression top quartile network (31%) versus areas with no overlap (14%), p < 2.2 x 10^-16
+      - MS disease within depression top quartile network (31%) versus outside depression top quartile network (24%), p 1.7 * 10^-5
       - Linear Regressions (essentially matching t tests above): 
@@
          +dep vs healthy within non dep network (p = 0.042)
+         +dep vs healthy within dep top quartile (p = 0.054);
       - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis.
@@
 fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-
+fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
 #write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
@@
 
+fascicle_names_dep_network_all_no_cranial_nerves <- fascicle_names_dep_network_all[grep("CN", fascicle_names_dep_network_all, invert = TRUE)]
+
 #fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
@@
   geom_bar(stat="identity") +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        axis.text.y = element_text(size=2),
+        plot.title = element_text(hjust = 0.5))
 print(a)
@@
 
-plot_all_fascicles_r_and_l_sep <-ggplot(data = fascicle_values_dep_sorted_by_overlap, aes(x = fascicle, y=prop_in_mask), main = "Proportion of Volume per Fascicle Affected Dep net 3_09") + 
-  geom_bar(stat="identity") +
+plot_all_fascicles_r_and_l_sep <-ggplot(data = fascicle_values_dep_sorted_by_overlap, aes(x = fascicle, y=prop_in_mask))+
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion in Depression Mask") +
   ylim(0,0.4) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
 print(plot_all_fascicles_r_and_l_sep)
 
-#get top quartile
+fascicles_r_and_l_sep_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+plot_all_fascicles_r_and_l_sep_no_cranial_nerves <-ggplot(data = fascicles_r_and_l_sep_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask))+
+  ggtitle("Proportion of Volume Overlap per Fascicle 3_09") + 
+  geom_bar(stat="identity") +
+  xlab("Fascicle") +
+  ylab("Overlap with Depression Mask") +
+  ylim(0,0.4) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
+print(plot_all_fascicles_r_and_l_sep_no_cranial_nerves)
+
+#### get top quartile; 
+#figure out how many fascicles that is, then 
 quartile <- round((dim(fascicle_values_dep_sorted_by_overlap)[1] / 4),0)
-dep_network_names_top_quartile<- (fascicle_values_dep_sorted_by_overlap[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile]
+
+#quartile <- floor((dim(fascicle_values_dep_sorted_by_overlap)[1] / 4))
+
+#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
+dep_network_names_top_quartile_preproc<- (fascicle_values_dep_sorted_by_overlap[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile]
+
+dep_network_names_bottom_quartile_preproc<- (fascicle_values_dep_sorted_by_overlap[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask, decreasing = FALSE),])$fascicle[1:quartile]
+#make sure it right format, character
+
+dep_network_names_top_quartile <- as.character(dep_network_names_top_quartile_preproc)
+
+#get a network that includes all of the fascicles NOT in dep 25% network
+dep_network_bottom_3_quartiles <- fascicle_names[!(fascicle_names %in% dep_network_names_top_quartile_preproc)]
+dep_network_names_bottom_quartile <- fascicle_names[(fascicle_names %in% dep_network_names_bottom_quartile_preproc)]
+
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_25_percent_for_plot <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_preproc),]
+
+plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=prop_in_mask))+
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, top quartile") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion in Depression Mask") +
+  ylim(0,0.4) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_top_quartile)
+
+#bottom 3 quartiles
+fascicles_dep_bottom_3_quartiles_for_plot <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_bottom_3_quartiles),]
+
+plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=prop_in_mask))+
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom 3 quartiles") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.4) +
+  ylab("Proportion in Depression Mask") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)
+
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_bottom_quartile_for_plot <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_quartile_preproc),]
+
+plot_all_fascicles_r_and_l_sep_bottom_quartile <-ggplot(data = fascicles_dep_bottom_quartile_for_plot, aes(x = fascicle, y=prop_in_mask)) +
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom quartile") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.4) +
+  ylab("Proportion in Depression Mask") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_quartile)
+
+##################
+### Sensitivity analysis, repeat this all but exclude cranial nerves
+##################
+
+### remove all rows with cranial nerves in them
+# of note, we do not have all cranial nerves represented, we only have, CNII (optic), III(oculomotor), V (trigeminal), VII(facial), VIII(vestibulocochlear)
+fascicle_values_dep_no_cranial_nerves = fascicle_values_dep_sorted_by_overlap[(grep(pattern = "CN", x = fascicle_values_dep_sorted_by_overlap$fascicle, invert = TRUE)),]
+
+#figure out how many fascicles that is, then 
+quartile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 4),0)
+#quartile <- floor((dim(fascicle_values_dep_no_cranial_nerves)[1] / 4))
+
+#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
+dep_network_names_top_quartile_preproc_no_cranial_nerves<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile] #n=19
+
+dep_network_names_bottom_quartile_preproc_no_cranial_nerves<- (fascicle_values_dep_sorted_by_overlap[order(fascicle_values_dep_no_cranial_nerves$prop_in_mask, decreasing = FALSE),])$fascicle[1:quartile]
+#make sure it right format, character
+
+#make sure it right format, character
+
+dep_network_names_top_quartile_no_cranial_nerves <- as.character(dep_network_names_top_quartile_preproc_no_cranial_nerves) #n=19
+
+#get a network that includes all of the fascicles NOT in dep 25% network
+dep_network_bottom_3_quartiles_no_cranial_nerves <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_preproc_no_cranial_nerves)] #n=68
+
+#bottom quartile
+dep_network_names_bottom_quartile_no_cranial_nerves <- fascicle_names[(fascicle_names %in% dep_network_names_bottom_quartile_preproc_no_cranial_nerves)]
+
+#get a network than includes all of the fascicles with NO overlap with depression network
+nondep_network_names_no_cranial_nerves <- fascicle_values_dep_no_cranial_nerves$fascicle[fascicle_values_dep_no_cranial_nerves$prop_in_mask == 0] #n=15
+
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_preproc_no_cranial_nerves),]
+
+plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
+  geom_bar(stat="identity") +
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, top quartile") + 
+  ylim(0,0.4) +
+  ylab("Proportion in Depression Mask") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
+print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)
+
+#bottom 3 quartiles
+fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_bottom_3_quartiles_no_cranial_nerves),]
+
+plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
+  geom_bar(stat="identity") +
+  ylab("Proportion in Depression Mask") +
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
+  ylim(0,0.4) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs
+
+#bottom quartile
+fascicles_dep_bottom_quartile_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_quartile_preproc_no_cranial_nerves),]
+
+plot_all_fascicles_r_and_l_sep_bottom_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_quartile_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask), main = "Proportion of Volume per Fascicle Affected Dep net 3_09, bottom quartile") +
+  geom_bar(stat="identity") +
+  ylab("Proportion in Depression Mask") +
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom quartile") + 
+  ylim(0,0.4) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_quartile_no_cranial_nerves)
+
+#non dep network, no cranial nerves
+nondep_network_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap %>% filter(fascicle_values_dep_sorted_by_overlap$fascicle %in% nondep_network_names_no_cranial_nerves)
+
+plot_all_fascicles_nondep_network_for_plot_no_cranial_nerves  <-ggplot(data = nondep_network_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
+  geom_bar(stat="identity") +
+  ggtitle("Proportion of Volume per Fascicle Affected nonDep net 3_09") + 
+  ylab("Proportion in Depression Mask") +
+  ylim(0,0.4) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_nondep_network_for_plot_no_cranial_nerves)
 
@@
 
-sorted_ggplot<-ggplot(data = fascicle_values_dep_sorted_by_overlap, aes(x = fascicle, y=prop_in_mask), main = "Proportion of Volume per Fascicle Affected Dep net 3_09") + 
+sorted_ggplot<-ggplot(data = fascicle_values_dep_sorted_by_overlap, aes(x = fascicle, y=prop_in_mask)) + 
   geom_bar(stat="identity") +
   ylim(0,0.4) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  xlab("Fascicle") +
+  ylab("Proportion Overlap with Depression Mask") +
+  ggtitle("Proportion of Volume that Overlaps with Depression Network by Fascicle 3_09")
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
 print(sorted_ggplot)
@@
 
+#GET RID OF CRANIAL NERVES
+fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
 #sort based on overlap
@@
 
-sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected") + 
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected") + 
+  ylab("Proportion") + 
   geom_bar(stat="identity") +
   ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
 print(sorted_ggplot)
@@
 
-sorted_ggplot_healthy<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected in Non Depressed") + 
-  geom_bar(stat="identity") +
+sorted_ggplot_healthy<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected in Non Depressed") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion") +
   ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
 print(sorted_ggplot_healthy)
@@
 fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
-sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
-  geom_bar(stat="identity") +
+sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion") +
   ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
 print(sorted_ggplot_healthy_ordered)
@@
 
-sorted_ggplot_depressed<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected Depressed") + 
-  geom_bar(stat="identity") +
+sorted_ggplot_depressed<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Depressed") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion Affected") +
   ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
 print(sorted_ggplot_depressed)
@@
 fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
-sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value), main = "Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
+sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
+  ylab("Proportion Affected") +
   geom_bar(stat="identity") +
   ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
 print(sorted_ggplot_depressed_ordered)
+
+#no cranial nerves
+fascicles_for_plot_no_cranial_nerves <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% fascicle_names_dep_network_all_no_cranial_nerves),]
+
+plot_all_fascicles_r_and_l_no_cranial_nerves <-ggplot(data = fascicles_for_plot_no_cranial_nerves, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Net 3_09") + 
+  xlab("Fascicle") + 
+  ylab("Proportion Affected") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
+print(plot_all_fascicles_r_and_l_no_cranial_nerves )
+
+##### look at mean fascicles affected in dep vs non dep network
+######
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_25_percent_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_preproc),]
+
+plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, top quartile") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_top_quartile)
+
+#bottom 3 quartiles
+fascicles_dep_bottom_3_quartiles_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_bottom_3_quartiles),]
+
+plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
+  ylab("Proportion") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)
+
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_bottom_quartile_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_bottom_quartile_preproc),]
+
+plot_all_fascicles_r_and_l_sep_bottom_quartile <-ggplot(data = fascicles_dep_bottom_quartile_for_plot, aes(x = fascicle, y=value)) +
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom quartile") + 
+  ylab("Proportion") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_quartile)
+
+#non dep network
+fascicles_nondep_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% fascicle_names_nondep_network),]
+
+plot_all_fascicles_r_and_l_nondep <-ggplot(data = fascicles_nondep_for_plot, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected nonDep net 3_09") + 
+  ylab("Proportion") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_nondep)
+
+```
+
+```{r proof_of_concept_unique}
+
+#this section creates histograms looking at the proportion of volume lost in each fascicle across subjects
+#we recapitulate earlier work and show that higher volume loss with our method reflects regions previously known to have higher volume loss
+#These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis
+
+data_empi_acc_f_phq_unique <- data_empi_acc_f_phq %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup() #n=3737
+
+#have to pull out just the pieces I'm interested in
+just_dep_and_empi <- subset(data_empi_acc_f_phq_unique, select = c("EMPI", "depGroupVar"))
+added_dep_to_fascicles <- unique(merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI")))
+
+melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
+
+#exclude regions where there was no lesion at all
+lesioned <- subset(melted_df, value > 0)
+
+#only include people in healthy or depressed group
+lesioned_dx <- subset(lesioned, depGroupVar != 0)
+
+#separate out healthy and depressed
+healthy <- subset(lesioned, depGroupVar == 1)
+depressed <- subset(lesioned, depGroupVar == 2)
+
+q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
+r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
+r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
+x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
+y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
+z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
+
+#print(q)
+#print(r)
+print(r_color)
+#print(x)
+#print(y)
+#print(z)
+
+###########################################################################################################
+##### Looking at distribution of disease across fascicles across group and within healthy and depressed ###
+###########################################################################################################
+
+#### look at mean fascicle volume across participants and transpose
+# drop exam date and empi, and transpose
+fascicles_alone_df <- t(added_dep_to_fascicles[,!(colnames(added_dep_to_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
+fascicle_names <- rownames(fascicles_alone_df)
+
+#take row means
+fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
+names(fascicles_alone_rowMeans) <- c("fascicle", "value")
+
+#GET RID OF CRANIAL NERVES
+fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
+
+#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
+order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected") + 
+  xlab("Fascicle") +
+  ylab("Mean Proportion") + 
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size =30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
+print(sorted_ggplot)
+
+######### 
+#############
+#### look at mean fascicle volume across healthy participants and transpose
+#############
+
+# drop exam date and empi, and transpose
+healthy_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 1),]
+fascicles_alone_df <- t(healthy_group_fascicles[,!(colnames(healthy_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
+fascicle_names <- rownames(fascicles_alone_df)
+
+#take row means
+fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
+names(fascicles_alone_rowMeans) <- c("fascicle", "value")
+
+#GET RID OF CRANIAL NERVES
+fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines in Non Depressed", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
+
+sorted_ggplot_healthy<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected in Non Depressed") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(sorted_ggplot_healthy)
+
+# ordered by mean value overall
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
+sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(sorted_ggplot_healthy_ordered)
+
+###########
+#### look at mean fascicle volume across depressed participants and transpose
+###########
+# drop exam date and empi, and transpose
+depressed_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 2),]
+fascicles_alone_df <- t(depressed_group_fascicles[,!(colnames(depressed_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
+fascicle_names <- rownames(fascicles_alone_df)
+
+#take row means
+fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
+names(fascicles_alone_rowMeans) <- c("fascicle", "value")
+
+#GET RID OF CRANIAL NERVES
+fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+#sort based on overlap
+fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
+
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
+
+######## Histograms/bargraphs
+#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
+hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected Depressed", breaks = 20)
+
+sorted_ggplot_depressed<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Depressed") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion Affected") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(sorted_ggplot_depressed)
+
+# ordered by mean value overall
+fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
+sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
+  ylab("Proportion Affected") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(sorted_ggplot_depressed_ordered)
+
+#no cranial nerves
+fascicles_for_plot_no_cranial_nerves <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% fascicle_names_dep_network_all_no_cranial_nerves),]
+
+plot_all_fascicles_r_and_l_no_cranial_nerves <-ggplot(data = fascicles_for_plot_no_cranial_nerves, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Net 3_09") + 
+  ylab("Proportion Affected") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_no_cranial_nerves )
+
+##### look at mean fascicles affected in dep vs non dep network
+######
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_25_percent_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_preproc),]
+
+plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, top quartile") + 
+  geom_bar(stat="identity") +
+  ylab("Proportion") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_top_quartile)
+
+#bottom 3 quartiles
+fascicles_dep_bottom_3_quartiles_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_bottom_3_quartiles),]
+
+plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
+  ylab("Proportion") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)
+
+#plot values within the depression network, and values outside of it
+
+fascicles_dep_bottom_quartile_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_bottom_quartile_preproc),]
+
+plot_all_fascicles_r_and_l_sep_bottom_quartile <-ggplot(data = fascicles_dep_bottom_quartile_for_plot, aes(x = fascicle, y=value)) +
+  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom quartile") + 
+  ylab("Proportion") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_sep_bottom_quartile)
+
+#non dep network
+fascicles_nondep_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% fascicle_names_nondep_network),]
+
+plot_all_fascicles_r_and_l_nondep <-ggplot(data = fascicles_nondep_for_plot, aes(x = fascicle, y=value))+
+  ggtitle("Proportion of Volume per Fascicle Affected nonDep net 3_09") + 
+  ylab("Proportion") +
+  geom_bar(stat="identity") +
+  ylim(0,0.8) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5))
+print(plot_all_fascicles_r_and_l_nondep)
+
 ```
@@
   filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
   group_by(EMPI) %>% 
@@
 #lm
-fascicle_age_lm <- lapply(fascicle_names, function(x) 
+fascicle_age_lm <- lapply(fascicle_names_no_cranial_nerves, function(x) 
 {
@@
 })
-names(fascicle_age_lm) <- fascicle_names
+names(fascicle_age_lm) <- fascicle_names_no_cranial_nerves
 
@@
 
+### Mean proportion volume loss x age
+age_x_prop_vol_affected <- lm(total_fascicle_prop_lost~ PAT_AGE_AT_EXAM, data=dep_and_healthy_groups_for_ICD_analysis )
+print(summary(age_x_prop_vol_affected))
+visreg(age_x_prop_vol_affected, gg=TRUE) + 
+  xlab("Age") + 
+  ylab("Mean Proportion of Fascicle Affected") + 
+  ggtitle("Mean Proportion of Fascicles Affected by Age, p = 2.2*10^-5") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+scatter_age_x_prop_affected <- ggplot(data=age_x_prop_vol_affected, aes(x=PAT_AGE_AT_EXAM, y=total_fascicle_prop_lost)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Age") + 
+  ylab("Mean Proportion of Fascicles Affected") + 
+  ggtitle("Mean Proportion of Fascicles Affected by Age, p = 2.2*10^-5") + 
+  theme(plot.title = element_text(hjust = 0.5, size=20),
+        axis.title = element_text(size=20),
+        axis.text = element_text(size=14),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(scatter_age_x_prop_affected)
+
+##### Ascending graph of all age regressions
+#get lm results, including uncorrected
+df_t_lm_age <- sapply(fascicle_age_lm, function(v) summary(v)$coefficients[2,3])
+
+#Convert to data frame
+df_t_lm_age <- data_frame(fascicle_names_no_cranial_nerves, df_t_lm_age)
+names(df_t_lm_age) = c("Fascicle", "tvalue")
+
+#remove items with cranial nerves
+df_t_lm_age_no_cranial_nerves <- df_t_lm_age  %>% filter(Fascicle %in% fascicle_names_no_cranial_nerves)
+
+df_t_lm_age_no_cranial_nerves$Fascicle <- factor(df_t_lm_age_no_cranial_nerves$Fascicle, levels = df_t_lm_age_no_cranial_nerves$Fascicle[order(df_t_lm_age_no_cranial_nerves$tvalue)])
+
+# bar graph of T-values of fascicles
+plot_fascicle_by_t_sorted <-ggplot(data = df_t_lm_age_no_cranial_nerves, aes(x = Fascicle, y=tvalue))+ 
+  geom_bar(stat="identity") +
+  xlab("Fascicle") + 
+  ylab("T-value of AgexFascile Affected Regression") + 
+  ylim(-2,8) +
+  ggtitle("Effects of age within each fascicle ") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
+print(plot_fascicle_by_t_sorted)
 ```
@@
 dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
   group_by(EMPI) %>% 
@@
 
+######## Mean male vs female, male = 1, female = 2, m = 30%, F = 29%
+males_vs_females_across_all_fascicles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[which(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[which(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)], var.equal = T)
+
+sex <- c("Male", "Female")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized ==1))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized ==2))))
+
+data <- data.frame(sex,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(sex, -value),fill=sex, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ggtitle("Mean Proportion of Fascicles Affected Males vs Females, p = 0.48, NS") +
+  ylab("Mean Proportion of Fasccicles Affected") +
+  theme(plot.title = element_text(size=14, hjust=0.5),
+        legend.position="none", 
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
+
+### Linear models
 #lm
@@
 print(fascicle_sex_anova_p05_unc)
+
+######
+# lapply to get results of t-test in each fascicle
+#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_sex_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
+{
+  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2], var.equal=F)")
+  eval(parse(text=t_test_command))
+})
+names(fascicle_sex_t) <- fascicle_names_no_cranial_nerves
+
+#pull t-statistic
+sex_t_values <- sapply(fascicle_sex_t, function(x) x$statistic)
+
+#convert to data frame
+df_t_values <- as.data.frame(sex_t_values) 
+row.names(df_t_values) <- fascicle_names_no_cranial_nerves
+
+#pull p-values
+sex_p_values <- sapply(fascicle_sex_t, function(x) x$p.value)
+
+#convert to data frame
+df_sex_p_values <- as.data.frame(sex_p_values)
+row.names(df_sex_p_values) <- fascicle_names_no_cranial_nerves
+
+#FDR correct p-values
+sex_pfdr <- p.adjust(sex_p_values,method="fdr")
+
+#Convert to data frame
+df_sex_pfdr <- as.data.frame(sex_pfdr)
+row.names(df_sex_pfdr) <- fascicle_names_no_cranial_nerves
+
+# Sort by t
+#Convert to data frame
+df_t_sex <- data_frame(fascicle_names_no_cranial_nerves, df_t_values)
+names(df_t_sex) = c("Fascicle", "tvalue")
+
+df_t_sex$Fascicle <- factor(df_t_sex$Fascicle, levels = df_t_sex$Fascicle[order(df_t_sex$tvalue)])
+
+# bar graph of T-values of fascicles
+plot_fascicle_by_t_sorted <-ggplot(data = df_t_sex, aes(x = Fascicle, y=tvalue))+ 
+  geom_bar(stat="identity") +
+  xlab("Fascicle") + 
+  ylab("T-value of Sex difference") + 
+  ylim(-2,8) + 
+  ggtitle("Effects of sex within each fascicle; all P=NS") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30),
+        plot.title = element_text(hjust = 0.5, size = 30))
+print(plot_fascicle_by_t_sorted)
+
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_overlap_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_t_values$sex_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
+names(df_fascicle_dep_mask_overlap_t_value) <- c("fascicle", "depression_mask_overlap", "sex_t", "volume_in_dep_mask")
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_t_value
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_t_value
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])
+
+# bar graph of T-values of fascicles, sorted by depression overlap
+plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=sex_t), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") +
+  geom_bar(stat="identity") +
+  ylim(0,5) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(plot_fascicle_by_t_sorted_by_dep_overlap)
+
+# plot fascicle overlap x t value
+#statistic, p = 0.010
+dep_overlap_x_sex_t <- lm(sex_t ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_t_value)
+print(summary(dep_overlap_x_sex_t))
+
+scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=depression_mask_overlap, y=sex_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Proportion of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask overlap by dx T, p = 0.010")
+print(scatter_t_val_by_dep_net_overlap)
+
+#statistic volume, p = 0.031
+dep_overlap_x_sex_t <- lm(sex_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_overlap_t_value)
+print(summary(dep_overlap_x_sex_t))
+
+scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=volume_in_dep_mask, y=sex_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask volume by dx T, p = 0.031")
+print(scatter_t_val_by_dep_net_overlap)
+
+#drop areas where there is no overlap with depression network, n = 62
+df_removed_areas_with_no_depression_overlap <- df_fascicle_dep_mask_overlap_t_value %>% filter(depression_mask_overlap!=0)
+
+#statistic, p = 0.040
+df_removed_areas_with_no_depression_overlap_t <- lm(sex_t ~ depression_mask_overlap, data=df_removed_areas_with_no_depression_overlap)
+print(summary(df_removed_areas_with_no_depression_overlap_t))
+
+scatter_t_val_by_dep_net_overlap_removed_areas_with_no_dep_overlap <- ggplot(data=df_removed_areas_with_no_depression_overlap, aes(x=depression_mask_overlap, y=sex_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Proportion of Fascicle in Depression Network (only areas that overlapped to some degree") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask overlap by dx T, removed fascicles with no dep mask overlap, p = 0.040")
+print(scatter_t_val_by_dep_net_overlap_removed_areas_with_no_dep_overlap)
+
+### look at number of voxels, not just proportion
 
@@
            rowMeans(across(fascicle_names_nondep_network))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_15_percent =
-           rowMeans(across(fascicle_names_dep_network_15_percent))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_bottom_quartile = 
+           rowMeans(across(dep_network_names_bottom_quartile))) %>%
   group_by(EMPI) %>% 
@@
 summary(dep_vs_healthy) #p=0.0494
+
+dep_vs_healthy_top_quartile <- lm(total_fascicle_prop_lost_indepnet_top_quartile ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_top_quartile) #p=0.054
 
@@
 
-#look at within depression network in fascicles where at least 15% of the volume of the fascicle was in dep network
-#yes, p=0.054, mean dep = 30%, mean nondep = 25%
-total_volume_lost_dep_vs_healthy_unique_in_dep_network_15percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_15_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_15_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
+#within 25% depression network; n = 0.053, depressed 33%, healthy 28%
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
+
+#within bottom quartile depression network; n = 0.053, depressed 8%, healthy 7%, p=0.12
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_bottom_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
 
@@
 #df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
+
+```
+
+```{r depression_unique_all_fascicles_together_in_dep_networks_no_cranial_nerves}
+
+####
+
+# so this is interesting.
+When you look at the top quartile, but exclude the cranial nerves, you drop a few more fascicles from the depression network (i.e.
+now it has 19 items, instead of 22).
+This actually pushes the depressed vs healthy within the depression network over the hump, becoming significant.
+Probably an even cooler finding.
+So if you look over all the fascicles, not significantly different.
+But if you look within the depression network (25%), depressed patients have more disease than healthy patients
+
+#### 
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_dep_network_all_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all_no_cranial_nerves))) %>% 
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowMeans(across(nondep_network_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_bottom_3_quartiles_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_bottom_quartile = 
+           rowMeans(across(dep_network_names_bottom_quartile_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy) #p=0.055
+
+dep_vs_healthy_top_quartile <- lm(total_fascicle_prop_lost_indepnet_top_quartile ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_top_quartile) #p=0.048
+
+dep_vs_healthy_nondep_bottom_3_quartiles <- lm(total_fascicle_prop_lost_nondep_bottom_3_quartiles ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_nondep_bottom_3_quartiles) #p=0.057
+
+#test whether proportion lost overall different between depressed and healthy- no, p = 0.054
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
+
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
+df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
+
+plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
+  geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))
+
+print(plot)
+
+#bar plot
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of Fascicles Affected") +
+  ylim(0,0.5) +
+  theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Proportion of Fascicles Affected") +
+  ggtitle("Total Prop Affected by dx") + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 14), 
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
+
+#look at within depression network using top quartiles, yes p = 0.047
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 26%, p = 0.047
+
+#look at outside depression network (i.e. areas that never overalpped dep network), yes p = 0.049
+total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 23%, mean non_dep = 20%, p = 0.049
+
+#look at 75%, no p = 0.057
+total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
+
+#look at bottom quartile
+
+total_volume_lost_dep_vs_healthy_unique_in_bottom_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.13
+
+############
+#bar plot
+############
+
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ggtitle("Mean Prop Fasc Affected w/in 25% dep mask by dx; p=0.047") +
+  ylab("Proportion of Fascicles Affected") +
+  ylim(0,0.5) +
+  theme(title = element_text(size=14),
+    axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ggtitle("Mean Prop Fasc Affected w/in 25% dep mask by dx; p=0.047") +
+  ylab("Proportion of Fascicles Affected") +
+  theme(title = element_text(size=14),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
 
@@
 
-#fdr corrected only dep 15% mask
-fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_volumes_dep_network$prop_in_mask >= 0.15)]
-fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
-print(fascicle_anova_fdr)
-
 #uncorrected
@@
 
-nondep15_net <- fascicle_names[!(fascicle_names %in% fascicle_names_dep_network_15_percent)]
-
+#some new subnetworks
+
+#filter
 dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
@@
            rowMeans(across(fascicle_names_nondep_network))) %>%
-  mutate(total_fascicle_prop_lost_dep_network_15_percent = 
-           rowMeans(across(fascicle_names_dep_network_15_percent))) %>%
-  mutate(total_fascicle_prop_lost_nondep_network_15_percent = 
-           rowMeans(across(nondep15_net))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_bottom_three_quartiles = 
+           rowMeans(across(dep_network_bottom_3_quartiles))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_bottom_quartile = 
+           rowMeans(across(dep_network_names_bottom_quartile))) %>%
   group_by(EMPI) %>% 
@@
 
-volume_lost_dep15_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_dep_network_15_percent,
-dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 28%, mean non_dep_net = 14%, p 2.2 * 10^-16
-
-volume_lost_dep15_vs_nondep_network_15_percent_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_dep_network_15_percent,
-dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network_15_percent, var.equal = F) #mean dep 15 net = 28%, mean all fascicles not in 15% net = 26%, p=0.13
+#compares top quartile of depression network with fascicles that NEVER overlapped w/depression network
+volume_lost_dep_vs_nondep_network_top_quartile_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 31%, mean non_dep_net = 20%, p 2.2 * 10^-16
+
+#compares top quartile of depression network with all fascicles that are NOT in the top quartile of depression network; i.e. everything not in top quartile is the "non"-depressed network
+volume_lost_dep_top_versus_bottom_3_quartiles_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_bottom_three_quartiles, var.equal = F) #mean dep net = 31%, mean bottom 3 quartiles = 24, p 1.6 * 10^-5
+
+#compares top quartile of depression network with bottom quartile of depression network
+volume_lost_dep_top_versus_bottom_quartile_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_bottom_quartile, var.equal = F) #mean dep net = 31%, mean bottom quartile = 7%, p 2.2 * 10^-16
 
@@
 loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.042
+
+```
+
+```{r testing_whether_ms_disease_more_likely_in_depression_mask_unique_no_cranial_nerves}
+### tests whether there is higher proportion of vertices lost in depression network versus not
+
+### Take home here - 
+## 1) in the absence of the cranial nerves, more disease in areas than interact w/depression network vs places that never had any overlap
+## 2) in the absence of the cranial nerves, more disease in the top 25% than in areas that never intersected with depression map 
+## 3) in the absense of cranial nerves, the difference between the top 25% and the bottom 75% is no longer significant
+
+#some new subnetworks
+
+#filter
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_nondep_network =
+           rowMeans(across(nondep_network_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_bottom_three_quartiles = 
+           rowMeans(across(dep_network_bottom_3_quartiles_no_cranial_nerves))) %>%
+  mutate(total_fascicle_prop_lost_indepnet_bottom_quartile = 
+           rowMeans(across(dep_network_names_bottom_quartile_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ### when I filter and remove fascicles where there was no loss at all, the depression vs healthy findings go away
+#  filter(total_fascicle_prop_lost > 0) %>%
+  ungroup()
+
+#test whether there is differentially more volume loss in depression versus nondepression network
+volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 31%, mean non_dep_net =  22%, p 8.58 x 10^-12
+
+#compares top quartile of depression network with fascicles that NEVER overlapped w/depression network
+volume_lost_dep_vs_nondep_network_top_quartile_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 22%, p 7.3* 10^-8
+
+#compares top quartile of depression network with all fascicles that are NOT in the top quartile of depression network; i.e. everything not in top quartile is the "non"-depressed network
+volume_lost_dep_top_versus_bottom_3_quartiles_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_bottom_three_quartiles, var.equal = F) #mean dep net = 30%, mean bottom 3 quartiles = 29, p = 0.52
+
+#compares top quartile of depression network with bottom quartile of depression network; 
+volume_lost_dep_top_versus_bottom_quartile_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile,
+dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_bottom_quartile, var.equal = F) #mean dep net = 30%, mean bottom quartile= 28, p = 0.4
+
+#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "prop_lost", "network")
+df_for_visreg$network <- as.factor(df_for_visreg$network)
+
+plot <- ggplot(df_for_visreg, aes(x=network, y = prop_lost, col=network)) +
+  geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="red") + 
+  scale_x_discrete(labels=c("1" = "Depression Network", "2" = "Non-Depression Network")) + 
+  ylab("Proportion of Fascicles Affected") + 
+  theme(axis.text =element_text(size=14), axis.title.x = element_blank())
+
+print(plot)
+
+########## Bar plot
+Network <- c("Depression Network","Non-Depression Network")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])))
+
+data <- data.frame(Network,value,sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Network,fill=Network, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("Proportion of Fascicles Affected") +
+  ylim(0,0.5) +
+  theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+# Black and white
+barplot <- ggplot(data, aes(x=Network,fill=Network, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("Proportion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
 
@@
 ```
+
+```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves}
+# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups
+
+##############################
+## FASCICLES TO BE INCLUDED ##
+##############################
+
+#define the depression network (all areas that intersect with depression network, no cranial nerves)
+fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
+
+#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
+fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
+fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
+
+#######################
+### depression map  ###
+#######################
+
+### define depression map
+fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
+
+#filter to exclude fascicles w/cranial nerves
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_dep %>%
+  replace(is.na(.), 0) %>%
+  filter(fascicle%in% fascicle_names_no_cranial_nerves)
+
+#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column
+
+df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))
+df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
+
+#######################
+### real data x dx  ###
+#######################
+
+#then, make the data frame
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+# lapply to get results of t-test in each fascicle
+#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
+{
+  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
+  eval(parse(text=t_test_command))
+})
+names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves
+
+#pull t-statistic
+dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)
+
+#convert to data frame
+df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
+row.names(df_t_values) <- fascicle_names_no_cranial_nerves
+
+#pull p-values
+dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)
+
+#convert to data frame
+df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
+row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves
+
+#FDR correct p-values
+dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
+
+#Convert to data frame
+df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
+row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves
+
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_overlap_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
+names(df_fascicle_dep_mask_overlap_t_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_t", "volume_in_dep_mask")
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_t_value
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])
+
+# bar graph of T-values of fascicles, sorted by depression overlap
+plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") +
+  geom_bar(stat="identity") +
+  ylim(0,5) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(plot_fascicle_by_t_sorted_by_dep_overlap)
+
+# plot fascicle overlap x t value
+#statistic, p = 0.010
+dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_t_value)
+print(summary(dep_overlap_x_dep_vs_healthy_t))
+
+scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Proportion of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask overlap by dx T, p = 0.010") + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_t_val_by_dep_net_overlap)
+
+#statistic volume, p = 0.031
+dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_overlap_t_value)
+print(summary(dep_overlap_x_dep_vs_healthy_t))
+
+scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask volume by dx T, p = 0.031") + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_t_val_by_dep_net_overlap)
+
+#drop areas where there is no overlap with depression network, n = 62
+df_removed_areas_with_no_depression_overlap <- df_fascicle_dep_mask_overlap_t_value %>% filter(depression_mask_overlap!=0)
+
+#statistic, p = 0.040
+df_removed_areas_with_no_depression_overlap_t <- lm(dep_vs_healthy_t ~ depression_mask_overlap, data=df_removed_areas_with_no_depression_overlap)
+print(summary(df_removed_areas_with_no_depression_overlap_t))
+
+scatter_t_val_by_dep_net_overlap_removed_areas_with_no_dep_overlap <- ggplot(data=df_removed_areas_with_no_depression_overlap, aes(x=depression_mask_overlap, y=dep_vs_healthy_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Proportion of Fascicle in Depression Network (only areas that overlapped to some degree") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask overlap by dx T, removed fascicles with no dep mask overlap, p = 0.040") + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_t_val_by_dep_net_overlap_removed_areas_with_no_dep_overlap)
+
+### look at number of voxels, not just proportion
+
+```
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected}
+
+#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#read in fiber volume values
+fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+
+#remove rows with cranial nerves
+#fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fiber_volume_values$fascicle %in% fascicle_names_no_cranial_nerves)
+
+#make dataframe and multiply values
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_dep_network_all_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all_no_cranial_nerves))) %>% 
+  mutate(total_fascicle_vol_lost_nondep_network =
+           rowMeans(across(nondep_network_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowMeans(across(dep_network_bottom_3_quartiles_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_nondep_bottom_quartile = 
+           rowMeans(across(dep_network_names_bottom_quartile_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+dep_vs_healthy <- lm(total_fascicle_vol_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy) #p=0.0456
+
+dep_vs_healthy_top_quartile <- lm(total_fascicle_vol_lost_indepnet_top_quartile ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_top_quartile) #p=0.0351
+
+dep_vs_healthy_nondep_bottom_3_quartiles <- lm(total_fascicle_vol_lost_nondep_bottom_3_quartiles ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_nondep_bottom_3_quartiles) #p=0.0535
+
+#test whether volortion lost overall different between depressed and healthy- no, p = 0.054
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
+
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "vol_lost", "Diagnosis")
+df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
+
+plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = vol_lost, col=Diagnosis)) +
+  geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))
+
+print(plot)
+
+#bar plot
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("volortion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("volortion of Fascicles Affected") +
+  ggtitle("Total vol Affected by dx") + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 14), 
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
+
+#look at within depression network using top quartiles, yes p = 0.047
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 26%, p = 0.047
+
+#look at outside depression network (i.e. areas that never overalpped dep network), yes p = 0.049
+total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 23%, mean non_dep = 20%, p = 0.049
+
+#look at 75%, no p = 0.057
+total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
+
+#look at bottom quartile
+
+total_volume_lost_dep_vs_healthy_unique_in_bottom_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.13
+
+############
+#bar plot
+############
+
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ggtitle("Mean vol Fasc Affected w/in 25% dep mask by dx") +
+  ylab("volortion of Fascicles Affected") +
+  theme(title = element_text(size=14),
+    axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ggtitle("Mean vol Fasc Affected w/in 25% dep mask by dx") +
+  ylab("Volume of Fascicles Affected") +
+  theme(title = element_text(size=14),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
+#test dep vs healthy across all volume
+
+```
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_sums}
+
+#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
+#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
+
+#read in fiber volume values
+fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
+df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
+
+#multiply each row by the vector of total volumes
+df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values$volume_full, FUN = "*")
+
+#combine back with data_fram
+df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+
+#remove rows with cranial nerves
+#fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fiber_volume_values$fascicle %in% fascicle_names_no_cranial_nerves)
+
+#make dataframe and multiply values
+
+#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_vol_lost = rowSums(across(fascicle_names_dep_network_all_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_indepnet = rowSums(across(fascicle_names_dep_network_all_no_cranial_nerves))) %>% 
+  mutate(total_fascicle_vol_lost_nondep_network =
+           rowSums(across(nondep_network_names_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
+           rowSums(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowSums(across(dep_network_bottom_3_quartiles_no_cranial_nerves))) %>%
+  mutate(total_fascicle_vol_lost_nondep_bottom_quartile = 
+           rowSums(across(dep_network_names_bottom_quartile_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+dep_vs_healthy <- lm(total_fascicle_vol_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy) #p=0.0456
+
+dep_vs_healthy_top_quartile <- lm(total_fascicle_vol_lost_indepnet_top_quartile ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_top_quartile) #p=0.0351
+
+dep_vs_healthy_nondep_bottom_3_quartiles <- lm(total_fascicle_vol_lost_nondep_bottom_3_quartiles ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy_nondep_bottom_3_quartiles) #p=0.0535
+
+#test whether vol lost overall different between depressed and healthy- no, p = 0.045
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #Sum dep = 33%, Sum non_dep = 28%, p = 0.045
+
+#look at within depression network using top quartiles, yes p = 0.034
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #sum dep = 32%, sum non_dep = 26%, p = 0.047
+
+#look at outside depression network (i.e. areas that never overalpped dep network), yes p = 0.048
+total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #sum dep = 23%, sum non_dep = 20%, p = 0.049
+
+#look at 75%, no p = 0.054
+total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #sum dep = 30%, sum non_dep = 27%, p = 0.057
+
+#look at bottom quartile, p = 0.14
+
+total_volume_lost_dep_vs_healthy_unique_in_bottom_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #sum dep = 18%, sum non_dep = 16%, p = 0.13
+
+df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
+names(df_for_visreg) <- c("EMPI", "vol_lost", "Diagnosis")
+df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)
+
+plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = vol_lost, col=Diagnosis)) +
+  geom_violin() + 
+  geom_jitter(shape=16, position=position_jitter(0.05)) + 
+  geom_boxplot(width=0.1) + theme_minimal() + 
+  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))
+
+print(plot)
+
+#bar plot
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ylab("volortion of Fascicles Affected") +
+  theme(axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ylab("volortion of Fascicles Affected") +
+  ggtitle("Total vol Affected by dx") + 
+  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        plot.title = element_text(hjust=0.5, size = 14), 
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+  #scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
+
+############
+#bar plot
+############
+
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ggtitle("sum vol Fasc Affected w/in 25% dep mask by dx") +
+  ylab("volortion of Fascicles Affected") +
+  theme(title = element_text(size=14),
+    axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ggtitle("sum vol Fasc Affected w/in 25% dep mask by dx") +
+  ylab("Volume of Fascicles Affected") +
+  theme(title = element_text(size=14),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+
+print(barplot)
+#test dep vs healthy across all volume
+
+################
+
+################
+#######################
+### depression map  ###
+#######################
+
+### define depression map
+fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
+
+#filter to exclude fascicles w/cranial nerves
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_dep %>%
+  replace(is.na(.), 0) %>%
+  filter(fascicle%in% fascicle_names_no_cranial_nerves)
+
+#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column
+
+df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))
+df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
+
+#######################
+### real data x dx  ###
+#######################
+
+# lapply to get results of t-test in each fascicle
+#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
+{
+  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
+  eval(parse(text=t_test_command))
+})
+names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves
+
+#pull t-statistic
+dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)
+
+#convert to data frame
+df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
+row.names(df_t_values) <- fascicle_names_no_cranial_nerves
+
+#pull p-values
+dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)
+
+#convert to data frame
+df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
+row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves
+
+#FDR correct p-values
+dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
+
+#Convert to data frame
+df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
+row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves
+
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_overlap_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
+names(df_fascicle_dep_mask_overlap_t_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_t", "volume_in_dep_mask")
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_t_value
+
+df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])
+
+# bar graph of T-values of fascicles, sorted by depression overlap
+plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") +
+  geom_bar(stat="identity") +
+  ylim(0,5) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+print(plot_fascicle_by_t_sorted_by_dep_overlap)
+
+# plot fascicle overlap x t value
+#statistic, p = 0.01
+dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_t_value)
+print(summary(dep_overlap_x_dep_vs_healthy_t))
+
+scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Proportion of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask overlap by dx T, p = 0.01") + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_t_val_by_dep_net_overlap)
+
+#statistic volume, p = 0.031
+dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_overlap_t_value)
+print(summary(dep_overlap_x_dep_vs_healthy_t))
+
+scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask volume by dx T, p = 0.031") + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_t_val_by_dep_net_overlap)
+
+#drop areas where there is no overlap with depression network, n = 62
+df_removed_areas_with_no_depression_overlap <- df_fascicle_dep_mask_overlap_t_value %>% filter(depression_mask_overlap!=0)
+
+#statistic, p = 0.040
+df_removed_areas_with_no_depression_overlap_t <- lm(dep_vs_healthy_t ~ depression_mask_overlap, data=df_removed_areas_with_no_depression_overlap)
+print(summary(df_removed_areas_with_no_depression_overlap_t))
+
+scatter_t_val_by_dep_net_overlap_removed_areas_with_no_dep_overlap <- ggplot(data=df_removed_areas_with_no_depression_overlap, aes(x=depression_mask_overlap, y=dep_vs_healthy_t)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Proportion of Fascicle in Depression Network (only areas that overlapped to some degree") + 
+  ylab("Effect of diagnosis (T)") + 
+  ggtitle("Fascicle/depression mask overlap by dx T, removed fascicles with no dep mask overlap, p = 0.040") + 
+  theme(plot.title=element_text(hjust=0.5))
+print(scatter_t_val_by_dep_net_overlap_removed_areas_with_no_dep_overlap)
+
+### look at number of voxels, not just proportion
+
+```
+
+```{r compare_overall_volume_of_lesions_dep_vs_non_dep}
+
+lesion_volumes <- read.csv(paste0(homedir, "results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE)
+
+#merge with full dataset
+df_full_dataset_plus_lesion_volumes = merge(df_demo_and_fascicles, lesion_volumes, by = c("EMPI", "EXAM_DATE"))
+
+#separate out only people in depression group, and take first instance
+dep_and_healthy_groups_for_ICD_analysis <- df_full_dataset_plus_lesion_volumes %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+dep_vs_healthy_volume <- lm(volume_of_mimosa_lesions ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
+summary(dep_vs_healthy) #p=0.08
+
+dep_vs_healthy_t_test_lesion_volume <-  t.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #p=0.07
+
+Diagnosis <- c("MS w/Depression", "MS No Depression")
+value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+           mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+
+sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
+           (sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
+
+data <- data.frame(Diagnosis,value, sem)
+#data$Fascicle <- pretty_names
+
+# Grouped
+barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", position=position_dodge(0.9)) + 
+  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
+  ggtitle("Volume of Mimosa Lesions by Diagnosis") +
+  ylab("Volume of Lesions") +
+  theme(title = element_text(size=14),
+    axis.text = element_text(size = 12, colour = "black"))
+
+print(barplot)
+
+# Grouped
+barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
+  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
+    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
+  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
+  ggtitle("Volume of Mimosa Lesions by Diagnosis") +
+  ylab("Volume of Lesion") +
+  theme(title = element_text(size=14),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=20),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
+#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
+print(barplot)
+
+```

### 2022-08-13 make_afni_map_w_fascicles_colored_by_prop_or_volume.sh

@@ -1,0 +1,66 @@
+#!/bin/bash
+
+##### Welcome to make_afni_map_w_fascicles_colored_by_prop_overlap ###
+### Pre: A file that has two columns, no headers. First column is the abbreviation of a fascicle, second is proportion overlap (or volume overlap)
+### Post: A nifti where each fascicle is colored by the proportion of overlap, or any value you give it
+### Uses: I wanted to find a way to show, in a brain, which fascicles have the most overlap with lesions (though this can be done w/depression, non-dep, volume). This will go through a table, and for each fascicle, multiply it times the value in the second column, add a suffix, and then at the end, sum them all to make a final image
+### Dependencies: Afni
+
+#change this if doing on pmacs
+homedir="/Users/eballer/BBL/msdepression/"
+
+default=${homedir}/results/fascicle_mean_proportion_overlap_nfasc_77.csv
+default_suffix="prop"
+
+if [ $# == 0 ]
+then
+	echo "We will use the default path file" $default
+    	echo "We will use the default suffix " $default_suffix
+    	fascicle_file=${default}
+	suffix=${default_suffix}
+
+else  
+	fascicle_file=$1
+	suffix=$2
+fi
+
+
+echo "File being read is "$fascicle_file
+echo "Suffix being used is " $suffix
+
+fascicles=$(cat $fascicle_file)
+
+cd ${homedir}/templates/dti/HCP_YA1065_tractography/niftis_excluding_cranial_nerves
+
+(pwd)
+for fascicle in ${fascicles}; do
+	# extract fascicle name and value
+	fascicle_name=$(echo ${fascicle} | perl -pe 's/(.*),.*/$1/')
+	fascicle_value=$(echo ${fascicle} | perl -pe 's/.*,(.*)/$1/')
+	
+	# remove nifti w/values if already there
+	rm -rf ${fascicle_name}_${suffix}.nii.gz
+
+	#take the fascicle .nii, multiply it with the value
+	command=$(echo 3dcalc -a ${fascicle_name}.nii.gz -expr \'a*${fascicle_value}\' -prefix ${fascicle_name}_${suffix}.nii.gz)
+	echo "Here is the command: "${command}
+	
+	#run command, have to do separately to get variable into afni command
+	eval "${command}";
+done
+
+#remove old file
+rm -rf fascicle_${suffix}_sum.nii.gz
[…]

### 2022-08-13 make_red_to_yellow_RGB_color_scheme.R

@@ -1,0 +1,48 @@
+#### Welcome to make_red_to_yellow_RGB_color_scheme! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle to be its proportion of loss, scaled on the RGB scale
+### In order to do this, I needed to figure out how to map my proportions to RGB
+
+### Pre: File of proportions
+### Post: Output of RGB values in Red to Yellow Scale for the Proportions
+### Uses: Step 0: Figure out how to scale from red to yellow 
+###           - Red = 255 0 0, yellow 255 255 0
+###           - Means that we need to vary the middle value (255) to get where we need to be
+###       Step 1) For red to yellow, calculate the step (i.e. how much each value will be)
+###             - because my proportions are out of 1, I wanted each step to be 1/100
+###             - I am setting this value as "max", so it could vary if you have another max #
+###       Step 2) Multiply the proportion by 100, so now each count is 1-100, in percents
+###       Step 3) Multiply the percent by the step
+###       Step 4) Save this value in that fascicle, 255 [new_value] 0
+###
+### Dependencies: Any R will do
+
+###
+homedir="/Users/eballer/BBL/msdepression/"
+
+#highest possible number in scale
+max<-100 
+
+#step that each value in our scale will correspond to
+step<-255/max 
+
+#read in file with fascicle name and proportion
+fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77.csv"), sep = ",", header = F)
+names(fascicle_name_and_prop) <- c("fascicle", "proportion")
+
+# make a column for red
+red <- rep(x=255, times = 77)
+
+fascicle_name_and_prop$red <- red
+#multiply proportion column by 100, then by step, and round
+fascicle_name_and_prop$green<- round((fascicle_name_and_prop$proportion *100*step),0)
+
+#make column for blue
+blue <- rep(x=0, times = 77)
+fascicle_name_and_prop$blue=blue
+
+#sort based on proportion to check, it works! Lower values are associated with lower RGB
+sorted <- fascicle_name_and_prop[order(fascicle_name_and_prop$proportion),]
+
+

### 2022-08-23 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
added t tests
including wilcoxon tests

@@
 
+dep_network_total_volume_top_quartile_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network$num_voxels_total[which(fascicle_volumes_dep_network$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)])
+
 #get a network that includes all of the fascicles NOT in dep 25% network
 dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)] #n=68
+
+bottom_75_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network$num_voxels_total[which(fascicle_volumes_dep_network$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol)])
 
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        axis.text.y = element_text(size=12),
         plot.title = element_text(hjust = 0.5, size=14),
@@
 #isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
+
+### of note, order of operations is that you take the mean first, and then you divide by the total volume of the network.
+One could imagine a situation where you divide by network first, and then take mean, but it seems reasonable to include network measure after calculating the network mean.
+
 dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
@@
   mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves_by_vol))) %>%
+  mutate(prop_fascicle_vol_lost_indepnet_top_quartile = (total_fascicle_vol_lost_indepnet_top_quartile / dep_network_total_volume_top_quartile_no_cranial_nerves)) %>%
   mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol))) %>%
+  mutate(total_fascicle_vol_lost_no_cc = rowMeans(across(fascicle_names_no_cranial_nerves[fascicle_names_no_cranial_nerves!="CC"]))) %>%
+  mutate(prop_fascicle_vol_lost_bottom_3_quartile = (total_fascicle_vol_lost_nondep_bottom_3_quartiles / bottom_75_total_volume_no_cranial_nerves)) %>%
+  mutate(total_fascicle_vol_lost_indepnet_top_quartile_no_cc = 
+           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves_by_vol[dep_network_names_top_quartile_no_cranial_nerves_by_vol != "CC"]))) %>%
+  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc = 
+           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol[dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol != "CC"]))) %>%
   group_by(EMPI) %>% 
@@
 
+#test whether ms preferentially hits within or outside depression network, p = 2.2 * 10^-16, mean lost in dep net = 18027, mean lost outside dep net = 4020.
+More disease outside versus inside, when you look at means
+
+msdisease_within_versus_outside_depression_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles, var.equal = F, paired = TRUE)
+
+#wilcox test because non normally distributed; p < 2.2*10^-16
+msdisease_within_versus_outside_depression_network_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles,paired = TRUE)
+
+#still 2.2*10^-16; mean of dep net: 6856, mean outside dep net = 4020, excluded corpus callosum b/c it is so very big. 
+msdisease_within_versus_outside_depression_network_no_cc <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile_no_cc, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc, var.equal = F, paired = TRUE)
+
+#2.2*1^-16
+msdisease_within_versus_outside_depression_network_no_cc_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile_no_cc, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc, paired = TRUE)
+
+#testing group differences if you scale by the size of the network, still significant, in top quartile, proportion lost = 0.021
+#bottom 3, proportion lost = 0.006, p=2.2 * 10^-16
+
+msdisease_within_versus_outside_depression_network_scaled_by_network_size <- t.test(dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_indepnet_top_quartile, dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_bottom_3_quartile, var.equal = F, paired = TRUE)
+
+#2.2 *10^-16
+msdisease_within_versus_outside_depression_network_scaled_by_network_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_indepnet_top_quartile, dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_bottom_3_quartile, paired = TRUE)
+
 #test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
-total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
-
-#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
+total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
+
+#Yes, p=0.046, Dep = 4952, nondep = 4284
+total_volume_lost_dep_vs_healthy_unique_no_cc <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
+
+#look at within depression network using top quartiles, yes p = 0.034, Dep = 18957, non dep = 16572
 total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
 
-#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
+#no corpus callosum, p = 0.028, Dep = 7380, nondep = 6036
+total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile_no_cc <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
+
+#look at bottom 75%, p = 0.07, dep = 4198, non dep = 3740
 total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
+
+#gut check, this should be identical, since CC not in bottom 3 quartiles; p = 0.07, dep = 4198, non dep = 3740
+total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles_no_cc <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
 
@@
 dplyr::select(total_fascicle_vol_lost, total_fascicle_vol_lost_indepnet_top_quartile, total_fascicle_vol_lost_nondep_bottom_3_quartiles, depGroupVar) %>%
+  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
+
+df_mean_fascicle_volumes_and_diagnosis_no_cc <- dep_and_healthy_groups_for_ICD_analysis %>% 
+dplyr::select(total_fascicle_vol_lost_no_cc, total_fascicle_vol_lost_indepnet_top_quartile_no_cc, total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc, depGroupVar) %>%
   mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
@@
 
+violin_mean_fascicle_volume_no_cc <- ggplot(df_mean_fascicle_volumes_and_diagnosis_no_cc, aes(x=depGroupVar, y=total_fascicle_vol_lost_no_cc, fill=depGroupVar)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected by Diagnosis, No CC")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,42000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_mean_fascicle_volume_no_cc) #p=0.067
+
 violin_mean_fascicle_volume_dep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_indepnet_top_quartile, fill=depGroupVar)) +
@@
         panel.background = element_rect(colour = "white", fill = "white"))
-print(violin_mean_fascicle_volume_dep) #p=0.067
+print(violin_mean_fascicle_volume_dep)
+
+violin_mean_fascicle_volume_dep_no_cc <- ggplot(df_mean_fascicle_volumes_and_diagnosis_no_cc, aes(x=depGroupVar, y=total_fascicle_vol_lost_indepnet_top_quartile_no_cc, fill=depGroupVar)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected Dep Net by Diagnosis")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,42000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_mean_fascicle_volume_dep_no_cc)
 
@@
 print(violin_mean_fascicle_volume_nondep)
+
+# geom violin within vs outside depression network
+dep_net_group <- c(rep("Dep_net", times = 380), rep("Non_dep_net",times = 380))
+net_values <- c(df_mean_fascicle_volumes_and_diagnosis$total_fascicle_vol_lost_indepnet_top_quartile, df_mean_fascicle_volumes_and_diagnosis$total_fascicle_vol_lost_nondep_bottom_3_quartiles)
+
+newdf <- data.frame(dep_net_group, net_values)
+
+#p = 2.2 x 10^-16.
+But it is kind of irrelevant because corpus callosum, which is periventricular and just so much bigger, is in the depression network so if you look at volume alone, there is really no way to compete.
+For this, you gotta use proportion.
+And when you drop the cranial nerves, doesn't work.
+COuld try dropping CC, but don't know why I coul dmake the case for that.
+Though honestly, will prob have to try
+violin_within_vs_outside_network_fascicle_volume <- ggplot(newdf, aes(x=dep_net_group, y=net_values, fill=dep_net_group)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Dep Net", "2" = "Non-Dep Net")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected by Network")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,42000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_within_vs_outside_network_fascicle_volume) #
+
+# geom violin within vs outside depression network, no cc
+dep_net_group <- c(rep("Dep_net", times = 380), rep("Non_dep_net",times = 380))
+net_values <- c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile_no_cc, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc)
+
+newdf_no_cc <- data.frame(dep_net_group, net_values)
+
+#p = 2.2 x 10^-16.
+But it is kind of irrelevant because corpus callosum, which is periventricular and just so much bigger, is in the depression network so if you look at volume alone, there is really no way to compete.
+For this, you gotta use proportion.
+And when you drop the cranial nerves, doesn't work.
+COuld try dropping CC, but don't know why I coul dmake the case for that.
+Though honestly, will prob have to try
+violin_within_vs_outside_network_fascicle_volume_no_cc <- ggplot(newdf_no_cc, aes(x=dep_net_group, y=net_values, fill=dep_net_group)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Dep Net", "2" = "Non-Dep Net")) + 
+  ggtitle(paste0("Mean Fascicle Volume Affected by Network, no CC")) +
+  ylab("Mean Volume of Fascicles") +
+  ylim(0,42000) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_within_vs_outside_network_fascicle_volume_no_cc) #
+
+### scaling by size of network
+
+dep_net_group <- c(rep("Dep_net", times = 380), rep("Non_dep_net",times = 380))
+net_values <- c(dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_indepnet_top_quartile, dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_bottom_3_quartile)
+
+newdf_no_cc <- data.frame(dep_net_group, net_values)
+
+#p = 2.2 x 10^-16. 
+violin_within_vs_outside_network_scaled_by_network_size <- ggplot(newdf_no_cc, aes(x=dep_net_group, y=net_values, fill=dep_net_group)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Dep Net", "2" = "Non-Dep Net")) + 
+  ggtitle(paste0("Proportion Fascicle Volume Affected by Network")) +
+  ylab("Proportion (tot volume lost/Net Vol) ") +
+  ylim(0,0.1) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_within_vs_outside_network_scaled_by_network_size) #
+
 ```
@@
 ```
+
+```{r age_effects_fascicles_unique_volume}
+########### age effect whole group
+
+#isolate out only the depressed group and healthy group
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+#### age histo ###
+hist(dep_and_healthy_groups_for_ICD_analysis$PAT_AGE_AT_EXAM,
+main = paste0("Age In Good Mimosa Group : n = ", dim(dep_and_healthy_groups_for_ICD_analysis)[1]), xlab = "Age", breaks = 10)
+
+#lm
+fascicle_age_lm <- lapply(fascicle_names_no_cranial_nerves, function(x) 
+{
+  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
+})
+names(fascicle_age_lm) <- fascicle_names_no_cranial_nerves
+
+#anova
+fascicle_age_anova <- lapply(fascicle_age_lm, anova)
+
+#fdr corrected
+fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
+print(fascicle_age_anova_fdr)
+
+#fdr corrected only association cortex
+fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
+fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
+print(fascicle_anova_fdr_association)
+
+#visreg
+#sapply(fascicle_age_lm,visreg)
+visreg(fascicle_age_lm$AF_R,ylab = "Volume of Affected Streamlines", xlab = "Age in Years",
+       line=list(col="red"),
+       points=list(col="black"))
+
+### Mean Volume loss x age
+age_x_vol_affected <- lm(total_fascicle_vol_lost~ PAT_AGE_AT_EXAM, data=dep_and_healthy_groups_for_ICD_analysis )
+print(summary(age_x_vol_affected))
+visreg(age_x_vol_affected, gg=TRUE) + 
+  xlab("Age") + 
+  ylab("Mean Volume of Fascicle Affected") + 
+  ggtitle("Mean Volume of Fascicles Affected by Age, p = 2.1*10^-5") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+scatter_age_x_vol_affected <- ggplot(data=age_x_vol_affected, aes(x=PAT_AGE_AT_EXAM, y=total_fascicle_vol_lost)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Age") + 
+  ylab("Mean Volume of Fascicles Affected") + 
+  ggtitle("Mean Volume of Fascicles Affected by Age, p = 2.1*10^-5") + 
+  theme(plot.title = element_text(hjust = 0.5, size=20),
+        axis.title = element_text(size=20),
+        axis.text = element_text(size=14),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(scatter_age_x_vol_affected)
+
+##### Ascending graph of all age regressions
+#get lm results, including uncorrected
+df_t_lm_age <- sapply(fascicle_age_lm, function(v) summary(v)$coefficients[2,3])
+
+#Convert to data frame
+df_t_lm_age <- data_frame(fascicle_names_no_cranial_nerves, df_t_lm_age)
+names(df_t_lm_age) = c("Fascicle", "tvalue")
+
+#remove items with cranial nerves
+df_t_lm_age_no_cranial_nerves <- df_t_lm_age  %>% filter(Fascicle %in% fascicle_names_no_cranial_nerves)
+
+df_t_lm_age_no_cranial_nerves$Fascicle <- factor(df_t_lm_age_no_cranial_nerves$Fascicle, levels = df_t_lm_age_no_cranial_nerves$Fascicle[order(df_t_lm_age_no_cranial_nerves$tvalue)])
+
+# bar graph of T-values of fascicles
+plot_fascicle_by_t_sorted <-ggplot(data = df_t_lm_age_no_cranial_nerves, aes(x = Fascicle, y=tvalue))+ 
+  geom_bar(stat="identity") +
+  xlab("Fascicle") + 
+  ylab("T-value of AgexFascile Affected Regression") + 
+  ylim(-2,8) +
+  ggtitle("Effects of age within each fascicle ") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=30),
+        axis.title.x = element_text(size=30),
+        axis.title.y = element_text(size=30))
+print(plot_fascicle_by_t_sorted)
+```

### 2022-08-26 msdepression/scripts/promis_score_cleaning_2010-2022.Rmd
Add files via upload

---
title: "Cleaning PROMIS scores"
author: "Erica Baller"
date: '2022-08-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Goal of this script is to clean the promis scores and assemble them in a better table format for later processing.

Number of non NA values for each of the promis scores:

General_Health                  Quality_of_Life                  Physical_Health 
                            1694                             1693                             1503                             1503 
          Mental_Health_and_Mood   Social_Activities_Satisfaction   Carrying_Out_Social_Activities Carrying_Out_Physical_Activities 
                            1503                             1503                             1501                             1499 
              Emotional_Problems                  Fatique_Average     PostOp.PROMIS.Physical.Score       PostOp.PROMIS.Mental.Score 
                            1499                             1499                             1440                             1442

```{r score_cleaning}
#before doing this, must save the promis score spreadsheet, /Users/eballer/BBL/msdepression/data/dac/copyofpromisflowsheetdata2010-2022.xlsx in two separate csv files (because it has two tabs). This was done manually and separated by date. New files are:
#/Users/eballer/BBL/msdepression/data/dac/promis_data_2015.csv
#/Users/eballer/BBL/msdepression/data/dac/promis_data_2016-2022.csv

#read in spreadsheets
promis_data_2015 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/promisflowsheetdataupdated_good_physical_health_scores_20220824_sheet1.csv", sep = ",", header = T) #n=2041
promis_data_2016_2022 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/promisflowsheetdataupdated_good_physical_health_scores_20220824_sheet2.csv", sep = ",", header = T)#n=

promis_data_combined <- rbind(promis_data_2015, promis_data_2016_2022) #n=6916

#promis score names, from 7th column all the way to the end of the spreadsheet, starting with general health to PostOp.PROMIS.Mental.Score

promis_score_names <- names(promis_data_combined[7:dim(promis_data_combined)[2]-1])

##### cleaning ####

#total scores n=1694
promis_data_cleaned <- promis_data_combined %>%
  mutate(across(everything(), gsub, pattern = "NULL", replacement = "NA")) %>% #replace null with NA for easier use
  filter(RECORDED_TIME != "NA") %>% #remove nulls in Recorded time
  mutate(date_promis_collected = as.Date(RECORDED_TIME, format = "%m/%d/%y")) %>% 
  mutate(date_promis_collected = gsub(date_promis_collected, pattern = "-", replacement = "")) %>% #get time into better format, now matches other spreadsheet
  mutate(year_promis_collected = gsub(date_promis_collected, pattern = "\\d{4}$", replacement = "")) %>% #make new column that just as date and change to numeric
  mutate(across(everything(), gsub, pattern = "\\s*-.*", replacement = "")) %>% #\\s = whitespace, .* any character, replace with blank, to clean the text up
  mutate_at(promis_score_names, as.numeric) %>% #convert columns from character to numeric
  mutate(year_promis_collected = as.numeric(year_promis_collected)) %>% #convert to numeric
  mutate(RECORDED_TIME = as.Date(RECORDED_TIME, format = "%m/%d/%y")) #turn back into date format

#histos
hist(promis_data_cleaned$year_promis_collected) #want to see spread of when scores collected

for (i in promis_score_names) {
  eval_for_parse <- paste0("hist(promis_data_cleaned$", i, ")")
  eval(parse(text=eval_for_parse))
}

#look at number of samples in each column promis_data_cleaned
colSums(!is.na(promis_data_cleaned))

unique_individuals <- promis_data_cleaned %>% 
  group_by(EMPI) %>% 
  slice(1) %>%
  ungroup #n = 501

total_num_promis_scores <- dim(promis_data_cleaned)[1] #n= 1694

num_unique_indiv <- dim(unique_individuals)[1] #n=501

saveRDS(promis_data_cleaned, file = paste0("/Users/eballer/BBL/msdepression/data/dac/promis_data_20220824_fixed_physical_health_cleaned_n", total_num_promis_scores, "_unique_indiv_n", num_unique_indiv, "_2015-2022.rds"))
```

### 2022-08-26 msdepression/scripts/make_binary_colored_depression_net_maps.R

@@ -1,0 +1,41 @@
+#### Welcome to make_binary_colored_depression_net_maps! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle according to whether or not it is in the depression map. Yellow is dep map, blue is non dep map (though this is flexible)
+
+### Pre: File of fascicle overlap with depression network proportions
+### Post: Output of RGB values in colors
+### Uses: Opens up proportions, sorts by overlap to get a name of a vector in top quartile
+###       iterate through each fascicle, if in the top quartile, color will be first color, if not, will be second color
+###       **** If you change colors!!! CHANGE SUFFIX***
+### Dependencies: Any R will do, need the names of the fibers in the depression network, and a file with proportions
+
+
+homedir="/Users/eballer/BBL/msdepression/"
+
+dep_network_names_top_quartile_no_cranial_nerves_by_vol <- read.csv(file = paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), header = F)
+
+files = c("fascicle_proportion_overlap_with_depression_network_n77")
+
+#color_depressed_net = "255 0 255" #pink
+#color_nondepressed_net = "0 255 255" #light blue
+color_depressed_net = "255 255 0" #yellow
+color_nondepressed_net = "0 0 255" #blue
+
+for (filename in files) {
+  #suffix, CHANGE IF YOU WANT A DIFFERENT FLAG
+  suffix = "yb"
+  
+  #read in file with fascicle name and proportion
+  fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
+  names(fascicle_name_and_prop) <- c("fascicle", "proportion")
+  
+  #output color file (77 lines, each represents a color), and color name vector. Not sure how to do 1:1 assignments rather than manually, but I'll see if I can figure that out!
+  fascicle_name_and_prop_rgb_only <- fascicle_name_and_prop %>%
+    mutate(rgb = ifelse(fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol$V1, color_depressed_net, color_nondepressed_net)) %>%
+    dplyr::select(fascicle, rgb)
+  
[…]

### 2022-08-26 msdepression/scripts/make_red_to_yellow_RGB_color_scheme.R

@@ -1,0 +1,75 @@
+#### Welcome to make_red_to_yellow_RGB_color_scheme! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle to be its proportion of loss, scaled on the RGB scale
+### In order to do this, I needed to figure out how to map my proportions to RGB
+
+### Pre: File of proportions
+### Post: Output of RGB values in Red to Yellow Scale for the Proportions
+### Uses: Step 0: Figure out how to scale from red to yellow 
+###           - Red = 255 0 0, yellow 255 255 0
+###           - Means that we need to vary the middle value (255) to get where we need to be
+###       Step 1) For red to yellow, calculate the step (i.e. how much each value will be)
+###             - because my proportions are out of 1, I wanted each step to be 1/100
+###             - I am setting this value as "max", so it could vary if you have another max #
+###       Step 2) Multiply the proportion by 100, so now each count is 1-100, in percents
+###       Step 3) Multiply the percent by the step
+###       Step 4) Save this value in that fascicle, 255 [new_value] 0
+###
+### Dependencies: Any R will do
+
+###
+homedir="/Users/eballer/BBL/msdepression/"
+
+files = c("fascicle_mean_proportion_overlap_nfasc_77", "fascicle_proportion_overlap_with_depression_network_n77")
+
+for (filename in files) {
+  #suffix
+  suffix = paste0("_", filename, "_RGB")
+  
+  #read in file with fascicle name and proportion
+  fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
+  names(fascicle_name_and_prop) <- c("fascicle", "proportion")
+  
+  #set scale dynamically
+  #highest possible number in scale
+  max<-100 * max(fascicle_name_and_prop$proportion)
+  
+  #step that each value in our scale will correspond to
+  step<-255/max 
+  
+  # make a column for red
+  red <- rep(x=255, times = 77)
+  
+  fascicle_name_and_prop$red <- red
+  #multiply proportion column by 100, then by step, and round
+  fascicle_name_and_prop$green<- round((fascicle_name_and_prop$proportion *100*step),0)
+  
+  #make column for blue
[…]

### 2022-08-26 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_just_volumes_started_20220824.Rmd
Add files via upload

--- 2022-08-05 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
@@
 ---
-title: "MS Depression Paper"
+title: "MS Depression Paper, started 2022 08 24"
 author: "Erica Baller"
-date: '2022-07-28'
-output:
-  pdf_document: default
-  html_document: default
+date: '2022-08-24'
+output: html_document
 ---
@@
 library(lsr)
+library(sjPlot)
 #functions
@@
 
-## MS Depression
-
-Welcome to the MS Depression Data, installment 1 analysis markdown!
-This project was designed to ask a relatively simple question - are patients with MS who are depressed more likely to have lesions in fascicles that connect areas previously associated with depression.
-The literature are sparse in this domain.
-It is very clear that depression occurs in up to 50% of patients with MS, and some studies have suggested that it relates to lesion burden.
-But to my knowledge there are no papers that evaluate how lesions disrupt white matter tracts that serve as the backbone for functional networks.
-Over the course of the next few chunks, I'll walk through our analysis.
-It combines imaging data from research grade scans acquired at 3T as part of a clinical protocol (T1 and T2 Flair), and clinical data from the electronic medical record.
-
-## R Markdown
-
-Part 1: 
-The first chunk here is reading in appropriate data frames and preprocessing.
-Two big data frames are read in.
-The first is the demographic, drug, lab info for the MS patients (from the DAC pull).
-The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle).
-The most important preprocessing occurs in a series of mutate commands.
-These essentially add new columns to the data frame that can later be used.
-The thrust of these mutate commands is to define who has depression and who does not.
-
-Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion.
-Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses.
-For these preliminary analysis, our inclusion/exclusion are more stringent.
-Of note, the graph below is irrespective of whether they had good/bad MIMoSA.
-As such, the MIMoSA subset is smaller.
-
-<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>
-
-Inclusion criteria below:
-
-Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
-  1) Scanned under MS protocol (42K)
-  2) MS Diagnosis (17K)
-  3) Seen by MS doctor (16K)
-
-Inclusion criteria for Depressed Group
-  1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
-        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder)
-        - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition
-  2) Screened positive for depression on PHQ2 or PHQ9 at any point
-  3) Prescribed antidepressant medications per NAMI website
-        - https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications
-
-Inclusion criteria for Healthy Group
-  1) No F* diagnoses
-  2) At least 1 PHQ2 or PHQ9 with a documented 0
-  3) Free of psychiatric medications per NAMI website
-
-Exclusions
-  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
-  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression.
-These subjects were excluded from future analyses.
-We could consider adding these to the healthy group maybe, but not too inclined to do that right now.
-
-Imaging exclusion
-  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.
-
-Final N demographics 
-  Depressed: 859 (232 unique)
-  Super clean healthy: 604 (148 unique)
-
-(unique, with good MIMoSA):
-
-                         Healthy          Depressed             p      test
-  n                                148            232                     
-  Race (%)         Caucasian       108 ( 73.0)    173 ( 74.6)   0.821     
-                   Non-caucasian    40 ( 27.0)     59 ( 25.4)             
-  Sex (%)          Female          117 ( 79.1)    199 ( 85.8)   0.117     
-                   Male             31 ( 20.9)     33 ( 14.2)             
-  Age (mean (SD))                46.57 (12.66)  48.75 (11.75)   0.089     
-  Depression (%)   1               148 (100.0)      0 (  0.0)  <0.001     
-                   2                 0 (  0.0)    232 (100.0)             
-  PHQ2 (mean (SD))                0.00 (0.00)    0.58 (1.53)   <0.001     
-  PHQ9 (mean (SD))                0.00 (NA)     10.43 (9.02)       NA
-
-Networks
-
-  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1).
-It is an unthresholded mask (he initially sent one where the unit is r, but then sent us one where unit is T).
-  Per Siddiqi's article: The location of each lesion or brain stimulation site (Fig.
-2a–c, top panels) was mapped to an underlying brain circuit using a large normative connectome database (n=1000) and previously validated methods (Fig.
-2a–c, bottom panels)5.
-The normative connectome was used to estimate connectivity of each lesion or stimulation site to every voxel in the brain.
-At each voxel, a Pearson r-value was computed for the correlation between depression score and lesion or stimulation site connectivity to that voxel (Fig.
-2a–c, right panels), yielding a population-derived “circuit map” for each of the 14 datasets (Supplementary Figure 1).
-We generated a combined depression circuit map by taking the mean of all 14 circuit maps, weighted by the sample size of each dataset (Fig.
-5a).
-Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted)
-
-  To use it, I thresholded the clusters (3.09), binarized it and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle.
-Most fascicles were either entirely overlapping or entirely outside the depression mask, though the range of overlap was anywhere from 1-30%.
-This required us to make some choices.
-We considered an option where the "depression mask" was anywhere that ever overlapped, and non-depression were areas where there was no overlap.
-We also looked at quartiles, and settled on the top 25% (top quartile), i.e.
-the top 25% of fascicles with the highest proportion of network overlap were considered in the depression network.
-Everything outside of that was considered "non_depression" network, or nondep_net.
-Cranial nerves are too small to reliably be assess with our method, and so they were removed from the analysis.In total, 77 fascicles were evaluated.
-
-Depression top quartile, r&l separated
- [1] "SLF3_R"  "AF_R"    "AF_L"    "SLF2_R"  "CBT_R"   "FAT_R"   "CS_P_R"  "C_PHP_R" "MdLF_L"  "FAT_L"   "SLF2_L"  "MdLF_R"  "CBT_L"   "TR_P_L"  "TR_S_R" 
-[16] "CPT_O_R" "TR_P_R"  "PAT_L"   "PAT_R"   "EMC_R"   "TR_S_L"  "CC"
-
-Notebook Overview
-
-Analyses
-  There are two main types of analyses I performed, t.tests (for example, to compare inside versus outside depression network) and linear models (for example, looking at the relationship of depression to each individual fascicle).
-With respect to T-tests, I assume unequal variance for all.
-For linear models, I look only at the first scan.
-The general pattern is to look at the total proportion of volume lost across fascicles, then within depression network, and within non depression network.
-Lastly, I look at overall mean lesion burden irrespective of location by counting the number of nonzero voxels in each individual's mimosa mask.
-
-Results
-  Depression group: 232 with depression, 148 healthy:
-
-  A) Lesion volume between depressed and controls
-
-    - P=0.067, NS
-
-  B) Depressed vs Controls Fascicle Affected - Proportions
-
-    - Dep vs non-dep, total fascicle proportion affected p = 0.050 (NS), Dep = 31%, Non dep = 27%
-    - Dep vs non, top 25%, **p = 0.047**, Dep = 32%, non dep = 26% 
-    - Dep vs non, bottom 75%, no p = 0.057, dep = 30%, non dep = 27%
-
-  C) Depressed vs Controls Fascicle Affected - Volume
-
-    - Dep vs nondep, total fascicle volume affected: **p = 0.045**, Dep = 9382, Non dep = 8260
-    - Dep vs nondep in top 25% yes **p = 0.034**, Dep = 4829, non dep = 3914
-    - Dep vs nondep in bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
-
-  D) Linear models of values in fascicle associated with Overlap in Depression map
-
-    - T values correlated w/proportion of overlap of fascicle in depression map
-        - p of lm (0.010),
-        - 95%   ( 0.701,  4.553 )
-        - boostrap p calc from CI - **0.0087**
-
-    - Cohen'd correlated w/proportion of overlap of fascicle w/depression map
-        - p of lm (0.001)
-        -
-
-    - T values correlated w/volume of overlap with depression network (i.e. # voxels that fascicle shares w/depression map),
-        - p = 0.031
-        - 95%   ( 1.052,  3.724 )
-        - bootstrap p calc from CI- **1*10^-16**
-
-    - Cohen'd correlated w/volume of overlap of fascicle w/depression map
-        - p of lm (p < 2.2 * 10^-16)
-        - Largest effect in corpus callosum
-I think it works best when sticking with *volumes* rather than proportions
-
-#### Promis data, late edition
-
-We decided to look at some of the promis data (which measures broad functional categories, though not specifically psychiatric diagnosis) to see if a) measures differed by diagnostic groups and b) whether there were dimensional effects.
-Unfortunately, the PROMIS picked up in 2018, which was also the same year that Melissa ended her pull, so our n's are small.
-
-Current ns for people with at least 1 promis score= 
-Total n promis scores: 1694 
-Total n unique individuals w/promis scores: 501
-
-Different scores have different # of overall scans
-                        Total                      General_Health                  Quality_of_Life                  Physical_Health 
-                            1694                             1693                             1503                             1503 
-          Mental_Health_and_Mood   Social_Activities_Satisfaction   Carrying_Out_Social_Activities Carrying_Out_Physical_Activities 
-                            1503                             1503                             1501                             1499 
-              Emotional_Problems                  Fatique_Average     PostOp.PROMIS.Physical.Score       PostOp.PROMIS.Mental.Score 
-                            1499                             1499                             1440                             1442
-
-W/good mimosa:
-Total n promis scores : 398
-Total n unique individuals w/promis scores: 116
-Total n unique indivs in nondepressed group = 36
-Total n unique individuals in depressed group = 46
-
-Method:
-To use the promis scores, I cleaned them in a separate script (promis_score_cleaning_2010-2022.Rmd)
-In this script, I link each scan with its most proximal promis score.
-The challenge here is that there is no time linking.
-Summary of the time (in days and years) between scanning and a promis (these are absolute values, do not indicate before or after).
-
-Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-    1.0   376.8   763.0   873.3  1229.8  2487.0 
-    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
-0.002747 1.035027 2.096154 2.399132 3.378434 6.832418
-
-Next, I looked at overall differences between diagnostic groups on each of the PROMIS measures, correcting for multiple comparisons w/FDR (Q<0.05).
-
-Then, I looked at the relationship between promis scores and fascicle measures (overall, 1620 analyses done):
-
-Proportion:
-Proportion within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
-    - all participants (116)
-    - Depressed + nondepressed (82)
-    - Non-depressed alone (36)
-    - Depressed alone (46)
-
-Proportion of individual fascicle affected related to promis scores (777 analyses per group (11 promis domains x 77 individual fascicles))
-  - all participants (116)
-  - Depressed + nondepressed (82)
-  - Non-depressed alone (36)
-  - Depressed alone (46)
-
-Volume:
-
-Volume within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
-   - all participants (116)
-    - Depressed + nondepressed (82)
-    - Non-depressed alone (36)
-    - Depressed alone (46)
-
-Volume of individual fascicle affected related to promis scores  (777 analyses per group (11 promis domains x 77 individual fascicles))
-  - all participants (116)
-  - Depressed + nondepressed (82)
-  - Non-depressed alone (36)
-  - Depressed alone (46)
-
-Promis results:
-Interestingly, across all 11 scores, with the exception of fatigue, Depressed patients have lower scores (lower = worse functioning), than non-depressed patients.
-controls.
-They all fdr correct.
-What I love about this is fatigue not being different.
-If fatigue was worse in depressed patients, one could argue that quality of life, health, emotional problems, social activities, etc are all worse because of underlying fatigue.
-Similarly, as fatigue is a component of the PHQ9 and depression screening, this could be confounding.
-That they were different in a lot of measures but not fatigue is informative.
-
-Quality_of_Life	0.015199144			
-Physical_Health	0.012579264			
-Mental_Health_and_Mood	0.012579264			
-Social_Activities_Satisfaction	0.001063449			
-Carrying_Out_Social_Activities	0.011375799			
-Carrying_Out_Physical_Activities	0.031616473			
-Emotional_Problems	0.011270683			
-Fatique_Average	0.231497853			
-PostOp.PROMIS.Physical.Score	0.041172446			
-PostOp.PROMIS.Mental.Score	0.001253404
-
-Proportion
-Proportion within versus outside depression map correlated w/promis scores - NO SIGNIFICANT FINDINGS
-    - all participants (116)
-    - Depressed + nondepressed (82)
-    - Non-depressed alone (36)
-    - Depressed alone (46)
-
-Proportion of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
-  - all participants (116)
-  - Depressed + nondepressed (82)
-  - Non-depressed alone (36)
-  - Depressed alone (46)
-
-Volume     
-Volume within versus outside depression map correlated w/promis scores -  NO SIGNIFICANT FINDINGS
-   - all participants (116)
-    - Depressed + nondepressed (82)
-    - Non-depressed alone (36)
-    - Depressed alone (46)
-
-Volume of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
-  - all participants (116)
-  - Depressed + nondepressed (82)
-  - Non-depressed alone (36)
-  - Depressed alone (46)
+Read in Data and do some preprocessing
 
@@
 fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
+fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77, drop cranial nerves
+
 #write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
+
+#add suffix to each measure
+fascicle_names_w_prop_suffix <- gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE, x = fascicle_names_no_cranial_nerves)
+fascicle_names_w_vol_suffix <- gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE, x = fascicle_names_no_cranial_nerves)
 
@@
 #add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
-fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
-  replace(is.na(.), 0) %>%
-  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))
+#fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
+ # replace(is.na(.), 0) %>%
+#  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))
 
 #Write out the new network groupings to a file, will be used later
-fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
-write.table(fascicle_volumes_dep_network, file=fascicle_names_dep_network_grouping, sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
+#fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
+#write.table(fascicle_volumes_dep_network, file=fascicle_names_dep_network_grouping, sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
 
 #fascicle_mapping, returns numerical mapping as well as names of tracts, and whether any overlap with depression network (anything above 1 is overlap), as well as whether the fascicle overlaps, and the proportion of overlap with yeo networks
-fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)
+#fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)
 
@@
 ##########################################################
+#  Add columns for volume and change proportion names    #
+##########################################################
+
+#read in fiber volume values
+fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_prop <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]
+df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]
+
+#multiply each row by the vector of total volumes to extract volume of disease within fascicle
+df_fascicles_volumes <- sweep(x = df_fascicles_prop, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values$volume_full, FUN = "*")
+
+# change name of df_fascicle_volumes to have the suffix, and add a suffix for the proportion measures
+names(df_fascicles_prop) <- fascicle_names_w_prop_suffix
+names(df_fascicles_volumes) <- fascicle_names_w_vol_suffix
+
+#combine back with data_fram
+#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+df_demo_and_fascicles_prop_and_vol <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes )
+
+```
+PROMIS DATA HANDLING to link a person to their most proximal promis
+
+```{r promis_data_handling}
+##########################################################
 #            Read in PROMIS Data              #
@@
 
-df_people_with_promis_data <- df_demo_and_fascicles %>%
+df_people_with_promis_data <- df_demo_and_fascicles_prop_and_vol %>%
   filter(EMPI %in% promis_data$EMPI) #n=398
@@
 
-```
-```{r make_df_with_volume_instead_of_proportion}
-#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
-
-#read in fiber volume values
-fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
-
-#extract data frame that only includes the fascicle proportion lost
-df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
-df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
-
-#multiply each row by the vector of total volumes
-df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
-
-#combine back with data_fram
-df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
-
-```
-
-```{r promis_data_handling}
 #go through each person w/promis data and append the columns to their row that reflects the closest collected PROMIS
@@
 #summarize the distance from scans
-
-summary(df_people_with_promis_data_scores_added$distance_from_scan) #mean 873.3 days, median 763 days, range 1 to 2487 days, convert to years
-hist(df_people_with_promis_data_scores_added$distance_from_scan) #convert to years
-
 summary(df_people_with_promis_data_scores_added$distance_from_scan/364) #mean 2.40 ys, median 2.10 years, range 0.003 y to 6.83 years
-hist(df_people_with_promis_data_scores_added$distance_from_scan/364) #convert to years
-
-#sub volume for proportion
-#extract data frame that only includes the fascicle proportion lost
-df_fascicles_only <- df_people_with_promis_data_scores_added[ , (names(df_people_with_promis_data_scores_added) %in% fascicle_names)]
-df_everything_else <- df_people_with_promis_data_scores_added[ , !(names(df_people_with_promis_data_scores_added) %in% fascicle_names)]
-
-#multiply each row by the vector of total volumes
-df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
-
-#combine back with data_fram
-df_demo_and_fascicle_volumes_full_w_promis <- cbind(df_everything_else, df_fascicle_volumes)
+hist(df_people_with_promis_data_scores_added$distance_from_scan/364, 
+     main = "Hisogram of years between Promis and Scan",
+     xlab = "Years from Scan") #convert to years
 
@@
 
+Demographics
 ```{r demographics}
@@
 
-make_demographics_table_ms(data_frame = data_empi_acc_f_phq)
-
-df_unique_empi <- data_empi_acc_f_phq %>% 
+df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles %>% 
+  filter(depGroupVar != 0) %>%
   group_by(EMPI) %>% 
@@
   ungroup() 
+make_demographics_table_ms(data_frame = df_good_mimosa_unique_dep_and_healthy)
 
-make_demographics_table_ms(data_frame = df_unique_empi)
-
-df_just_healthy_and_depressed <- df_unique_empi %>% filter(depGroupVar != 0)
-
-make_demographics_table_ms(data_frame = df_just_healthy_and_depressed)
-
-#just people with good mimosa
-df_good_mimosa <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
-make_demographics_table_ms(data_frame = df_good_mimosa)
-
-df_good_mimosa_unique <- df_demo_and_fascicles %>% 
-#  filter(depGroupVar != 0) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() 
-make_demographics_table_ms(data_frame = df_good_mimosa_unique)
-
-### with promis data
-make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added)
-
-## overall
 #just good mimosa w/promis
-df_good_mimosa_promis <- df_people_with_promis_data_scores_added %>% filter(depGroupVar != 0)
-make_demographics_table_ms_w_promis(df_good_mimosa_promis) #n=287 scans
-
-#unique
 df_people_with_promis_data_scores_added_unique <- df_people_with_promis_data_scores_added %>%
@@
   ungroup() 
-
 make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added_unique)
@@
 ```
+Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups
 
@@
 #separate out only people in depression group, and take first instance
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
   filter(depGroupVar != 0)  %>% 
@@
 
-Diagnosis <- c("MS w/Depression", "MS No Depression")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
+#extract overall volume and diagnosis
+df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
+  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
+  filter(depGroupVar != 0) %>% 
+  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
 
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
-
-data <- data.frame(Diagnosis,value, sem)
-
-# Grouped
-barplot_lesion_volume <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
-  ylab("Volume of Lesion") +
-  theme(title = element_text(size=14),
+violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
+  geom_violin() + 
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
+  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
+  ylab("Volume of Lesions") +
+  theme(plot.title = element_text(size=16, hjust = 0.8),
         axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=20),
+        axis.title.y = element_text(size=16),
         axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
         panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
-#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
-print(barplot_lesion_volume) #p=0.067
+print(violin_lesion_volume) #p=0.067
 
 ```
-
-```{r visualize_depression_network}
+Define Depression Network by Volume of Overlap
+```{r visualize_depression_network_volume}
 ##########################################################
@@
 ######## Histograms for depression network
-#sort based on overlap
-fascicle_values_dep_sorted_by_overlap <- fascicle_volumes_dep
-fascicle_values_dep_sorted_by_overlap$prop_in_mask[is.na(fascicle_values_dep_sorted_by_overlap$prop_in_mask)] = 0
-fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask)])
+#first set the values to be the volumes
+fascicle_volumes_dep_network_sorted_by_overlap <- fascicle_volumes_dep
+
+#set NAs to zero
+fascicle_volumes_dep_network_sorted_by_overlap$non_zero_voxels_in_dep_map[is.na(fascicle_volumes_dep_network_sorted_by_overlap$non_zero_voxels_in_dep_map)] = 0
+
+#sort based on overlap with depression depression map
+fascicle_volumes_dep_network_sorted_by_overlap$fascicle <- factor(fascicle_volumes_dep_network_sorted_by_overlap$fascicle, levels = fascicle_volumes_dep_network_sorted_by_overlap$fascicle[order(fascicle_volumes_dep_network_sorted_by_overlap$non_zero_voxels_in_dep_map)])
 
 #remove fascicles with cranial nerves
-fascicles_r_and_l_sep_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_dep_network_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
 
-plot_all_fascicles_r_and_l_sep_no_cranial_nerves <-ggplot(data = fascicles_r_and_l_sep_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask))+
-  ggtitle("Proportion of Volume Overlap per Fascicle 3_09") + 
-  geom_bar(stat="identity") +
+plot_all_fascicle_volumes_dep_network_no_cranial_nerves <-ggplot(data = fascicle_volumes_dep_network_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
+  ggtitle("Volume Overlap per Fascicle 3_09") + 
+  geom_bar(stat="identity", fill="006300") +
   xlab("Fascicle") +
   ylab("Overlap with Depression Mask") +
-  ylim(0,0.4) +
+  ylim(0,40000) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
@@
         axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
-print(plot_all_fascicles_r_and_l_sep_no_cranial_nerves)
-
-### remove all rows with cranial nerves in them
-# of note, we do not have all cranial nerves represented, we only have, CNII (optic), III(oculomotor), V (trigeminal), VII(facial), VIII(vestibulocochlear)
-fascicle_values_dep_no_cranial_nerves = fascicle_values_dep_sorted_by_overlap[(grep(pattern = "CN", x = fascicle_values_dep_sorted_by_overlap$fascicle, invert = TRUE)),]
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(plot_all_fascicle_volumes_dep_network_no_cranial_nerves)
 
 #figure out how many fascicles that is, then 
-quartile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 4),0)
+quartile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 4),0)
 
 #sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
-dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile] #n=19
+dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile] #n=19
+
+dep_network_total_volume_top_quartile_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)])
 
 #get a network that includes all of the fascicles NOT in dep 25% network
-dep_network_names_bottom_3_quartiles_no_cranial_nerves <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)] #n=68
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol <- fascicle_volumes_dep_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)] #n=68
 
-#plot values within the depression network, and values outside of it
+bottom_75_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol)])
 
-fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves),]
+#### Calculate total volume in all non-cranial_nerve_network
+total_volume_all_full_fascicles <- dep_network_total_volume_top_quartile_no_cranial_nerves + bottom_75_total_volume_no_cranial_nerves
+#plot volumes within the depression network, and volumes outside of it
 
-plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
-  geom_bar(stat="identity") +
-  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, top quartile") + 
-  ylim(0,0.4) +
-  ylab("Proportion in Depression Mask") +
+fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_volumes_dep_sorted_by_overlap[(fascicle_volumes_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol),]
+
+plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
+  geom_bar(stat="identity", fill="violet") +
+  ggtitle("Volume per Fascicle Affected Dep net 3_09, top quartile") + 
+  ylab("Volume in Depression Mask") +
+  ylim(0,40000) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
@@
         axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
 print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)
@@
 #bottom 3 quartiles
-fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves),]
+fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves<- fascicle_volumes_dep_sorted_by_overlap[(fascicle_volumes_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol),]
 
-plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
-  geom_bar(stat="identity") +
-  ylab("Proportion in Depression Mask") +
-  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
-  ylim(0,0.4) +
+plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
+  geom_bar(stat="identity", fill="darkgreen") +
+  ylab("Volume in Depression Mask") +
+  ylim(0,40000) +
+  ggtitle("Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5))
+        plot.title = element_text(hjust = 0.5),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
 print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs
@@
 ```
-
-```{r proof_of_concept}
-
-#this section creates histograms looking at the proportion of volume lost in each fascicle across subjects
-#we recapitulate earlier work and show that higher volume loss with our method reflects regions previously known to have higher volume loss
-#These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis
-
-data_empi_acc_f_phq_unique <- data_empi_acc_f_phq %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-#have to pull out just the pieces I'm interested in
-just_dep_and_empi <- subset(data_empi_acc_f_phq, select = c("EMPI", "depGroupVar"))
-added_dep_to_fascicles <- unique(merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI")))
-
-melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))
-
-#exclude regions where there was no lesion at all
-lesioned <- subset(melted_df, value > 0)
-
-#only include people in healthy or depressed group
-lesioned_dx <- subset(lesioned, depGroupVar != 0)
-
-#separate out healthy and depressed
-healthy <- subset(lesioned, depGroupVar == 1)
-depressed <- subset(lesioned, depGroupVar == 2)
-
-q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
-r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
-r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
-x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
-y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
-z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
-
-#print(q)
-#print(r)
-print(r_color)
-#print(x)
-#print(y)
-#print(z)
-
-###########################################################################################################
-##### Looking at distribution of disease across fascicles across group and within healthy and depressed ###
-###########################################################################################################
-
-#### look at mean fascicle volume across participants and transpose
-# drop exam date and empi, and transpose
-fascicles_alone_df <- t(added_dep_to_fascicles[,!(colnames(added_dep_to_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
-fascicle_names <- rownames(fascicles_alone_df)
-
-#take row means
-fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
-names(fascicles_alone_rowMeans) <- c("fascicle", "value")
-
-#GET RID OF CRANIAL NERVES
-fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
-
-#sort based on overlap
-fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
-
-#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
-order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]
-
-fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
-
-######## Histograms/bargraphs
-#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
-hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)
-
-sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
-  ggtitle("Proportion of Volume per Fascicle Affected") + 
-  ylab("Proportion") + 
-  geom_bar(stat="identity") +
-  ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5, size=14),
-        axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
-print(sorted_ggplot)
-
-######### 
-#############
-#### look at mean fascicle volume across healthy participants and transpose
-#############
-
-# drop exam date and empi, and transpose
-healthy_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 1),]
-fascicles_alone_df <- t(healthy_group_fascicles[,!(colnames(healthy_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
-fascicle_names <- rownames(fascicles_alone_df)
-
-#take row means
-fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
-names(fascicles_alone_rowMeans) <- c("fascicle", "value")
-
-#GET RID OF CRANIAL NERVES
-fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
-
-#sort based on overlap
-fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
-
-fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
-
-######## Histograms/bargraphs
-
-# ordered by mean value overall
-fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
-sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
-  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
-  geom_bar(stat="identity") +
-  ylab("Proportion") +
-  ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5, size=14),
-        axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
-print(sorted_ggplot_healthy_ordered)
-
-###########
-#### look at mean fascicle volume across depressed participants and transpose
-###########
-# drop exam date and empi, and transpose
-depressed_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 2),]
-fascicles_alone_df <- t(depressed_group_fascicles[,!(colnames(depressed_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
-fascicle_names <- rownames(fascicles_alone_df)
-
-#take row means
-fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
-names(fascicles_alone_rowMeans) <- c("fascicle", "value")
-
-#GET RID OF CRANIAL NERVES
-fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
-
-#sort based on overlap
-fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans
-
-fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])
-
-######## Histograms/bargraphs
-
-# ordered by mean value overall
-fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
-sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
-  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
-  ylab("Proportion Affected") +
-  geom_bar(stat="identity") +
-  ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5, size=14),
-        axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
-print(sorted_ggplot_depressed_ordered)
-
-##### look at mean fascicles affected in dep vs non dep network
-######
-#plot values within the depression network, and values outside of it
-
-fascicles_dep_25_percent_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves),]
-
-plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=value))+
-  ggtitle("Proportion of Volume per Fascicle Affected Dep net, top quartile") + 
-  geom_bar(stat="identity") +
-  ylab("Proportion") +
-  ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5, size=14),
-        axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
-print(plot_all_fascicles_r_and_l_sep_top_quartile)
-
-#bottom 3 quartiles
-fascicles_dep_bottom_3_quartiles_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves),]
-
-plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=value))+
-  ggtitle("Proportion of Volume per Fascicle Affected, bottom three quartiles") + 
-  ylab("Proportion") +
-  geom_bar(stat="identity") +
-  ylim(0,0.8) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5, size=14),
-        axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14))
-print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)
-
-```
-
-```{r depression_vs_healthy_overall_proportion_of_fascicles_affected}
-
-#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
-#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
-  filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-#test whether proportion lost overall different between depressed and healthy- No, p = 0.050, Dep = 31%, Non dep = 27%
-total_prop_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
-
-#look at within depression network using top quartiles, yes p = 0.047, Dep = 32%, non dep = 26%
-total_prop_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
-
-#look at bottom 75%, no p = 0.057, dep = 30%, non dep = 27%
-total_prop_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
-
-############
-#bar plot
-############
-
-#over all fascicles
-Diagnosis <- c("MS w/Depression", "MS No Depression")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
-
-data <- data.frame(Diagnosis,value, sem)
-barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-  ylab("Mean Proportion of Fascicles Affected") +
-  ggtitle(paste0("Total Proportion Affected by dx, p=", round(total_prop_lost_dep_vs_healthy_unique$p.value,3))) + 
-  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=20),
-        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-        plot.title = element_text(hjust=0.5, size = 14), 
-        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
-
-print(barplot_overall)
-
-#in top quartile
-
-Diagnosis <- c("MS w/Depression", "MS No Depression")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
-
-data <- data.frame(Diagnosis,value, sem)
-
-# Grouped
-
-barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-  ylab("Mean proportion of Fascicles Affected") +
-    ggtitle(paste0("Total Propotion Affected by dx in Dep 25%, p=", round(total_prop_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
-  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=20),
-        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-        plot.title = element_text(hjust=0.5, size = 14), 
-        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) 
-print(barplot_top_quartile)
-
-#in bottom 3 quartiles
-Diagnosis <- c("MS w/Depression", "MS No Depression")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
-
-data <- data.frame(Diagnosis,value, sem)
-
-# Grouped
-
-barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-  ylab("Mean Proportion of Fascicles Affected") +
-    ggtitle(paste0("Total Proportion Affected by dx, Bottom 75% p=", round(total_prop_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
-  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=20),
-        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-        plot.title = element_text(hjust=0.5, size = 14), 
-        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52))
-
-print(barplot_bottom_3_quartiles)
-
-```
-
-```{r depression_vs_healthy_overall_volume_of_fascicles_affected}
-
-#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
-#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
-
-#read in fiber volume values
-#fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
-
-#extract data frame that only includes the fascicle proportion lost
-#df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
-#df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
-
-#multiply each row by the vector of total volumes
-#df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
- #                 STATS = fiber_volume_values$volume_full, FUN = "*")
-
-#combine back with data_fram
-#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
-  filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-#test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
-total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054
-
-#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
-total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)
-
-#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
-total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
-
-############
-#bar plot
-############
-
-#over all fascicles
-Diagnosis <- c("MS w/Depression", "MS No Depression")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
-
-data <- data.frame(Diagnosis,value, sem)
-barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-  ylab("Mean volume of Fascicles Affected") +
-  ggtitle(paste0("Total Volume Affected by dx, p=", round(total_volume_lost_dep_vs_healthy_unique$p.value,3))) + 
-  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=20),
-        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-        plot.title = element_text(hjust=0.5, size = 14), 
-        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) #comment out if you want colored graph
-
-print(barplot_overall)
-
-#in top quartile
-
-Diagnosis <- c("MS w/Depression", "MS No Depression")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
-
-data <- data.frame(Diagnosis,value, sem)
-
-# Grouped
-
-barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-  ylab("Mean volume of Fascicles Affected") +
-    ggtitle(paste0("Total Volume Affected by dx in Dep 25%, p=", round(total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
-  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=20),
-        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-        plot.title = element_text(hjust=0.5, size = 14), 
-        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) 
-print(barplot_top_quartile)
-
-#in bottom 3 quartiles
-Diagnosis <- c("MS w/Depression", "MS No Depression")
-value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))
-
-sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
-           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))
-
-data <- data.frame(Diagnosis,value, sem)
-
-# Grouped
-
-barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-  ylab("Mean volume of Fascicles Affected") +
-    ggtitle(paste0("Total Volume Affected by dx, Bottom 75% p=", round(total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
-  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=20),
-        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-        plot.title = element_text(hjust=0.5, size = 14), 
-        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000))
-
-print(barplot_bottom_3_quartiles)
-
-```
-
-```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop}
-# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups
-
-##############################
-## FASCICLES TO BE INCLUDED ##
-##############################
-
-#define the depression network (all areas that intersect with depression network, no cranial nerves)
-fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
-
-#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
-fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
-
-#######################
-### depression map  ###
-#######################
-
-### define depression map
-fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
-
-#filter to exclude fascicles w/cranial nerves
-fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
-  replace(is.na(.), 0) %>%
-  filter(fascicle%in% fascicle_names_no_cranial_nerves)
-
-#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column
-
-df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))
-
-#######################
-### real data x dx  ###
-#######################
-
-#then, make the data frame
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-# lapply to get results of t-test in each fascicle
-#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
-fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
-{
-
-  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
- eval(parse(text=t_test_command))
-})
-names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves
-
-#pull t-statistic
-dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)
-
-#convert to data frame
-df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
-row.names(df_t_values) <- fascicle_names_no_cranial_nerves
-
-#pull p-values
-dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)
-
-#convert to data frame
-df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
-row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves
-
-#FDR correct p-values
-dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
-
-#Convert to data frame
-df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
-row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves
-
-#######################
-### plot differences ##
-#######################
-
-#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
-
-df_fascicle_dep_mask_overlap_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_t_values$dep_vs_healthy_t_values)
-names(df_fascicle_dep_mask_overlap_t_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_t")
-
-#make copy to sort
-df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_t_value
-
-df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])
-
-# bar graph of T-values of fascicles, sorted by depression overlap
-plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") +
-  geom_bar(stat="identity") +
-  ylim(0,5) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(plot_fascicle_by_t_sorted_by_dep_overlap)
-
-# plot fascicle overlap x t value
-#statistic, p = 0.010
-dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_t_value)
-print(summary(dep_overlap_x_dep_vs_healthy_t))
-
-scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_t)) +
-  geom_point(size=1) + 
-  geom_smooth(method=lm)+
-  xlab("Proportion of Fascicle in Depression Network") + 
-  ylab("Effect of diagnosis calculated in proportions (T)") + 
-  ggtitle(paste0("Fascicle/depression mask Proportion overlap by dx T, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t)$coefficients[2,4],3))) + 
-  theme(plot.title=element_text(hjust=0.5))
-print(scatter_t_val_by_dep_net_overlap)
-
-```
-
-```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop_effect_size}
-# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups
-
-##############################
-## FASCICLES TO BE INCLUDED ##
-##############################
-
-#define the depression network (all areas that intersect with depression network, no cranial nerves)
-fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
-
-#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
-fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
-
-#######################
-### depression map  ###
-#######################
-
-### define depression map
-fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
-
-#filter to exclude fascicles w/cranial nerves
-fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
-  replace(is.na(.), 0) %>%
-  filter(fascicle%in% fascicle_names_no_cranial_nerves)
-
-#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column
-
-df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))
-
-#######################
-### real data x dx  ###
-#######################
-
-#then, make the data frame
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-# lapply to get results of t-test in each fascicle
-#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
-fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
-{
-  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
-  eval(parse(text=text_for_eval))
-})
-names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves
-
-#convert to data frame
-df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)
-
-#######################
-### plot differences ##
-#######################
-
-#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
-
-df_fascicle_dep_mask_overlap_d_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_d_values)
-names(df_fascicle_dep_mask_overlap_d_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_d")
-
-#make copy to sort
-df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_d_value
-
-df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])
-
-# bar graph of T-values of fascicles, sorted by depression overlap
-plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") +
-  geom_bar(stat="identity") +
-  ylim(-0.1,0.5) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(plot_fascicle_by_d_sorted_by_dep_overlap)
-
-# plot fascicle overlap x t value
-#statistic, p = 0.010
-dep_overlap_x_dep_vs_healthy_d <- lm(dep_vs_healthy_d ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_d_value)
-print(summary(dep_overlap_x_dep_vs_healthy_d))
-
-scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_d_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_d)) +
-  geom_point(size=1) + 
-  geom_smooth(method=lm)+
-  xlab("Proportion of Fascicle in Depression Network") + 
-  ylab("Effect of diagnosis calculated in proportions (D)") + 
-  ggtitle(paste0("Fascicle/depression mask Proportion overlap by dx Eff Size, p = ", round(summary(dep_overlap_x_dep_vs_healthy_d)$coefficients[2,4],3))) + 
-  theme(plot.title=element_text(hjust=0.5))
-print(scatter_d_val_by_dep_net_overlap)
-
-```
-
-```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume}
-
-#define the depression network (all areas that intersect with depression network, no cranial nerves)
-fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
-
-#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
-fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
-
-#######################
-### depression map  ###
-#######################
-
-### define depression map
-fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
-
-#filter to exclude fascicles w/cranial nerves
-fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
-  replace(is.na(.), 0) %>%
-  filter(fascicle%in% fascicle_names_no_cranial_nerves)
-
-df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
-
-########################
-## Get Volume Values ###
-########################
-
-#read in fiber volume values
-fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
-
-#extract data frame that only includes the fascicle proportion lost
-df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
-df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
-
-#multiply each row by the vector of total volumes
-df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
-
-#combine back with data_fram
-df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
-
-#isolate out only the depressed group and healthy group, fascicles should be in volumes rather than proportions
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-# lapply to get results of t-test in each fascicle
-#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
-fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
-{
-  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
-  eval(parse(text=t_test_command))
-})
-names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves
-
-#pull t-statistic
-dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)
-
-#convert to data frame
-df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
-row.names(df_t_values) <- fascicle_names_no_cranial_nerves
-
-#pull p-values
-dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)
-
-#convert to data frame
-df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
-row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves
-
-#FDR correct p-values
-dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
-
-#Convert to data frame
-df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
-row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves
-
-#######################
-### plot differences ##
-#######################
-
-#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
-
-df_fascicle_dep_mask_volume_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
-names(df_fascicle_dep_mask_volume_t_value) <- c("fascicle", "dep_vs_healthy_t", "volume_in_dep_mask")
-
-df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask <- df_fascicle_dep_mask_volume_t_value
-
-#order by volume in depression mask
-df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$volume_in_dep_mask)])
-
-# bar graph of T-values of fascicles, sorted by depression overlap
-plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by volume of dep network") +
-  geom_bar(stat="identity") +
-  ylim(0,5) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(plot_fascicle_by_t_sorted_by_dep_overlap)
-
-#statistic volume, p = 0.031
-dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value)
-print(summary(dep_overlap_x_dep_vs_healthy_t))
-
-scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_t_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
-  geom_point(size=1) + 
-  geom_smooth(method=lm)+
-  xlab("Volume of Fascicle in Depression Network") + 
-  ylab("Effect of diagnosis (T)") + 
-  ggtitle(paste0("Fascicle/depression mask volume by dx T, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t)$coefficients[2,4],3))) + 
-  theme(plot.title=element_text(hjust=0.5))
-print(scatter_t_val_by_dep_net_overlap)
-
-```
-
-```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume_no_CC}
-
-#define the depression network (all areas that intersect with depression network, no cranial nerves)
-fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
-
-#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
-fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
-
-#######################
-### depression map  ###
-#######################
-
-### define depression map
-fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
-
-#filter to exclude fascicles w/cranial nerves
-fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
-  replace(is.na(.), 0) %>%
-  filter(fascicle%in% fascicle_names_no_cranial_nerves)
-
-df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
-
-########################
-## Get Volume Values ###
-########################
-
-#read in fiber volume values
-fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
-
-#extract data frame that only includes the fascicle proportion lost
-df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
-df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
-
-#multiply each row by the vector of total volumes
-df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
-
-#combine back with data_fram
-df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
-
-#isolate out only the depressed group and healthy group, fascicles should be in volumes rather than proportions
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-# lapply to get results of t-test in each fascicle
-#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
-fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
-{
-  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
-  eval(parse(text=t_test_command))
-})
-names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves
-
-#pull t-statistic
-dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)
-
-#convert to data frame
-df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
-row.names(df_t_values) <- fascicle_names_no_cranial_nerves
-
-#pull p-values
-dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)
-
-#convert to data frame
-df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
-row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves
-
-#FDR correct p-values
-dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
-
-#Convert to data frame
-df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
-row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves
-
-#######################
-### plot differences ##
-#######################
-
-#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
-
-df_fascicle_dep_mask_volume_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
-names(df_fascicle_dep_mask_volume_t_value) <- c("fascicle", "dep_vs_healthy_t", "volume_in_dep_mask")
-
-#remove corpus callosum
-df_fascicle_dep_mask_volume_t_value_no_cc <- df_fascicle_dep_mask_volume_t_value %>%
-  filter(fascicle != "CC")
-
-df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask <- df_fascicle_dep_mask_volume_t_value_no_cc
-
-#order by volume in depression mask
-df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$volume_in_dep_mask)])
-
-# bar graph of T-values of fascicles, sorted by depression overlap
-plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by volume of dep network, no CC") +
-  geom_bar(stat="identity") +
-  ylim(0,5) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(plot_fascicle_by_t_sorted_by_dep_overlap)
-
-#statistic volume, p = 0.031
-dep_overlap_x_dep_vs_healthy_t_no_cc <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value_no_cc)
-print(summary(dep_overlap_x_dep_vs_healthy_t_no_cc))
-
-scatter_t_val_by_dep_net_overlap_no_cc <- ggplot(data=df_fascicle_dep_mask_volume_t_value_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
-  geom_point(size=1) + 
-  geom_smooth(method=lm)+
-  xlab("Volume of Fascicle in Depression Network") + 
-  ylab("Effect of diagnosis (T)") + 
-  ggtitle(paste0("Fascicle/depression mask volume by dx T, no CC, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t_no_cc)$coefficients[2,4],3))) + 
-  theme(plot.title=element_text(hjust=0.5))
-print(scatter_t_val_by_dep_net_overlap_no_cc)
-
-```
-
-```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_vol_effect_size_no_CC}
-# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups
-
-##############################
-## FASCICLES TO BE INCLUDED ##
-##############################
-
-#define the depression network (all areas that intersect with depression network, no cranial nerves)
-fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
-
-#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
-fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
-fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
-
-#######################
-### depression map  ###
-#######################
-
-### define depression map
-fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
-
-#filter to exclude fascicles w/cranial nerves
-fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
-  replace(is.na(.), 0) %>%
-  filter(fascicle%in% fascicle_names_no_cranial_nerves)
-
-df_fascicle_and_vol_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))
-
-########################
-## Get Volume Values ###
-########################
-
-#read in fiber volume values
-fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
-
-#extract data frame that only includes the fascicle proportion lost
-df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
-df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]
-
-#multiply each row by the vector of total volumes
-df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
-
-#combine back with data_fram
-df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
-
-###############################
-#######################
-### real data x dx  ###
-#######################
-
-#then, make the data frame
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-# lapply to get results of t-test in each fascicle
-#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
-fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
-{
-  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
-  eval(parse(text=text_for_eval))
-})
-names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves
-
-#convert to data frame
-df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)
-
-#######################
-### plot differences ##
-#######################
-
-#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
-
-df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(df_fascicle_and_vol_in_mask$fascicle, df_fascicle_and_vol_in_mask$non_zero_voxels_in_dep_map, df_d_values)
-names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")
-
-df_fascicle_dep_mask_volume_overlap_d_value_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>%
-  filter(fascicle != "CC")
-
-#make copy to sort
-df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask <- df_fascicle_dep_mask_volume_overlap_d_value_no_cc
-
-df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$volume_in_dep_mask)])
-
-# bar graph of eff size of fascicles, sorted by depression overlap
-plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network, no CC") +
-  geom_bar(stat="identity") +
-  #ylim(-0.1,0.5) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(plot_fascicle_by_d_sorted_by_dep_overlap)
-
-# plot fascicle overlap x Eff size value
-#statistic, p=001
-dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc )
-print(summary(dep_overlap_x_dep_vs_healthy_d))
-
-scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc , aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
-  geom_point(size=1) + 
-  geom_smooth(method=lm)+
-  xlab("Volume of Fascicle in Depression Network") + 
-  ylab("Effect of diagnosis calculated in Volumes (D)") + 
-  ggtitle(paste0("Fascicle/depression mask Volume overlap by dx Eff Size, No CC p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
-  theme(plot.title=element_text(hjust=0.5))
-print(scatter_d_val_by_dep_net_overlap)
-
-```
-
-```{r bootstrapping_proportion_and_volume_x_dep_t_analyses}
-
-######### functions
-tval <- function(formula, data, indices) {
-  df <- data[indices,] # allows boot to select sample
-  fit <- lm(formula, data=df)
-  return(summary(fit)$coefficients[2,3])
-}
-
-##### proportion
-# 95% confidence intervals for the correlation
-
-set.seed(1)
-data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_t_value, select = c("depression_mask_overlap","dep_vs_healthy_t"))
-results_prop <- boot(data=data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap, statistic=tval,
-   R=10000, formula=dep_vs_healthy_t~depression_mask_overlap)
-print(results_prop) #actual 2.62, CI 0.701, 4.553)
-plot(results_prop)
-boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="perc", index=1)
-print(boot_results_prop.ci)
-pval_prop<-boot.pval(boot_res=results_prop, theta_null = 0, type = "perc", pval_precision = NULL)
-print(pval_prop) #p=0.0087
-
-##### effect size
-# 95% confidence intervals for the correlation
-
-set.seed(1)
-data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_d_value, select = c("depression_mask_overlap","dep_vs_healthy_d"))
-results_prop <- boot(data=data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap, statistic=tval,
-   R=10000, formula=dep_vs_healthy_d~depression_mask_overlap)
-print(results_prop) #actual 2.62, CI 0.589, 4.669)
-plot(results_prop)
-boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="perc", index=1)
-print(boot_results_prop.ci)
-pval_prop<-boot.pval(boot_res=results_prop, theta_null = 0, type = "perc", pval_precision = NULL)
-print(pval_prop) #p=0.0083
-
-##### volume
-# 95% confidence intervals for the correlation
-set.seed(1)
-data_frame_dep_v_nondep_t_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_t_value, select = c("dep_vs_healthy_t", "volume_in_dep_mask"))
-
-results_vol <- boot(data=data_frame_dep_v_nondep_t_x_vol_dep_mask, statistic=tval,
-   R=10000, formula=dep_vs_healthy_t~volume_in_dep_mask)
-print(results_vol)
-plot(results_vol)
-boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
-print(boot_results_vol.ci)
-pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
-print(pval_vol) #p=1*10^-16
-
-##### volume
-#effect sizes
-# 95% confidence intervals for the correlation
-set.seed(1)
-data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))
-
-results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
-   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
-print(results_vol)
-plot(results_vol)
-boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
-print(boot_results_vol.ci)
-pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
-print(pval_vol) #p=1*10^-16
-
-### with no corpus callosumm p = 1* 10^-16, CI(1.727, 5.363)
-# 95% confidence intervals for the correlation
-set.seed(1)
-results_vol <- boot(data=df_fascicle_dep_mask_volume_overlap_d_value_no_cc, statistic=tval,
-   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
-print(results_vol)
-plot(results_vol)
-boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
-print(boot_results_vol.ci)
-pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
-print(pval_vol) #p=1*10^-16
-
-```
-
-##### PROMIS DATA
-
-```{r compare_overall_promis_scores_by_diagnosis}
-
-#separate out only people in depression group, and take first instance
-dep_and_healthy_groups_for_ICD_analysis <-  df_people_with_promis_data_scores_added %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=82
-
-dep_vs_healthy_t <- lapply(promis_score_names, function(x) 
-{
-  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
-  eval(parse(text=t_test_command))
-})
-names(dep_vs_healthy_t) <- promis_score_names
-print(dep_vs_healthy_t)
-
-#pull t-statistic
-dep_vs_healthy_t_values <- sapply(dep_vs_healthy_t, function(x) x$statistic)
-
-#convert to data frame
-df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
-row.names(df_t_values) <- promis_score_names
-print(df_t_values)
-
-#pull p-values
-dep_vs_healthy_p_values <- sapply(dep_vs_healthy_t, function(x) x$p.value)
-
-#convert to data frame
-df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
-row.names(df_dep_vs_healthy_p_values) <- promis_score_names
-print(df_dep_vs_healthy_p_values)
-#FDR correct p-values
-dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")
-
-plots <- lapply(promis_score_names, function(x) 
-{
-
-  Diagnosis <- c("MS w/Depression", "MS No Depression")
-  dep_group <- eval(parse(text=paste0("dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]")))
-  nondep_group <- eval(parse(text=paste0("dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]")))
-  value <- c(mean(dep_group, na.rm=TRUE),
-           mean(nondep_group, na.rm=TRUE))
-  sem=c((sd(dep_group, na.rm = TRUE)/sqrt(sum(dep_group, na.rm = TRUE))), (sd(nondep_group, na.rm = TRUE)/sqrt(sum(nondep_group, na.rm = TRUE))))
-
-  data <- data.frame(Diagnosis,value,sem)
-
-  # Grouped
-  barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
-    geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
-      geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
-    scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
-    ggtitle(paste0(x, " by Diagnosis")) + #, p=", round(dep_vs_healthy_t_test$p.value,3))) +
-    ylab(paste0("Promis ", x)) +
-    theme(title = element_text(size=14),
-          axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-          axis.title.y = element_text(size=20),
-          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-          panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
-  #  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
-  print(barplot) #
-
-})
-
-```
-
-#### PROMIS Proportion analyses
-
-```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
- # filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_dep_and_nondep}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
-  filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_nondep}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
-  filter(depGroupVar == 1)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_dep}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
-  filter(depGroupVar == 2)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_irrespective_of_dx}
-
-#total n = 116, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=116
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_depressed_and_nondepressed}
-
-#total n = 82, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=82
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_nondepressed}
-
-#total n = 36, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
-  filter(depGroupVar == 1)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=36
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_depressed}
-
-#total n = 46, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
-  filter(depGroupVar == 2)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=46
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-#### PROMIS VOLUME ANALYSES
-
-```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
- # filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-```
-
-```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_dep_and_nondep}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
-  filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_nondep}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
-  filter(depGroupVar == 1)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-```
-
-```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_dep}
-
-#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
-
-#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
-  filter(depGroupVar == 2)  %>% 
-  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
-           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
-  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
-           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup()
-
-fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")
-
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_proportion_lost_categories
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_irrespective_of_dx_vol}
-
-#total n = 116, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=116
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_depressed_and_nondepressed_vol}
-
-#total n = 82, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
-  filter(depGroupVar != 0)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=82
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_nondepressed_vol}
-
-#total n = 36, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
-  filter(depGroupVar == 1)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=36
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```
-
-```{r dimensional_promis_effects_unique_depressed_vol}
-
-#total n = 46, nothing significant
-
-dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
-  filter(depGroupVar == 2)  %>% 
-  group_by(EMPI) %>% 
-  arrange(EXAM_DATE) %>% 
-  slice(1) %>% 
-  ungroup() #n=46
-
-#ones to try
-#PostOp.PROMIS.Mental.Score
-#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
-# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
-# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"
-
-#lm
-promis_lm <- lapply(promis_score_names, function(promis_score) {
-
-  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
-  {
-    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
-       data = dep_and_healthy_groups_for_ICD_analysis)
-  })
-  names(fascicle_lm) <- fascicle_names_no_cranial_nerves
-
-  #anova
-  fascicle_anova <- lapply(fascicle_lm, anova)
-
-  #fdr corrected
-  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
-  print(fascicle_anova_fdr)
-})
-
-```

### 2022-08-26 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
Add files via upload

@@
 library(lsr)
+library(sjPlot)
+#library(merTools)
 #functions
@@
 Current ns for people with at least 1 promis score= 
-Total n promis scores: 1694 
-Total n unique individuals w/promis scores: 501
+Total n promis scores: 1694-> 1813 
+Total n unique individuals w/promis scores: 501->532
 
 Different scores have different # of overall scans
-                        Total                      General_Health                  Quality_of_Life                  Physical_Health 
-                            1694                             1693                             1503                             1503 
-          Mental_Health_and_Mood   Social_Activities_Satisfaction   Carrying_Out_Social_Activities Carrying_Out_Physical_Activities 
-                            1503                             1503                             1501                             1499 
-              Emotional_Problems                  Fatique_Average     PostOp.PROMIS.Physical.Score       PostOp.PROMIS.Mental.Score 
-                            1499                             1499                             1440                             1442
+(x<-promis_data[6:15] ; x[!is.na(x)] = 1; x[is.na(x)] = 0, colSums(x))
+
+                  General_Health                  Quality_of_Life                  Physical_Health           Mental_Health_and_Mood 
+                            1812                             1567                             1567                             1567 
+  Social_Activities_Satisfaction   Carrying_Out_Social_Activities Carrying_Out_Physical_Activities               Emotional_Problems 
+                            1567                             1565                             1563                             1563 
+                 Fatique_Average     PostOp.PROMIS.Physical.Score 
+                            1563                             1504
 
 W/good mimosa:
-Total n promis scores : 398
-Total n unique individuals w/promis scores: 116
+Total n promis scores : 398 -> 412
+Total n unique individuals w/promis scores: 116 -> 121
 Total n unique indivs in nondepressed group = 36
-Total n unique individuals in depressed group = 46
+Total n unique individuals in depressed group = 46->49
 
@@
 Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-    1.0   376.8   763.0   873.3  1229.8  2487.0 
+  1.0   396.0   800.5   933.8  1325.5  2606.0 
     Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
-0.002747 1.035027 2.096154 2.399132 3.378434 6.832418
+0.002747 1.087912 2.199176 2.565346 3.641484 7.159341
 
@@
 Proportion within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
-    - all participants (116)
-    - Depressed + nondepressed (82)
+    - all participants (121)
+    - Depressed + nondepressed (85)
     - Non-depressed alone (36)
-    - Depressed alone (46)
+    - Depressed alone (49)
 
@@
   - all participants (116)
-  - Depressed + nondepressed (82)
+  - Depressed + nondepressed (85)
   - Non-depressed alone (36)
-  - Depressed alone (46)
+  - Depressed alone (49)
 
@@
 Volume within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
-   - all participants (116)
-    - Depressed + nondepressed (82)
+   - all participants (121)
+    - Depressed + nondepressed (85)
     - Non-depressed alone (36)
-    - Depressed alone (46)
+    - Depressed alone (49)
 
 Volume of individual fascicle affected related to promis scores  (777 analyses per group (11 promis domains x 77 individual fascicles))
-  - all participants (116)
-  - Depressed + nondepressed (82)
+  - all participants (121)
+  - Depressed + nondepressed (85)
   - Non-depressed alone (36)
-  - Depressed alone (46)
+  - Depressed alone (49)
 
 Promis results:
-Interestingly, across all 11 scores, with the exception of fatigue, Depressed patients have lower scores (lower = worse functioning), than non-depressed patients.
+Interestingly, across all 10 scores, with the exception of fatigue, Depressed patients have lower scores (lower = worse functioning), than non-depressed patients.
 controls.
@@
 
-Quality_of_Life	0.015199144			
-Physical_Health	0.012579264			
-Mental_Health_and_Mood	0.012579264			
-Social_Activities_Satisfaction	0.001063449			
-Carrying_Out_Social_Activities	0.011375799			
-Carrying_Out_Physical_Activities	0.031616473			
-Emotional_Problems	0.011270683			
-Fatique_Average	0.231497853			
-PostOp.PROMIS.Physical.Score	0.041172446			
-PostOp.PROMIS.Mental.Score	0.001253404
+Quality_of_Life	0.010			
+Physical_Health	0.019			
+Mental_Health_and_Mood	0.011		
+Social_Activities_Satisfaction	0.001		
+Carrying_Out_Social_Activities	0.007			
+Carrying_Out_Physical_Activities	0.032			
+Emotional_Problems	0.011			
+Fatique_Average	0.181			
+PostOp.PROMIS.Physical.Score	0.032			
+PostOp.PROMIS.Mental.Score	0.001
 
@@
 Proportion within versus outside depression map correlated w/promis scores - NO SIGNIFICANT FINDINGS
-    - all participants (116)
-    - Depressed + nondepressed (82)
+  - all participants (121)
+  - Depressed + nondepressed (85)
   - Non-depressed alone (36)
-    - Depressed alone (46)
+  - Depressed alone (49)
 
 Proportion of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
-  - all participants (116)
-  - Depressed + nondepressed (82)
+  - all participants (121)
+  - Depressed + nondepressed (85)
   - Non-depressed alone (36)
-  - Depressed alone (46)
+  - Depressed alone (49)
 
@@
 Volume within versus outside depression map correlated w/promis scores -  NO SIGNIFICANT FINDINGS
-   - all participants (116)
-    - Depressed + nondepressed (82)
+  - all participants (121)
+  - Depressed + nondepressed (85)
   - Non-depressed alone (36)
-    - Depressed alone (46)
+  - Depressed alone (49)
 
 Volume of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
-  - all participants (116)
-  - Depressed + nondepressed (82)
+  - all participants (121)
+  - Depressed + nondepressed (85)
   - Non-depressed alone (36)
-  - Depressed alone (46)
+  - Depressed alone (49)
 
@@
 ##########################################################
-promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_cleaned_n1694_unique_indiv_n501_2015-2022.rds"))
+#promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_cleaned_n1694_unique_indiv_n501_2015-2022.rds"))
+promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_20220824_fixed_physical_health_cleaned_n1813_unique_indiv_n532_2015-2022.rds"))
 promis_score_names <- names(promis_data[7:16])
@@
 df_people_with_promis_data <- df_demo_and_fascicles %>%
-  filter(EMPI %in% promis_data$EMPI) #n=398
-
-unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=116
-
-```
+  filter(EMPI %in% promis_data$EMPI) #n=416
+
+unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=121
+
+```
+
 ```{r make_df_with_volume_instead_of_proportion}
@@
 
+just_fascicle_and_proportion <- fascicle_values_dep_no_cranial_nerves %>% dplyr::select(fascicle, prop_in_mask)
+
+write.table(file = paste0(homedir, "/results/fascicle_proportion_overlap_with_depression_network_n77.csv"), x = just_fascicle_and_proportion, sep = ",", row.names = F, col.names = FALSE, quote=F)
+
 #figure out how many fascicles that is, then 
@@
 
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_values_dep_no_cranial_nerves %>%
+filter(fascicle%in% fascicle_names_no_cranial_nerves)
+
 #figure out how many fascicles that is, then 
@@
 dep_network_names_top_quartile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile] #n=19
+write.table(dep_network_names_top_quartile_no_cranial_nerves_by_vol, file = paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), quote = F, row.names = F, col.names = F)
 
 dep_network_total_volume_top_quartile_no_cranial_nerves <-
-sum(fascicle_volumes_dep_network$num_voxels_total[which(fascicle_volumes_dep_network$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)])
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)])
 
@@
 bottom_75_total_volume_no_cranial_nerves <-
-sum(fascicle_volumes_dep_network$num_voxels_total[which(fascicle_volumes_dep_network$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol)])
-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol)])
+
+#### Calculate total volume in all non-cranial_nerve_network
+total_volume_all_full_fascicles <- dep_network_total_volume_top_quartile_no_cranial_nerves + bottom_75_total_volume_no_cranial_nerves
 #plot values within the depression network, and values outside of it
@@
   mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(sum_fascicle_vol_lost = rowSums(across(fascicle_names_no_cranial_nerves))) %>%
+  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_volume_all_full_fascicles) %>%
+
   mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
            rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves_by_vol))) %>%
+  mutate(sum_fascicle_vol_lost_indepnet_top_quartile = 
+           rowSums(across(dep_network_names_top_quartile_no_cranial_nerves_by_vol))) %>%
   mutate(prop_fascicle_vol_lost_indepnet_top_quartile = (total_fascicle_vol_lost_indepnet_top_quartile / dep_network_total_volume_top_quartile_no_cranial_nerves)) %>%
+  mutate(proportion_volume_lost_dep_net_by_network_size_sum = sum_fascicle_vol_lost_indepnet_top_quartile /dep_network_total_volume_top_quartile_no_cranial_nerves) %>%
+
   mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
            rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol))) %>%
+  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowSums(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol))) %>%
+  mutate(prop_fascicle_vol_lost_bottom_3_quartile = (total_fascicle_vol_lost_nondep_bottom_3_quartiles / bottom_75_total_volume_no_cranial_nerves)) %>% 
+   mutate(proportion_volume_lost_nondep_net_by_network_size_sum = sum_fascicle_vol_lost_nondep_bottom_3_quartiles /bottom_75_total_volume_no_cranial_nerves) %>%
+
   mutate(total_fascicle_vol_lost_no_cc = rowMeans(across(fascicle_names_no_cranial_nerves[fascicle_names_no_cranial_nerves!="CC"]))) %>%
-  mutate(prop_fascicle_vol_lost_bottom_3_quartile = (total_fascicle_vol_lost_nondep_bottom_3_quartiles / bottom_75_total_volume_no_cranial_nerves)) %>%
   mutate(total_fascicle_vol_lost_indepnet_top_quartile_no_cc = 
@@
            rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol[dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol != "CC"]))) %>%
+
+  #remove people who didn't lose anything, i.e. bad data
+ # filter(total_fascicle_vol_lost != 0) %>%
   group_by(EMPI) %>% 
@@
 
+### use sums instead of means
+#wilcox test because non normally distributed; sums instead of means, scaled by network size p < 2.2*10^-16
+msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum,paired = TRUE)
+
+#wilcox test because non normally distributed; sums instead of means, scaled by network size, only in non_dep people p < 2.2*10^-16
+msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_nondep <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], paired = TRUE)
+
+#wilcox test because non normally distributed; sums instead of means, scaled by network size, only in dep people p < 2.2*10^-16
+msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_dep <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = TRUE)
+
 #test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
@@
 total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles_no_cc <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057
+
+#### interaction of diagnosis by network
+
+scale=1
+
+df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(diagnosis = as.factor(depGroupVar)) %>%
+  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
+  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
+  #mutate(Dep_network_affected = total_fascicle_vol_lost_indepnet_top_quartile) %>%
+  #mutate(Nondep_network_affected = total_fascicle_vol_lost_nondep_bottom_3_quartiles) %>%
+  dplyr::select(subject, diagnosis, Dep_network, Nondep_network) %>% 
+  gather(network, proportion_affected, -subject, -diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+ # mutate(network_size = ifelse(network == "Dep_network", dep_network_total_volume_top_quartile_no_cranial_nerves/scale, bottom_75_total_volume_no_cranial_nerves/scale)) %>%
+#  mutate(total_affected_volume_in_net = (proportion_affected * network_size)) %>% 
+  ungroup()
+
+#df_empi_dx_network_proportions <- as.data.frame(df_empi_dx_network_proportions)
+
+diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ network*diagnosis + (1|subject), data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer)
+
+diagnosis_by_network_interaction_lmer_main_effects <- lmer(proportion_affected ~ network+diagnosis + (1|subject), data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_main_effects)
+
+diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, type = "int")
+print(diagnosis_by_network_interaction_lmer_plot)
+
+#diagnosis_by_network_interaction_glmer <- glmer(cbind(total_affected_volume_in_net, network_size-total_affected_volume_in_net) ~ network*diagnosis + (1|subject), data = df_empi_dx_network_proportions, family = binomial(link="logit"))
+#summary(diagnosis_by_network_interaction_glmer )
+
+#diagnosis_by_network_interaction_glmer_plot  <- plot_model(diagnosis_by_network_interaction_glmer , type = "int")
+#print(diagnosis_by_network_interaction_glmer_plot)
+
+#pred_df <- df_empi_dx_network_proportions %>% 
+ #          dplyr::select(network,diagnosis) %>% 
+  #         expand.grid() %>% distinct() %>% mutate(subject = rep(df_empi_dx_network_proportions$subject[1], 4))
+
+#predicted_values_means_glmer <- predict(object=diagnosis_by_network_interaction_glmer,newdata=pred_df,se.fit=TRUE, re.form=NA, type = "response")
+
+#predicted_values_means_lmer <- predict(object=diagnosis_by_network_interaction_lmer,newdata=pred_df,se.fit=TRUE, re.form=NA, type = "response")
+
+#pred_lmer <- merTools::predictInterval(diagnosis_by_network_interaction_lmer, newdata =pred_df, n.sims = 999, which = "fixed")
+
+#pred_glmer <- merTools::predictInterval(diagnosis_by_network_interaction_glmer, newdata = pred_df, n.sims = 999, which="fixed")
+
+#pred_df_lmer_w_ci <- cbind(pred_df, pred_lmer)
+#pred_df_glmer_w_ci <- cbind(pred_df, pred_glmer)
+
+#plot_pred_df_lmer_plot <- ggplot(pred_df_lmer_w_ci,aes(x=network,y=fit,color=diagnosis, group=diagnosis, ymax=upr,ymin=lwr)) +
+ #  geom_point(position = position_dodge(width=0.5),size=1) + geom_errorbar(size=1,position = position_dodge(width=0.5),width=.5)+geom_line(position = position_dodge(width=0.5),size=1) 
+#  ylim(0.2,0.5)
+
+#plot_pred_df_glmer_plot <- ggplot(pred_df_glmer_w_ci,aes(x=network,y=fit,color=diagnosis, group=diagnosis, ymax=upr,ymin=lwr)) +
+   #geom_point() + geom_errorbar()+geom_line()
+
+#geom_point(aes(color=diagnosis)) + geom_line(aes(color=diagnosis)) + geom_errorbar()
+
+#ggplot(pred_df,aes(x=network,y=mean,color=diagnosis,ymax=upper_ci,ymin=lower_ci)) + geom_point() + geom_errorbar()+geom_line()
+#predicted_values_ces <- predict(object=diagnosis_by_network_interaction_glmer,newdata=pred_df,se.fit=TRUE, re.form=NA)
+
+#binomial, model needs to account for the fact that you can't have values outside of the 0-1 range. Change families because outcome variable not from a normal distribution.
 
@@
 # geom violin within vs outside depression network
-dep_net_group <- c(rep("Dep_net", times = 380), rep("Non_dep_net",times = 380))
+dep_net_group <- c(rep("Dep_net", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Non_dep_net",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
 net_values <- c(df_mean_fascicle_volumes_and_diagnosis$total_fascicle_vol_lost_indepnet_top_quartile, df_mean_fascicle_volumes_and_diagnosis$total_fascicle_vol_lost_nondep_bottom_3_quartiles)
@@
 # geom violin within vs outside depression network, no cc
-dep_net_group <- c(rep("Dep_net", times = 380), rep("Non_dep_net",times = 380))
+dep_net_group <- c(rep("Dep_net", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Non_dep_net",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
 net_values <- c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile_no_cc, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles_no_cc)
@@
 
-dep_net_group <- c(rep("Dep_net", times = 380), rep("Non_dep_net",times = 380))
+dep_net_group <- c(rep("Dep_net", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Non_dep_net",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
 net_values <- c(dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_indepnet_top_quartile, dep_and_healthy_groups_for_ICD_analysis$prop_fascicle_vol_lost_bottom_3_quartile)
@@
   ggtitle(paste0("Proportion Fascicle Volume Affected by Network")) +
-  ylab("Proportion (tot volume lost/Net Vol) ") +
+  ylab("Proportion (mean tot volume lost/Net Vol) ") +
   ylim(0,0.1) + 
@@
 print(violin_within_vs_outside_network_scaled_by_network_size) #
+
+#sums instead of means, scaling by size of network
+dep_net_group <- c(rep("Dep_net", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Non_dep_net",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
+net_values <- c(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum)
+
+newdf_no_cc <- data.frame(dep_net_group, net_values)
+
+#p = 2.2 x 10^-16. 
+violin_within_vs_outside_network_sum_proportion_scaled_by_network_size <- ggplot(newdf_no_cc, aes(x=dep_net_group, y=net_values, fill=dep_net_group)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Dep Net", "2" = "Non-Dep Net")) + 
+  ggtitle(paste0("Proportion Fascicle Volume Affected by Network")) +
+  ylab("Proportion (sum total volume lost/Net Vol) ") +
+ # ylim(0,0.1) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_within_vs_outside_network_sum_proportion_scaled_by_network_size) #
 
@@
 
-#define the depression network (all areas that intersect with depression network, no cranial nerves)
+#dfectine the depression network (all areas that intersect with depression network, no cranial nerves)
 fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
@@
   filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
+#  filter(total_fascicle_vol_lost != 0) %>%
   group_by(EMPI) %>% 

### 2022-08-26 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
 Corpus_Callosum_Forceps_Major
+
+#### 20220819
+Try the within/between without cc
+make talk pretty
+
+#### 20220823
+[x] sarah weinstein about network enrichment on slack
+[x] scale by number of fibers in the depression network vs outside, rather than by fascicle. - works p < 2.2 *10^-16
+
+#### 20220824
+[x] keep talking with sarah
+  - Sarah and Taki like option three (sum all disease within each network, divide by size of network)
+[x] double check that wilcoxon is the right test to do given non normality. 
+[x] there are 2 people who have no disease at all on their scans. WIll check them, but it is possible that I should remove them, maybe the mimosa step broke.
+  - if i exclude those two people, findings are interesting - total overall volume lost not different between groups (p=0.06). Depression (p=0.046), nondepressed people = 0.096)
+  - I did check, they do have mimosas, but the mimosas seem to be in perhaps cranial nerve areas. I am going to keep as is for now.
+[] make clean notebook. Current one is excessive.
+
+#### 20220825
+[x] figure out how to do interaction
+       - need a data frame with two colums, diagnosis (1s and 2s), and dep network (1s and 2s). What is the measure? 
+       - Diagnosis    Dep Network   Proportion     Person
+           1            1               .39           1
+           1            2               .61           1
+           2            1               .22           2
+           2            2               .34           2
+
+           model:
+           proportion disease ~ diagnosis*dep_net + (1 | person)
+[x] bart decided he wanted to do binomial instead of linear, realize that this is quite unstable given how big the numbers are
+- tried with a scale factor
+- tried 100000 and 10000, they give VERY big differences (i.e. dividing the number of total voxels in the network divided by the scale)
+- might be related to the false consideration that each voxel is independent, and the size of the data
+- in the end, will probably just do the linear model
+- everyone got greedy
+
+GOING BACK TO JUST LMER
+
+#### 20220826
+[] colors 
+  - redownloaded dsistudio from frank. Currently, fascicles need to have colors manually added
+  - script to make individual files for colors (rscript make_red_to_yellow_RGB_color_scheme.R)
+  - Also, if you give load a table with multiple lines, it will assign it to multiple fascicles in order! So I am just going to also just output the RGB values into a spreadsheet for reading in. 
+    [x] made table wmake_red_to_yello_RBG_color_scheme
+    [x] in dsi studio, added each tract manually
+      [x] double checked TWICE that fascicle order from fascicle_mean_proportion_overlap_nfasc_77_fascicle_names.txt is the same order that the fascicles are loaded in /Users/eballer/BBL/msdepression/templates/dti
+      [x] saved it All_77_tracts_no_cranial_nerves.gz.txt
+      [x] loaded colors (tracts misc/open cluster colors), navigate to /Users/eballer/BBL/msdepression/templates/dti/colors, and selected: fascicle_mean_proportion_overlap_nfasc_77_rgb.txt
+      [x] looked at the pretty pictures!, red is lower proportion, yellow is higher proportion
+      [x] wow, so pretty!
+Challenge is that it assumes there are equal values from 0-1, so if the max is low, it scales incorrectly.
+Will try to set max to be the max proportion, and scale based on that.
+      [x] now you can scale!
+*********
+  - To open the collection of fascicles: in dsistudio, in t3D tracts (right next to autotracking), click the button. 
+  - open /Users/eballer/BBL/msdepression/templates/dti/colors/All_77_tracts_colored_by_overlap_w_dep_map.tt.gz
+  - it gives you each cluster in order, labeled 0 through 76. These were cross referenced with size of the named fascicles to confirm THEY ARE IN ORDER OF RGB MAPS
+
+  **********
+  - Another way to do this
+      - Open dsistudio
+      - click on fib file to open: /Users/eballer/BBL/msdepression/templates/dti/1904002815.src.gz.odf8.f5.gqi.1.25.fib.gz
+      - once in studio, in t3d on right hand corner, click the folder
+      - open /Users/eballer/BBL/msdepression/templates/dti/colors/All_77_tracts_colored_by_overlap_w_dep_map.tt.gz (should load in the order I want)
+      - click each of the boxes
+      - confirm order is what I want
+      - load colors : /Users/eballer/BBL/msdepression/templates/dti/colors, and selected: fascicle_mean_proportion_overlap_nfasc_77_rgb.txt (or whatever color I want)
+      - available color maps at this time (constructed in make_red_to_yellow_RGB_color_scheme):
+      "fascicle_mean_proportion_overlap_nfasc_77", "fascicle_proportion_overlap_with_depression_network_n77"
+  *********
+
+***********
+   - Now just need to find a way to make a color scale like in afni, a bar of sorts
+[x] made separate spreadsheet for fascicles and their overlap with the depression network so I could get colors for them
+``` {r writing}
+#line 660 through 664 msdepression_n380_started_20220726
+
+just_fascicle_and_proportion <- fascicle_values_dep_no_cranial_nerves %>% dplyr::select(fascicle, prop_in_mask)
+
+write.table(file = paste0(homedir, "/results/fascicle_proportion_overlap_with_depression_network_n77.csv"), x = just_fascicle_and_proportion, sep = ",", row.names = F, col.names = FALSE, quote=F)
+
+```
+
+[x] promis - updated, everything looks good, fatigue not different, all other findings going in expected direction, physical and mental health are now different
+
+``` {r make_red_to_yellow_RGB_color_scheme.R}
+#### Welcome to make_red_to_yellow_RGB_color_scheme! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle to be its proportion of loss, scaled on the RGB scale
+### In order to do this, I needed to figure out how to map my proportions to RGB
+
+### Pre: File of proportions
+### Post: Output of RGB values in Red to Yellow Scale for the Proportions
+### Uses: Step 0: Figure out how to scale from red to yellow 
+###           - Red = 255 0 0, yellow 255 255 0
+###           - Means that we need to vary the middle value (255) to get where we need to be
+###       Step 1) For red to yellow, calculate the step (i.e. how much each value will be)
+###             - because my proportions are out of 1, I wanted each step to be 1/100
+###             - I am setting this value as "max", so it could vary if you have another max #
+###       Step 2) Multiply the proportion by 100, so now each count is 1-100, in percents
+###       Step 3) Multiply the percent by the step
+###       Step 4) Save this value in that fascicle, 255 [new_value] 0
+###
+### Dependencies: Any R will do
+
+###
+homedir="/Users/eballer/BBL/msdepression/"
+
+files = c("fascicle_mean_proportion_overlap_nfasc_77", "fascicle_proportion_overlap_with_depression_network_n77")
+
+for (filename in files) {
+  #suffix
+  suffix = paste0("_", filename, "_RGB")
+
+  #read in file with fascicle name and proportion
+  fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
+  names(fascicle_name_and_prop) <- c("fascicle", "proportion")
+
+  #set scale dynamically
+  #highest possible number in scale
+  max<-100 * max(fascicle_name_and_prop$proportion)
+
+  #step that each value in our scale will correspond to
+  step<-255/max
+
+  # make a column for red
+  red <- rep(x=255, times = 77)
+
+  fascicle_name_and_prop$red <- red
+  #multiply proportion column by 100, then by step, and round
+  fascicle_name_and_prop$green<- round((fascicle_name_and_prop$proportion *100*step),0)
+
+  #make column for blue
+  blue <- rep(x=0, times = 77)
+  fascicle_name_and_prop$blue=blue
+
+  #sort based on proportion to check, it works! Lower values are associated with lower RGB
+  #sorted <- fascicle_name_and_prop[order(fascicle_name_and_prop$proportion),]
+
+  #make individual file colors for loading
+  for (fascicle in 1:dim(fascicle_name_and_prop)[1]) {
+
+    fascicle_name = fascicle_name_and_prop$fascicle[fascicle]
+    fascicle_rgb = paste0(fascicle_name_and_prop$red[fascicle], " ", fascicle_name_and_prop$green[fascicle], " ", fascicle_name_and_prop$blue[fascicle])
+    print(paste0(fascicle_name, " ", fascicle_rgb))
+    write.table(file = paste0(homedir, "templates/dti/colors/", fascicle_name, suffix,".txt"), x = fascicle_rgb, quote = FALSE, col.names = FALSE, row.names = FALSE)
+
+  }
+
+  #output color file (77 lines, each represents a color), and color name vector. Not sure how to do 1:1 assignments rather than manually, but I'll see if I can figure that out!
+  fascicle_name_and_prop_rgb_only <- fascicle_name_and_prop %>%
+    mutate(rgb = paste0(fascicle_name_and_prop$red, " ", fascicle_name_and_prop$green, " ", fascicle_name_and_prop$blue)) %>%
+    dplyr::select(fascicle, rgb)
+
+  write.table(file = paste0(homedir, "/templates/dti/colors/", filename, "_fascicle_names.txt"), x = fascicle_name_and_prop_rgb_only$fascicle, row.names = FALSE, col.names = FALSE, quote = FALSE)
+
+  write.table(file = paste0(homedir, "/templates/dti/colors/", filename, "_rgb.txt"), x = fascicle_name_and_prop_rgb_only$rgb, row.names = FALSE, col.names = FALSE, quote = FALSE)
+}
+```
+
+#### 20220830
+
+[] play with visualizations. Could probably have a map that has blue for non dep, and yellow for dep

### 2022-08-31 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
Add files via upload

@@
 
+#sensitivity analysis
+tertile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 3),0)
+quintile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 5),0)
+
+dep_network_names_top_tertile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile] #n=26
+
+dep_network_names_top_quintile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile] #n=15
+
+dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)]
+
+dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)]
+
+top_tertile_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)])
+
+top_quintile_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)])
+
+bottom_tertiles_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol)])
+
+bottom_quintiles_total_volume_no_cranial_nerves<-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol)])
+
 ```
@@
 
+  # tertile and quintiles
+  mutate(sum_fascicle_vol_lost_top_tertile = 
+           rowSums(across(dep_network_names_top_tertile_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_top_tertile_by_network_size_sum = sum_fascicle_vol_lost_top_tertile /top_tertile_total_volume_no_cranial_nerves) %>%
+    mutate(sum_fascicle_vol_lost_bottom_tertiles = 
+           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%
+    mutate(sum_fascicle_vol_lost_top_quintile = 
+           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%
+    mutate(sum_fascicle_vol_lost_bottom_quintiles = 
+           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves) %>%
+
   #remove people who didn't lose anything, i.e. bad data
@@
 
+#### Save a data frame that contains the mean volume of disease within network, scaled by the size of the network, for visualization
+
+diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
+network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
+mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))
+
+df_means_volume_scaled_by_network_size <-data.frame(diagnosis_vector, network_vector, mean_diagnosis_by_network_vector)
+names(df_means_volume_scaled_by_network_size) <- c("diagnosis", "network", "mean")
+write.table(df_means_volume_scaled_by_network_size, paste0(homedir, "/results/df_means_of_volumes_scaled_by_network_size_dx_net_for_color.csv"), row.names=F, col.names = T, quote = F, sep=",")
+
 #test whether ms preferentially hits within or outside depression network, p = 2.2 * 10^-16, mean lost in dep net = 18027, mean lost outside dep net = 4020.
@@
 ### use sums instead of means
-#wilcox test because non normally distributed; sums instead of means, scaled by network size p < 2.2*10^-16
+#wilcox test because non normally distributed; sums instead of means, scaled by network size p < 2.2*10^-16, dep net prop = 0.400, non dep = 0.339
 msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum,paired = TRUE)
@@
 
+#### tertile/quintile
+
+#p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
+msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)
+
+#p=0.0328, dep = 0.4407, nondep = 0.387
+msdisease_diagnostic_effects_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.058, dep = 0.31, nondep= 0.28
+msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=2.2*10^-16, top quintile prop vol affected = 0.419, bottom 4 quintiles = 0.322
+msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)
+
+#p=0.0325, dep = 0.4411, nondep = 0.386
+msdisease_diagnostic_effects_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.048, dep = 0.34, nondep = 0.30
+msdisease_diagnostic_effects_wilcoxon_bottom_quintiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
 #test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
@@
   mutate(subject = as.factor(EMPI)) %>%
-  mutate(diagnosis = as.factor(depGroupVar)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
   mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
@@
   #mutate(Nondep_network_affected = total_fascicle_vol_lost_nondep_bottom_3_quartiles) %>%
-  dplyr::select(subject, diagnosis, Dep_network, Nondep_network) %>% 
-  gather(network, proportion_affected, -subject, -diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  # mutate(network_size = ifelse(network == "Dep_network", dep_network_total_volume_top_quartile_no_cranial_nerves/scale, bottom_75_total_volume_no_cranial_nerves/scale)) %>%
 #  mutate(total_affected_volume_in_net = (proportion_affected * network_size)) %>% 
+  mutate(Network = as.factor(Network)) %>%
   ungroup()
@@
 
-diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ network*diagnosis + (1|subject), data = df_empi_dx_network_proportions)
+diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
 summary(diagnosis_by_network_interaction_lmer)
 
-diagnosis_by_network_interaction_lmer_main_effects <- lmer(proportion_affected ~ network+diagnosis + (1|subject), data = df_empi_dx_network_proportions)
+diagnosis_by_network_interaction_lmer_main_effects <- lmer(proportion_affected ~ Network+Diagnosis + (1|subject), data = df_empi_dx_network_proportions)
 summary(diagnosis_by_network_interaction_lmer_main_effects)
 
-diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, type = "int")
+diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, 
+                                                         type = "int",
+                                                         axis.lim = c(0.2,0.6))
+
+diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, 
+                                                         type = "int",
+                                                         axis.lim = c(0.2,0.5),
+                                                         colors = c("#ff00ff", "#3df2f2"),
+                                                         line.size = 1.3,
+                                                         dodge=0.2) + 
+  theme(axis.text.x = element_text(size=12),
+        axis.text.y = element_text(size=12),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        aspect.ratio = 5/3,
+        panel.background = element_rect(colour = "white", fill = "white"))
+
 print(diagnosis_by_network_interaction_lmer_plot)
@@
 print(scatter_d_val_by_dep_net_overlap)
+df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_d_value %>% dplyr::select(fascicle, dep_vs_healthy_d)
+write.table(x = df_for_making_colors, file = paste0(homedir, "results/fascicle_and_effect_size_for_volxdx_analysis_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")
 

### 2022-08-31 msdepression/scripts/make_red_to_yellow_RGB_color_scheme.R
Add files via upload

@@ -19,13 +19,18 @@
 ### Dependencies: Any R will do
 
 ###
+
+### gradients
+## red-yellow - R always 255, green scaled (0 is red, 255 is yellow), blue always 0
+## pink gradient (pg) - Red always 255, green scaled 0 to 255 , blue always 255
+
 homedir="/Users/eballer/BBL/msdepression/"
 
-files = c("fascicle_mean_proportion_overlap_nfasc_77", "fascicle_proportion_overlap_with_depression_network_n77")
+files = c("fascicle_mean_proportion_overlap_nfasc_77", "fascicle_proportion_overlap_with_depression_network_n77", "fascicle_and_effect_size_for_volxdx_analysis_n77")
 
 for (filename in files) {
   #suffix
-  suffix = paste0("_", filename, "_RGB")
+  suffix = "_pg" #pink gradient
   
   #read in file with fascicle name and proportion
   fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
@@ -39,29 +44,34 @@ for (filename in files) {
   step<-255/max 
   
   # make a column for red
+  
   red <- rep(x=255, times = 77)
   
   fascicle_name_and_prop$red <- red
   #multiply proportion column by 100, then by step, and round
-  fascicle_name_and_prop$green<- round((fascicle_name_and_prop$proportion *100*step),0)
+ # fascicle_name_and_prop$green<- round((fascicle_name_and_prop$proportion *100*step),0)
+  fascicle_name_and_prop$green<- 255-round((fascicle_name_and_prop$proportion *100*step),0) #for pg, I want darker colors to be more sig
+  
   
   #make column for blue
-  blue <- rep(x=0, times = 77)
+ # blue <- rep(x=0, times = 77)
+  blue <- rep(x=255, times = 77)
   fascicle_name_and_prop$blue=blue
   
   #sort based on proportion to check, it works! Lower values are associated with lower RGB
   #sorted <- fascicle_name_and_prop[order(fascicle_name_and_prop$proportion),]
   
   #make individual file colors for loading
-  for (fascicle in 1:dim(fascicle_name_and_prop)[1]) {
+  #for (fascicle in 1:dim(fascicle_name_and_prop)[1]) {
     
-    fascicle_name = fascicle_name_and_prop$fascicle[fascicle]
[…]

### 2022-08-31 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
       - confirm order is what I want
-      - load colors : /Users/eballer/BBL/msdepression/templates/dti/colors, and selected: fascicle_mean_proportion_overlap_nfasc_77_rgb.txt (or whatever color I want)
+      - load colors : tracts misc -> open cluster colors -> 
+                -/Users/eballer/BBL/msdepression/templates/dti/colors, and selected: fascicle_mean_proportion_overlap_nfasc_77_rgb.txt (or whatever color I want)
                 - available color maps at this time (constructed in make_red_to_yellow_RGB_color_scheme):
@@
 
-[] play with visualizations. Could probably have a map that has blue for non dep, and yellow for dep
+[]
+
+[x] Talk about figures with Ted
+  [] 1. Flow chart for patient selection (fig 1)
+  [] 2. Demographics (or just list in text, or make A with flow chart and B with demographics) (table 1)
+  [] 3. Pipeline figure
+  [] 4. Two part image of depression network
+    [] a: Rank of fascicles by volume of overlap w/dep network (cortical rendering of depression network)
+    [] b: DSI fascicle of volume of overlap w/network (how to scale for visualizations) (new hamphire forest in the fall colors)
+  [] 5. In vs outside dep net
+    a. pink and blue fascicle pcicture part B
+    b. inside vs outside dep net violin
+  [] 6. Diagnosis analysis
+    a. no diff by lesion volum
+    b. but difference in fascicle volume
+  [] 7. Dep Network Interaction bar graph
+    [] is it possible to have a 3 part figure? 
+      - a. interaction graph
+      - b.
+picture of A.
+In depressed patients, volume of disease in dep network in depressed patients in one color, volume of disease in non depressed network in another color, divided by overall network size; B.
+Same thing in nondepressed patients.
+All on same color scale so you can hopefully see brighter colors/contrast in depressed as compared to non-depressed
+  [] 8. Depression network overlap 
+    a. scatter
+    b. Fascicles colored by effect size
+
+is there a way to make a graphic of the histogram? Like median prop overlap rather than mean? I think this would be nice
+
+[] sensitivity analysis using a different cutoff (do top third or quintile).
+ - made the quintile and tertile groups
+ - double checked
+   - names (15 in top quintile, 62 in bottom 4; 26 in top tertile, 51 in bottom tertile)
+   - volume sums (top tert + bottom 2 tertiles = top quintile + bottom 4 quintiles)
+
+  - Added columns in that big mutate command
+  - checked analysis: wilcox disease within depression network versus outside, calcuated by taking a sum of the affected volume in the depression network for each person, and dividing by the size overall size of the depression network, repeating for non-depression network, and using those proportions as inputs.
+
+  - Great news! Works with tertiles and quintiles, more disease in depression network. 
+  - With depression diagnosis, more disease consistently in depressed vs nondepressed
+      - with tertiles, p < 0.05 in dep network, but not outside (trend)
+      - with quintiles, p < 0.05 in both dep and non-dep network
+  - haven't tested interaction, but seems fair to say that different ways of slicing the cake, depressed people always have more disease in depression network, sometimes also in nondep network, and that more disease in depressed vs nondepressed network
+```{r sensitivity analyses}
+ #### tertile/quintile
+
+#p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
+msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)
+
+#p=0.0328, dep = 0.4407, nondep = 0.387
+msdisease_diagnostic_effects_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.058, dep = 0.31, nondep= 0.28
+msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=2.2*10^-16, top quintile prop vol affected = 0.419, bottom 4 quintiles = 0.322
+msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)
+
+#p=0.0325, dep = 0.4411, nondep = 0.386
+msdisease_diagnostic_effects_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.048, dep = 0.34, nondep = 0.30
+msdisease_diagnostic_effects_wilcoxon_bottom_quintiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+```
+
+#### 20220831
+[x] ping ted with updates re: sensitivity analysis and confirm that analyses are done
+
+[x] make effect size graphics
+[o] make graphics separating out depressed patients and non depressed patients 
+   - the scaling I am using right now just isn't wide enough to give me the contrast I'd like.
+
+[] make interaction graph prettier. Can actually hard code some things
+> get_model_data(diagnosis_by_network_interaction_lmer, type="int")
+# Predicted values of proportion_affected
+
+# Network = Dep_network
+
+Diagnosis | Predicted |   group_col |       95% CI
+--------------------------------------------------
+        1 |      0.37 | Dep_network | [0.33, 0.40]
+        2 |      0.42 | Dep_network | [0.39, 0.45]
+
+# Network = Nondep_network
+
+Diagnosis | Predicted |      group_col |       95% CI
+-----------------------------------------------------
+        1 |      0.32 | Nondep_network | [0.28, 0.35]
+        2 |      0.35 | Nondep_network | [0.33, 0.38]
+
+Adjusted for:
+* subject = 0 (population-level)
+
+#### 20220901
+[] separate to make a clean document once confirm analyses w/Ted

### 2022-09-02 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_just_volumes_started_20220824.Rmd
massive cleanup - just need to fix sensis

@@
 library(sjPlot)
+library(coin)
+library(rstatix)
+
 #functions
 source("~/BBL/msdepression/scripts/msdepression_functions.R")
+#source("~/BBL/msdepression/scripts/wilcox_effsize.R")
 ```
@@
 
-#write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)
-
 #add suffix to each measure
@@
 fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
-
-#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
-#fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
- # replace(is.na(.), 0) %>%
-#  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))
-
-#Write out the new network groupings to a file, will be used later
-#fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
-#write.table(fascicle_volumes_dep_network, file=fascicle_names_dep_network_grouping, sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)
-
-#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether any overlap with depression network (anything above 1 is overlap), as well as whether the fascicle overlaps, and the proportion of overlap with yeo networks
-#fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)
 
@@
 
+#remove cranial nerves
+fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
 #extract data frame that only includes the fascicle proportion lost
@@
 df_fascicles_volumes <- sweep(x = df_fascicles_prop, MARGIN = 2,                        # Sweep with Complex STATS
-                  STATS = fiber_volume_values$volume_full, FUN = "*")
+                  STATS = fiber_volume_values_no_cranial_nerves$volume_full, FUN = "*") %>%
+  mutate_if(is.numeric, round) #round all values
 
@@
 ```
-PROMIS DATA HANDLING to link a person to their most proximal promis
+PROMIS DATA HANDLING to link a person to their most proximal PROMIS
 
@@
 ##########################################################
-promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_cleaned_n1694_unique_indiv_n501_2015-2022.rds"))
+promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_20220824_fixed_physical_health_cleaned_n1813_unique_indiv_n532_2015-2022.rds"))
 promis_score_names <- names(promis_data[7:16])
@@
 
-#function, take an empi and date, returns a column of promis scores
+#function, take an empi and date, returns the row of data w/promis scores
 get_closest_promis_to_ms_exam_date <- function(subj_empi, subj_exam_date) {
@@
   #pull out subject empi and exam date
-  print(line)
   subj_empi <- df_people_with_promis_data$EMPI[line]
   subj_exam_date <- df_people_with_promis_data$BEGIN_EXAM_DTTM[line]
+
+  #returns whole line, promis scores, id, empi, everything
   closest_promis_data <- get_closest_promis_to_ms_exam_date(subj_empi = subj_empi, subj_exam_date = subj_exam_date)
+
+  #store as new line
   promis_info_closest_proximity[line,] = closest_promis_data
@@
 summary(df_people_with_promis_data_scores_added$distance_from_scan/364) #mean 2.40 ys, median 2.10 years, range 0.003 y to 6.83 years
+
 hist(df_people_with_promis_data_scores_added$distance_from_scan/364, 
@@
 ```
+
+Define Depression Network by Volume of Overlap
+```{r visualize_depression_network_volume}
+###################################################
+### Defining and Visualizing Depression Network ###
+###################################################
+
+#This chunk reads in a file where columns are fascicle, num_voxels_total, non_zero_voxels_in_dep_map, prop_in_mask
+# by the end, the depression mask will be defined
+# We consider fascicles in the depression network to be the ones that share the most VOLUME with the depression mask
+# Of course, it is possible that huge fascicles contribute a ton to the depression network, but also the non-depression network
+# And to get at specificity (i.e.
+which fibers most exclusively send information within the depression network), we could look at proportion of the fiber (volume that intersects / total volume of the fascicle).
+But then you have a situation where tiny fibers get lots of credit.
+We have actually tried this as well and the findings aren't that different.
+#But for the sake of this paper, will do volume because the question is all volume based (volume of lesions to volume of fascicles)
+
+######## Histograms for depression network
+#first set the values to be the volumes
+fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_dep %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+#set NAs to zero
+fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map[is.na(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)] = 0
+
+#sort based on overlap with depression depression map
+fascicle_volumes_dep_network_no_cranial_nerves$fascicle <- factor(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, levels = fascicle_volumes_dep_network_no_cranial_nerves$fascicle[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)])
+
+#figure out how many fascicles are in top quartile
+quartile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 4),0)
+
+#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take names of the top quartile
+dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile]
+
+#get names of bottom 3 quartiles
+
+#get a network that includes all of the fascicles NOT in dep 25% network
+dep_network_names_bottom_3_quartiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)]
+
+#### Get sums of the volumes of the fascicles in each network, and total sum
+
+#top quartile, depression network
+dep_network_total_volume_top_quartile_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)])
+
+#bottom 3 quartiles, nondepression network
+bottom_75_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves)])
+
+#### Calculate total volume in all non-cranial_nerve_network
+total_volume_all_full_fascicles <- dep_network_total_volume_top_quartile_no_cranial_nerves + bottom_75_total_volume_no_cranial_nerves
+
+###
+
+plot_all_fascicle_volumes_dep_network_no_cranial_nerves <-ggplot(data = fascicle_volumes_dep_network_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
+  ggtitle("Volume Overlap per Fascicle 3_09") + 
+  geom_bar(stat="identity", fill="006300") +
+  xlab("Fascicle") +
+  ylab("Overlap with Depression Mask") +
+  ylim(0,40000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(plot_all_fascicle_volumes_dep_network_no_cranial_nerves)
+
+### Change names to add suffix
+#replace each fascicle with vol suffix
+dep_network_names_top_quartile_no_cranial_nerves <- dep_network_names_top_quartile_no_cranial_nerves %>% 
+  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
+
+dep_network_names_bottom_3_quartiles_no_cranial_nerves <- dep_network_names_bottom_3_quartiles_no_cranial_nerves %>%
+   gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
+```
+
 Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups
@@
 
-dep_vs_healthy_t_test_lesion_volume <-  t.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #p=0.07
+# Assessing Normality w/histograms and Shapiro_Wilk test for normality
+hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])
+non_dep_lesion_vol_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]) #W=0.77, p = 8.2 * 10^-14, NON-NORMAL
+
+hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+dep_lesion_vol_shapiro <- hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.80, p = 2.2 * 10^-16, NON-NORMAL
+
+dep_vs_healthy_wilcoxon_lesion_volume_results <-  wilcox.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = F) #p=0.036; depressed = 15,871; nondepressed = 12,886
+
+dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
 
@@
 ```
-Define Depression Network by Volume of Overlap
-```{r visualize_depression_network_volume}
-##########################################################
-#                 Histos                        #
-##########################################################
 
-######## Histograms for depression network
-#first set the values to be the volumes
-fascicle_volumes_dep_network_sorted_by_overlap <- fascicle_volumes_dep
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected}
+## compare disease within and outside depression network, between diagnoses, and interactions
 
-#set NAs to zero
-fascicle_volumes_dep_network_sorted_by_overlap$non_zero_voxels_in_dep_map[is.na(fascicle_volumes_dep_network_sorted_by_overlap$non_zero_voxels_in_dep_map)] = 0
+###############################
+### Constructing Data Frame ###
+###############################
 
-#sort based on overlap with depression depression map
-fascicle_volumes_dep_network_sorted_by_overlap$fascicle <- factor(fascicle_volumes_dep_network_sorted_by_overlap$fascicle, levels = fascicle_volumes_dep_network_sorted_by_overlap$fascicle[order(fascicle_volumes_dep_network_sorted_by_overlap$non_zero_voxels_in_dep_map)])
+#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>%
 
-#remove fascicles with cranial nerves
-fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_dep_network_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+  #sum of total volume lost and proportion of volume lost 
+  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
+  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_volume_all_full_fascicles) %>%
 
-plot_all_fascicle_volumes_dep_network_no_cranial_nerves <-ggplot(data = fascicle_volumes_dep_network_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
-  ggtitle("Volume Overlap per Fascicle 3_09") + 
-  geom_bar(stat="identity", fill="006300") +
-  xlab("Fascicle") +
-  ylab("Overlap with Depression Mask") +
-  ylim(0,40000) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5, size=14),
-        axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14),
-        axis.line = element_line(size=1),
-        panel.background = element_rect(colour = "white", fill = "white"))
-print(plot_all_fascicle_volumes_dep_network_no_cranial_nerves)
+  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_indepnet_top_quartile = 
+           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves)))) %>%
+  mutate(proportion_volume_lost_dep_net_by_network_size_sum = sum_fascicle_vol_lost_indepnet_top_quartile /dep_network_total_volume_top_quartile_no_cranial_nerves) %>%
 
-#figure out how many fascicles that is, then 
-quartile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 4),0)
+  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves)))) %>%
+  mutate(proportion_volume_lost_nondep_net_by_network_size_sum = sum_fascicle_vol_lost_nondep_bottom_3_quartiles /bottom_75_total_volume_no_cranial_nerves) %>%
 
-#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
-dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile] #n=19
+  ungroup()
 
-dep_network_total_volume_top_quartile_no_cranial_nerves <-
-sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)])
+###################
 
-#get a network that includes all of the fascicles NOT in dep 25% network
-dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol <- fascicle_volumes_dep_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)] #n=68
+##### Assessing Normality w/histograms and Shapiro_Wilk test for normality ####
 
-bottom_75_total_volume_no_cranial_nerves <-
-sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol)])
+### Test for normality within each network, within each diagnosis
 
-#### Calculate total volume in all non-cranial_nerve_network
-total_volume_all_full_fascicles <- dep_network_total_volume_top_quartile_no_cranial_nerves + bottom_75_total_volume_no_cranial_nerves
-#plot volumes within the depression network, and volumes outside of it
+disease_overall_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum) #W=0.97, p = 3.4 * 10^-7, NON-NORMAL
 
-fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_volumes_dep_sorted_by_overlap[(fascicle_volumes_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol),]
+disease_dep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum) #W=0.97, p = 5.0 * 10^-7, NON-NORMAL
 
-plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
-  geom_bar(stat="identity", fill="violet") +
-  ggtitle("Volume per Fascicle Affected Dep net 3_09, top quartile") + 
-  ylab("Volume in Depression Mask") +
-  ylim(0,40000) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5, size=14),
-        axis.title.x = element_text(size=14),
-        axis.title.y = element_text(size=14),
-        axis.line = element_line(size=1),
-        panel.background = element_rect(colour = "white", fill = "white"))
-print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)
+disease_nondep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum) #W=0.97, p = 1.9 * 10^-7, NON-NORMAL
 
-#bottom 3 quartiles
-fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves<- fascicle_volumes_dep_sorted_by_overlap[(fascicle_volumes_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol),]
+disease_dep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.97, p = 0.0002 NON-NORMAL
 
-plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
-  geom_bar(stat="identity", fill="darkgreen") +
-  ylab("Volume in Depression Mask") +
-  ylim(0,40000) +
-  ggtitle("Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.title = element_text(hjust = 0.5),
-        axis.line = element_line(size=1),
-        panel.background = element_rect(colour = "white", fill = "white"))
-print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs
+disease_nondep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]) #W=0.96, p = 0.0001 NON-NORMAL
+
+## All are non-normal, will use wilcoxon for comparisons
+
+### Within/outside depression network, p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339
+msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum,paired = TRUE)
+
+### Diagnostic effects
+
+#proportion volume lost over whole network = W=14995, p=0.04, dep = 0.39, nondep = 0.30
+proportion_volume_lost_over_whole_brain_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
+
+#effsize 0.11
+proportion_volume_lost_over_whole_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_per_total_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+
+#proportion volume lost within depression network, W = 14941, p-value = 0.03; dep = 0.42, nondep = 0.38
+proportion_volume_lost_dep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
+
+#effsize 0.11
+proportion_volume_lost_over_dep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_dep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+
+#proportion volume lost outside depression network W = 15120, p-value 0.05; dep = 0.36, nondep = 0.33
+proportion_volume_lost_nondep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
+
+#effsize 0.10
+proportion_volume_lost_over_nondep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_nondep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+
+### Interaction Network x Diagnosis
+scale=1
+
+df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
+  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
+  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
+  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  mutate(Network = as.factor(Network)) %>%
+  ungroup()
+
+#eff size = 0.73
+msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions, formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+
+#diagnosis: Dep > nondep T = 2.285, p = 0.023; Network dep>nondep = T=16.6, p = 2*10^-16; intx = T = 2.21, p = 0.027
+diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer)
 
 ```
+
+``` {r plots_for_dx_net_intx}
+########## PLOTS #############
+
+#Within vs outside network
+dep_net_group <- c(rep("Dep_net", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Non_dep_net",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
+net_values <- c(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum)
+
+newdf <- data.frame(dep_net_group, net_values)
+
+violin_within_vs_outside_network_sum_proportion_scaled_by_network_size <- ggplot(newdf, aes(x=dep_net_group, y=net_values, fill=dep_net_group)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "Dep Net", "2" = "Non-Dep Net")) + 
+  ggtitle(paste0("Proportion Fascicle Volume Affected by Network")) +
+  ylab("Proportion (sum total volume lost/Net Vol) ") +
+  ylim(0,1) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_within_vs_outside_network_sum_proportion_scaled_by_network_size) #
+
+#diagnosis
+
+df_proportion_network_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
+dplyr::select(proportion_volume_lost_per_total_network_size_sum, depGroupVar) %>%
+  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
+
+violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
+  geom_violin() +  
+  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
+  scale_x_discrete(labels=c("1" = "NonDepressed", "2" = "Depressed")) + 
+  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
+  ylab("Proportion (sum total volume lost/Net Vol) ") +
+  ylim(0,1) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(violin_dep_vs_nondep_scaled_by_network_size) #
+
+#interaction
+diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, 
+                                                         type = "int",
+                                                         axis.lim = c(0.2,0.5),
+                                                         colors = c("#ff00ff", "#3df2f2"),
+                                                         line.size = 1.3,
+                                                         dodge=0.2,
+                                                         axis.title = c("Diagnosis", "Proportion Affected"),
+                                                         title = "Diagnosis x Depression Network Disease")+
+
+  theme(axis.text.x = element_text(size=12),
+        axis.text.y = element_text(size=12),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        aspect.ratio = 3/2,
+        panel.background = element_rect(colour = "white", fill = "white"))
+
+print(diagnosis_by_network_interaction_lmer_plot)
+
+```
+
+``` {r effect_size_regression}
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
+  filter(depGroupVar != 0)  %>% 
+  mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>% 
+  ungroup()
+
+# lapply to get results of t-test in each fascicle
+#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_w_vol_suffix, function(x) 
+{
+  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
+  eval(parse(text=text_for_eval))
+})
+names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_w_vol_suffix
+
+#convert to data frame
+df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)
+
+```
+
+``` {r bootstrap}
+#### bootstrap
+
+######### functions
+tval <- function(formula, data, indices) {
+  df <- data[indices,] # allows boot to select sample
+  fit <- lm(formula, data=df)
+  return(summary(fit)$coefficients[2,3])
+}
+
+#effect sizes
+# 95% confidence intervals for the correlation
+set.seed(1)
+data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))
+
+results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
+   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
+print(results_vol)
+plot(results_vol)
+boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
+print(boot_results_vol.ci)
+pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
+print(pval_vol) #p=1*10^-16
+
+```
+
+``` {r plot_scatter}
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_d_values)
+names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")
+
+# plot fascicle overlap x Eff size value
+#statistic, p=001
+dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )
+
+scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
+  geom_point(size=1) + 
+  geom_smooth(method=lm)+
+  xlab("Volume of Fascicle in Depression Network") + 
+  ylab("Effect Size (D)") + 
+  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", pval_vol)) + 
+                 #round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
+#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
+  theme(plot.title=element_text(hjust=0.5),
+        axis.text = element_text(size=12, color = "black"),
+        axis.title = element_text(size=14, color = "black"),
+        axis.line = element_line(color="black", size=1.3),
+        panel.background = element_blank())
+print(scatter_d_val_by_dep_net_overlap)
+
+df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_d_value %>% dplyr::select(fascicle, dep_vs_healthy_d)
+write.table(x = df_for_making_colors, file = paste0(homedir, "results/fascicle_and_effect_size_for_volxdx_analysis_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")
+
+```
+
+```{r sensitivity analyses}
+tertile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 3),0)
+quintile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 5),0)
+
+dep_network_names_top_tertile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile] #n=26
+
+dep_network_names_top_quintile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile] #n=15
+
+dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)]
+
+dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)]
+
+top_tertile_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)])
+
+top_quintile_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)])
+
+bottom_tertiles_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol)])
+
+bottom_quintiles_total_volume_no_cranial_nerves<-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol)])
+
+dep_and_healthy_groups_for_ICD_analysis <- dep_and_healthy_groups_for_ICD_analysis %>%
+
+  #tertile and quintiles
+  mutate(sum_fascicle_vol_lost_top_tertile = 
+           rowSums(across(dep_network_names_top_tertile_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_top_tertile_by_network_size_sum = sum_fascicle_vol_lost_top_tertile /top_tertile_total_volume_no_cranial_nerves) %>%
+
+  mutate(sum_fascicle_vol_lost_bottom_tertiles = 
+           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%
+
+  mutate(sum_fascicle_vol_lost_top_quintile = 
+           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%
+
+  mutate(sum_fascicle_vol_lost_bottom_quintiles = 
+           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol))) %>%
+  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves) %>%
+
+#### Save a data frame that contains the mean volume of disease within network, scaled by the size of the network, for visualization
+
+diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
+network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
+mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))
+
+#df_means_volume_scaled_by_network_size <-data.frame(diagnosis_vector, network_vector, mean_diagnosis_by_network_vector)
+#names(df_means_volume_scaled_by_network_size) <- c("diagnosis", "network", "mean")
+#write.table(df_means_volume_scaled_by_network_size, paste0(homedir, "/results/df_means_of_volumes_scaled_by_network_size_dx_net_for_color.csv"), row.names=F, col.names = T, quote = F, sep=",")
+
+#### tertile/quintile
+
+#p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
+msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)
+
+#p=0.0328, dep = 0.4407, nondep = 0.387
+msdisease_diagnostic_effects_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.058, dep = 0.31, nondep= 0.28
+msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=2.2*10^-16, top quintile prop vol affected = 0.419, bottom 4 quintiles = 0.322
+msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)
+
+#p=0.0325, dep = 0.4411, nondep = 0.386
+msdisease_diagnostic_effects_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.048, dep = 0.34, nondep = 0.30
+msdisease_diagnostic_effects_wilcoxon_bottom_quintiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+```

### 2022-09-02 lab_notebook/lab_notebook.Rmd
Lab notebook

@@
 
-[] make interaction graph prettier. Can actually hard code some things
-> get_model_data(diagnosis_by_network_interaction_lmer, type="int")
-# Predicted values of proportion_affected
-
 # Network = Dep_network
@@
 [] separate to make a clean document once confirm analyses w/Ted
+  - making a better spreadsheet that separates out proportion and volume
+
+```{r fiber_volumes_values}
+
+#read in fiber volume values
+fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)
+
+#remove cranial nerves
+fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+#extract data frame that only includes the fascicle proportion lost
+df_fascicles_prop <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]
+df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]
+
+#multiply each row by the vector of total volumes to extract volume of disease within fascicle
+df_fascicles_volumes <- sweep(x = df_fascicles_prop, MARGIN = 2,                        # Sweep with Complex STATS
+                  STATS = fiber_volume_values_no_cranial_nerves$volume_full, FUN = "*")
+
+# change name of df_fascicle_volumes to have the suffix, and add a suffix for the proportion measures
+names(df_fascicles_prop) <- fascicle_names_w_prop_suffix
+names(df_fascicles_volumes) <- fascicle_names_w_vol_suffix
+
+#combine back with data_fram
+#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
+df_demo_and_fascicles_prop_and_vol <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes )
+
+```
+
+  [x] THIS WAS DOUBLE CHECKED with a few spot checks
+>df_demo_and_fascicles_prop_and_vol$CC_prop[4]
+[1] 0.5956769
+> df_demo_and_fascicles_prop_and_vol$CC_vol[4]
+[1] 232672
+> fiber_volume_values_no_cranial_nerves$volume_full[fiber_volume_values_no_cranial_nerves$fascicle=="CC"]
+[1] 390601
+> 0.5956769*390601
+[1] 232672
+> df_demo_and_fascicles_prop_and_vol$AF_L_prop[4]
+[1] 0.1407772
+> df_demo_and_fascicles_prop_and_vol$AF_L_vol[4]
+[1] 5028
+> fiber_volume_values_no_cranial_nerves$volume_full[fiber_volume_values_no_cranial_nerves$fascicle=="AF_L"]
+[1] 35716
+> 0.1407772*35716
+[1] 5027.998
+
+[] make interaction plot prettier
+  - awaiting bart - no other feedback. Will need to do this manually
+[x] try making fascicle maps in dsi studio with different fib template (that better maps onto the template we used (fib))
+  - /Users/eballer/BBL/msdepression/templates/dsi/hcp1065.1mm.fib.gz
+  - Open T3d 
+  - Select all files in HCP_YA1065_tractography/ in each section, except for cranial nerves
+    [x] association [x] cerebellum [x] projection [x] commissural
+  [x] make sure in right order
+  [x] click all the boxes
+  [x] save as one big file
+  - load color file -> tracts misc -> open cluster colors
+  - save this file
+
+[] make depression bold map - need to either apply a brain mask or chose slices selectively b/c of the whole template thing
+
+[] Ted wants me to try for JAMA Neurology
+- 3000 words
+≤5 tables and/or figures
+Structured abstract
+Key Points
+Study Type
+Follow EQUATOR Reporting Guidelines
+Do not submit figures with more than 4 panels unless otherwise justified.
+See the AMA Manual of Style for more guidance on figure types and components.
+
+For JAMA Neurology, 5 figures:
+
+  [] 1. Flow chart for patient selection (fig 1, required)
+  [] 2. Demographics (or just list in text, or make A with flow chart and B with demographics) (table 1)
+  [] 3. Pipeline figure (maybe will be e-file)
+  [] 4. Two part image of depression network
+    [] a: Rank of fascicles by volume of overlap w/dep network (cortical rendering of depression network)
+    [] b: DSI fascicle of volume of overlap w/network (how to scale for visualizations) (new hamphire forest in the fall colors)
+  [] 5. In vs outside dep net
+    a. pink and blue fascicle pcicture part B
+    b. inside vs outside dep net violin
+  [] 6. Diagnosis analysis
+    a. no diff by lesion volum
+    b. but difference in fascicle volume
+  [] 7. Dep Network Interaction bar graph
+    [] is it possible to have a 3 part figure? 
+      - a. interaction graph
+      - b.
+picture of A.
+In depressed patients, volume of disease in dep network in depressed patients in one color, volume of disease in non depressed network in another color, divided by overall network size; B.
+Same thing in nondepressed patients.
+All on same color scale so you can hopefully see brighter colors/contrast in depressed as compared to non-depressed
+  [] 8. Depression network overlap 
+    a. scatter
+    b. Fascicles colored by effect size
+
+Could prob combine 5,6, 7&8
+1. flow chart
+2. Pipeline
+3. Depression network
+4. In vs out network, diagnosis analysis
+5. Interaction and scatter
+
+FIGURE POTENTIAL
+1. Flow chart
+  - demographics (+- promis scores) e-file (?)
+  - pipeline is e -file (can only have max 4 pt figure)
+2. Making depression network 
+  a: Cortical rendering of depression network
+  b: DSI fascicle of volume of overlap w/network (how to scale for visualizations) (new hampshire forest in the fall colors)
+3.  In vs outside dep net
+    a. inside vs outside dep net violin
+    b. pink and blue fascicle picture part B
+4. Diagnosis analysis
+    a. diff by lesion volume (violin)
+    b. but difference in fascicle volume (violin)
+        - just dep vs non dep (put stuff about within/outside network for each diagnostic group in text)
+5. Dep Network Interaction bar graph
+     a. interaction graph
+     b. scatter (effect size by overlap)
+     c. Fascicles colored by effect size
+
+#### 20220902
+[] Cleaning script
+  [] Sections
+    [x] Initialize
+    [x] read in data/subset
+    [x] read in and handle promis scores, link to scan
+    [x] Demographics
+    [x] Define Depression Network
+    [x] Network enrichment/disease within/outside depression network
+    [x] Lesion Volume between diagnostic groups
+    [x] Lesion burden between diagnostic groups
+    [x] Interaction
+    [x] Effect size x fascicle volume of overlap
+
+#### 20220906
+    [] Sensitivity Analyses

### 2022-09-03 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
     [x] Effect size x fascicle volume of overlap
+    [x] Sensitivity Analyses
 
 #### 20220906
-    [] Sensitivity Analyses
+
+    - add vol suffix

### 2022-09-03 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_just_volumes_started_20220824.Rmd
runs all the way through!

@@
 
-``` {r bootstrap}
+``` {r plot_scatter}
+#######################
+### plot differences ##
+#######################
+
+#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
+
+df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_d_values)
+names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")
+
+# plot fascicle overlap x Eff size value
+#statistic, p=001
+dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )
+
+##### bootstrap
+
 #### bootstrap
@@
 
-```
-
-``` {r plot_scatter}
-#######################
-### plot differences ##
-#######################
-
-#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle
-
-df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_d_values)
-names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")
-
-# plot fascicle overlap x Eff size value
-#statistic, p=001
-dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )
+###
 
@@
 ```{r sensitivity analyses}
-tertile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 3),0)
-quintile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 5),0)
 
-dep_network_names_top_tertile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile] #n=26
+## to check different splits of the networks
 
-dep_network_names_top_quintile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile] #n=15
+tertile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 3),0)
+quintile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 5),0)
 
-dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)]
+## get names of each network, and add suffix
+dep_network_names_top_tertile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile]  #n=26
 
-dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)]
+dep_network_names_bottom_tertiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves)]
 
+dep_network_names_top_quintile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile]  #n=15
+
+dep_network_names_bottom_quintiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves)]
+
+### get volumes of each network
 top_tertile_total_volume_no_cranial_nerves <-
-sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)])
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves)])
+
+bottom_tertiles_total_volume_no_cranial_nerves <-
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves)])
 
 top_quintile_total_volume_no_cranial_nerves <-
-sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)])
-
-bottom_tertiles_total_volume_no_cranial_nerves <-
-sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol)])
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves)])
 
 bottom_quintiles_total_volume_no_cranial_nerves<-
-sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol)])
+sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves)])
 
+## change suffix to _vol
+
+dep_network_names_top_tertile_no_cranial_nerves <- dep_network_names_top_tertile_no_cranial_nerves %>% 
+  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)
+
+dep_network_names_bottom_tertiles_no_cranial_nerves <- dep_network_names_bottom_tertiles_no_cranial_nerves %>% 
+  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)
+
+dep_network_names_top_quintile_no_cranial_nerves <- dep_network_names_top_quintile_no_cranial_nerves %>%
+  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)
+
+dep_network_names_bottom_quintiles_no_cranial_nerves <- 
+dep_network_names_bottom_quintiles_no_cranial_nerves %>% 
+  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)
+
+### make new columns
 dep_and_healthy_groups_for_ICD_analysis <- dep_and_healthy_groups_for_ICD_analysis %>%
@@
   mutate(sum_fascicle_vol_lost_top_tertile = 
-           rowSums(across(dep_network_names_top_tertile_no_cranial_nerves_by_vol))) %>%
+           rowSums(across(all_of(dep_network_names_top_tertile_no_cranial_nerves)))) %>%
   mutate(proportion_volume_lost_top_tertile_by_network_size_sum = sum_fascicle_vol_lost_top_tertile /top_tertile_total_volume_no_cranial_nerves) %>%
@@
   mutate(sum_fascicle_vol_lost_bottom_tertiles = 
-           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol))) %>%
+           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves))) %>%
   mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%
@@
   mutate(sum_fascicle_vol_lost_top_quintile = 
-           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves_by_vol))) %>%
+           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves))) %>%
   mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%
@@
   mutate(sum_fascicle_vol_lost_bottom_quintiles = 
-           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol))) %>%
-  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves) %>%
+           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves))) %>%
+  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves)
 
@@
 
-diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
-network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
-mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
-                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
-                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))
+#diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
+#network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
+#mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+ #                            mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))
 
@@
 
+##### within versus outside network
 #p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
 msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)
+
+#p=2.2*10^-16, top quintile prop vol affected = 0.419, bottom 4 quintiles = 0.322
+msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)
+
+#Diagnostic effects
 
@@
 msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
-
-#p=2.2*10^-16, top quintile prop vol affected = 0.419, bottom 4 quintiles = 0.322
-msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)
 

### 2022-09-06 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220824_for_replication.Rmd
Add files via upload

--- 2022-09-03 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_just_volumes_started_20220824.Rmd
@@
 date: '2022-08-24'
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
+
+## MS Depression
+
+Welcome to the MS Depression Data!
+This project was designed to ask a relatively simple question - are patients with MS who are depressed more likely to have lesions in fascicles that connect areas previously associated with depression.
+The literature are sparse in this domain.
+It is very clear that depression occurs in up to 50% of patients with MS, and some studies have suggested that it relates to lesion burden.
+But to my knowledge there are no papers that evaluate how lesions disrupt white matter tracts that serve as the backbone for functional networks.
+Over the course of the next few chunks, I'll walk through our analysis.
+It combines imaging data from research grade scans acquired at 3T as part of a clinical protocol (T1 and T2 Flair), and clinical data from the electronic medical record.
+
+The first chunk here is reading in appropriate data frames and preprocessing.
+A few big data frames are read in.
+The first is the demographic, drug, lab info for the MS patients (from the DAC pull).
+The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle).
+The most important preprocessing occurs in a series of mutate commands.
+These essentially add new columns to the data frame that can later be used.
+The thrust of these mutate commands is to define who has depression and who does not.
+
+Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion.
+Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses.
+For these preliminary analysis, our inclusion/exclusion are more stringent.
+Of note, the graph below is irrespective of whether they had good/bad MIMoSA.
+As such, the MIMoSA subset is smaller.
+
+<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>
+
+### Inclusion/Exclusion
+
+#### Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
+
+  1) Scanned under MS protocol (42K)
+
+  2) MS Diagnosis (17K)
+
+  3) Seen by MS doctor (16K)
+
+#### Inclusion criteria for Depressed Group
+
+  1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
+        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder)
+        - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition
+
+  2) Screened positive for depression on PHQ2 or PHQ9 at any point
+
+  3) Prescribed antidepressant medications per NAMI website
+        - https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications
+
+#### Inclusion criteria for Nondepressed Group
+
+  1) No F* diagnoses
+
+  2) At least 1 PHQ2 or PHQ9 with a documented 0
+
+  3) Free of psychiatric medications per NAMI website
+
+#### EMR Exclusions
+
+  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
+  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression.
+These subjects were excluded from future analyses.
+We could consider adding these to the healthy group maybe, but not too inclined to do that right now.
+
+#### Imaging exclusion
+
+  We are currently only using people who had good MIMOSA (75 or 100).
+That severely limits our numbers, but that is okay!
+When we re-QC, we will probably get a lot more into this group.
+We are specifically using their FIRST scan.
+
+### Group validation and depression network definition
+
+#### Functional Impairment Scores
+
+  To validate our grouping, we wanted to compare symptom burden between patients with depression and nondepressed comparators.
+One challenge of EMR research is that we are limited to what has been coded in particular fields, and could not readily assess our patients, or access their notes.
+As PHQ2 and 9 were used in dignostic categorization, they, by default differ between groups and are less informative.
+MS impairment is often assessed and recorded via EDSS, but this is not currently coded in the EMR.
+However, the Penn EMR has started recording Patient-Reported Outcomes Measurement Information System (PROMIS).
+Per the NIH, "PROMIS addressed a need in the clinical research community for a rigorously tested patient reported outcome (PRO) measurement tool that uses recent advances in information technology, psychometrics, and qualitative, cognitive, and health survey research to measure PROs such as pain, fatigue, physical functioning, emotional distress, and social role participation that have a major impact on quality-of-life across a variety of chronic diseases." PROMIS scores, for reference, are scored where higher numbers mean better functioning, and lower numbers mean worse functioning.
+
+  A small subset of our participents (36 in non-depressed group and 49 in depressed group) had PROMIS scores.
+For participants who had multiple PROMIS scores, the score that was analyzed was the closest in time proximity to the day of the scan.
+PROMIS score cleaning and linking of score to scan is performed in a separate markdown **(promis_score_cleaning_2010-2022.Rmd).
+**
+
+  Based on previous MS literature, we hypothesized that our patients with depression would overall show more functional impairment than patients without depression.
+
+#### Depression Network Construction
+
+  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1).
+It is an unthresholded mask (he initially sent one where the unit is r, but then sent us one where unit is T).
+  Per Siddiqi's article: The location of each lesion or brain stimulation site (Fig.
+2a–c, top panels) was mapped to an underlying brain circuit using a large normative connectome database (n=1000) and previously validated methods (Fig.
+2a–c, bottom panels)5.
+The normative connectome was used to estimate connectivity of each lesion or stimulation site to every voxel in the brain.
+At each voxel, a Pearson r-value was computed for the correlation between depression score and lesion or stimulation site connectivity to that voxel (Fig.
+2a–c, right panels), yielding a population-derived “circuit map” for each of the 14 datasets (Supplementary Figure 1).
+We generated a combined depression circuit map by taking the mean of all 14 circuit maps, weighted by the sample size of each dataset (Fig.
+5a).
+Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted)
+
+  To use it, I thresholded the clusters (3.09), binarized it and then used it as an ROI and calculated, per fascicle, the volume occupied by the fascicle that intersected with the depression mask.
+Most fascicles were either entirely overlapping or entirely outside the depression mask, though the range of overlap was anywhere from 1-30% (volume of fascicle overlap/total volume of fascicle).
+This required us to make some choices.
+We considered an option where the "depression mask" was anywhere that ever overlapped, and non-depression were areas where there was no overlap.
+We also looked at quartiles, and settled on the top 25% (top quartile), i.e.
+the top 25% of fascicles with the highest volume of network overlap were considered in the depression network.
+Everything outside of that was considered "non_depression" network, or nondep_net.
+Cranial nerves are too small to reliably be assess with our method, and so they were removed from the analysis.In total, 77 fascicles were evaluated.
+
+### Analyses
+  Demographic differences between groups were assessed with T-tests and chi-square for categorical variables w/the command CreateTableOne.
+
+#### *Disease burden measures per individual (done in the shell):*
+
+    For each individual, the number of 1s in the MIMoSA binary mask were summed to get a measure of overall lesion burden (to be distinguished from fascicle burden).
+
+    To assess fascicle burden for each individual, streamline filtering was performed.
+
+    Streamline filtering is an interative process performed in DSI studio.
+For each individual, the MIMoSA binary map was considered a region of interest.
+For each of the 77 fascicles, streamlines that ran through the lesion were "filtered" or kept, whereas the fascicles that avoided the MIMoSA mask were eliminated.
+Streamlines that passed through the MIMoSA were then saved binary .nii files, where 1 indicated that disease was present in that voxel, and 0 indicated either 1) that fascicle did not cross through that voxel or there was no disease in it.
+
+    I was then able to calculate the "volume" of the disease in a fascicle (i.e.
+volume of the streamlines that were affected) by summing the # of 1s in the map.
+At the end, each individual had 77 single values that represented the volume of affected streamlines within each fascicle.
+
+    Full fascicle volumes were also calculated and saved as .niis.
+
+####  *Disease burden summary measures (done in this R script):*
+
+    Having computed disease measures at the individual fascicle, I wanted to assess network effects for all future analyses. To do this, I calculated three summary measures per individual.
+
+    1) Total disease : Sum of all 77 volume measures of disease, divided by the total volume of "healthy" fascicles (i.e.
+taking a sum of all of the full fascicle volumes).
+This yields a proportion of the overall burden of disease in the brain
+
+    2) Depression network: Sum of all 19 volume measures of disease in the fascicles within the depression network, divided by the sum of total full fascicle volumes in the depression network.
+
+    3) Nondepression network: Sum of all 58 volume measures of disease in the fascicles outside the depression network, divided by the sum of total full fascicle volumes in the nondepression network.
+
+####  *Network effects:*
+
+   Given that depression is so common in multiple sclerosis, we first wanted to check whether disease was diffuse across the brain, or whether there was network specificity (i.e.
+is MS random, or do MS lesions happen to strike fascicles that support the depression network, irrespective of diagnosis).
+To assess this, I took the following steps.
+
+    1) Tests for normality (Shapiro–Wilk test) of proportion of disease in depression network and nondepression network. Both not normally distributed
+
+    2) Used wilcoxon tests to compare disease (paired = T) between depression and nondepression network.
+As each individual contributed one measure for each network (i.e.
+2 measures per person), paired was used
+
+####  *Diagnostic effects:*
+
+  Per our inclusion/exclusion criteria, we were able to identify a subset of patients that had depression, as well as a subset who did not meet criteria per EMR review.
+We next wanted to assess whether patients with depression had more overall disease than patients without depression.
+
+    First, we assessed whether lesion burden (i.e. the volume occupied by white matter disease), irrespective of which fascicles were impacted, differed between patients and controls.
+
+    1) Tests for normality (Shapiro–Wilk test), volume of lesions not normally distributed.
+
+    2) Used wilcoxon tests to compare lesion burden between depressed and non-depressed patients. As each person only contributed one measure, paired = F.
+
+  Next, we assessed whether fascicle burden (i.e. total volume of disease divided by total volume of all fascicles) differed between diagnostic groups.
+
+    1) Tests for normality (Shapiro–Wilk test), fascicle burden not normally distributed.
+
+    2) Used wilcoxon tests to compare fascicle burden between depressed and non-depressed patients. As each person only contributed one measure, paired = F.
+
+  We next aimed to see whether there was an interaction of Diagnosis and Network.
+Specifically, we hypothesized that patients with depression would have more disease specifically in the depression network.
+
+     1) Model: diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
+
+  Given the somewhat arbitrary definition of depression network (25%/75%), we next assessed whether the relationship between diagnosis and network was continuous.
+
+    1) For each fascicle, two values were computed
+
+      a) Effect size (r) from the wilcoxon statistic comparing volume of disease in the fascicle between depressed vs nondepressed
+
+      b) The volume of the overlap of that fascicle with the depression network
+
+    2) A linear model relating the overlap of volume of the fascicle w/the depression network to the effect size from the depressed v nondepressed wilcoxon analysis.
+
+    3) This relationship was permuted 10,000 w/boot (w/replacement)
+
+    4) Plots were generated
+
+####  *Sensitivity Analyses*
+
+  Though the effect size by overlap analysis was reassuring, we wanted to further guarantee that our findings were not driven by arbitrary distinctions between the depression and non-depression networks.
+To assess this, we repeated the network analysis using tertiles and quintiles.
+LMER for assessing interactions was used.
+
+    1) Tertiles:
+
+      a) Depression network: Top Tertile
+      b) Nondepression network: Bottom 2 Tertiles
+
+    2) Quintiles:
+
+      a) Depression network: Top Quintile
+      b) Nondepression network: Bottom 4 Quintiles
+
+## Results:
+
+#### Demographics
+
+232 with depression (depGroupVar == 2), 148 healthy (depGroupVar == 1)
+
+                         Healthy          Depressed             p      test
+  n                                148            232                     
+  Race (%)         Caucasian       108 ( 73.0)    173 ( 74.6)   0.821     
+                   Non-caucasian    40 ( 27.0)     59 ( 25.4)             
+  Sex (%)          Female          117 ( 79.1)    199 ( 85.8)   0.117     
+                   Male             31 ( 20.9)     33 ( 14.2)             
+  Age (mean (SD))                46.57 (12.66)  48.75 (11.75)   0.089     
+  Depression (%)   1               148 (100.0)      0 (  0.0)  <0.001     
+                   2                 0 (  0.0)    232 (100.0)             
+  PHQ2 (mean (SD))                0.00 (0.00)    0.58 (1.53)   <0.001     
+  PHQ9 (mean (SD))                0.00 (NA)     10.43 (9.02)       NA
+
+#### PROMIS Scores
+                                               level   1              2              p      test
+  n                                                       36             49                     
+  Quality_of_Life (mean (SD))                           3.59 (1.02)    2.95 (0.99)    0.010     
+  Physical_Health (mean (SD))                           2.86 (0.83)    2.41 (0.76)    0.019     
+  Mental_Health_and_Mood (mean (SD))                    3.97 (0.82)    3.39 (0.99)    0.011     
+  Social_Activities_Satisfaction (mean (SD))            3.83 (1.00)    2.84 (1.36)    0.001     
+  Carrying_Out_Social_Activities (mean (SD))            3.71 (1.12)    2.93 (1.21)    0.007     
+  Carrying_Out_Physical_Activities (mean (SD))          3.68 (1.06)    3.14 (1.00)    0.032     
+  Emotional_Problems (mean (SD))                        3.67 (1.00)    3.00 (1.08)    0.011     
+  **Fatique_Average (mean (SD))                           3.48 (1.25)    3.09 (1.14)    0.181  *   
+  PostOp.PROMIS.Physical.Score (mean (SD))             13.15 (3.03)   11.67 (2.55)    0.032     
+  PostOp.PROMIS.Mental.Score (mean (SD))               15.00 (2.92)   12.31 (3.31)    0.001
+
+  This is actually great news - if depressed/nondepressed individuals differed on fatigue, you could make the case that of COURSE they'd feel worse as a consequence of their fatigue.
+However, that they weren't different here lends some specificity to the measures.
+
+#### Depression Network Construction
+
+dep_network_names_top_quartile_no_cranial_nerves
+ [1] CC      AF_L    SLF2_R  SLF3_R  FAT_L   TR_S_R  SLF2_L  AF_R    TR_S_L  FAT_R   CS_S_R  CPT_F_R CPT_P_L ILF_R   CPT_P_R CS_S_L  PAT_R   MdLF_L  CBT_R
+
+dep_network_names_bottom_3_quartiles_no_cranial_nerves
+  [1] C_FPH_L C_FPH_R C_FP_L  C_FP_R  C_PH_L  C_PHP_L C_PHP_R C_PH_R  C_R_L   C_R_R   EMC_L   EMC_R   IFOF_L  IFOF_R  ILF_L   MdLF_R  PAT_L   SLF1_L  SLF1_R 
+[20] SLF3_L  UF_L    UF_R    VOF_L   VOF_R   CB_L    CB_R    ICP_L   ICP_R   MCP     SCP     V       AR_L    AR_R    CBT_L   CPT_F_L CPT_O_L CPT_O_R CS_A_L 
+[39] CS_A_R  CS_P_L  CS_P_R  CST_L   CST_R   DRTT_L  DRTT_R  F_L     F_R     ML_L    ML_R    OR_L    OR_R    RST_L   RST_R   TR_A_L  TR_A_R  TR_P_L  TR_P_R 
+[58] AC
+
+#### Network Effects
+
+    - Disease within versus outside depression network: p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339, eff size 0.73
+
+#### Diagnostic Effect
+
+    - Volume of MIMoSA lesion burden: p=0.036; depressed = 15,871; nondepressed = 12,886, eff size 0.107
+
+    - Proportion Volume of whole brain fascicle disease burden: W=14995, p=0.04, dep = 0.39, nondep = 0.30
+
+#### Network x Diagnosis Interaction
+
+    - Main effect of Diagnosis: T = 2.29, p = 0.02 (Depressed patient disease > Nondepressed patient disease)
+
+    - Main effect of Network: T = 16.58, p <2.2 * 10^-16 (Depression Network more disease than nondepression network)
+
+    - Interaction= T = 2.2, p = 0.03; finding driven by worse disease in depressed patients, within depression network
+
+#### Effect Size x Fascicle overlap with depression network scatter
+
+    - p before bootstrap: p=1.8*10^-6; effect increases as overlap with depression network increases
+
+    - p after bootstrap: p = < 1*10^-16
+
+#### Sensitivity Analyses
+
+      1) Tertiles
+
+        - Main effect of network (Depression > Nondepresison network): T=23.2, p = 2*10^-16
+
+        - Main effect of diagnosis (Depressed > Nondepressed people):  T = 2.37, p = 0.018
+
+        - Interaction of network and diagnosis:  T = 2.17, p = 0.031
+
+      2) Quintiles
+
+        - Main effect of network (Depression > Nondepresison network): T=22.56, p = 2*10^-16;
+
+        - Main effect of diagnosis (Depressed > Nondepressed people): T = 2.285, p = 0.019
+
+        - Interaction of network and diagnosis, T = 2.34, p = 0.020
 
@@
 
+homedir <- "/Users/eballer/BBL/msdepression/"
+
 #functions
-source("~/BBL/msdepression/scripts/msdepression_functions.R")
-#source("~/BBL/msdepression/scripts/wilcox_effsize.R")
+source(paste0(homedir,"scripts/msdepression_functions.R"))
+
 ```
 
-Read in Data and do some preprocessing
+### Read in Data and do some preprocessing
 
 ```{r read_in_df_and_subset}
-homedir <- "/Users/eballer/BBL/msdepression/"
 
@@
 ```
-PROMIS DATA HANDLING to link a person to their most proximal PROMIS
+### PROMIS DATA HANDLING to link a person to their most proximal PROMIS
 
@@
 
-Demographics
+### Demographics
 ```{r demographics}
@@
 
-Define Depression Network by Volume of Overlap
+### Define Depression Network by Volume of Overlap
 ```{r visualize_depression_network_volume}
@@
 dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile]
+
+write.table(dep_network_names_top_quartile_no_cranial_nerves, paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), row.names = F, col.names = F, quote = F)
+
+df_for_making_binary_color_maps <- fascicle_volumes_dep_network_no_cranial_nerves %>% dplyr::select(fascicle, non_zero_voxels_in_dep_map)
+
+write.table(df_for_making_binary_color_maps, paste0(homedir, "/results/volume_of_overlap_w_dep_net_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")
 
@@
    gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
+
 ```
 
-Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups
+### Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups
 
@@
 
-dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000) #eff size 0.107 (small)
 
@@
 ```
+
+### Compare between networks, diagnostic groups, and interaction
 
@@
 ```
-
+### Plot interaction
 ``` {r plots_for_dx_net_intx}
@@
 ```
-
+### Effect size regression (effect size x overlap with dep net)
 ``` {r effect_size_regression}
@@
 
-# lapply to get results of t-test in each fascicle
-#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
-fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_w_vol_suffix, function(x) 
+# 
+#depGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
+fascicle_dep_vs_healthy_effsize <- sapply(fascicle_names_w_vol_suffix, function(fascicle) 
 {
-  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
+
+  text_for_eval <- paste0("rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = ", fascicle, " ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = \"perc\", nboot = 1000)$effsize")
   eval(parse(text=text_for_eval))
 })
-names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_w_vol_suffix
+names(fascicle_dep_vs_healthy_effsize) <- fascicle_names_w_vol_suffix
 
 #convert to data frame
-df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)
+df_effsize_values <- as.data.frame(fascicle_dep_vs_healthy_effsize)
+
+write.table(x = df_effsize_values, file = paste0(homedir, "results/fascicle_and_wilcox_effect_size_for_volxdx_analysis_n77.csv"), row.names = T, col.names = F, quote = F, sep = ",")
 
@@
 
+### Plot effect size x depression net overlap
 ``` {r plot_scatter}
@@
 
-df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_d_values)
-names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")
+df_fascicle_dep_mask_volume_overlap_effsize_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_effsize_values)
+names(df_fascicle_dep_mask_volume_overlap_effsize_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_effsize")
 
 # plot fascicle overlap x Eff size value
-#statistic, p=001
-dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )
-
-##### bootstrap
+#statistic, p=1.8*10^-6
+dep_overlap_x_dep_vs_healthy_vol_effsize <- lm(dep_vs_healthy_effsize ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_effsize_value )
 
@@
 set.seed(1)
-data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))
+data_frame_dep_v_nondep_effsize_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_effsize_value, select = c("dep_vs_healthy_effsize", "volume_in_dep_mask"))
 
-results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
-   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
+results_vol <- boot(data=data_frame_dep_v_nondep_effsize_x_vol_dep_mask, statistic=tval,
+   R=10000, formula=dep_vs_healthy_effsize~volume_in_dep_mask)
 print(results_vol)
@@
 pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
-print(pval_vol) #p=1*10^-16
+print(pval_vol) #p=1*10^-16; CI (1.774-5.144)
 
@@
 
-scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
+scatter_effsize_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize)) +
   geom_point(size=1) + 
@@
   xlab("Volume of Fascicle in Depression Network") + 
-  ylab("Effect Size (D)") + 
+  ylab("Effect Size (r)") + 
   ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", pval_vol)) + 
-                 #round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
+                 #round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize )$coefficients[2,4],3))) + 
 #  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
@@
         panel.background = element_blank())
-print(scatter_d_val_by_dep_net_overlap)
+print(scatter_effsize_val_by_dep_net_overlap)
 
-df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_d_value %>% dplyr::select(fascicle, dep_vs_healthy_d)
+df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_effsize_value %>% dplyr::select(fascicle, dep_vs_healthy_effsize)
 write.table(x = df_for_making_colors, file = paste0(homedir, "results/fascicle_and_effect_size_for_volxdx_analysis_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")
@@
 ```
+
+### Sensitivity Analyses
 
@@
 
-#### Save a data frame that contains the mean volume of disease within network, scaled by the size of the network, for visualization
-
-#diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
-#network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
-#mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
-#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
-#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
- #                            mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))
-
-#df_means_volume_scaled_by_network_size <-data.frame(diagnosis_vector, network_vector, mean_diagnosis_by_network_vector)
-#names(df_means_volume_scaled_by_network_size) <- c("diagnosis", "network", "mean")
-#write.table(df_means_volume_scaled_by_network_size, paste0(homedir, "/results/df_means_of_volumes_scaled_by_network_size_dx_net_for_color.csv"), row.names=F, col.names = T, quote = F, sep=",")
-
 #### tertile/quintile
@@
 
-#Diagnostic effects
+#################
+## Interaction ##
+#################
 
-#p=0.0328, dep = 0.4407, nondep = 0.387
-msdisease_diagnostic_effects_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+#### tertile
+df_empi_dx_network_proportions_tertiles <- dep_and_healthy_groups_for_ICD_analysis %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
+  mutate(Dep_network = proportion_volume_lost_top_tertile_by_network_size_sum) %>%
+  mutate(Nondep_network = proportion_volume_lost_bottom_tertiles_by_network_size_sum) %>%
+  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  mutate(Network = as.factor(Network)) %>%
+  ungroup()
 
-#p=0.058, dep = 0.31, nondep= 0.28
-msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+#diagnosis: Dep > nondep T = 2.37, p = 0.018; Network dep>nondep = T=23.2, p = 2*10^-16; intx = T = 2.17, p = 0.031
+diagnosis_by_network_interaction_lmer_tertile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_tertiles)
+summary(diagnosis_by_network_interaction_lmer_tertile)
 
-#p=0.0325, dep = 0.4411, nondep = 0.386
-msdisease_diagnostic_effects_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+#### quintile
+df_empi_dx_network_proportions_quintiles <- dep_and_healthy_groups_for_ICD_analysis %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
+  mutate(Dep_network = proportion_volume_lost_top_quintile_by_network_size_sum) %>%
+  mutate(Nondep_network = proportion_volume_lost_bottom_quintiles_by_network_size_sum) %>%
+  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  mutate(Network = as.factor(Network)) %>%
+  ungroup()
 
-#p=0.048, dep = 0.34, nondep = 0.30
-msdisease_diagnostic_effects_wilcoxon_bottom_quintiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+#diagnosis: Dep > nondep T = 2.285, p = 0.019; Network dep>nondep = T=22.56, p = 2*10^-16; intx = T = 2.34, p = 0.020
+diagnosis_by_network_interaction_lmer_quintile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_quintiles)
+summary(diagnosis_by_network_interaction_lmer_quintile)
 

### 2022-09-06 msdepression/scripts/make_binary_colored_depression_net_maps.R
Add files via upload

@@ -15,16 +15,16 @@ homedir="/Users/eballer/BBL/msdepression/"
 
 dep_network_names_top_quartile_no_cranial_nerves_by_vol <- read.csv(file = paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), header = F)
 
-files = c("fascicle_proportion_overlap_with_depression_network_n77")
+files = c("volume_of_overlap_w_dep_net_n77") #c("fascicle_proportion_overlap_with_depression_network_n77")
 
-#color_depressed_net = "255 0 255" #pink
-#color_nondepressed_net = "0 255 255" #light blue
-color_depressed_net = "255 255 0" #yellow
-color_nondepressed_net = "0 0 255" #blue
+color_depressed_net = "255 0 255" #pink
+color_nondepressed_net = "0 255 255" #light blue
+#color_depressed_net = "255 255 0" #yellow
+#color_nondepressed_net = "0 0 255" #blue
 
 for (filename in files) {
   #suffix, CHANGE IF YOU WANT A DIFFERENT FLAG
-  suffix = "yb"
+  suffix = "pinkblue"
   
   #read in file with fascicle name and proportion
   fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)

### 2022-09-06 msdepression/scripts/make_red_to_yellow_RGB_color_scheme.R

@@ -26,12 +26,13 @@
 
 homedir="/Users/eballer/BBL/msdepression/"
 
-files = c("fascicle_mean_proportion_overlap_nfasc_77", "fascicle_proportion_overlap_with_depression_network_n77", "fascicle_and_effect_size_for_volxdx_analysis_n77")
+files = c("fascicle_mean_proportion_overlap_nfasc_77", "fascicle_proportion_overlap_with_depression_network_n77", "fascicle_and_wilcox_effect_size_for_volxdx_analysis_n77")  # "fascicle_and_effect_size_for_volxdx_analysis_n77")
 
 for (filename in files) {
   #suffix
-  suffix = "_pg" #pink gradient
-  
+  #suffix = "_ry" #red-yello gradient
+
+    suffix = "_pg"
   #read in file with fascicle name and proportion
   fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
   names(fascicle_name_and_prop) <- c("fascicle", "proportion")
@@ -48,6 +49,7 @@ for (filename in files) {
   red <- rep(x=255, times = 77)
   
   fascicle_name_and_prop$red <- red
+
   #multiply proportion column by 100, then by step, and round
  # fascicle_name_and_prop$green<- round((fascicle_name_and_prop$proportion *100*step),0)
   fascicle_name_and_prop$green<- 255-round((fascicle_name_and_prop$proportion *100*step),0) #for pg, I want darker colors to be more sig
@@ -55,24 +57,12 @@ for (filename in files) {
   
   #make column for blue
  # blue <- rep(x=0, times = 77)
-  blue <- rep(x=255, times = 77)
+
+    blue <- rep(x=255, times = 77)
   fascicle_name_and_prop$blue=blue
   
-  #sort based on proportion to check, it works! Lower values are associated with lower RGB
-  #sorted <- fascicle_name_and_prop[order(fascicle_name_and_prop$proportion),]
-  
-  #make individual file colors for loading
-  #for (fascicle in 1:dim(fascicle_name_and_prop)[1]) {
-    
-   # fascicle_name = fascicle_name_and_prop$fascicle[fascicle]
-  #  fascicle_rgb = paste0(fascicle_name_and_prop$red[fascicle], " ", fascicle_name_and_prop$green[fascicle], " ", fascicle_name_and_prop$blue[fascicle])
-  #  print(paste0(fascicle_name, " ", fascicle_rgb))
[…]

### 2022-09-09 msdepression/scripts/make_red_to_yellow_RGB_color_scheme_accounting_for_cc.R
changed to account for CC being so big

@@ -1,0 +1,87 @@
+#### Welcome to make_red_to_yellow_RGB_color_scheme_accounting_for_corpus callosum! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle to be its volume of loss, scaled on the RGB scale
+### In order to do this, I needed to figure out how to map my volumes to RGB
+
+### Pre: File of volumes
+### Post: Output of RGB values in Red to Yellow Scale for the Volumes
+### Uses: Step 0: Figure out how to scale from red to yellow 
+###           - Red = 255 0 0, yellow 255 255 0
+###           - Means that we need to vary the middle value (255) to get where we need to be
+###       Step 1) For red to yellow, calculate the step (i.e. how much each value will be)
+###             - because my volumes are out of 1, I wanted each step to be 1/100
+###             - Usually I set this value as "max"
+###             - But with CC being so big, things scale poorly. So I will set cc to full yellow, and set max to be the next biggest
+###             - This is similar to normal thresholding, where everything above a value will be a specific color
+###       Step 2) Multiply the volume by 100, so now each count is 1-100, in percents
+###       Step 3) Multiply the percent by the step
+###       Step 4) Save this value in that fascicle, 255 [new_value] 0
+###       Step 5) if you want to change how the max value is scaled, change SLICE to be the # (slice(1) - max value, slice(2) - second value, etc)
+###
+### Dependencies: Any R will do
+
+###
+
+### gradients
+## red-yellow - R always 255, green scaled (0 is red, 255 is yellow), blue always 0
+## pink gradient (pg) - Red always 255, green scaled 0 to 255 , blue always 255
+
+homedir="/Users/eballer/BBL/msdepression/"
+
+files = c("volume_of_overlap_w_dep_net_n77")  # "fascicle_mean_volume_overlap_nfasc_77", "fascicle_volume_overlap_with_depression_network_n77", "fascicle_and_effect_size_for_volxdx_analysis_n77")
+
+for (filename in files) {
+  #suffix
+  suffix = "_ry_scaling_cc" #red-yello gradient
+
+  #  suffix = "_pg"
+  #read in file with fascicle name and volume
[…]

### 2022-09-09 lab_notebook/lab_notebook.Rmd
with figures

@@
     - add vol suffix
+
+#### 20220906
+
+- Getting rid of some of the within network analyses in the replication. Just going to do interactions because they look great.
+
+``` {r teriles_and_quintiles}
+
+#Diagnostic effects
+
+#p=0.0328, dep = 0.4407, nondep = 0.387
+msdisease_diagnostic_effects_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.058, dep = 0.31, nondep= 0.28
+msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.0325, dep = 0.4411, nondep = 0.386
+msdisease_diagnostic_effects_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+#p=0.048, dep = 0.34, nondep = 0.30
+msdisease_diagnostic_effects_wilcoxon_bottom_quintiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
+
+```
+
+also getting rid of some commented out code for saving things
+
+``` {r commented_out_stuff_from_quintile_tertile_section}
+#### Save a data frame that contains the mean volume of disease within network, scaled by the size of the network, for visualization
+
+#diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
+#network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
+#mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
+#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
+ #                            mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))
+
+#df_means_volume_scaled_by_network_size <-data.frame(diagnosis_vector, network_vector, mean_diagnosis_by_network_vector)
+#names(df_means_volume_scaled_by_network_size) <- c("diagnosis", "network", "mean")
+#write.table(df_means_volume_scaled_by_network_size, paste0(homedir, "/results/df_means_of_volumes_scaled_by_network_size_dx_net_for_color.csv"), row.names=F, col.names = T, quote = F, sep=",")
+
+```
+
+Getting rid of effect size regression with cohen d (b/c fascicles aren't normally distributed)
+``` {r effect_size_regression}
+#  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", fascicle, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", fascicle, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
+```
+
+#### 20220908
+
+[x] got replication off to Elizabeth on 9/7
+[x] Making figures
+  [x] For brains, made red to yellow map for volume of overlap w/depression network (scaled w AF, not cc because it is too big)
+    [x] brain results file: /Users/eballer/BBL/msdepression/volume_of_overlap_w_dep_net_n77
+    [x] color file: /Users/eballer/BBL/msdepression/templates/dti/colors/volume_of_overlap_w_dep_net_n77_ry_scaling_cc.txt
+  [x] made pink gradient map for the vol/effect size map
+    [x] brain results file: /Users/eballer/BBL/msdepression/fascicle_and_wilcox_effect_size_for_volxdx_analysis_n77
+    [x] color file: /Users/eballer/BBL/msdepression/templates/dti/colors/fascicle_and_wilcox_effect_size_for_volxdx_analysis_n77_pg.txt
+  [x] made binary map of volume
+    [x] brain results file: /Users/eballer/BBL/msdepression/volume_of_overlap_w_dep_net_n77
+    [x] color file: /Users/eballer/BBL/msdepression/templates/dti/colors/volume_of_overlap_w_dep_net_n77_pinkblue.txt
+
+[] make figure of depression network from bold
+  [x] resampled the harvard map to mni_icbm_2009b template: 
+  [x] multiplied the resampled image with the mask
+  [x] displayed in fsleyes and looks amazing
+    underlay: mni_icbm152_t1_tal_nlin_asym_09a_mask.nii (drop intensity a bit to get the background not to have that ring around it)
+    overlay: DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii (set to 3.09/3.09 for yellow, 3.09/5 for colored)
+
+```{bash resampling}
+
+ #resample to get in same space
+
+ 3dresample -master mni_icbm152_t1_tal_nlin_asym_09a.nii -input DepressionCircuit_t.nii -prefix DepressionCircuit_t_resampled_to_mni_icbm152_asym09a.nii
+
+ #mask
+  3dcalc -a DepressionCircuit_t_resampled_to_mni_icbm152_asym09a.nii -b mni_icbm152_t1_tal_nlin_asym_09a_mask.nii -expr 'a*b' -prefix DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii
+
+# display
+ fsleyes
+ #underlay: mni_icbm152_t1_tal_nlin_asym_09a_mask.nii
+#overlay: DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii
+```
+
+```{r make_red_to_yello_RGB_color_scheme_accounting_for_cc} 
+#### Welcome to make_red_to_yellow_RGB_color_scheme_accounting_for_corpus callosum! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle to be its volume of loss, scaled on the RGB scale
+### In order to do this, I needed to figure out how to map my volumes to RGB
+
+### Pre: File of volumes
+### Post: Output of RGB values in Red to Yellow Scale for the Volumes
+### Uses: Step 0: Figure out how to scale from red to yellow 
+###           - Red = 255 0 0, yellow 255 255 0
+###           - Means that we need to vary the middle value (255) to get where we need to be
+###       Step 1) For red to yellow, calculate the step (i.e. how much each value will be)
+###             - because my volumes are out of 1, I wanted each step to be 1/100
+###             - Usually I set this value as "max"
+###             - But with CC being so big, things scale poorly. So I will set cc to full yellow, and set max to be the next biggest
+###             - This is similar to normal thresholding, where everything above a value will be a specific color
+###       Step 2) Multiply the volume by 100, so now each count is 1-100, in percents
+###       Step 3) Multiply the percent by the step
+###       Step 4) Save this value in that fascicle, 255 [new_value] 0
+###       Step 5) if you want to change how the max value is scaled, change SLICE to be the # (slice(1) - max value, slice(2) - second value, etc)
+###
+### Dependencies: Any R will do
+
+###
+
+### gradients
+## red-yellow - R always 255, green scaled (0 is red, 255 is yellow), blue always 0
+## pink gradient (pg) - Red always 255, green scaled 0 to 255 , blue always 255
+
+homedir="/Users/eballer/BBL/msdepression/"
+
+files = c("volume_of_overlap_w_dep_net_n77")  # "fascicle_mean_volume_overlap_nfasc_77", "fascicle_volume_overlap_with_depression_network_n77", "fascicle_and_effect_size_for_volxdx_analysis_n77")
+
+for (filename in files) {
+  #suffix
+  suffix = "_ry_scaling_cc" #red-yello gradient
+
+  #  suffix = "_pg"
+  #read in file with fascicle name and volume
+  fascicle_names_and_vol <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
+  names(fascicle_names_and_vol) <- c("fascicle", "volume")
+
+  #Make each value into a volume by dividing the volume by the total volume of all values
+  total_volume_of_all_overlapping_fascicles <- sum(fascicle_names_and_vol$volume)
+
+  fascicle_names_and_vol$proportion <- fascicle_name_and_vol$volume/total_volume_of_all_overlapping_fascicles
+
+  #set scale dynamically
+  #highest possible number in scale
+
+  #sort by volume, take 2nd highest, extract proportion
+  max_value<- fascicle_names_and_vol %>% 
+    dplyr::select(proportion) %>%
+    arrange(desc(proportion)) %>%
+    slice(2) * 100
+
+  #step that each value in our scale will correspond to
+  max_value <- as.numeric(max_value)
+  step <- 255/max_value
+
+  # make a column for red
+
+  red <- rep(x=255, times = 77)
+
+  fascicle_names_and_vol$red <- red
+
+  #if proportion greater than max, set it to max value, otherwise, scale
+  fascicle_names_and_vol$green<-  ifelse(fascicle_names_and_vol$proportion *100 > max_value, 255, round((fascicle_names_and_vol$proportion *100*step),0)) #ry
+
+  #make column for blue
+  blue <- rep(x=0, times = 77) #ry
+
+ #blue <- rep(x=255, times = 77) #pg
+  fascicle_names_and_vol$blue=blue
+
+#output color file (77 lines, each represents a color), and color name vector. Not sure how to do 1:1 assignments rather than manually, but I'll see if I can figure that out!
+  fascicle_names_and_vol_rgb_only <- fascicle_names_and_vol %>%
+    mutate(rgb = paste0(fascicle_names_and_vol$red, " ", fascicle_names_and_vol$green, " ", fascicle_names_and_vol$blue)) %>%
+    dplyr::select(fascicle, rgb)
+
+  write.table(file = paste0(homedir, "/templates/dti/colors/", filename, suffix, "_fascicle_names.txt"), x = fascicle_names_and_vol_rgb_only$fascicle, row.names = FALSE, col.names = FALSE, quote = FALSE)
+
+  write.table(file = paste0(homedir, "/templates/dti/colors/", filename, suffix, ".txt"), x = fascicle_names_and_vol_rgb_only$rgb, row.names = FALSE, col.names = FALSE, quote = FALSE)
+}
+
+```

### 2022-09-16 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220824_for_replication.Rmd
Changed scatter

@@
 The first is the demographic, drug, lab info for the MS patients (from the DAC pull).
-The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle).
+The second is the data frame of fascicle proportion affected for people with good MIMoSA (# volume affected/total volume of fascicle).
 The most important preprocessing occurs in a series of mutate commands.
@@
 
-#### Inclusion criteria for Depressed Group
+#### Inclusion criteria for Depressed Group (one of the three criteria must be met)
 
   1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
-        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder)
+        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder
         - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition
@@
 
-#### Inclusion criteria for Nondepressed Group
+#### Inclusion criteria for Nondepressed Group (ALL three criteria must be met)
 
@@
 
-  We are currently only using people who had good MIMOSA (75 or 100).
+  We are currently only using people who had good MIMoSA (75 or 100).
 That severely limits our numbers, but that is okay!
@@
 5a).
-Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted)
+Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted).
 
@@
 As each individual contributed one measure for each network (i.e.
-2 measures per person), paired was used
+2 measures per person), paired was used.
 
@@
 #functions
-source(paste0(homedir,"scripts/msdepression_functions.R"))
+source(paste0(homedir,"/scripts/msdepression_functions.R"))
 
@@
 
-columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")
+columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE")
+                             #,"PHQ.2","PHQ.9")
+
+#columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")
 
@@
 
-fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
+fascicle_volumes_dep <- read.csv(paste0(homedir, "/results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")
 
@@
 
-lesion_volumes <- read.csv(paste0(homedir, "results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE)
+lesion_volumes <- read.csv(paste0(homedir, "/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE)
 
@@
 #remove cranial nerves
-fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
+  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves)))
 
@@
 #df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
-df_demo_and_fascicles_prop_and_vol <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes )
+df_demo_and_fascicles_prop_and_vol <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes)
 
@@
 
-unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=116
+unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=121
 
@@
   ylab("Volume of Lesions") +
-  theme(plot.title = element_text(size=16, hjust = 0.8),
+  theme(plot.title = element_text(size=16, hjust = 0.55),
         axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
@@
 #Within vs outside network
-dep_net_group <- c(rep("Dep_net", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Non_dep_net",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
+dep_net_group <- c(rep("Depression Network", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Nondepression Network",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
 net_values <- c(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum)
@@
 newdf <- data.frame(dep_net_group, net_values)
+
+newdf <- newdf %>%
+  mutate( dep_net_group=factor(dep_net_group,levels=c( "Nondepression Network", "Depression Network")))
 
@@
   geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
-  scale_x_discrete(labels=c("1" = "Dep Net", "2" = "Non-Dep Net")) + 
-  ggtitle(paste0("Proportion Fascicle Volume Affected by Network")) +
-  ylab("Proportion (sum total volume lost/Net Vol) ") +
+  ggtitle(paste0("Volume of Fascicle Disease Burden")) +
+  ylab("Proportion of network affected") +
   ylim(0,1) + 
+  scale_fill_manual(values=c("#80ffff", "#ff40ff")) +
   theme(plot.title = element_text(size=16, hjust = 0.55),
-        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=16),
+        axis.text = element_text(size = 16, colour = "black"), axis.title.y  = element_blank(), 
+        axis.title.x = element_text(size=16),
         axis.line = element_line(size = 1.5, colour = "black", linetype=1),
         legend.position = "none",
-        aspect.ratio=7/5,
-        panel.background = element_rect(colour = "white", fill = "white"))
+        aspect.ratio=4/7,
+        panel.background = element_rect(colour = "white", fill = "white")) + 
+  coord_flip()
+
 print(violin_within_vs_outside_network_sum_proportion_scaled_by_network_size) #
@@
   filter(depGroupVar != 0)  %>% 
-  mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
+ # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
   group_by(EMPI) %>% 
@@
 names(df_fascicle_dep_mask_volume_overlap_effsize_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_effsize")
+df_fascicle_dep_mask_volume_overlap_effsize_value$incremented_by_1 <- df_fascicle_dep_mask_volume_overlap_effsize_value$volume_in_dep_mask + 1 
+df_fascicle_dep_mask_volume_overlap_effsize_value$log <- log10(df_fascicle_dep_mask_volume_overlap_effsize_value$incremented_by_1)
 
@@
 dep_overlap_x_dep_vs_healthy_vol_effsize <- lm(dep_vs_healthy_effsize ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_effsize_value )
+
+#log
+dep_overlap_x_dep_vs_healthy_logvol_effsize <- lm(dep_vs_healthy_effsize ~ log, data=df_fascicle_dep_mask_volume_overlap_effsize_value)
 
@@
 
-scatter_effsize_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize)) +
+scatter_effsize_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=incremented_by_1, 
+                                                                                                             y=dep_vs_healthy_effsize)) +
   geom_point(size=1) + 
-  geom_smooth(method=lm)+
+  geom_smooth(formula = y ~ x, method=lm)+
   xlab("Volume of Fascicle in Depression Network") + 
   ylab("Effect Size (r)") + 
+  scale_x_continuous(trans = 'log10') + 
   ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", pval_vol)) + 
-                 #round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize )$coefficients[2,4],3))) + 
-#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
   theme(plot.title=element_text(hjust=0.5),
@@
   mutate(sum_fascicle_vol_lost_bottom_tertiles = 
-           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves))) %>%
+           rowSums(across(all_of(dep_network_names_bottom_tertiles_no_cranial_nerves)))) %>%
   mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%
@@
   mutate(sum_fascicle_vol_lost_top_quintile = 
-           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves))) %>%
+           rowSums(across(all_of(dep_network_names_top_quintile_no_cranial_nerves)))) %>%
   mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%
@@
   mutate(sum_fascicle_vol_lost_bottom_quintiles = 
-           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves))) %>%
+           rowSums(across(all_of(dep_network_names_bottom_quintiles_no_cranial_nerves)))) %>%
   mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves)

### 2022-09-16 msdepression/scripts/visualization/vol2surf_mapping.Rmd
Volume to surface mapping from Val

---
title: "Project an MNI volume to the fslr surface for visualization"
author: "Valerie Jill Sydnor, adapted by Erica Baller"
date: "9/16/2022"
output: html_document
---

```{r setup, include=FALSE}
library(ciftiTools)
ciftiTools.setOption('wb_path', '/Users/eballer/Software/workbench/')
#set the path to your installation of connectome workbench as 'wb_path'
library(rgl) #to use ciftiTools graphics
library(rmarkdown)
library(gifti) #to read in your surface giftis 
```

Map volume data to left and right cortical surfaces
```{r}
#left hemisphere
command1=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii -trilinear")
ciftiTools::run_wb_cmd(command1, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)

#right hemisphere
command2=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii -trilinear")
ciftiTools::run_wb_cmd(command2, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)
```

Read in your surface giftis
```{r}
test_l <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii")
test_r <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii")
```

Combine your giftis into one cifti surface (for visualization with ciftiTools)
```{r}
test.ciftimap <- as_cifti(cortexL = test_l$data$normal, cortexR = test_r$data$normal) #the datas are stored in file$data$normal
test.ciftimap$data$cortex_left[test.ciftimap$data$cortex_left > 0] <- 1
test.ciftimap$data$cortex_right[test.ciftimap$data$cortex_right > 0] <- 1

```

You can also save out this cifti for viewing in connectome workbench
```{r}
write_cifti(test.ciftimap, "/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf.dscalar.nii")
```

Choose what surface you want to visualize your data on (midthickness, inflated, sphere, etc.)
```{r}
#lets use midthickness
surfL_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii")
surfR_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii")

#gotta add these surfs to your cifti from the step above
test.ciftimap <- add_surf(test.ciftimap, surfL=surfL_name, surfR=surfR_name)
```

Visualize!
```{r, rgl=TRUE, format="png"} 
view_cifti(test.ciftimap, colors=c("magma"), widget=TRUE)
```

### 2022-09-16 msdepression/scripts/make_binary_colored_depression_net_maps.R
Add files via upload

@@ -18,13 +18,14 @@ dep_network_names_top_quartile_no_cranial_nerves_by_vol <- read.csv(file = paste
 files = c("volume_of_overlap_w_dep_net_n77") #c("fascicle_proportion_overlap_with_depression_network_n77")
 
 color_depressed_net = "255 0 255" #pink
-color_nondepressed_net = "0 255 255" #light blue
+color_nondepressed_net = "220 220 220" #grey
+#color_nondepressed_net = "0 255 255" #light blue
 #color_depressed_net = "255 255 0" #yellow
 #color_nondepressed_net = "0 0 255" #blue
 
 for (filename in files) {
   #suffix, CHANGE IF YOU WANT A DIFFERENT FLAG
-  suffix = "pinkblue"
+  suffix = "pinkgrey"
   
   #read in file with fascicle name and proportion
   fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)

### 2022-09-16 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
 
-[] make figure of depression network from bold
+[x] make figure of depression network from bold
   [x] resampled the harvard map to mni_icbm_2009b template: 
@@
 ```
+
+#### 20220913
+1. [o] Flow chart (text is small, do I need to make bigger?)
+  efile a.[x] demographics full
+  efile b.[x] (promis scores)
+  efile c. [] pipeline is e -file (can only have max 4 pt figure), will need to work with Maddie for this.
+2. Making depression network
+  [x]: Cortical rendering of depression network
+  [x]: DSI fascicle of volume of overlap w/network (how to scale for visualizations) (new hampshire forest in the fall colors)
+3.  In vs outside dep net
+    a. inside vs outside dep net violin
+    b. pink and blue fascicle picture part B
+4. Diagnosis analysis
+    a. no diff by lesion volume (violin)
+    b. but difference in fascicle volume (violin)
+        - just in/outside of network (put stuff about within/outside network for each diagnostic group in text)
+5. Dep Network/dx interactions
+     a. interaction graph
+     b. scatter (effect size by overlap)
+     c. Fascicles colored by effect size (edited)
+
+#### 20220914
+Ted meeting
+
+[] Elizabeth is currently replicating
+  - found a bug in the naming order of one of the files. We have changed it, and the subsequent results are the same, but a little better
+  - The mask stays the same (dep vs non), so no need to regenerate those pictures
+  - Coloring for new hamphire colors and pink effect size are slightly different
+    - [x] redid the new hampshire color plot
+    - [x] redid pink plot
+  - Will need to remake the main effect of dx and network violins with new data
+    [x] main effect of network
+    [x] main effect of diagnosis
+  - Remake scatter and interaction plots
+    [x] scatter
+    [x] interaction
+[x] review some sample figures
+[x] talk about science for LRP
+  - spoke with Mark, feasibility important
+  - In line with that, do I convert aim 1 to 2 aims (have sex differences be one)?
+  - Or (my preference), basically keep everything the same but state that aim 2 will be the beginning of a collection of data (50 people per year), rather than 250. 
+    - would be much eaasier to do this rather than rewrite that whole chunk.
+    - Also, mentorship team from K, but without mike fox and tom. Substitute sex differences for EMR training?
+  - I have also included that I applied for a K already, and what my plans are if I don't get it
+  -
+
+#### 20220915
+  [] cleaning up scatter with Elizabeth - incremented the volume by 1, and log transformed for a much prettier graph!
+  [x] Making the binary brain map on an inflated brain in workbench
+   - first, binarized it 
+```{bash}
+3dcalc -a DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii -expr 'ispositive(a-3.09)' -prefix DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii
+```
+   - tried to use command line tools for workbench, couldn't do it, but val passed on a script that was just amazing, and I adapted it for my stuff
+```{r}
+
+---
+title: "Project an MNI volume to the fslr surface for visualization"
+author: "Valerie Jill Sydnor, adapted by Erica Baller"
+date: "9/16/2022"
+output: html_document
+---
+library(ciftiTools)
+ciftiTools.setOption('wb_path', '/Users/eballer/Software/workbench/')
+#set the path to your installation of connectome workbench as 'wb_path'
+library(rgl) #to use ciftiTools graphics
+library(rmarkdown)
+library(gifti) #to read in your surface giftis
+
+#Map volume data to left and right cortical surfaces
+#left hemisphere
+command1=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii -trilinear")
+ciftiTools::run_wb_cmd(command1, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)
+
+#right hemisphere
+command2=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii -trilinear")
+ciftiTools::run_wb_cmd(command2, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)
+
+#Read in your surface giftis
+test_l <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii")
+test_r <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii")
+
+#Combine your giftis into one cifti surface (for visualization with ciftiTools)
+test.ciftimap <- as_cifti(cortexL = test_l$data$normal, cortexR = test_r$data$normal) #the datas are stored in file$data$normal
+test.ciftimap$data$cortex_left[test.ciftimap$data$cortex_left > 0] <- 1
+test.ciftimap$data$cortex_right[test.ciftimap$data$cortex_right > 0] <- 1
+
+#You can also save out this cifti for viewing in connectome workbench
+write_cifti(test.ciftimap, "/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf.dscalar.nii")
+
+#Choose what surface you want to visualize your data on (midthickness, inflated, sphere, etc.)
+#lets use midthickness
+surfL_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii")
+surfR_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii")
+
+#gotta add these surfs to your cifti from the step above
+test.ciftimap <- add_surf(test.ciftimap, surfL=surfL_name, surfR=surfR_name)
+
+#Visualize!
+#```{r, rgl=TRUE, format="png"} 
+view_cifti(test.ciftimap, colors=c("magma"), widget=TRUE)
+```
+
+- I then had to take this into workbench, and futz with the colors
+
+  - open application: wb_view
+  - open file: /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf.dscalar.nii
+  - open file: /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii
+  - open file: /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii
+  - click montage
+  - under settings, click the wrench
+  - Threshold: Off
+  - Palette
+    - Pick the palette HSB8_clrmid
+    - Invert off
+    - click interpolate colors
+    - click fixed
+    - setting 40/0/0/0
+  - click negative and positive
+  - Click bars under histogram
+  - change bars color and envelope to fuchsia (don't know if this does anything)
+  - Color bar and palette settings: apply to all maps
+
+- Save, screenshot, bring to illustrator and have at it
+
+- Also, will do log scatter, looks so much better. Have to increment by 1 in order to make it pretty though, otherwise all the 0 values will screw you over

### 2022-09-23 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022092022_post_replication.Rmd
Add files via upload

--- 2022-09-16 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220824_for_replication.Rmd
@@
 ---
-title: "MS Depression Paper, started 2022 08 24"
+title: "MS Depression Paper, post replication, started 2022 09 21"
 author: "Erica Baller"
@@
 
-    - Disease within versus outside depression network: p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339, eff size 0.73
+    - Disease within versus outside depression network: p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339, eff size 0.78 (large)
 
@@
 
-    - Volume of MIMoSA lesion burden: p=0.036; depressed = 15,871; nondepressed = 12,886, eff size 0.107
-
-    - Proportion Volume of whole brain fascicle disease burden: W=14995, p=0.04, dep = 0.39, nondep = 0.30
+    - Volume of MIMoSA lesion burden: W = 14982, p=0.036; depressed = 15,871; nondepressed = 12,886, eff size 0.107 (small)
+
+    - Proportion Volume of whole brain fascicle disease burden: W=14996, p=0.04, dep = 0.39, nondep = 0.34, eff size 0.11 (small)
 
@@
 
-    - Main effect of Diagnosis: T = 2.29, p = 0.02 (Depressed patient disease > Nondepressed patient disease)
-
-    - Main effect of Network: T = 16.58, p <2.2 * 10^-16 (Depression Network more disease than nondepression network)
-
-    - Interaction= T = 2.2, p = 0.03; finding driven by worse disease in depressed patients, within depression network
+    - Main effect of Diagnosis: T = 2.4, p = 0.015 (Depressed patient disease > Nondepressed patient disease)
+
+    - Main effect of Network: T = 19.2, p <2 * 10^-16 (Depression Network more disease than nondepression network)
+
+    - Interaction= T = 2.3, p = 0.02; finding driven by worse disease in depressed patients, within depression network
 
@@
 
-    - p before bootstrap: p=1.8*10^-6; effect increases as overlap with depression network increases
-
-    - p after bootstrap: p = < 1*10^-16
+    - T = 2.4, p=0.02, adjusted R^2 = 0.06 (no boot)
 
@@
 
-        - Main effect of network (Depression > Nondepresison network): T=23.2, p = 2*10^-16
-
-        - Main effect of diagnosis (Depressed > Nondepressed people):  T = 2.37, p = 0.018
-
-        - Interaction of network and diagnosis:  T = 2.17, p = 0.031
+        - Main effect of network (Depression > Nondepresison network): T=22.7, p = 2*10^-16
+
+        - Main effect of diagnosis (Depressed > Nondepressed people):  T = 2.4, p = 0.015
+
+        - Interaction of network and diagnosis:  T = 2.2, p = 0.029
 
@@
 
-        - Main effect of network (Depression > Nondepresison network): T=22.56, p = 2*10^-16;
-
-        - Main effect of diagnosis (Depressed > Nondepressed people): T = 2.285, p = 0.019
-
-        - Interaction of network and diagnosis, T = 2.34, p = 0.020
+        - Main effect of network (Depression > Nondepresison network): T=23.0, p = 2*10^-16;
+
+        - Main effect of diagnosis (Depressed > Nondepressed people): T = 2.4, p = 0.048
+
+        - Interaction of network and diagnosis, T = 2.4, p = 0.018
 
@@
 library(rstatix)
+library(effectsize)
 
@@
 
+#save pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/plot_all_fascicle_volumes_dep_network_no_cranial_nerves.pdf"))
+  plot_all_fascicle_volumes_dep_network_no_cranial_nerves
+dev.off()
+
 ### Change names to add suffix
@@
 hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
-dep_lesion_vol_shapiro <- hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
-shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.80, p = 2.2 * 10^-16, NON-NORMAL
-
-dep_vs_healthy_wilcoxon_lesion_volume_results <-  wilcox.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = F) #p=0.036; depressed = 15,871; nondepressed = 12,886
+dep_lesion_vol_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.80, p = 2.2 * 10^-16, NON-NORMAL
+
+dep_vs_healthy_wilcoxon_lesion_volume_results <-  wilcox.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = F) #W = 14982, p=0.036; mean depressed = 15,871; mean nondepressed = 12,886
 
@@
 
+#save pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/violin_lesion_volume.pdf"))
+  violin_lesion_volume
+dev.off()
+
 ```
@@
 
-disease_overall_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum) #W=0.97, p = 3.4 * 10^-7, NON-NORMAL
-
-disease_dep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum) #W=0.97, p = 5.0 * 10^-7, NON-NORMAL
-
-disease_nondep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum) #W=0.97, p = 1.9 * 10^-7, NON-NORMAL
-
-disease_dep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.97, p = 0.0002 NON-NORMAL
+disease_overall_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum) #W=0.97, p = 4.1 * 10^-7, NON-NORMAL
+
+disease_dep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum) #W=0.97, p = 2.0 * 10^-7, NON-NORMAL
+
+disease_nondep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum) #W=0.97, p = 2.6 * 10^-7, NON-NORMAL
+
+disease_dep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.97, p = 0.0003 NON-NORMAL
 
@@
 
-### Within/outside depression network, p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339
+### Within/outside depression network, V = 65897, p = 2.2*10^-16, mean dep net prop = 0.41, mean non dep = 0.31
 msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum,paired = TRUE)
@@
 
-#proportion volume lost over whole network = W=14995, p=0.04, dep = 0.39, nondep = 0.30
+#proportion volume lost over whole network = W=14996, p=0.04, mean dep = 0.39, mean nondep = 0.34
 proportion_volume_lost_over_whole_brain_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
 
-#effsize 0.11
+#effsize 0.11 (small)
 proportion_volume_lost_over_whole_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_per_total_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
 
-#proportion volume lost within depression network, W = 14941, p-value = 0.03; dep = 0.42, nondep = 0.38
+#proportion volume lost within depression network, W = 14969, p-value = 0.04; dep = 0.44, nondep = 0.38
 proportion_volume_lost_dep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
 
-#effsize 0.11
+#effsize 0.11 (small)
 proportion_volume_lost_over_dep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_dep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
 
-#proportion volume lost outside depression network W = 15120, p-value 0.05; dep = 0.36, nondep = 0.33
+#proportion volume lost outside depression network W = 15134, p-value 0.051; mean dep = 0.32, mean nondep = 0.29
 proportion_volume_lost_nondep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
 
-#effsize 0.10
+#effsize 0.10 (0.099) (small)
 proportion_volume_lost_over_nondep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_nondep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
@@
 ### Interaction Network x Diagnosis
-scale=1
+#scale=1
 
@@
 
-#eff size = 0.73
+#eff size = 0.78 (large)
 msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions, formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
 
-#diagnosis: Dep > nondep T = 2.285, p = 0.023; Network dep>nondep = T=16.6, p = 2*10^-16; intx = T = 2.21, p = 0.027
+#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=19.2, p = 2*10^-16; intx = T = 2.3, p = 0.020
 diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
@@
 
+#save pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/violin_within_vs_outside_network_sum_proportion_scaled_by_network_size.pdf"))
+  violin_within_vs_outside_network_sum_proportion_scaled_by_network_size
+dev.off()
+
 #diagnosis
@@
   ylim(0,1) + 
+  scale_fill_manual(values=c("#AB19E6", "#1954E6")) + 
   theme(plot.title = element_text(size=16, hjust = 0.55),
@@
 
+#save pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/violin_dep_vs_nondep_scaled_by_network_size.pdf"))
+  violin_dep_vs_nondep_scaled_by_network_size
+dev.off()
+
 #interaction
+
 diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, 
                                                          type = "int",
-                                                         axis.lim = c(0.2,0.5),
+                                                         axis.lim = c(0,1),
                                                          colors = c("#ff00ff", "#3df2f2"),
-                                                         line.size = 1.3,
+                                                         line.size = 2,
                                                          dodge=0.2,
                                                          axis.title = c("Diagnosis", "Proportion Affected"),
-                                                         title = "Diagnosis x Depression Network Disease")+
-
-  theme(axis.text.x = element_text(size=12),
-        axis.text.y = element_text(size=12),
+                                                         title = "Diagnosis x Depression Network Disease") +
+
+  theme(axis.text.x = element_text(size=16, color="black"),
+        axis.text.y = element_text(size=16, color="black"),
+        axis.title.x = element_text(size =16), 
+        axis.title.y = element_text(size =16), 
+        title = element_text(size=18),
         axis.line = element_line(size = 1.5, colour = "black", linetype=1),
-        aspect.ratio = 3/2,
+        aspect.ratio = 5/3,
         panel.background = element_rect(colour = "white", fill = "white"))
@@
 print(diagnosis_by_network_interaction_lmer_plot)
+
+#save as pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_lmer_range0to1.pdf"))
+  diagnosis_by_network_interaction_lmer_plot
+dev.off()
 
@@
 names(df_fascicle_dep_mask_volume_overlap_effsize_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_effsize")
-df_fascicle_dep_mask_volume_overlap_effsize_value$incremented_by_1 <- df_fascicle_dep_mask_volume_overlap_effsize_value$volume_in_dep_mask + 1 
-df_fascicle_dep_mask_volume_overlap_effsize_value$log <- log10(df_fascicle_dep_mask_volume_overlap_effsize_value$incremented_by_1)
 
 # plot fascicle overlap x Eff size value
-#statistic, p=1.8*10^-6
+#statistic, p=0.02, T = 2.4, eff size (adjusted R^2 = 0.06)
 dep_overlap_x_dep_vs_healthy_vol_effsize <- lm(dep_vs_healthy_effsize ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_effsize_value )
-
-#log
-dep_overlap_x_dep_vs_healthy_logvol_effsize <- lm(dep_vs_healthy_effsize ~ log, data=df_fascicle_dep_mask_volume_overlap_effsize_value)
-
-#### bootstrap
-
-######### functions
-tval <- function(formula, data, indices) {
-  df <- data[indices,] # allows boot to select sample
-  fit <- lm(formula, data=df)
-  return(summary(fit)$coefficients[2,3])
-}
-
-#effect sizes
-# 95% confidence intervals for the correlation
-set.seed(1)
-data_frame_dep_v_nondep_effsize_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_effsize_value, select = c("dep_vs_healthy_effsize", "volume_in_dep_mask"))
-
-results_vol <- boot(data=data_frame_dep_v_nondep_effsize_x_vol_dep_mask, statistic=tval,
-   R=10000, formula=dep_vs_healthy_effsize~volume_in_dep_mask)
-print(results_vol)
-plot(results_vol)
-boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
-print(boot_results_vol.ci)
-pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
-print(pval_vol) #p=1*10^-16; CI (1.774-5.144)
-
-###
-
-scatter_effsize_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=incremented_by_1, 
-                                                                                                             y=dep_vs_healthy_effsize)) +
-  geom_point(size=1) + 
-  geom_smooth(formula = y ~ x, method=lm)+
+summary(dep_overlap_x_dep_vs_healthy_vol_effsize)
+
+scatter_effsize_val_by_dep_net_overlap_log <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize, color=dep_vs_healthy_effsize)) +
+  geom_point(size=2) + 
+  scale_color_gradient(low = "#FFFFFF", high = "#ff00ff") +
+  geom_smooth(formula = y ~ x, method=lm, color="black")+
+  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(0, 10, 100, 1000, 10000, 100000)) + 
   xlab("Volume of Fascicle in Depression Network") + 
   ylab("Effect Size (r)") + 
-  scale_x_continuous(trans = 'log10') + 
-  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", pval_vol)) + 
   theme(plot.title=element_text(hjust=0.5),
-        axis.text = element_text(size=12, color = "black"),
-        axis.title = element_text(size=14, color = "black"),
+        axis.text = element_text(size=16, color = "black"),
+        axis.title = element_text(size=18, color = "black"),
         axis.line = element_line(color="black", size=1.3),
-        panel.background = element_blank())
-print(scatter_effsize_val_by_dep_net_overlap)
+        panel.background = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 4/4)
+print(scatter_effsize_val_by_dep_net_overlap_log)
+
+#save as pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/scatter_effsize_val_by_dep_net_overlap_log.pdf"))
+  scatter_effsize_val_by_dep_net_overlap_log
+dev.off()
 
@@
 ##### within versus outside network
-#p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
+#V=67727, p=2.2*10^16, top tertile prop vol affected = 0.42, bottom 2 tertiles = 0.28
 msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)
 
-#p=2.2*10^-16, top quintile prop vol affected = 0.419, bottom 4 quintiles = 0.322
+#p=2.2*10^-16, top quintile prop vol affected = 0.43, bottom 4 quintiles = 0.30
 msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)
@@
 
-#diagnosis: Dep > nondep T = 2.37, p = 0.018; Network dep>nondep = T=23.2, p = 2*10^-16; intx = T = 2.17, p = 0.031
+#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=22.7, p = 2*10^-16; intx = T = 2.19, p = 0.029
 diagnosis_by_network_interaction_lmer_tertile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_tertiles)
@@
 
-#diagnosis: Dep > nondep T = 2.285, p = 0.019; Network dep>nondep = T=22.56, p = 2*10^-16; intx = T = 2.34, p = 0.020
+#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=23.0, p = 2*10^-16; intx = T = 2.37, p = 0.018
 diagnosis_by_network_interaction_lmer_quintile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_quintiles)

### 2022-09-23 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
 - Also, will do log scatter, looks so much better. Have to increment by 1 in order to make it pretty though, otherwise all the 0 values will screw you over
+
+#### 20220921
+
+[x] replicated!
+[x] Updated the stats in my markdown and changed the name to reflect that it is post replication
+[x] big changes - correcting the volume multiplication error
+[x] Also changed the effect size by volume plot to be scaled logarithmicly
+  [x] Now p value of this regression is 0.0187 (Adjusted R2 0.06).
+  [x] Elizabeth thinks we can drop the Boot. Ted??
+[x] colors for depressed vs nondepressed plots
+[x] figure alternatives (move interaction to figure 4, and have fig 5 just brain + scatter)
+
+#### 20220922
+
+[] no cerebellum -> should I find a way to do a flat representation?
+
+Need to install python and pennlinc toolkit
+
+First installed pip (python3 is what i am using)
+
+1. navigate to your git directory of choice (/Users/eballer/BBL/msdepression/erica_local_git/)
+2. git clone https://github.com/PennLINC/PennLINC-Kit.git 
+3. Activate your python environment of choice [python3]
+4. in the cloned directory: pip install -e . (/Users/eballer/BBL/msdepression/erica_local_git/PennLINC-Kit)
+5. That should do it.. you can then load up your python interpreter and should be able to import pennlinckit, or from pennlinckit import … where … is whatever function or subpackage you want to import
+
+```{bash}
+
+```

### 2022-09-29 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022092022_post_replication.Rmd
visualizations

@@
 library(effectsize)
+library(colormod)
+library(ggpubr)
 
@@
 ### Interaction Network x Diagnosis
-#scale=1
 
@@
 ########## PLOTS #############
+
+#### color scheme
+Diagnosis_colors <- c("#AB19E6", "#1954E6")
+Diagnosis_colors_faded <- as.vector(sapply(Diagnosis_colors, function(color) {decrease_saturation(color, 0.3)})) 
+Network_colors <- c("#ff00ff", "#3df2f2")
+Network_colors_faded <- as.vector(sapply(Network_colors, function(color) {decrease_saturation(color, 0.3)}))
+
+#label scheme
+Network_labels <- c("Depression Network", "Nondepression Network")
+Network_labels_wrapped <- c("Depression\nNetwork", "Nondepression\n Network")
+Diagnosis_labels <- c("Depressed", "Nondepressed")
 
@@
   ylim(0,1) + 
-  scale_fill_manual(values=c("#AB19E6", "#1954E6")) + 
+  scale_fill_manual(values=Diagnosis_colors) + 
   theme(plot.title = element_text(size=16, hjust = 0.55),
@@
 
-#interaction
-
-diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, 
-                                                         type = "int",
-                                                         axis.lim = c(0,1),
-                                                         colors = c("#ff00ff", "#3df2f2"),
-                                                         line.size = 2,
-                                                         dodge=0.2,
-                                                         axis.title = c("Diagnosis", "Proportion Affected"),
-                                                         title = "Diagnosis x Depression Network Disease") +
-
+# bar graph instead of violin
+
+df_proportion_network_volume_and_diagnosis_summary <- df_proportion_network_volume_and_diagnosis %>%
+  group_by(depGroupVar) %>%
+  summarise(
+    mean = mean(proportion_volume_lost_per_total_network_size_sum),
+    sd = sd(proportion_volume_lost_per_total_network_size_sum),
+    n = n(),
+    se = sd / sqrt(n)
+  )
+
+bar_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis_summary, aes(x=depGroupVar, y=mean, fill=depGroupVar)) +
+  geom_bar(stat = "identity", fill=Diagnosis_colors_faded, color = Diagnosis_colors, size = 1.5, width = 0.8) +  
+  geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.2, size=1.5, color = 'black') + #, position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
+  scale_x_discrete(labels=c("Non-depressed" = "Nondepressed", "Depressed" = "Depressed")) + 
+  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
+  ylab("Proportion (sum total volume lost/Net Vol) ") +
+  ylim(0,0.5) + 
+  theme(plot.title = element_text(size=16, hjust = 0.55),
+        axis.text = element_text(size = 16, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=16),
+        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+        legend.position = "none",
+        aspect.ratio=7/5,
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(bar_dep_vs_nondep_scaled_by_network_size) #
+
+#save pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/bar_dep_vs_nondep_scaled_by_network_size.pdf"))
+  bar_dep_vs_nondep_scaled_by_network_size
+dev.off()
+
+#plot interaction with ggplot rather than plot_model
+
+df_interaction <- expand.grid(Network = levels(df_empi_dx_network_proportions$Network), Diagnosis = levels(df_empi_dx_network_proportions$Diagnosis), subject = "8450972724") # just pick a random person
+
+model_data <- get_model_data(model = diagnosis_by_network_interaction_lmer, type = "int")
+
+df_interaction_fit_upper_lower <- cbind(df_interaction, model_data$predicted, model_data$conf.high, model_data$conf.low)
+
+### line
+
+interaction_line <- ggplot(data = df_interaction_fit_upper_lower, aes(x = Network, y = model_data$predicted, color = Diagnosis)) +
+  geom_line(aes(group = Diagnosis), size = 1.5) +
+  scale_color_manual(values=Diagnosis_colors, labels=Diagnosis_labels)+ 
+  geom_point(size = 3) + 
+ # geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), width = 0.2,size = 1, color='black') +
+  scale_x_discrete(labels=Network_labels) +
+  ylim(0,0.5) + 
+    xlab("Diagnosis") + 
+    ylab("Proportion Affected") + 
     theme(axis.text.x = element_text(size=16, color="black"),
           axis.text.y = element_text(size=16, color="black"),
-        axis.title.x = element_text(size =16), 
+          axis.title.x = element_blank(), #element_text(size =16), 
+          axis.title.y = element_text(size =16), 
+          title = element_text(size=18),
+          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+          aspect.ratio = 7/5,
+          panel.background = element_rect(colour = "white", fill = "white"),
+          legend.key=element_blank(),
+          legend.title = element_blank(),
+          legend.text = element_text(size = 14), legend.position = "top")
+
+#save as pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_line_graph_range0to05_point.pdf"))
+  interaction_line
+dev.off()
+
+#geom_point
+interaction_ggplot_point <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
+    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
+    geom_point(position = position_dodge(width=0.5), size = 5) + 
+    scale_color_manual(values=Network_colors, labels = c("Depression Network", "Nondepression Network")) + 
+    ylim(0,1) + 
+    xlab("Diagnosis") + 
+    ylab("Proportion Affected") + 
+    theme(axis.text.x = element_text(size=16, color="black"),
+          axis.text.y = element_text(size=16, color="black"),
+          axis.title.x = element_blank(), #element_text(size =16), 
           axis.title.y = element_text(size =16), 
@@
           aspect.ratio = 5/3,
-        panel.background = element_rect(colour = "white", fill = "white"))
-
-print(diagnosis_by_network_interaction_lmer_plot)
+          panel.background = element_rect(colour = "white", fill = "white"),
+          legend.key=element_blank(),
+          legend.title = element_blank(),
+          legend.text = element_text(size = 14), #,
+          legend.position = "top") #+ 
+ #   ggtitle("Diagnosis x Depression Network Disease")
+print(interaction_ggplot_point)
 
 #save as pdf
-pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_lmer_range0to1.pdf"))
-  diagnosis_by_network_interaction_lmer_plot
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to1_point.pdf"))
+  interaction_ggplot_point
 dev.off()
 
+#geom_point 0.2-0.5
+interaction_ggplot_point_02_05 <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
+    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
+    geom_point(position = position_dodge(width=0.5), size = 5) + 
+    scale_color_manual(values=Network_colors, labels = c("Depression Network", "Nondepression Network")) + 
+    ylim(0.2,0.5) + 
+    xlab("Diagnosis") + 
+    ylab("Proportion Affected") + 
+    theme(axis.text.x = element_text(size=16, color="black"),
+          axis.text.y = element_text(size=16, color="black"),
+          axis.title.x = element_blank(), #element_text(size =16), 
+          axis.title.y = element_text(size =16), 
+          title = element_text(size=18),
+          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+          aspect.ratio = 5/3,
+          panel.background = element_rect(colour = "white", fill = "white"),
+          legend.key=element_blank(),
+          legend.title = element_blank(),
+          legend.text = element_text(size = 14), #,
+          legend.position = "top") #+ 
+ #   ggtitle("Diagnosis x Depression Network Disease")
+print(interaction_ggplot_point_02_05)
+
+#save as pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0_2to0_5_point.pdf"))
+  interaction_ggplot_point_02_05
+dev.off()
+
+#geom_bar
+interaction_ggplot_bar <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network,  group = interaction(Diagnosis, Network))) + 
+    geom_col(position = position_dodge(width=0.6), fill = c(Network_colors_faded, Network_colors_faded), color = c(Network_colors, Network_colors), width = 0.5, size = 1.5) + 
+    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.6), width = 0.2,size = 1.5, color='black') +
+    #geom_point(position = position_dodge(width=0.5), size = 5) + 
+    scale_color_manual(values=Network_colors) + #, labels = c("Depression Network", "Nondepression Network"))+ 
+    ylim(0,0.5) + 
+    xlab("Diagnosis") + 
+    ylab("Proportion Affected") + 
+    theme(axis.text.x = element_text(size=16, color="black"),
+          axis.text.y = element_text(size=16, color="black"),
+          axis.title.x = element_blank(), #element_text(size =16), 
+          axis.title.y = element_text(size =16), 
+          plot.title = element_text(size=16, hjust = 0.5),
+          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
+          aspect.ratio = 7/5,
+          panel.background = element_rect(colour = "white", fill = "white"),
+          legend.key=element_blank(),
+          legend.title = element_blank(),
+          legend.text = element_text(size = 14),
+          legend.position = "top")+
+
+    ggtitle("Diagnosis x Depression Network Disease")
+print(interaction_ggplot_bar)
+
+#save as pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to05_bar.pdf"))
+  interaction_ggplot_bar
+dev.off()
+
+### wrap
+
+diagnosis_and_interaction_bar_combined <- ggarrange(bar_dep_vs_nondep_scaled_by_network_size,interaction_ggplot_bar, 
+                                                labels = c("A", "B"),
+                                                ncol = 2, nrow = 1)
+
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to05.pdf"))
+  diagnosis_and_interaction_bar_combined
+
+diagnosis_and_interaction_all_combined <- ggarrange(bar_dep_vs_nondep_scaled_by_network_size, interaction_line,interaction_ggplot_bar, 
+                                                labels = c("A", "B", "C"),
+                                                ncol = 3, nrow = 1)
+
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to05.pdf"))
+  diagnosis_and_interaction_all_combined
+dev.off()
 ```
+
 ### Effect size regression (effect size x overlap with dep net)

### 2022-09-29 lab_notebook/lab_notebook.Rmd
Add files via upload

--- 2022-09-16 lab_notebook/lab_notebook.Rmd
@@
 - Also, will do log scatter, looks so much better. Have to increment by 1 in order to make it pretty though, otherwise all the 0 values will screw you over
+
+#### 20220921
+
+[x] replicated!
+[x] Updated the stats in my markdown and changed the name to reflect that it is post replication
+[x] big changes - correcting the volume multiplication error
+[x] Also changed the effect size by volume plot to be scaled logarithmicly
+  [x] Now p value of this regression is 0.0187 (Adjusted R2 0.06).
+  [x] Elizabeth thinks we can drop the Boot. Ted??
+[x] colors for depressed vs nondepressed plots
+[x] figure alternatives (move interaction to figure 4, and have fig 5 just brain + scatter)
+
+#### 20220922
+
+[x] no cerebellum -> should I find a way to do a flat representation? (NO)
+
+Need to install python and pennlinc toolkit - did not work. It is broken. Don't use it
+
+#### 20220930 or next week
+[] add post hoc */bars to figure 4
+[] add legend
+[] call it a win!
+[] add statistics to other graphs (p/effect size)

### 2022-09-29 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022092022_post_replication.Rmd
Add files via upload

@@
 
+#within nondepressed group, V=9661, p <2.2*10^-16
+msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1],paired = TRUE)
+
+#within depressed group, V= 25143, p<2.2*10^-16
+msdisease_within_versus_outside_depression_network_dep_group_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2],paired = TRUE)
+
 ### Diagnostic effects
@@
 
-#eff size = 0.78 (large)
+#effect size within vs outside dep network, eff size = 0.78 (large)
 msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions, formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+
+#effect size dep vs nondep network within nondepressed, eff size = 0.75 (large)
+msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Nondepressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+
+#effect size dep vs dep network within depressed, eff size = 0.80, large
+msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Depressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
 

### 2022-10-06 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022092022_post_replication.Rmd
Add files via upload

@@
 library(ggpubr)
+library(ggsignif)
 
@@
 
-#goal is to keep people who were seen by 
  data_empi_acc_f_phq <- data %>% 
@@
             sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
-   ungroup()
+   ungroup() #n=3,737 unique people, n= 16,830 total
 
@@
 
-fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
+fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336, n unique = 958
+
+#change empi to factor
+
+fascicle_proportions$EMPI = as.factor(fascicle_proportions$EMPI)
 
@@
 
-lesion_volumes <- read.csv(paste0(homedir, "/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE)
+lesion_volumes <- read.csv(paste0(homedir, "/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE) #n=958 unique, identical to fascicle unique
+
+lesion_volumes$EMPI <- as.factor(lesion_volumes$EMPI)
 
@@
 #merge whole data set with fascicle proportions
-df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962
+df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962, unique = 783
 
@@
 ```
+
+```{r get_mimosa_QC_info}
+
+#This section checks how many unique individuals had QA data, and then how many met MIMoSA QC. Some additional people lost at registration.
+
+#1 Total unique individuals who also have MS - 890
+#2 Total unique who had good MIMoSA - 786
+#3 Total unique with bad MIMoSA - 890-786 = 104
+#4 Total lost in streamline filtering - 3
+#5 Final N going into inclusion/exclusion - 783
+
+#read in mimosa data
+mimosa_qc <- read.csv(paste0(homedir, "/data/melissa_martin_files/csv/mimosa_dataframe"), sep = ",", header = FALSE) #n=2840
+
+#preprocess data frame to extract dates, empis, and score
+empis_with_qc <- mimosa_qc %>% 
+  mutate(EMPI = as.factor(gsub(x = V5, pattern = ".*sub-(.*)/ses-.*/.*", replacement="\\1", perl = TRUE))) %>%
+  mutate(EXAM_DATE = gsub(x=V5, pattern = ".*/ses-(.{8})/.*", replacement = "\\1\\2\\3")) %>% 
+  mutate(score = V3)
+
+#merge with whole data frame to make sure we exclude people w/out MS
+mimosa_qc_with_data_empi_acc_f_phq <- merge(data_empi_acc_f_phq, empis_with_qc, by = c("EMPI", "EXAM_DATE")) #n = 3592
+
+#get list of unique patients, n = 890
+mimosa_qc_with_data_empi_acc_f_phq_unique <- mimosa_qc_with_data_empi_acc_f_phq %>%
+  group_by(EMPI) %>%
+  arrange(EXAM_DATE) %>%
+  slice(1) %>%
+  ungroup #n=890
+
+#get list of unique patients with good QC, n = 786
+mimosa_unique_score_gre_75 <- mimosa_qc_with_data_empi_acc_f_phq %>%
+  filter(score >= 75) %>%
+  group_by(EMPI) %>%
+  arrange(EXAM_DATE) %>%
+  slice(1) %>%
+  ungroup #n=786
+
+#merge with fascicle proportions, we lose 3
+mimosa_qc_with_data_empi_acc_f_phq_fascicle_props <- merge(mimosa_qc_with_data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n=2962
+
+mimosa_qc_with_data_empi_acc_f_phq_fascicle_props_unique <- mimosa_qc_with_data_empi_acc_f_phq_fascicle_props %>% 
+  group_by(EMPI) %>%
+  arrange(EXAM_DATE) %>%
+  slice(1) %>%
+  ungroup #n=783
+```
+
 ### PROMIS DATA HANDLING to link a person to their most proximal PROMIS
@@
 
-df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles %>% 
+df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles_prop_and_vol %>% 
   filter(depGroupVar != 0) %>%
@@
   scale_x_discrete(labels=c("Non-depressed" = "Nondepressed", "Depressed" = "Depressed")) + 
-  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
-  ylab("Proportion (sum total volume lost/Net Vol) ") +
-  ylim(0,0.5) + 
-  theme(plot.title = element_text(size=16, hjust = 0.55),
-        axis.text = element_text(size = 16, colour = "black"), axis.title.x = element_blank(), 
-        axis.title.y = element_text(size=16),
+#  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
+  ylab("Proportion of Whole Brain White Matter Volume Affected") +
+  ylim(0,0.6) + 
+  theme(plot.title = element_text(size=14, hjust = 0.55),
+        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
+        axis.title.y = element_text(size=14),
         axis.line = element_line(size = 1.5, colour = "black", linetype=1),
@@
   scale_x_discrete(labels=Network_labels) +
-  ylim(0,0.5) + 
+  ylim(0,0.6) + 
     xlab("Diagnosis") + 
@@
 #save as pdf
-pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_line_graph_range0to05_point.pdf"))
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_line_graph_range0to06_point.pdf"))
   interaction_line
@@
 #geom_point 0.2-0.5
-interaction_ggplot_point_02_05 <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
+interaction_ggplot_point_02_06 <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
     geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
@@
     scale_color_manual(values=Network_colors, labels = c("Depression Network", "Nondepression Network")) + 
-    ylim(0.2,0.5) + 
+    ylim(0.2,0.6) + 
     xlab("Diagnosis") + 
@@
  #   ggtitle("Diagnosis x Depression Network Disease")
-print(interaction_ggplot_point_02_05)
+print(interaction_ggplot_point_02_06)
 
 #save as pdf
-pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0_2to0_5_point.pdf"))
-  interaction_ggplot_point_02_05
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0_2to0_6_point.pdf"))
+  interaction_ggplot_point_02_06
 dev.off()
@@
     geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.6), width = 0.2,size = 1.5, color='black') +
-    #geom_point(position = position_dodge(width=0.5), size = 5) + 
-    scale_color_manual(values=Network_colors) + #, labels = c("Depression Network", "Nondepression Network"))+ 
-    ylim(0,0.5) + 
+    scale_fill_manual(values=Network_colors) + #, labels = c("Depression Network", "Nondepression Network"))+ 
+    ylim(0,0.6) + 
     xlab("Diagnosis") + 
-    ylab("Proportion Affected") + 
-    theme(axis.text.x = element_text(size=16, color="black"),
-          axis.text.y = element_text(size=16, color="black"),
+    ylab("Proportion of White Matter Network Affected") + 
+    theme(axis.text.x = element_text(size=14, color="black"),
+          axis.text.y = element_text(size=14, color="black"),
           axis.title.x = element_blank(), #element_text(size =16), 
-          axis.title.y = element_text(size =16), 
+          axis.title.y = element_text(size =14), 
           plot.title = element_text(size=16, hjust = 0.5),
@@
           aspect.ratio = 7/5,
-          panel.background = element_rect(colour = "white", fill = "white"),
-          legend.key=element_blank(),
-          legend.title = element_blank(),
-          legend.text = element_text(size = 14),
-          legend.position = "top")+
-
-    ggtitle("Diagnosis x Depression Network Disease")
+          panel.background = element_rect(colour = "white", fill = "white"))+
+
+    geom_signif(
+      y_position = c(0.48, 0.48, 0.53, 0.58), 
+      xmin = c(0.85, 1.85, 0.85, 1.15),
+      xmax = c(1.15, 2.15, 1.85, 2.15),
+      annotation = c("***", "***", "*", "NS"), 
+      tip_length = 0.03, color = "black",
+      size = 1.1, textsize = 5) #+
+  #  ggtitle("Diagnosis x Depression Network Disease")
 print(interaction_ggplot_bar)
@@
 #save as pdf
-pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to05_bar.pdf"))
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to06_bar.pdf"))
   interaction_ggplot_bar
@@
 
-pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to05.pdf"))
+print(diagnosis_and_interaction_bar_combined)
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_and_intx_bargraph_range0to06.pdf"))
   diagnosis_and_interaction_bar_combined
@@
 
-pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to05.pdf"))
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to06.pdf"))
   diagnosis_and_interaction_all_combined

### 2022-11-04 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022110222_post_replication_playing_w_graphs.Rmd
changed names to MS+Dep/MS-Dep

--- 2022-10-06 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022092022_post_replication.Rmd
@@
   filter(depGroupVar != 0) %>% 
-  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
-
-violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
+  mutate(depGroupVar = ifelse(depGroupVar == 1, "MS-Depression", "MS+Depression"))
+
+violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=factor(depGroupVar, level = c('MS+Depression', 'MS-Depression')), y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) +
   geom_violin() + 
   geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
-  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
+ # scale_x_discrete(labels=c("1" = "MS+Depression", "2" = "MS-Depression")) + 
   ggtitle(paste0("Volume of Lesions by Diagnosis")) +
@@
   mutate(subject = as.factor(EMPI)) %>%
-  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
   mutate(Diagnosis = as.factor(Diagnosis)) %>%
@@
 #effect size dep vs nondep network within nondepressed, eff size = 0.75 (large)
-msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Nondepressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "MS-Depression"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
 
 #effect size dep vs dep network within depressed, eff size = 0.80, large
-msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Depressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
+msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "MS+Depression"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
 
@@
 Network_labels_wrapped <- c("Depression\nNetwork", "Nondepression\n Network")
-Diagnosis_labels <- c("Depressed", "Nondepressed")
+#Diagnosis_labels <- c("MS+Depression", "MS-Depression")
+Diagnosis_labels <- c("MS+Depression", "MS-Depression")
 
@@
 dplyr::select(proportion_volume_lost_per_total_network_size_sum, depGroupVar) %>%
-  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))
-
-violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
+   mutate(depGroupVar = ifelse(depGroupVar == 1, "MS-Depression", "MS+Depression"))
+
+ # mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "MS+Depression"))
+
+violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=factor(depGroupVar, level = c('MS+Depression', 'MS-Depression')), y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
   geom_violin() +  
   geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
-  scale_x_discrete(labels=c("1" = "NonDepressed", "2" = "Depressed")) + 
+  #scale_x_discrete(labels=c("1" = "MS-Depression", "2" = "MS+Depression")) +
+
+#  scale_x_discrete(labels=c("1" = "MS-Depression", "2" = "MS+Depression")) + 
   ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
@@
 
+#factor(depGroupVar, level = c('MS+Depression', 'MS-Depression'))
 bar_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis_summary, aes(x=depGroupVar, y=mean, fill=depGroupVar)) +
@@
   geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.2, size=1.5, color = 'black') + #, position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
-  scale_x_discrete(labels=c("Non-depressed" = "Nondepressed", "Depressed" = "Depressed")) + 
+
+  #scale_x_discrete(labels=c("Non-depressed" = "MS-Depression", "MS+Depression" = "MS+Depression")) + 
 #  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
@@
   ylim(0,0.6) + 
-  theme(plot.title = element_text(size=14, hjust = 0.55),
+  theme(plot.title = element_text(size=12, hjust = 0.55),
         axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
@@
 #geom_bar
+
+#factor(Diagnosis, level = c('MS+Depression', 'MS-Depression'))
 interaction_ggplot_bar <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network,  group = interaction(Diagnosis, Network))) + 
@@
     geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.6), width = 0.2,size = 1.5, color='black') +
-    scale_fill_manual(values=Network_colors) + #, labels = c("Depression Network", "Nondepression Network"))+ 
+    scale_fill_manual(values=Network_colors) + #, labels = c("MS+Depression", "MS-Depression"))+ 
     ylim(0,0.6) + 
@@
       size = 1.1, textsize = 5) #+
-  #  ggtitle("Diagnosis x Depression Network Disease")
 print(interaction_ggplot_bar)
@@
         legend.position = "none",
-        aspect.ratio = 4/4)
+        aspect.ratio = 4/4) + 
+  annotate("text", x=1.5, y=0.14, size = 5, 
+           label = paste0("P = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize)$coef[2,4],3)))
+
 print(scatter_effsize_val_by_dep_net_overlap_log)

### 2022-11-05 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022110222_post_replication_playing_w_graphs.Rmd
changing wrap graph

@@
                                                 labels = c("A", "B"),
+                                                font.label = list(size = 16, color = "black", family = NULL),
+                                                vjust = 9,
                                                 ncol = 2, nrow = 1)

### 2023-03-01 msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
Add files via upload

--- 2022-11-04 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022110222_post_replication_playing_w_graphs.Rmd
@@
 
+#### Network, Diagnosis, and Network * Diagnosis
+
+Model: diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
+
 ####  *Network effects:*
@@
 is MS random, or do MS lesions happen to strike fascicles that support the depression network, irrespective of diagnosis).
-To assess this, I took the following steps.
-
-    1) Tests for normality (Shapiro–Wilk test) of proportion of disease in depression network and nondepression network. Both not normally distributed
-
-    2) Used wilcoxon tests to compare disease (paired = T) between depression and nondepression network.
-As each individual contributed one measure for each network (i.e.
-2 measures per person), paired was used.
+To assess this, we looked at the main effect of network in the linear mixed model.
+Effect sizes were calculated with confint.
 
@@
 We next wanted to assess whether patients with depression had more overall disease than patients without depression.
-
-    First, we assessed whether lesion burden (i.e. the volume occupied by white matter disease), irrespective of which fascicles were impacted, differed between patients and controls.
+To assess this, we looked at the main effect of diagnosis in the linear mixed model.
+Effect sizes were calculated with confint.
+
+    We also assessed whether lesion burden (i.e. the volume occupied by white matter disease), irrespective of which fascicles were impacted, differed between patients and controls.
 
@@
 
-  Next, we assessed whether fascicle burden (i.e. total volume of disease divided by total volume of all fascicles) differed between diagnostic groups.
-
-    1) Tests for normality (Shapiro–Wilk test), fascicle burden not normally distributed.
-
-    2) Used wilcoxon tests to compare fascicle burden between depressed and non-depressed patients. As each person only contributed one measure, paired = F.
+#### *NetworkxDiagnosis effects: *
 
@@
     - Disease within versus outside depression network: p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339, eff size 0.78 (large)
+     - Main effect of Network: T = 12.4, p <2 * 10^-16 (Depression Network more disease than nondepression network)
+
+     - Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.076-0.104, P = 2*10^-16)
 
@@
 
-    - Proportion Volume of whole brain fascicle disease burden: W=14996, p=0.04, dep = 0.39, nondep = 0.34, eff size 0.11 (small)
+    - Main effect of Diagnosis: T = 2.4, p = 0.015 (Depressed patient disease > Nondepressed patient disease)
+
+    - Main effect of diagnosis MS+Depression > MS-Depression (Estimate = 0.056, 95% CI 0.011-0.100, P= 0.015)
 
@@
 
-    - Main effect of Diagnosis: T = 2.4, p = 0.015 (Depressed patient disease > Nondepressed patient disease)
-
-    - Main effect of Network: T = 19.2, p <2 * 10^-16 (Depression Network more disease than nondepression network)
-
     - Interaction= T = 2.3, p = 0.02; finding driven by worse disease in depressed patients, within depression network
 
+    - Main effect of interaction (Estimate = 0.022, 95% CI 0.003 - 0.040, P=0.020)
+
 #### Effect Size x Fascicle overlap with depression network scatter
@@
 
+        - Network (.122, CI = 0.107-0.138, P < 2*10^-16)
+
+        - Diagnosis (0.055, CI = 0.011-0.100, P = 0.015)
+
+        - interaction (0.022, CI = 0.002-0.042, P = 0.029)
+
       2) Quintiles
@@
 
-        - Main effect of diagnosis (Depressed > Nondepressed people): T = 2.4, p = 0.048
+        - Main effect of diagnosis (Depressed > Nondepressed people): T = 2.4, p = 0.0148)
 
         - Interaction of network and diagnosis, T = 2.4, p = 0.018
+
+        - Network (.110, CI = 0.10-0.12, P < 2*10^-16)
+
+        - Diagnosis (0.057, CI = 0.011-0.102, P = 0.0148)
+
+        - interaction (0.022, CI = 0.004-0.040, P = 0.018)
 
@@
 library(rstatix)
-library(effectsize)
+#library(effectsize)
 library(colormod)
@@
 
-dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000) #eff size 0.107 (small)
+#dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000) #eff size 0.107 (small)
+
+dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, nboot = 1000) #eff size 0.107 (small)
 
@@
 
-##### Assessing Normality w/histograms and Shapiro_Wilk test for normality ####
-
-### Test for normality within each network, within each diagnosis
-
-disease_overall_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum) #W=0.97, p = 4.1 * 10^-7, NON-NORMAL
-
-disease_dep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum) #W=0.97, p = 2.0 * 10^-7, NON-NORMAL
-
-disease_nondep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum) #W=0.97, p = 2.6 * 10^-7, NON-NORMAL
-
-disease_dep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.97, p = 0.0003 NON-NORMAL
-
-disease_nondep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]) #W=0.96, p = 0.0001 NON-NORMAL
-
-## All are non-normal, will use wilcoxon for comparisons
-
-### Within/outside depression network, V = 65897, p = 2.2*10^-16, mean dep net prop = 0.41, mean non dep = 0.31
-msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum,paired = TRUE)
-
-#within nondepressed group, V=9661, p <2.2*10^-16
-msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1],paired = TRUE)
-
-#within depressed group, V= 25143, p<2.2*10^-16
-msdisease_within_versus_outside_depression_network_dep_group_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2],paired = TRUE)
-
-### Diagnostic effects
-
-#proportion volume lost over whole network = W=14996, p=0.04, mean dep = 0.39, mean nondep = 0.34
-proportion_volume_lost_over_whole_brain_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
-
-#effsize 0.11 (small)
-proportion_volume_lost_over_whole_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_per_total_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
-
-#proportion volume lost within depression network, W = 14969, p-value = 0.04; dep = 0.44, nondep = 0.38
-proportion_volume_lost_dep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
-
-#effsize 0.11 (small)
-proportion_volume_lost_over_dep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_dep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
-
-#proportion volume lost outside depression network W = 15134, p-value 0.051; mean dep = 0.32, mean nondep = 0.29
-proportion_volume_lost_nondep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)
-
-#effsize 0.10 (0.099) (small)
-proportion_volume_lost_over_nondep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_nondep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)
-
 ### Interaction Network x Diagnosis
 
+#clean up data frame
 df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
@@
 
-#effect size within vs outside dep network, eff size = 0.78 (large)
-msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions, formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
-
-#effect size dep vs nondep network within nondepressed, eff size = 0.75 (large)
-msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "MS-Depression"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
-
-#effect size dep vs dep network within depressed, eff size = 0.80, large
-msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "MS+Depression"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)
-
-#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=19.2, p = 2*10^-16; intx = T = 2.3, p = 0.020
+#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020
 diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
 summary(diagnosis_by_network_interaction_lmer)
+
+#Diagnosis: MS+ > MS- 0.011-0.1; Network dep>nondep -.104--.076; intx: -0.040--0.003
+confint(diagnosis_by_network_interaction_lmer)
+
+#Final data: 
+#Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.076-0.104, P = 2*10^-16)
+#Main effecet of diagnosis MS+Depression > MS-Depression (Estimate = 0.056, 95% CI 0.011-0.100, P= 0.015)
+#Main effect of interaction (Estimate = 0.022, 95% CI 0.003 - 0.040, P=0.020)
 
@@
 #  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
-  ylab("Proportion of Whole Brain White Matter Volume Affected") +
+  ylab("Proportion of Whole Brain White Matter Affected") +
   ylim(0,0.6) + 
@@
           aspect.ratio = 7/5,
-          panel.background = element_rect(colour = "white", fill = "white"))+
+          panel.background = element_rect(colour = "white", fill = "white")) +
 
@@
                                                 labels = c("A", "B"),
+                                                font.label = list(size = 16, color = "black", family = NULL),
+                                                vjust = 7,
                                                 ncol = 2, nrow = 1)
@@
 print(diagnosis_and_interaction_bar_combined)
-pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_and_intx_bargraph_range0to06.pdf"))
-  diagnosis_and_interaction_bar_combined
+ggsave(plot = diagnosis_and_interaction_bar_combined, filename =paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_and_intx_bargraph_range0to06.pdf"), height = 8, width = 9)
+#pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_and_intx_bargraph_range0to06.pdf"))
+ # diagnosis_and_interaction_bar_combined
+#dev.off()
 
@@
         aspect.ratio = 4/4) + 
-  annotate("text", x=1.5, y=0.14, size = 5, 
-           label = paste0("P = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize)$coef[2,4],3)))
+  annotate("text", x=1.6, y=0.14, size = 5, 
+           label = paste0("P = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize)$coef[2,4],2)))
 
@@
   mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves)
-
-#### tertile/quintile
-
-##### within versus outside network
-#V=67727, p=2.2*10^16, top tertile prop vol affected = 0.42, bottom 2 tertiles = 0.28
-msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)
-
-#p=2.2*10^-16, top quintile prop vol affected = 0.43, bottom 4 quintiles = 0.30
-msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)
 
@@
 
+confint(diagnosis_by_network_interaction_lmer_tertile)
+
+#Final for tertiles
+#Network (.122, CI = 0.107-0.138, P < 2*10^-16)
+#Diagnosis (0.055, CI = 0.011-0.100, P = 0.015)
+#interaction (0.022, CI = 0.002-0.042, P = 0.029)
+
 #### quintile
@@
 
+confint(diagnosis_by_network_interaction_lmer_quintile)
+
+#Final for quintile
+#Network (.110, CI = 0.10-0.12, P < 2*10^-16)
+#Diagnosis (0.057, CI = 0.011-0.102, P = 0.0148)
+#interaction (0.022, CI = 0.004-0.040, P = 0.018)
+
 ```

### 2023-03-03 get_volume_of_mimosa_lesions.sh
Add files via upload

@@ -1,0 +1,61 @@
+#!/bin/bash
+
+#########     Welcome to  get_volume_of_mimosa_lesions.sh !  ##########3
+## The purpose of this script is to extract the overall volume of lesion burden for each mimosa mask. 
+## The goal is to use the output to answer the question: Does overall lesion burden (quantified by volume of lesions), irrespective of location, relate to depression diagnosis?
+
+### Pre: File containing full paths to the mimosa binary masks:  /project/msdepression/data/melissa_martin_files/csv/mimosa_binary_masks_hcp_space_20211026_n2336
+### Post: File containing ACCESSION_NUM, EXAM_DATA, volume_of_mimosa_lesions
+### Uses: 	1) Reads in the file with full paths
+###		2) Extracts the accession number and exam date
+###		3) Performs 3dmaskave -quiet -mask SELF -sum ${binary_map} and writes to a file
+### Dependencies: 1) Requires afni, so you need to xbash, and module load afni<tab complete>
+
+### Opening Script: If nothing is entered, let person know we are using default mimosa path file, otherwise use file on command line
+
+module load afni_openmp/20.1
+
+echo "Greetings and welcome to the get_volume_of_mimosa_lesions script"
+echo "This script takes a file with full paths to mimosa binary maps, and returns a file that has the empi, exam data, and overall volume of lesions in that map"
+#set default path
+default_mimosa_path="/project/msdepression/data/melissa_martin_files/csv/mimosa_binary_masks_hcp_space_20211026_n2336"
+
+echo "Default mimosa path $default_mimosa_path"
+#set output directory for the volumes file
+output_direc="/project/msdepression/results/"
+
+if [ $# == 0 ]
+then
+    echo "We will use the default mimosa files path" $default_mimosa_path
+    mimosa_files_path=$default_mimosa_path
+else  
+    mimosa_files_path=$1
+fi
+
+echo "Mimosa files path" $mimosa_files_path
+#### initiate new file
+
+#append "_volumes" to the end of the path
+suffix=$(echo ${mimosa_files_path} | perl -pe s'/.*\/(.*)/$1/g')
+output_csv=$output_direc/$suffix"_volumes"
+
+#get rid of old file if it exists, write new one and put in headers
+rm -f $output_csv
+touch $output_csv
[…]

### 2023-03-03 make_binary_colored_depression_net_maps.R
Add files via upload

@@ -1,0 +1,42 @@
+#### Welcome to make_binary_colored_depression_net_maps! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle according to whether or not it is in the depression map. Yellow is dep map, blue is non dep map (though this is flexible)
+
+### Pre: File of fascicle overlap with depression network proportions
+### Post: Output of RGB values in colors
+### Uses: Opens up proportions, sorts by overlap to get a name of a vector in top quartile
+###       iterate through each fascicle, if in the top quartile, color will be first color, if not, will be second color
+###       **** If you change colors!!! CHANGE SUFFIX***
+### Dependencies: Any R will do, need the names of the fibers in the depression network, and a file with proportions
+
+
+homedir="/Users/eballer/BBL/msdepression/"
+
+dep_network_names_top_quartile_no_cranial_nerves_by_vol <- read.csv(file = paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), header = F)
+
+files = c("volume_of_overlap_w_dep_net_n77") #c("fascicle_proportion_overlap_with_depression_network_n77")
+
+color_depressed_net = "255 0 255" #pink
+color_nondepressed_net = "220 220 220" #grey
+#color_nondepressed_net = "0 255 255" #light blue
+#color_depressed_net = "255 255 0" #yellow
+#color_nondepressed_net = "0 0 255" #blue
+
+for (filename in files) {
+  #suffix, CHANGE IF YOU WANT A DIFFERENT FLAG
+  suffix = "pinkgrey"
+  
+  #read in file with fascicle name and proportion
+  fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
+  names(fascicle_name_and_prop) <- c("fascicle", "proportion")
+  
+  #output color file (77 lines, each represents a color), and color name vector. Not sure how to do 1:1 assignments rather than manually, but I'll see if I can figure that out!
+  fascicle_name_and_prop_rgb_only <- fascicle_name_and_prop %>%
+    mutate(rgb = ifelse(fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol$V1, color_depressed_net, color_nondepressed_net)) %>%
+    dplyr::select(fascicle, rgb)
+  
[…]

### 2023-03-03 make_binary_colored_depression_net_maps_by_dx.R

@@ -1,0 +1,90 @@
+#### Welcome to make_binary_colored_depression_net_maps! ####
+
+### I developed this script because I wanted to make my own RGB color palettes for DSI studio. 
+### In short, I want to color each fascicle according to whether or not it is in the depression map. I want to make a separate map for depressed and nondepressed patients, but put them on the same scale
+
+### Pre: File with 4 rows, 3 columns, diagnosis (1 no dep, 2 dep), network (1 dep, 2 nondep) and mean of the proportion of overlap (calculated by summing volumes in individual subjects, dividing by network size)
+
+### Post: Two output of RGB files that can be fed to dsi studio
+### Uses: Opens up file, determines max saturation of pink to be max in dep net, and max blue to be non-dep net
+###       - Makes two color files, depression file where pink/blue triplets are weighted based on previous calculation
+###       - same for nondepressed patients
+###       - hope is that by doing the colors this way, you can see the contrast between groups, where there is a bigger spread of color in depressed, vs nondep, sorts by overlap to get a name of a vector in top quartile
+### Dependencies: Any R will do, need the names of the fibers in the depression network, and a file with proportions
+
+
+homedir="/Users/eballer/BBL/msdepression/"
+
+dep_network_names_top_quartile_no_cranial_nerves_by_vol <- read.csv(file = paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), header = F)
+
+files = c("fascicle_proportion_overlap_with_depression_network_n77")
+
+fascicle_name_and_prop <- read.csv(paste0(homedir, "/results/", filename, ".csv"), sep = ",", header = F)
+names(fascicle_name_and_prop) <- c("fascicle", "prop")
+
+
+df_means_volume_scaled_by_network_size<- read.csv(file = paste0(homedir, "/results/df_means_of_volumes_scaled_by_network_size_dx_net_for_color.csv"), header = T)
+
+df_means_volume_scaled_by_network_size$mean <- df_means_volume_scaled_by_network_size$mean * 100 #easier to scale colors this way
+
[…]

### 2023-03-03 afni_commands_to_construct_yeo_indiv_vol_masks.sh
Delete afni_commands_to_construct_yeo_indiv_vol_masks.sh

@@ -1,54 +1,0 @@
-#!/bin/sh
-
-# pre: Yeo cortical parcellations (7 network) in volumetric space: https://surfer.nmr.mgh.harvard.edu/fswiki/CorticalParcellation_Yeo2011
-# post: 7 individual network maps: VIS, MOT, DA, VA, LIM, FP, DM, labeled yeo_7_vol_[network abbreviation]
-# uses: This script takes the parcellations in volume (Yeo2011_7Networks_MNI152_FreeSurferConformed1mm_LiberalMask.nii) and makes 7 individual maps. In the Yeo2011 volume mask, each network is given a number (1-7), with 1 corresponding to VIS, and 7 corresponding to DM. In order to make individual maps, I need to separate out each network in the map based on intensity. Will try to code in the following way: 1) make a vector where each item is the name of the network 2) make a loop that iterates 1-7, have a second counter that goes in the reverse direction 3) For each iteration, the count corresponds to both the intensity in the mask as well as its position in the name vector 4) To separate out the values with a specific intensity, need to do some weird math (b/c I can't just pull out intensity directly: If my count is 1, I need to pull out only values that are 1 from the map. This is equivalent to ispositive(a-[1-1])+isnegative(a-[1+1]). For 2, ispositive(a-[2-1]+isnegative(a-[2+1]) 
-# dependencies: Afni 
-
-#make list of yeo network abbreviations
-yeo_7_net_list=("VIS"
-		"MOT"
-		"DA"
-		"VA"
-		"LIM"
-		"FP"
-		"DM")
-
-#network path
-network_path=/Users/eballer/BBL/msdepression/templates/Yeo_JNeurophysiol11_MNI152/
-
-#current working directory
-cwd=`pwd`
-echo $cwd
-
-#set volume mask [change this if I ever want to use a diff mask
-volume_mask=Yeo2011_7Networks_MNI152_FreeSurferConformed1mm_LiberalMask.nii
-
-#assign generic network_prefix_list variable; this is in case I decide I want to do this for 17 networks
-network_prefix=yeo_7_net
-network_list=$(echo ${network_prefix}_list)
-
-echo "network list = ${network_list}"
-
-#set counter, this will be used in loop
-network_counter=1
-
-echo "Volume mask = ${volume_mask}"
-
-echo "Deleting old masks"
-rm -f ${network_path}/${network_prefix}_*.nii
-
-echo "Making individual network maps"
[…]

### 2023-03-03 calc_vol_fascicles_within_depression_mask.sh
Delete calc_vol_fascicles_within_depression_mask.sh

@@ -1,58 +1,0 @@
-#### calculate volume of fascicles within the depression network ####
-### Pre: All streamlines must have been made in dsi studio (these are the nii files)
-### Post: A depression mask resampled to template space, file <streamline_volume_within_dep_network>, containing each fascicle and its corresponding % within depression network
-### Uses: Trying to figure out which fascicles are in the depression network for dimensionality reduction for use in later regressions (i.e. decrease # comparisons by only looking at fascicles within depression network)
-### Dependencies: This script requires afni to be installed (currently using 20.1)
-
-#set outfile
-outfile="/project/msdepression/results/streamline_volume_within_dep_network.csv"
-
-#set template directories, will be making our resampled mask from within this
-template="Depression_Clust_mask_roi_binarized.nii.gz"
-template_dir="/project/msdepression/templates/harvard_depression"
-resampled_filename="resampled_${template}"
-template_fullpath="/project/msdepression/templates/harvard_depression/${template}"
-
-#set fascicle directory
-fascicle_direc="/project/msdepression/templates/dti/HCP_YA1065_tractography/"
-fascicle_type="association"
-fascicle_for_resampling="AF_L.nii.gz"
-
-#set home directory
-homedir="/project/msdepression/scripts/"
-
-#initiate csvs
-rm -f $outfile
-touch $outfile
-echo "fascicle,num_voxels_total,non_zero_voxels_in_dep_map,prop_in_mask" >> $outfile
-
-#resample Depression_Clust_mask_roi_binarized.nii.gz to match sample fascicle - AF_L.nii.gz
-if [ -f "${template_dir}/${resampled_filename}" ]; then
-	rm -f ${template_dir}/${resampled_filename}
-fi
-3dresample -master ${fascicle_direc}/${fascicle_type}/${fascicle_for_resampling} -prefix ${template_dir}/${resampled_filename} -input ${template_fullpath} -rmode NN
-
-#go through each fiber, get volume, and store it in csv
-fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"
-for fiber_bundle_type in ${fiber_bundle_type_list}; do
[…]

### 2023-03-03 calc_vol_fascicles_within_yeo_7_net.sh
Delete calc_vol_fascicles_within_yeo_7_net.sh

@@ -1,100 +1,0 @@
-#### calculate volume of fascicles within the yeo 7 networks ####
-### Pre: All streamlines must have been made in dsi studio (these are the nii files)
-### Post: Yeo7 network mask resampled to template space, file <streamline_volume_within_${network}_network>, containing each fascicle and its corresponding % within yeo networ. Additionally, will have a file that contains all of the overlaps per each network, and another with a binarized (yes/no)k
-### Uses: Trying to figure out which fascicles are in the yeo 7 network for dimensionality reduction for use in later regressions (i.e. decrease # comparisons by only looking at fascicles within yeo 7 networks)
-### Dependencies: This script requires afni to be installed (currently using 20.1)
-
-#output directory
-outdir="/project/msdepression/results/"
-
-#assemble yeo 7 networks
-
-yeo_7_net_list=("VIS"
-                "MOT"
-                "DA"
-                "VA"
-                "LIM"
-                "FP"
-                "DM")
-
-#set up output file names
-all_net_overlap_proportions="${outdir}/yeo_7_network_overlap_proportions.txt"
-all_net_overlap_proportions_binarized="${outdir}/yeo_7_network_overlap_proportions_binarized.txt"
-
-#set counter to keep track of whether it is the first iteration (and you need to do a special column add), or just take the last line and pr it
-counter=1
-for network in "${yeo_7_net_list[@]}"; do
-	#set outfile
-	outfile="/project/msdepression/results/streamline_volume_within_${network}_network.csv"
-
-	#set template directories, will be making our resampled mask from within this
-	template="yeo_7_net_${network}.nii"
-	template_dir="/project/msdepression/templates/Yeo_JNeurophysiol11_MNI152/"
-	resampled_filename="resampled_${template}"
-	template_fullpath="${template_dir}/${template}"
-
-	#set fascicle directory
-	fascicle_direc="/project/msdepression/templates/dti/HCP_YA1065_tractography/"
-	fascicle_type="association"
-	fascicle_for_resampling="AF_L.nii.gz"
-
-	#set home directory
-	homedir="/project/msdepression/scripts/"
-
[…]

### 2023-03-03 dacSummary.R
Delete dacSummary.R

@@ -1,1 +1,0 @@
-# This script takes in data from DAC output and summarizes statistics

### 2023-03-03 make_afni_map_w_fascicles_colored_by_prop_or_volume.sh
Delete make_afni_map_w_fascicles_colored_by_prop_or_volume.sh

@@ -1,66 +1,0 @@
-#!/bin/bash
-
-##### Welcome to make_afni_map_w_fascicles_colored_by_prop_overlap ###
-### Pre: A file that has two columns, no headers. First column is the abbreviation of a fascicle, second is proportion overlap (or volume overlap)
-### Post: A nifti where each fascicle is colored by the proportion of overlap, or any value you give it
-### Uses: I wanted to find a way to show, in a brain, which fascicles have the most overlap with lesions (though this can be done w/depression, non-dep, volume). This will go through a table, and for each fascicle, multiply it times the value in the second column, add a suffix, and then at the end, sum them all to make a final image
-### Dependencies: Afni
-
-#change this if doing on pmacs
-homedir="/Users/eballer/BBL/msdepression/"
-
-default=${homedir}/results/fascicle_mean_proportion_overlap_nfasc_77.csv
-default_suffix="prop"
-
-if [ $# == 0 ]
-then
-	echo "We will use the default path file" $default
-    	echo "We will use the default suffix " $default_suffix
-    	fascicle_file=${default}
-	suffix=${default_suffix}
-
-else  
-	fascicle_file=$1
-	suffix=$2
-fi
-
-
-echo "File being read is "$fascicle_file
-echo "Suffix being used is " $suffix
-
-fascicles=$(cat $fascicle_file)
-
-cd ${homedir}/templates/dti/HCP_YA1065_tractography/niftis_excluding_cranial_nerves
-
-(pwd)
-for fascicle in ${fascicles}; do
-	# extract fascicle name and value
-	fascicle_name=$(echo ${fascicle} | perl -pe 's/(.*),.*/$1/')
-	fascicle_value=$(echo ${fascicle} | perl -pe 's/.*,(.*)/$1/')
-	
-	# remove nifti w/values if already there
-	rm -rf ${fascicle_name}_${suffix}.nii.gz
-
-	#take the fascicle .nii, multiply it with the value
-	command=$(echo 3dcalc -a ${fascicle_name}.nii.gz -expr \'a*${fascicle_value}\' -prefix ${fascicle_name}_${suffix}.nii.gz)
-	echo "Here is the command: "${command}
-	
-	#run command, have to do separately to get variable into afni command
-	eval "${command}";
-done
-
-#remove old file
-rm -rf fascicle_${suffix}_sum.nii.gz
[…]

### 2023-03-03 make_streamline_volumes_single_subj_local_comp.sh
Delete make_streamline_volumes_single_subj_local_comp.sh

@@ -1,60 +1,0 @@
-### Make streamline volumes ###
-
-### Pre: All streamlines must have been made in dsi studio (these are the nii files)
-### Post: One csv that contains the name of the streamline as one column, and the volume of the streamline in a second column, separated by a ','
-### Uses: Will be getting the volumes of the streamlines for use in later density analyses
-### Dependencies: This script requires afni to be installed.
-
-#module load afni
-
-#module load singularity
-#export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
-#export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
-num_cores=1
-
-#set default directory/templates. This will change when we use local, personal dwi
-#template='/project/msdepression/templates/dti/HCP1065.1mm.fib.gz'
-#fascicle_directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
-fascicle_directory='/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/'
-
-output_csv="${fascicle_directory}/fiber_volume_values.csv"
-
-#initiate csv
-rm -f $output_csv
-touch $output_csv
-echo "fascicle,volume_full" >> $output_csv
-#if [ $# == 0 ]
-#then
- #   echo "You did not enter a file. Make sure your call makes sense"
-  #  echo "for now, using default $fascicle_directory"
- #else  
- #   lesion=$1
-    #parent_dir=$(echo ${lesion} | perl -pe s'/(.*)\/.*/$1/g')
-    parent_dir=${fascicle_directory}
-#	echo "File being read is "$l
-#	echo "... Starting to make tracts"
-# make directory within parent directory for tracts
-   
-    echo "$parent_dir"
-	#mask_prefix=$(echo ${lesion} | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
-    #echo $mask_prefix
-	#if [ ! -d $parent_dir/fiber_tracking_maps ] 
-    #	then
-     #   	mkdir $parent_dir/fiber_tracking_maps
-    #	fi
-
-    #make sub directories for each bundle type, so within fiber_tracking_maps, there will be subdirectories for each type of map
-	fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
-    	for fiber_bundle_type in ${fiber_bundle_type_list}; do
[…]

### 2023-03-03 scripts/ants_registration_code.sh
Create ants_registration_code.sh

@@ -1,0 +1,114 @@
+#!/bin/bash
+### ANTS registration for MS Depression
+##pre- requires ms mimosa file that contains the directory, mni brain: mni_icbm152_t1_tal_nlin_asym_09a.nii
+##Post - ms participant T1w and mimosa registered to icbm_template space
+##Uses - We have all of these great ms depression scans but they are in native space. This will get them into mni space so they can be used with HCP templates, both structural and diffusion
+#Steps
+## 1 - take all file paths to data with good qc, extract file paths, and make shadow directories with linksin out local data directory
+## 2 - make the dilated mask
+     #####mask (currently using the one that was dilated in afni, d3
+     ## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
+## 3 - for each subject, run the registration, and then apply the transpforms
+##Dependencies: Using ANTs/2.3.5
+# Set up paths
+module load ANTs/2.3.5
+export ANTSPATH=/appl/ANTs-2.3.5/bin
+export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
+#######if it doesn't work, fix this ###
+#export PATH=${ANTSPATH}:$PATH
+
+
+
+
+#set default templatesfiles/paths
+
+base_dir='/project/msdepression/scripts'
+
+#Primary directorie - CHANGE THESE IF ANYTHING EVER CHANGES, IT WILL UPDATE FILE NAMES BELOW
+
+qc_file_path='/project/msdepression/data/melissa_martin_files/csv/minimelissa_list_n1'
+#qc_file_path='/project/msdepression/data/melissa_martin_files/csv/melissa_mimosa_100_and_75_paths'
+mni_t1_root='mni_icbm152_t1_tal_nlin_asym_09a'
+mask_dilation=3
+brain_mask_root="t1_n4_brainmask"
+t1="t1_n4_reg_brain_ws.nii.gz"
+mimosa_root="mimosa_binary_mask_0.25"
+out_data_path='/project/msdepression/data/subj_directories/'
+mni_t1_direc='/project/msdepression/templates/mni_icbm152_nlin_asym_09a/'
+
+#set outfile prefix
+outfile_prefix=ms_t1_to_mni_icbm152
+
+#for transform
[…]

### 2023-03-03 scripts/ants_registration_code.sh
Delete ants_registration_code.sh

@@ -1,114 +1,0 @@
-#!/bin/bash
-### ANTS registration for MS Depression
-##pre- requires ms mimosa file that contains the directory, mni brain: mni_icbm152_t1_tal_nlin_asym_09a.nii
-##Post - ms participant T1w and mimosa registered to icbm_template space
-##Uses - We have all of these great ms depression scans but they are in native space. This will get them into mni space so they can be used with HCP templates, both structural and diffusion
-#Steps
-## 1 - take all file paths to data with good qc, extract file paths, and make shadow directories with linksin out local data directory
-## 2 - make the dilated mask
-     #####mask (currently using the one that was dilated in afni, d3
-     ## 3dmask_tool -input t1_n4_brainmask.nii.gz -prefix t1_n4_brainmask_d3.nii.gz -dilate_input 3
-## 3 - for each subject, run the registration, and then apply the transpforms
-##Dependencies: Using ANTs/2.3.5
-# Set up paths
-module load ANTs/2.3.5
-export ANTSPATH=/appl/ANTs-2.3.5/bin
-export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
-#######if it doesn't work, fix this ###
-#export PATH=${ANTSPATH}:$PATH
-
-
-
-
-#set default templatesfiles/paths
-
-base_dir='/project/msdepression/scripts'
-
-#Primary directorie - CHANGE THESE IF ANYTHING EVER CHANGES, IT WILL UPDATE FILE NAMES BELOW
-
-qc_file_path='/project/msdepression/data/melissa_martin_files/csv/minimelissa_list_n1'
-#qc_file_path='/project/msdepression/data/melissa_martin_files/csv/melissa_mimosa_100_and_75_paths'
-mni_t1_root='mni_icbm152_t1_tal_nlin_asym_09a'
-mask_dilation=3
-brain_mask_root="t1_n4_brainmask"
-t1="t1_n4_reg_brain_ws.nii.gz"
-mimosa_root="mimosa_binary_mask_0.25"
-out_data_path='/project/msdepression/data/subj_directories/'
-mni_t1_direc='/project/msdepression/templates/mni_icbm152_nlin_asym_09a/'
-
-#set outfile prefix
-outfile_prefix=ms_t1_to_mni_icbm152
-
-#for transform
[…]

### 2023-03-03 scripts/parse_mimosa_files.sh
Create parse_mimosa_files.sh

@@ -1,0 +1,5 @@
+
+
+more 1yr_report_address.csv | grep "mimosa" | grep ",,75.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_75_paths
+more 1yr_report_address.csv | grep "mimosa" | grep ",,100.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_100_paths
+cat mimosa_100_paths mimosa_75_paths > mimosa_100_and_75_paths

### 2023-03-03 parse_mimosa_file.sh
Delete parse_mimosa_file.sh

@@ -1,3 +1,0 @@
-more 1yr_report_address.csv | grep "mimosa" | grep ",,75.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_75_paths
-more 1yr_report_address.csv | grep "mimosa" | grep ",,100.0" | perl -pe 's/(.*)\/mimosa.*/$1/' > mimosa_100_paths
-cat mimosa_100_paths mimosa_75_paths > mimosa_100_and_75_paths

### 2023-03-08 scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
Rename msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd to scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
Add files via upload

--- 2023-03-01 msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
@@
 ##########################################################
+full_fascicle_names <- read.csv(paste0(homedir, "/templates/dti/HCP_YA1065_tractography/fascicle_abbreviation.csv"), header = T, sep = ",")
 
@@
 fascicle_volumes_dep_network_no_cranial_nerves$fascicle <- factor(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, levels = fascicle_volumes_dep_network_no_cranial_nerves$fascicle[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)])
+
+#create data frame with fascicles sorted by overlap with depression network, replace abbreviations with full fascicle names and swap _ for space in the full name
+df_fascicles_sorted_by_volume_in_mask <- fascicle_volumes_dep_network_no_cranial_nerves %>%  
+  replace(is.na(.), 0) %>% 
+  arrange(desc(non_zero_voxels_in_dep_map)) %>% 
+  left_join(full_fascicle_names, by = c("fascicle" = "Abbreviation")) %>%
+  mutate(full_fascicle_name_space = str_replace_all(Full_name, pattern = "_", replacement = " ")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " R$", replacement = " Right")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " L$", replacement = " Left")) %>%
+  mutate(full_fascicle_name_space = gsub(pattern = "([0-9])", replacement = " \\1", x = full_fascicle_name_space))
+
+write.table(df_fascicles_sorted_by_volume_in_mask, paste0(homedir, "/results/df_fascicles_sorted_by_volume_in_mask_for_eTable3.csv"), quote = F, sep = ",") #had to manually change the spreadsheet because the dictionary is based on a spelling mistake - C_R_R is Cingulum Rarolfactory in the dictionary, but should be Parolfactory
 

### 2023-03-09 scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
Add files via upload
Formatted (Left) (Right)

--- 2023-03-01 msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
@@
 fascicle_volumes_dep_network_no_cranial_nerves$fascicle <- factor(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, levels = fascicle_volumes_dep_network_no_cranial_nerves$fascicle[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)])
+
+#create data frame with fascicles sorted by overlap with depression network, replace abbreviations with full fascicle names and swap _ for space in the full name
+df_fascicles_sorted_by_volume_in_mask <- fascicle_volumes_dep_network_no_cranial_nerves %>%  
+  replace(is.na(.), 0) %>% 
+  arrange(desc(non_zero_voxels_in_dep_map)) %>% 
+  left_join(full_fascicle_names, by = c("fascicle" = "Abbreviation")) %>%
+  mutate(full_fascicle_name_space = str_replace_all(Full_name, pattern = "_", replacement = " ")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " R$", replacement = " (Right)")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " L$", replacement = " (Left)")) %>%
+  mutate(full_fascicle_name_space = gsub(pattern = "([0-9])", replacement = " \\1", x = full_fascicle_name_space))
+
+write.table(df_fascicles_sorted_by_volume_in_mask, paste0(homedir, "/results/df_fascicles_sorted_by_volume_in_mask_for_eTable3.csv"), quote = F, sep = ",") #had to manually change the spreadsheet because the dictionary is based on a spelling mistake - C_R_R is Cingulum Rarolfactory in the dictionary, but should be Parolfactory
 

### 2023-05-17 scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
added corpus callosum sensitivity analyses

@@
 
+#scatter for Taki
+scatter_net <- ggplot(newdf,  aes(x=dep_net_group, y=net_values)) +  geom_jitter(shape=16, position=position_jitter(0.2))
+
+scatter_dx <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum)) +  geom_jitter(shape=16, position=position_jitter(0.2))
+
 # bar graph instead of violin
@@
 
-```{r sensitivity analyses}
+#### tercile/quintile cutoffs
+```{r sensitivity_analyses_diff_depression_cutoffs}
 
@@
 ```
+
+#### dropping corpus callosum
+```{r sensitivity_analysis_no_cc}
+## compare disease within and outside depression network, between diagnoses, and interactions
+
+## volume of corpus collosum
+volume_of_healthy_corpus_callosum = fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[fascicle_volumes_dep_network_no_cranial_nerves$fascicle == "CC"]
+
+###############################
+### Constructing Data Frame ###
+###############################
+
+#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
+dep_and_healthy_groups_for_ICD_analysis_no_cc <- df_demo_and_fascicles_prop_and_vol %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>%
+
+  #sum of total volume lost and proportion of volume lost 
+  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix[fascicle_names_w_vol_suffix != "CC_vol"])))) %>%
+  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/(total_volume_all_full_fascicles-volume_of_healthy_corpus_callosum)) %>%
+
+  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_indepnet_top_quartile = 
+           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves[dep_network_names_top_quartile_no_cranial_nerves != "CC_vol"])))) %>%
+  mutate(proportion_volume_lost_dep_net_by_network_size_sum = sum_fascicle_vol_lost_indepnet_top_quartile /(dep_network_total_volume_top_quartile_no_cranial_nerves-volume_of_healthy_corpus_callosum)) %>%
+
+  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves)))) %>%
+  mutate(proportion_volume_lost_nondep_net_by_network_size_sum = sum_fascicle_vol_lost_nondep_bottom_3_quartiles/bottom_75_total_volume_no_cranial_nerves) %>%
+
+  ungroup()
+
+###################
+
+### Interaction Network x Diagnosis
+
+#clean up data frame
+df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis_no_cc %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
+  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
+  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
+  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  mutate(Network = as.factor(Network)) %>%
+  ungroup()
+
+#diagnosis: Dep > nondep, P= 0.013, T 2.49, Estimate 0.06 ; Network dep>nondep P = 9.48*10^-5, T = 3.95, estimate 0.03; intx  P=.03, T = -2.12, Estimate = -0.02
+diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer)
+
+#Diagnosis: MS+ > MS- 0.01-0.10; Network dep>nondep 0.02-0.05; intx: -0.04--0.002 
+confint(diagnosis_by_network_interaction_lmer)
+
+#### Volume x Effect size analysis
+
+fascicle_dep_vs_healthy_effsize <- sapply(fascicle_names_w_vol_suffix[fascicle_names_w_vol_suffix != "CC_vol"], function(fascicle) 
+{
+
+  text_for_eval <- paste0("rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis_no_cc, formula = ", fascicle, " ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = \"perc\", nboot = 1000)$effsize")
+  eval(parse(text=text_for_eval))
+})
+names(fascicle_dep_vs_healthy_effsize) <- fascicle_names_w_vol_suffix[fascicle_names_w_vol_suffix != "CC_vol"]
+
+#convert to data frame
+df_effsize_values <- as.data.frame(fascicle_dep_vs_healthy_effsize)
+
+df_fascicle_dep_mask_volume_overlap_effsize_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle[fascicle_volumes_dep_network_no_cranial_nerves$fascicle != "CC"], fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map[fascicle_volumes_dep_network_no_cranial_nerves$fascicle != "CC"], df_effsize_values)
+names(df_fascicle_dep_mask_volume_overlap_effsize_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_effsize")
+
+# plot fascicle overlap x Eff size value
+#statistic, p=0.0002, T =3.865 , eff size (adjusted R^2 = 0.16)
+dep_overlap_x_dep_vs_healthy_vol_effsize <- lm(dep_vs_healthy_effsize ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_effsize_value )
+summary(dep_overlap_x_dep_vs_healthy_vol_effsize)
+
+scatter_effsize_val_by_dep_net_overlap_log <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize, color=dep_vs_healthy_effsize)) +
+  geom_point(size=2) + 
+  scale_color_gradient(low = "#FFFFFF", high = "#ff00ff") +
+  geom_smooth(formula = y ~ x, method=lm, color="black")+
+  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(0, 10, 100, 1000, 10000, 100000)) + 
+  xlab("Volume of Fascicle in Depression Network, excluding CC") + 
+  ylab("Effect Size (r)") + 
+  theme(plot.title=element_text(hjust=0.5),
+        axis.text = element_text(size=16, color = "black"),
+        axis.title = element_text(size=18, color = "black"),
+        axis.line = element_line(color="black", size=1.3),
+        panel.background = element_blank(),
+        legend.position = "none",
+        aspect.ratio = 4/4) + 
+  annotate("text", x=1.6, y=0.14, size = 5, 
+           label = paste0("P = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize)$coef[2,4],4)))
+
+print(scatter_effsize_val_by_dep_net_overlap_log)
+
+#Final data: 
+#Main effect of network Dep > Nondep (Estimate = 0.03, 95% CI 0.02-0.05 , P= 9.48*10^-5)
+#Main effecet of diagnosis MS+Depression > MS-Depression (Estimate = 0.06, 95% CI 0.01-0.10 , P = 0.013)
+#Main effect of interaction (Estimate = 0.02, 95% CI 0.002 -0.04 , P=0.03)
+
+```

### 2023-07-21 scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
added age as covariate

@@
   mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
-  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
-  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network, PAT_AGE_AT_EXAM) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
   mutate(Network = as.factor(Network)) %>%
@@
 confint(diagnosis_by_network_interaction_lmer)
+
+#with age; diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020
+diagnosis_by_network_interaction_lmer_age <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_age)
+
+#Diagnosis: MS+ > MS- 0.011-0.1; Network dep>nondep -.104--.076; intx: -0.040--0.003
+confint(diagnosis_by_network_interaction_lmer_age)
 

### 2023-07-21 scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
with age

@@
 
-#with age; diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020
+#with age; diagnosis: Dep > nondep T = 2.1, p = 0.035; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020
 diagnosis_by_network_interaction_lmer_age <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
@@
 
-#Diagnosis: MS+ > MS- 0.011-0.1; Network dep>nondep -.104--.076; intx: -0.040--0.003
+#Diagnosis: MS+ > MS- 0.004-0.9; Network dep>nondep -.104--.076; intx: -0.040--0.003
 confint(diagnosis_by_network_interaction_lmer_age)

### 2023-09-27 Biological_Psychiatry_Revision_Scripts/msdepression_n380_started_20230927_mm_BP_r1.Rmd
Add files via upload

--- 2023-05-17 scripts/msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
@@
         - interaction (0.022, CI = 0.004-0.040, P = 0.018)
+
+#### New analyses per Biological Psychiatry Reviewer Recommendations from 9/26/2023
+
+[x] Add sex into model - Reviewer 1
+
+    - Main effect of network Dep > Nondep (Estimate = 0.090, 95% CI 0.076-0.10, P = 2*10^-16)
+    - Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.058, 95% CI  0.013-0.10, P= 0.012)
+    - Main effect of sex M>F (Estimate = 0.021, 95% CI -0.061-0.020, P = 0.32)
+    - Main effect of interaction (Estimate = 0.022, 95% CI 0.0035 - 0.040, P=0.020)
+
+[x] Different thresholds for defining the depression network (using quartile breakdown for in/out) - Reviewer 1
+  [x] Threshold the Harvard Mask at T>2.3 (P<0.01)
+
+    - Main effect of network Dep > Nondep (Estimate = 0.048, 95% CI 0.037-0.058, P = 2*10^-16)
+    - Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.053, 95% CI 0.0085-0.098, P= 0.020)
+    -Main effect of interaction (Estimate = 0.016, 95% CI 0.0028-0.030, P=0.019)
+
+  [x] Threshold the Harvard Mask at T>2.6 (P<0.005)
+
+    -Main effect of network Dep > Nondep (Estimate = 0.084, 95% CI 0.070-0.10, P = 2*10^-16)
+    -Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.056, 95% CI 0.011-0.10, P= 0.016)
+    -Main effect of interaction (Estimate = 0.021, 95% CI 0.0032 - 0.039, P=0.021)
+
+[x] Prevalence of depression in the PROMIS cohort - Reviewer 3
+
+    - chi sq for prevalence? 49 promis scores in depressed group (out of 232); 36 promis scores in nondepressed group (out of 148)
+    - X-squared = 0.21, p = 0.65, not significant
 
@@
 library(ggsignif)
+#library(lspline)
 
@@
 
+###new for BP
+fascicle_volumes_dep_2_6 <- read.csv(paste0(homedir, "/results/streamline_volume_within_dep_network_2_6.csv"), header = T, sep = ",")
+
+fascicle_volumes_dep_2_3 <- read.csv(paste0(homedir, "/results/streamline_volume_within_dep_network_2_3.csv"), header = T, sep = ",")
+
 ##########################################################
@@
 df_demo_and_fascicles_prop_and_vol <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes)
+
+#repeat for 2.3 and 2.6
 
@@
 
+#testing whether PROMIS sub sample is equally representative with respect to prevalence of depression/non-depression.
+I.e.
+do more people with depression get PROMIS scores?
+or are we equally drawing from both groups?
+
+#first number is # with depression, second is number without depression
+num_w_promis <- c(sum(df_people_with_promis_data_scores_added_unique$depGroupVar == 2), sum(df_people_with_promis_data_scores_added_unique$depGroupVar == 1))
+
+#first number is # with depression, second is number without depression
+num_total <- c(sum(df_good_mimosa_unique_dep_and_healthy$depGroupVar == 2), sum(df_good_mimosa_unique_dep_and_healthy$depGroupVar == 1))
+
+df_2x2_promis_score_by_total_people_per_diagnosis <- data.frame(cbind(num_w_promis, num_total), row.names = c("MS+Depression", "MS-Depression"))
+
+#X-squared = 0.21, p = 0.65, not significant)
+chisq <- chisq.test(df_2x2_promis_score_by_total_people_per_diagnosis)
 ```
@@
   mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
-  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
-  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  dplyr::select(subject, osex, Diagnosis, Dep_network, Nondep_network, PAT_AGE_AT_EXAM) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM, -osex) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
   mutate(Network = as.factor(Network)) %>%
@@
 
+#with age; diagnosis: Dep > nondep T = 2.1, p = 0.035; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020; also tried with + lspline(PAT_AGE_AT_EXAM, knots = c(5)) instead of PAT_AGE_AT_EXAM, findings stayed the same so probs not modeling true spline
+diagnosis_by_network_interaction_lmer_age <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_age)
+
+#Diagnosis: MS+ > MS- 0.004-0.9; Network dep>nondep -.104--.076; intx: -0.040--0.003
+confint(diagnosis_by_network_interaction_lmer_age)
+
+#with sex; diagnosis: Dep > nondep T = 2.5, p = 0.012; sex: M>F = 1, p = 0.32; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020
+diagnosis_by_network_interaction_lmer_sex <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + osex, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_sex)
+
+#Diagnosis: MS+ > MS- 0.013-0.10; sex: M>F 0.06 - -0.02 Network dep>nondep -.104--.076; intx: -0.040--0.003
+confint(diagnosis_by_network_interaction_lmer_sex)
+
 #Final data: 
 #Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.076-0.104, P = 2*10^-16)
-#Main effecet of diagnosis MS+Depression > MS-Depression (Estimate = 0.056, 95% CI 0.011-0.100, P= 0.015)
+#Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.056, 95% CI 0.011-0.100, P= 0.015)
 #Main effect of interaction (Estimate = 0.022, 95% CI 0.003 - 0.040, P=0.020)
+
+#final data with sex included:
+#Main effect of network Dep > Nondep (Estimate = 0.090, 95% CI 0.076-0.10, P = 2*10^-16)
+#Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.058, 95% CI  0.013-0.10, P= 0.012)
+#Main effect of sex M>F (Estimate = 0.021, 95% CI -0.061-0.020, P = 0.32)
+#Main effect of interaction (Estimate = 0.022, 95% CI 0.0035 - 0.040, P=0.020)
 
@@
 ```
+
+#### BP repeat analysis with depression threshold of 2.3 and 2.6
+
+### Define Depression Network by Volume of Overlap for 2.3
+```{r visualize_depression_network_volume}
+###################################################
+### Defining and Visualizing Depression Network ###
+###################################################
+
+#This chunk reads in a file where columns are fascicle, num_voxels_total, non_zero_voxels_in_dep_map, prop_in_mask
+# by the end, the depression mask will be defined
+# We consider fascicles in the depression network to be the ones that share the most VOLUME with the depression mask
+# Of course, it is possible that huge fascicles contribute a ton to the depression network, but also the non-depression network
+# And to get at specificity (i.e.
+which fibers most exclusively send information within the depression network), we could look at proportion of the fiber (volume that intersects / total volume of the fascicle).
+But then you have a situation where tiny fibers get lots of credit.
+We have actually tried this as well and the findings aren't that different.
+#But for the sake of this paper, will do volume because the question is all volume based (volume of lesions to volume of fascicles)
+
+######## Histograms for depression network
+#first set the values to be the volumes
+fascicle_volumes_dep_2_3_network_no_cranial_nerves <- fascicle_volumes_dep_2_3 %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+#set NAs to zero
+fascicle_volumes_dep_2_3_network_no_cranial_nerves$non_zero_voxels_in_dep_map[is.na(fascicle_volumes_dep_2_3_network_no_cranial_nerves$non_zero_voxels_in_dep_map)] = 0
+
+#sort based on overlap with depression depression map
+fascicle_volumes_dep_2_3_network_no_cranial_nerves$fascicle <- factor(fascicle_volumes_dep_2_3_network_no_cranial_nerves$fascicle, levels = fascicle_volumes_dep_2_3_network_no_cranial_nerves$fascicle[order(fascicle_volumes_dep_2_3_network_no_cranial_nerves$non_zero_voxels_in_dep_map)])
+
+#create data frame with fascicles sorted by overlap with depression network, replace abbreviations with full fascicle names and swap _ for space in the full name
+df_fascicles_sorted_by_volume_in_mask <- fascicle_volumes_dep_2_3_network_no_cranial_nerves %>%  
+  replace(is.na(.), 0) %>% 
+  arrange(desc(non_zero_voxels_in_dep_map)) %>% 
+  left_join(full_fascicle_names, by = c("fascicle" = "Abbreviation")) %>%
+  mutate(full_fascicle_name_space = str_replace_all(Full_name, pattern = "_", replacement = " ")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " R$", replacement = " (Right)")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " L$", replacement = " (Left)")) %>%
+  mutate(full_fascicle_name_space = gsub(pattern = "([0-9])", replacement = " \\1", x = full_fascicle_name_space))
+
+write.table(df_fascicles_sorted_by_volume_in_mask, paste0(homedir, "/results/df_fascicles_sorted_by_volume_in_mask_2_3.csv"), quote = F, sep = ",") #had to manually change the spreadsheet because the dictionary is based on a spelling mistake - C_R_R is Cingulum Rarolfactory in the dictionary, but should be Parolfactory
+
+#figure out how many fascicles are in top quartile
+quartile <- round((dim(fascicle_volumes_dep_2_3_network_no_cranial_nerves)[1] / 4),0)
+
+#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take names of the top quartile
+dep_network_names_top_quartile_no_cranial_nerves_2_3<- (fascicle_volumes_dep_2_3_network_no_cranial_nerves[order(fascicle_volumes_dep_2_3_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile]
+
+write.table(dep_network_names_top_quartile_no_cranial_nerves, paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol_2.3.txt"), row.names = F, col.names = F, quote = F)
+
+df_for_making_binary_color_maps <- fascicle_volumes_dep_2_3_network_no_cranial_nerves %>% dplyr::select(fascicle, non_zero_voxels_in_dep_map)
+
+write.table(df_for_making_binary_color_maps, paste0(homedir, "/results/volume_of_overlap_w_dep_net_n77_T_2_3.csv"), row.names = F, col.names = F, quote = F, sep = ",")
+
+#get names of bottom 3 quartiles
+
+#get a network that includes all of the fascicles NOT in dep 25% network
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_3 <- fascicle_volumes_dep_2_3_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_2_3_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_2_3)]
+
+#### Get sums of the volumes of the fascicles in each network, and total sum
+
+#top quartile, depression network
+dep_network_total_volume_top_quartile_no_cranial_nerves_2_3 <-
+sum(fascicle_volumes_dep_2_3_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_2_3_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_2_3)])
+
+#bottom 3 quartiles, nondepression network
+bottom_75_total_volume_no_cranial_nerves_2_3 <-
+sum(fascicle_volumes_dep_2_3_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_2_3_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_3)])
+
+#### Calculate total volume in all non-cranial_nerve_network
+total_volume_all_full_fascicles_2_3 <- dep_network_total_volume_top_quartile_no_cranial_nerves_2_3 + bottom_75_total_volume_no_cranial_nerves_2_3
+
+###
+
+plot_all_fascicle_volumes_dep_2_3_network_no_cranial_nerves <-ggplot(data = fascicle_volumes_dep_2_3_network_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
+  ggtitle("Volume Overlap per Fascicle 2_3") + 
+  geom_bar(stat="identity", fill="006300") +
+  xlab("Fascicle") +
+  ylab("Overlap with Depression Mask") +
+  ylim(0,65000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(plot_all_fascicle_volumes_dep_2_3_network_no_cranial_nerves)
+
+#save pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/plot_all_fascicle_volumes_dep_2_3_network_no_cranial_nerves.pdf"))
+  plot_all_fascicle_volumes_dep_2_3_network_no_cranial_nerves
+dev.off()
+
+### Change names to add suffix
+#replace each fascicle with vol suffix
+dep_network_names_top_quartile_no_cranial_nerves_2_3 <- dep_network_names_top_quartile_no_cranial_nerves_2_3 %>% 
+  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
+
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_3 <- dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_3 %>%
+   gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
+
+```
+
+### Compare between networks, diagnostic groups, and interaction
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected}
+## compare disease within and outside depression network, between diagnoses, and interactions
+
+###############################
+### Constructing Data Frame ###
+###############################
+
+#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>%
+
+  #sum of total volume lost and proportion of volume lost 
+  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
+  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_volume_all_full_fascicles) %>%
+
+  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_indepnet_top_quartile_2_3 = 
+           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves_2_3)))) %>%
+  mutate(proportion_volume_lost_dep_net_by_network_size_sum_2_3 = sum_fascicle_vol_lost_indepnet_top_quartile_2_3 /dep_network_total_volume_top_quartile_no_cranial_nerves_2_3) %>%
+
+  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles_2_3 = 
+           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_3)))) %>%
+  mutate(proportion_volume_lost_nondep_net_by_network_size_sum_2_3 = sum_fascicle_vol_lost_nondep_bottom_3_quartiles_2_3 /bottom_75_total_volume_no_cranial_nerves_2_3) %>%
+
+  ungroup()
+
+###################
+
+### Interaction Network x Diagnosis
+
+#clean up data frame
+df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
+  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum_2_3) %>%
+  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum_2_3) %>%
+  dplyr::select(subject, osex, Diagnosis, Dep_network, Nondep_network, PAT_AGE_AT_EXAM) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM, -osex) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  mutate(Network = as.factor(Network)) %>%
+  ungroup()
+
+#diagnosis: Dep > nondep T = 2.3, p = 0.02; Network dep>nondep = T=8.7, p = 2*10^-16; intx = T = 2.4, p = 0.019
+diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer)
+
+#Diagnosis: MS+ > MS- 0.009-0.10; Network dep>nondep -.04--.06; intx: -0.030--0.003
+confint(diagnosis_by_network_interaction_lmer)
+
+#with age; diagnosis: Dep > nondep T = 2.01, p = 0.046; Network dep>nondep = T=-8.7, p = 2*10^-16; intx = T = 2.4, p = 0.019
+diagnosis_by_network_interaction_lmer_age <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_age)
+
+#Diagnosis: MS+ > MS- 0.001-0.9; Network dep>nondep -.04--.06; intx: -0.030--0.003; Age = 0.002=0.005
+confint(diagnosis_by_network_interaction_lmer_age)
+
+#with sex; diagnosis: Dep > nondep T = 2.4, p = 0.016; sex: M>F = 1, p = 0.32; Network dep>nondep = T=8.7, p = 2*10^-16; intx = T = 2.4, p = 0.020
+diagnosis_by_network_interaction_lmer_sex <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + osex, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_sex)
+
+#Diagnosis: MS+ > MS- 0.01-0.10; sex: M>F 0.06 - -0.02 Network dep>nondep -.04--.06; intx: -0.030--0.003
+confint(diagnosis_by_network_interaction_lmer_sex)
+
+#Final data: 
+#Main effect of network Dep > Nondep (Estimate = 0.048, 95% CI 0.037-0.058, P = 2*10^-16)
+#Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.053, 95% CI 0.0085-0.098, P= 0.020)
+#Main effect of interaction (Estimate = 0.016, 95% CI 0.0028-0.030, P=0.019)
+
+```
+
+#### BP repeat analysis with depression threshold of 2.6
+
+### Define Depression Network by Volume of Overlap for 2.6
+```{r visualize_depression_network_volume}
+###################################################
+### Defining and Visualizing Depression Network ###
+###################################################
+
+#This chunk reads in a file where columns are fascicle, num_voxels_total, non_zero_voxels_in_dep_map, prop_in_mask
+# by the end, the depression mask will be defined
+# We consider fascicles in the depression network to be the ones that share the most VOLUME with the depression mask
+# Of course, it is possible that huge fascicles contribute a ton to the depression network, but also the non-depression network
+# And to get at specificity (i.e.
+which fibers most exclusively send information within the depression network), we could look at proportion of the fiber (volume that intersects / total volume of the fascicle).
+But then you have a situation where tiny fibers get lots of credit.
+We have actually tried this as well and the findings aren't that different.
+#But for the sake of this paper, will do volume because the question is all volume based (volume of lesions to volume of fascicles)
+
+######## Histograms for depression network
+#first set the values to be the volumes
+fascicle_volumes_dep_2_6_network_no_cranial_nerves <- fascicle_volumes_dep_2_6 %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)
+
+#set NAs to zero
+fascicle_volumes_dep_2_6_network_no_cranial_nerves$non_zero_voxels_in_dep_map[is.na(fascicle_volumes_dep_2_6_network_no_cranial_nerves$non_zero_voxels_in_dep_map)] = 0
+
+#sort based on overlap with depression depression map
+fascicle_volumes_dep_2_6_network_no_cranial_nerves$fascicle <- factor(fascicle_volumes_dep_2_6_network_no_cranial_nerves$fascicle, levels = fascicle_volumes_dep_2_6_network_no_cranial_nerves$fascicle[order(fascicle_volumes_dep_2_6_network_no_cranial_nerves$non_zero_voxels_in_dep_map)])
+
+#create data frame with fascicles sorted by overlap with depression network, replace abbreviations with full fascicle names and swap _ for space in the full name
+df_fascicles_sorted_by_volume_in_mask <- fascicle_volumes_dep_2_6_network_no_cranial_nerves %>%  
+  replace(is.na(.), 0) %>% 
+  arrange(desc(non_zero_voxels_in_dep_map)) %>% 
+  left_join(full_fascicle_names, by = c("fascicle" = "Abbreviation")) %>%
+  mutate(full_fascicle_name_space = str_replace_all(Full_name, pattern = "_", replacement = " ")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " R$", replacement = " (Right)")) %>%
+  mutate(full_fascicle_name_space = str_replace(full_fascicle_name_space, pattern = " L$", replacement = " (Left)")) %>%
+  mutate(full_fascicle_name_space = gsub(pattern = "([0-9])", replacement = " \\1", x = full_fascicle_name_space))
+
+write.table(df_fascicles_sorted_by_volume_in_mask, paste0(homedir, "/results/df_fascicles_sorted_by_volume_in_mask_2_6.csv"), quote = F, sep = ",") #had to manually change the spreadsheet because the dictionary is based on a spelling mistake - C_R_R is Cingulum Rarolfactory in the dictionary, but should be Parolfactory
+
+#figure out how many fascicles are in top quartile
+quartile <- round((dim(fascicle_volumes_dep_2_6_network_no_cranial_nerves)[1] / 4),0)
+
+#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take names of the top quartile
+dep_network_names_top_quartile_no_cranial_nerves_2_6<- (fascicle_volumes_dep_2_6_network_no_cranial_nerves[order(fascicle_volumes_dep_2_6_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile]
+
+write.table(dep_network_names_top_quartile_no_cranial_nerves, paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol_2.6.txt"), row.names = F, col.names = F, quote = F)
+
+df_for_making_binary_color_maps <- fascicle_volumes_dep_2_6_network_no_cranial_nerves %>% dplyr::select(fascicle, non_zero_voxels_in_dep_map)
+
+write.table(df_for_making_binary_color_maps, paste0(homedir, "/results/volume_of_overlap_w_dep_net_n77_T_2_6.csv"), row.names = F, col.names = F, quote = F, sep = ",")
+
+#get names of bottom 3 quartiles
+
+#get a network that includes all of the fascicles NOT in dep 25% network
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_6 <- fascicle_volumes_dep_2_6_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_2_6_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_2_6)]
+
+#### Get sums of the volumes of the fascicles in each network, and total sum
+
+#top quartile, depression network
+dep_network_total_volume_top_quartile_no_cranial_nerves_2_6 <-
+sum(fascicle_volumes_dep_2_6_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_2_6_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_2_6)])
+
+#bottom 3 quartiles, nondepression network
+bottom_75_total_volume_no_cranial_nerves_2_6 <-
+sum(fascicle_volumes_dep_2_6_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_2_6_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_6)])
+
+#### Calculate total volume in all non-cranial_nerve_network
+total_volume_all_full_fascicles_2_6 <- dep_network_total_volume_top_quartile_no_cranial_nerves_2_6 + bottom_75_total_volume_no_cranial_nerves_2_6
+
+###
+
+plot_all_fascicle_volumes_dep_2_6_network_no_cranial_nerves <-ggplot(data = fascicle_volumes_dep_2_6_network_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
+  ggtitle("Volume Overlap per Fascicle 2_6") + 
+  geom_bar(stat="identity", fill="006300") +
+  xlab("Fascicle") +
+  ylab("Overlap with Depression Mask") +
+  ylim(0,65000) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(hjust = 0.5, size=14),
+        axis.title.x = element_text(size=14),
+        axis.title.y = element_text(size=14),
+        axis.line = element_line(size=1),
+        panel.background = element_rect(colour = "white", fill = "white"))
+print(plot_all_fascicle_volumes_dep_2_6_network_no_cranial_nerves)
+
+#save pdf
+pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/plot_all_fascicle_volumes_dep_2_6_network_no_cranial_nerves.pdf"))
+  plot_all_fascicle_volumes_dep_2_6_network_no_cranial_nerves
+dev.off()
+
+### Change names to add suffix
+#replace each fascicle with vol suffix
+dep_network_names_top_quartile_no_cranial_nerves_2_6 <- dep_network_names_top_quartile_no_cranial_nerves_2_6 %>% 
+  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
+
+dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_6 <- dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_6 %>%
+   gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
+
+```
+
+### Compare between networks, diagnostic groups, and interaction
+
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected}
+## compare disease within and outside depression network, between diagnoses, and interactions
+
+###############################
+### Constructing Data Frame ###
+###############################
+
+#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>%
+
+  #sum of total volume lost and proportion of volume lost 
+  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
+  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_volume_all_full_fascicles) %>%
+
+  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_indepnet_top_quartile_2_6 = 
+           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves_2_6)))) %>%
+  mutate(proportion_volume_lost_dep_net_by_network_size_sum_2_6 = sum_fascicle_vol_lost_indepnet_top_quartile_2_6 /dep_network_total_volume_top_quartile_no_cranial_nerves_2_6) %>%
+
+  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles_2_6 = 
+           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves_2_6)))) %>%
+  mutate(proportion_volume_lost_nondep_net_by_network_size_sum_2_6 = sum_fascicle_vol_lost_nondep_bottom_3_quartiles_2_6 /bottom_75_total_volume_no_cranial_nerves_2_6) %>%
+
+  ungroup()
+
+###################
+
+### Interaction Network x Diagnosis
+
+#clean up data frame
+df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
+  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum_2_6) %>%
+  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum_2_6) %>%
+  dplyr::select(subject, osex, Diagnosis, Dep_network, Nondep_network, PAT_AGE_AT_EXAM) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM, -osex) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  mutate(Network = as.factor(Network)) %>%
+  ungroup()
+
+#diagnosis: Dep > nondep T = 2.4, p = 0.016; Network dep>nondep = T=11.9, p = 2*10^-16; intx = T = 2.3, p = 0.021
+diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer)
+
+#Diagnosis: MS+ > MS- 0.01-0.10; Network dep>nondep -0.07-0.10; intx: -0.003--0.04
+confint(diagnosis_by_network_interaction_lmer)
+
+#with age; diagnosis: Dep > nondep T = 2.01, p = 0.036; Network dep>nondep = T=-11.9, p = 2*10^-16; intx = T = 2.3, p = 0.021, age T = 4.3, P 2*1-^-5
+diagnosis_by_network_interaction_lmer_age <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_age)
+
+#Diagnosis: MS+ > MS- 0.003-0.9; Network dep>nondep -.07--.10; intx: -0.04--0.003; Age = 0.002=0.006
+confint(diagnosis_by_network_interaction_lmer_age)
+
+#with sex; diagnosis: Dep > nondep T = 2.5, p = 0.013; sex: M>F = 1, p = 0.32; Network dep>nondep = T=11.9, p = 2*10^-16; intx = T = 2.3, p = 0.021
+diagnosis_by_network_interaction_lmer_sex <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + osex, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_sex)
+
+#Diagnosis: MS+ > MS- 0.01-0.10; sex: M>F 0.06 - -0.02 Network dep>nondep -.07--.09; intx: -0.04--0.003
+confint(diagnosis_by_network_interaction_lmer_sex)
+
+#Final data: 
+#Main effect of network Dep > Nondep (Estimate = 0.084, 95% CI 0.070-0.10, P = 2*10^-16)
+#Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.056, 95% CI 0.011-0.10, P= 0.016)
+#Main effect of interaction (Estimate = 0.021, 95% CI 0.0032 - 0.039, P=0.021)
+
+```

### 2023-09-27 lab_notebook/lab_notebook.Rmd
BP revisions

@@
 #### 20220930 or next week
-[] add post hoc */bars to figure 4
-[] add legend
-[] call it a win!
-[] add statistics to other graphs (p/effect size)
+[x] add post hoc */bars to figure 4
+  - things that are significant - depressed vs nondepressed network for each (big sig), dep net between dep & nondep (p=0.04); non dep network nonsig (p = 0.051)
+[x] add legend
+[x] call it a win!
+[x] add statistics to other graphs (p/effect size)
+[x] fix flow chart
+  - why is it that when you combine a data frame where the EMPIs overlap by 905 people, do you end up with only 783 unique EMPIs?
+  - perhaps the data_empi_ etc pulls people with scans that aren't on the same day as the ones that are mimosa'ed
+  - So what that would mean is when the merge of subject and exam date happens, some things are dropped.
+it is possible that there are more subjects that overlap in EMPI, but that scans were pulled on different dates (suggesting that there are missing scans, or scans that aren't in the big spreadsheet but are in the mimosa files).
+  - So i will merge data_empi with the mimosa qc, and check to see how many people were available in that overlap. And that will become my number I think.
+[x] start writing science
+
+#### 20230113
+[x] Wilcoxon eff size not working, can report everything from the mixed effect model
+[x] Made new code msdepression_n380_started_20230113_post_replication_just_simple_mm.Rmd
+[x] removing wilcoxon code
+[x] getting confidence intervals for mixed model with confint
+[x] in reporting, beta (estimate), CI (from confint), p value
+
+#### 20230510
+[x] going back to do some Yeo stuff. Made sure I could sum up each individual's Yeo network injury
+
+#### 20230511
+[x] make proportions of yeo injury by dividing by total volume within network
+[x] compare disease across networks 
+[x] compare disease within network across diagnoses
+
+#### 20230517
+[x] added sensitivity analyses where I dropped corpus callosum. Findings still great.
+[x] need to update paper with this analysis.
+
+#### 2023 0621
+
+[x] Interested in whether lesions to fibers involved in serotonergic projections (primarily UF) are more associated with depression (or response to SSRIs, I imagine people who have completely obliterated it - bad, people who have some left - okay, people who have healthy - maybe not as much impact.
+Inverted u-shaped curve).
+No differences in UF
+
+```{r}
+
+#same results when using volume. No difference with UF; W = 16268, p-value = 0.3762
+dep_vs_healthy_wilcoxon_lesion_volume_results <-  wilcox.test(dep_and_healthy_groups_for_ICD_analysis$UF_L_prop[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$UF_L_prop[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = F)
+
+#same results when using volume. No difference with UF; W = 16076, p-value = 0.2914
+dep_vs_healthy_wilcoxon_lesion_volume_results <-  wilcox.test(dep_and_healthy_groups_for_ICD_analysis$UF_R_prop[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$UF_R_prop[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = F)
+```
+[] Same with NE
+
+[] What if you ranked fascicles based on their overlap with GM areas that have 5HT receptors
+ - need a dasb mask essentially
+ - 5ht1a mask
+ - 5ht2a?
+ - SERT (dasb)
+
+[] Birth control study
+  - disease differences in people on/off bcp?
+  - longitudinal outcome models?
+
+20230824
+SOBP analyses
+[] age x sex, age x dep 
+[x] looking at indiv fascicles in dep network, figure out which ones independently separate dep vs non - theoretically do prop lost ~ dep diagnosis
+   - only uncorrected findings
+[]
+
+20230829
+[] take each map from golia, project to surface, spin test for overlap with depression map from shan.
+Rank by overlap.
+Make white matter backbones.
+Test for lesion network enrichment in depression group.
+correlate effect sizes from spin correlation.
+Idea would be that maps that were more highly related to depression network are also more associated with pt vs control differences.
+And also, hypothesize that these areas are enriched for major depression neurotransmitters
+[] also consider receptor density studies. 
+[] if you find that fascicles that feed into areas with higher receptor densities, could be data for PET
+[] also wonder whether medication selection using brain imaging data could be useful for depression
+
+#### 20230927 analyses for BP revision
+
+[x] Add sex into model
+
+#with sex; diagnosis: Dep > nondep T = 2.5, p = 0.012; sex: M>F = 1, p = 0.32; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020
+diagnosis_by_network_interaction_lmer_sex <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject) + osex, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer_sex)
+
+#Diagnosis: MS+ > MS- 0.013-0.10; sex: M>F 0.06 - -0.02 Network dep>nondep -.104--.076; intx: -0.040--0.003
+confint(diagnosis_by_network_interaction_lmer_sex)
+
+[x] Prevalence of depression in the PROMIS cohort
+ - chi sq for prevalence? 49 promis scores in depressed group (out of 232); 36 promis scores in nondepressed group (out of 148)
+ - #X-squared = 0.21, p = 0.65, not significant)
+
+[] use different threshold of 
+  [] 2.3
+  [] 2.6
+```{bash}
+cd /Users/eballer/BBL/msdepression/templates/harvard_depression
+#we are going to threshold the DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii at 2.3 and 2.6
+3dcalc -a DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii -expr 'ispositive(a-2.3)' -prefix DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_2_3.nii
+3dcalc -a DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii -expr 'ispositive(a-2.6)' -prefix DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_2_6.nii
+
+#send to pmacs
+scp DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_2_* eballer@bblsub.pmacs.upenn.edu:/project/msdepression/templates/harvard_depression/.
+
+#on pmacs
+
+#change the scripts
+more calc_vol_fascicles_within_depression_mask_3_09.sh  | perl -pe 's/3\.09/2\.3/' | perl -pe 's/3_09/2_3/' > calc_vol_fascicles_within_depression_mask_2_3.sh
+
+more calc_vol_fascicles_within_depression_mask_3_09.sh  | perl -pe 's/3\.09/2\.6/' | perl -pe 's/3_09/2_6/' > calc_vol_fascicles_within_depression_mask_2_6.sh
+
+#go in to each script individually and update the mask to be DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_2_3.nii and 2_6 individually
+
+#run the scripts
+sh calc_vol_fascicles_within_depression_mask_2_3.sh
+sh calc_vol_fascicles_within_depression_mask_2_6.sh
+
+#send output files back to home; from local, needed to do without wildcard because zsh was giving me a hard time
+scp eballer@bblsub.pmacs.upenn.edu:/project/msdepression/results/streamline_volume_within_dep_network_2_6.csv /Users/eballer/BBL/msdepression/results/.
+scp eballer@bblsub.pmacs.upenn.edu:/project/msdepression/results/streamline_volume_within_dep_network_2_3.csv /Users/eballer/BBL/msdepression/results/.
+
+```

### 2023-09-29 Biological_Psychiatry_Revision_Scripts/msdepression_n380_started_20230927_mm_BP_r1.Rmd
with sensitivity analysis removing meds

@@
     - X-squared = 0.21, p = 0.65, not significant
+
+[x] sensitivity analysis removing people on methylpred/pred/interferone B (lose 20 people, 5% of sample; MS+Dep = 221 (-11, 4.7%); MS-Dep = 139(-9, 6%))
+
+    -Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.070-0.10, P <0.001)
+    -Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.05, 95% CI 0.001-0.095, P= 0.046)
+    -Main effect of interaction (Estimate = 0.018, 95% CI 0.00 - 0.030, P=0.060)
 
@@
 ```
+
+#repeat analysis but remove people on neuroactive MS meds
+```{r depression_vs_healthy_overall_volume_of_fascicles_affected_noifnbeta_steroids}
+## compare disease within and outside depression network, between diagnoses, and interactions
+
+###############################
+### Constructing Data Frame ###
+###############################
+
+#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
+dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
+  filter(depGroupVar != 0)  %>% 
+  group_by(EMPI) %>% 
+  arrange(EXAM_DATE) %>% 
+  slice(1) %>%
+
+ #sum of total volume lost and proportion of volume lost 
+  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
+  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_volume_all_full_fascicles) %>%
+
+  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_indepnet_top_quartile = 
+           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves)))) %>%
+  mutate(proportion_volume_lost_dep_net_by_network_size_sum = sum_fascicle_vol_lost_indepnet_top_quartile /dep_network_total_volume_top_quartile_no_cranial_nerves) %>%
+
+  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
+  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles = 
+           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves)))) %>%
+  mutate(proportion_volume_lost_nondep_net_by_network_size_sum = sum_fascicle_vol_lost_nondep_bottom_3_quartiles /bottom_75_total_volume_no_cranial_nerves) %>%
+
+  #remove people on ms medications that are known to worsen mood
+ # filter(!str_detect(Medication, 'INTERFERON|PREDNISONE|METHYLPREDNISOLONE')) %>% 
+  #filter(!str_detect(Medication, 'INTERFERON')) %>%
+  filter(!str_detect(Medication, 'PREDNISONE')) %>%
+ # filter(!str_detect(Medication, 'METHYLPREDNISOLONE')) %>%
+  ungroup()
+
+###################
+
+### Interaction Network x Diagnosis
+
+#clean up data frame
+df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
+  mutate(subject = as.factor(EMPI)) %>%
+  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
+  mutate(Diagnosis = as.factor(Diagnosis)) %>%
+  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
+  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
+  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network, PAT_AGE_AT_EXAM) %>% 
+  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
+  mutate(Network = as.factor(Network)) %>%
+  ungroup()
+
+#diagnosis: Dep > nondep T = 2.0, p = 0.046; Network dep>nondep = T=12.3, p = 2*10^-16; intx = T = 1.89, p = 0.06
+diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lmer)
+
+#Diagnosis: MS+ > MS- 0.001-0.9; Network dep>nondep -.104--.078; intx: -0.040--0.00
+confint(diagnosis_by_network_interaction_lmer)
+
+#Final data: 
+#Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.070-0.10, P <0.001)
+#Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.05, 95% CI 0.001-0.095, P= 0.046)
+#Main effect of interaction (Estimate = 0.018, 95% CI 0.00 - 0.030, P=0.060)
+
+```

### 2023-09-29 lab_notebook/lab_notebook.Rmd
Add files via upload

@@
 
-[] use different threshold of 
-  [] 2.3
-  [] 2.6
+[x] use different threshold of 
+  [x] 2.3
+  [x] 2.6
 ```{bash}
@@
 ```
+
+[] a couple of other analyses
+- people on Interferon B (11 total, 5 in nondepressed, 6 in depressed; solumedrol 2/2)
+```{r}
+table(dep_and_healthy_groups_for_ICD_analysis$depGroupVar[grep(pattern = "INTERF", x = dep_and_healthy_groups_for_ICD_analysis$Medication)])
+```
+- people on steroids (prednisone 1 in nondep, 1 in dep)
+- 
+```{r}
+```
+- people on ocrelizumab (17 in healthy group, 14 in depressed group)
+```{r}
+table(dep_and_healthy_groups_for_ICD_analysis$depGroupVar[grep(pattern = "OCRELIZUMAB", x = dep_and_healthy_groups_for_ICD_analysis$Medication)])
+```
+
+Could also do sensitivity analysis and remove the people on interferon Beta, prednisone, methylprednisolone
+```{r}
+dim(dep_and_healthy_groups_for_ICD_analysis %>% filter(!str_detect(Medication, 'INTERFERON|PREDNISON|METHYLPREDNISOLONE')))  #remove
+
+```

### 2023-10-05 Biological_Psychiatry_Revision_Scripts/msdepression_n380_started_20230927_mm_BP_r1.Rmd
add diagnosis*age interaction

@@
     -Main effect of interaction (Estimate = 0.018, 95% CI 0.00 - 0.030, P=0.060)
+
+[x] agexdep to address duration issue: lm(proportion_affected ~ Diagnosis*Age)
+
+  - Age x dep interaction:  (Estimate<0.001, CI -0.003-0.002 , P=0.84)
 
@@
 
+#Agexdepression interaction - Not significant
+diagnosis_by_network_interaction_lm_agexdep <- lm(proportion_affected ~  Diagnosis*PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
+summary(diagnosis_by_network_interaction_lm_agexdep)
+
+#Age x dep interaction:  (Estimate<0.001, CI -0.003-0.002 , P=0.84)
+confint(diagnosis_by_network_interaction_lm_agexdep)
+
 #with sex; diagnosis: Dep > nondep T = 2.5, p = 0.012; sex: M>F = 1, p = 0.32; Network dep>nondep = T=12.4, p = 2*10^-16; intx = T = 2.3, p = 0.020

### 2023-10-21 scripts/msdepression_n380_BP_r1_script.Rmd
Create msdepression_n380_BP_r1_script.Rmd

--- 2023-10-05 Biological_Psychiatry_Revision_Scripts/msdepression_n380_started_20230927_mm_BP_r1.Rmd
@@
 
-    -Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.070-0.10, P <0.001)
+    -Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.08-0.11, P <0.001)
     -Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.05, 95% CI 0.001-0.095, P= 0.046)
@@
 #Final data: 
-#Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.070-0.10, P <0.001)
+#Main effect of network Dep > Nondep (Estimate = 0.09, 95% CI 0.08-0.11, P <0.001)
 #Main effects of diagnosis MS+Depression > MS-Depression (Estimate = 0.05, 95% CI 0.001-0.095, P= 0.046)

### 2023-10-21 scripts/msdepression_n380_BP_r1_script.Rmd
Update msdepression_n380_BP_r1_script.Rmd

@@
 author: "Erica Baller"
-date: '2022-08-24'
+date: '2023-10-24'
 output:
